### 2020-05-01 physcraper_ms.Rmd
starting :bowtie:

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Physcraper: A Python package for continually updated gene trees'
tags:
  - Python
  - phylogenetics
  - update
  - species tree
  - Open Tree
authors:
  - name: Luna L. Sanchez Reyes
    orcid: 0000-0000-0000-0000
    affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
  - name: Emily Jane McTavish
    orcid: 0000-0000-0000-0000
    affiliation: 1
affiliations:
 - name: University of California, Merced
   index: 1
 - name: Institution 2
   index: 2
citation_author: Sanchez-Reyes et. al.
date: 24 April 2020
year: 2020
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

Update trees that were reported on a published peer-reviewed study, uploaded to Treebase and to the Open Tree of Life database

# Examples

Updating trees from @schoch2009ascomycota

1. Get study id from Open Tree of Life using R
  - Getting the treebase ID (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
```{r}
rotl::studies_find_studies(property = "treebaseId", value = "S2137")
```

1. Download the alignment from Treebase
  - If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.

```
```

```{bash}
opentree_scrape.py -s pg_238 -t tree109 -a treebase_alns/pg_238-LSU-M3956.nex -as "nexus" -o tmp
```

# Citations

Tools that do similar things:
pyPhlawd [@smith2019pyphlawd]

For a quick reference, the following citation commands can be used:
- `@author:2001`  ->  "Author et al. (2001)"
- `[@author:2001]` -> "(Author et al., 2001)"
- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"

# Rendered R Figures

Figures can be plotted like so:

```{r}
plot(1:10)
```

# Acknowledgements

We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
Oh, and support from Kathryn Johnston during the genesis of this project.

# References

### 2020-05-03 physcraper_ms.Rmd
intro

@@
 
+# Intro
+
+It is well known that GenBank holds enormous amounts of genetic data, and it continues to grow.
+A lot of this genetic data has the potential to be used to reconstruct the phylogenetic history of various organisms [@sanderson2008phylota].
+Pipelines that harness this potential have been available for over a decade now, such
+as the Phylota brower, and PHLAWD. New ones keep on being developed, such as SUPERSMART
+and the upgraded version of PHLAWD, PyPHLAWD.
+Notably large phylogenies have been constructed using some of these tools,
+Some other have not been used that much. So, how well accepted is this approach in the community?
+
+Concerns:
+Errors in identification of sequences
+Little control along the process
+Too much of a black box?
+
+Most of these phylogenies are being constructed by people learning about the methods,
+so they want to know what is going on.
+
+The pipelines are so powerful and they will give you an answer, but there
+is no way to assess if it is better than previous answers, it just assumes it is better because it used more data.
+
+All these pipelines start tree consruction from zero?
+
+The goal of Physcraper is to build upon previous phylogenegtic knowledge,
+allowing a direct comparison of existing phylogenies to phylogenies constructed
+using new information obtained from GenBank
+
+To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
+to the TreeBase database, to use the original DNA dataset matrices that produced
+a phylogeny in the OToL database as
+a starting point to get new genetic data belonging to the focal group of study, and
+update the phylogenetic relationships displayed by that dataset in particular.
+
+Physcraper also allows node by node comparison of the trees, using the conflict API of OToL.
+
+A less automated workflow is one in which the alignments that generated the published
+phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
+are provided by the users.
+
+The original tree is used as starting tree for the phylogenetic searches, but it
+can also be set as a full topological constraint or not used at all, depending on
+the goals of the user.
+
 # Examples
@@
 
-1. Get study id from Open Tree of Life using R
-  - Getting the treebase ID (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
+1. Get study id from Open Tree of Life. There are some options to do this:
+  - You can go to the Open Tree of Life web browser, or
+  - you can do it using R:
+    - By getting the treebase ID of the study (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
 ```{r}
@@
 1. Download the alignment from Treebase
-  - If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.
+If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.
+  To make this task easier, you can use a command to download everything into your working folder:
+```
+auto_scrape.py -s pg_238 -t tree109 -o ../physcraper_example/treebase_alns
+```
+In this example, all alignments posted on Treebase were used to reconstruct both trees.
 
-```
-```
+1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:
 
@@
 
-# Citations
+# Tools on a similar track:
 
@@
 pyPhlawd [@smith2019pyphlawd]
-
-For a quick reference, the following citation commands can be used:
-- `@author:2001`  ->  "Author et al. (2001)"
-- `[@author:2001]` -> "(Author et al., 2001)"
-- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"
-
-# Rendered R Figures
-
-Figures can be plotted like so:
-
-```{r}
-plot(1:10)
-```
+SUPERSMART [@antonelli]
 
@@
 
-We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
-Oh, and support from Kathryn Johnston during the genesis of this project.
+We acknowledge contributions from
 

### 2020-05-04 physcraper_ms.Rmd
summary

@@
 
-Update trees that were reported on a published peer-reviewed study, uploaded to Treebase and to the Open Tree of Life database
+1. Phylogenies are still a key part of research in all areas of biology. The nonspecialist
+can use one of the many tools available to automatize phylogenetic reconstruction.
+However, interpretation of results, comparison with previously available phylogenetic
+hypotheses, and choosing of one phylogeny for discussion still impose difficulties
+to one that is not a specialist on phylogenetic methods or on a particular group of study.
+
+1. Physcraper is an open‐source Python program that automatizes the update of public
+phylogenies by making use of public DNA sequence data and taxonomic information.
+
+1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
+hypothesis based on public expert phylogentic knowledge.
+The phylogenetics and group specialist will find it useful as a tool to facilitate comparison
+of alternative phylogenetic hypotheses (topologies).
+***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
+the ones that do not know about phylogenetic methods and the ones that might know
+about phylogenetic methods but do not know of a certain biological group.***
+
+1. Physcraper implements node by node comparison of the the original and the updated
+trees using the conflict API of OToL.
+
+1. We hope the physcraper workflow encourages researchers on their data sharing practices.
+
+1. Physcraper can be used with any OS. Detailed instructions for installation and
+use are available at <https://github.com/McTavishLab/physcraper>.
 
@@
 
+There is a lot of phylogenetic knowledge already available.
+Some of it has been made public at OToL.
+Most or all public phylogenies have the potential to be updated with the ever increasing
+amount of genetic data that is available on GenBank.
+There are now various methods to automatize use of GenBank data for phylogenetic reconstruction, which are
+generally intended to facilitate the use of phylogenetic knowledge for the nonspecialist.
+However, none of them provide a means to compare with previously available expert
+knowledge on a particular group.
+
+A key aspect of the standard phylogenetic workflow is comparison with already existing
+phylogenetic hypotheses and with phylogenies that are considered "best" by experts
+not only in phylogenetics, but also experts of the focal group of study.
+
 It is well known that GenBank holds enormous amounts of genetic data, and it continues to grow.
-A lot of this genetic data has the potential to be used to reconstruct the phylogenetic history of various organisms [@sanderson2008phylota].
+A lot of this genetic data has the potential to be used to reconstruct the phylogenetic
+history of various organisms [@sanderson2008phylota].
 Pipelines that harness this potential have been available for over a decade now, such
@@
 
-Concerns:
+Concerns with these tools:
 Errors in identification of sequences
@@
 The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is better because it used more data.
+is no way to assess if it is better than previous answers, it just assumes it is
+better because it used more data.
 
@@
 
-The goal of Physcraper is to build upon previous phylogenegtic knowledge,
-allowing a direct comparison of existing phylogenies to phylogenies constructed
-using new information obtained from GenBank
+The goal of Physcraper is to build upon previous phylogenetic knowledge,
+allowing a direct comparison of existing phylogenies to phylogenies that are constructed
+using new genetic data from GenBank
 
 To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to use the original DNA dataset matrices that produced
-a phylogeny in the OToL database as
-a starting point to get new genetic data belonging to the focal group of study, and
-update the phylogenetic relationships displayed by that dataset in particular.
-
-Physcraper also allows node by node comparison of the trees, using the conflict API of OToL.
+to the TreeBase database, to (1) get the original DNA dataset matrices (alignments) that produced
+a phylogeny that was published and then made available in the OToL database, (2)
+use this DNA alignments as a starting point to get new genetic data belonging
+to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
 
@@
 
-The original tree is used as starting tree for the phylogenetic searches, but it
+The original tree is by defualt used as starting tree for the phylogenetic searches, but it
 can also be set as a full topological constraint or not used at all, depending on
 the goals of the user.
+
+Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
+
+# Physcraper
 

### 2020-05-07 phylota-overview.Rmd
comparison with other tools

# Phylota overview

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database [@sanderson2008phylota].

Since then, it has been cited 122 times for different reasons.

1. As an example study of very large and comprehensive phylogeny from minedDNA sequence data, (even if no phylogeny was published there, only the method to do so):
  - @beaulieu2012modeling
  -

1. As an example tool that mines GenBak data for phylogenetic reconstruction:
  - original publication of PHLAWD [@smith2009mega]
  -

1. When the software was actually used to construct a DNA data set for phylogenetic reconstruction of:
  - A 1000 tip phylogeny of the family of the nightshades [@sarkinen2013solanaceae]
  - A 56 tip phylogeny of crustacean zooplancton [@helmus2010communities] -- ecological study
  - A 63 tip phylogeny of the Salmonidae family [@crete2012salmonidae]
  - A 321 tip phylogeny of Testudines [@thomson2010sparse]

1. When the website was used to identify sequences and markers available in
GenBank for a particular group, In this cases, the dataset mining was performed with other tools, or not performed at all and just used for discussion:
  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset constructed with PHLAWD
  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed by hand (I think??)
  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil [@simon2012cerrado] -- just for discussion of the lack of markers available for these species on GenBank

1. Sometimes, it was cited by mistake:
    - In this 630 tip phylogeny of the Caryophyllaceae study [@greenberg2011caryophyllaceae] it might have been originally
    cited as an example of large phylogenies that reflect well supported relationships
    from previous smaller phylogenies. However, it was removed from the text but
    not from the final list of references. The DNA data set was constructed by hand most probably.

### 2020-05-07 physcraper_ms.Rmd
more examples

@@
 ---
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "##",
+  eval = TRUE
+)
+```
 
@@
 
-1. We hope the physcraper workflow encourages researchers on their data sharing practices.
+1. We hope the physcraper workflow demonstrates the benefits of open results in phylogenetics and encourages researchers to strive for better data sharing practices.
 
@@
 
-# Physcraper
+# Physcraper overview
 
-# Examples
+# Use case/ example
 
-Updating trees from @schoch2009ascomycota
+Imagine you are starting to work on a new biological group X. You have not much of an idea about
+its phylogenetic relationships, you are a newly established researcher, and the group
+is not anything any of your collaborators have worked on before.
+A good idea is to start an intensive litt review on the phylogenetics of the group.
+Rapidly, you find out there are 5 different phylogeneis, that used different markers, and that
+the papers, published at different times, do not discuss which phylogeny is
+the one accepted by the expert community on X. You might need
+to go to the annual conference of X, and even then, you might only find
+different and contrasting opinions.
+Somewhere along these months or even years doing this task, you looked into the
+the OToL database. You found in there some or all the published trees of X, along
+with a tree that has been deemed the best tree by curators and ideally experts on X?
 
-1. Get study id from Open Tree of Life. There are some options to do this:
-  - You can go to the Open Tree of Life web browser, or
-  - you can do it using R:
-    - By getting the treebase ID of the study (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
-```{r}
+## Ascomycota Example
+
+Let's be more specific now about our X group and say it is the Ascomycota.
+The best tree currently available in OToL was published by @schoch2009ascomycota.
+The first step, is to get the Open Tree of Life study id. There are some options to do this:
+- You can go to the Open Tree of Life website and browse until you find it, or
+- you can get the study id using R tools:
+  - By using the treebase ID of the study (which is not fully exposed on the
+    Treebase website home page of the study, so you have to really look it up manually):
+```{r, treebase, eval = TRUE}
 rotl::studies_find_studies(property = "treebaseId", value = "S2137")
+```
+  - By using the name of the focal clade of study (but this behaved very differently):
+```{r, focal, eval = FALSE}
+rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
+```
+
+Once we have the study id, we can gather the trees published on that study:
+```{r, treeids, eval = TRUE}
+rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
+rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
+my_trees <- rotl::get_study("pg_238")
+```
+Both trees from this study have `r ape::Ntip(my_trees[[1]])` tips.
+
+Let's check what one of the trees looks like:
+
+```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.height = 4}
+ape::plot.phylo(my_trees[[1]], type = "fan", cex = 0.1)
 ```
@@
 ```
-auto_scrape.py -s pg_238 -t tree109 -o ../physcraper_example/treebase_alns
+physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
 ```
@@
 
-```{bash}
-opentree_scrape.py -s pg_238 -t tree109 -a treebase_alns/pg_238-LSU-M3956.nex -as "nexus" -o tmp
+```
+physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
+```
+
+## Testudines example
+
+Phylogeny of the Testudines 6 tips from @crawford2012more
+There is just one tree in OToL.
+There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.
+
+```
+physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
 ```
@@
 pyPhlawd [@smith2019pyphlawd]
-SUPERSMART [@antonelli]
+SUPERSMART [@antonelli2017toward]
 

### 2020-05-08 joss_submission.Rmd
modularization

--- 2020-05-03 physcraper_ms.Rmd
@@
    index: 1
- - name: Institution 2
-   index: 2
 citation_author: Sanchez-Reyes et. al.
@@
 ---
-
-# Summary
-
-Update trees that were reported on a published peer-reviewed study, uploaded to Treebase and to the Open Tree of Life database
-
-# Intro
-
-It is well known that GenBank holds enormous amounts of genetic data, and it continues to grow.
-A lot of this genetic data has the potential to be used to reconstruct the phylogenetic history of various organisms [@sanderson2008phylota].
-Pipelines that harness this potential have been available for over a decade now, such
-as the Phylota brower, and PHLAWD. New ones keep on being developed, such as SUPERSMART
-and the upgraded version of PHLAWD, PyPHLAWD.
-Notably large phylogenies have been constructed using some of these tools,
-Some other have not been used that much. So, how well accepted is this approach in the community?
-
-Concerns:
-Errors in identification of sequences
-Little control along the process
-Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is better because it used more data.
-
-All these pipelines start tree consruction from zero?
-
-The goal of Physcraper is to build upon previous phylogenegtic knowledge,
-allowing a direct comparison of existing phylogenies to phylogenies constructed
-using new information obtained from GenBank
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to use the original DNA dataset matrices that produced
-a phylogeny in the OToL database as
-a starting point to get new genetic data belonging to the focal group of study, and
-update the phylogenetic relationships displayed by that dataset in particular.
-
-Physcraper also allows node by node comparison of the trees, using the conflict API of OToL.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-The original tree is used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
-
-# Examples
-
-Updating trees from @schoch2009ascomycota
-
-1. Get study id from Open Tree of Life. There are some options to do this:
-  - You can go to the Open Tree of Life web browser, or
-  - you can do it using R:
-    - By getting the treebase ID of the study (which is not fully exposed on the Treebase website home page of the study, so you have to look it up manually)
-```{r}
-rotl::studies_find_studies(property = "treebaseId", value = "S2137")
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "##",
+  eval = TRUE
+)
 ```
 
-1. Download the alignment from Treebase
-If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.
-  To make this task easier, you can use a command to download everything into your working folder:
+```{r body, child="physcraper_ms.Rmd"}
 ```
-auto_scrape.py -s pg_238 -t tree109 -o ../physcraper_example/treebase_alns
-```
-In this example, all alignments posted on Treebase were used to reconstruct both trees.
-
-1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:
-
-```{bash}
-opentree_scrape.py -s pg_238 -t tree109 -a treebase_alns/pg_238-LSU-M3956.nex -as "nexus" -o tmp
-```
-
-# Tools on a similar track:
-
-Tools that do similar things:
-pyPhlawd [@smith2019pyphlawd]
-SUPERSMART [@antonelli]
-
-# Acknowledgements
-
-We acknowledge contributions from
-
-# References

### 2020-05-08 physcraper_ms_ 2020-05-05.tex
modularization

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
% \renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/3.6/Resources/library/rticles/rmarkdown/templates/joss_article/resources/JOSS-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\footnotesize{\sffamily Sanchez-Reyes et. al., (2020). Physcraper: A Python package for continually updated gene trees. \textit{Journal of Open Source Software}, (), . \href{https://doi.org/}{https://doi.org/}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Physcraper: A Python package for continually updated gene trees},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{Physcraper: A Python package for continually updated gene trees}

        \author[1]{Luna L. Sanchez Reyes}
          \author[1]{Emily Jane McTavish}

      \affil[1]{University of California, Merced}
      \affil[2]{Institution 2}
  \date{\vspace{-5ex}}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{}{\color{linky}{Repository}} \ExternalLink
    \item \href{}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  {\bfseries Submitted:} \\
  {\bfseries Published:}

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Phylogenies are still a key part of research in all areas of biology.
  The nonspecialist can use one of the many tools available to
  automatize phylogenetic reconstruction. However, interpretation of
  results, comparison with previously available phylogenetic hypotheses,
  and choosing of one phylogeny for discussion still impose difficulties
  to one that is not a specialist on phylogenetic methods or on a
  particular group of study.
\item
  Physcraper is an open‐source Python program that automatizes the
  update of public phylogenies by making use of public DNA sequence data
  and taxonomic information.
\item
  Physcraper can be used by the nonspecialist, as a tool to generate
  phylogenetic hypothesis based on public expert phylogentic knowledge.
  The phylogenetics and group specialist will find it useful as a tool
  to facilitate comparison of alternative phylogenetic hypotheses
  (topologies). \textbf{\emph{Is physcraper intended for the
  nonspecialist?? We have two types of nonspecialists: the ones that do
  not know about phylogenetic methods and the ones that might know about
  phylogenetic methods but do not know of a certain biological group.}}
\item
  Physcraper implements node by node comparison of the the original and
  the updated trees using the conflict API of OToL.
\item
  We hope the physcraper workflow demonstrates the benefits of open
  results in phylogenetics and encourages researchers to strive for
  better data sharing practices.
\item
  Physcraper can be used with any OS. Detailed instructions for
  installation and use are available at
  \url{https://github.com/McTavishLab/physcraper}.
\end{enumerate}

\hypertarget{intro}{%
\section{Intro}\label{intro}}

There is a lot of phylogenetic knowledge already available. Some of it
has been made public at OToL. Most or all public phylogenies have the
potential to be updated with the ever increasing amount of genetic data
that is available on GenBank. There are now various methods to
automatize use of GenBank data for phylogenetic reconstruction, which
are generally intended to facilitate the use of phylogenetic knowledge
for the nonspecialist. However, none of them provide a means to compare
with previously available expert knowledge on a particular group.

A key aspect of the standard phylogenetic workflow is comparison with
already existing phylogenetic hypotheses and with phylogenies that are
considered ``best'' by experts not only in phylogenetics, but also
experts of the focal group of study.

It is well known that GenBank holds enormous amounts of genetic data,
and it continues to grow. A lot of this genetic data has the potential
to be used to reconstruct the phylogenetic history of various organisms
(Sanderson, Boss, Chen, Cranston, \& Wehe, 2008). Pipelines that harness
this potential have been available for over a decade now, such as the
Phylota brower, and PHLAWD. New ones keep on being developed, such as
SUPERSMART and the upgraded version of PHLAWD, PyPHLAWD. Notably large
phylogenies have been constructed using some of these tools, Some other
have not been used that much. So, how well accepted is this approach in
the community?

Concerns with these tools: Errors in identification of sequences Little
control along the process Too much of a black box?

Most of these phylogenies are being constructed by people learning about
the methods, so they want to know what is going on.

The pipelines are so powerful and they will give you an answer, but
there is no way to assess if it is better than previous answers, it just
assumes it is better because it used more data.

All these pipelines start tree consruction from zero?

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison of existing phylogenies to phylogenies that
are constructed using new genetic data from GenBank

To achieve this, Physcraper uses the Open Tree of Life phylesystem and
connects it to the TreeBase database, to (1) get the original DNA
dataset matrices (alignments) that produced a phylogeny that was
published and then made available in the OToL database, (2) use this DNA
alignments as a starting point to get new genetic data belonging to the
focal group of study, to (3) finally update the phylogenetic
relationships in the group.

A less automated workflow is one in which the alignments that generated
the published phylogeny are stored in other public database (such as
DRYAD) or elsewhere (the users computer), and are provided by the users.

The original tree is by defualt used as starting tree for the
phylogenetic searches, but it can also be set as a full topological
constraint or not used at all, depending on the goals of the user.

Physcraper implements node by node comparison of the the original and
the updated trees, using the conflict API of OToL.

\hypertarget{physcraper-overview}{%
\section{Physcraper overview}\label{physcraper-overview}}

\hypertarget{examples}{%
\section{Examples}\label{examples}}

Imagine you are starting to work on a new biological group X. You have
not much of an idea about its phylogenetic relationships, you are a
newly established researcher, and the group is not anything any of your
collaborators have worked on before. A good idea is to start an
intensive litt review on the phylogenetics of the group. Rapidly, you
find out there are 5 different phylogeneis, that used different markers,
and that the papers, published at different times, do not discuss which
phylogeny is the one accepted by the expert community on X. You might
need to go to the annual conference of X, and even then, you might only
find different and contrasting opinions. Somewhere along these months or
even years doing this task, you looked into the the OToL database. You
found in there some or all the published trees of X, along with a tree
that has been deemed the best tree by curators and ideally experts on X?

Let's be more specific now about our X group and say it is the
Ascomycota. The best tree currently available in OToL was published by
Schoch et al. (2009). The first step, is to get the Open Tree of Life
study id. There are some options to do this: - You can go to the Open
Tree of Life website and browse until you find it, or - you can get the
study id using R tools: - By using the treebase ID of the study (which
is not fully exposed on the Treebase website home page of the study, so
you have to really look it up manually):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property =} \StringTok{"treebaseId"}\NormalTok{, }\DataTypeTok{value =} \StringTok{"S2137"}\NormalTok{)}
\CommentTok{##   study_ids n_trees         tree_ids candidate study_year title}
\CommentTok{## 1    pg_238       2 tree109, tree110                 2009      }
\CommentTok{##                                 study_doi}
\CommentTok{## 1 http://dx.doi.org/10.1093/sysbio/syp020}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  By using the name of the focal clade of study (but this behaved very
  differently):
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property=}\StringTok{"ot:focalCladeOTTTaxonName"}\NormalTok{, }\DataTypeTok{value=}\StringTok{"Ascomycota"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Once we have the study id, we can gather the trees published on that
study:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_tree_ids}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
\CommentTok{## [1] "tree109" "tree110"}
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{candidate_for_synth}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
\CommentTok{## NULL}
\NormalTok{my_trees <-}\StringTok{ }\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_study}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Both trees from this study have 434 tips.

Let's check what one of the trees looks like:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ape}\OperatorTok{::}\KeywordTok{plot.phylo}\NormalTok{(my_trees[[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{type =} \StringTok{"fan"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics[width=468px]{physcraper_ms_ 2020-05-05_files/figure-latex/fungi-phylo-1}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Download the alignment from Treebase If you are on the Treebase home
  page of the study, you can navigate to the matrix tab, and manually
  download the alignments that were used to reconstruct the trees
  reported on the study that were also uploaded to Treebase and to the
  Open Tree of Life repository. To make this task easier, you can use a
  command to download everything into your working folder:
\end{enumerate}

\begin{verbatim}
physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
\end{verbatim}

In this example, all alignments posted on Treebase were used to
reconstruct both trees.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  With the study id and the alignment files saved locally, we can do a
  physcraper run with the command:
\end{enumerate}

\begin{verbatim}
physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
\end{verbatim}

\hypertarget{tools-on-a-similar-track}{%
\section{Tools on a similar track:}\label{tools-on-a-similar-track}}

Tools that do similar things: pyPhlawd (Smith \& Walker, 2019)
SUPERSMART (Antonelli et al., 2017)

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We acknowledge contributions from

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-antonelli2017toward}{}%
Antonelli, A., Hettling, H., Condamine, F. L., Vos, K., Nilsson, R. H.,
Sanderson, M. J., Sauquet, H., et al. (2017). Toward a self-updating
platform for estimating rates of speciation and migration, ages, and
relationships of taxa. \emph{Systematic Biology}, \emph{66}(2),
152--166.

\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
Sanderson, M. J., Boss, D., Chen, D., Cranston, K. A., \& Wehe, A.
(2008). The PhyLoTA Browser: Processing GenBank for Molecular
Phylogenetics Research. \emph{Systematic Biology}, \emph{57}(3),
335--346.
doi:\href{https://doi.org/10.1080/10635150802158688}{10.1080/10635150802158688}

\leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
Schoch, C. L., Sung, G.-H., López-Giráldez, F., Townsend, J. P.,
Miadlikowska, J., Hofstetter, V., Robbertse, B., et al. (2009). The
ascomycota tree of life: A phylum-wide phylogeny clarifies the origin
and evolution of fundamental reproductive and ecological traits.
\emph{Systematic biology}, \emph{58}(2), 224--239.

\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
Smith, S. A., \& Walker, J. F. (2019). PyPHLAWD: A python tool for
phylogenetic dataset construction. \emph{Methods in Ecology and
Evolution}, \emph{10}(1), 104--108.

\end{document}

### 2020-05-08 sysbio_submission.Rmd
modularization

--- 2020-05-08 joss_submission.Rmd
@@
 bibliography: paper.bib
-output: rticles::joss_article
+output:
+  pdf_document:
+    template: template.tex
 csl: apa.csl
-journal: JOSS
+links-as-notes: true
 ---

### 2020-05-11 phylota-overview.Rmd
more phylota examples

@@
 
-1. As an example study of very large and comprehensive phylogeny from minedDNA sequence data, (even if no phylogeny was published there, only the method to do so):
+1. As an example study of very large and comprehensive phylogeny from mined DNA sequence data, (even if no phylogeny was really published there, only the method to do so):
   - @beaulieu2012modeling
@@
 
-1. As an example tool that mines GenBak data for phylogenetic reconstruction:
+1. As an example of a tool that mines GenBank data for phylogenetic reconstruction:
   - original publication of PHLAWD [@smith2009mega]
-  -
+  - an analysis identifying research priorities and data requirements for resolving
+  the red algal tree of life [@verbruggen2010data]
+  - a review for ecologists about phylogenetic tools [@roquet2013building]
 
-1. When the software was actually used to construct a DNA data set for phylogenetic reconstruction of:
+1. When the software was actually used to construct (partially or in full) a DNA
+data set to be used for phylogenetic reconstruction:
   - A 1000 tip phylogeny of the family of the nightshades [@sarkinen2013solanaceae]
@@
   - A 321 tip phylogeny of Testudines [@thomson2010sparse]
+  - A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish [@martin2011trophic]
+  - A 2,957 taxa phylogeny of the class Moniloformopses of living ferns [@lehtonen2011towards]
+  - A 2,573 species phylogeny of the Papilionoidea [@hardy2014specialization]
+  - A 23 taxa phylogeny of the California flora [@anacker2011origins]
+  - Phylogenies of 6 different clades of flowering plants representing an independent
+  evolutionary origin of extrafloral nectaries: *Byttneria* (Malvaceae), *Pleopeltis* (Polypodiaceae),
+  *Polygoneae* (Polygoneaceae), *Senna* (Fabaceae), *Turnera* (Passifloraceae), and *Viburnum*
+  (Adoxaceae) [@weber2014defense].
+  - To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
+  sampled at different taxonomic levels [@faurby2015species]
+  -
 
 1. When the website was used to identify sequences and markers available in
-GenBank for a particular group, In this cases, the dataset mining was performed with other tools, or not performed at all and just used for discussion:
-  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset constructed with PHLAWD
-  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed by hand (I think??)
-  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil [@simon2012cerrado] -- just for discussion of the lack of markers available for these species on GenBank
+GenBank for a particular group. In this cases, the dataset mining was either performed
+with other tools, or not performed at all and just used for discussion:
+  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset
+  constructed with PHLAWD
+  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed
+  by hand (I think??)
+  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil
+  [@simon2012cerrado] -- just for discussion of the lack of markers available for
+  these species on GenBank
+  - A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
+  protists [@morrison2009apicomplexa].
+  - Three  data sets from phylota (the suborder Pleurodira of side-necked turtles;
+    the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
+    to demonstrate and exemplify phylogenetic decisiveness [@sanderson2010phylogenomics]
 

### 2020-05-11 joss_submission.Rmd
:computer:

@@
     affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
+  - name: Martha Kandziora
+    orcid: 0000-0000-0000-0000
+    affiliation: 1
   - name: Emily Jane McTavish

### 2020-05-11 physcraper_ms.Rmd
methods

@@
----
-# Example from https://joss.readthedocs.io/en/latest/submitting.html
-title: 'Physcraper: A Python package for continually updated gene trees'
-tags:
-  - Python
-  - phylogenetics
-  - update
-  - species tree
-  - Open Tree
-authors:
-  - name: Luna L. Sanchez Reyes
-    orcid: 0000-0000-0000-0000
-    affiliation: 1 # (Multiple affiliations must be quoted, like this: "1, 2")
-  - name: Emily Jane McTavish
-    orcid: 0000-0000-0000-0000
-    affiliation: 1
-affiliations:
- - name: University of California, Merced
-   index: 1
- - name: Institution 2
-   index: 2
-citation_author: Sanchez-Reyes et. al.
-date: 24 April 2020
-year: 2020
-bibliography: paper.bib
-output: rticles::joss_article
-csl: apa.csl
-journal: JOSS
----
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "##",
-  eval = TRUE
-)
-```
+# Abstract
 
-# Summary
+1. Phylogenies are a key part of research in all areas of biology. Tools that automatize
+some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
+have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
+However, interpretation of results, comparison with previously available phylogenetic
+hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
+to one that is not a specialist either on phylogenetic methods or on a particular group of study.
 
-1. Phylogenies are still a key part of research in all areas of biology. The nonspecialist
-can use one of the many tools available to automatize phylogenetic reconstruction.
-However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choosing of one phylogeny for discussion still impose difficulties
-to one that is not a specialist on phylogenetic methods or on a particular group of study.
-
-1. Physcraper is an open‐source Python program that automatizes the update of public
-phylogenies by making use of public DNA sequence data and taxonomic information.
+1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
+phylogenies by making use of public DNA sequence data and taxonomic information,
+providing a framework for comparison of phylogenies.
 
 1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on public expert phylogentic knowledge.
-The phylogenetics and group specialist will find it useful as a tool to facilitate comparison
+hypothesis based on already available expert phylogentic knowledge.
+Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
 of alternative phylogenetic hypotheses (topologies).
@@
 the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know of a certain biological group.***
+about phylogenetic methods but do not know much acbout a certain biological group.***
 
@@
 
-1. We hope the physcraper workflow demonstrates the benefits of open results in phylogenetics and encourages researchers to strive for better data sharing practices.
+1. We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
 
@@
 
-There is a lot of phylogenetic knowledge already available.
-Some of it has been made public at OToL.
-Most or all public phylogenies have the potential to be updated with the ever increasing
-amount of genetic data that is available on GenBank.
-There are now various methods to automatize use of GenBank data for phylogenetic reconstruction, which are
-generally intended to facilitate the use of phylogenetic knowledge for the nonspecialist.
-However, none of them provide a means to compare with previously available expert
-knowledge on a particular group.
+Phylogenies are important.
+Generating phylogenies is not easy.
+The process of phylogenetic reconstruction implies many steps that can be generalized to the following:
+1. Obtention of molecular or morphological character data -- get DNA from some organisms and sequence it, or get it from an online repository, such as GenBank.
+1. Assemble a hypothesis of homology -- Create a matrix of your character data, by aligning the sequences, in the case of molecular data.
+1.
+Analyse this hypothesis of homology to infer phylogenetic relationships among the organisms you are studying -- Use different available programs to infer molecular evolution, trees and times of divergence.
+1.
+Discuss the inferred relationships in the context of previous hypothesis, the biology and biogeography of the organisms, etc.
+-- Answer the question, *is this phylogenetic solution fair/reasonable?*
+
+Each of these steps require different types of specialized training: in the field,
+in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
+All of these steps also require considerable amounts of time for training and implementation.
+
+In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
+pipelines that mine already available molecular data from the GenBank repository,
+to obtain homologous characters that can be used for phylogenetic reconstruction.
+These tools have been presented as aid for the nonspecialist to decrease some of
+the difficulties in the generation of phylogenetic knowledge. However, they are
+not that often used as so, suggesting that there are still difficulties for the nonspecialist.
+The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
+However, automatizating the assembly of the character data set is a crucial step towards
+reproducibility for a task that was otherwise primarily artisanal and hence largely
+non-reproducible.
+
+Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
+They are crucial to solve problems such as food security, global warming, global health.
+There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
+In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
+
+Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank.
+
+A way to automatize the comparison of phylogenetic hypotheses and to allow reproducibility of the last step of the process.
 
@@
 phylogenetic hypotheses and with phylogenies that are considered "best" by experts
-not only in phylogenetics, but also experts of the focal group of study.
+not only in phylogenetics, but also experts on the focal group of study.
 
@@
 
-# Physcraper overview
+# How does Physcraper work?
+
+## Study tree and alignment
+
+- The phylogenetic tree has to be in Open Tree of Life. You can choose form a variety of ytrees supporting any node of the Tree of Life.
+- The alignment should be a gene alignment that was used to generate the tree. The alignments are usually stored in treebase or dryad.
+  If it is stored in Treebase, it can be downloaded by pyscraper. If it is on dryad, it has to be provided by the user.
+- Taxon name matching step: check that taxon names on the tips of the trees are in the character matrix and visceversa.
+- A csv file with the summary of taxon name matching is produced for the user.
+- Unmatched taxon names are dropped from the tree and alignment. ***What is the minimum amount of tips necessary for a physcraper run??***
+- A .tre and .aln files are generated for a physcraper run.
+
+## DNA sequence search and cleaning
+
+- The first step is to identify the taxonomic MRCA of all species in the tree. The taxonomic MRCA
+  is the closest phylogenetic MRCA that has a taxonomic name assigned in the NCBI database. ***Does
+  physcraper take into account the focal clade from Open Tree in any way???***
+- The taxonomic MRCA is used to constrain the DNA sequence search within that taxonomic group on the GenBank database.
+- The DNA sequence search can be done on a local database that is easily setup by the user, or remotely.
+- The BLAST algorithm is used to identify all DNA sequences in the database within
+  the taxonomic MRCA that are similar to each sequence in the alignment ***How does the blast is setup? Is it an all-blast-all??***
+- Reverse complement sequences are identified and translated.
+- A fasta file containing all sequences resulting from the search is generated for the user.
+
+## DNA sequence alignment
+
+- MUSCLE
+
+## Tree reconstruction
+
+- A gene tree is reconstructed for each alignment provided, using ***what software?***
+
+## Tree comparison
+
+- Conflict information can only be generated in the context of the whole Open Tree of
+Life. Otherwise, it is not really possible to get conflict data.
+***- One way to compare two independent phylogenetic trees is to compare them both to
+the synthetic OToL and then measure how well they do agianst each other***
 
@@
 
-```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.height = 4}
-ape::plot.phylo(my_trees[[1]], type = "fan", cex = 0.1)
-```
+<!-- ```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.width = 2} -->
+<!-- ape::plot.phylo(ape::ladderize(my_trees[[1]]), type = "phylogram", cex = 0.1) -->
+<!-- ``` -->
 

### 2020-05-12 physcraper_ms.Rmd
more methods

@@
 1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on already available expert phylogentic knowledge.
+hypothesis based on already available expert phylogenetic knowledge.
 Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
@@
 the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know much acbout a certain biological group.***
+about phylogenetic methods but do not know much about a certain biological group.***
 
@@
 The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizating the assembly of the character data set is a crucial step towards
+However, automatizing the assembly of the character data set is a crucial step towards
 reproducibility for a task that was otherwise primarily artisanal and hence largely
@@
 Pipelines that harness this potential have been available for over a decade now, such
-as the Phylota brower, and PHLAWD. New ones keep on being developed, such as SUPERSMART
+as the Phylota browser, and PHLAWD. New ones keep on being developed, such as SUPERSMART
 and the upgraded version of PHLAWD, PyPHLAWD.
@@
 
-All these pipelines start tree consruction from zero?
+All these pipelines start tree construction from zero?
 
@@
 To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA dataset matrices (alignments) that produced
+to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
 a phylogeny that was published and then made available in the OToL database, (2)
@@
 
-The original tree is by defualt used as starting tree for the phylogenetic searches, but it
+The original tree is by default used as starting tree for the phylogenetic searches, but it
 can also be set as a full topological constraint or not used at all, depending on
@@
 
-## Study tree and alignment
+## The input: a study tree and an alignment
 
-- The phylogenetic tree has to be in Open Tree of Life. You can choose form a variety of ytrees supporting any node of the Tree of Life.
-- The alignment should be a gene alignment that was used to generate the tree. The alignments are usually stored in treebase or dryad.
-  If it is stored in Treebase, it can be downloaded by pyscraper. If it is on dryad, it has to be provided by the user.
-- Taxon name matching step: check that taxon names on the tips of the trees are in the character matrix and visceversa.
-- A csv file with the summary of taxon name matching is produced for the user.
+- The phylogenetic tree has to be in the Open Tree of Life store [@mctavish2015phylesystem].
+  You can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
+  interested in is not in Open Tree of Life, you can easily upload it via the curator
+  tool.
+- The alignment should be a gene alignment that was used to generate the tree. The
+  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml]
+  or DRYAD (<http://datadryad.org/>). If the alignment is stored in TreeBase,
+  `physcraper` can download it directly either from the TreeBASE website (<https://treebase.org/>)
+  or through its GitHub repositiry, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
+  If the alignment is on another repository, a copy of it has to be
+  downloaded by the user, and it's location has to be provided as an argument.
+- A taxon name matching step is performed to verify that all taxon names on the tips
+  of the tree are in the character matrix and vice versa.
+- A ".csv" file with the summary of taxon name matching is produced for the user.
 - Unmatched taxon names are dropped from the tree and alignment. ***What is the minimum amount of tips necessary for a physcraper run??***
-- A .tre and .aln files are generated for a physcraper run.
+- A ".tre" and ".aln" files are generated and saved for a  `physcraper` run.
 
@@
 
-- The first step is to identify the taxonomic MRCA of all species in the tree. The taxonomic MRCA
-  is the closest phylogenetic MRCA that has a taxonomic name assigned in the NCBI database. ***Does
+- The first step is to identify the MRCA of all taxa in the tree that is also a
+  named clade in the NCBI taxonomy. This "taxonomic MRCA" can be different from
+  the phylogenetic MRCA when the latter is an unnamed clade. ***Does
   physcraper take into account the focal clade from Open Tree in any way???***
@@
 
-- MUSCLE
+- The software MUSCLE [@edgar2004muscle] is implemented for profile alignment, in
+  which the original alignment is used as a template to align all new sequences.
 
@@
 ***- One way to compare two independent phylogenetic trees is to compare them both to
-the synthetic OToL and then measure how well they do agianst each other***
+the synthetic OToL and then measure how well they do against each other***
 
@@
 is not anything any of your collaborators have worked on before.
-A good idea is to start an intensive litt review on the phylogenetics of the group.
-Rapidly, you find out there are 5 different phylogeneis, that used different markers, and that
+A good idea is to start an intensive literature review on the phylogenetics of the group.
+Rapidly, you find out there are 5 different phylogenies, that used different markers, and that
 the papers, published at different times, do not discuss which phylogeny is
@@
 - you can get the study id using R tools:
-  - By using the treebase ID of the study (which is not fully exposed on the
-    Treebase website home page of the study, so you have to really look it up manually):
+  - By using the TreeBase ID of the study (which is not fully exposed on the
+    TreeBase website home page of the study, so you have to really look it up manually):
 ```{r, treebase, eval = TRUE}
@@
 
-1. Download the alignment from Treebase
-If you are on the Treebase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to Treebase and to the Open Tree of Life repository.
+1. Download the alignment from TreeBase
+If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
 To make this task easier, you can use a command to download everything into your working folder:
@@
 ```
-In this example, all alignments posted on Treebase were used to reconstruct both trees.
+In this example, all alignments posted on TreeBase were used to reconstruct both trees.
 

### 2020-05-12 physcraper_ms.Rmd
doc version

@@
+---
+title: 'Physcraper: A Python package for continually updated gene trees'
+date: 'Manuscript rendered on `r format(Sys.time(), "%d %B, %Y")`'
+author:
+  - Luna L. Sanchez Reyes, University of California, Merced
+  - Martha Kandziora, University of California, Merced
+  - Emily Jane McTavish, University of California, Merced
+output:
+  redoc::redoc
+bibliography: paper.bib
+csl: apa.csl
+---
+
 # Abstract
@@
 
-# Intro
+# Introduction
 
@@
 - A ".csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from the tree and alignment. ***What is the minimum amount of tips necessary for a physcraper run??***
+- Unmatched taxon names are dropped from the tree and alignment.
+  ***What is the minimum amount of tips necessary for a physcraper run??*** Just one!
 - A ".tre" and ".aln" files are generated and saved for a  `physcraper` run.
@@
   the phylogenetic MRCA when the latter is an unnamed clade. ***Does
-  physcraper take into account the focal clade from Open Tree in any way???***
+  physcraper take into account the focal clade from Open Tree in any way???*** Not really, it is a search MRCA more than a taxonomic, the default is to get the ingroup from Otol, which is not the focal group.
 - The taxonomic MRCA is used to constrain the DNA sequence search within that taxonomic group on the GenBank database.
@@
   the taxonomic MRCA that are similar to each sequence in the alignment ***How does the blast is setup? Is it an all-blast-all??***
+  The rmote BLAST uses bioPython and the local BLAST uses BLASTn
 - Reverse complement sequences are identified and translated.
+- Sequences found for each
 - A fasta file containing all sequences resulting from the search is generated for the user.
+- ***Add cycles of blast searches with new sequences***
+- How many sequences per species? Was hardcoded to 5, but it will be an argument now, default to 5. ***How does pyphlawd, or phylota does it?***
 
@@
 
-- A gene tree is reconstructed for each alignment provided, using ***what software?***
+- A gene tree is reconstructed for each alignment provided, using RAxML with bootstrap replicates.
+- The final result is a gene tree couples to the conlict info.
 

### 2020-05-13 phylota-overview.Rmd
some more phylota examples

@@
   - a review for ecologists about phylogenetic tools [@roquet2013building]
+  - a study constructing a dated seed plant phylogeny using pyPHLAWD [@smith2018constructing]
+  - a study presenting an assembly and alignment free method for phylogenetic reconstruction
+    using genomic data, that aims to be incorporated in a tool as phylota some day [@fan2015assembly]
+  - nexml format presentation [@vos2012nexml] - cites phylota as a tool that uses
+    stored phyloinformatic data that could benefit from adopting nexml, to increase
+    interoperability
+  - a study of fruit evolution, analysing a previously published phylogeny of 8911
+    tips of the Campanulidae, constructed with PHLAWD [@beaulieu2013fruit]
+  - a study of Southeast Asia plant biodiversity inventory [@webb2010biodiversity] -
+    cites phylota as a tool that would allow rapid phylogentic placing of newly
+    discovered species, and generation of phylogenetically informed guides for field
+    identification.
 
@@
   sampled at different taxonomic levels [@faurby2015species]
-  -
+  - A 900 species tree of muroid rodents, Muroidea [@steppan2017muroid], where 300
+    species were newly added by the study and the rest obtained using phylota.
+  - A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales [@ickert2009fossil]
+  - A 1061 genera phylogeny of the Oscine birds [@selvatti2015paleogene]
+  - A 268 species phylogeny of sharks, representing all orders and 32 families [@sorenson2014effect]
 

### 2020-05-13 physcraper_ms.Rmd
some more phylota examples

@@
 title: 'Physcraper: A Python package for continually updated gene trees'
-date: 'Manuscript rendered on `r format(Sys.time(), "%d %B, %Y")`'
 author:
-  - Luna L. Sanchez Reyes, University of California, Merced
-  - Martha Kandziora, University of California, Merced
-  - Emily Jane McTavish, University of California, Merced
+  - Luna L. Sanchez Reyes
+  - Martha Kandziora
+  - Emily Jane McTavish
+  - University of California, Merced
+date: '`r format(Sys.time(), "%d %B, %Y")`'
+bibliography: paper.bib
+csl: apa.csl
 output:
   redoc::redoc
-bibliography: paper.bib
-csl: apa.csl
 ---
@@
 
-# Tools on a similar track:
+# Tools on a similar track
 
-Tools that do similar things:
-pyPhlawd [@smith2019pyphlawd]
-SUPERSMART [@antonelli2017toward]
+Tools that do similar things at different levels:
+
+Phylota [@sanderson2008phylota] - cited by 122 studies.
+
+PHLAWD [@smith2009mega] and pyPhlawd [@smith2019pyphlawd] - baited analyses
+
+A [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals),
+only available from the [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP)
+of the journal [[@peters2011taming]](https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55#Sec21)
+
+PUmPER [@izquierdo2014pumper] - perpetual updating with newly added sequences to
+GenBank
+
+SUMAC [@freyman2015sumac] -  both “baited” analyses and single‐linkage clustering
+methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix
+
+SUPERSMART [@antonelli2017toward] - baited analyses up to bayesian divergente time
+estimation
+
+SOPHI - [@chesters2017construction] - Searches DNA sequence data from repos other
+than GenBank, such as transcriptomic and barcoding repos.
+
+PhySpeTre [@fang2019physpetree] - no sequence retrieval, just phylogenetic reconstruction
+pipeline.
 

### 2020-05-17 phylota-overview.Rmd
blast methodology

--- 2020-05-07 phylota-overview.Rmd
@@
+---
+output:
+  redoc::redoc
+---
+
 # Phylota overview
@@
 
-1. As an example study of very large and comprehensive phylogeny from minedDNA sequence data, (even if no phylogeny was published there, only the method to do so):
-  - @beaulieu2012modeling
-  -
-
-1. As an example tool that mines GenBak data for phylogenetic reconstruction:
+1. As an example of a tool that mines GenBank data for phylogenetic reconstruction,
+or that is useful in any way for phylogenetics:
     - original publication of PHLAWD [@smith2009mega]
-  -
-
-1. When the software was actually used to construct a DNA data set for phylogenetic reconstruction of:
+    - an analysis identifying research priorities and data requirements for resolving
+    the red algal tree of life [@verbruggen2010data]
+    - @beaulieu2012modeling cites phylota As an example study of very large and comprehensive
+    phylogeny from mined DNA sequence data, (even if no phylogeny was really published
+    there, only the method to do so)
+    - a review for ecologists about phylogenetic tools [@roquet2013building]
+    - a study constructing a dated seed plant phylogeny using pyPHLAWD [@smith2018constructing]
+    - a study presenting an assembly and alignment free method for phylogenetic reconstruction
+    using genomic data, that aims to be incorporated in a tool as phylota some day [@fan2015assembly].
+    - nexml format presentation [@vos2012nexml] - cites phylota as a tool that uses
+    stored phyloinformatic data that could benefit from adopting nexml, to increase
+    interoperability.
+    - a study of fruit evolution, analysing a previously published phylogeny of 8911
+    tips of the Campanulidae, constructed with PHLAWD [@beaulieu2013fruit]
+    - a study of Southeast Asia plant biodiversity inventory [@webb2010biodiversity] -
+    cites phylota as a tool that would allow rapid phylogentic placing of newly
+    discovered species, and generation of phylogenetically informed guides for field
+    identification.
+    - a study of wood density for carbon stock assessments [@flores2011estimating],
+    cites phylota as an initiative to "get supertrees resolved up to species level".
+    - a study proposing something similar to Open tree but applied only to land plants [@beaulieu2012synthesizing]
+    - an analysis of the phylogenetic diversity-area curve [@helmus2012phylogenetic],
+    cited phylota as a method alternative to phylomatic to "obtain plant phylogenetic
+    trees for ecophylogenetic studies".
+    - a study generating a phylogeny of 6,098 species of vascular plants from China
+    [@chen2016tree] - uses DarwinTree [@meng2015darwintree] and generates sequence
+    data *de novo* for 781 genera.
+    - a review of the state of methods and knowledge generated by molecular systematics
+    [@san2010molecular] cites phylota as a tool "intended to systematize GenBank information
+    for large-scale molecular phylogenetics analysis".
+    - the first phylotastic paper [@stoltzfus2013phylotastic] cites phylota as a "phylogeny
+    related resource that provides ways to generate custom species trees for downstream use".
+    - @antonelli2017toward cites phylota as a "pipeline that pre-processes entire GenBank
+    releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
+    then clustered into candidate data sets". I also uses the PHYLOTA database in its
+    own pipeline.
+    - @deepak2014evominer present an algorithm for mining of frequent subtrees (common patterns)
+    in collections of phylogenetic trees, as a way to extract meaningful phylogenetic
+    information from collections of trees when compared to maximum agreement subtrees
+    and majority-rule trees. They cite phylota as one of such tree collections available
+    along with TreeBASE [@piel2009treebase].
+    - @ranwez2009phyloexplorer cites phylota as a "program providing basic statistics
+    on data availability for molecular datasets". They propose a tool to upload and
+    explore user phylogenies to obtain detailed summary statistics on user tree collections.
+    - @freyman2015sumac cites phylota as a tool that "provides a web interface to view
+    all GenBank sequences within ta xonomic groups clustered into homologs" but that
+    does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They
+    compare the performance of SUMAC to Phylota.
+    - @chesters2013resolving cites phylota as a data mining tool that compiles metadata
+    from mining of public DNA databases "for construction of large phylogenetic trees
+    and multiple gene sets" and that the authors have recognised that gene annotations
+    in public databases are insufficient and that careful partitioning of orthologous
+    sequences is needed for supermatrix construction. @chesters2013resolving present
+    a procedure that minimizes the problem of forming multilocus species units in
+    a large phylogenetic data set using algorithms from graph theory.
+1. When the software was actually used to construct (partially or in full) a DNA
+data set to be used for phylogenetic reconstruction:
     - A 1000 tip phylogeny of the family of the nightshades [@sarkinen2013solanaceae]
@@
     - A 321 tip phylogeny of Testudines [@thomson2010sparse]
-
+    - A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish [@martin2011trophic]
+    - A 2,957 taxa phylogeny of the class Moniloformopses of living ferns [@lehtonen2011towards]
+    - A 2,573 species phylogeny of the Papilionoidea [@hardy2014specialization]
+    - A 23 taxa phylogeny of the California flora [@anacker2011origins]
+    - Phylogenies of 6 different clades of flowering plants representing an independent
+    evolutionary origin of extrafloral nectaries: *Byttneria* (Malvaceae), *Pleopeltis* (Polypodiaceae),
+    *Polygoneae* (Polygoneaceae), *Senna* (Fabaceae), *Turnera* (Passifloraceae), and *Viburnum*
+    (Adoxaceae) [@weber2014defense].
+    - To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
+    sampled at different taxonomic levels [@faurby2015species]
+    - A 900 species tree of muroid rodents, Muroidea [@steppan2017muroid], where 300
+    species were newly added by the study and the rest obtained using phylota.
+    - A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales [@ickert2009fossil]
+    - A 1061 genera phylogeny of the Oscine birds [@selvatti2015paleogene]
+    - A 268 species phylogeny of sharks, representing all orders and 32 families [@sorenson2014effect]
+    - A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region [@tucker2012incorporating].
+    - A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects [@hardy2010gall].
+    - A 196 species phylogeny of the family Boraginaceae [@nazaire2012broad]. The authors
+    actually found data for 318 Boraginaceae spp using phylota, but decided to reduce
+    their data set to focus on the monophyly of genus *Mertensia*.
+    - A phylogeny of 401 species of scale insects Coccoidea, Hemiptera [@ross2013large],
+    with some sequences generated *de novo*.
+    - Two phylogenies sampling all species of two different clades of insectivorous
+    lizards, agamids and diplodactyline geckos, groups considered to be radiating
+    in the Australia’s Great Victoria Desert [@rabosky2011species]
+    - A phylogeny of 91 species of sparid and centracanthid fishes, Sparidae, Percomorpha,
+    plus 2 outgroups, a lethrinid and a nemipterid exemplar [@santini2014first].
+    - Updating a phylogeny of Arecaceae, constructing relationships in 6 cldes within
+    the group: subfamilies Calamoideae and Coryphoideae, the tribe Ceroxyleae within
+    subfamily Ceroxyloideae and three groups within subfamily Arecoideae: (1) Iriarteeae,
+    (2) Cocoseae: Attaleinae except Beccariophoenix and (3) a group containing six
+    tribes; Euterpeae, Leopoldinieae, Pelagodoxeae, Manicarieae, Geonomateae and Areceae
+    [@faurby2016all].
+    - A phylogeny of 768 Gesneriaceae species and 58 outgroups for a total species
+    sampling of 826 taxa [@roalson2016distinct] some sequence were generated *de novo*.
+    - A phylogeny of 47 species of scombrid fishes, with 2 outgroups, a gempylid and
+    a trichiurid [@santini2013first].
+    - to update a dataset underlying a large-scale fern phylogeny [@lehtonen2017environmentally],
+    data set in <https://zenodo.org/record/345670#.Xr9QFRPYqqg>, also in TreeBASE,
+    but it is one of those studies that is broken.
+    - A phylogeny of 13 species of billfishes, order Istiophoriformes: Acanthomorpha,
+    and four outgroups [@santini2013first]
+    - A phylogeny of 765 aphid species, family Aphididae [@hardy2015evolution]
+    - A phylogeny of less than 100 taxa of the family Ranunculaceae [@lehtonen2016sensitive],
+    even though they retrieved info from phylota for 194 taxa within the family, they
+    reduced their data set because of low sampling of markers for some taxa.
+    - A phylogeny of 144 neobatrachian genera, assuming the monophyletic status of
+    genera to increase matrix-filling levels [@frazao2015gondwana].
+    - A 179 species phylogeny of the bird family Picidae (woodpeckers, piculets,
+    and wrynecks) [@dufort2016augmented], augmented with data from an updated GenBank
+    release and newly sequenced data.
+    - A phylogeny of species of freshwater fish endemic to NorthAmerica [@strecker2014fish],
+    phylota found data for 54 out of 66 spp.
+    - A phylogeny of 520 species of the order Ericales [@hardy2012testing]
 1. When the website was used to identify sequences and markers available in
-GenBank for a particular group, In this cases, the dataset mining was performed with other tools, or not performed at all and just used for discussion:
-  - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset constructed with PHLAWD
-  - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed by hand (I think??)
-  - A review of dated phylogenies of fire-prone tropical savanna species from Brazil [@simon2012cerrado] -- just for discussion of the lack of markers available for these species on GenBank
-
+GenBank for a particular group. In this cases, the dataset mining was either performed
+with other tools, or not performed at all and just used for discussion:
+    - A 812 tips phylogeny of the Order Chiroptera [@shi2015speciation] -- dataset
+    constructed with PHLAWD
+    - A 1276 tips phylogeny of the Fabaceae [@legume2013legume] -- dataset constructed
+    by hand (I think??)
+    - A review of dated phylogenies of fire-prone tropical savanna species from Brazil
+    [@simon2012cerrado] -- just for discussion of the lack of markers available for
+    these species on GenBank
+    - A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
+    protists [@morrison2009apicomplexa].
+    - Three data sets from phylota (the suborder Pleurodira of side-necked turtles;
+      the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
+      to demonstrate and exemplify phylogenetic decisiveness [@sanderson2010phylogenomics]
 1. Sometimes, it was cited by mistake:
@@
     from previous smaller phylogenies. However, it was removed from the text but
-    not from the final list of references. The DNA data set was constructed by hand most probably.
+    not from the final list of references. The DNA data set was constructed by hand
+    most probably.
+    - a study reconstructing the insect tree of life with 49,358 species, 13,865
+    genera, and 760 families within the order Insecta [@chesters2017construction],
+    uses its own algorithm (SOPHI) to mine public DNA databases [@chesters2014protocol].
+    It does not cite phylota as it should, but includes it in their references.
+1. Other miscellaneous uses of phylota:
+    - @page2013bionames uses it to generate phylogenies for the [bionames website](http://bionames.org),
+    a "database linking taxonomic names to their original descriptions, to taxa, and
+    to phylogenies" generated with phylota.

### 2020-05-17 physcraper_ms.Rmd
blast methodology

@@
 - The alignment should be a gene alignment that was used to generate the tree. The
-  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml]
-  or DRYAD (<http://datadryad.org/>). If the alignment is stored in TreeBase,
-  `physcraper` can download it directly either from the TreeBASE website (<https://treebase.org/>)
-  or through its GitHub repositiry, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
+  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
+  DRYAD (<http://datadryad.org/>), or th ejournal were the tree was originally published.
+  If the alignment is stored in TreeBase, `physcraper` can download it directly
+  either from the TreeBASE website (<https://treebase.org/>)
+  or through TreeBASE GitHub repository, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
   If the alignment is on another repository, a copy of it has to be
-  downloaded by the user, and it's location has to be provided as an argument.
+  downloaded by the user, and it's local path has to be provided as an argument.
 - A taxon name matching step is performed to verify that all taxon names on the tips
@@
 - Unmatched taxon names are dropped from the tree and alignment.
-  ***What is the minimum amount of tips necessary for a physcraper run??*** Just one!
+  <!--***What is the minimum amount of tips necessary for a physcraper run??*** Just one!-->
+  Technically, just one matching name is needed to perform the searches. See below.
 - A ".tre" and ".aln" files are generated and saved for a `physcraper` run.
@@
 
-- The first step is to identify the MRCA of all taxa in the tree that is also a
-  named clade in the NCBI taxonomy. This "taxonomic MRCA" can be different from
-  the phylogenetic MRCA when the latter is an unnamed clade. ***Does
-  physcraper take into account the focal clade from Open Tree in any way???*** Not really, it is a search MRCA more than a taxonomic, the default is to get the ingroup from Otol, which is not the focal group.
-- The taxonomic MRCA is used to constrain the DNA sequence search within that taxonomic group on the GenBank database.
-- The DNA sequence search can be done on a local database that is easily setup by the user, or remotely.
-- The BLAST algorithm is used to identify all DNA sequences in the database within
-  the taxonomic MRCA that are similar to each sequence in the alignment ***How does the blast is setup? Is it an all-blast-all??***
-  The rmote BLAST uses bioPython and the local BLAST uses BLASTn
+- The next step is to identify the search taxon. This must be a taxon (a named clade)
+  from the NCBI taxonomy. It will be used to constraint the DNA sequence search on
+  the GenBank database within that taxonomic group.
+  By default, the search taxon is the most recent common ancestor (MRCA) of the
+  matched taxa that is also a named clade in the NCBI taxonomy. This is refered to as the most recent common
+  ancestral taxon (MRCAT) or the least inclusive common ancestral taxon (LICA).
+  It can be different from the phylogenetic MRCA when the latter is an unnamed clade.
+  This is done using the Open Tree API [taxonomy/mrca](https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca).
+  <!-- ***Does physcraper takes into account the focal clade from Open Tree in any way???***
+  Not really, it is a search MRCA more than a taxonomic, the default is to get the ingroup from Otol, which is not the focal group.-->
+  A search taxon can also be given by the user. It can be a more inclusive
+  clade, if the user wants to perform a wider search, outside the MRCAT of the matched taxa, e.g., including all taxa within
+  the family or the order. It can also be a less inclusive clade, if the user only
+  wants to focus on enriching a particular clade/region within the tree.
+  When only one taxon is matched in both the tree and alignment, an MRCAT can be
+  found for that single taxon, and thus a DNA sequence search can be performed even
+  with only one sequence in the alignment.
+- The BLAST algorithm is used to identify similarity among DNA sequences in the GenBank nucleotide
+  database within the search taxon and the remaining sequences on the alignment.
+  <!--***How does the blast is setup? Is it an all-blast-all??***-->
+- The DNA sequence similarity search can be done on a local database that is easily setup by the user. In this case it uses the BLASTn algorithm.
+- The search can also be performed remotely, using the bioPython BLASt algorithm.
+- A BLAST run is performed for each sequence in the alignment. Results of each BLAST
+  are recorded. All matched sequences are saved with their corresponding GenBank
+  accesion numbers that will be used to download the whole sequences into a local library.
+- Matched sequences below an e-value, percentage similarity, and outside a minimum
+  and maximum length threshold are discarded. This will leave out genomic sequences.
+  All acepted sequences are asigned an internal identifier, and are further filtered.
+- Because we usually do not have the accession number from sequences in the original
+  alignment, a filtering process is needed. Accepted sequences that belong to the
+  same taxon in the query sequence and that are either identical or shorter than
+  the original sequence are also discarded. Only longer sequences belonging to the
+  same taxon as the orignal sequence will be considered for further analyses.
+- Among the remaining filtered sequences, there are usually several exemplars per taxon.
+  Although it can be useful to keep some of them to, for example, investigate monophyly
+  within species, there can be hundreds of exemplar sequences per taxon for some markers.
+  To control the number of sequences per taxon kept for further analyses, by default
+  5 sequences per taxon are chosen at random. This number can be controlled by the user.
+  <!--***How does pyphlawd, or phylota does the exemplar per sepcies choosing?***-->
 - Reverse complement sequences are identified and translated.
-- Sequences found for each
-- A fasta file containing all sequences resulting from the search is generated for the user.
-- ***Add cycles of blast searches with new sequences***
-- How many sequences per species? Was hardcoded to 5, but it will be an argument now, default to 5. ***How does pyphlawd, or phylota does it?***
+- This cycle of sequence search is performed two times. ***Is there an arhument to control the number of cycles of blast searches with new sequences***
+- A fasta file containing all sequences resulting from the BLAST searches is generated for the user.
 
@@
 - A gene tree is reconstructed for each alignment provided, using RAxML with bootstrap replicates.
-- The final result is a gene tree couples to the conlict info.
+- The final result is a gene tree coupled to the conlict info.
 
@@
 
+DarwinTree [@meng2015darwintree] predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP) - takes data from GenBank, EMBL and DDBJ for land plants only.
+
+NCBIminer [@xu2015ncbiminer]
+
 A [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals),
@@
 
+```{r phylota, child="phylota-overview.Rmd"}
+```
+
 # Acknowledgements

### 2020-05-18 phylota-overview.Rmd
other similar tools and phylota examples

@@
     a large phylogenetic data set using algorithms from graph theory.
+    - @chesters2014protocol present an algorithm to delineate species form GenBank
+    DNA data, and cites phylota as a tool that partitions "the contents of a database
+    according to homology", by "grouping of database sequences according to internal
+    criteria", searching "from a standardized set of references [...] patterns in
+    sequence similarity and overlap."
+    - the paper presenting phylotaR, a pipeline that recreates the phylota output
+    but uses the most updated GenBank release, and is available in R [@bennett2018phylotar],
+    cites phylota as its predecessor and inspiration. the authors mention that phylotaR
+    pipeline mimics phylota's pipeline but with improvememnts.
+    - The paper presenging PhyloBase [@jamil2016visual], cites phylota as one of
+    its resources to get phylogenies, along with TreeBASE and others.
+    - The paper presenting STBase, a database of one million precomputed species
+    trees [@mcmahon2015stbase], cites phylota as a databse of gene trees or multrees,
+    "trees having multiple sequences with the same taxon name".
+    - @drori2018onetwotree present a Web‐based, user-friendly, online tool for species-tree
+    reconstruction, based on the *supermatrix paradigm* and retrieves all available
+    sequence data from NCBI GenBank. They cite phylota in the intro as a tool that is "designed to provide
+    users with precomputed sets of clusters that were assembled through a single‐linkage
+    clustering approach and additionally provides precomputed gene trees that were
+    reconstructed for each cluster. In particular, the results obtained by PhyLoTa
+    are taxonomically constrained; that is, all sequences of the most recent common
+    ancestor are collected even if one specifies only part of a clade".
+    - A study developing a tool to link wikipedia data to NCBI taxonomy [@page2011linking]
+    cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
+    - the study that present DarwinTree [@meng2015darwintree] as well as the study
+    presenting an approach to screen sequence data for The Platform
+    for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
+    to parallelize BLAST [@yong2010screening], both cite phylota as one among other
+    "studies based on data mining large numbers of taxa or loci".
 1. When the software was actually used to construct (partially or in full) a DNA
@@
     - A phylogeny of 520 species of the order Ericales [@hardy2012testing]
+    - A phylgeny of 16 fish species of the family Sphyraenidae (Percomorpha), as well
+    as two outgroup species of the Centropomidae (barracudas) [@santini2015first]
+    - A phylogeny of 34 vole species, Arvicolinae, Rodentia [@garcia2016role]
 1. When the website was used to identify sequences and markers available in

### 2020-05-18 physcraper_ms.Rmd
other similar tools and phylota examples

@@
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- The DNA sequence similarity search can be done on a local database that is easily setup by the user. In this case it uses the BLASTn algorithm.
-- The search can also be performed remotely, using the bioPython BLASt algorithm.
-- A BLAST run is performed for each sequence in the alignment. Results of each BLAST
-  are recorded. All matched sequences are saved with their corresponding GenBank
-  accesion numbers that will be used to download the whole sequences into a local library.
+- The DNA sequence similarity search can be done on a local database that is easily
+  setup by the user. In this case it uses the BLASTn algorithm.
+- The search can also be performed remotely, using the bioPython BLAST algorithm.
+- A pairwise all-against-all BLAST search is performed. This means that each sequence
+  in the alignment is BLASTed against DNA sequences in the database within the search
+  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
+  along with their corresponding GenBank accesion numbers. This information will be
+  used later to download the whole sequences into a local library.
 - Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This will leave out genomic sequences.
+  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
   All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because we usually do not have the accession number from sequences in the original
-  alignment, a filtering process is needed. Accepted sequences that belong to the
-  same taxon in the query sequence and that are either identical or shorter than
+- Because the original alignments usually do not have the GenBank accession numbers
+  on the sequence names, a filtering process is needed. Accepted sequences that belong to the
+  same taxon of the query sequence, and that are either identical or shorter than
   the original sequence are also discarded. Only longer sequences belonging to the
@@
 - Reverse complement sequences are identified and translated.
-- This cycle of sequence search is performed two times. ***Is there an arhument to control the number of cycles of blast searches with new sequences***
+- This cycle of sequence search is performed two times. ***Is there an argument to control the number of cycles of blast searches with new sequences***
 - A fasta file containing all sequences resulting from the BLAST searches is generated for the user.
@@
 
+A [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals),
+only available from the [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP)
+of the journal [[@peters2011taming]](https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55#Sec21)
+
+@chesters2014protocol presents an algorithm that mines GenBank data to delineate species in the insecta.
+The authors present a nice comparison with the phylota algorithm.
+
+PUmPER [@izquierdo2014pumper] - perpetual updating with newly added sequences to
+GenBank
+
 DarwinTree [@meng2015darwintree] predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP) - takes data from GenBank, EMBL and DDBJ for land plants only.
@@
 
-A [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals),
-only available from the [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP)
-of the journal [[@peters2011taming]](https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55#Sec21)
-
-PUmPER [@izquierdo2014pumper] - perpetual updating with newly added sequences to
-GenBank
-
 SUMAC [@freyman2015sumac] -  both “baited” analyses and single‐linkage clustering
 methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix
+
+STBase - @mcmahon2015stbase present a pipeline for species tree construction and
+the public database of one million precomputed species trees
 
@@
 than GenBank, such as transcriptomic and barcoding repos.
+
+OneTwoTree [@drori2018onetwotree] present a Web‐based, user-friendly, online tool for species-tree
+reconstruction, based on the *supermatrix paradigm* and retrieves all available
+sequence data from NCBI GenBank.
 

### 2020-05-19 phylota-overview.Rmd
starting to outline discussion

@@
     "studies based on data mining large numbers of taxa or loci".
+    - A review on online plant databases aiming to "provide recommendations for current
+    information managers and developers concerning the user interface and experience;
+    and to provide a picture about the possible directions to take for those in charge
+    of the creation of information at all levels". They cite phylota as a tool allowing
+    researchers "to acces equally and globally, without travel, a [phylogenetic?] model
+    of plants at the kingdom level".
+    - A study presenting a tool to asses gene sequence quality for automatic
+    construction of databases [@meng2012gsqct], as well as their parallelized version using MapReduce
+    [@meng2012cloud], cite phylota (along with @yong2010screening) as a tool that relies on sequence
+    similarity (BLAST) and not taxon name annotations in the database, for mining
+    large numbers of taxa or loci, without making any control on the quality of the
+    sequencing. On the same line
+    - a paper aiming to establish an online information system for the legumes and
+    to outline "best practices for development of a legume portal to enable data
+    sharing and a better understanding of what data are available, missing, or erroneous,
+    and ultimately facilitate cross-analyses and collaboration within the legume-systematics
+    community and with other stakeholders" [@bruneau2019towards], cites phylota (along with supersmart and pyphlawd) as a
+    "pipeline for large-scale retrieval of GenBank data of particular taxa or clades".
+    In their Table 1, they also list phylota as a potential data source for developing a legume portal.
 1. When the software was actually used to construct (partially or in full) a DNA
@@
     - A phylogeny of 34 vole species, Arvicolinae, Rodentia [@garcia2016role]
+    - @kolmann2017dna uses phylota to download all 1691 co1 sequences belonging to
+    the order Carchariniformes, to place phylogenetically DNA samples obtained from
+    fish markets.
+    - A phylogeny of 329 bird species in the Tyrannidae (77% of the species in the
+      family) [@gomez2020speciation]
 1. When the website was used to identify sequences and markers available in
@@
     It does not cite phylota as it should, but includes it in their references.
-1. Other miscellaneous uses of phylota:
+1. When phylota was used to extract full trees (not only DNA data sets or markers):
     - @page2013bionames uses it to generate phylogenies for the [bionames website](http://bionames.org),
@@
     to phylogenies" generated with phylota.
+    - @deepak2013extracting uses a sample of phylota trees to test their method to
+    remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
+    trees with two or more leaves sharing a label, e.g., a species name, which can
+    imply multiple conflicting phylogenetic relationships for the same set of taxa.

### 2020-05-19 physcraper_ms.Rmd
starting to outline discussion

@@
 The process of phylogenetic reconstruction implies many steps that can be generalized to the following:
-1. Obtention of molecular or morphological character data -- get DNA from some organisms and sequence it, or get it from an online repository, such as GenBank.
-1. Assemble a hypothesis of homology -- Create a matrix of your character data, by aligning the sequences, in the case of molecular data.
-1.
-Analyse this hypothesis of homology to infer phylogenetic relationships among the organisms you are studying -- Use different available programs to infer molecular evolution, trees and times of divergence.
-1.
-Discuss the inferred relationships in the context of previous hypothesis, the biology and biogeography of the organisms, etc.
--- Answer the question, *is this phylogenetic solution fair/reasonable?*
+
+1. Obtention of molecular or morphological character data -- get DNA from some organisms
+and sequence it, or get it from an online repository, such as GenBank.   
+1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
+aligning the sequences, in the case of molecular data.
+1. Analyse this hypothesis of homology to infer phylogenetic relationships among
+the organisms you are studying -- Use different available programs to infer molecular
+evolution, trees and times of divergence.
+1. Discuss the inferred relationships in the context of previous hypothesis, the
+biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
 
@@
 ```
+# Discussion
 
-# Tools on a similar track
+Mention statistics provided by PhyloExplorer [@ranwez2009phyloexplorer]
 
-Tools that do similar things at different levels:
+## Tools that do similar things at different levels
+
+PhyloFinder [@chen2008phylofinder] - a search engine for phylogenetic databases using
+trees from TreeBASE - MEH
 
@@
 PHLAWD [@smith2009mega] and pyPhlawd [@smith2019pyphlawd] - baited analyses
+
+PhyloExplorer [@ranwez2009phyloexplorer] - a python and MySQL based website to facilitate
+assessment and management of phylogenetic tree collections.
+It provides "statistics describing the collection,
+correcting invalid taxon names, extracting taxonomically relevant parts of the collection
+using a dedicated query language, and identifying related trees in the TreeBASE database".
 

### 2020-05-20 phylota-overview.Rmd
phylota citations 81-90

@@
     - The paper presenting STBase, a database of one million precomputed species
-    trees [@mcmahon2015stbase], cites phylota as a databse of gene trees or multrees,
+    trees [@mcmahon2015stbase], cites phylota as a databse of gene trees or mul-trees,
     "trees having multiple sequences with the same taxon name".
@@
     cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
-    - the study that present DarwinTree [@meng2015darwintree] as well as the study
+    - the study that present DarwinTree [@meng2015darwintree], the study
     presenting an approach to screen sequence data for The Platform
     for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST [@yong2010screening], both cite phylota as one among other
-    "studies based on data mining large numbers of taxa or loci".
+    to parallelize BLAST [@yong2010screening], as well as [@meng2014rapidtree] all
+    cite phylota as one among other "studies based on data mining large numbers of taxa or loci".
+    - A study presenting a tool to asses gene sequence quality for automatic
+    construction of databases [@meng2012gsqct], as well as their parallelized version using MapReduce
+    [@meng2012cloud], cite phylota (along with @yong2010screening) as a tool that relies on sequence
+    similarity (BLAST) and not taxon name annotations in the database, for mining
+    large numbers of taxa or loci, without making any control on the quality of the
+    sequencing.
     - A review on online plant databases aiming to "provide recommendations for current
@@
     researchers "to acces equally and globally, without travel, a [phylogenetic?] model
-    of plants at the kingdom level".
-    - A study presenting a tool to asses gene sequence quality for automatic
-    construction of databases [@meng2012gsqct], as well as their parallelized version using MapReduce
-    [@meng2012cloud], cite phylota (along with @yong2010screening) as a tool that relies on sequence
-    similarity (BLAST) and not taxon name annotations in the database, for mining
-    large numbers of taxa or loci, without making any control on the quality of the
-    sequencing. On the same line
+    of plants at the kingdom level" [@jones2014trends].
     - a paper aiming to establish an online information system for the legumes and
@@
     In their Table 1, they also list phylota as a potential data source for developing a legume portal.
+    - A study on morphological evolution of electric fish skull, that uses phylotaR
+    to retrieve sequences of the family Apteronotidae, order Gymnotiformes [@evans2019bony],
+    cites phylota as the inspiration and fundament of phylotaR.
+    - A master thesis on SearchTree, a "software tool that allows users to query
+    efficiently on an arbitrary user taxon list and returns high scoring matches
+    from approximately one billion phylogenetic trees being constructed from molecular
+    sequence data in GenBank" [@deepak2010searchtree], that seems to be the preliminary
+    work for STBase [@mcmahon2015stbase], cites phylota as "a standard strategy, to assemble sets of homologous sequences
+    (clusters) from a database of all-against-all BLAST searches, [in which] clusters
+    are constructed in the context of the NCBI taxonomy tree for convenience of display,
+    thus child clusters are contained within parent clusters, following the NCBI hierarchy".
+    In opposition, SearchTree uses true agglomerative hierarchical clustering
+    (AHC: @day1984efficient) based on the BLAST estimates of sequence dissimilarity
+    rather than the NCBI tree".
+    - a recent review on the state of large phylogeny (namely insects) generation
+    using tools of the data-driven era [@chesters2019phylogeny] cites phylota as
+    a tool for homology inference and retrieval.
+    - the study presenting phylotol [@ceron2019phylotol], cites phylota as a tool
+    that "focus on the identification and collection of homologous genes from public
+    databases".
 1. When the software was actually used to construct (partially or in full) a DNA
@@
     - A phylogeny of 329 bird species in the Tyrannidae (77% of the species in the
-      family) [@gomez2020speciation]
+    family) [@gomez2020speciation; @gomez2015behavioral]
+    - Retrive 145 sequences registered as *Holothuria* species, but kept 84 as ingroup,
+    plus 4 outgroup sequences from *Stichopus ocellatus*, all belonging to the order
+    Apodida of sea cucumbers [@kamarudin2016phylogenetic]
+    - On a master thesis, to get the sequences of the outgroups of Melinidinae, family Poaceae, namely several spp of the
+    subfamily Panicoideae, plus *Gynerium sagittatum*, *Chasmanthium latifolium*,
+    and *Zea mays*, [@salariato2010filogenia]. Interestingly, phylota was not used
+    in the published study of the thesis [@salariato2010molecular]. Ingroup sequences were generated *de novo*.
+    - On a PhD thesis, to construct a phylogeny of Platyrrhini (internal group),
+    Catarrhini (outgroup), and Tarsiiformes @pereira2013padroes. Have not found a published study.
 1. When the website was used to identify sequences and markers available in
@@
     imply multiple conflicting phylogenetic relationships for the same set of taxa.
+    - A review by @sanderson2016perspective, takes 134 595 gene trees from phylota
+    GenBank rel. 176 and estimates its degree of resolutin, calculating that less
+    than half of clades are supported with minilam statistical support (0.53 ± 0.32).

### 2020-05-20 physcraper_ms.Rmd
phylota citations 81-90

@@
 
+in @ceron2019phylotol: "pipelines such as AMPHORA (Wu and Eisen 2008) and Mega-phylogeny (Smith et al.
+2009) focus on the construction and refinement of robust alignments rather than the collection of homologs"
+
 PhyloExplorer [@ranwez2009phyloexplorer] - a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections.
-It provides "statistics describing the collection,
+assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
 correcting invalid taxon names, extracting taxonomically relevant parts of the collection
@@
 
-SUPERSMART [@antonelli2017toward] - baited analyses up to bayesian divergente time
+@papadopoulou2015automated - Automated DNA-based plant identification for large-scale biodiversity assessment
+
+SUPERSMART [@antonelli2017toward] - baited analyses up to bayesian divergence time
 estimation
@@
 
+Phylotol [@ceron2019phylotol] - "phylogenomic pipeline to allow easy incorporation
+of data from  high-throughput sequencing studies, to automate production of both
+multiple sequence alignments and gene trees, and to identify and remove contaminants.
+PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree
+of life", i.e., bacteria and unicellular eukaryotes.
+
 ```{r phylota, child="phylota-overview.Rmd"}

### 2020-05-20 physcraper_ms.Rmd
profile alignment et al

@@
 phylogenies by making use of public DNA sequence data and taxonomic information,
-providing a framework for comparison of phylogenies.
+providing a framework for comparison of published phylogenies with their updated versions.
 
@@
 
-- The phylogenetic tree has to be in the Open Tree of Life store [@mctavish2015phylesystem].
-  You can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
+- The phylogenetic tree should be already in the Open Tree of Life store, or submitted
+  via the curator system [@mctavish2015phylesystem].
+  A user can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
   interested in is not in Open Tree of Life, you can easily upload it via the curator
-  tool.
+  tool (ADD URL).
 - The alignment should be a gene alignment that was used to generate the tree. The
   alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD (<http://datadryad.org/>), or th ejournal were the tree was originally published.
+  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
   If the alignment is stored in TreeBase, `physcraper` can download it directly
@@
 - A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the character matrix and vice versa.
+  of the tree are in the DNA character matrix and vice versa.
 - A ".csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from the tree and alignment.
+- Unmatched taxon names are dropped from both the tree and alignment.
   <!--***What is the minimum amount of tips necessary for a physcraper run??*** Just one!-->
@@
 
-- The next step is to identify the search taxon. This must be a taxon (a named clade)
+- The next step is to identify and validate the search taxon. This must be a taxon (a named clade)
   from the NCBI taxonomy. It will be used to constraint the DNA sequence search on
@@
   <!-- ***Does physcraper takes into account the focal clade from Open Tree in any way???***
-  Not really, it is a search MRCA more than a taxonomic, the default is to get the ingroup from Otol, which is not the focal group.-->
+  Not really, the focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
+  In the case that only one taxon is matched in both the tree and alignment, the MRCAT
+  for that single taxon would be determined as HOW?,
   A search taxon can also be given by the user. It can be a more inclusive
@@
   wants to focus on enriching a particular clade/region within the tree.
-  When only one taxon is matched in both the tree and alignment, an MRCAT can be
-  found for that single taxon, and thus a DNA sequence search can be performed even
-  with only one sequence in the alignment.
-- The BLAST algorithm is used to identify similarity among DNA sequences in the GenBank nucleotide
-  database within the search taxon and the remaining sequences on the alignment.
+- The BLAST algorithm is used to identify similarity between DNA sequences in a GenBank nucleotide
+  database within the search taxon, and the remaining sequences on the alignment.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case it uses the BLASTn algorithm.
-- The search can also be performed remotely, using the bioPython BLAST algorithm.
 - A pairwise all-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in the database within the search
+  in the alignment is BLASTed against DNA sequences in a GenBank database constrained to the search
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
   along with their corresponding GenBank accesion numbers. This information will be
-  used later to download the whole sequences into a local library.
+  used later to store the whole sequences into a new library.
+- The DNA sequence similarity search can be done on a local database that is easily
+  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
+- The search can also be performed remotely, on the NCBI database. In this case, the
+  bioPython BLAST algorithm is used to perform the similarity search.
 - Matched sequences below an e-value, percentage similarity, and outside a minimum
@@
 
-- The software MUSCLE [@edgar2004muscle] is implemented for profile alignment, in
-  which the original alignment is used as a template to align all new sequences.
+- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
+- First, all new seweunces are aligned using default MUSCLE options.
+- Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
+  is aligned against the original alignment, working as a template. This ensures that
+  the final alignment follows the homology criteria established by the original alignment.
+- The final alignment is not further processed automatically. We encourage users
+  to check it by eye and eliminate columns with no information.
 

### 2020-05-22 physcraper_ms.Rmd
outline discussion

@@
 
-Mention statistics provided by PhyloExplorer [@ranwez2009phyloexplorer]
+There are many tools that are making use of DNA data repositories in different ways.
+Most of them focus on efficient ways to mine the data -- getting the most homologs.
+Some focus on accurate ways of mining the data - getting real and clean homologs.
+Others focus on refinement of the alignment.
+Most focus on generating full trees *de novo*, mainly for regions of the Tree of
+Life that have no phylogenetic assessment yet in published studies, but also for
+regions that have been already studied and that have phylogenetic data already.
+
+All these tools are great efforts for advancing towards reproducibility in phylogenetics,
+a field that has been largely recognised as somewhat artisanal.
+We propose adding focus to other sources of information available from data repositories.
+DNA data bases have been the focus for long time, but phylogenetic knowledge is
+also accumulating fast in open repositories.
+Certainly, parts of other tools that might work faster in some ways could be incorporated into the physcraper pipeline.
+
+We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
+community to build upon this phylogenetic knowledge, using not only data accumulated in DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
+This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*
+
+Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
+Mention statistics provided by other tools: PhyloExplorer [@ranwez2009phyloexplorer].
+Compare and discuss.
 
@@
 PhyloFinder [@chen2008phylofinder] - a search engine for phylogenetic databases using
-trees from TreeBASE - MEH
+trees from TreeBASE - more related to phylotastic's goal than to updating phylogenies
 

### 2020-05-22 phylota-overview.Rmd
phylota citations 91-100

@@
     - The paper presenting STBase, a database of one million precomputed species
-    trees [@mcmahon2015stbase], cites phylota as a databse of gene trees or mul-trees,
+    trees [@mcmahon2015stbase; @deepak2013managing], cites phylota as a databse of gene trees or mul-trees,
     "trees having multiple sequences with the same taxon name".
@@
     for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST [@yong2010screening], as well as [@meng2014rapidtree] all
-    cite phylota as one among other "studies based on data mining large numbers of taxa or loci".
+    to parallelize BLAST [@yong2010screening], as well as @gao2011solution, @li2013partfasttree, and @meng2014rapidtree, all
+    cite phylota using the exact same phrase, as one among other "studies based on data mining large numbers of taxa or loci".
     - A study presenting a tool to asses gene sequence quality for automatic
@@
     databases".
+    - The [iPTOL project](https://www.researchgate.net/profile/Douglas_Soltis/publication/228815637_Assembling_the_Tree_of_Life_for_the_Plant_Sciences_iPTOL/links/56a7c6bc08aeded22e3700dc.pdf)
+    cites phylota as a resource of phylogenetic trees.
 1. When the software was actually used to construct (partially or in full) a DNA
@@
     - A 1061 genera phylogeny of the Oscine birds [@selvatti2015paleogene]
-    - A 268 species phylogeny of sharks, representing all orders and 32 families [@sorenson2014effect]
+    - A 268 species phylogeny of sharks, representing all 8 orders and 32 families [@sorenson2014effect; @sorenson2014evolution]
     - A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region [@tucker2012incorporating].
@@
     Catarrhini (outgroup), and Tarsiiformes @pereira2013padroes. Have not found a published study.
+    - The 10k trees project [@arnold201010ktrees] uses phylota to construct a tree of 301 primate species
+    and the outgroup species *Galeopterus variegates*, a tree of 17 extant odd-toed
+    ungulates species and the outgroup species *Bos taurus*, and a tree of 70 different
+    species of carnivorans and *Equus caballus* as outgroup. However, the do not
+    cite it on the paper, but only on their documentation <http://www.academia.edu/download/49690788/10kTrees_Documentation.pdf>.
 1. When the website was used to identify sequences and markers available in
@@
       to demonstrate and exemplify phylogenetic decisiveness [@sanderson2010phylogenomics]
+    - Mentioned in a PHD thesis [@gagnon2016systematique], but not on the final publication [@gagnon2016new],
+    phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
 1. Sometimes, it was cited by mistake:

### 2020-05-23 phylota-overview.Rmd
phylota citations 101-110

@@
     does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They
-    compare the performance of SUMAC to Phylota.
+    compare the performance of SUMAC to Phylota. This is also presented in their PhD dissertation [@freyman2017phylogenetic].
     - @chesters2013resolving cites phylota as a data mining tool that compiles metadata
@@
     for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST [@yong2010screening], as well as @gao2011solution, @li2013partfasttree, and @meng2014rapidtree, all
-    cite phylota using the exact same phrase, as one among other "studies based on data mining large numbers of taxa or loci".
+    to parallelize BLAST [@yong2010screening], as well as @gao2011solution, @li2013partfasttree,
+    @meng2015sotree, and @meng2014rapidtree, all cite phylota using the exact same
+    introduction and sentence: as one among other "studies based on data mining large numbers of taxa or loci".
     - A study presenting a tool to asses gene sequence quality for automatic
@@
     cites phylota as a resource of phylogenetic trees.
-1. When the software was actually used to construct (partially or in full) a DNA
+    - @mahmood2015avian PhD dissertation presents a database of avian Raptor sequences
+    (raptorbase), based on the phylota pipeline.
+    - @ruiz2019datataxa develops datataxa and cite phylota as "software that has
+    been developed to mine the massive amount of information stored in GenBank", along with its R version [phylotaR; @bennett2018phylotar] and restez <https://www.rdocumen-tation.org/packages/restez/versions/1.0.0>.
+2. When the software was actually used to construct (partially or in full) a DNA
 data set to be used for phylogenetic reconstruction:
@@
     - A 179 species phylogeny of the bird family Picidae (woodpeckers, piculets,
-    and wrynecks) [@dufort2016augmented], augmented with data from an updated GenBank
+    and wrynecks) [@dufort2016augmented; @dufort2015coexistence], augmented with data from an updated GenBank
     release and newly sequenced data.
@@
     cite it on the paper, but only on their documentation <http://www.academia.edu/download/49690788/10kTrees_Documentation.pdf>.
-1. When the website was used to identify sequences and markers available in
+    - @freyman2015sumac [also in @freyman2017phylogenetic], use phylota to construct
+    a phylogeny (or maybe only mine genbank???) of the Onagraceae and Lythracea,
+    and compare it to the tool they propose, SUMAC.
+    - @blackmon2017synthesis PhD study applies phylota to reconstruct a 822 mite
+    species tree.
+    - A study of the effect of poliploidy on niche evolution [@baniaga2018polyploid],
+    uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
+    16 families and 25 genera,and a tree of 33 genera from 20 different families
+    comprising 1706 taxa.
+3. When the website was used to identify sequences and markers available in
 GenBank for a particular group. In this cases, the dataset mining was either performed
@@
     phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
-1. Sometimes, it was cited by mistake:
+4. Sometimes, it was cited by mistake:
     - In this 630 tip phylogeny of the Caryophyllaceae study [@greenberg2011caryophyllaceae] it might have been originally
@@
     It does not cite phylota as it should, but includes it in their references.
-1. When phylota was used to extract full trees (not only DNA data sets or markers):
+5. When phylota was used to extract full trees (not only DNA data sets or markers):
     - @page2013bionames uses it to generate phylogenies for the [bionames website](http://bionames.org),

### 2020-05-23 phylota-overview.Rmd
finished reviewing phylota citations :tada:

@@
     but uses the most updated GenBank release, and is available in R [@bennett2018phylotar],
-    cites phylota as its predecessor and inspiration. the authors mention that phylotaR
-    pipeline mimics phylota's pipeline but with improvememnts.
+    cites phylota as its predecessor and inspiration. The authors mention that phylotaR
+    pipeline mimics phylota's pipeline but with improvements.
     - The paper presenging PhyloBase [@jamil2016visual], cites phylota as one of
@@
     cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
-    - the study that present DarwinTree [@meng2015darwintree], the study
+    - the study that present DarwinTree [@meng2015darwintree], and all derived studies: the study
     presenting an approach to screen sequence data for The Platform
@@
     to parallelize BLAST [@yong2010screening], as well as @gao2011solution, @li2013partfasttree,
-    @meng2015sotree, and @meng2014rapidtree, all cite phylota using the exact same
+    @meng2014rapidtree, @meng2015sotree, and @meng2015solution, all cite phylota using the exact same
     introduction and sentence: as one among other "studies based on data mining large numbers of taxa or loci".
@@
     cites phylota as the inspiration and fundament of phylotaR.
+    - A phylogenetic revision of the Gymnotidae fish (Teleostei: Gymnotiformes),
+    uses phylotaR to retrieve seqeunces, but cites phylota as "a pipeline that implements
+    BLAST searches to both identify and download sequence clusters for listed taxonomic
+    groups to assemble a robust collection of sequences in a reproducible way based
+    on publicly-available gene sequences while avoiding selection bias on the part
+    of the assembler".
     - A master thesis on SearchTree, a "software tool that allows users to query
@@
     - @ruiz2019datataxa develops datataxa and cite phylota as "software that has
-    been developed to mine the massive amount of information stored in GenBank", along with its R version [phylotaR; @bennett2018phylotar] and restez <https://www.rdocumen-tation.org/packages/restez/versions/1.0.0>.
+    been developed to mine the massive amount of information stored in GenBank",
+    along with its R version [phylotaR; @bennett2018phylotar] and restez <https://www.rdocumen-tation.org/packages/restez/versions/1.0.0>.
+    - The phylotastic project [@stoltzfus2013phylotastic] cites phylota as a "phylogeny-related resource providing ways to generate custom species trees *de novo* for downstream use" along with CIPRES.
 2. When the software was actually used to construct (partially or in full) a DNA

### 2020-05-28 physcraper_ms.Rmd
phylota citations 101-110
summarizing phylota papers for discussion

@@
 - The final alignment is not further processed automatically. We encourage users
-  to check it by eye and eliminate columns with no information.
+  to check it using any of the many tools for alignment processing, and eliminate
+  columns with no information.
 
@@
 
-- A gene tree is reconstructed for each alignment provided, using RAxML with bootstrap replicates.
+- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with bootstrap replicates.
 - The final result is a gene tree coupled to the conlict info.
@@
 
+Data repositories hold more information than meets the eye.
+Besides the actual data, they have other types of information that can be used for the advantage of science.
+
+Usually, initial ideas about the data are changed by analyses.
+We expect that this new ideas on the data can be registered on data bases,
+exposing new comers to expert understanding about the data.
+
 There are many tools that are making use of DNA data repositories in different ways.
@@
 We propose adding focus to other sources of information available from data repositories.
-DNA data bases have been the focus for long time, but phylogenetic knowledge is
-also accumulating fast in open repositories.
-Certainly, parts of other tools that might work faster in some ways could be incorporated into the physcraper pipeline.
+Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
+also accumulating fast in public and open repositories.
+In this way, the physcraper pipeline can be complemented with other tools that have
+been developed for other purposes.
 
 We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon this phylogenetic knowledge, using not only data accumulated in DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*
+community to build upon phylogenetic knowledge, using not only data accumulated in
+DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
+This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.
 
@@
 
+How is physcraper already useful:
+- mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think.
+-
+
+How can it be used for the advantage of the field:
+- rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
+- obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
+- one day could be used to sistematize GenBank data, as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
+- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
+
+Things that physcraper does not do:
+- analyse the whole GenBank database to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward. There are already some very good tools that do that.
+- provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
+- it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
+
 ## Tools that do similar things at different levels
 
-PhyloFinder [@chen2008phylofinder] - a search engine for phylogenetic databases using
+### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
+
+| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
+|-----|------|:------:|:------:|:------:|
+| Phylota | @sanderson2008phylota | 122 studies | finding homologs on GenBank database | Supermatrix |
+| AMPHORA | @wu2008simple | 458 studies | mines public genomic data and constructs phylogenies using whole genomic data | Supermatrix?? |
+| PHLAWD | @smith2009mega | 234 studies | Baited search | Supermatrix |
+| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
+| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
+The authors present a nice comparison with the phylota algorithm | Species trees??
+|
+| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
+| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
+| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
+|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
+|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
+| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
+| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
+| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
+| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
+| pyPhlawd | @smith2019pyphlawd | 6 studies | baited analyses | Supermatrix or gene tree |
+| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
+PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
+
+According to @ceron2019phylotol, both AMPHORA and PHLAWD
+"focus on the construction and refinement of robust alignments rather than the collection of homologs."
+
+### 2. Searching phylogenetic tree databases
+
+PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
 trees from TreeBASE - more related to phylotastic's goal than to updating phylogenies
 
-Phylota [@sanderson2008phylota] - cited by 122 studies.
+### 3. Mining phylogenetic tree databases
 
-PHLAWD [@smith2009mega] and pyPhlawd [@smith2019pyphlawd] - baited analyses
-
-in @ceron2019phylotol: "pipelines such as AMPHORA (Wu and Eisen 2008) and Mega-phylogeny (Smith et al.
-2009) focus on the construction and refinement of robust alignments rather than the collection of homologs"
-
-PhyloExplorer [@ranwez2009phyloexplorer] - a python and MySQL based website to facilitate
+PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
 assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
@@
 
-A [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals),
-only available from the [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP)
-of the journal [[@peters2011taming]](https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55#Sec21)
+### 4. Pipeline for phylogenetic reconstruction
 
-@chesters2014protocol presents an algorithm that mines GenBank data to delineate species in the insecta.
-The authors present a nice comparison with the phylota algorithm.
-
-PUmPER [@izquierdo2014pumper] - perpetual updating with newly added sequences to
-GenBank
-
-DarwinTree [@meng2015darwintree] predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP) - takes data from GenBank, EMBL and DDBJ for land plants only.
-
-NCBIminer [@xu2015ncbiminer]
-
-SUMAC [@freyman2015sumac] -  both “baited” analyses and single‐linkage clustering
-methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix
-
-STBase - @mcmahon2015stbase present a pipeline for species tree construction and
-the public database of one million precomputed species trees
-
-@papadopoulou2015automated - Automated DNA-based plant identification for large-scale biodiversity assessment
-
-SUPERSMART [@antonelli2017toward] - baited analyses up to bayesian divergence time
-estimation
-
-SOPHI - [@chesters2017construction] - Searches DNA sequence data from repos other
-than GenBank, such as transcriptomic and barcoding repos.
-
-OneTwoTree [@drori2018onetwotree] present a Web‐based, user-friendly, online tool for species-tree
-reconstruction, based on the *supermatrix paradigm* and retrieves all available
-sequence data from NCBI GenBank.
-
-PhySpeTre [@fang2019physpetree] - no sequence retrieval, just phylogenetic reconstruction
+PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
 pipeline.
 
-Phylotol [@ceron2019phylotol] - "phylogenomic pipeline to allow easy incorporation
-of data from  high-throughput sequencing studies, to automate production of both
-multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree
-of life", i.e., bacteria and unicellular eukaryotes.
+### 5. getting metadata and not sequences from GenBank.
+
+Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank seqeunce information.
+
+## Phylota overview
 

### 2020-05-29 physcraper_ms.Rmd
more automatization tools

@@
 
-1. Physcraper implements node by node comparison of the the original and the updated
-trees using the conflict API of OToL.
+1. Physcraper implements node by node/topology comparison of the the original and the updated
+trees using the conflict API of OToL, and summarizes differences.
 
@@
 Phylogenies are important.
-Generating phylogenies is not easy.
-The process of phylogenetic reconstruction implies many steps that can be generalized to the following:
+
+Generating phylogenies is not easy and it is largely artisanal.
+Although many efforts to automatize the process have been done, and the community is using those more and more, iautomatizatio of phylogeentic reconstruction is still not a widespread practice and it would be important for reproducibility.
+***paragraph better to end discussion???
+***
+
+The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
 
@@
 1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
-aligning the sequences, in the case of molecular data.
+aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
 1. Analyse this hypothesis of homology to infer phylogenetic relationships among
@@
 Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-They are crucial to solve problems such as food security, global warming, global health.
+Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
 There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
@@
 
-A way to automatize the comparison of phylogenetic hypotheses and to allow reproducibility of the last step of the process.
+A way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
 
@@
 Notably large phylogenies have been constructed using some of these tools,
-Some other have not been used that much. So, how well accepted is this approach in the community?
+Some others have not been used that much. So, how well accepted is this approach in the community?
 
-Concerns with these tools:
-Errors in identification of sequences
-Little control along the process
-Too much of a black box?
+Concerns I think people have about these tools:
+- Errors in identification of sequences
+- Little control along the process
+- Too much of a black box?
 
@@
 
-All these pipelines start tree construction from zero?
+All these pipelines start tree construction from zero? Yes.
 
@@
   All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually do not have the GenBank accession numbers
-  on the sequence names, a filtering process is needed. Accepted sequences that belong to the
+- Because the original alignments usually lack GenBank accession numbers, a filtering
+  step is needed. Accepted sequences that belong to the
   same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are also discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered for further analyses.
+  the original sequence are discarded. Only longer sequences belonging to the
+  same taxon as the orignal sequence will be considered further for analysis.
 - Among the remaining filtered sequences, there are usually several exemplars per taxon.
@@
   within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon kept for further analyses, by default
-  5 sequences per taxon are chosen at random. This number can be controlled by the user.
+  To control the number of sequences per taxon in downstream analyses,
+  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
   <!--***How does pyphlawd, or phylota does the exemplar per sepcies choosing?***-->
 - Reverse complement sequences are identified and translated.
-- This cycle of sequence search is performed two times. ***Is there an argument to control the number of cycles of blast searches with new sequences***
-- A fasta file containing all sequences resulting from the BLAST searches is generated for the user.
+- This cycle of sequence search is performed once. Users can BLAST newly found sequences. ***Is there an argument to control the number of cycles of blast searches with new sequences?***
+- A local database of full sequences is downloaded to the software directory, and
+ a fasta file containing filtered and processed sequences resulting from the BLAST
+ search is generated for the user.
 
@@
 - The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
-- First, all new seweunces are aligned using default MUSCLE options.
+- First, all new sequences are aligned using default MUSCLE options.
 - Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
@@
 
-## Tools that do similar things at different levels
+## Tools that automatize any part of the process of phylogenetic reconstruction:
 
@@
 |-----|------|:------:|:------:|:------:|
-| Phylota | @sanderson2008phylota | 122 studies | finding homologs on GenBank database | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | mines public genomic data and constructs phylogenies using whole genomic data | Supermatrix?? |
-| PHLAWD | @smith2009mega | 234 studies | Baited search | Supermatrix |
+| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
+| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
+| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
 | Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
+| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
 | Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
@@
 | Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
+| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
 | SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
 | SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
+| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
 | OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
-| pyPhlawd | @smith2019pyphlawd | 6 studies | baited analyses | Supermatrix or gene tree |
+| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
 | Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
@@
 
-According to @ceron2019phylotol, both AMPHORA and PHLAWD
+According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
+of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
 "focus on the construction and refinement of robust alignments rather than the collection of homologs."

### 2020-06-01 physcraper_ms.Rmd
ilex example

@@
 
-# Use case/ example
+# Examples
+
+## The hollies
+
+The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
+It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
+The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
+The second has not been made available anywhere, not even in supplementary data of the journal.
+***Contact authors? They seem old school, probably do not wanna share their data.***
+The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
+it has not been added to the syntehtic tree yet.
+This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on GenBank.
+Updating the oldest tree, we should get something very similar to the newest tree.
 
@@
 the one accepted by the expert community on X. You might need
-to go to the annual conference of X, and even then, you might only find
-different and contrasting opinions.
-Somewhere along these months or even years doing this task, you looked into the
-the OToL database. You found in there some or all the published trees of X, along
-with a tree that has been deemed the best tree by curators and ideally experts on X?
+to go to the annual conference of X, and probably you will only find
+different and contrasting opinions in there, too.
+Somewhere along the months or even years you've been on this task, you landed into the
+the OToL website. You found in there some or all the published trees of X, along
+with a tree that has been labeled as the best tree by volunteered curators (and ideally experts on X). What do you do?
 
-## Ascomycota Example
+## The Ascomycota
 

### 2020-06-03 physcraper_ms.Rmd
better intro

@@
 
+Various efforts to automatize parts of the process in order to be able to take advantage of data in public databases.
+
+Most of this tools were envisioned for phylogenomics, were the analyses and processing of huge amounts of genetic data is naturally needed.
+Without automatization, curation of a phylogenomics database is a task that could not be finished in human time.
+Some pipelines sacrificed precision and accuracy for quantity. But many tools to check
+Examples of pipelines for phylogenomics:
+BIR [Blast, Identify, Realign; @kumar2015bir] takes curated alignments as seed to find more data on genomic databases and create new phylogenomic hypotheses
+
+Among other things, one of the major benefits stemming from these efforts are pipelines that make it easier to implement a reproducibile workflow in phylogenetic studies.
+
+A widespread practice is a thorough visual inspection and manual curation of alignments, that are checked for "errors" that can usually result from similarity algorithms ***(give examples)***.
+Commonly, alignment curators also take into account how the nature of the genetic marker, the taxon sampling and the organisms themselves, might affect the alignments.
+This hypotheses of homology are charged with expert knowledge, and hours of study, making them a scientifically reliable source if biological information.
+As such, they represent an ideal platform to build up new phylogenetic knowledge.
+Not many studies report its use.
+Which might be an indication that the amount of manual effort that is required to curate gene alignments is still a bottleneck.
+
 Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, iautomatizatio of phylogeentic reconstruction is still not a widespread practice and it would be important for reproducibility.
+Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and it would be key for adoption of bettr reproducibility practices in the scientific community.
 ***paragraph better to end discussion???
@@
 
-A way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
+We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
 
@@
 Updating the oldest tree, we should get something very similar to the newest tree.
-
-Imagine you are starting to work on a new biological group X. You have not much of an idea about
-its phylogenetic relationships, you are a newly established researcher, and the group
-is not anything any of your collaborators have worked on before.
-A good idea is to start an intensive literature review on the phylogenetics of the group.
-Rapidly, you find out there are 5 different phylogenies, that used different markers, and that
-the papers, published at different times, do not discuss which phylogeny is
-the one accepted by the expert community on X. You might need
-to go to the annual conference of X, and probably you will only find
-different and contrasting opinions in there, too.
-Somewhere along the months or even years you've been on this task, you landed into the
-the OToL website. You found in there some or all the published trees of X, along
-with a tree that has been labeled as the best tree by volunteered curators (and ideally experts on X). What do you do?
 

### 2020-06-03 physcraper_ms.Rmd
ingroup is used to determine taxonomic  mrca

@@
 Which might be an indication that the amount of manual effort that is required to curate gene alignments is still a bottleneck.
+
+*** Introduce the Open Tree of Life***
 
@@
 
-- The phylogenetic tree should be already in the Open Tree of Life store, or submitted
-  via the curator system [@mctavish2015phylesystem].
-  A user can choose from a variety of published trees supporting any node of the Tree of Life. If the tree you are
-  interested in is not in Open Tree of Life, you can easily upload it via the curator
-  tool (ADD URL).
+- The study tree should be in the OToL store [phylesystem; @mctavish2015phylesystem]. The main
+  reason for this is that trees in OToL have a set of user defined characteristics
+  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup.
+  A user can choose from a variety of published trees supporting any node of the synthetic OToL. If the tree you are
+  interested in is not in there, you should upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
 - The alignment should be a gene alignment that was used to generate the tree. The
@@
   DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, `physcraper` can download it directly
+  If the alignment is stored in TreeBase, `physcraper` can download it directly,
   either from the TreeBASE website (<https://treebase.org/>)
-  or through TreeBASE GitHub repository, SuperTreeBASE (<https://github.com/TreeBASE/supertreebase>).
-  If the alignment is on another repository, a copy of it has to be
+  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
+  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
   downloaded by the user, and it's local path has to be provided as an argument.
@@
 - Unmatched taxon names are dropped from both the tree and alignment.
-  <!--***What is the minimum amount of tips necessary for a physcraper run??*** Just one!-->
-  Technically, just one matching name is needed to perform the searches. See below.
-- A ".tre" and ".aln" files are generated and saved for a `physcraper` run.
+  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
+  Technically, just one matching name is needed to perform the searches. Please, see next section.
+- A ".tre" and ".aln" files are generated and saved as input for a `physcraper` run.
 
@@
 - The next step is to identify and validate the search taxon. This must be a taxon (a named clade)
-  from the NCBI taxonomy. It will be used to constraint the DNA sequence search on
+  on the NCBI taxonomy. It will be used to constraint the DNA sequence search on
   the GenBank database within that taxonomic group.
-  By default, the search taxon is the most recent common ancestor (MRCA) of the
-  matched taxa that is also a named clade in the NCBI taxonomy. This is refered to as the most recent common
-  ancestral taxon (MRCAT) or the least inclusive common ancestral taxon (LICA).
-  It can be different from the phylogenetic MRCA when the latter is an unnamed clade.
-  This is done using the Open Tree API [taxonomy/mrca](https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca).
-  <!-- ***Does physcraper takes into account the focal clade from Open Tree in any way???***
-  Not really, the focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  In the case that only one taxon is matched in both the tree and alignment, the MRCAT
-  for that single taxon would be determined as HOW?,
+  By default, the search taxon is the Most Recent Common Ancestor (MRCA) of the
+  matched taxa belonging to the ingroup, that is also a named clade in the NCBI taxonomy. This is refered to as the Most Recent Common
+  Ancestral Taxon (MRCAT; also reffered the Least Inclusive Common Ancestral Taxon - LICA).
+  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade.
+  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
+  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT
+  for that single taxon would be determined as ***HOW?***
+  By default, the set of taxon names that are searched for their MRCAT to be used as search taxon belong to the ingroup defined by the curator in the OToL phylesystem.
+  <!-- ***Physcraper does not take into account the focal clade from Open Tree.
+Because that is arbitrarily assigned or not by a curator.
+The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
   A search taxon can also be given by the user. It can be a more inclusive
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-  <!--***How does pyphlawd, or phylota does the exemplar per sepcies choosing?***-->
+  <!--***How does pyphlawd, or phylota does the exemplar per species choosing?***-->
 - Reverse complement sequences are identified and translated.
-- This cycle of sequence search is performed once. Users can BLAST newly found sequences. ***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- A local database of full sequences is downloaded to the software directory, and
- a fasta file containing filtered and processed sequences resulting from the BLAST
- search is generated for the user.
+- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
+This can be done iteratively, but by default only sequences in the alignment are blasted.
+***Is there an argument to control the number of cycles of blast searches with new sequences?***
+- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
+- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
 

### 2020-06-13 physcraper_ms.Rmd
genbank and blast cittaions

@@
 1. Obtention of molecular or morphological character data -- get DNA from some organisms
-and sequence it, or get it from an online repository, such as GenBank.
+and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
 1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
@@
 In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository,
+pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
 to obtain homologous characters that can be used for phylogenetic reconstruction.
@@
 
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank.
+Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
 
@@
 
-It is well known that GenBank holds enormous amounts of genetic data, and it continues to grow.
+It is well known that GenBank [@benson2000genbank; @wheeler2000database] holds enormous amounts of genetic data, and it continues to grow.
 A lot of this genetic data has the potential to be used to reconstruct the phylogenetic
@@
 The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison of existing phylogenies to phylogenies that are constructed
-using new genetic data from GenBank
+allowing a direct comparison between existing phylogenies and phylogenies that are constructed
+using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).
 
@@
   on the NCBI taxonomy. It will be used to constraint the DNA sequence search on
-  the GenBank database within that taxonomic group.
+  the nucleotide database within that taxonomic group.
   By default, the search taxon is the Most Recent Common Ancestor (MRCA) of the
@@
   wants to focus on enriching a particular clade/region within the tree.
-- The BLAST algorithm is used to identify similarity between DNA sequences in a GenBank nucleotide
-  database within the search taxon, and the remaining sequences on the alignment.
+- The Basic local alignment search tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
+  similarity between DNA sequences within the search taxon in a nucleotide
+  database, and the remaining sequences on the alignment.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
 - A pairwise all-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a GenBank database constrained to the search
+  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding GenBank accesion numbers. This information will be
+  along with their corresponding ide numbers (accesion numbers in the case of the GenBank database). This information will be
   used later to store the whole sequences into a new library.
@@
   All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack GenBank accession numbers, a filtering
+- Because the original alignments usually lack database id numbers, a filtering
   step is needed. Accepted sequences that belong to the
@@
 it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on GenBank.
+This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
 Updating the oldest tree, we should get something very similar to the newest tree.
@@
 - obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
-- one day could be used to sistematize GenBank data, as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
+- one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
 - allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
@@
 Things that physcraper does not do:
-- analyse the whole GenBank database to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward. There are already some very good tools that do that.
+- analyse the whole GenBank database [@benson2000genbank; @wheeler2000database] to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward.
+There are already some very good tools that do that.
 - provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
@@
 PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating phylogenies
+trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
 
@@
 
-Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank seqeunce information.
+Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.
 
@@
 
+The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
+
 # References

### 2020-06-14 physcraper_ms.Rmd
root methods

@@
 
-Various efforts to automatize parts of the process in order to be able to take advantage of data in public databases.
-
-Most of this tools were envisioned for phylogenomics, were the analyses and processing of huge amounts of genetic data is naturally needed.
-Without automatization, curation of a phylogenomics database is a task that could not be finished in human time.
-Some pipelines sacrificed precision and accuracy for quantity. But many tools to check
-Examples of pipelines for phylogenomics:
-BIR [Blast, Identify, Realign; @kumar2015bir] takes curated alignments as seed to find more data on genomic databases and create new phylogenomic hypotheses
-
-Among other things, one of the major benefits stemming from these efforts are pipelines that make it easier to implement a reproducibile workflow in phylogenetic studies.
-
-A widespread practice is a thorough visual inspection and manual curation of alignments, that are checked for "errors" that can usually result from similarity algorithms ***(give examples)***.
-Commonly, alignment curators also take into account how the nature of the genetic marker, the taxon sampling and the organisms themselves, might affect the alignments.
-This hypotheses of homology are charged with expert knowledge, and hours of study, making them a scientifically reliable source if biological information.
-As such, they represent an ideal platform to build up new phylogenetic knowledge.
-Not many studies report its use.
-Which might be an indication that the amount of manual effort that is required to curate gene alignments is still a bottleneck.
-
-*** Introduce the Open Tree of Life***
-
 Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and it would be key for adoption of bettr reproducibility practices in the scientific community.
+Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
 ***paragraph better to end discussion???
@@
 
-It is well known that GenBank [@benson2000genbank; @wheeler2000database] holds enormous amounts of genetic data, and it continues to grow.
-A lot of this genetic data has the potential to be used to reconstruct the phylogenetic
-history of various organisms [@sanderson2008phylota].
-Pipelines that harness this potential have been available for over a decade now, such
-as the Phylota browser, and PHLAWD. New ones keep on being developed, such as SUPERSMART
-and the upgraded version of PHLAWD, PyPHLAWD.
-Notably large phylogenies have been constructed using some of these tools,
-Some others have not been used that much. So, how well accepted is this approach in the community?
-
 Concerns I think people have about these tools:
@@
 
-- The study tree should be in the OToL store [phylesystem; @mctavish2015phylesystem]. The main
-  reason for this is that trees in OToL have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup.
-  A user can choose from a variety of published trees supporting any node of the synthetic OToL. If the tree you are
-  interested in is not in there, you should upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
-- The alignment should be a gene alignment that was used to generate the tree. The
+- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
+  reason for this is that trees in phylesystem have a set of user defined characteristics
+  that are essential for automatizing the phylogeny update process.
+The most relevant of these being the definition of ingroup and outgroup.
+Outgroup and ingroup taxa in the original tree are identified and tagged.
+This allows to automatically set the root for the updated tree on the next steps of the pipeline.
+  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
+If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
+- The alignment should be a gene alignment that was used to generate the tree. The original
   alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
@@
   Technically, just one matching name is needed to perform the searches. Please, see next section.
-- A ".tre" and ".aln" files are generated and saved as input for a `physcraper` run.
+- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.
 
@@
 
-- The next step is to identify and validate the search taxon. This must be a taxon (a named clade)
-  on the NCBI taxonomy. It will be used to constraint the DNA sequence search on
-  the nucleotide database within that taxonomic group.
-  By default, the search taxon is the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup, that is also a named clade in the NCBI taxonomy. This is refered to as the Most Recent Common
-  Ancestral Taxon (MRCAT; also reffered the Least Inclusive Common Ancestral Taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade.
+- The next step is to identify the search taxon within the reference taxonomy.
+The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
+Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
+The search taxon can be provided by the user.
+If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
+  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred to as the Least Inclusive Common Ancestral Taxon - LICA).
+  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
   To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
+
   In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT
   for that single taxon would be determined as ***HOW?***
-  By default, the set of taxon names that are searched for their MRCAT to be used as search taxon belong to the ingroup defined by the curator in the OToL phylesystem.
   <!-- ***Physcraper does not take into account the focal clade from Open Tree.
@@
 The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  A search taxon can also be given by the user. It can be a more inclusive
+  If the search taxon is provided by the user, it can be a more inclusive
   clade, if the user wants to perform a wider search, outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order. It can also be a less inclusive clade, if the user only
-  wants to focus on enriching a particular clade/region within the tree.
+  the family or the order of the ingroup. It can also be a less inclusive clade, if the user only
+  wants to focus on enriching a particular clade/region within the ingroup.
 - The Basic local alignment search tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
@@
 
-- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with bootstrap replicates.
-- The final result is a gene tree coupled to the conlict info.
+- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with 100 bootstrap replicates.
+- The root of the tree is determined by finding the parent node of the sequences that do not belong to the ingroup.
+This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
+- The final result is a tree for eac of the genes provided in the alignment.
 
@@
 
+- Gene trees are then coupled to conlict info.
 - Conflict information can only be generated in the context of the whole Open Tree of

### 2020-06-23 physcraper_ms.Rmd
blast command tools

@@
   matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred to as the Least Inclusive Common Ancestral Taxon - LICA).
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
@@
 The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  If the search taxon is provided by the user, it can be a more inclusive
-  clade, if the user wants to perform a wider search, outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order of the ingroup. It can also be a less inclusive clade, if the user only
-  wants to focus on enriching a particular clade/region within the ingroup.
-- The Basic local alignment search tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
+  Users can provide a search taxon that is either a more or a less inclusive
+  clade relative to the ingroup of the original phylogeny.
+If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
+  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
+- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide
   database, and the remaining sequences on the alignment.
+  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise all-against-all BLAST search is performed. This means that each sequence
+- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding ide numbers (accesion numbers in the case of the GenBank database). This information will be
-  used later to store the whole sequences into a new library.
+  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
+This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
 - The DNA sequence similarity search can be done on a local database that is easily

### 2020-06-23 physcraper_ms.Rmd
added bootstrap to methods

@@
 - First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
-  is aligned against the original alignment, working as a template. This ensures that
-  the final alignment follows the homology criteria established by the original alignment.
+- Then, a MUSCLE profile alignment is performed, in which the original alignment
+  is used as a template to align new sequences. This ensures that the final alignment
+  follows the homology criteria established by the original alignment.
 - The final alignment is not further processed automatically. We encourage users
@@
 
-## Tree reconstruction
+## Tree reconstruction and comparison
 
-- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with 100 bootstrap replicates.
+- A gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml]
+  with 100 classic bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
+  Other bootstrap types are transfer bootstrap expectation (TBE) recently proposed in @lemoine2018renewing.
+- The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
+- A conflict analysis is then run
 - The root of the tree is determined by finding the parent node of the sequences that do not belong to the ingroup.
 This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- The final result is a tree for eac of the genes provided in the alignment.
-
-## Tree comparison
-
 - Gene trees are then coupled to conlict info.

### 2020-06-24 physcraper_ms.Rmd
take out to dos from main ms

--- 2020-06-14 physcraper_ms.Rmd
@@
 
-  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT
-  for that single taxon would be determined as ***HOW?***
-  <!-- ***Physcraper does not take into account the focal clade from Open Tree.
-Because that is arbitrarily assigned or not by a curator.
-The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  If the search taxon is provided by the user, it can be a more inclusive
-  clade, if the user wants to perform a wider search, outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order of the ingroup. It can also be a less inclusive clade, if the user only
-  wants to focus on enriching a particular clade/region within the ingroup.
-- The Basic local alignment search tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
+  Users can provide a search taxon that is either a more or a less inclusive
+  clade relative to the ingroup of the original phylogeny.
+If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
+  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
+- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide
   database, and the remaining sequences on the alignment.
+  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise all-against-all BLAST search is performed. This means that each sequence
+- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding ide numbers (accesion numbers in the case of the GenBank database). This information will be
-  used later to store the whole sequences into a new library.
+  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
+This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
 - The DNA sequence similarity search can be done on a local database that is easily
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-  <!--***How does pyphlawd, or phylota does the exemplar per species choosing?***-->
 - Reverse complement sequences are identified and translated.
@@
 - First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
-  is aligned against the original alignment, working as a template. This ensures that
-  the final alignment follows the homology criteria established by the original alignment.
+- Then, a MUSCLE profile alignment is performed, in which the original alignment
+  is used as a template to align new sequences. This ensures that the final alignment
+  follows the homology criteria established by the original alignment.
 - The final alignment is not further processed automatically. We encourage users
-  to check it using any of the many tools for alignment processing, and eliminate
-  columns with no information.
+  to check it either by eye and perform manual refinement or using any of the many
+  tools for alignment processing, to eliminate columns with no information.
 
-## Tree reconstruction
+## Tree reconstruction and comparison
 
-- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with 100 bootstrap replicates.
-- The root of the tree is determined by finding the parent node of the sequences that do not belong to the ingroup.
+- A gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml]
+  with 100 classic bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
+  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+- The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
+- Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
+- Robinson Foulds metrics
+- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
+- For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
+- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
 This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- The final result is a tree for eac of the genes provided in the alignment.
-
-## Tree comparison
-
-- Gene trees are then coupled to conlict info.
 - Conflict information can only be generated in the context of the whole Open Tree of
@@
 How is physcraper already useful:
-- mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think.
+- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
 -

### 2020-06-24 _TODO.Rmd
take out to dos from main ms

---
title: Daily goals
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ilex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  global.par = TRUE
  )
```

# next:

- [ ] Finding a tree to update
  - [ ]  Using the curator website. Mention that there is a differnce between focal clade and ingroup. Physcraper does not take into account the focal clade from Open Tree. Because that is arbitrarily assigned by a curator. The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.
  - [ ] The link between the ingroup and the search taxon.

- [ ] How does the blast cycles results compare to other tools?

- [ ] Intro: why python? Introduce the taxon name space and the possibility of cross conectivity between platforms.

- [ ] How to create taxon name spaces for new trees so we can perform a comparison?

- [ ] How does pyphlawd, or phylota do the exemplar per species choosing for the baited analyses?

- [ ] bootsrap: is it classic [@felsenstein1985confidence] or transfer bootstrap expectation [@lemoine2018renewing]

- [ ] check the all-vs-all blast method, whre does it come from?

- [ ] add phylotaR to the table of pipelines for mining
- [ ] check paper: A System for Phylogenetic Analyses over Alignments of Next Generation Sequence Data

- [ ] order of trees for synthesis https://files.opentreeoflife.org/synthesis/opentree12.3/opentree12.3/phylo_input/study_tree_pairs.txt

- [ ] comment on things needed for automatization of tree comparison
  - [ ] validating taxon names with a reference taxonomy
  - [ ] taxon sampling statistics
  - [ ] finding trees in a collection
  - [ ] pruning source trees to contain only taxa of interest

- [ ] the case for taxon name standardization
  - [ ] cite studies that have pointed out before that it is a necessary step for effective automated tree comparison @page2005taxonomic; @ranwez2009phyloexplorer; @stoltzfus2013phylotastic; treebase

- [ ] statistics on taxon coverage
  - [ ] dataset sparseness
  - [ ] taxonomic coverage of the datasets
  - [ ] taxonomic sampling among groups
  - [ ] detect under-represented taxa

- [ ] finding trees to update:
  - [ ] from a collection? NOT NOW
  - [ ] containing taxa from a specific taxon
  - [ ] trees from a study

- [ ] pruning trees

- [ ] the root
  - [ ] write something about the need to set the root to be able to do comparisons

- [ ]  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT for that single taxon would be determined as ***HOW?***

# Done!

- [x] add bootsrap to methods

- [x] add BLAST command line tools to methods [@camacho2009blast]

- [X] explain how we root the updated trees

- [X] add blast citation!

- [x] add merced cluster nsf grant to acknowledgments
- [X] add genbank citations
  - [x] refer to it as a nucleotode database when decsribing the methods, bc it could be any molecular database, not just GenBank.

- [x] the ingroup
  - [x] how is it determined
  - [x] how is it relevant for blast results

### 2020-06-24 1-abstract.Rmd
refactoring

--- 2020-06-23 physcraper_ms.Rmd
@@
 date: '`r format(Sys.time(), "%d %B, %Y")`'
-bibliography: paper.bib
 csl: apa.csl
 output:
-  redoc::redoc
+  bookdown::pdf_document2:
+    keep_tex: true
 ---
@@
 use are available at <https://github.com/McTavishLab/physcraper>.
-
-# Introduction
-
-Phylogenies are important.
-
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-***paragraph better to end discussion???
-***
-
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
-
-1. Obtention of molecular or morphological character data -- get DNA from some organisms
-and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
-1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
-aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-1. Analyse this hypothesis of homology to infer phylogenetic relationships among
-the organisms you are studying -- Use different available programs to infer molecular
-evolution, trees and times of divergence.
-1. Discuss the inferred relationships in the context of previous hypothesis, the
-biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
-
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
-
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
-
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
-
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
-
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
-
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered "best" by experts
-not only in phylogenetics, but also experts on the focal group of study.
-
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control along the process
-- Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
-
-All these pipelines start tree construction from zero? Yes.
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-The original tree is by default used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
-
-# How does Physcraper work?
-
-## The input: a study tree and an alignment
-
-- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
-- The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, `physcraper` can download it directly,
-  either from the TreeBASE website (<https://treebase.org/>)
-  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
-- A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-- A ".csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from both the tree and alignment.
-  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.
-
-## DNA sequence search and cleaning
-
-- The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be provided by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
-
-  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT
-  for that single taxon would be determined as ***HOW?***
-  <!-- ***Physcraper does not take into account the focal clade from Open Tree.
-Because that is arbitrarily assigned or not by a curator.
-The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
-  <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-- The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
-- Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack database id numbers, a filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
-- Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-  <!--***How does pyphlawd, or phylota does the exemplar per species choosing?***-->
-- Reverse complement sequences are identified and translated.
-- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
-***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
-- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
-
-## DNA sequence alignment
-
-- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
-- First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
-  is aligned against the original alignment, working as a template. This ensures that
-  the final alignment follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. We encourage users
-  to check it using any of the many tools for alignment processing, and eliminate
-  columns with no information.
-
-## Tree reconstruction
-
-- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with 100 bootstrap replicates.
-- The root of the tree is determined by finding the parent node of the sequences that do not belong to the ingroup.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- The final result is a tree for eac of the genes provided in the alignment.
-
-## Tree comparison
-
-- Gene trees are then coupled to conlict info.
-- Conflict information can only be generated in the context of the whole Open Tree of
-Life. Otherwise, it is not really possible to get conflict data.
-***- One way to compare two independent phylogenetic trees is to compare them both to
-the synthetic OToL and then measure how well they do against each other***
-
-# Examples
-
-## The hollies
-
-The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-***Contact authors? They seem old school, probably do not wanna share their data.***
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
-Updating the oldest tree, we should get something very similar to the newest tree.
-
-## The Ascomycota
-
-Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by @schoch2009ascomycota.
-The first step, is to get the Open Tree of Life study id. There are some options to do this:
-- You can go to the Open Tree of Life website and browse until you find it, or
-- you can get the study id using R tools:
-  - By using the TreeBase ID of the study (which is not fully exposed on the
-    TreeBase website home page of the study, so you have to really look it up manually):
-```{r, treebase, eval = TRUE}
-rotl::studies_find_studies(property = "treebaseId", value = "S2137")
-```
-  - By using the name of the focal clade of study (but this behaved very differently):
-```{r, focal, eval = FALSE}
-rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
-```
-
-Once we have the study id, we can gather the trees published on that study:
-```{r, treeids, eval = TRUE}
-rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
-rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
-my_trees <- rotl::get_study("pg_238")
-```
-Both trees from this study have `r ape::Ntip(my_trees[[1]])` tips.
-
-Let's check what one of the trees looks like:
-
-<!-- ```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.width = 2} -->
-<!-- ape::plot.phylo(ape::ladderize(my_trees[[1]]), type = "phylogram", cex = 0.1) -->
-<!-- ``` -->
-
-1. Download the alignment from TreeBase
-If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
-To make this task easier, you can use a command to download everything into your working folder:
-```
-physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
-```
-In this example, all alignments posted on TreeBase were used to reconstruct both trees.
-
-1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:
-
-```
-physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
-```
-
-## Testudines example
-
-Phylogeny of the Testudines 6 tips from @crawford2012more
-There is just one tree in OToL.
-There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.
-
-```
-physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
-```
-# Discussion
-
-Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
-
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
-exposing new comers to expert understanding about the data.
-
-There are many tools that are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
-Most focus on generating full trees *de novo*, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have been already studied and that have phylogenetic data already.
-
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
-
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.
-
-Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer [@ranwez2009phyloexplorer].
-Compare and discuss.
-
-How is physcraper already useful:
-- mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think.
--
-
-How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
-- obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
-- one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
-
-Things that physcraper does not do:
-- analyse the whole GenBank database [@benson2000genbank; @wheeler2000database] to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward.
-There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
-
-## Tools that automatize any part of the process of phylogenetic reconstruction:
-
-### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
-
-| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
-|-----|------|:------:|:------:|:------:|
-| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
-| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
-| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
-| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
-The authors present a nice comparison with the phylota algorithm | Species trees??
-|
-| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
-| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
-| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
-|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
-|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
-| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
-| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
-| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
-| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
-| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
-| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
-| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
-| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
-
-According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
-"focus on the construction and refinement of robust alignments rather than the collection of homologs."
-
-### 2. Searching phylogenetic tree databases
-
-PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
-
-### 3. Mining phylogenetic tree databases
-
-PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database".
-
-### 4. Pipeline for phylogenetic reconstruction
-
-PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
-
-### 5. getting metadata and not sequences from GenBank.
-
-Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.
-
-## Phylota overview
-
-```{r phylota, child="phylota-overview.Rmd"}
-```
-
-# Acknowledgements
-
-We acknowledge contributions from
-
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
-
-# References

### 2020-06-24 1-abstract.tex
refactoring

--- 2020-05-08 physcraper_ms_ 2020-05-05.tex
@@
-\documentclass[10pt,a4paper,onecolumn]{article}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage{xcolor}
-\usepackage{authblk,etoolbox}
-\usepackage{titlesec}
-\usepackage{calc}
-\usepackage{tikz}
-\usepackage{hyperref}
-\hypersetup{colorlinks,breaklinks,
-            urlcolor=[rgb]{0.0, 0.5, 1.0},
-            linkcolor=[rgb]{0.0, 0.5, 1.0}}
-\usepackage{caption}
-\usepackage{tcolorbox}
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[]{article}
+\usepackage{lmodern}
 \usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
-\usepackage{seqsplit}
 \usepackage{fixltx2e} % provides \textsubscript
-\usepackage[
-  backend=biber,
-%  style=alphabetic,
-%  citestyle=numeric
-]{biblatex}
-\bibliography{paper.bib}
-
-% --- Page layout -------------------------------------------------------------
-\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
-            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}
-
-% --- Default font ------------------------------------------------------------
-% \renewcommand\familydefault{\sfdefault}
-
-% --- Style -------------------------------------------------------------------
-\renewcommand{\bibfont}{\small \sffamily}
-\renewcommand{\captionfont}{\small\sffamily}
-\renewcommand{\captionlabelfont}{\bfseries}
-
-% --- Section/SubSection/SubSubSection ----------------------------------------
-\titleformat{\section}
-  {\normalfont\sffamily\Large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsection}
-  {\normalfont\sffamily\large\bfseries}
-  {}{0pt}{}
-\titleformat{\subsubsection}
-  {\normalfont\sffamily\bfseries}
-  {}{0pt}{}
-\titleformat*{\paragraph}
-  {\sffamily\normalsize}
-
-% --- Header / Footer ---------------------------------------------------------
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyhf{}
-%\renewcommand{\headrulewidth}{0.50pt}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/3.6/Resources/library/rticles/rmarkdown/templates/joss_article/resources/JOSS-logo.png}}
-\fancyhead[C]{}
-\fancyhead[R]{}
-\renewcommand{\footrulewidth}{0.25pt}
-
-\fancyfoot[L]{\footnotesize{\sffamily Sanchez-Reyes et.
-al., (2020).
-Physcraper: A Python package for continually updated gene trees.
-\textit{Journal of Open Source Software}, (), .
-\href{https://doi.org/}{https://doi.org/}}}
-
-\fancyfoot[R]{\sffamily \thepage}
-\makeatletter
-\let\ps@plain\ps@fancy
-\fancyheadoffset[L]{4.5cm}
-\fancyfootoffset[L]{4.5cm}
-
-% --- Macros ---------
-
-\definecolor{linky}{rgb}{0.0, 0.5, 1.0}
-
-\newtcolorbox{repobox}
-   {colback=red, colframe=red!75!black,
-     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}
-
-\newcommand{\ExternalLink}{%
-   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
-       \begin{scope}[x=1ex, y=1ex]
-           \clip (-0.1,-0.1)
-               --++ (-0, 1.2)
-               --++ (0.6, 0)
-               --++ (0, -0.6)
-               --++ (0.6, 0)
-               --++ (0, -1);
-           \path[draw,
-               line width = 0.5,
-               rounded corners=0.5]
-               (0,0) rectangle (1,1);
-       \end{scope}
-       \path[draw, line width = 0.5] (0.5, 0.5)
-           -- (1, 1);
-       \path[draw, line width = 0.5] (0.6, 1)
-           -- (1, 1) -- (1, 0.6);
-       }
-   }
-
-% --- Title / Authors ---------------------------------------------------------
-% patch \maketitle so that it doesn't center
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-\patchcmd{\@maketitle}{center}{flushleft}{}{}
-% patch \maketitle so that the font size for the title is normal
-\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
-% patch the patch by authblk so that the author block is flush left
-\def\maketitle{{%
-  \renewenvironment{tabular}[2][]
-    {\begin{flushleft}}
-    {\end{flushleft}}
-  \AB@maketitle}}
-\makeatletter
-\renewcommand\AB@affilsepx{ \protect\Affilfont}
-%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
-\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
-\makeatother
-\renewcommand\Authfont{\sffamily\bfseries}
-\renewcommand\Affilfont{\sffamily\small\mdseries}
-\setlength{\affilsep}{1em}
-
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-
+  \usepackage{textcomp} % provides euro and other symbols
 \else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
+  \usepackage{unicode-math}
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-
 \fi
@@
 \IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \usepackage{hyperref}
-\hypersetup{unicode=true,
+\hypersetup{
             pdftitle={Physcraper: A Python package for continually updated gene trees},
+            pdfauthor={Luna L. Sanchez Reyes; Martha Kandziora; Emily Jane McTavish; University of California, Merced},
             pdfborder={0 0 0},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{5}
 % Redefines (sub)paragraphs to behave more like sections
@@
 
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
 \title{Physcraper: A Python package for continually updated gene trees}
-
-        \author[1]{Luna L. Sanchez Reyes}
-          \author[1]{Emily Jane McTavish}
-
-      \affil[1]{University of California, Merced}
-      \affil[2]{Institution 2}
-  \date{\vspace{-5ex}}
+\author{Luna L. Sanchez Reyes \and Martha Kandziora \and Emily Jane McTavish \and University of California, Merced}
+\date{24 June, 2020}
 
@@
 
-\marginpar{
-  %\hrule
-  \sffamily\small
-
-  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}
-
-  \vspace{2mm}
-
-  {\bfseries Software}
-  \begin{itemize}
-    \setlength\itemsep{0em}
-    \item \href{}{\color{linky}{Review}} \ExternalLink
-    \item \href{}{\color{linky}{Repository}} \ExternalLink
-    \item \href{}{\color{linky}{Archive}} \ExternalLink
-  \end{itemize}
-
-  \vspace{2mm}
-
-  {\bfseries Submitted:} \\
-  {\bfseries Published:}
-
-  \vspace{2mm}
-  {\bfseries License}\\
-  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
 }
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
 
@@
 \item
-  Phylogenies are still a key part of research in all areas of biology.
-  The nonspecialist can use one of the many tools available to
-  automatize phylogenetic reconstruction. However, interpretation of
-  results, comparison with previously available phylogenetic hypotheses,
-  and choosing of one phylogeny for discussion still impose difficulties
-  to one that is not a specialist on phylogenetic methods or on a
-  particular group of study.
+  Phylogenies are a key part of research in all areas of biology. Tools that automatize
+  some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
+  have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
+  However, interpretation of results, comparison with previously available phylogenetic
+  hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
+  to one that is not a specialist either on phylogenetic methods or on a particular group of study.
 \item
-  Physcraper is an open‐source Python program that automatizes the
-  update of public phylogenies by making use of public DNA sequence data
-  and taxonomic information.
+  Physcraper is an open‐source, command-line Python program that automatizes the update of published
+  phylogenies by making use of public DNA sequence data and taxonomic information,
+  providing a framework for comparison of published phylogenies with their updated versions.
 \item
-  Physcraper can be used by the nonspecialist, as a tool to generate
-  phylogenetic hypothesis based on public expert phylogentic knowledge.
-  The phylogenetics and group specialist will find it useful as a tool
-  to facilitate comparison of alternative phylogenetic hypotheses
-  (topologies). \textbf{\emph{Is physcraper intended for the
-  nonspecialist?? We have two types of nonspecialists: the ones that do
-  not know about phylogenetic methods and the ones that might know about
-  phylogenetic methods but do not know of a certain biological group.}}
+  Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
+  hypothesis based on already available expert phylogenetic knowledge.
+  Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
+  of alternative phylogenetic hypotheses (topologies).
+  \textbf{\emph{Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
+  the ones that do not know about phylogenetic methods and the ones that might know
+  about phylogenetic methods but do not know much about a certain biological group.}}
 \item
-  Physcraper implements node by node comparison of the the original and
-  the updated trees using the conflict API of OToL.
+  Physcraper implements node by node/topology comparison of the the original and the updated
+  trees using the conflict API of OToL, and summarizes differences.
 \item
-  We hope the physcraper workflow demonstrates the benefits of open
-  results in phylogenetics and encourages researchers to strive for
-  better data sharing practices.
+  We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
 \item
-  Physcraper can be used with any OS. Detailed instructions for
-  installation and use are available at
-  \url{https://github.com/McTavishLab/physcraper}.
+  Physcraper can be used with any OS. Detailed instructions for installation and
+  use are available at \url{https://github.com/McTavishLab/physcraper}.
 \end{enumerate}
 
-\hypertarget{intro}{%
-\section{Intro}\label{intro}}
-
-There is a lot of phylogenetic knowledge already available. Some of it
-has been made public at OToL. Most or all public phylogenies have the
-potential to be updated with the ever increasing amount of genetic data
-that is available on GenBank. There are now various methods to
-automatize use of GenBank data for phylogenetic reconstruction, which
-are generally intended to facilitate the use of phylogenetic knowledge
-for the nonspecialist. However, none of them provide a means to compare
-with previously available expert knowledge on a particular group.
-
-A key aspect of the standard phylogenetic workflow is comparison with
-already existing phylogenetic hypotheses and with phylogenies that are
-considered ``best'' by experts not only in phylogenetics, but also
-experts of the focal group of study.
-
-It is well known that GenBank holds enormous amounts of genetic data,
-and it continues to grow. A lot of this genetic data has the potential
-to be used to reconstruct the phylogenetic history of various organisms
-(Sanderson, Boss, Chen, Cranston, \& Wehe, 2008). Pipelines that harness
-this potential have been available for over a decade now, such as the
-Phylota brower, and PHLAWD. New ones keep on being developed, such as
-SUPERSMART and the upgraded version of PHLAWD, PyPHLAWD. Notably large
-phylogenies have been constructed using some of these tools, Some other
-have not been used that much. So, how well accepted is this approach in
-the community?
-
-Concerns with these tools: Errors in identification of sequences Little
-control along the process Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about
-the methods, so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but
-there is no way to assess if it is better than previous answers, it just
-assumes it is better because it used more data.
-
-All these pipelines start tree consruction from zero?
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison of existing phylogenies to phylogenies that
-are constructed using new genetic data from GenBank
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and
-connects it to the TreeBase database, to (1) get the original DNA
-dataset matrices (alignments) that produced a phylogeny that was
-published and then made available in the OToL database, (2) use this DNA
-alignments as a starting point to get new genetic data belonging to the
-focal group of study, to (3) finally update the phylogenetic
-relationships in the group.
-
-A less automated workflow is one in which the alignments that generated
-the published phylogeny are stored in other public database (such as
-DRYAD) or elsewhere (the users computer), and are provided by the users.
-
-The original tree is by defualt used as starting tree for the
-phylogenetic searches, but it can also be set as a full topological
-constraint or not used at all, depending on the goals of the user.
-
-Physcraper implements node by node comparison of the the original and
-the updated trees, using the conflict API of OToL.
-
-\hypertarget{physcraper-overview}{%
-\section{Physcraper overview}\label{physcraper-overview}}
-
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
-
-Imagine you are starting to work on a new biological group X. You have
-not much of an idea about its phylogenetic relationships, you are a
-newly established researcher, and the group is not anything any of your
-collaborators have worked on before. A good idea is to start an
-intensive litt review on the phylogenetics of the group. Rapidly, you
-find out there are 5 different phylogeneis, that used different markers,
-and that the papers, published at different times, do not discuss which
-phylogeny is the one accepted by the expert community on X. You might
-need to go to the annual conference of X, and even then, you might only
-find different and contrasting opinions. Somewhere along these months or
-even years doing this task, you looked into the the OToL database. You
-found in there some or all the published trees of X, along with a tree
-that has been deemed the best tree by curators and ideally experts on X?
-
-Let's be more specific now about our X group and say it is the
-Ascomycota. The best tree currently available in OToL was published by
-Schoch et al. (2009). The first step, is to get the Open Tree of Life
-study id. There are some options to do this: - You can go to the Open
-Tree of Life website and browse until you find it, or - you can get the
-study id using R tools: - By using the treebase ID of the study (which
-is not fully exposed on the Treebase website home page of the study, so
-you have to really look it up manually):
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property =} \StringTok{"treebaseId"}\NormalTok{, }\DataTypeTok{value =} \StringTok{"S2137"}\NormalTok{)}
-\CommentTok{##   study_ids n_trees         tree_ids candidate study_year title}
-\CommentTok{## 1    pg_238       2 tree109, tree110                 2009      }
-\CommentTok{##                                 study_doi}
-\CommentTok{## 1 http://dx.doi.org/10.1093/sysbio/syp020}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{itemize}
-\tightlist
-\item
-  By using the name of the focal clade of study (but this behaved very
-  differently):
-\end{itemize}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property=}\StringTok{"ot:focalCladeOTTTaxonName"}\NormalTok{, }\DataTypeTok{value=}\StringTok{"Ascomycota"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Once we have the study id, we can gather the trees published on that
-study:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_tree_ids}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\CommentTok{## [1] "tree109" "tree110"}
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{candidate_for_synth}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\CommentTok{## NULL}
-\NormalTok{my_trees <-}\StringTok{ }\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_study}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Both trees from this study have 434 tips.
-
-Let's check what one of the trees looks like:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ape}\OperatorTok{::}\KeywordTok{plot.phylo}\NormalTok{(my_trees[[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{type =} \StringTok{"fan"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics[width=468px]{physcraper_ms_ 2020-05-05_files/figure-latex/fungi-phylo-1}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Download the alignment from Treebase If you are on the Treebase home
-  page of the study, you can navigate to the matrix tab, and manually
-  download the alignments that were used to reconstruct the trees
-  reported on the study that were also uploaded to Treebase and to the
-  Open Tree of Life repository. To make this task easier, you can use a
-  command to download everything into your working folder:
-\end{enumerate}
-
-\begin{verbatim}
-physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
-\end{verbatim}
-
-In this example, all alignments posted on Treebase were used to
-reconstruct both trees.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  With the study id and the alignment files saved locally, we can do a
-  physcraper run with the command:
-\end{enumerate}
-
-\begin{verbatim}
-physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
-\end{verbatim}
-
-\hypertarget{tools-on-a-similar-track}{%
-\section{Tools on a similar track:}\label{tools-on-a-similar-track}}
-
-Tools that do similar things: pyPhlawd (Smith \& Walker, 2019)
-SUPERSMART (Antonelli et al., 2017)
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We acknowledge contributions from
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli, A., Hettling, H., Condamine, F. L., Vos, K., Nilsson, R. H.,
-Sanderson, M. J., Sauquet, H., et al. (2017). Toward a self-updating
-platform for estimating rates of speciation and migration, ages, and
-relationships of taxa. \emph{Systematic Biology}, \emph{66}(2),
-152--166.
-
-\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M. J., Boss, D., Chen, D., Cranston, K. A., \& Wehe, A.
-(2008). The PhyLoTA Browser: Processing GenBank for Molecular
-Phylogenetics Research. \emph{Systematic Biology}, \emph{57}(3),
-335--346.
-doi:\href{https://doi.org/10.1080/10635150802158688}{10.1080/10635150802158688}
-
-\leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
-Schoch, C. L., Sung, G.-H., López-Giráldez, F., Townsend, J. P.,
-Miadlikowska, J., Hofstetter, V., Robbertse, B., et al. (2009). The
-ascomycota tree of life: A phylum-wide phylogeny clarifies the origin
-and evolution of fundamental reproductive and ecological traits.
-\emph{Systematic biology}, \emph{58}(2), 224--239.
-
-\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
-Smith, S. A., \& Walker, J. F. (2019). PyPHLAWD: A python tool for
-phylogenetic dataset construction. \emph{Methods in Ecology and
-Evolution}, \emph{10}(1), 104--108.
-
 \end{document}

### 2020-06-24 100-phylota-overview.Rmd
refactoring

--- 2020-05-23 phylota-overview.Rmd
@@
 ---
-
-# Phylota overview
 
@@
     the red algal tree of life [@verbruggen2010data]
-    - @beaulieu2012modeling cites phylota As an example study of very large and comprehensive
+    - @beaulieu2012modeling cites phylota as an example study of very large and comprehensive
     phylogeny from mined DNA sequence data, (even if no phylogeny was really published
@@
     - a study constructing a dated seed plant phylogeny using pyPHLAWD [@smith2018constructing]
-    - a study presenting an assembly and alignment free method for phylogenetic reconstruction
-    using genomic data, that aims to be incorporated in a tool as phylota some day [@fan2015assembly].
+    - a study presenting an "assembly and alignment free" method for phylogenetic reconstruction
+    using genomic data. It aims to be incorporated into a pipeline such as phylota some day [@fan2015assembly].
     - nexml format presentation [@vos2012nexml] - cites phylota as a tool that uses
@@
     - a study of Southeast Asia plant biodiversity inventory [@webb2010biodiversity] -
-    cites phylota as a tool that would allow rapid phylogentic placing of newly
+    cites phylota as a tool that would allow rapid phylogenetic placing of newly
     discovered species, and generation of phylogenetically informed guides for field
@@
     releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
-    then clustered into candidate data sets". I also uses the PHYLOTA database in its
+    then clustered into candidate data sets". They also use the PHYLOTA database in its
     own pipeline.
@@
     - @freyman2015sumac cites phylota as a tool that "provides a web interface to view
-    all GenBank sequences within ta xonomic groups clustered into homologs" but that
+    all GenBank sequences within taxonomic groups clustered into homologs" but that
     does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They

### 2020-06-24 2-body.Rmd
refactoring

--- 2020-06-24 physcraper_ms.Rmd
@@
 ---
-
-# Abstract
-
-1. Phylogenies are a key part of research in all areas of biology. Tools that automatize
-some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
-have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
-However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
-to one that is not a specialist either on phylogenetic methods or on a particular group of study.
-
-1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
-phylogenies by making use of public DNA sequence data and taxonomic information,
-providing a framework for comparison of published phylogenies with their updated versions.
-
-1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on already available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
-of alternative phylogenetic hypotheses (topologies).
-***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know much about a certain biological group.***
-
-1. Physcraper implements node by node/topology comparison of the the original and the updated
-trees using the conflict API of OToL, and summarizes differences.
-
-1. We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-
-1. Physcraper can be used with any OS. Detailed instructions for installation and
-use are available at <https://github.com/McTavishLab/physcraper>.
 
@@
 ```
-Both trees from this study have `r ape::Ntip(my_trees[[1]])` tips.
+Both trees from this study have `r tryCatch(ape::Ntip(my_trees[[1]]), error = function(e){return(NA)})` tips.
 
@@
 - to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
--
+- Finding
 
@@
 PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
+| phylotaR | @bennett2018phylotar | studies |  |  |
 
@@
 
-```{r phylota, child="phylota-overview.Rmd"}
+```{r phylota, child="100-phylota-overview.Rmd"}
 ```

### 2020-06-24 2.1-intro-ideas.Rmd
refactoring

It is well known that GenBank [@benson2000genbank; @wheeler2000database] holds enormous amounts of genetic data, and it continues to grow. **Too cliché**

A lot of this genetic data has the potential to be used to reconstruct the phylogenetic
history of various organisms [@sanderson2008phylota].
Pipelines that harness this potential have been available for over a decade now, such
as the Phylota browser, and PHLAWD. New ones keep on being developed, such as SUPERSMART
and the upgraded version of PHLAWD, PyPHLAWD.
Notably large phylogenies have been constructed using some of these tools,
Some others have not been used that much. So, how well accepted is this approach in the community?

Various efforts to automatize parts of the process in order to be able to take advantage of data in public databases.

Most of this tools were envisioned for phylogenomics, were the analyses and processing of huge amounts of genetic data is naturally needed. Without automatization, curation of a phylogenomics database is a task that could not be finished in human time.
Some pipelines sacrificed precision and accuracy for quantity. But many tools to check
Examples of pipelines for phylogenomics:
BIR [Blast, Identify, Realign; @kumar2015bir] takes curated alignments as seed to find more data on genomic databases and create new phylogenomic hypotheses

Among other things, one of the major benefits stemming from these efforts are pipelines that make it easier to implement a reproducibile workflow in phylogenetic studies.

A widespread practice is a thorough visual inspection and manual curation of alignments, that are checked for "errors" that can usually result from similarity algorithms ***(give examples)***.
Commonly, alignment curators also take into account how the nature of the genetic marker, the taxon sampling and the organisms themselves, might affect the alignments. This hypotheses of homology are charged with expert knowledge, and hours of study, making them a scientifically reliable source if biological information.
As such, they represent an ideal platform to build up new phylogenetic knowledge.
Not many studies report its use.
Which might be an indication that the amount of manual effort that is required to curate gene alignments is still a bottleneck.

*** Introduce the Open Tree of Life***

### 2020-06-24 docs/submission-sysbio-2020-06-24.tex
refactoring

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin = 1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{caption}
\usepackage{blindtext}
\usepackage[left]{lineno}
\linenumbers

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\newpage

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Phylogenies are a key part of research in all areas of biology. Tools that automatize
  some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
  have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
  However, interpretation of results, comparison with previously available phylogenetic
  hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
  to one that is not a specialist either on phylogenetic methods or on a particular group of study.
\item
  Physcraper is an open‐source, command-line Python program that automatizes the update of published
  phylogenies by making use of public DNA sequence data and taxonomic information,
  providing a framework for comparison of published phylogenies with their updated versions.
\item
  Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
  hypothesis based on already available expert phylogenetic knowledge.
  Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
  of alternative phylogenetic hypotheses (topologies).
  \textbf{\emph{Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
  the ones that do not know about phylogenetic methods and the ones that might know
  about phylogenetic methods but do not know much about a certain biological group.}}
\item
  Physcraper implements node by node/topology comparison of the the original and the updated
  trees using the conflict API of OToL, and summarizes differences.
\item
  We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
\item
  Physcraper can be used with any OS. Detailed instructions for installation and
  use are available at \url{https://github.com/McTavishLab/physcraper}.
\end{enumerate}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Phylogenies are important.

Generating phylogenies is not easy and it is largely artisanal. Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community. \textbf{\emph{paragraph better to end discussion??? }}

The process of phylogenetic reconstruction implies many steps (that I generalize to the following):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtention of molecular or morphological character data -- get DNA from some organisms
  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}).
\item
  Assemble a hypothesis of homology -- Create a matrix of your character data, by
  aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
\item
  Analyse this hypothesis of homology to infer phylogenetic relationships among
  the organisms you are studying -- Use different available programs to infer molecular
  evolution, trees and times of divergence.
\item
  Discuss the inferred relationships in the context of previous hypothesis, the
  biology and biogeography of the organisms, etc. -- Answer the question, \emph{is this phylogenetic solution fair/reasonable?}
\end{enumerate}

Each of these steps require different types of specialized training: in the field,
in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
All of these steps also require considerable amounts of time for training and implementation.

In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
pipelines that mine already available molecular data from the GenBank repository (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}),
to obtain homologous characters that can be used for phylogenetic reconstruction.
These tools have been presented as aid for the nonspecialist to decrease some of
the difficulties in the generation of phylogenetic knowledge. However, they are
not that often used as so, suggesting that there are still difficulties for the nonspecialist.
The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
However, automatizing the assembly of the character data set is a crucial step towards
reproducibility for a task that was otherwise primarily artisanal and hence largely
non-reproducible.

Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.

Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}).

We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.

A key aspect of the standard phylogenetic workflow is comparison with already existing
phylogenetic hypotheses and with phylogenies that are considered ``best'' by experts
not only in phylogenetics, but also experts on the focal group of study.

Concerns I think people have about these tools:
- Errors in identification of sequences
- Little control along the process
- Too much of a black box?

Most of these phylogenies are being constructed by people learning about the methods,
so they want to know what is going on.

The pipelines are so powerful and they will give you an answer, but there
is no way to assess if it is better than previous answers, it just assumes it is
better because it used more data.

All these pipelines start tree construction from zero? Yes.

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison between existing phylogenies and phylogenies that are constructed
using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000})).

To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
a phylogeny that was published and then made available in the OToL database, (2)
use this DNA alignments as a starting point to get new genetic data belonging
to the focal group of study, to (3) finally update the phylogenetic relationships in the group.

A less automated workflow is one in which the alignments that generated the published
phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
are provided by the users.

The original tree is by default used as starting tree for the phylogenetic searches, but it
can also be set as a full topological constraint or not used at all, depending on
the goals of the user.

Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.

\hypertarget{how-does-physcraper-work}{%
\section{How does Physcraper work?}\label{how-does-physcraper-work}}

\hypertarget{the-input-a-study-tree-and-an-alignment}{%
\subsection{The input: a study tree and an alignment}\label{the-input-a-study-tree-and-an-alignment}}

\begin{itemize}
\tightlist
\item
  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish et al. \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
  reason for this is that trees in phylesystem have a set of user defined characteristics
  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup. Outgroup and ingroup taxa in the original tree are identified and tagged. This allows to automatically set the root for the updated tree on the next steps of the pipeline.
  A user can choose from the `r rotl::tol\_about()\$num\_source\_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}). If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
\item
  The alignment should be a gene alignment that was used to generate the tree. The original
  alignments are usually stored in a public repository such as TreeBase (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}; Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}),
  DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
  If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
  either from the TreeBASE website (\url{https://treebase.org/})
  or through the TreeBASE GitHub repository (SuperTreeBASE; \url{https://github.com/TreeBASE/supertreebase}).
  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
  downloaded by the user, and it's local path has to be provided as an argument.
\item
  A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the DNA character matrix and vice versa.
\item
  A ``.csv'' file with the summary of taxon name matching is produced for the user.
\item
  Unmatched taxon names are dropped from both the tree and alignment.
  Technically, just one matching name is needed to perform the searches. Please, see next section.
\item
  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
\end{itemize}

\hypertarget{dna-sequence-search-and-cleaning}{%
\subsection{DNA sequence search and cleaning}\label{dna-sequence-search-and-cleaning}}

\begin{itemize}
\item
  The next step is to identify the search taxon within the reference taxonomy. The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group. Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy. The search taxon can be provided by the user. If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (\url{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}).

  Users can provide a search taxon that is either a more or a less inclusive
  clade relative to the ingroup of the original phylogeny. If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
\item
  The Basic Local Alignment Search Tool, BLAST {[}Altschul et al. (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
  similarity between DNA sequences within the search taxon in a nucleotide
  database, and the remaining sequences on the alignment.
  The BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
\item
  A pairwise alignment-against-all BLAST search is performed. This means that each sequence
  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database). This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
\item
  The DNA sequence similarity search can be done on a local database that is easily
  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
\item
  The search can also be performed remotely, on the NCBI database. In this case, the
  bioPython BLAST algorithm is used to perform the similarity search.
\item
  Matched sequences below an e-value, percentage similarity, and outside a minimum
  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
  All acepted sequences are asigned an internal identifier, and are further filtered.
\item
  Because the original alignments usually lack database id numbers, a filtering
  step is needed. Accepted sequences that belong to the
  same taxon of the query sequence, and that are either identical or shorter than
  the original sequence are discarded. Only longer sequences belonging to the
  same taxon as the orignal sequence will be considered further for analysis.
\item
  Among the remaining filtered sequences, there are usually several exemplars per taxon.
  Although it can be useful to keep some of them to, for example, investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
\item
  Reverse complement sequences are identified and translated.
\item
  Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences. This can be done iteratively, but by default only sequences in the alignment are blasted. \textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
\item
  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
\item
  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
\end{itemize}

\hypertarget{dna-sequence-alignment}{%
\subsection{DNA sequence alignment}\label{dna-sequence-alignment}}

\begin{itemize}
\tightlist
\item
  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is implemented to perform alignments.
\item
  First, all new sequences are aligned using default MUSCLE options.
\item
  Then, a MUSCLE profile alignment is performed, in which the original alignment
  is used as a template to align new sequences. This ensures that the final alignment
  follows the homology criteria established by the original alignment.
\item
  The final alignment is not further processed automatically. We encourage users
  to check it either by eye and perform manual refinement or using any of the many
  tools for alignment processing, to eliminate columns with no information.
\end{itemize}

\hypertarget{tree-reconstruction-and-comparison}{%
\subsection{Tree reconstruction and comparison}\label{tree-reconstruction-and-comparison}}

\begin{itemize}
\tightlist
\item
  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine et al. (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
\item
  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
\item
  Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
\item
  Robinson Foulds metrics
\item
  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
\item
  For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
\item
  Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon. This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
\item
  Conflict information can only be generated in the context of the whole Open Tree of
  Life. Otherwise, it is not really possible to get conflict data.
  \textbf{\emph{- One way to compare two independent phylogenetic trees is to compare them both to
  the synthetic OToL and then measure how well they do against each other}}
\end{itemize}

\hypertarget{examples}{%
\section{Examples}\label{examples}}

\hypertarget{the-hollies}{%
\subsection{The hollies}\label{the-hollies}}

The genus \emph{Ilex} is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
The second has not been made available anywhere, not even in supplementary data of the journal.
\textbf{\emph{Contact authors? They seem old school, probably do not wanna share their data.}}
The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
it has not been added to the syntehtic tree yet.
This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}). Updating the oldest tree, we should get something very similar to the newest tree.

\hypertarget{the-ascomycota}{%
\subsection{The Ascomycota}\label{the-ascomycota}}

Let's be more specific now about our X group and say it is the Ascomycota.
The best tree currently available in OToL was published by Schoch et al. (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
The first step, is to get the Open Tree of Life study id. There are some options to do this:
- You can go to the Open Tree of Life website and browse until you find it, or
- you can get the study id using R tools:
- By using the TreeBase ID of the study (which is not fully exposed on the
TreeBase website home page of the study, so you have to really look it up manually):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property =} \StringTok{"treebaseId"}\NormalTok{, }\DataTypeTok{value =} \StringTok{"S2137"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   study_ids n_trees tree_ids candidate study_year title
## 1    pg_238       1  tree109                 2009      
##                                 study_doi
## 1 http://dx.doi.org/10.1093/sysbio/syp020
\end{verbatim}

\begin{itemize}
\tightlist
\item
  By using the name of the focal clade of study (but this behaved very differently):
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property=}\StringTok{"ot:focalCladeOTTTaxonName"}\NormalTok{, }\DataTypeTok{value=}\StringTok{"Ascomycota"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Once we have the study id, we can gather the trees published on that study:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_tree_ids}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "tree109"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rotl}\OperatorTok{::}\KeywordTok{candidate_for_synth}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## NULL
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{my_trees <-}\StringTok{ }\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_study}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Both trees from this study have NA tips.

Let's check what one of the trees looks like:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Download the alignment from TreeBase
  If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
  To make this task easier, you can use a command to download everything into your working folder:
\end{enumerate}

\begin{verbatim}
physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
\end{verbatim}

In this example, all alignments posted on TreeBase were used to reconstruct both trees.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  With the study id and the alignment files saved locally, we can do a physcraper run with the command:
\end{enumerate}

\begin{verbatim}
physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
\end{verbatim}

\hypertarget{testudines-example}{%
\subsection{Testudines example}\label{testudines-example}}

Phylogeny of the Testudines 6 tips from Crawford et al. (\protect\hyperlink{ref-crawford2012more}{2012})
There is just one tree in OToL.
There is just one alignment on \href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase} with all the 1 145 loci.

\begin{verbatim}
physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
\end{verbatim}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Data repositories hold more information than meets the eye.
Besides the actual data, they have other types of information that can be used for the advantage of science.

Usually, initial ideas about the data are changed by analyses.
We expect that this new ideas on the data can be registered on data bases,
exposing new comers to expert understanding about the data.

There are many tools that are making use of DNA data repositories in different ways.
Most of them focus on efficient ways to mine the data -- getting the most homologs.
Some focus on accurate ways of mining the data - getting real and clean homologs.
Others focus on refinement of the alignment.
Most focus on generating full trees \emph{de novo}, mainly for regions of the Tree of
Life that have no phylogenetic assessment yet in published studies, but also for
regions that have been already studied and that have phylogenetic data already.

All these tools are great efforts for advancing towards reproducibility in phylogenetics,
a field that has been largely recognised as somewhat artisanal.
We propose adding focus to other sources of information available from data repositories.
Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
also accumulating fast in public and open repositories.
In this way, the physcraper pipeline can be complemented with other tools that have
been developed for other purposes.

We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
community to build upon phylogenetic knowledge, using not only data accumulated in
DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies \emph{de novo}.

Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
Mention statistics provided by other tools: PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
Compare and discuss.

How is physcraper already useful:
- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
- Finding

How can it be used for the advantage of the field:
- rapid phylogenetic placing of newly discovered species, as mentioned in Webb et al. (\protect\hyperlink{ref-webb2010biodiversity}{2010})
- obtain trees for ecophylogenetic studies, as mentioned in Helmus and Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
- one day could be used to sistematize nucleotide databases, such as Genbank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro and Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus et al. (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})

Things that physcraper does not do:
- analyse the whole GenBank database (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017}). There are already some very good tools that do that.
- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
- it is not a tree repo, as phylota is, mentioned in Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014})

\hypertarget{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}{%
\subsection{Tools that automatize any part of the process of phylogenetic reconstruction:}\label{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}}

\hypertarget{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}{%
\subsubsection{1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction}\label{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}}

\begin{longtable}[]{@{}llccc@{}}
\toprule
\begin{minipage}[b]{0.12\columnwidth}\raggedright
Tool\strut
\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright
Citation\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
Cited by\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
Description\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
Supermatrix/gene tree/species tree\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Phylota\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Sanderson et al. (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
122 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
AMPHORA\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Wu and Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
458 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
PHLAWD\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Smith et al. (\protect\hyperlink{ref-smith2009mega}{2009})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
234 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Baited search of DNA markers on the GenBank database; phylogenetic reconstruction\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Unnamed \href{https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals}{ruby pipeline}, only available from \href{https://static-content.springer.com/esm/art\%3A10.1186\%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP}{supplementary data} of the journal\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Peters et al. (\protect\hyperlink{ref-peters2011taming}{2011})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
64 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established ``criteria of compositional homogeneity and defined levels of density and overlap''\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Unnamed\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Grant and Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
38 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Unnamed\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
10 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
algorithm that mines GenBank data to delineate species in the insecta. The authors present a nice comparison with the phylota algorithm\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Species trees??\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
PUmPER\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Izquierdo-Carrasco et al. (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
14 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
perpetual updating with newly added sequences to GenBank\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure yet\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
DarwinTree\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Meng et al. (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
6 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
NCBIminer\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Xu et al. (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
4 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
part of darwintree\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
SUMAC\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
19 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
both ``baited'' analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
STBase\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
McMahon et al. (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
7 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
pipeline for species tree construction and the public database of one million precomputed species trees\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
species trees\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Unnamed\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Papadopoulou et al. (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
17 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Automated DNA-based plant identification for large-scale biodiversity assessment\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
BIR\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Kumar et al. (\protect\hyperlink{ref-kumar2015bir}{2015})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
6 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
blast, align, identify homologs via constructed trees, curate and realign\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
SUPERSMART\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
35 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
baited analyses up to bayesian divergence time estimation\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
SOPHI\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
{[}Chesters (\protect\hyperlink{ref-chesters2017construction}{2017})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
17 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
not sure\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
phyloSkeleton\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Guy (\protect\hyperlink{ref-guy2017phyloskeleton}{2017})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
5 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
OneTwoTree\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
7 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Web‐based, user-friendly, online tool for species-tree reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available sequence data from NCBI GenBank\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
pyPhlawd\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Smith and Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
6 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
baited and clustering analyses\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
Supermatrix or gene tree\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
Phylotol\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
5 studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
``phylogenomic pipeline to allow easy incorporation of data from high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants. PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life'', i.e., bacteria and unicellular eukaryotes\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
supermatrix and gene trees\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.12\columnwidth}\raggedright
phylotaR\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
Bennett et al. (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
studies\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

According to Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
``focus on the construction and refinement of robust alignments rather than the collection of homologs.''

\hypertarget{searching-phylogenetic-tree-databases}{%
\subsubsection{2. Searching phylogenetic tree databases}\label{searching-phylogenetic-tree-databases}}

PhyloFinder (Chen et al. \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies

\hypertarget{mining-phylogenetic-tree-databases}{%
\subsubsection{3. Mining phylogenetic tree databases}\label{mining-phylogenetic-tree-databases}}

PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database''.

\hypertarget{pipeline-for-phylogenetic-reconstruction}{%
\subsubsection{4. Pipeline for phylogenetic reconstruction}\label{pipeline-for-phylogenetic-reconstruction}}

PhySpeTre (Fang et al. \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
pipeline.

\hypertarget{getting-metadata-and-not-sequences-from-genbank.}{%
\subsubsection{5. getting metadata and not sequences from GenBank.}\label{getting-metadata-and-not-sequences-from-genbank.}}

Datataxa Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.

\hypertarget{phylota-overview}{%
\subsection{Phylota overview}\label{phylota-overview}}

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database (Sanderson et al. \protect\hyperlink{ref-sanderson2008phylota}{2008}).

Since then, it has been cited 122 times for different reasons.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  As an example of a tool that mines GenBank data for phylogenetic reconstruction,
  or that is useful in any way for phylogenetics:

  \begin{itemize}
  \tightlist
  \item
    original publication of PHLAWD (Smith et al. \protect\hyperlink{ref-smith2009mega}{2009})
  \item
    an analysis identifying research priorities and data requirements for resolving
    the red algal tree of life (Verbruggen et al. \protect\hyperlink{ref-verbruggen2010data}{2010})
  \item
    Beaulieu et al. (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive
    phylogeny from mined DNA sequence data, (even if no phylogeny was really published
    there, only the method to do so)
  \item
    a review for ecologists about phylogenetic tools (Roquet et al. \protect\hyperlink{ref-roquet2013building}{2013})
  \item
    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith and Brown \protect\hyperlink{ref-smith2018constructing}{2018})
  \item
    a study presenting an ``assembly and alignment free'' method for phylogenetic reconstruction
    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan et al. \protect\hyperlink{ref-fan2015assembly}{2015}).
  \item
    nexml format presentation (Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses
    stored phyloinformatic data that could benefit from adopting nexml, to increase
    interoperability.
  \item
    a study of fruit evolution, analysing a previously published phylogeny of 8911
    tips of the Campanulidae, constructed with PHLAWD (Beaulieu and Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})
  \item
    a study of Southeast Asia plant biodiversity inventory (Webb et al. \protect\hyperlink{ref-webb2010biodiversity}{2010}) -
    cites phylota as a tool that would allow rapid phylogenetic placing of newly
    discovered species, and generation of phylogenetically informed guides for field
    identification.
  \item
    a study of wood density for carbon stock assessments (Flores and Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),
    cites phylota as an initiative to ``get supertrees resolved up to species level''.
  \item
    a study proposing something similar to Open tree but applied only to land plants (Beaulieu et al. \protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})
  \item
    an analysis of the phylogenetic diversity-area curve (Helmus and Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
    cited phylota as a method alternative to phylomatic to ``obtain plant phylogenetic
    trees for ecophylogenetic studies''.
  \item
    a study generating a phylogeny of 6,098 species of vascular plants from China
    (Chen et al. \protect\hyperlink{ref-chen2016tree}{2016}) - uses DarwinTree (Meng et al. \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}) and generates sequence
    data \emph{de novo} for 781 genera.
  \item
    a review of the state of methods and knowledge generated by molecular systematics
    (San Mauro and Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
    for large-scale molecular phylogenetics analysis''.
  \item
    the first phylotastic paper (Stoltzfus et al. \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny
    related resource that provides ways to generate custom species trees for downstream use''.
  \item
    Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017}) cites phylota as a ``pipeline that pre-processes entire GenBank
    releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
    then clustered into candidate data sets''. They also use the PHYLOTA database in its
    own pipeline.
  \item
    Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014}) present an algorithm for mining of frequent subtrees (common patterns)
    in collections of phylogenetic trees, as a way to extract meaningful phylogenetic
    information from collections of trees when compared to maximum agreement subtrees
    and majority-rule trees. They cite phylota as one of such tree collections available
    along with TreeBASE (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}).
  \item
    Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) cites phylota as a ``program providing basic statistics
    on data availability for molecular datasets''. They propose a tool to upload and
    explore user phylogenies to obtain detailed summary statistics on user tree collections.
  \item
    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}) cites phylota as a tool that ``provides a web interface to view
    all GenBank sequences within taxonomic groups clustered into homologs'' but that
    does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They
    compare the performance of SUMAC to Phylota. This is also presented in their PhD dissertation (Freyman \protect\hyperlink{ref-freyman2017phylogenetic}{2017}).
  \item
    Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) cites phylota as a data mining tool that compiles metadata
    from mining of public DNA databases ``for construction of large phylogenetic trees
    and multiple gene sets'' and that the authors have recognised that gene annotations
    in public databases are insufficient and that careful partitioning of orthologous
    sequences is needed for supermatrix construction. Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) present
    a procedure that minimizes the problem of forming multilocus species units in
    a large phylogenetic data set using algorithms from graph theory.
  \item
    Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014}) present an algorithm to delineate species form GenBank
    DNA data, and cites phylota as a tool that partitions ``the contents of a database
    according to homology'', by ``grouping of database sequences according to internal
    criteria'', searching ``from a standardized set of references {[}\ldots{}{]} patterns in
    sequence similarity and overlap.''
  \item
    the paper presenting phylotaR, a pipeline that recreates the phylota output
    but uses the most updated GenBank release, and is available in R (Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}),
    cites phylota as its predecessor and inspiration. The authors mention that phylotaR
    pipeline mimics phylota's pipeline but with improvements.
  \item
    The paper presenging PhyloBase (Jamil \protect\hyperlink{ref-jamil2016visual}{2016}), cites phylota as one of
    its resources to get phylogenies, along with TreeBASE and others.
  \item
    The paper presenting STBase, a database of one million precomputed species
    trees (Deepak \protect\hyperlink{ref-deepak2013managing}{2013}; McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as a databse of gene trees or mul-trees,
    ``trees having multiple sequences with the same taxon name''.
  \item
    Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018}) present a Web‐based, user-friendly, online tool for species-tree
    reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available
    sequence data from NCBI GenBank. They cite phylota in the intro as a tool that is ``designed to provide
    users with precomputed sets of clusters that were assembled through a single‐linkage
    clustering approach and additionally provides precomputed gene trees that were
    reconstructed for each cluster. In particular, the results obtained by PhyLoTa
    are taxonomically constrained; that is, all sequences of the most recent common
    ancestor are collected even if one specifies only part of a clade''.
  \item
    A study developing a tool to link wikipedia data to NCBI taxonomy (Page \protect\hyperlink{ref-page2011linking}{2011})
    cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
  \item
    the study that present DarwinTree (Meng et al. \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}), and all derived studies: the study
    presenting an approach to screen sequence data for The Platform
    for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
    to parallelize BLAST (Yong et al. \protect\hyperlink{ref-yong2010screening}{2010}), as well as Gao et al. (\protect\hyperlink{ref-gao2011solution}{2011}), Li et al. (\protect\hyperlink{ref-li2013partfasttree}{2013}),
    Meng et al. (\protect\hyperlink{ref-meng2014rapidtree}{2014}), Meng et al. (\protect\hyperlink{ref-meng2015sotree}{2015}\protect\hyperlink{ref-meng2015sotree}{c}), and Meng et al. (\protect\hyperlink{ref-meng2015solution}{2015}\protect\hyperlink{ref-meng2015solution}{b}), all cite phylota using the exact same
    introduction and sentence: as one among other ``studies based on data mining large numbers of taxa or loci''.
  \item
    A study presenting a tool to asses gene sequence quality for automatic
    construction of databases (Meng et al. \protect\hyperlink{ref-meng2012gsqct}{2012}\protect\hyperlink{ref-meng2012gsqct}{a}), as well as their parallelized version using MapReduce
    (Meng et al. \protect\hyperlink{ref-meng2012cloud}{2012}\protect\hyperlink{ref-meng2012cloud}{b}), cite phylota (along with Yong et al. (\protect\hyperlink{ref-yong2010screening}{2010})) as a tool that relies on sequence
    similarity (BLAST) and not taxon name annotations in the database, for mining
    large numbers of taxa or loci, without making any control on the quality of the
    sequencing.
  \item
    A review on online plant databases aiming to ``provide recommendations for current
    information managers and developers concerning the user interface and experience;
    and to provide a picture about the possible directions to take for those in charge
    of the creation of information at all levels''. They cite phylota as a tool allowing
    researchers ``to acces equally and globally, without travel, a {[}phylogenetic?{]} model
    of plants at the kingdom level'' (Jones et al. \protect\hyperlink{ref-jones2014trends}{2014}).
  \item
    a paper aiming to establish an online information system for the legumes and
    to outline ``best practices for development of a legume portal to enable data
    sharing and a better understanding of what data are available, missing, or erroneous,
    and ultimately facilitate cross-analyses and collaboration within the legume-systematics
    community and with other stakeholders'' (Bruneau et al. \protect\hyperlink{ref-bruneau2019towards}{2019}), cites phylota (along with supersmart and pyphlawd) as a
    ``pipeline for large-scale retrieval of GenBank data of particular taxa or clades''.
    In their Table 1, they also list phylota as a potential data source for developing a legume portal.
  \item
    A study on morphological evolution of electric fish skull, that uses phylotaR
    to retrieve sequences of the family Apteronotidae, order Gymnotiformes (Evans et al. \protect\hyperlink{ref-evans2019bony}{2019}),
    cites phylota as the inspiration and fundament of phylotaR.
  \item
    A phylogenetic revision of the Gymnotidae fish (Teleostei: Gymnotiformes),
    uses phylotaR to retrieve seqeunces, but cites phylota as ``a pipeline that implements
    BLAST searches to both identify and download sequence clusters for listed taxonomic
    groups to assemble a robust collection of sequences in a reproducible way based
    on publicly-available gene sequences while avoiding selection bias on the part
    of the assembler''.
  \item
    A master thesis on SearchTree, a ``software tool that allows users to query
    efficiently on an arbitrary user taxon list and returns high scoring matches
    from approximately one billion phylogenetic trees being constructed from molecular
    sequence data in GenBank'' (Deepak \protect\hyperlink{ref-deepak2010searchtree}{2010}), that seems to be the preliminary
    work for STBase (McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as ``a standard strategy, to assemble sets of homologous sequences
    (clusters) from a database of all-against-all BLAST searches, {[}in which{]} clusters
    are constructed in the context of the NCBI taxonomy tree for convenience of display,
    thus child clusters are contained within parent clusters, following the NCBI hierarchy''.
    In opposition, SearchTree uses true agglomerative hierarchical clustering
    (AHC: Day and Edelsbrunner (\protect\hyperlink{ref-day1984efficient}{1984})) based on the BLAST estimates of sequence dissimilarity
    rather than the NCBI tree".
  \item
    a recent review on the state of large phylogeny (namely insects) generation
    using tools of the data-driven era (Chesters \protect\hyperlink{ref-chesters2019phylogeny}{2019}) cites phylota as
    a tool for homology inference and retrieval.
  \item
    the study presenting phylotol (Cerón-Romero et al. \protect\hyperlink{ref-ceron2019phylotol}{2019}), cites phylota as a tool
    that ``focus on the identification and collection of homologous genes from public
    databases''.
  \item
    The \href{https://www.researchgate.net/profile/Douglas_Soltis/publication/228815637_Assembling_the_Tree_of_Life_for_the_Plant_Sciences_iPTOL/links/56a7c6bc08aeded22e3700dc.pdf}{iPTOL project}
    cites phylota as a resource of phylogenetic trees.
  \item
    Mahmood (\protect\hyperlink{ref-mahmood2015avian}{2015}) PhD dissertation presents a database of avian Raptor sequences
    (raptorbase), based on the phylota pipeline.
  \item
    Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) develops datataxa and cite phylota as ``software that has
    been developed to mine the massive amount of information stored in GenBank'',
    along with its R version (phylotaR; Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}) and restez \url{https://www.rdocumen-tation.org/packages/restez/versions/1.0.0}.
  \item
    The phylotastic project (Stoltzfus et al. \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny-related resource providing ways to generate custom species trees \emph{de novo} for downstream use'' along with CIPRES.
  \end{itemize}
\item
  When the software was actually used to construct (partially or in full) a DNA
  data set to be used for phylogenetic reconstruction:

  \begin{itemize}
  \tightlist
  \item
    A 1000 tip phylogeny of the family of the nightshades (Särkinen et al. \protect\hyperlink{ref-sarkinen2013solanaceae}{2013})
  \item
    A 56 tip phylogeny of crustacean zooplancton (Helmus et al. \protect\hyperlink{ref-helmus2010communities}{2010}) -- ecological study
  \item
    A 63 tip phylogeny of the Salmonidae family (Crête-Lafrenière et al. \protect\hyperlink{ref-crete2012salmonidae}{2012})
  \item
    A 321 tip phylogeny of Testudines (Thomson and Shaffer \protect\hyperlink{ref-thomson2010sparse}{2010})
  \item
    A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish (Martin and Wainwright \protect\hyperlink{ref-martin2011trophic}{2011})
  \item
    A 2,957 taxa phylogeny of the class Moniloformopses of living ferns (Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011})
  \item
    A 2,573 species phylogeny of the Papilionoidea (Hardy and Otto \protect\hyperlink{ref-hardy2014specialization}{2014})
  \item
    A 23 taxa phylogeny of the California flora (Anacker et al. \protect\hyperlink{ref-anacker2011origins}{2011})
  \item
    Phylogenies of 6 different clades of flowering plants representing an independent
    evolutionary origin of extrafloral nectaries: \emph{Byttneria} (Malvaceae), \emph{Pleopeltis} (Polypodiaceae),
    \emph{Polygoneae} (Polygoneaceae), \emph{Senna} (Fabaceae), \emph{Turnera} (Passifloraceae), and \emph{Viburnum}
    (Adoxaceae) (Weber and Agrawal \protect\hyperlink{ref-weber2014defense}{2014}).
  \item
    To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
    sampled at different taxonomic levels (Faurby and Svenning \protect\hyperlink{ref-faurby2015species}{2015})
  \item
    A 900 species tree of muroid rodents, Muroidea (Steppan and Schenk \protect\hyperlink{ref-steppan2017muroid}{2017}), where 300
    species were newly added by the study and the rest obtained using phylota.
  \item
    A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales (Ickert-Bond et al. \protect\hyperlink{ref-ickert2009fossil}{2009})
  \item
    A 1061 genera phylogeny of the Oscine birds (Selvatti et al. \protect\hyperlink{ref-selvatti2015paleogene}{2015})
  \item
    A 268 species phylogeny of sharks, representing all 8 orders and 32 families (Sorenson et al. \protect\hyperlink{ref-sorenson2014effect}{2014}; Sorenson \protect\hyperlink{ref-sorenson2014evolution}{2014})
  \item
    A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region (Tucker et al. \protect\hyperlink{ref-tucker2012incorporating}{2012}).
  \item
    A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects (Hardy and Cook \protect\hyperlink{ref-hardy2010gall}{2010}).
  \item
    A 196 species phylogeny of the family Boraginaceae (Nazaire and Hufford \protect\hyperlink{ref-nazaire2012broad}{2012}). The authors
    actually found data for 318 Boraginaceae spp using phylota, but decided to reduce
    their data set to focus on the monophyly of genus \emph{Mertensia}.
  \item
    A phylogeny of 401 species of scale insects Coccoidea, Hemiptera (Ross et al. \protect\hyperlink{ref-ross2013large}{2013}),
    with some sequences generated \emph{de novo}.
  \item
    Two phylogenies sampling all species of two different clades of insectivorous
    lizards, agamids and diplodactyline geckos, groups considered to be radiating
    in the Australia's Great Victoria Desert (Rabosky et al. \protect\hyperlink{ref-rabosky2011species}{2011})
  \item
    A phylogeny of 91 species of sparid and centracanthid fishes, Sparidae, Percomorpha,
    plus 2 outgroups, a lethrinid and a nemipterid exemplar (Santini et al. \protect\hyperlink{ref-santini2014first}{2014}).
  \item
    Updating a phylogeny of Arecaceae, constructing relationships in 6 cldes within
    the group: subfamilies Calamoideae and Coryphoideae, the tribe Ceroxyleae within
    subfamily Ceroxyloideae and three groups within subfamily Arecoideae: (1) Iriarteeae,
  \end{itemize}

  \begin{enumerate}
  \def\labelenumii{(\arabic{enumii})}
  \setcounter{enumii}{1}
  \tightlist
  \item
    Cocoseae: Attaleinae except Beccariophoenix and (3) a group containing six
    tribes; Euterpeae, Leopoldinieae, Pelagodoxeae, Manicarieae, Geonomateae and Areceae
    (Faurby et al. \protect\hyperlink{ref-faurby2016all}{2016}).
  \end{enumerate}

  \begin{itemize}
  \tightlist
  \item
    A phylogeny of 768 Gesneriaceae species and 58 outgroups for a total species
    sampling of 826 taxa (Roalson and Roberts \protect\hyperlink{ref-roalson2016distinct}{2016}) some sequence were generated \emph{de novo}.
  \item
    A phylogeny of 47 species of scombrid fishes, with 2 outgroups, a gempylid and
    a trichiurid (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013}).
  \item
    to update a dataset underlying a large-scale fern phylogeny (Lehtonen et al. \protect\hyperlink{ref-lehtonen2017environmentally}{2017}),
    data set in \url{https://zenodo.org/record/345670\#.Xr9QFRPYqqg}, also in TreeBASE,
    but it is one of those studies that is broken.
  \item
    A phylogeny of 13 species of billfishes, order Istiophoriformes: Acanthomorpha,
    and four outgroups (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013})
  \item
    A phylogeny of 765 aphid species, family Aphididae (Hardy et al. \protect\hyperlink{ref-hardy2015evolution}{2015})
  \item
    A phylogeny of less than 100 taxa of the family Ranunculaceae (Lehtonen et al. \protect\hyperlink{ref-lehtonen2016sensitive}{2016}),
    even though they retrieved info from phylota for 194 taxa within the family, they
    reduced their data set because of low sampling of markers for some taxa.
  \item
    A phylogeny of 144 neobatrachian genera, assuming the monophyletic status of
    genera to increase matrix-filling levels (Frazao et al. \protect\hyperlink{ref-frazao2015gondwana}{2015}).
  \item
    A 179 species phylogeny of the bird family Picidae (woodpeckers, piculets,
    and wrynecks) (Dufort \protect\hyperlink{ref-dufort2015coexistence}{2015}, \protect\hyperlink{ref-dufort2016augmented}{2016}), augmented with data from an updated GenBank
    release and newly sequenced data.
  \item
    A phylogeny of species of freshwater fish endemic to NorthAmerica (Strecker and Olden \protect\hyperlink{ref-strecker2014fish}{2014}),
    phylota found data for 54 out of 66 spp.
  \item
    A phylogeny of 520 species of the order Ericales (Hardy and Cook \protect\hyperlink{ref-hardy2012testing}{2012})
  \item
    A phylgeny of 16 fish species of the family Sphyraenidae (Percomorpha), as well
    as two outgroup species of the Centropomidae (barracudas) (Santini et al. \protect\hyperlink{ref-santini2015first}{2015})
  \item
    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas et al. \protect\hyperlink{ref-garcia2016role}{2016})
  \item
    Kolmann et al. (\protect\hyperlink{ref-kolmann2017dna}{2017}) uses phylota to download all 1691 co1 sequences belonging to
    the order Carchariniformes, to place phylogenetically DNA samples obtained from
    fish markets.
  \item
    A phylogeny of 329 bird species in the Tyrannidae (77\% of the species in the
    family) (Gómez Bahamón and others \protect\hyperlink{ref-gomez2015behavioral}{2015}; Gómez-Bahamón et al. \protect\hyperlink{ref-gomez2020speciation}{2020})
  \item
    Retrive 145 sequences registered as \emph{Holothuria} species, but kept 84 as ingroup,
    plus 4 outgroup sequences from \emph{Stichopus ocellatus}, all belonging to the order
    Apodida of sea cucumbers (Kamarudin et al. \protect\hyperlink{ref-kamarudin2016phylogenetic}{2016})
  \item
    On a master thesis, to get the sequences of the outgroups of Melinidinae, family Poaceae, namely several spp of the
    subfamily Panicoideae, plus \emph{Gynerium sagittatum}, \emph{Chasmanthium latifolium},
    and \emph{Zea mays}, (Salariato \protect\hyperlink{ref-salariato2010filogenia}{2010}). Interestingly, phylota was not used
    in the published study of the thesis (Salariato et al. \protect\hyperlink{ref-salariato2010molecular}{2010}). Ingroup sequences were generated \emph{de novo}.
  \item
    On a PhD thesis, to construct a phylogeny of Platyrrhini (internal group),
    Catarrhini (outgroup), and Tarsiiformes Pereira (\protect\hyperlink{ref-pereira2013padroes}{2013}). Have not found a published study.
  \item
    The 10k trees project (Arnold et al. \protect\hyperlink{ref-arnold201010ktrees}{2010}) uses phylota to construct a tree of 301 primate species
    and the outgroup species \emph{Galeopterus variegates}, a tree of 17 extant odd-toed
    ungulates species and the outgroup species \emph{Bos taurus}, and a tree of 70 different
    species of carnivorans and \emph{Equus caballus} as outgroup. However, the do not
    cite it on the paper, but only on their documentation \url{http://www.academia.edu/download/49690788/10kTrees_Documentation.pdf}.
  \item
    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}, also in \protect\hyperlink{ref-freyman2017phylogenetic}{2017}), use phylota to construct
    a phylogeny (or maybe only mine genbank???) of the Onagraceae and Lythracea,
    and compare it to the tool they propose, SUMAC.
  \item
    Blackmon (\protect\hyperlink{ref-blackmon2017synthesis}{2017}) PhD study applies phylota to reconstruct a 822 mite
    species tree.
  \item
    A study of the effect of poliploidy on niche evolution (Baniaga et al. \protect\hyperlink{ref-baniaga2018polyploid}{2018}),
    uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
    16 families and 25 genera,and a tree of 33 genera from 20 different families
    comprising 1706 taxa.
  \end{itemize}
\item
  When the website was used to identify sequences and markers available in
  GenBank for a particular group. In this cases, the dataset mining was either performed
  with other tools, or not performed at all and just used for discussion:

  \begin{itemize}
  \tightlist
  \item
    A 812 tips phylogeny of the Order Chiroptera (Shi and Rabosky \protect\hyperlink{ref-shi2015speciation}{2015}) -- dataset
    constructed with PHLAWD
  \item
    A 1276 tips phylogeny of the Fabaceae (Group et al. \protect\hyperlink{ref-legume2013legume}{2013}) -- dataset constructed
    by hand (I think??)
  \item
    A review of dated phylogenies of fire-prone tropical savanna species from Brazil
    (Simon and Pennington \protect\hyperlink{ref-simon2012cerrado}{2012}) -- just for discussion of the lack of markers available for
    these species on GenBank
  \item
    A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
    protists (Morrison \protect\hyperlink{ref-morrison2009apicomplexa}{2009}).
  \item
    Three data sets from phylota (the suborder Pleurodira of side-necked turtles;
    the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
    to demonstrate and exemplify phylogenetic decisiveness (Sanderson et al. \protect\hyperlink{ref-sanderson2010phylogenomics}{2010})
  \item
    Mentioned in a PHD thesis (Gagnon and others \protect\hyperlink{ref-gagnon2016systematique}{2016}), but not on the final publication (Gagnon et al. \protect\hyperlink{ref-gagnon2016new}{2016}),
    phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
  \end{itemize}
\item
  Sometimes, it was cited by mistake:

  \begin{itemize}
  \tightlist
  \item
    In this 630 tip phylogeny of the Caryophyllaceae study (Greenberg and Donoghue \protect\hyperlink{ref-greenberg2011caryophyllaceae}{2011}) it might have been originally
    cited as an example of large phylogenies that reflect well supported relationships
    from previous smaller phylogenies. However, it was removed from the text but
    not from the final list of references. The DNA data set was constructed by hand
    most probably.
  \item
    a study reconstructing the insect tree of life with 49,358 species, 13,865
    genera, and 760 families within the order Insecta (Chesters \protect\hyperlink{ref-chesters2017construction}{2017}),
    uses its own algorithm (SOPHI) to mine public DNA databases (Chesters and Zhu \protect\hyperlink{ref-chesters2014protocol}{2014}).
    It does not cite phylota as it should, but includes it in their references.
  \end{itemize}
\item
  When phylota was used to extract full trees (not only DNA data sets or markers):

  \begin{itemize}
  \tightlist
  \item
    Page (\protect\hyperlink{ref-page2013bionames}{2013}) uses it to generate phylogenies for the \href{http://bionames.org}{bionames website},
    a ``database linking taxonomic names to their original descriptions, to taxa, and
    to phylogenies'' generated with phylota.
  \item
    Deepak et al. (\protect\hyperlink{ref-deepak2013extracting}{2013}) uses a sample of phylota trees to test their method to
    remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
    trees with two or more leaves sharing a label, e.g., a species name, which can
    imply multiple conflicting phylogenetic relationships for the same set of taxa.
  \item
    A review by Sanderson et al. (\protect\hyperlink{ref-sanderson2016perspective}{2016}), takes 134 595 gene trees from phylota
    GenBank rel. 176 and estimates its degree of resolutin, calculating that less
    than half of clades are supported with minilam statistical support (0.53 ± 0.32).
  \end{itemize}
\end{enumerate}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We acknowledge contributions from

The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No.~ACI-1429783).

\hypertarget{references}{%
\section{References}\label{references}}

\newpage
\begin{center}
\textsc{References}
\end{center}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-altschul1990basic}{}%
Altschul S.F., Gish W., Miller W., Myers E.W., Lipman D.J. 1990. Basic local alignment search tool. Journal of molecular biology. 215:403--410.

\leavevmode\hypertarget{ref-anacker2011origins}{}%
Anacker B.L., Whittall J.B., Goldberg E.E., Harrison S.P. 2011. Origins and consequences of serpentine endemism in the california flora. Evolution: International Journal of Organic Evolution. 65:365--376.

\leavevmode\hypertarget{ref-antonelli2017toward}{}%
Antonelli A., Hettling H., Condamine F.L., Vos K., Nilsson R.H., Sanderson M.J., Sauquet H., Scharn R., Silvestro D., Töpel M., others. 2017. Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa. Systematic Biology. 66:152--166.

\leavevmode\hypertarget{ref-arnold201010ktrees}{}%
Arnold C., Matthews L.J., Nunn C.L. 2010. The 10kTrees website: A new online resource for primate phylogeny. Evolutionary Anthropology: Issues, News, and Reviews. 19:114--118.

\leavevmode\hypertarget{ref-baniaga2018polyploid}{}%
Baniaga A.E., Marx H.E., Arrigo N., Barker M.S. 2018. Polyploid plants have faster rates of multivariate climatic niche evolution than their diploid relatives. BioRxiv.:406314.

\leavevmode\hypertarget{ref-beaulieu2013fruit}{}%
Beaulieu J.M., Donoghue M.J. 2013. Fruit evolution and diversification in campanulid angiosperms. Evolution. 67:3132--3144.

\leavevmode\hypertarget{ref-beaulieu2012modeling}{}%
Beaulieu J.M., Jhwueng D.-C., Boettiger C., O'Meara B.C. 2012a. Modeling stabilizing selection: Expanding the ornstein--uhlenbeck model of adaptive evolution. Evolution: International Journal of Organic Evolution. 66:2369--2383.

\leavevmode\hypertarget{ref-beaulieu2012synthesizing}{}%
Beaulieu J.M., Ree R.H., Cavender-Bares J., Weiblen G.D., Donoghue M.J. 2012b. Synthesizing phylogenetic knowledge for ecological research. Ecology. 93:S4--S13.

\leavevmode\hypertarget{ref-bennett2018phylotar}{}%
Bennett D.J., Hettling H., Silvestro D., Zizka A., Bacon C.D., Faurby S., Vos R.A., Antonelli A. 2018. PhylotaR: An automated pipeline for retrieving orthologous dna sequences from genbank in r. Life. 8:20.

\leavevmode\hypertarget{ref-benson2000genbank}{}%
Benson D.A., Karsch-Mizrachi I., Lipman D.J., Ostell J., Rapp B.A., Wheeler D.L. 2000. GenBank. Nucleic acids research. 28:15--18.

\leavevmode\hypertarget{ref-blackmon2017synthesis}{}%
Blackmon H. 2017. Synthesis and phylogenetic comparative analyses of the causes and consequences of karyotype evolution in arthropods..

\leavevmode\hypertarget{ref-bruneau2019towards}{}%
Bruneau A., Borges L.M., Allkin R., Egan A.N., De La Estrella M., Javadi F., Klitgaard B., Miller J.T., Murphy D.J., Sinou C., others. 2019. Towards a new online species-information system for legumes. Australian Systematic Botany. 32:495--518.

\leavevmode\hypertarget{ref-camacho2009blast}{}%
Camacho C., George C., Vahram A., Ning M., Jason P., Kevin B., Thomas L. 2009. BLAST+: Architecture and applications. BMC bioinformatics. 10:421.

\leavevmode\hypertarget{ref-ceron2019phylotol}{}%
Cerón-Romero M.A., Maurer-Alcalá X.X., Grattepanche J.-D., Yan Y., Fonseca M.M., Katz L. 2019. PhyloToL: A taxon/gene-rich phylogenomic pipeline to explore genome evolution of diverse eukaryotes. Molecular biology and evolution. 36:1831--1842.

\leavevmode\hypertarget{ref-chen2008phylofinder}{}%
Chen D., Burleigh J.G., Bansal M.S., Fernández-Baca D. 2008. PhyloFinder: An intelligent search engine for phylogenetic tree databases. BMC Evolutionary Biology. 8:90.

\leavevmode\hypertarget{ref-chen2016tree}{}%
Chen Z.-D., Yang T., Lin L., Lu L.-M., Li H.-L., Sun M., Liu B., Chen M., Niu Y.-T., Ye J.-F., others. 2016. Tree of life for the genera of chinese vascular plants. Journal of Systematics and Evolution. 54:277--306.

\leavevmode\hypertarget{ref-chesters2017construction}{}%
Chesters D. 2017. Construction of a species-level tree of life for the insects and utility in taxonomic profiling. Systematic biology. 66:426--439.

\leavevmode\hypertarget{ref-chesters2019phylogeny}{}%
Chesters D. 2019. The phylogeny of insects in the data-driven era. Systematic Entomology.

\leavevmode\hypertarget{ref-chesters2013resolving}{}%
Chesters D., Vogler A.P. 2013. Resolving ambiguity of species limits and concatenation in multilocus sequence data for the construction of phylogenetic supermatrices. Systematic Biology. 62:456--466.

\leavevmode\hypertarget{ref-chesters2014protocol}{}%
Chesters D., Zhu C.-D. 2014. A protocol for species delineation of public dna databases, applied to the insecta. Systematic biology. 63:712--725.

\leavevmode\hypertarget{ref-crawford2012more}{}%
Crawford N.G., Faircloth B.C., McCormack J.E., Brumfield R.T., Winker K., Glenn T.C. 2012. More than 1000 ultraconserved elements provide evidence that turtles are the sister group of archosaurs. Biology letters. 8:783--786.

\leavevmode\hypertarget{ref-crete2012salmonidae}{}%
Crête-Lafrenière A., Weir L.K., Bernatchez L. 2012. Framing the salmonidae family phylogenetic portrait: A more complete picture from increased taxon sampling. PloS one. 7.

\leavevmode\hypertarget{ref-day1984efficient}{}%
Day W.H., Edelsbrunner H. 1984. Efficient algorithms for agglomerative hierarchical clustering methods. Journal of classification. 1:7--24.

\leavevmode\hypertarget{ref-deepak2010searchtree}{}%
Deepak A. 2010. SearchTree: Mining robust phylogenetic trees..

\leavevmode\hypertarget{ref-deepak2013managing}{}%
Deepak A. 2013. Managing and analyzing phylogenetic databases..

\leavevmode\hypertarget{ref-deepak2013extracting}{}%
Deepak A., Fernández-Baca D., McMahon M.M. 2013. Extracting conflict-free information from multi-labeled trees. Algorithms for Molecular Biology. 8:18.

\leavevmode\hypertarget{ref-deepak2014evominer}{}%
Deepak A., Fernández-Baca D., Tirthapura S., Sanderson M.J., McMahon M.M. 2014. EvoMiner: Frequent subtree mining in phylogenetic databases. Knowledge and Information Systems. 41:559--590.

\leavevmode\hypertarget{ref-drori2018onetwotree}{}%
Drori M., Rice A., Einhorn M., Chay O., Glick L., Mayrose I. 2018. OneTwoTree: An online tool for phylogeny reconstruction. Molecular ecology resources. 18:1492--1499.

\leavevmode\hypertarget{ref-dufort2015coexistence}{}%
Dufort M. 2015. Coexistence, ecomorphology, and diversification in the avian family picidae (woodpeckers and allies)..

\leavevmode\hypertarget{ref-dufort2016augmented}{}%
Dufort M.J. 2016. An augmented supermatrix phylogeny of the avian family picidae reveals uncertainty deep in the family tree. Molecular phylogenetics and evolution. 94:313--326.

\leavevmode\hypertarget{ref-edgar2004muscle}{}%
Edgar R.C. 2004. MUSCLE: Multiple sequence alignment with high accuracy and high throughput. Nucleic acids research. 32:1792--1797.

\leavevmode\hypertarget{ref-evans2019bony}{}%
Evans K.M., Vidal-García M., Tagliacollo V.A., Taylor S.J., Fenolio D.B. 2019. Bony patchwork: Mosaic patterns of evolution in the skull of electric fishes (apteronotidae: Gymnotiformes). Integrative and comparative biology. 59:420--431.

\leavevmode\hypertarget{ref-fan2015assembly}{}%
Fan H., Ives A.R., Surget-Groba Y., Cannon C.H. 2015. An assembly and alignment-free method of phylogeny reconstruction from next-generation sequencing data. BMC genomics. 16:522.

\leavevmode\hypertarget{ref-fang2019physpetree}{}%
Fang Y., Liu C., Lin J., Li X., Alavian K.N., Yang Y., Niu Y. 2019. PhySpeTree: An automated pipeline for reconstructing phylogenetic species trees. BMC evolutionary biology. 19:1--8.

\leavevmode\hypertarget{ref-faurby2016all}{}%
Faurby S., Eiserhardt W.L., Baker W.J., Svenning J.-C. 2016. An all-evidence species-level supertree for the palms (arecaceae). Molecular Phylogenetics and Evolution. 100:57--69.

\leavevmode\hypertarget{ref-faurby2015species}{}%
Faurby S., Svenning J.-C. 2015. A species-level phylogeny of all extant and late quaternary extinct mammals using a novel heuristic-hierarchical bayesian approach. Molecular phylogenetics and evolution. 84:14--26.

\leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
Felsenstein J. 1985. Confidence intervals on phylogenetics: An approach using bootstrap. Evolution. 39:783--791.

\leavevmode\hypertarget{ref-flores2011estimating}{}%
Flores O., Coomes D.A. 2011. Estimating the wood density of species for carbon stock assessments. Methods in Ecology and Evolution. 2:214--220.

\leavevmode\hypertarget{ref-frazao2015gondwana}{}%
Frazao A., Silva H.R. da, Moraes Russo C.A. de. 2015. The gondwana breakup and the history of the atlantic and indian oceans unveils two new clades for early neobatrachian diversification. PloS one. 10.

\leavevmode\hypertarget{ref-freyman2015sumac}{}%
Freyman W.A. 2015. SUMAC: Constructing phylogenetic supermatrices and assessing partially decisive taxon coverage. Evolutionary Bioinformatics. 11:EBO--S35384.

\leavevmode\hypertarget{ref-freyman2017phylogenetic}{}%
Freyman W.A. 2017. Phylogenetic models linking speciation and extinction to chromosome and mating system evolution..

\leavevmode\hypertarget{ref-gagnon2016new}{}%
Gagnon E., Bruneau A., Hughes C.E., Queiroz L.P. de, Lewis G.P. 2016. A new generic system for the pantropical caesalpinia group (leguminosae). PhytoKeys.:1.

\leavevmode\hypertarget{ref-gagnon2016systematique}{}%
Gagnon E., others. 2016. Systématique et biogéographie du groupe caesalpinia (famille leguminosae)..

\leavevmode\hypertarget{ref-gao2011solution}{}%
Gao Y., Meng Z., He X., Liu Y., Zhou Y., Li J. 2011. A solution to integrate data for phylogenetic research. 2011 5th international conference on bioinformatics and biomedical engineering.:1--4.

\leavevmode\hypertarget{ref-garcia2016role}{}%
García-Navas V., Bonnet T., Bonal R., Postma E. 2016. The role of fecundity and sexual selection in the evolution of size and sexual size dimorphism in new world and old world voles (rodentia: Arvicolinae). Oikos. 125:1250--1260.

\leavevmode\hypertarget{ref-gomez2020speciation}{}%
Gómez-Bahamón V., Márquez R., Jahn A.E., Miyaki C.Y., Tuero D.T., Laverde-R O., Restrepo S., Cadena C.D. 2020. Speciation associated with shifts in migratory behavior in an avian radiation. Current Biology.

\leavevmode\hypertarget{ref-gomez2015behavioral}{}%
Gómez Bahamón V., others. 2015. A behavioral polymorphism as an intermediate stage in the evolution of divergent forms-partial migration in new world flycatchers (aves, tyrannidae)..

\leavevmode\hypertarget{ref-grant2014building}{}%
Grant J.R., Katz L.A. 2014. Building a phylogenomic pipeline for the eukaryotic tree of life-addressing deep phylogenies with genome-scale data. PLoS currents. 6.

\leavevmode\hypertarget{ref-greenberg2011caryophyllaceae}{}%
Greenberg A.K., Donoghue M.J. 2011. Molecular systematics and character evolution in caryophyllaceae. Taxon. 60:1637--1652.

\leavevmode\hypertarget{ref-legume2013legume}{}%
Group L.P.W., Bruneau A., Doyle J.J., Herendeen P., Hughes C., Kenicer G., Lewis G., Mackinder B., Pennington R.T., Sanderson M.J., others. 2013. Legume phylogeny and classification in the 21st century: Progress, prospects and lessons for other species--rich clades. Taxon. 62:217--248.

\leavevmode\hypertarget{ref-guy2017phyloskeleton}{}%
Guy L. 2017. PhyloSkeleton: Taxon selection, data retrieval and marker identification for phylogenomics. Bioinformatics. 33:1230--1232.

\leavevmode\hypertarget{ref-hardy2010gall}{}%
Hardy N.B., Cook L.G. 2010. Gall-induction in insects: Evolutionary dead-end or speciation driver? BMC evolutionary biology. 10:257.

\leavevmode\hypertarget{ref-hardy2012testing}{}%
Hardy N.B., Cook L.G. 2012. Testing for ecological limitation of diversification: A case study using parasitic plants. The American Naturalist. 180:438--449.

\leavevmode\hypertarget{ref-hardy2014specialization}{}%
Hardy N.B., Otto S.P. 2014. Specialization and generalization in the diversification of phytophagous insects: Tests of the musical chairs and oscillation hypotheses. Proceedings of the Royal Society B: Biological Sciences. 281:20132960.

\leavevmode\hypertarget{ref-hardy2015evolution}{}%
Hardy N.B., Peterson D.A., Dohlen C.D. von. 2015. The evolution of life cycle complexity in aphids: Ecological optimization or historical constraint? Evolution. 69:1423--1432.

\leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
Helmus M.R., Ives A.R. 2012. Phylogenetic diversity--area curves. Ecology. 93:S31--S43.

\leavevmode\hypertarget{ref-helmus2010communities}{}%
Helmus M.R., Keller W., Paterson M.J., Yan N.D., Cannon C.H., Rusak J.A. 2010. Communities contain closely related species during ecosystem disturbance. Ecology letters. 13:162--174.

\leavevmode\hypertarget{ref-ickert2009fossil}{}%
Ickert-Bond S.M., Rydin C., Renner S.S. 2009. A fossil-calibrated relaxed clock for ephedra indicates an oligocene age for the divergence of asian and new world clades and miocene dispersal into south america. Journal of Systematics and Evolution. 47:444--456.

\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
Izquierdo-Carrasco F., Cazes J., Smith S.A., Stamatakis A. 2014. PUmPER: Phylogenies updated perpetually. Bioinformatics. 30:1476--1477.

\leavevmode\hypertarget{ref-jamil2016visual}{}%
Jamil H.M. 2016. A visual interface for querying heterogeneous phylogenetic databases. IEEE/ACM transactions on computational biology and bioinformatics. 14:131--144.

\leavevmode\hypertarget{ref-jones2014trends}{}%
Jones T.M., Baxter D.G., Hagedorn G., Legler B., Gilbert E., Thiele K., Vargas-Rodriguez Y., Urbatsch L.E. 2014. Trends in access of plant biodiversity data revealed by google analytics. Biodiversity data journal.

\leavevmode\hypertarget{ref-kamarudin2016phylogenetic}{}%
Kamarudin K.R., Rehan A.M., Hashim R., Usup G., Rehan M.M. 2016. Phylogenetic relationships within the genus holothuria inferred from 16S mitochondiral rRNA gene sequences. Sains Malaysiana. 45:1079--1087.

\leavevmode\hypertarget{ref-kolmann2017dna}{}%
Kolmann M.A., Elbassiouny A.A., Liverpool E.A., Lovejoy N.R. 2017. DNA barcoding reveals the diversity of sharks in guyana coastal markets. Neotropical Ichthyology. 15.

\leavevmode\hypertarget{ref-kumar2015bir}{}%
Kumar S., Krabberød A.K., Neumann R.S., Michalickova K., Zhao S., Zhang X., Shalchian-Tabrizi K. 2015. BIR pipeline for preparation of phylogenomic data. Evolutionary Bioinformatics. 11:EBO--S10189.

\leavevmode\hypertarget{ref-lehtonen2011towards}{}%
Lehtonen S. 2011. Towards resolving the complete fern tree of life. PLoS One. 6.

\leavevmode\hypertarget{ref-lehtonen2016sensitive}{}%
Lehtonen S., Christenhusz M.J., Falck D. 2016. Sensitive phylogenetics of clematis and its position in ranunculaceae. Botanical Journal of the Linnean Society. 182:825--867.

\leavevmode\hypertarget{ref-lehtonen2017environmentally}{}%
Lehtonen S., Silvestro D., Karger D.N., Scotese C., Tuomisto H., Kessler M., Peña C., Wahlberg N., Antonelli A. 2017. Environmentally driven extinction and opportunistic origination explain fern diversification patterns. Scientific Reports. 7:1--12.

\leavevmode\hypertarget{ref-lemoine2018renewing}{}%
Lemoine F., Entfellner J.-B.D., Wilkinson E., Correia D., Felipe M.D., De Oliveira T., Gascuel O. 2018. Renewing felsenstein's phylogenetic bootstrap in the era of big data. Nature. 556:452--456.

\leavevmode\hypertarget{ref-li2013partfasttree}{}%
Li J., Meng Z., Hou Y., Zhou Y., Gao Y. 2013. PartFastTree: Constructing large phylogenetic trees and estimating their reliability. 2013 ninth international conference on natural computation (icnc).:1052--1056.

\leavevmode\hypertarget{ref-mahmood2015avian}{}%
Mahmood M.T. 2015. Avian raptor evolution..

\leavevmode\hypertarget{ref-martin2011trophic}{}%
Martin C.H., Wainwright P.C. 2011. Trophic novelty is linked to exceptional rates of morphological diversification in two adaptive radiations of cyprinodon pupfish. Evolution: International Journal of Organic Evolution. 65:2197--2212.

\leavevmode\hypertarget{ref-mcmahon2015stbase}{}%
McMahon M.M., Deepak A., Fernández-Baca D., Boss D., Sanderson M.J. 2015. STBase: One million species trees for comparative biology. PloS one. 10.

\leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
McTavish E.J., Hinchliff C.E., Allman J.F., Brown J.W., Cranston K.A., Holder M.T., Rees J.A., Smith S.A. 2015. Phylesystem: A git-based data store for community-curated phylogenetic estimates. Bioinformatics. 31:2794--2800.

\leavevmode\hypertarget{ref-meng2015darwintree}{}%
Meng Z., Dong H., Li J., Chen Z., Zhou Y., Wang X., Zhang S. 2015a. Darwintree: A molecular data analysis and application environment for phylogenetic study. Data Science Journal. 14.

\leavevmode\hypertarget{ref-meng2015solution}{}%
Meng Z., Li J., Chen Z. 2015b. A solution to phylogeny assembly for ecologists. 2015 12th international conference on fuzzy systems and knowledge discovery (fskd).:1103--1107.

\leavevmode\hypertarget{ref-meng2015sotree}{}%
Meng Z., Li J., Yang T., Lin L., Chen Z. 2015c. SoTree: An automated phylogeny assembly tool for ecologists from big tree. 2015 ieee international conference on smart city/socialcom/sustaincom (smartcity).:792--797.

\leavevmode\hypertarget{ref-meng2012gsqct}{}%
Meng Z., Li J., Zhou Y., Cao W., Xiao X., Zhao J., Dong H., Zhang S. 2012a. GSQCT: A solution to screening gene sequences for phylogenetics analysis. 2012 9th international conference on fuzzy systems and knowledge discovery.:2929--2933.

\leavevmode\hypertarget{ref-meng2014rapidtree}{}%
Meng Z., Shao J., Cao W., Li J., Zhou Y., Wang X. 2014. RapidTree: A solution to rapid reconstruction phylogenetic tree. 2014 11th international conference on fuzzy systems and knowledge discovery (fskd).:513--517.

\leavevmode\hypertarget{ref-meng2012cloud}{}%
Meng Z., Xiao X., Li J., Zhou Y., Cao W., Shen G. 2012b. Cloud-gsqct: A parallel approach to screen gene sequences for phylogenetics analysis. 2012 international conference on computer science and information processing (csip).:660--663.

\leavevmode\hypertarget{ref-morrison2009apicomplexa}{}%
Morrison D.A. 2009. Evolution of the apicomplexa: Where are we now? Trends in parasitology. 25:375--382.

\leavevmode\hypertarget{ref-nazaire2012broad}{}%
Nazaire M., Hufford L. 2012. A broad phylogenetic analysis of boraginaceae: Implications for the relationships of mertensia. Systematic Botany. 37:758--783.

\leavevmode\hypertarget{ref-page2011linking}{}%
Page R.D. 2011. Linking ncbi to wikipedia: A wiki-based approach. PLoS currents. 3.

\leavevmode\hypertarget{ref-page2013bionames}{}%
Page R.D. 2013. BioNames: Linking taxonomy, texts, and trees. PeerJ. 1:e190.

\leavevmode\hypertarget{ref-papadopoulou2015automated}{}%
Papadopoulou A., Chesters D., Coronado I., De la Cadena G., Cardoso A., Reyes J.C., Maes J.-M., Rueda R.M., Gómez-Zurita J. 2015. Automated dna-based plant identification for large-scale biodiversity assessment. Molecular ecology resources. 15:136--152.

\leavevmode\hypertarget{ref-pereira2013padroes}{}%
Pereira J.E.S. 2013. Padrões e processos na evolução de primatas neotropicais (platyrrhini, primates)..

\leavevmode\hypertarget{ref-peters2011taming}{}%
Peters R.S., Meyer B., Krogmann L., Borner J., Meusemann K., Schütte K., Niehuis O., Misof B. 2011. The taming of an impossible child: A standardized all-in approach to the phylogeny of hymenoptera using public database sequences. BMC biology. 9:55.

\leavevmode\hypertarget{ref-piel2009treebase}{}%
Piel W., Chan L., Dominus M., Ruan J., Vos R., Tannen V. 2009. Treebase v. 2: A database of phylogenetic knowledge. E-biosphere..

\leavevmode\hypertarget{ref-rabosky2011species}{}%
Rabosky D.L., Cowan M.A., Talaba A.L., Lovette I.J. 2011. Species interactions mediate phylogenetic community structure in a hyperdiverse lizard assemblage from arid australia. The American Naturalist. 178:579--595.

\leavevmode\hypertarget{ref-ranwez2009phyloexplorer}{}%
Ranwez V., Clairon N., Delsuc F., Pourali S., Auberval N., Diser S., Berry V. 2009. PhyloExplorer: A web server to validate, explore and query phylogenetic trees. BMC evolutionary biology. 9:108.

\leavevmode\hypertarget{ref-roalson2016distinct}{}%
Roalson E.H., Roberts W.R. 2016. Distinct processes drive diversification in different clades of gesneriaceae. Systematic Biology. 65:662--684.

\leavevmode\hypertarget{ref-roquet2013building}{}%
Roquet C., Thuiller W., Lavergne S. 2013. Building megaphylogenies for macroecology: Taking up the challenge. Ecography. 36:13--26.

\leavevmode\hypertarget{ref-ross2013large}{}%
Ross L., Hardy N.B., Okusu A., Normark B.B. 2013. Large population size predicts the distribution of asexuality in scale insects. Evolution: International Journal of Organic Evolution. 67:196--206.

\leavevmode\hypertarget{ref-ruiz2019datataxa}{}%
Ruiz-Sanchez E., Maya-Lastra C.A., Steinmann V.W., Zamudio S., Carranza E., Murillo R.M., Rzedowski J. 2019. Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajío as a case study. Botanical Sciences. 97:754--760.

\leavevmode\hypertarget{ref-salariato2010filogenia}{}%
Salariato D.L. 2010. Filogenia y evolución de la subtribu melinidinae (paniceae: Panicoideae: Poaceae)..

\leavevmode\hypertarget{ref-salariato2010molecular}{}%
Salariato D.L., Zuloaga F.O., Giussani L.M., Morrone O. 2010. Molecular phylogeny of the subtribe melinidinae (poaceae: Panicoideae: Paniceae) and evolutionary trends in the homogenization of inflorescences. Molecular Phylogenetics and Evolution. 56:355--369.

\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
Sanderson M.J., Boss D., Chen D., Cranston K.A., Wehe A. 2008. The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. Systematic Biology. 57:335--346.

\leavevmode\hypertarget{ref-sanderson2010phylogenomics}{}%
Sanderson M.J., McMahon M.M., Steel M. 2010. Phylogenomics with incomplete taxon coverage: The limits to inference. BMC Evolutionary Biology. 10:155.

\leavevmode\hypertarget{ref-sanderson2016perspective}{}%
Sanderson M.J., Olson P., Hughes J., Cotton J. 2016. Perspective: Challenges in assembling the `next generation'Tree of life. Olson PD, Hughes J and Cotton JA.:13--27.

\leavevmode\hypertarget{ref-san2010molecular}{}%
San Mauro D., Agorreta A. 2010. Molecular systematics: A synthesis of the common methods and the state of knowledge. Cellular \& Molecular Biology Letters. 15:311.

\leavevmode\hypertarget{ref-santini2014first}{}%
Santini F., Carnevale G., Sorenson L. 2014. First multi-locus timetree of seabreams and porgies (percomorpha: Sparidae). Italian Journal of Zoology. 81:55--71.

\leavevmode\hypertarget{ref-santini2015first}{}%
Santini F., Carnevale G., Sorenson L. 2015. First timetree of sphyraenidae (percomorpha) reveals a middle eocene crown age and an oligo--miocene radiation of barracudas. Italian Journal of Zoology. 82:133--142.

\leavevmode\hypertarget{ref-santini2013first}{}%
Santini F., Sorenson L. 2013. First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies. Italian journal of zoology. 80:481--489.

\leavevmode\hypertarget{ref-sarkinen2013solanaceae}{}%
Särkinen T., Bohs L., Olmstead R.G., Knapp S. 2013. A phylogenetic framework for evolutionary study of the nightshades (solanaceae): A dated 1000-tip tree. BMC evolutionary biology. 13:214.

\leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
Schoch C.L., Sung G.-H., López-Giráldez F., Townsend J.P., Miadlikowska J., Hofstetter V., Robbertse B., Matheny P.B., Kauff F., Wang Z., others. 2009. The ascomycota tree of life: A phylum-wide phylogeny clarifies the origin and evolution of fundamental reproductive and ecological traits. Systematic biology. 58:224--239.

\leavevmode\hypertarget{ref-selvatti2015paleogene}{}%
Selvatti A.P., Gonzaga L.P., Moraes Russo C.A. de. 2015. A paleogene origin for crown passerines and the diversification of the oscines in the new world. Molecular phylogenetics and evolution. 88:1--15.

\leavevmode\hypertarget{ref-shi2015speciation}{}%
Shi J.J., Rabosky D.L. 2015. Speciation dynamics during the global radiation of extant bats. Evolution. 69:1528--1545.

\leavevmode\hypertarget{ref-simon2012cerrado}{}%
Simon M.F., Pennington T. 2012. Evidence for adaptation to fire regimes in the tropical savannas of the brazilian cerrado. International Journal of Plant Sciences. 173:711--723.

\leavevmode\hypertarget{ref-smith2009mega}{}%
Smith S.A., Beaulieu J.M., Donoghue M.J. 2009. Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches. BMC evolutionary biology. 9:37.

\leavevmode\hypertarget{ref-smith2018constructing}{}%
Smith S.A., Brown J.W. 2018. Constructing a broadly inclusive seed plant phylogeny. American Journal of Botany. 105:302--314.

\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
Smith S.A., Walker J.F. 2019. PyPHLAWD: A python tool for phylogenetic dataset construction. Methods in Ecology and Evolution. 10:104--108.

\leavevmode\hypertarget{ref-sorenson2014evolution}{}%
Sorenson L. 2014. Evolution of marine fish biodiversity: Phylogenomics and ecological processes shaping diversification..

\leavevmode\hypertarget{ref-sorenson2014effect}{}%
Sorenson L., Santini F., Alfaro M. 2014. The effect of habitat on modern shark diversification. Journal of Evolutionary Biology. 27:1536--1548.

\leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
Stamatakis A. 2014. RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. Bioinformatics. 30:1312--1313.

\leavevmode\hypertarget{ref-steppan2017muroid}{}%
Steppan S.J., Schenk J.J. 2017. Muroid rodent phylogenetics: 900-species tree reveals increasing diversification rates. PLoS One. 12.

\leavevmode\hypertarget{ref-stoltzfus2013phylotastic}{}%
Stoltzfus A., Lapp H., Matasci N., Deus H., Sidlauskas B., Zmasek C.M., Vaidya G., Pontelli E., Cranston K., Vos R., others. 2013. Phylotastic! Making tree-of-life knowledge accessible, reusable and convenient. BMC bioinformatics. 14:158.

\leavevmode\hypertarget{ref-strecker2014fish}{}%
Strecker A.L., Olden J.D. 2014. Fish species introductions provide novel insights into the patterns and drivers of phylogenetic structure in freshwaters. Proceedings of the Royal Society B: Biological Sciences. 281:20133003.

\leavevmode\hypertarget{ref-thomson2010sparse}{}%
Thomson R.C., Shaffer H.B. 2010. Sparse supermatrices for phylogenetic inference: Taxonomy, alignment, rogue taxa, and the phylogeny of living turtles. Systematic biology. 59:42--58.

\leavevmode\hypertarget{ref-tucker2012incorporating}{}%
Tucker C.M., Cadotte M.W., Davies T.J., Rebelo T.G. 2012. Incorporating geographical and evolutionary rarity into conservation prioritization. Conservation Biology. 26:593--601.

\leavevmode\hypertarget{ref-verbruggen2010data}{}%
Verbruggen H., Maggs C.A., Saunders G.W., Le Gall L., Yoon H.S., De Clerck O. 2010. Data mining approach identifies research priorities and data requirements for resolving the red algal tree of life. BMC evolutionary biology. 10:16.

\leavevmode\hypertarget{ref-vos2012nexml}{}%
Vos R.A., Balhoff J.P., Caravas J.A., Holder M.T., Lapp H., Maddison W.P., Midford P.E., Priyam A., Sukumaran J., Xia X., others. 2012. NeXML: Rich, extensible, and verifiable representation of comparative data and metadata. Systematic biology. 61:675--689.

\leavevmode\hypertarget{ref-webb2010biodiversity}{}%
Webb C.O., Slik J.F., Triono T. 2010. Biodiversity inventory and informatics in southeast asia. Biodiversity and Conservation. 19:955--972.

\leavevmode\hypertarget{ref-weber2014defense}{}%
Weber M.G., Agrawal A.A. 2014. Defense mutualisms enhance plant diversification. Proceedings of the National Academy of Sciences. 111:16442--16447.

\leavevmode\hypertarget{ref-wheeler2000database}{}%
Wheeler D.L., Chappey C., Lash A.E., Leipe D.D., Madden T.L., Schuler G.D., Tatusova T.A., Rapp B.A. 2000. Database resources of the national center for biotechnology information. Nucleic acids research. 28:10--14.

\leavevmode\hypertarget{ref-wu2008simple}{}%
Wu M., Eisen J.A. 2008. A simple, fast, and accurate method of phylogenomic inference. Genome biology. 9:R151.

\leavevmode\hypertarget{ref-xu2015ncbiminer}{}%
Xu X., Dimitrov D., Rahbek C., Wang Z. 2015. NCBIminer: Sequences harvest from genbank. Ecography. 38:426--430.

\leavevmode\hypertarget{ref-yong2010screening}{}%
Yong L., Zhen M., Qi L., Yanping G., Yuanchun Z., Jianhui L. 2010. Screening data for phylogenetic analysis of land plants: A parallel approach. 2010 first international conference on networking and distributed computing.:305--308.

\end{document}

### 2020-06-24 render2-submission-sysbio.tex
refactoring

--- 2020-06-24 docs/submission-sysbio-2020-06-24.tex
@@
   \item
-    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas et al. \protect\hyperlink{ref-garcia2016role}{2016})
+    A phylogeny of 34 vole species, Arvicolinae, Rodentia (Garcı́a-Navas et al. \protect\hyperlink{ref-garcia2016role}{2016})
   \item
@@
 \leavevmode\hypertarget{ref-evans2019bony}{}%
-Evans K.M., Vidal-García M., Tagliacollo V.A., Taylor S.J., Fenolio D.B.
+Evans K.M., Vidal-Garcı́a M., Tagliacollo V.A., Taylor S.J., Fenolio D.B.
 2019.
@@
 \leavevmode\hypertarget{ref-garcia2016role}{}%
-García-Navas V., Bonnet T., Bonal R., Postma E.
+Garcı́a-Navas V., Bonnet T., Bonal R., Postma E.
 2016.
@@
 2019.
-Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajío as a case study.
+Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajı́o as a case study.
 Botanical Sciences.

### 2020-06-24 submission_ams/submission_ams.Rmd
refactoring

---
journal: jhm
layout: draft
title: Title here
author1: Andrew N. Other
author2: Fred T. Secondauthor
currentaddress: "Current address: Some other place, Germany"
affiliation: "American Meteorological Society,Boston, Massachusetts"
exauthors: 
  - name: Ping Lu
    exaffiliation: Princeton University
    correspondingauthor: "American Meteorological Society, 45 Beacon St., Boston, MA 02108."
    email: \email{groupleader@unknown.uu}
  - name: Miao Yu
    exaffiliation: University of Waterloo
    currentaddress: "Current address: Some other place, Canada"
abstract: |
  some abstract
bibliography: ../paper.bib
csl: american-meteorological-society.csl
output: rticles::ams_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r body, child="../2-body.Rmd"}
```

### 2020-06-24 submission_ams/submission_ams.tex
refactoring

--- 2020-06-24 docs/submission-sysbio-2020-06-24.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
+\documentclass[draft]{ametsoc}
+\usepackage{color}
+\usepackage{hyperref}
+\journal{jhm}
+%  Please choose a journal abbreviation to use above from the following list:
 %
-\documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{setspace}
-\setstretch{2}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
+%   jamc     (Journal of Applied Meteorology and Climatology)
+%   jtech     (Journal of Atmospheric and Oceanic Technology)
+%   jhm      (Journal of Hydrometeorology)
+%   jpo     (Journal of Physical Oceanography)
+%   jas      (Journal of Atmospheric Sciences)
+%   jcli      (Journal of Climate)
+%   mwr      (Monthly Weather Review)
+%   wcas      (Weather, Climate, and Society)
+%   waf       (Weather and Forecasting)
+%   bams (Bulletin of the American Meteorological Society)
+%   ei    (Earth Interactions)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Citations should be of the form ``author year''  not ``author, year''
+\bibpunct{(}{)}{;}{a}{}{,}
+\title{Title here}
+
+\authors{Andrew N. Other
+and Fred T. Secondauthor
+\thanks{Current address: Some other place, Germany}}
+\affiliation{American Meteorological Society,Boston, Massachusetts}
+
+\extraauthor{Ping Lu
+\correspondingauthor{Ping Lu,American Meteorological Society, 45 Beacon St., Boston, MA 02108.}
 }
-\usepackage{hyperref}
-\hypersetup{
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage[margin = 1in]{geometry}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\usepackage{caption}
-\usepackage{blindtext}
-\usepackage[left]{lineno}
-\linenumbers
-
-\author{}
-\date{\vspace{-2.5em}}
-
+\extraaffil{Princeton University}
+\email{groupleader@unknown.uu}
+\extraauthor{Miao Yu
+\thanks{Current address: Some other place, Canada}
+}
+\extraaffil{University of Waterloo}
+
+% pandoc header
+\abstract{some abstract}
 \begin{document}
-
-\newpage
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Phylogenies are a key part of research in all areas of biology. Tools that automatize
-  some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
-  have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
-  However, interpretation of results, comparison with previously available phylogenetic
-  hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
-  to one that is not a specialist either on phylogenetic methods or on a particular group of study.
-\item
-  Physcraper is an open‐source, command-line Python program that automatizes the update of published
-  phylogenies by making use of public DNA sequence data and taxonomic information,
-  providing a framework for comparison of published phylogenies with their updated versions.
-\item
-  Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-  hypothesis based on already available expert phylogenetic knowledge.
-  Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
-  of alternative phylogenetic hypotheses (topologies).
-  \textbf{\emph{Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-  the ones that do not know about phylogenetic methods and the ones that might know
-  about phylogenetic methods but do not know much about a certain biological group.}}
-\item
-  Physcraper implements node by node/topology comparison of the the original and the updated
-  trees using the conflict API of OToL, and summarizes differences.
-\item
-  We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-\item
-  Physcraper can be used with any OS. Detailed instructions for installation and
-  use are available at \url{https://github.com/McTavishLab/physcraper}.
-\end{enumerate}
-
-\newpage
-
+\maketitle
 \hypertarget{introduction}{%
@@
 
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-\textbf{\emph{paragraph better to end discussion???
-}}
-
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
+Generating phylogenies is not easy and it is largely artisanal. Although
+many efforts to automatize the process have been done, and the community
+is using those more and more, automatization of phylogenetic
+reconstruction is still not a widespread practice and among other
+benefits, it might be key for adoption of better reproducibility
+practices in the phylogenetics community. \textbf{\emph{paragraph better
+to end discussion??? }}
+
+The process of phylogenetic reconstruction implies many steps (that I
+generalize to the following):
 
@@
 \item
-  Obtention of molecular or morphological character data -- get DNA from some organisms
-  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
-\item
-  Assemble a hypothesis of homology -- Create a matrix of your character data, by
-  aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-\item
-  Analyse this hypothesis of homology to infer phylogenetic relationships among
-  the organisms you are studying -- Use different available programs to infer molecular
-  evolution, trees and times of divergence.
-\item
-  Discuss the inferred relationships in the context of previous hypothesis, the
-  biology and biogeography of the organisms, etc. -- Answer the question, \emph{is this phylogenetic solution fair/reasonable?}
+  Obtention of molecular or morphological character data -- get DNA from
+  some organisms and sequence it, or get it from an online nucleotide
+  data repository, such as GenBank (Benson et al. 2000; Wheeler et al.
+  2000).
+\item
+  Assemble a hypothesis of homology -- Create a matrix of your character
+  data, by aligning the sequences, in the case of molecular data. Make
+  sure thay are paralogs!
+\item
+  Analyse this hypothesis of homology to infer phylogenetic
+  relationships among the organisms you are studying -- Use different
+  available programs to infer molecular evolution, trees and times of
+  divergence.
+\item
+  Discuss the inferred relationships in the context of previous
+  hypothesis, the biology and biogeography of the organisms, etc. --
+  Answer the question, \emph{is this phylogenetic solution
+  fair/reasonable?}
 \end{enumerate}
 
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
-
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}),
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
-
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
-
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
-
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
-
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered ``best'' by experts
-not only in phylogenetics, but also experts on the focal group of study.
-
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control along the process
-- Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
+Each of these steps require different types of specialized training: in
+the field, in the lab, in front of a computer, discussions with experts
+in the methods, and/or in the biological group of study. All of these
+steps also require considerable amounts of time for training and
+implementation.
+
+In the past decade, various studies have developed solutions to
+automatize the first and second steps, by creating pipelines that mine
+already available molecular data from the GenBank repository (Benson et
+al. 2000; Wheeler et al. 2000), to obtain homologous characters that can
+be used for phylogenetic reconstruction. These tools have been presented
+as aid for the nonspecialist to decrease some of the difficulties in the
+generation of phylogenetic knowledge. However, they are not that often
+used as so, suggesting that there are still difficulties for the
+nonspecialist. The phylogenetic community has some reserves towards
+these tools, too. Mainly because they sometimes act as a black box.
+However, automatizing the assembly of the character data set is a
+crucial step towards reproducibility for a task that was otherwise
+primarily artisanal and hence largely non-reproducible.
+
+Even if it is hard to obtain phylogenies, we invest copious amounts of
+time and energy in generating them. Issues such as food security, global
+warming, global health are crucial to solve and phylogenies might help.
+There is a lot of phylogenetic knowledge already available in published
+peer-reviewed studies. In this sense, the non-specialists (and also the
+specialist) face a new problem: how do I choose the best phylogeny.
+
+Public phylogenies can be updated with the ever increasing amount of
+genetic data that is available on GenBank (Benson et al. 2000; Wheeler
+et al. 2000).
+
+We present a way to automatize and standardize the comparison of
+phylogenetic hypotheses and to allow reproducibility of this last step
+of the research process.
+
+A key aspect of the standard phylogenetic workflow is comparison with
+already existing phylogenetic hypotheses and with phylogenies that are
+considered ``best'' by experts not only in phylogenetics, but also
+experts on the focal group of study.
+
+Concerns I think people have about these tools: - Errors in
+identification of sequences - Little control along the process - Too
+much of a black box?
+
+Most of these phylogenies are being constructed by people learning about
+the methods, so they want to know what is going on.
+
+The pipelines are so powerful and they will give you an answer, but
+there is no way to assess if it is better than previous answers, it just
+assumes it is better because it used more data.
 
@@
 The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000})).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-The original tree is by default used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
+allowing a direct comparison between existing phylogenies and
+phylogenies that are constructed using new genetic data retrieved from a
+public nucleotide database (i.e., GenBank (Benson et al. 2000; Wheeler
+et al. 2000)).
+
+To achieve this, Physcraper uses the Open Tree of Life phylesystem and
+connects it to the TreeBase database, to (1) get the original DNA data
+set matrices (alignments) that produced a phylogeny that was published
+and then made available in the OToL database, (2) use this DNA
+alignments as a starting point to get new genetic data belonging to the
+focal group of study, to (3) finally update the phylogenetic
+relationships in the group.
+
+A less automated workflow is one in which the alignments that generated
+the published phylogeny are stored in other public database (such as
+DRYAD) or elsewhere (the users computer), and are provided by the users.
+
+The original tree is by default used as starting tree for the
+phylogenetic searches, but it can also be set as a full topological
+constraint or not used at all, depending on the goals of the user.
+
+Physcraper implements node by node comparison of the the original and
+the updated trees, using the conflict API of OToL.
 
@@
 \hypertarget{the-input-a-study-tree-and-an-alignment}{%
-\subsection{The input: a study tree and an alignment}\label{the-input-a-study-tree-and-an-alignment}}
+\subsection{The input: a study tree and an
+alignment}\label{the-input-a-study-tree-and-an-alignment}}
 
@@
 \item
-  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish et al. \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the `r rotl::tol\_about()\$num\_source\_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
-\item
-  The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}; Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}),
-  DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
-  either from the TreeBASE website (\url{https://treebase.org/})
-  or through the TreeBASE GitHub repository (SuperTreeBASE; \url{https://github.com/TreeBASE/supertreebase}).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
-\item
-  A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-\item
-  A ``.csv'' file with the summary of taxon name matching is produced for the user.
+  The study tree is a published phylogenetic tree stored in the OToL
+  database, phylesystem (McTavish et al. 2015). The main reason for this
+  is that trees in phylesystem have a set of user defined
+  characteristics that are essential for automatizing the phylogeny
+  update process. The most relevant of these being the definition of
+  ingroup and outgroup. Outgroup and ingroup taxa in the original tree
+  are identified and tagged. This allows to automatically set the root
+  for the updated tree on the next steps of the pipeline. A user can
+  choose from the `r rotl::tol\_about()\$num\_source\_trees' published
+  trees supporting the resolved node of the synthetic tree in the OToL
+  website (\textless{}\textgreater{}). If the tree you are interested in
+  updating is not in there, you can upload it via OToL's curator tool
+  (\textless{}https://tree.opentreeoflife.org/curator).
+\item
+  The alignment should be a gene alignment that was used to generate the
+  tree. The original alignments are usually stored in a public
+  repository such as TreeBase (Piel et al. 2009; Vos et al. 2012), DRYAD
+  (\url{http://datadryad.org/}), or the journal were the tree was
+  originally published. If the alignment is stored in TreeBase,
+  \texttt{physcraper} can download it directly, either from the TreeBASE
+  website (\url{https://treebase.org/}) or through the TreeBASE GitHub
+  repository (SuperTreeBASE;
+  \url{https://github.com/TreeBASE/supertreebase}). If the alignment is
+  on another repository, or provided personally by the owner, a copy of
+  it has to be downloaded by the user, and it's local path has to be
+  provided as an argument.
+\item
+  A taxon name matching step is performed to verify that all taxon names
+  on the tips of the tree are in the DNA character matrix and vice
+  versa.
+\item
+  A ``.csv'' file with the summary of taxon name matching is produced
+  for the user.
 \item
   Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-\item
-  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
+  Technically, just one matching name is needed to perform the searches.
+  Please, see next section.
+\item
+  A ``.tre'' file and a ``.fas'' file containing only the matched taxa
+  are generated and saved in the \texttt{inputs} folder to be used in
+  the following steps.
 \end{itemize}
@@
 \hypertarget{dna-sequence-search-and-cleaning}{%
-\subsection{DNA sequence search and cleaning}\label{dna-sequence-search-and-cleaning}}
+\subsection{DNA sequence search and
+cleaning}\label{dna-sequence-search-and-cleaning}}
 
@@
 \item
-  The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be provided by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (\url{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}).
-
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-\item
-  The Basic Local Alignment Search Tool, BLAST {[}Altschul et al. (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
-\item
-  A pairwise alignment-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-\item
-  The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-\item
-  The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
-\item
-  Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
-\item
-  Because the original alignments usually lack database id numbers, a filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
-\item
-  Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
+  The next step is to identify the search taxon within the reference
+  taxonomy. The search taxon will be used to constraint the DNA sequence
+  search on the nucleotide database within that taxonomic group. Because
+  we are using the NCBI nucleotide database, by default the reference
+  taxonomy is the NCBI taxonomy. The search taxon can be provided by the
+  user. If none is provided, then the search taxon is identified as the
+  Most Recent Common Ancestor (MRCA) of the matched taxa belonging to
+  the ingroup in the tree, that is also a named clade in the reference
+  taxonomy. This is known as the Most Recent Common Ancestral Taxon
+  (MRCAT; also referred in the literature as the Least Inclusive Common
+  Ancestral Taxon - LICA). The MRCAT can be different from the
+  phylogenetic MRCA when the latter is an unnamed clade in the reference
+  taxonomy. To automatically identify the MRCAT of a group of taxon
+  names, we make use of the OToL taxonomy tool
+  (\url{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}).
+
+  Users can provide a search taxon that is either a more or a less
+  inclusive clade relative to the ingroup of the original phylogeny. If
+  the search taxon is more inclusive, the sequence search will be
+  performed outside the MRCAT of the matched taxa, e.g., including all
+  taxa within the family or the order that the ingroup belongs to. If
+  the search taxon is a less inclusive clade, the users can focus on
+  enriching a particular clade/region within the ingroup of the
+  phylogeny.
+\item
+  The Basic Local Alignment Search Tool, BLAST {[}Altschul et al.
+  (1990); altschul1997gapped{]} is used to identify similarity between
+  DNA sequences within the search taxon in a nucleotide database, and
+  the remaining sequences on the alignment. The BLAST command line tools
+  (Camacho et al. 2009) are used for both web-based and local-database
+  searches. 
+\item
+  A pairwise alignment-against-all BLAST search is performed. This means
+  that each sequence in the alignment is BLASTed against DNA sequences
+  in a nucleotide database constrained to the search taxon. Results from
+  each one of these BLAST runs are recorded, and matched sequences are
+  saved along with their corresponding identification numbers (accesion
+  numbers in the case of the GenBank database). This information will be
+  used later to store the whole sequences in a dedicated library within
+  the physcraper folder, allowing for secondary analyses to run
+  significantly faster.
+\item
+  The DNA sequence similarity search can be done on a local database
+  that is easily setup by the user. In this case, the BLASTn algorithm
+  is used to performs the similarity search.
+\item
+  The search can also be performed remotely, on the NCBI database. In
+  this case, the bioPython BLAST algorithm is used to perform the
+  similarity search.
+\item
+  Matched sequences below an e-value, percentage similarity, and outside
+  a minimum and maximum length threshold are discarded. This filtering
+  leaves out genomic sequences. All acepted sequences are asigned an
+  internal identifier, and are further filtered.
+\item
+  Because the original alignments usually lack database id numbers, a
+  filtering step is needed. Accepted sequences that belong to the same
+  taxon of the query sequence, and that are either identical or shorter
+  than the original sequence are discarded. Only longer sequences
+  belonging to the same taxon as the orignal sequence will be considered
+  further for analysis.
+\item
+  Among the remaining filtered sequences, there are usually several
+  exemplars per taxon. Although it can be useful to keep some of them
+  to, for example, investigate monophyly within species, there can be
+  hundreds of exemplar sequences per taxon for some markers. To control
+  the number of sequences per taxon in downstream analyses, 5 sequences
+  per taxon are chosen at random. This number is set by default but can
+  be modified by the user.
 \item
@@
 \item
-  Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
-\textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
-\item
-  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
-\item
-  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
+  Users can choose to perform a more ``cycles'' of sequence similarity
+  search, by blasting the newly found sequences. This can be done
+  iteratively, but by default only sequences in the alignment are
+  blasted. \textbf{\emph{Is there an argument to control the number of
+  cycles of blast searches with new sequences?}}
+\item
+  Accepted sequences are downloaded in full, and stored as a local
+  database in a directory that is globally accesible
+  (physcraper/taxonomy), so they are accesible for further runs.
+\item
+  A fasta file containing all filtered and processed sequences resulting
+  from the BLAST search is generated for the user.
 \end{itemize}
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is implemented to perform alignments.
+  The software MUSCLE (Edgar 2004) is implemented to perform alignments.
 \item
@@
 \item
-  Then, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align new sequences. This ensures that the final alignment
-  follows the homology criteria established by the original alignment.
-\item
-  The final alignment is not further processed automatically. We encourage users
-  to check it either by eye and perform manual refinement or using any of the many
-  tools for alignment processing, to eliminate columns with no information.
+  Then, a MUSCLE profile alignment is performed, in which the original
+  alignment is used as a template to align new sequences. This ensures
+  that the final alignment follows the homology criteria established by
+  the original alignment.
+\item
+  The final alignment is not further processed automatically. We
+  encourage users to check it either by eye and perform manual
+  refinement or using any of the many tools for alignment processing, to
+  eliminate columns with no information.
 \end{itemize}
@@
 \hypertarget{tree-reconstruction-and-comparison}{%
-\subsection{Tree reconstruction and comparison}\label{tree-reconstruction-and-comparison}}
+\subsection{Tree reconstruction and
+comparison}\label{tree-reconstruction-and-comparison}}
 
@@
 \item
-  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
-  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
-  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine et al.
-(\protect\hyperlink{ref-lemoine2018renewing}{2018}).
-\item
-  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
-\item
-  Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
+  A gene tree is reconstructed for each alignment provided, using the
+  software RAxML (Stamatakis 2014) with 100 classic bootstrap
+  (Felsenstein 1985) replicates by default. The number of bootsrap
+  replicates can be modified by the user. Other type of bootstrap that I
+  think is not yet incorporated into physcraper is the Transfer
+  Bootstrap Expectation (TBE) recently proposed in Lemoine et al.
+  (2018).
+\item
+  The final result is an updated phylogenetic hypothesis for each of the
+  genes provided in the alignment.
+\item
+  Tips on all trees generated by physcraper are defined by a taxon name
+  space, allowing to perform comparisons and conflict analyses.
 \item
@@
 \item
-  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
-\item
-  For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
-\item
-  Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-\item
-  Conflict information can only be generated in the context of the whole Open Tree of
-  Life. Otherwise, it is not really possible to get conflict data.
-  \textbf{\emph{- One way to compare two independent phylogenetic trees is to compare them both to
-  the synthetic OToL and then measure how well they do against each other}}
+  Describe what a conflict analysis is: Node by node comparison of the
+  resulting clades compared to
+\item
+  For the conflict analysis to be meaningful, the root of the tree
+  ineeds to be accurately defined.
+\item
+  Currently, the root is determined by finding the parent node of the
+  sequences that do not belong to the ingroup/ search taxon. This
+  ensures a correct rooting of the tree even when the search taxon is
+  more inclusive than the ingroup.
+\item
+  Conflict information can only be generated in the context of the whole
+  Open Tree of Life. Otherwise, it is not really possible to get
+  conflict data. \textbf{\emph{- One way to compare two independent
+  phylogenetic trees is to compare them both to the synthetic OToL and
+  then measure how well they do against each other}}
 \end{itemize}
@@
 
-The genus \emph{Ilex} is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-\textbf{\emph{Contact authors? They seem old school, probably do not wanna share their data.}}
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
-Updating the oldest tree, we should get something very similar to the newest tree.
+The genus \emph{Ilex} is the only extant clade within the family
+Aquifoliaceae, order Aquifoliales of flowering plants. It encompasses
+between 400-600 living species. A review of litterature shows that there
+are three published phylogenetic trees, showing relationships within the
+hollies. The first one has been made available both on OToL phylesystem
+and synth tree, and on treeBASE, it samples 48 species. The second has
+not been made available anywhere, not even in supplementary data of the
+journal. \textbf{\emph{Contact authors? They seem old school, probably
+do not wanna share their data.}} The most recent one has been made
+available in the OToL Phylesystem and DRYAD. It is the best sampled yet,
+with 200 species. However, it has not been added to the syntehtic tree
+yet. This makes it a perfect case to test the basic functionalities of
+physcraper: we know that the sequences of the most recently published
+tree have been made available on the GenBank database (Benson et al.
+2000; Wheeler et al. 2000). Updating the oldest tree, we should get
+something very similar to the newest tree.
 
@@
 
-Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by Schoch et al. (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
-The first step, is to get the Open Tree of Life study id. There are some options to do this:
-- You can go to the Open Tree of Life website and browse until you find it, or
-- you can get the study id using R tools:
-- By using the TreeBase ID of the study (which is not fully exposed on the
-TreeBase website home page of the study, so you have to really look it up manually):
+Let's be more specific now about our X group and say it is the
+Ascomycota. The best tree currently available in OToL was published by
+Schoch et al. (2009). The first step, is to get the Open Tree of Life
+study id. There are some options to do this: - You can go to the Open
+Tree of Life website and browse until you find it, or - you can get the
+study id using R tools: - By using the TreeBase ID of the study (which
+is not fully exposed on the TreeBase website home page of the study, so
+you have to really look it up manually):
 
@@
 \item
-  By using the name of the focal clade of study (but this behaved very differently):
+  By using the name of the focal clade of study (but this behaved very
+  differently):
 \end{itemize}
@@
 
-Once we have the study id, we can gather the trees published on that study:
+Once we have the study id, we can gather the trees published on that
+study:
 
@@
 \item
-  Download the alignment from TreeBase
-  If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
-  To make this task easier, you can use a command to download everything into your working folder:
+  Download the alignment from TreeBase If you are on the TreeBase home
+  page of the study, you can navigate to the matrix tab, and manually
+  download the alignments that were used to reconstruct the trees
+  reported on the study that were also uploaded to TreeBase and to the
+  Open Tree of Life repository. To make this task easier, you can use a
+  command to download everything into your working folder:
 \end{enumerate}
@@
 
-In this example, all alignments posted on TreeBase were used to reconstruct both trees.
+In this example, all alignments posted on TreeBase were used to
+reconstruct both trees.
 
@@
 \item
-  With the study id and the alignment files saved locally, we can do a physcraper run with the command:
+  With the study id and the alignment files saved locally, we can do a
+  physcraper run with the command:
 \end{enumerate}
@@
 
-Phylogeny of the Testudines 6 tips from Crawford et al. (\protect\hyperlink{ref-crawford2012more}{2012})
-There is just one tree in OToL.
-There is just one alignment on \href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase} with all the 1 145 loci.
+Phylogeny of the Testudines 6 tips from Crawford et al. (2012) There is
+just one tree in OToL. There is just one alignment on
+\href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase}
+with all the 1 145 loci.
 
@@
 
-Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
-
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
+Data repositories hold more information than meets the eye. Besides the
+actual data, they have other types of information that can be used for
+the advantage of science.
+
+Usually, initial ideas about the data are changed by analyses. We expect
+that this new ideas on the data can be registered on data bases,
 exposing new comers to expert understanding about the data.
 
-There are many tools that are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
-Most focus on generating full trees \emph{de novo}, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have been already studied and that have phylogenetic data already.
-
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
-
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies \emph{de novo}.
-
-Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
-Compare and discuss.
-
-How is physcraper already useful:
-- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
-- Finding
-
-How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in Webb et al. (\protect\hyperlink{ref-webb2010biodiversity}{2010})
-- obtain trees for ecophylogenetic studies, as mentioned in Helmus and Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
-- one day could be used to sistematize nucleotide databases, such as Genbank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro and Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus et al. (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})
-
-Things that physcraper does not do:
-- analyse the whole GenBank database (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli et al.
-(\protect\hyperlink{ref-antonelli2017toward}{2017}).
-There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014})
+There are many tools that are making use of DNA data repositories in
+different ways. Most of them focus on efficient ways to mine the data --
+getting the most homologs. Some focus on accurate ways of mining the
+data - getting real and clean homologs. Others focus on refinement of
+the alignment. Most focus on generating full trees \emph{de novo},
+mainly for regions of the Tree of Life that have no phylogenetic
+assessment yet in published studies, but also for regions that have been
+already studied and that have phylogenetic data already.
+
+All these tools are great efforts for advancing towards reproducibility
+in phylogenetics, a field that has been largely recognised as somewhat
+artisanal. We propose adding focus to other sources of information
+available from data repositories. Taking advantage of public DNA data
+bases have been the main focus. However, phylogenetic knowledge is also
+accumulating fast in public and open repositories. In this way, the
+physcraper pipeline can be complemented with other tools that have been
+developed for other purposes.
+
+We emphasize that physcraper takes advantage of the knowledge and
+intuition of the expert community to build upon phylogenetic knowledge,
+using not only data accumulated in DNA repositories, but phylogenetic
+knowledge accumulated in tree repositories. This might help generate new
+phylogenetic data. But physcraper does not seek to generate full
+phylogenies \emph{de novo}.
+
+Describe again statistics to compare phylogenies provided by physcraper
+via OpenTreeOfLife. Mention statistics provided by other tools:
+PhyloExplorer (Ranwez et al. 2009). Compare and discuss.
+
+How is physcraper already useful: - to mine targeted sequences, in this
+way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota
+does not do baited analyses, I think, only clustered analyses. - Finding
+
+How can it be used for the advantage of the field: - rapid phylogenetic
+placing of newly discovered species, as mentioned in Webb et al. (2010)
+- obtain trees for ecophylogenetic studies, as mentioned in Helmus and
+Ives (2012) - one day could be used to sistematize nucleotide databases,
+such as Genbank (Benson et al. 2000; Wheeler et al. 2000), as mentioned
+in San Mauro and Agorreta (2010), i.e., curate ncbi taxonomic
+assignations. - allows to generate custom species trees for downstream
+analyses, as mentioned in Stoltzfus et al. (2013)
+
+Things that physcraper does not do: - analyse the whole GenBank database
+(Benson et al. 2000; Wheeler et al. 2000) to find homolog regions
+suitable to reconstruct phylogenies, as mentioned in Antonelli et al.
+(2017). There are already some very good tools that do that. - provide
+basic statistics on data availability to assemble molecular datasets, as
+mentioned by Ranwez et al. (2009). Phyloexplorer does this? - it is not
+a tree repo, as phylota is, mentioned in Deepak et al. (2014)
 
 \hypertarget{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}{%
-\subsection{Tools that automatize any part of the process of phylogenetic reconstruction:}\label{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}}
+\subsection{Tools that automatize any part of the process of
+phylogenetic
+reconstruction:}\label{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}}
 
 \hypertarget{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}{%
-\subsubsection{1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction}\label{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}}
+\subsubsection{1. Mining DNA databases to generate datasets suitable for
+phylogenetic
+reconstruction}\label{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}}
 
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Sanderson et al. (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut
+Sanderson et al. (2008)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction\strut
+finds sets of DNA homologs on the GenBank database; phylogenetic
+reconstruction\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Wu and Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut
+Wu and Eisen (2008)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction\strut
+baited search; protein markers on phylogenomic data; personal database
+of genomes or metagenomic data, manually downloaded either from a public
+database or from private data; phylogenetic reconstruction\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith et al. (\protect\hyperlink{ref-smith2009mega}{2009})\strut
+Smith et al. (2009)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Baited search of DNA markers on the GenBank database; phylogenetic reconstruction\strut
+Baited search of DNA markers on the GenBank database; phylogenetic
+reconstruction\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \begin{minipage}[t]{0.12\columnwidth}\raggedright
-Unnamed \href{https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals}{ruby pipeline}, only available from \href{https://static-content.springer.com/esm/art\%3A10.1186\%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP}{supplementary data} of the journal\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Peters et al. (\protect\hyperlink{ref-peters2011taming}{2011})\strut
+Unnamed
+\href{https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals}{ruby
+pipeline}, only available from
+\href{https://static-content.springer.com/esm/art\%3A10.1186\%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP}{supplementary
+data} of the journal\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
+Peters et al. (2011)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established ``criteria of compositional homogeneity and defined levels of density and overlap''\strut
+mining public DNA databases, focuses on filtering massive amounts of
+mined sequences by using established ``criteria of compositional
+homogeneity and defined levels of density and overlap''\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Grant and Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut
+Grant and Katz (2014)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position\strut
+predecessor of phylotol; homolog clustering; public and/or personal DNA
+database; phylogenetic reconstruction; broad taxon analyses; remove
+contaminant sequences, based on similarity and on phylogenetic
+position\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut
+Chesters and Zhu (2014)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-algorithm that mines GenBank data to delineate species in the insecta. The authors present a nice comparison with the phylota algorithm\strut
+algorithm that mines GenBank data to delineate species in the insecta.
+The authors present a nice comparison with the phylota algorithm\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Izquierdo-Carrasco et al. (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut
+Izquierdo-Carrasco et al. (2014)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Meng et al. (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut
+Meng et al. (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only\strut
+predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP),
+takes data from GenBank, EMBL and DDBJ for land plants only\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Xu et al. (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut
+Xu et al. (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015})\strut
+Freyman (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-both ``baited'' analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix\strut
+both ``baited'' analyses and single‐linkage clustering methods, as well
+as a novel means of determining when there are enough overlapping data
+in the DNA matrix\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-McMahon et al. (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut
+McMahon et al. (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-pipeline for species tree construction and the public database of one million precomputed species trees\strut
+pipeline for species tree construction and the public database of one
+million precomputed species trees\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Papadopoulou et al. (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut
+Papadopoulou et al. (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Automated DNA-based plant identification for large-scale biodiversity assessment\strut
+Automated DNA-based plant identification for large-scale biodiversity
+assessment\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Kumar et al. (\protect\hyperlink{ref-kumar2015bir}{2015})\strut
+Kumar et al. (2015)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-blast, align, identify homologs via constructed trees, curate and realign\strut
+blast, align, identify homologs via constructed trees, curate and
+realign\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut
+Antonelli et al. (2017)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-{[}Chesters (\protect\hyperlink{ref-chesters2017construction}{2017})\strut
+{[}Chesters (2017)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos\strut
+Searches DNA sequence data from repos other than GenBank, such as
+transcriptomic and barcoding repos\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Guy (\protect\hyperlink{ref-guy2017phyloskeleton}{2017})\strut
+Guy (2017)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification\strut
+focuses on taxon sampling; baited genomic sequences; public database
+(NCBI and JGI); marker identification\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut
+Drori et al. (2018)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Web‐based, user-friendly, online tool for species-tree reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available sequence data from NCBI GenBank\strut
+Web‐based, user-friendly, online tool for species-tree reconstruction,
+based on the \emph{supermatrix paradigm} and retrieves all available
+sequence data from NCBI GenBank\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith and Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut
+Smith and Walker (2019)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut
+Cerón-Romero et al. (2019)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-``phylogenomic pipeline to allow easy incorporation of data from high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life'', i.e., bacteria and unicellular eukaryotes\strut
+``phylogenomic pipeline to allow easy incorporation of data from
+high-throughput sequencing studies, to automate production of both
+multiple sequence alignments and gene trees, and to identify and remove
+contaminants. PhyloToL is designed for phylogenomic analyses of diverse
+lineages across the tree of life'', i.e., bacteria and unicellular
+eukaryotes\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Bennett et al. (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut
+Bennett et al. (2018)\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 
-According to Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
-``focus on the construction and refinement of robust alignments rather than the collection of homologs.''
+According to Cerón-Romero et al. (2019), PhyLoTA and BIR ``focus on the
+identification and collection of homologous and paralog genes from
+public databases such as GenBank'', while both AMPHORA and PHLAWD
+``focus on the construction and refinement of robust alignments rather
+than the collection of homologs.''
 
 \hypertarget{searching-phylogenetic-tree-databases}{%
-\subsubsection{2. Searching phylogenetic tree databases}\label{searching-phylogenetic-tree-databases}}
-
-PhyloFinder (Chen et al. \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
+\subsubsection{2. Searching phylogenetic tree
+databases}\label{searching-phylogenetic-tree-databases}}
+
+PhyloFinder (Chen et al. 2008) - cited by 18: a search engine for
+phylogenetic databases, using trees from TreeBASE - more related to
+phylotastic's goal than to updating/creating phylogenies
 
 \hypertarget{mining-phylogenetic-tree-databases}{%
-\subsubsection{3. Mining phylogenetic tree databases}\label{mining-phylogenetic-tree-databases}}
-
-PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database''.
+\subsubsection{3. Mining phylogenetic tree
+databases}\label{mining-phylogenetic-tree-databases}}
+
+PhyloExplorer (Ranwez et al. 2009) - cited by 21: a python and MySQL
+based website to facilitate assessment and management of phylogenetic
+tree collections. It provides ``statistics describing the collection,
+correcting invalid taxon names, extracting taxonomically relevant parts
+of the collection using a dedicated query language, and identifying
+related trees in the TreeBASE database''.
 
 \hypertarget{pipeline-for-phylogenetic-reconstruction}{%
-\subsubsection{4. Pipeline for phylogenetic reconstruction}\label{pipeline-for-phylogenetic-reconstruction}}
-
-PhySpeTre (Fang et al. \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
+\subsubsection{4. Pipeline for phylogenetic
+reconstruction}\label{pipeline-for-phylogenetic-reconstruction}}
+
+PhySpeTre (Fang et al. 2019) - no citations yet - no sequence retrieval,
+just phylogenetic reconstruction pipeline.
 
 \hypertarget{getting-metadata-and-not-sequences-from-genbank.}{%
-\subsubsection{5. getting metadata and not sequences from GenBank.}\label{getting-metadata-and-not-sequences-from-genbank.}}
-
-Datataxa Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.
+\subsubsection{5. getting metadata and not sequences from
+GenBank.}\label{getting-metadata-and-not-sequences-from-genbank.}}
+
+Datataxa Ruiz-Sanchez et al. (2019) - no citations yet - focus on
+extracting metadata from GenBank sequence information.
 
@@
 
-Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
-database (Sanderson et al. \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-
-Since then, it has been cited 122 times for different reasons.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  As an example of a tool that mines GenBank data for phylogenetic reconstruction,
-  or that is useful in any way for phylogenetics:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    original publication of PHLAWD (Smith et al. \protect\hyperlink{ref-smith2009mega}{2009})
-  \item
-    an analysis identifying research priorities and data requirements for resolving
-    the red algal tree of life (Verbruggen et al. \protect\hyperlink{ref-verbruggen2010data}{2010})
-  \item
-    Beaulieu et al. (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive
-    phylogeny from mined DNA sequence data, (even if no phylogeny was really published
-    there, only the method to do so)
-  \item
-    a review for ecologists about phylogenetic tools (Roquet et al. \protect\hyperlink{ref-roquet2013building}{2013})
-  \item
-    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith and Brown \protect\hyperlink{ref-smith2018constructing}{2018})
-  \item
-    a study presenting an ``assembly and alignment free'' method for phylogenetic reconstruction
-    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan et al. \protect\hyperlink{ref-fan2015assembly}{2015}).
-  \item
-    nexml format presentation (Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses
-    stored phyloinformatic data that could benefit from adopting nexml, to increase
-    interoperability.
-  \item
-    a study of fruit evolution, analysing a previously published phylogeny of 8911
-    tips of the Campanulidae, constructed with PHLAWD (Beaulieu and Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})
-  \item
-    a study of Southeast Asia plant biodiversity inventory (Webb et al. \protect\hyperlink{ref-webb2010biodiversity}{2010}) -
-    cites phylota as a tool that would allow rapid phylogenetic placing of newly
-    discovered species, and generation of phylogenetically informed guides for field
-    identification.
-  \item
-    a study of wood density for carbon stock assessments (Flores and Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),
-    cites phylota as an initiative to ``get supertrees resolved up to species level''.
-  \item
-    a study proposing something similar to Open tree but applied only to land plants (Beaulieu et al.
-\protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})
-  \item
-    an analysis of the phylogenetic diversity-area curve (Helmus and Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-    cited phylota as a method alternative to phylomatic to ``obtain plant phylogenetic
-    trees for ecophylogenetic studies''.
-  \item
-    a study generating a phylogeny of 6,098 species of vascular plants from China
-    (Chen et al.
-\protect\hyperlink{ref-chen2016tree}{2016}) - uses DarwinTree (Meng et al.
-\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}) and generates sequence
-    data \emph{de novo} for 781 genera.
-  \item
-    a review of the state of methods and knowledge generated by molecular systematics
-    (San Mauro and Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
-    for large-scale molecular phylogenetics analysis''.
-  \item
-    the first phylotastic paper (Stoltzfus et al. \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny
-    related resource that provides ways to generate custom species trees for downstream use''.
-  \item
-    Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017}) cites phylota as a ``pipeline that pre-processes entire GenBank
-    releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
-    then clustered into candidate data sets''. They also use the PHYLOTA database in its
-    own pipeline.
-  \item
-    Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014}) present an algorithm for mining of frequent subtrees (common patterns)
-    in collections of phylogenetic trees, as a way to extract meaningful phylogenetic
-    information from collections of trees when compared to maximum agreement subtrees
-    and majority-rule trees. They cite phylota as one of such tree collections available
-    along with TreeBASE (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}).
-  \item
-    Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) cites phylota as a ``program providing basic statistics
-    on data availability for molecular datasets''. They propose a tool to upload and
-    explore user phylogenies to obtain detailed summary statistics on user tree collections.
-  \item
-    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}) cites phylota as a tool that ``provides a web interface to view
-    all GenBank sequences within taxonomic groups clustered into homologs'' but that
-    does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They
-    compare the performance of SUMAC to Phylota. This is also presented in their PhD dissertation (Freyman \protect\hyperlink{ref-freyman2017phylogenetic}{2017}).
-  \item
-    Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) cites phylota as a data mining tool that compiles metadata
-    from mining of public DNA databases ``for construction of large phylogenetic trees
-    and multiple gene sets'' and that the authors have recognised that gene annotations
-    in public databases are insufficient and that careful partitioning of orthologous
-    sequences is needed for supermatrix construction. Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) present
-    a procedure that minimizes the problem of forming multilocus species units in
-    a large phylogenetic data set using algorithms from graph theory.
-  \item
-    Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014}) present an algorithm to delineate species form GenBank
-    DNA data, and cites phylota as a tool that partitions ``the contents of a database
-    according to homology'', by ``grouping of database sequences according to internal
-    criteria'', searching ``from a standardized set of references {[}\ldots{}{]} patterns in
-    sequence similarity and overlap.''
-  \item
-    the paper presenting phylotaR, a pipeline that recreates the phylota output
-    but uses the most updated GenBank release, and is available in R (Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}),
-    cites phylota as its predecessor and inspiration. The authors mention that phylotaR
-    pipeline mimics phylota's pipeline but with improvements.
-  \item
-    The paper presenging PhyloBase (Jamil \protect\hyperlink{ref-jamil2016visual}{2016}), cites phylota as one of
-    its resources to get phylogenies, along with TreeBASE and others.
-  \item
-    The paper presenting STBase, a database of one million precomputed species
-    trees (Deepak \protect\hyperlink{ref-deepak2013managing}{2013}; McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as a databse of gene trees or mul-trees,
-    ``trees having multiple sequences with the same taxon name''.
-  \item
-    Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018}) present a Web‐based, user-friendly, online tool for species-tree
-    reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available
-    sequence data from NCBI GenBank. They cite phylota in the intro as a tool that is ``designed to provide
-    users with precomputed sets of clusters that were assembled through a single‐linkage
-    clustering approach and additionally provides precomputed gene trees that were
-    reconstructed for each cluster. In particular, the results obtained by PhyLoTa
-    are taxonomically constrained; that is, all sequences of the most recent common
-    ancestor are collected even if one specifies only part of a clade''.
-  \item
-    A study developing a tool to link wikipedia data to NCBI taxonomy (Page \protect\hyperlink{ref-page2011linking}{2011})
-    cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
-  \item
-    the study that present DarwinTree (Meng et al. \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}), and all derived studies: the study
-    presenting an approach to screen sequence data for The Platform
-    for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST (Yong et al.
-\protect\hyperlink{ref-yong2010screening}{2010}), as well as Gao et al.
-(\protect\hyperlink{ref-gao2011solution}{2011}), Li et al.
-(\protect\hyperlink{ref-li2013partfasttree}{2013}),
-    Meng et al.
-(\protect\hyperlink{ref-meng2014rapidtree}{2014}), Meng et al.
-(\protect\hyperlink{ref-meng2015sotree}{2015}\protect\hyperlink{ref-meng2015sotree}{c}), and Meng et al.
-(\protect\hyperlink{ref-meng2015solution}{2015}\protect\hyperlink{ref-meng2015solution}{b}), all cite phylota using the exact same
-    introduction and sentence: as one among other ``studies based on data mining large numbers of taxa or loci''.
-  \item
-    A study presenting a tool to asses gene sequence quality for automatic
-    construction of databases (Meng et al. \protect\hyperlink{ref-meng2012gsqct}{2012}\protect\hyperlink{ref-meng2012gsqct}{a}), as well as their parallelized version using MapReduce
-    (Meng et al.
-\protect\hyperlink{ref-meng2012cloud}{2012}\protect\hyperlink{ref-meng2012cloud}{b}), cite phylota (along with Yong et al.
-(\protect\hyperlink{ref-yong2010screening}{2010})) as a tool that relies on sequence
-    similarity (BLAST) and not taxon name annotations in the database, for mining
-    large numbers of taxa or loci, without making any control on the quality of the
-    sequencing.
-  \item
-    A review on online plant databases aiming to ``provide recommendations for current
-    information managers and developers concerning the user interface and experience;
-    and to provide a picture about the possible directions to take for those in charge
-    of the creation of information at all levels''. They cite phylota as a tool allowing
-    researchers ``to acces equally and globally, without travel, a {[}phylogenetic?{]} model
-    of plants at the kingdom level'' (Jones et al. \protect\hyperlink{ref-jones2014trends}{2014}).
-  \item
-    a paper aiming to establish an online information system for the legumes and
-    to outline ``best practices for development of a legume portal to enable data
-    sharing and a better understanding of what data are available, missing, or erroneous,
-    and ultimately facilitate cross-analyses and collaboration within the legume-systematics
-    community and with other stakeholders'' (Bruneau et al. \protect\hyperlink{ref-bruneau2019towards}{2019}), cites phylota (along with supersmart and pyphlawd) as a
-    ``pipeline for large-scale retrieval of GenBank data of particular taxa or clades''.
-    In their Table 1, they also list phylota as a potential data source for developing a legume portal.
-  \item
-    A study on morphological evolution of electric fish skull, that uses phylotaR
-    to retrieve sequences of the family Apteronotidae, order Gymnotiformes (Evans et al. \protect\hyperlink{ref-evans2019bony}{2019}),
-    cites phylota as the inspiration and fundament of phylotaR.
-  \item
-    A phylogenetic revision of the Gymnotidae fish (Teleostei: Gymnotiformes),
-    uses phylotaR to retrieve seqeunces, but cites phylota as ``a pipeline that implements
-    BLAST searches to both identify and download sequence clusters for listed taxonomic
-    groups to assemble a robust collection of sequences in a reproducible way based
-    on publicly-available gene sequences while avoiding selection bias on the part
-    of the assembler''.
-  \item
-    A master thesis on SearchTree, a ``software tool that allows users to query
-    efficiently on an arbitrary user taxon list and returns high scoring matches
-    from approximately one billion phylogenetic trees being constructed from molecular
-    sequence data in GenBank'' (Deepak \protect\hyperlink{ref-deepak2010searchtree}{2010}), that seems to be the preliminary
-    work for STBase (McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as ``a standard strategy, to assemble sets of homologous sequences
-    (clusters) from a database of all-against-all BLAST searches, {[}in which{]} clusters
-    are constructed in the context of the NCBI taxonomy tree for convenience of display,
-    thus child clusters are contained within parent clusters, following the NCBI hierarchy''.
-    In opposition, SearchTree uses true agglomerative hierarchical clustering
-    (AHC: Day and Edelsbrunner (\protect\hyperlink{ref-day1984efficient}{1984})) based on the BLAST estimates of sequence dissimilarity
-    rather than the NCBI tree".
-  \item
-    a recent review on the state of large phylogeny (namely insects) generation
-    using tools of the data-driven era (Chesters \protect\hyperlink{ref-chesters2019phylogeny}{2019}) cites phylota as
-    a tool for homology inference and retrieval.
-  \item
-    the study presenting phylotol (Cerón-Romero et al. \protect\hyperlink{ref-ceron2019phylotol}{2019}), cites phylota as a tool
-    that ``focus on the identification and collection of homologous genes from public
-    databases''.
-  \item
-    The \href{https://www.researchgate.net/profile/Douglas_Soltis/publication/228815637_Assembling_the_Tree_of_Life_for_the_Plant_Sciences_iPTOL/links/56a7c6bc08aeded22e3700dc.pdf}{iPTOL project}
-    cites phylota as a resource of phylogenetic trees.
-  \item
-    Mahmood (\protect\hyperlink{ref-mahmood2015avian}{2015}) PhD dissertation presents a database of avian Raptor sequences
-    (raptorbase), based on the phylota pipeline.
-  \item
-    Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) develops datataxa and cite phylota as ``software that has
-    been developed to mine the massive amount of information stored in GenBank'',
-    along with its R version (phylotaR; Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}) and restez \url{https://www.rdocumen-tation.org/packages/restez/versions/1.0.0}.
-  \item
-    The phylotastic project (Stoltzfus et al.
-\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny-related resource providing ways to generate custom species trees \emph{de novo} for downstream use'' along with CIPRES.
-  \end{itemize}
-\item
-  When the software was actually used to construct (partially or in full) a DNA
-  data set to be used for phylogenetic reconstruction:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A 1000 tip phylogeny of the family of the nightshades (Särkinen et al. \protect\hyperlink{ref-sarkinen2013solanaceae}{2013})
-  \item
-    A 56 tip phylogeny of crustacean zooplancton (Helmus et al. \protect\hyperlink{ref-helmus2010communities}{2010}) -- ecological study
-  \item
-    A 63 tip phylogeny of the Salmonidae family (Crête-Lafrenière et al. \protect\hyperlink{ref-crete2012salmonidae}{2012})
-  \item
-    A 321 tip phylogeny of Testudines (Thomson and Shaffer \protect\hyperlink{ref-thomson2010sparse}{2010})
-  \item
-    A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish (Martin and Wainwright \protect\hyperlink{ref-martin2011trophic}{2011})
-  \item
-    A 2,957 taxa phylogeny of the class Moniloformopses of living ferns (Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011})
-  \item
-    A 2,573 species phylogeny of the Papilionoidea (Hardy and Otto \protect\hyperlink{ref-hardy2014specialization}{2014})
-  \item
-    A 23 taxa phylogeny of the California flora (Anacker et al. \protect\hyperlink{ref-anacker2011origins}{2011})
-  \item
-    Phylogenies of 6 different clades of flowering plants representing an independent
-    evolutionary origin of extrafloral nectaries: \emph{Byttneria} (Malvaceae), \emph{Pleopeltis} (Polypodiaceae),
-    \emph{Polygoneae} (Polygoneaceae), \emph{Senna} (Fabaceae), \emph{Turnera} (Passifloraceae), and \emph{Viburnum}
-    (Adoxaceae) (Weber and Agrawal \protect\hyperlink{ref-weber2014defense}{2014}).
-  \item
-    To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
-    sampled at different taxonomic levels (Faurby and Svenning \protect\hyperlink{ref-faurby2015species}{2015})
-  \item
-    A 900 species tree of muroid rodents, Muroidea (Steppan and Schenk \protect\hyperlink{ref-steppan2017muroid}{2017}), where 300
-    species were newly added by the study and the rest obtained using phylota.
-  \item
-    A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales (Ickert-Bond et al. \protect\hyperlink{ref-ickert2009fossil}{2009})
-  \item
-    A 1061 genera phylogeny of the Oscine birds (Selvatti et al. \protect\hyperlink{ref-selvatti2015paleogene}{2015})
-  \item
-    A 268 species phylogeny of sharks, representing all 8 orders and 32 families (Sorenson et al.
-\protect\hyperlink{ref-sorenson2014effect}{2014}; Sorenson \protect\hyperlink{ref-sorenson2014evolution}{2014})
-  \item
-    A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region (Tucker et al. \protect\hyperlink{ref-tucker2012incorporating}{2012}).
-  \item
-    A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects (Hardy and Cook \protect\hyperlink{ref-hardy2010gall}{2010}).
-  \item
-    A 196 species phylogeny of the family Boraginaceae (Nazaire and Hufford \protect\hyperlink{ref-nazaire2012broad}{2012}). The authors
-    actually found data for 318 Boraginaceae spp using phylota, but decided to reduce
-    their data set to focus on the monophyly of genus \emph{Mertensia}.
-  \item
-    A phylogeny of 401 species of scale insects Coccoidea, Hemiptera (Ross et al. \protect\hyperlink{ref-ross2013large}{2013}),
-    with some sequences generated \emph{de novo}.
-  \item
-    Two phylogenies sampling all species of two different clades of insectivorous
-    lizards, agamids and diplodactyline geckos, groups considered to be radiating
-    in the Australia's Great Victoria Desert (Rabosky et al. \protect\hyperlink{ref-rabosky2011species}{2011})
-  \item
-    A phylogeny of 91 species of sparid and centracanthid fishes, Sparidae, Percomorpha,
-    plus 2 outgroups, a lethrinid and a nemipterid exemplar (Santini et al. \protect\hyperlink{ref-santini2014first}{2014}).
-  \item
-    Updating a phylogeny of Arecaceae, constructing relationships in 6 cldes within
-    the group: subfamilies Calamoideae and Coryphoideae, the tribe Ceroxyleae within
-    subfamily Ceroxyloideae and three groups within subfamily Arecoideae: (1) Iriarteeae,
-  \end{itemize}
-
-  \begin{enumerate}
-  \def\labelenumii{(\arabic{enumii})}
-  \setcounter{enumii}{1}
-  \tightlist
-  \item
-    Cocoseae: Attaleinae except Beccariophoenix and (3) a group containing six
-    tribes; Euterpeae, Leopoldinieae, Pelagodoxeae, Manicarieae, Geonomateae and Areceae
-    (Faurby et al. \protect\hyperlink{ref-faurby2016all}{2016}).
-  \end{enumerate}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A phylogeny of 768 Gesneriaceae species and 58 outgroups for a total species
-    sampling of 826 taxa (Roalson and Roberts \protect\hyperlink{ref-roalson2016distinct}{2016}) some sequence were generated \emph{de novo}.
-  \item
-    A phylogeny of 47 species of scombrid fishes, with 2 outgroups, a gempylid and
-    a trichiurid (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013}).
-  \item
-    to update a dataset underlying a large-scale fern phylogeny (Lehtonen et al. \protect\hyperlink{ref-lehtonen2017environmentally}{2017}),
-    data set in \url{https://zenodo.org/record/345670\#.Xr9QFRPYqqg}, also in TreeBASE,
-    but it is one of those studies that is broken.
-  \item
-    A phylogeny of 13 species of billfishes, order Istiophoriformes: Acanthomorpha,
-    and four outgroups (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013})
-  \item
-    A phylogeny of 765 aphid species, family Aphididae (Hardy et al. \protect\hyperlink{ref-hardy2015evolution}{2015})
-  \item
-    A phylogeny of less than 100 taxa of the family Ranunculaceae (Lehtonen et al. \protect\hyperlink{ref-lehtonen2016sensitive}{2016}),
-    even though they retrieved info from phylota for 194 taxa within the family, they
-    reduced their data set because of low sampling of markers for some taxa.
-  \item
-    A phylogeny of 144 neobatrachian genera, assuming the monophyletic status of
-    genera to increase matrix-filling levels (Frazao et al. \protect\hyperlink{ref-frazao2015gondwana}{2015}).
-  \item
-    A 179 species phylogeny of the bird family Picidae (woodpeckers, piculets,
-    and wrynecks) (Dufort \protect\hyperlink{ref-dufort2015coexistence}{2015}, \protect\hyperlink{ref-dufort2016augmented}{2016}), augmented with data from an updated GenBank
-    release and newly sequenced data.
-  \item
-    A phylogeny of species of freshwater fish endemic to NorthAmerica (Strecker and Olden \protect\hyperlink{ref-strecker2014fish}{2014}),
-    phylota found data for 54 out of 66 spp.
-  \item
-    A phylogeny of 520 species of the order Ericales (Hardy and Cook \protect\hyperlink{ref-hardy2012testing}{2012})
-  \item
-    A phylgeny of 16 fish species of the family Sphyraenidae (Percomorpha), as well
-    as two outgroup species of the Centropomidae (barracudas) (Santini et al. \protect\hyperlink{ref-santini2015first}{2015})
-  \item
-    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas et al. \protect\hyperlink{ref-garcia2016role}{2016})
-  \item
-    Kolmann et al. (\protect\hyperlink{ref-kolmann2017dna}{2017}) uses phylota to download all 1691 co1 sequences belonging to
-    the order Carchariniformes, to place phylogenetically DNA samples obtained from
-    fish markets.
-  \item
-    A phylogeny of 329 bird species in the Tyrannidae (77\% of the species in the
-    family) (Gómez Bahamón and others \protect\hyperlink{ref-gomez2015behavioral}{2015}; Gómez-Bahamón et al. \protect\hyperlink{ref-gomez2020speciation}{2020})
-  \item
-    Retrive 145 sequences registered as \emph{Holothuria} species, but kept 84 as ingroup,
-    plus 4 outgroup sequences from \emph{Stichopus ocellatus}, all belonging to the order
-    Apodida of sea cucumbers (Kamarudin et al. \protect\hyperlink{ref-kamarudin2016phylogenetic}{2016})
-  \item
-    On a master thesis, to get the sequences of the outgroups of Melinidinae, family Poaceae, namely several spp of the
-    subfamily Panicoideae, plus \emph{Gynerium sagittatum}, \emph{Chasmanthium latifolium},
-    and \emph{Zea mays}, (Salariato \protect\hyperlink{ref-salariato2010filogenia}{2010}). Interestingly, phylota was not used
-    in the published study of the thesis (Salariato et al. \protect\hyperlink{ref-salariato2010molecular}{2010}). Ingroup sequences were generated \emph{de novo}.
-  \item
-    On a PhD thesis, to construct a phylogeny of Platyrrhini (internal group),
-    Catarrhini (outgroup), and Tarsiiformes Pereira (\protect\hyperlink{ref-pereira2013padroes}{2013}). Have not found a published study.
-  \item
-    The 10k trees project (Arnold et al. \protect\hyperlink{ref-arnold201010ktrees}{2010}) uses phylota to construct a tree of 301 primate species
-    and the outgroup species \emph{Galeopterus variegates}, a tree of 17 extant odd-toed
-    ungulates species and the outgroup species \emph{Bos taurus}, and a tree of 70 different
-    species of carnivorans and \emph{Equus caballus} as outgroup. However, the do not
-    cite it on the paper, but only on their documentation \url{http://www.academia.edu/download/49690788/10kTrees_Documentation.pdf}.
-  \item
-    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}, also in \protect\hyperlink{ref-freyman2017phylogenetic}{2017}), use phylota to construct
-    a phylogeny (or maybe only mine genbank???) of the Onagraceae and Lythracea,
-    and compare it to the tool they propose, SUMAC.
-  \item
-    Blackmon (\protect\hyperlink{ref-blackmon2017synthesis}{2017}) PhD study applies phylota to reconstruct a 822 mite
-    species tree.
-  \item
-    A study of the effect of poliploidy on niche evolution (Baniaga et al. \protect\hyperlink{ref-baniaga2018polyploid}{2018}),
-    uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
-    16 families and 25 genera,and a tree of 33 genera from 20 different families
-    comprising 1706 taxa.
-  \end{itemize}
-\item
-  When the website was used to identify sequences and markers available in
-  GenBank for a particular group. In this cases, the dataset mining was either performed
-  with other tools, or not performed at all and just used for discussion:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A 812 tips phylogeny of the Order Chiroptera (Shi and Rabosky \protect\hyperlink{ref-shi2015speciation}{2015}) -- dataset
-    constructed with PHLAWD
-  \item
-    A 1276 tips phylogeny of the Fabaceae (Group et al. \protect\hyperlink{ref-legume2013legume}{2013}) -- dataset constructed
-    by hand (I think??)
-  \item
-    A review of dated phylogenies of fire-prone tropical savanna species from Brazil
-    (Simon and Pennington \protect\hyperlink{ref-simon2012cerrado}{2012}) -- just for discussion of the lack of markers available for
-    these species on GenBank
-  \item
-    A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
-    protists (Morrison \protect\hyperlink{ref-morrison2009apicomplexa}{2009}).
-  \item
-    Three data sets from phylota (the suborder Pleurodira of side-necked turtles;
-    the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
-    to demonstrate and exemplify phylogenetic decisiveness (Sanderson et al. \protect\hyperlink{ref-sanderson2010phylogenomics}{2010})
-  \item
-    Mentioned in a PHD thesis (Gagnon and others \protect\hyperlink{ref-gagnon2016systematique}{2016}), but not on the final publication (Gagnon et al. \protect\hyperlink{ref-gagnon2016new}{2016}),
-    phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
-  \end{itemize}
-\item
-  Sometimes, it was cited by mistake:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    In this 630 tip phylogeny of the Caryophyllaceae study (Greenberg and Donoghue \protect\hyperlink{ref-greenberg2011caryophyllaceae}{2011}) it might have been originally
-    cited as an example of large phylogenies that reflect well supported relationships
-    from previous smaller phylogenies. However, it was removed from the text but
-    not from the final list of references. The DNA data set was constructed by hand
-    most probably.
-  \item
-    a study reconstructing the insect tree of life with 49,358 species, 13,865
-    genera, and 760 families within the order Insecta (Chesters \protect\hyperlink{ref-chesters2017construction}{2017}),
-    uses its own algorithm (SOPHI) to mine public DNA databases (Chesters and Zhu \protect\hyperlink{ref-chesters2014protocol}{2014}).
-    It does not cite phylota as it should, but includes it in their references.
-  \end{itemize}
-\item
-  When phylota was used to extract full trees (not only DNA data sets or markers):
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Page (\protect\hyperlink{ref-page2013bionames}{2013}) uses it to generate phylogenies for the \href{http://bionames.org}{bionames website},
-    a ``database linking taxonomic names to their original descriptions, to taxa, and
-    to phylogenies'' generated with phylota.
-  \item
-    Deepak et al. (\protect\hyperlink{ref-deepak2013extracting}{2013}) uses a sample of phylota trees to test their method to
-    remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
-    trees with two or more leaves sharing a label, e.g., a species name, which can
-    imply multiple conflicting phylogenetic relationships for the same set of taxa.
-  \item
-    A review by Sanderson et al. (\protect\hyperlink{ref-sanderson2016perspective}{2016}), takes 134 595 gene trees from phylota
-    GenBank rel. 176 and estimates its degree of resolutin, calculating that less
-    than half of clades are supported with minilam statistical support (0.53 ± 0.32).
-  \end{itemize}
-\end{enumerate}
-
 \hypertarget{acknowledgements}{%
@@
 
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No.~ACI-1429783).
+The University of California, Merced cluster, MERCED (Multi-Environment
+Research Computer for Exploration and Discovery) supported by the
+National Science Foundation (Grant No.~ACI-1429783).
 
 \hypertarget{references}{%
-\section{References}\label{references}}
-
-\newpage
-\begin{center}
-\textsc{References}
-\end{center}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 \leavevmode\hypertarget{ref-altschul1990basic}{}%
-Altschul S.F., Gish W., Miller W., Myers E.W., Lipman D.J. 1990. Basic local alignment search tool. Journal of molecular biology. 215:403--410.
-
-\leavevmode\hypertarget{ref-anacker2011origins}{}%
-Anacker B.L., Whittall J.B., Goldberg E.E., Harrison S.P.
-2011.
-Origins and consequences of serpentine endemism in the california flora.
-Evolution: International Journal of Organic Evolution.
-65:365--376.
+Altschul, S. F., W. Gish, W. Miller, E. W. Myers, and D. J. Lipman,
+1990: Basic local alignment search tool. \emph{Journal of molecular
+biology}, \textbf{215}, 403--410,
+doi:\href{https://doi.org/10.1016/S0022-2836(05)80360-2}{10.1016/S0022-2836(05)80360-2}.
 
 \leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli A., Hettling H., Condamine F.L., Vos K., Nilsson R.H., Sanderson M.J., Sauquet H., Scharn R., Silvestro D., Töpel M., others.
-2017.
-Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
-Systematic Biology.
-66:152--166.
-
-\leavevmode\hypertarget{ref-arnold201010ktrees}{}%
-Arnold C., Matthews L.J., Nunn C.L. 2010. The 10kTrees website: A new online resource for primate phylogeny. Evolutionary Anthropology: Issues, News, and Reviews. 19:114--118.
-
-\leavevmode\hypertarget{ref-baniaga2018polyploid}{}%
-Baniaga A.E., Marx H.E., Arrigo N., Barker M.S. 2018. Polyploid plants have faster rates of multivariate climatic niche evolution than their diploid relatives. BioRxiv.:406314.
-
-\leavevmode\hypertarget{ref-beaulieu2013fruit}{}%
-Beaulieu J.M., Donoghue M.J. 2013. Fruit evolution and diversification in campanulid angiosperms. Evolution. 67:3132--3144.
-
-\leavevmode\hypertarget{ref-beaulieu2012modeling}{}%
-Beaulieu J.M., Jhwueng D.-C., Boettiger C., O'Meara B.C.
-2012a.
-Modeling stabilizing selection: Expanding the ornstein--uhlenbeck model of adaptive evolution.
-Evolution: International Journal of Organic Evolution.
-66:2369--2383.
-
-\leavevmode\hypertarget{ref-beaulieu2012synthesizing}{}%
-Beaulieu J.M., Ree R.H., Cavender-Bares J., Weiblen G.D., Donoghue M.J. 2012b. Synthesizing phylogenetic knowledge for ecological research. Ecology. 93:S4--S13.
+Antonelli, A. and Coauthors, 2017: Toward a self-updating platform for
+estimating rates of speciation and migration, ages, and relationships of
+taxa. \emph{Systematic Biology}, \textbf{66}, 152--166,
+doi:\href{https://doi.org/10.1093/sysbio/syw066}{10.1093/sysbio/syw066}.
 
 \leavevmode\hypertarget{ref-bennett2018phylotar}{}%
-Bennett D.J., Hettling H., Silvestro D., Zizka A., Bacon C.D., Faurby S., Vos R.A., Antonelli A.
-2018.
-PhylotaR: An automated pipeline for retrieving orthologous dna sequences from genbank in r.
-Life.
-8:20.
+Bennett, D. J., H. Hettling, D. Silvestro, A. Zizka, C. D. Bacon, S.
+Faurby, R. A. Vos, and A. Antonelli, 2018: PhylotaR: An automated
+pipeline for retrieving orthologous dna sequences from genbank in r.
+\emph{Life}, \textbf{8}, 20,
+doi:\href{https://doi.org/10.3390/life8020020}{10.3390/life8020020}.
 
 \leavevmode\hypertarget{ref-benson2000genbank}{}%
-Benson D.A., Karsch-Mizrachi I., Lipman D.J., Ostell J., Rapp B.A., Wheeler D.L. 2000. GenBank. Nucleic acids research. 28:15--18.
-
-\leavevmode\hypertarget{ref-blackmon2017synthesis}{}%
-Blackmon H. 2017. Synthesis and phylogenetic comparative analyses of the causes and consequences of karyotype evolution in arthropods..
-
-\leavevmode\hypertarget{ref-bruneau2019towards}{}%
-Bruneau A., Borges L.M., Allkin R., Egan A.N., De La Estrella M., Javadi F., Klitgaard B., Miller J.T., Murphy D.J., Sinou C., others.
-2019.
-Towards a new online species-information system for legumes.
-Australian Systematic Botany.
-32:495--518.
+Benson, D. A., I. Karsch-Mizrachi, D. J. Lipman, J. Ostell, B. A. Rapp,
+and D. L. Wheeler, 2000: GenBank. \emph{Nucleic acids research},
+\textbf{28}, 15--18,
+doi:\href{https://doi.org/10.1093/nar/28.1.15}{10.1093/nar/28.1.15}.
 
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
-Camacho C., George C., Vahram A., Ning M., Jason P., Kevin B., Thomas L. 2009. BLAST+: Architecture and applications. BMC bioinformatics. 10:421.
+Camacho, C., C. George, A. Vahram, M. Ning, P. Jason, B. Kevin, and L.
+Thomas, 2009: BLAST+: Architecture and applications. \emph{BMC
+bioinformatics}, \textbf{10}, 421,
+doi:\href{https://doi.org/10.1186/1471-2105-10-421}{10.1186/1471-2105-10-421}.
 
 \leavevmode\hypertarget{ref-ceron2019phylotol}{}%
-Cerón-Romero M.A., Maurer-Alcalá X.X., Grattepanche J.-D., Yan Y., Fonseca M.M., Katz L.
-2019.
-PhyloToL: A taxon/gene-rich phylogenomic pipeline to explore genome evolution of diverse eukaryotes.
-Molecular biology and evolution.
-36:1831--1842.
+Cerón-Romero, M. A., X. X. Maurer-Alcalá, J.-D. Grattepanche, Y. Yan, M.
+M. Fonseca, and L. Katz, 2019: PhyloToL: A taxon/gene-rich phylogenomic
+pipeline to explore genome evolution of diverse eukaryotes.
+\emph{Molecular biology and evolution}, \textbf{36}, 1831--1842,
+doi:\href{https://doi.org/10.1093/molbev/msz103}{10.1093/molbev/msz103}.
 
 \leavevmode\hypertarget{ref-chen2008phylofinder}{}%
-Chen D., Burleigh J.G., Bansal M.S., Fernández-Baca D. 2008. PhyloFinder: An intelligent search engine for phylogenetic tree databases. BMC Evolutionary Biology. 8:90.
-
-\leavevmode\hypertarget{ref-chen2016tree}{}%
-Chen Z.-D., Yang T., Lin L., Lu L.-M., Li H.-L., Sun M., Liu B., Chen M., Niu Y.-T., Ye J.-F., others.
-2016.
-Tree of life for the genera of chinese vascular plants.
-Journal of Systematics and Evolution.
-54:277--306.
+Chen, D., J. G. Burleigh, M. S. Bansal, and D. Fernández-Baca, 2008:
+PhyloFinder: An intelligent search engine for phylogenetic tree
+databases. \emph{BMC Evolutionary Biology}, \textbf{8}, 90,
+doi:\href{https://doi.org/10.1186/1471-2148-8-90.}{10.1186/1471-2148-8-90.}
 
 \leavevmode\hypertarget{ref-chesters2017construction}{}%
-Chesters D. 2017. Construction of a species-level tree of life for the insects and utility in taxonomic profiling. Systematic biology. 66:426--439.
-
-\leavevmode\hypertarget{ref-chesters2019phylogeny}{}%
-Chesters D. 2019. The phylogeny of insects in the data-driven era. Systematic Entomology.
-
-\leavevmode\hypertarget{ref-chesters2013resolving}{}%
-Chesters D., Vogler A.P. 2013. Resolving ambiguity of species limits and concatenation in multilocus sequence data for the construction of phylogenetic supermatrices. Systematic Biology. 62:456--466.
+Chesters, D., 2017: Construction of a species-level tree of life for the
+insects and utility in taxonomic profiling. \emph{Systematic biology},
+\textbf{66}, 426--439,
+doi:\href{https://doi.org/10.1093/sysbio/syw099}{10.1093/sysbio/syw099}.
 
 \leavevmode\hypertarget{ref-chesters2014protocol}{}%
-Chesters D., Zhu C.-D. 2014. A protocol for species delineation of public dna databases, applied to the insecta. Systematic biology. 63:712--725.
+------, and C.-D. Zhu, 2014: A protocol for species delineation of
+public dna databases, applied to the insecta. \emph{Systematic biology},
+\textbf{63}, 712--725,
+doi:\href{https://doi.org/10.1093/sysbio/syu038}{10.1093/sysbio/syu038}.
 
 \leavevmode\hypertarget{ref-crawford2012more}{}%
-Crawford N.G., Faircloth B.C., McCormack J.E., Brumfield R.T., Winker K., Glenn T.C.
-2012.
-More than 1000 ultraconserved elements provide evidence that turtles are the sister group of archosaurs.
-Biology letters.
-8:783--786.
-
-\leavevmode\hypertarget{ref-crete2012salmonidae}{}%
-Crête-Lafrenière A., Weir L.K., Bernatchez L. 2012. Framing the salmonidae family phylogenetic portrait: A more complete picture from increased taxon sampling. PloS one. 7.
-
-\leavevmode\hypertarget{ref-day1984efficient}{}%
-Day W.H., Edelsbrunner H. 1984. Efficient algorithms for agglomerative hierarchical clustering methods. Journal of classification. 1:7--24.
-
-\leavevmode\hypertarget{ref-deepak2010searchtree}{}%
-Deepak A. 2010. SearchTree: Mining robust phylogenetic trees..
-
-\leavevmode\hypertarget{ref-deepak2013managing}{}%
-Deepak A. 2013. Managing and analyzing phylogenetic databases..
-
-\leavevmode\hypertarget{ref-deepak2013extracting}{}%
-Deepak A., Fernández-Baca D., McMahon M.M. 2013. Extracting conflict-free information from multi-labeled trees. Algorithms for Molecular Biology. 8:18.
+Crawford, N. G., B. C. Faircloth, J. E. McCormack, R. T. Brumfield, K.
+Winker, and T. C. Glenn, 2012: More than 1000 ultraconserved elements
+provide evidence that turtles are the sister group of archosaurs.
+\emph{Biology letters}, \textbf{8}, 783--786,
+doi:\href{https://doi.org/10.1098/rsbl.2012.0331}{10.1098/rsbl.2012.0331}.
 
 \leavevmode\hypertarget{ref-deepak2014evominer}{}%
-Deepak A., Fernández-Baca D., Tirthapura S., Sanderson M.J., McMahon M.M. 2014. EvoMiner: Frequent subtree mining in phylogenetic databases. Knowledge and Information Systems. 41:559--590.
+Deepak, A., D. Fernández-Baca, S. Tirthapura, M. J. Sanderson, and M. M.
+McMahon, 2014: EvoMiner: Frequent subtree mining in phylogenetic
+databases. \emph{Knowledge and Information Systems}, \textbf{41},
+559--590,
+doi:\href{https://doi.org/10.1007/s10115-013-0676-0}{10.1007/s10115-013-0676-0}.
 
 \leavevmode\hypertarget{ref-drori2018onetwotree}{}%
-Drori M., Rice A., Einhorn M., Chay O., Glick L., Mayrose I. 2018. OneTwoTree: An online tool for phylogeny reconstruction. Molecular ecology resources. 18:1492--1499.
-
-\leavevmode\hypertarget{ref-dufort2015coexistence}{}%
-Dufort M. 2015. Coexistence, ecomorphology, and diversification in the avian family picidae (woodpeckers and allies)..
-
-\leavevmode\hypertarget{ref-dufort2016augmented}{}%
-Dufort M.J. 2016. An augmented supermatrix phylogeny of the avian family picidae reveals uncertainty deep in the family tree. Molecular phylogenetics and evolution. 94:313--326.
+Drori, M., A. Rice, M. Einhorn, O. Chay, L. Glick, and I. Mayrose, 2018:
+OneTwoTree: An online tool for phylogeny reconstruction. \emph{Molecular
+ecology resources}, \textbf{18}, 1492--1499,
+doi:\href{https://doi.org/10.1111/1755-0998.12927}{10.1111/1755-0998.12927}.
 
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
-Edgar R.C. 2004. MUSCLE: Multiple sequence alignment with high accuracy and high throughput. Nucleic acids research. 32:1792--1797.
-
-\leavevmode\hypertarget{ref-evans2019bony}{}%
-Evans K.M., Vidal-García M., Tagliacollo V.A., Taylor S.J., Fenolio D.B.
-2019.
-Bony patchwork: Mosaic patterns of evolution in the skull of electric fishes (apteronotidae: Gymnotiformes).
-Integrative and comparative biology.
-59:420--431.
-
-\leavevmode\hypertarget{ref-fan2015assembly}{}%
-Fan H., Ives A.R., Surget-Groba Y., Cannon C.H. 2015. An assembly and alignment-free method of phylogeny reconstruction from next-generation sequencing data. BMC genomics. 16:522.
+Edgar, R. C., 2004: MUSCLE: Multiple sequence alignment with high
+accuracy and high throughput. \emph{Nucleic acids research},
+\textbf{32}, 1792--1797,
+doi:\href{https://doi.org/10.1093/nar/gkh340}{10.1093/nar/gkh340}.
 
 \leavevmode\hypertarget{ref-fang2019physpetree}{}%
-Fang Y., Liu C., Lin J., Li X., Alavian K.N., Yang Y., Niu Y. 2019. PhySpeTree: An automated pipeline for reconstructing phylogenetic species trees. BMC evolutionary biology. 19:1--8.
-
-\leavevmode\hypertarget{ref-faurby2016all}{}%
-Faurby S., Eiserhardt W.L., Baker W.J., Svenning J.-C. 2016. An all-evidence species-level supertree for the palms (arecaceae). Molecular Phylogenetics and Evolution. 100:57--69.
-
-\leavevmode\hypertarget{ref-faurby2015species}{}%
-Faurby S., Svenning J.-C.
-2015.
-A species-level phylogeny of all extant and late quaternary extinct mammals using a novel heuristic-hierarchical bayesian approach.
-Molecular phylogenetics and evolution.
-84:14--26.
+Fang, Y., C. Liu, J. Lin, X. Li, K. N. Alavian, Y. Yang, and Y. Niu,
+2019: PhySpeTree: An automated pipeline for reconstructing phylogenetic
+species trees. \emph{BMC evolutionary biology}, \textbf{19}, 1--8,
+doi:\href{https://doi.org/10.1186/s12862-019-1541-x}{10.1186/s12862-019-1541-x}.
 
 \leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
-Felsenstein J. 1985. Confidence intervals on phylogenetics: An approach using bootstrap. Evolution. 39:783--791.
-
-\leavevmode\hypertarget{ref-flores2011estimating}{}%
-Flores O., Coomes D.A. 2011. Estimating the wood density of species for carbon stock assessments. Methods in Ecology and Evolution. 2:214--220.
-
-\leavevmode\hypertarget{ref-frazao2015gondwana}{}%
-Frazao A., Silva H.R.
-da, Moraes Russo C.A.
-de.
-2015.
-The gondwana breakup and the history of the atlantic and indian oceans unveils two new clades for early neobatrachian diversification.
-PloS one.
-10.
+Felsenstein, J., 1985: Confidence intervals on phylogenetics: An
+approach using bootstrap. \emph{Evolution}, \textbf{39}, 783--791.
 
 \leavevmode\hypertarget{ref-freyman2015sumac}{}%
-Freyman W.A. 2015. SUMAC: Constructing phylogenetic supermatrices and assessing partially decisive taxon coverage. Evolutionary Bioinformatics. 11:EBO--S35384.
-
-\leavevmode\hypertarget{ref-freyman2017phylogenetic}{}%
-Freyman W.A. 2017. Phylogenetic models linking speciation and extinction to chromosome and mating system evolution..
-
-\leavevmode\hypertarget{ref-gagnon2016new}{}%
-Gagnon E., Bruneau A., Hughes C.E., Queiroz L.P. de, Lewis G.P. 2016. A new generic system for the pantropical caesalpinia group (leguminosae). PhytoKeys.:1.
-
-\leavevmode\hypertarget{ref-gagnon2016systematique}{}%
-Gagnon E., others. 2016. Systématique et biogéographie du groupe caesalpinia (famille leguminosae)..
-
-\leavevmode\hypertarget{ref-gao2011solution}{}%
-Gao Y., Meng Z., He X., Liu Y., Zhou Y., Li J. 2011. A solution to integrate data for phylogenetic research. 2011 5th international conference on bioinformatics and biomedical engineering.:1--4.
-
-\leavevmode\hypertarget{ref-garcia2016role}{}%
-García-Navas V., Bonnet T., Bonal R., Postma E.
-2016.
-The role of fecundity and sexual selection in the evolution of size and sexual size dimorphism in new world and old world voles (rodentia: Arvicolinae).
-Oikos.
-125:1250--1260.
-
-\leavevmode\hypertarget{ref-gomez2020speciation}{}%
-Gómez-Bahamón V., Márquez R., Jahn A.E., Miyaki C.Y., Tuero D.T., Laverde-R O., Restrepo S., Cadena C.D.
-2020.
-Speciation associated with shifts in migratory behavior in an avian radiation.
-Current Biology.
-
-\leavevmode\hypertarget{ref-gomez2015behavioral}{}%
-Gómez Bahamón V., others. 2015. A behavioral polymorphism as an intermediate stage in the evolution of divergent forms-partial migration in new world flycatchers (aves, tyrannidae)..
+Freyman, W. A., 2015: SUMAC: Constructing phylogenetic supermatrices and
+assessing partially decisive taxon coverage. \emph{Evolutionary
+Bioinformatics}, \textbf{11}, EBO--S35384,
+doi:\href{https://doi.org/10.4137/EBO.S35384}{10.4137/EBO.S35384}.
 
 \leavevmode\hypertarget{ref-grant2014building}{}%
-Grant J.R., Katz L.A. 2014. Building a phylogenomic pipeline for the eukaryotic tree of life-addressing deep phylogenies with genome-scale data. PLoS currents. 6.
-
-\leavevmode\hypertarget{ref-greenberg2011caryophyllaceae}{}%
-Greenberg A.K., Donoghue M.J. 2011. Molecular systematics and character evolution in caryophyllaceae. Taxon. 60:1637--1652.
-
-\leavevmode\hypertarget{ref-legume2013legume}{}%
-Group L.P.W., Bruneau A., Doyle J.J., Herendeen P., Hughes C., Kenicer G., Lewis G., Mackinder B., Pennington R.T., Sanderson M.J., others.
-2013.
-Legume phylogeny and classification in the 21st century: Progress, prospects and lessons for other species--rich clades.
-Taxon.
-62:217--248.
+Grant, J. R., and L. A. Katz, 2014: Building a phylogenomic pipeline for
+the eukaryotic tree of life-addressing deep phylogenies with
+genome-scale data. \emph{PLoS currents}, \textbf{6},
+doi:\href{https://doi.org/10.1371/currents.tol.c24b6054aebf3602748ac042ccc8f2e9}{10.1371/currents.tol.c24b6054aebf3602748ac042ccc8f2e9}.
 
 \leavevmode\hypertarget{ref-guy2017phyloskeleton}{}%
-Guy L. 2017. PhyloSkeleton: Taxon selection, data retrieval and marker identification for phylogenomics. Bioinformatics. 33:1230--1232.
-
-\leavevmode\hypertarget{ref-hardy2010gall}{}%
-Hardy N.B., Cook L.G. 2010. Gall-induction in insects: Evolutionary dead-end or speciation driver? BMC evolutionary biology. 10:257.
-
-\leavevmode\hypertarget{ref-hardy2012testing}{}%
-Hardy N.B., Cook L.G. 2012. Testing for ecological limitation of diversification: A case study using parasitic plants. The American Naturalist. 180:438--449.
-
-\leavevmode\hypertarget{ref-hardy2014specialization}{}%
-Hardy N.B., Otto S.P.
-2014.
-Specialization and generalization in the diversification of phytophagous insects: Tests of the musical chairs and oscillation hypotheses.
-Proceedings of the Royal Society B: Biological Sciences.
-281:20132960.
-
-\leavevmode\hypertarget{ref-hardy2015evolution}{}%
-Hardy N.B., Peterson D.A., Dohlen C.D. von. 2015. The evolution of life cycle complexity in aphids: Ecological optimization or historical constraint? Evolution. 69:1423--1432.
+Guy, L., 2017: PhyloSkeleton: Taxon selection, data retrieval and marker
+identification for phylogenomics. \emph{Bioinformatics}, \textbf{33},
+1230--1232,
+doi:\href{https://doi.org/10.1093/bioinformatics/btw824}{10.1093/bioinformatics/btw824}.
 
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
-Helmus M.R., Ives A.R. 2012. Phylogenetic diversity--area curves. Ecology. 93:S31--S43.
-
-\leavevmode\hypertarget{ref-helmus2010communities}{}%
-Helmus M.R., Keller W., Paterson M.J., Yan N.D., Cannon C.H., Rusak J.A. 2010. Communities contain closely related species during ecosystem disturbance. Ecology letters. 13:162--174.
-
-\leavevmode\hypertarget{ref-ickert2009fossil}{}%
-Ickert-Bond S.M., Rydin C., Renner S.S.
-2009.
-A fossil-calibrated relaxed clock for ephedra indicates an oligocene age for the divergence of asian and new world clades and miocene dispersal into south america.
-Journal of Systematics and Evolution.
-47:444--456.
+Helmus, M. R., and A. R. Ives, 2012: Phylogenetic diversity--area
+curves. \emph{Ecology}, \textbf{93}, S31--S43,
+doi:\href{https://doi.org/10.1890/11-0435.1}{10.1890/11-0435.1}.
 
 \leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco F., Cazes J., Smith S.A., Stamatakis A. 2014. PUmPER: Phylogenies updated perpetually. Bioinformatics. 30:1476--1477.
-
-\leavevmode\hypertarget{ref-jamil2016visual}{}%
-Jamil H.M. 2016. A visual interface for querying heterogeneous phylogenetic databases. IEEE/ACM transactions on computational biology and bioinformatics. 14:131--144.
-
-\leavevmode\hypertarget{ref-jones2014trends}{}%
-Jones T.M., Baxter D.G., Hagedorn G., Legler B., Gilbert E., Thiele K., Vargas-Rodriguez Y., Urbatsch L.E.
-2014.
-Trends in access of plant biodiversity data revealed by google analytics.
-Biodiversity data journal.
-
-\leavevmode\hypertarget{ref-kamarudin2016phylogenetic}{}%
-Kamarudin K.R., Rehan A.M., Hashim R., Usup G., Rehan M.M.
-2016.
-Phylogenetic relationships within the genus holothuria inferred from 16S mitochondiral rRNA gene sequences.
-Sains Malaysiana.
-45:1079--1087.
-
-\leavevmode\hypertarget{ref-kolmann2017dna}{}%
-Kolmann M.A., Elbassiouny A.A., Liverpool E.A., Lovejoy N.R. 2017. DNA barcoding reveals the diversity of sharks in guyana coastal markets. Neotropical Ichthyology. 15.
+Izquierdo-Carrasco, F., J. Cazes, S. A. Smith, and A. Stamatakis, 2014:
+PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics},
+\textbf{30}, 1476--1477,
+doi:\href{https://doi.org/10.1093/bioinformatics/btu053}{10.1093/bioinformatics/btu053}.
 
 \leavevmode\hypertarget{ref-kumar2015bir}{}%
-Kumar S., Krabberød A.K., Neumann R.S., Michalickova K., Zhao S., Zhang X., Shalchian-Tabrizi K. 2015. BIR pipeline for preparation of phylogenomic data. Evolutionary Bioinformatics. 11:EBO--S10189.
-
-\leavevmode\hypertarget{ref-lehtonen2011towards}{}%
-Lehtonen S. 2011. Towards resolving the complete fern tree of life. PLoS One. 6.
-
-\leavevmode\hypertarget{ref-lehtonen2016sensitive}{}%
-Lehtonen S., Christenhusz M.J., Falck D. 2016. Sensitive phylogenetics of clematis and its position in ranunculaceae. Botanical Journal of the Linnean Society. 182:825--867.
-
-\leavevmode\hypertarget{ref-lehtonen2017environmentally}{}%
-Lehtonen S., Silvestro D., Karger D.N., Scotese C., Tuomisto H., Kessler M., Peña C., Wahlberg N., Antonelli A.
-2017.
-Environmentally driven extinction and opportunistic origination explain fern diversification patterns.
-Scientific Reports.
-7:1--12.
+Kumar, S., A. K. Krabberød, R. S. Neumann, K. Michalickova, S. Zhao, X.
+Zhang, and K. Shalchian-Tabrizi, 2015: BIR pipeline for preparation of
+phylogenomic data. \emph{Evolutionary Bioinformatics}, \textbf{11},
+EBO--S10189,
+doi:\href{https://doi.org/10.4137/EBO.S10189}{10.4137/EBO.S10189}.
 
 \leavevmode\hypertarget{ref-lemoine2018renewing}{}%
-Lemoine F., Entfellner J.-B.D., Wilkinson E., Correia D., Felipe M.D., De Oliveira T., Gascuel O. 2018. Renewing felsenstein's phylogenetic bootstrap in the era of big data. Nature. 556:452--456.
-
-\leavevmode\hypertarget{ref-li2013partfasttree}{}%
-Li J., Meng Z., Hou Y., Zhou Y., Gao Y.
-2013.
-PartFastTree: Constructing large phylogenetic trees and estimating their reliability.
-2013 ninth international conference on natural computation (icnc).:1052--1056.
-
-\leavevmode\hypertarget{ref-mahmood2015avian}{}%
-Mahmood M.T. 2015. Avian raptor evolution..
-
-\leavevmode\hypertarget{ref-martin2011trophic}{}%
-Martin C.H., Wainwright P.C.
-2011.
-Trophic novelty is linked to exceptional rates of morphological diversification in two adaptive radiations of cyprinodon pupfish.
-Evolution: International Journal of Organic Evolution.
-65:2197--2212.
+Lemoine, F., J.-B. D. Entfellner, E. Wilkinson, D. Correia, M. D.
+Felipe, T. De Oliveira, and O. Gascuel, 2018: Renewing felsenstein's
+phylogenetic bootstrap in the era of big data. \emph{Nature},
+\textbf{556}, 452--456.
 
 \leavevmode\hypertarget{ref-mcmahon2015stbase}{}%
-McMahon M.M., Deepak A., Fernández-Baca D., Boss D., Sanderson M.J. 2015. STBase: One million species trees for comparative biology. PloS one. 10.
+McMahon, M. M., A. Deepak, D. Fernández-Baca, D. Boss, and M. J.
+Sanderson, 2015: STBase: One million species trees for comparative
+biology. \emph{PloS one}, \textbf{10},
+doi:\href{https://doi.org/10.1371/journal.pone.0117987}{10.1371/journal.pone.0117987}.
 
 \leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
-McTavish E.J., Hinchliff C.E., Allman J.F., Brown J.W., Cranston K.A., Holder M.T., Rees J.A., Smith S.A.
-2015.
-Phylesystem: A git-based data store for community-curated phylogenetic estimates.
-Bioinformatics.
-31:2794--2800.
+McTavish, E. J., C. E. Hinchliff, J. F. Allman, J. W. Brown, K. A.
+Cranston, M. T. Holder, J. A. Rees, and S. A. Smith, 2015: Phylesystem:
+A git-based data store for community-curated phylogenetic estimates.
+\emph{Bioinformatics}, \textbf{31}, 2794--2800,
+doi:\href{https://doi.org/10.1093/bioinformatics/btv276}{10.1093/bioinformatics/btv276}.
 
 \leavevmode\hypertarget{ref-meng2015darwintree}{}%
-Meng Z., Dong H., Li J., Chen Z., Zhou Y., Wang X., Zhang S. 2015a. Darwintree: A molecular data analysis and application environment for phylogenetic study. Data Science Journal. 14.
-
-\leavevmode\hypertarget{ref-meng2015solution}{}%
-Meng Z., Li J., Chen Z. 2015b. A solution to phylogeny assembly for ecologists. 2015 12th international conference on fuzzy systems and knowledge discovery (fskd).:1103--1107.
-
-\leavevmode\hypertarget{ref-meng2015sotree}{}%
-Meng Z., Li J., Yang T., Lin L., Chen Z.
-2015c.
-SoTree: An automated phylogeny assembly tool for ecologists from big tree.
-2015 ieee international conference on smart city/socialcom/sustaincom (smartcity).:792--797.
-
-\leavevmode\hypertarget{ref-meng2012gsqct}{}%
-Meng Z., Li J., Zhou Y., Cao W., Xiao X., Zhao J., Dong H., Zhang S.
-2012a.
-GSQCT: A solution to screening gene sequences for phylogenetics analysis.
-2012 9th international conference on fuzzy systems and knowledge discovery.:2929--2933.
-
-\leavevmode\hypertarget{ref-meng2014rapidtree}{}%
-Meng Z., Shao J., Cao W., Li J., Zhou Y., Wang X.
-2014.
-RapidTree: A solution to rapid reconstruction phylogenetic tree.
-2014 11th international conference on fuzzy systems and knowledge discovery (fskd).:513--517.
-
-\leavevmode\hypertarget{ref-meng2012cloud}{}%
-Meng Z., Xiao X., Li J., Zhou Y., Cao W., Shen G.
-2012b.
-Cloud-gsqct: A parallel approach to screen gene sequences for phylogenetics analysis.
-2012 international conference on computer science and information processing (csip).:660--663.
-
-\leavevmode\hypertarget{ref-morrison2009apicomplexa}{}%
-Morrison D.A. 2009. Evolution of the apicomplexa: Where are we now? Trends in parasitology. 25:375--382.
-
-\leavevmode\hypertarget{ref-nazaire2012broad}{}%
-Nazaire M., Hufford L. 2012. A broad phylogenetic analysis of boraginaceae: Implications for the relationships of mertensia. Systematic Botany. 37:758--783.
-
-\leavevmode\hypertarget{ref-page2011linking}{}%
-Page R.D. 2011. Linking ncbi to wikipedia: A wiki-based approach. PLoS currents. 3.
-
-\leavevmode\hypertarget{ref-page2013bionames}{}%
-Page R.D. 2013. BioNames: Linking taxonomy, texts, and trees. PeerJ. 1:e190.
+Meng, Z., H. Dong, J. Li, Z. Chen, Y. Zhou, X. Wang, and S. Zhang, 2015:
+Darwintree: A molecular data analysis and application environment for
+phylogenetic study. \emph{Data Science Journal}, \textbf{14},
+doi:\href{https://doi.org/10.5334/dsj-2015-010}{10.5334/dsj-2015-010}.
 
 \leavevmode\hypertarget{ref-papadopoulou2015automated}{}%
-Papadopoulou A., Chesters D., Coronado I., De la Cadena G., Cardoso A., Reyes J.C., Maes J.-M., Rueda R.M., Gómez-Zurita J.
-2015.
-Automated dna-based plant identification for large-scale biodiversity assessment.
-Molecular ecology resources.
-15:136--152.
-
-\leavevmode\hypertarget{ref-pereira2013padroes}{}%
-Pereira J.E.S. 2013. Padrões e processos na evolução de primatas neotropicais (platyrrhini, primates)..
+Papadopoulou, A. and Coauthors, 2015: Automated dna-based plant
+identification for large-scale biodiversity assessment. \emph{Molecular
+ecology resources}, \textbf{15}, 136--152,
+doi:\href{https://doi.org/10.1111/1755-0998.12256}{10.1111/1755-0998.12256}.
 
 \leavevmode\hypertarget{ref-peters2011taming}{}%
-Peters R.S., Meyer B., Krogmann L., Borner J., Meusemann K., Schütte K., Niehuis O., Misof B.
-2011.
-The taming of an impossible child: A standardized all-in approach to the phylogeny of hymenoptera using public database sequences.
-BMC biology.
-9:55.
+Peters, R. S., B. Meyer, L. Krogmann, J. Borner, K. Meusemann, K.
+Schütte, O. Niehuis, and B. Misof, 2011: The taming of an impossible
+child: A standardized all-in approach to the phylogeny of hymenoptera
+using public database sequences. \emph{BMC biology}, \textbf{9}, 55,
+doi:\href{https://doi.org/10.1186/1741-7007-9-55}{10.1186/1741-7007-9-55}.
+\url{https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55\#Sec21}.
 
 \leavevmode\hypertarget{ref-piel2009treebase}{}%
-Piel W., Chan L., Dominus M., Ruan J., Vos R., Tannen V. 2009. Treebase v. 2: A database of phylogenetic knowledge. E-biosphere..
-
-\leavevmode\hypertarget{ref-rabosky2011species}{}%
-Rabosky D.L., Cowan M.A., Talaba A.L., Lovette I.J.
-2011.
-Species interactions mediate phylogenetic community structure in a hyperdiverse lizard assemblage from arid australia.
-The American Naturalist.
-178:579--595.
+Piel, W., L. Chan, M. Dominus, J. Ruan, R. Vos, and V. Tannen, 2009:
+Treebase v. 2: A database of phylogenetic knowledge. E-biosphere.
 
 \leavevmode\hypertarget{ref-ranwez2009phyloexplorer}{}%
-Ranwez V., Clairon N., Delsuc F., Pourali S., Auberval N., Diser S., Berry V. 2009. PhyloExplorer: A web server to validate, explore and query phylogenetic trees. BMC evolutionary biology. 9:108.
-
-\leavevmode\hypertarget{ref-roalson2016distinct}{}%
-Roalson E.H., Roberts W.R. 2016. Distinct processes drive diversification in different clades of gesneriaceae. Systematic Biology. 65:662--684.
-
-\leavevmode\hypertarget{ref-roquet2013building}{}%
-Roquet C., Thuiller W., Lavergne S. 2013. Building megaphylogenies for macroecology: Taking up the challenge. Ecography. 36:13--26.
-
-\leavevmode\hypertarget{ref-ross2013large}{}%
-Ross L., Hardy N.B., Okusu A., Normark B.B. 2013. Large population size predicts the distribution of asexuality in scale insects. Evolution: International Journal of Organic Evolution. 67:196--206.
+Ranwez, V., N. Clairon, F. Delsuc, S. Pourali, N. Auberval, S. Diser,
+and V. Berry, 2009: PhyloExplorer: A web server to validate, explore and
+query phylogenetic trees. \emph{BMC evolutionary biology}, \textbf{9},
+108,
+doi:\href{https://doi.org/10.1186/1471-2148-9-108}{10.1186/1471-2148-9-108}.
 
 \leavevmode\hypertarget{ref-ruiz2019datataxa}{}%
-Ruiz-Sanchez E., Maya-Lastra C.A., Steinmann V.W., Zamudio S., Carranza E., Murillo R.M., Rzedowski J.
-2019.
-Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajío as a case study.
-Botanical Sciences.
-97:754--760.
-
-\leavevmode\hypertarget{ref-salariato2010filogenia}{}%
-Salariato D.L. 2010. Filogenia y evolución de la subtribu melinidinae (paniceae: Panicoideae: Poaceae)..
-
-\leavevmode\hypertarget{ref-salariato2010molecular}{}%
-Salariato D.L., Zuloaga F.O., Giussani L.M., Morrone O.
-2010.
-Molecular phylogeny of the subtribe melinidinae (poaceae: Panicoideae: Paniceae) and evolutionary trends in the homogenization of inflorescences.
-Molecular Phylogenetics and Evolution.
-56:355--369.
+Ruiz-Sanchez, E., C. A. Maya-Lastra, V. W. Steinmann, S. Zamudio, E.
+Carranza, R. M. Murillo, and J. Rzedowski, 2019: Datataxa: A new script
+to extract metadata sequence information from genbank, the flora of
+bajı́o as a case study. \emph{Botanical Sciences}, \textbf{97}, 754--760,
+doi:\href{https://doi.org/10.17129/botsci.2226}{10.17129/botsci.2226}.
 
 \leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson M.J., Boss D., Chen D., Cranston K.A., Wehe A. 2008. The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. Systematic Biology. 57:335--346.
-
-\leavevmode\hypertarget{ref-sanderson2010phylogenomics}{}%
-Sanderson M.J., McMahon M.M., Steel M. 2010. Phylogenomics with incomplete taxon coverage: The limits to inference. BMC Evolutionary Biology. 10:155.
-
-\leavevmode\hypertarget{ref-sanderson2016perspective}{}%
-Sanderson M.J., Olson P., Hughes J., Cotton J. 2016. Perspective: Challenges in assembling the `next generation'Tree of life. Olson PD, Hughes J and Cotton JA.:13--27.
+Sanderson, M. J., D. Boss, D. Chen, K. A. Cranston, and A. Wehe, 2008:
+The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics
+Research. \emph{Systematic Biology}, \textbf{57}, 335--346,
+doi:\href{https://doi.org/10.1080/10635150802158688}{10.1080/10635150802158688}.
+\url{https://doi.org/10.1080/10635150802158688}.
 
 \leavevmode\hypertarget{ref-san2010molecular}{}%
-San Mauro D., Agorreta A. 2010. Molecular systematics: A synthesis of the common methods and the state of knowledge. Cellular \& Molecular Biology Letters. 15:311.
-
-\leavevmode\hypertarget{ref-santini2014first}{}%
-Santini F., Carnevale G., Sorenson L. 2014. First multi-locus timetree of seabreams and porgies (percomorpha: Sparidae). Italian Journal of Zoology. 81:55--71.
-
-\leavevmode\hypertarget{ref-santini2015first}{}%
-Santini F., Carnevale G., Sorenson L.
-2015.
-First timetree of sphyraenidae (percomorpha) reveals a middle eocene crown age and an oligo--miocene radiation of barracudas.
-Italian Journal of Zoology.
-82:133--142.
-
-\leavevmode\hypertarget{ref-santini2013first}{}%
-Santini F., Sorenson L. 2013. First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies. Italian journal of zoology. 80:481--489.
-
-\leavevmode\hypertarget{ref-sarkinen2013solanaceae}{}%
-Särkinen T., Bohs L., Olmstead R.G., Knapp S. 2013. A phylogenetic framework for evolutionary study of the nightshades (solanaceae): A dated 1000-tip tree. BMC evolutionary biology. 13:214.
+San Mauro, D., and A. Agorreta, 2010: Molecular systematics: A synthesis
+of the common methods and the state of knowledge. \emph{Cellular \&
+Molecular Biology Letters}, \textbf{15}, 311,
+doi:\href{https://doi.org/10.2478/s11658-010-0010-8}{10.2478/s11658-010-0010-8}.
 
 \leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
-Schoch C.L., Sung G.-H., López-Giráldez F., Townsend J.P., Miadlikowska J., Hofstetter V., Robbertse B., Matheny P.B., Kauff F., Wang Z., others.
-2009.
-The ascomycota tree of life: A phylum-wide phylogeny clarifies the origin and evolution of fundamental reproductive and ecological traits.
-Systematic biology.
-58:224--239.
-
-\leavevmode\hypertarget{ref-selvatti2015paleogene}{}%
-Selvatti A.P., Gonzaga L.P., Moraes Russo C.A.
-de.
-2015.
-A paleogene origin for crown passerines and the diversification of the oscines in the new world.
-Molecular phylogenetics and evolution.
-88:1--15.
-
-\leavevmode\hypertarget{ref-shi2015speciation}{}%
-Shi J.J., Rabosky D.L. 2015. Speciation dynamics during the global radiation of extant bats. Evolution. 69:1528--1545.
-
-\leavevmode\hypertarget{ref-simon2012cerrado}{}%
-Simon M.F., Pennington T. 2012. Evidence for adaptation to fire regimes in the tropical savannas of the brazilian cerrado. International Journal of Plant Sciences. 173:711--723.
+Schoch, C. L. and Coauthors, 2009: The ascomycota tree of life: A
+phylum-wide phylogeny clarifies the origin and evolution of fundamental
+reproductive and ecological traits. \emph{Systematic biology},
+\textbf{58}, 224--239.
+
+\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
+Smith, S. A., and J. F. Walker, 2019: PyPHLAWD: A python tool for
+phylogenetic dataset construction. \emph{Methods in Ecology and
+Evolution}, \textbf{10}, 104--108.
 
 \leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith S.A., Beaulieu J.M., Donoghue M.J. 2009. Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches. BMC evolutionary biology. 9:37.
-
-\leavevmode\hypertarget{ref-smith2018constructing}{}%
-Smith S.A., Brown J.W. 2018. Constructing a broadly inclusive seed plant phylogeny. American Journal of Botany. 105:302--314.
-
-\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
-Smith S.A., Walker J.F. 2019. PyPHLAWD: A python tool for phylogenetic dataset construction. Methods in Ecology and Evolution. 10:104--108.
-
-\leavevmode\hypertarget{ref-sorenson2014evolution}{}%
-Sorenson L. 2014. Evolution of marine fish biodiversity: Phylogenomics and ecological processes shaping diversification..
-
-\leavevmode\hypertarget{ref-sorenson2014effect}{}%
-Sorenson L., Santini F., Alfaro M. 2014. The effect of habitat on modern shark diversification. Journal of Evolutionary Biology. 27:1536--1548.
+------, J. M. Beaulieu, and M. J. Donoghue, 2009: Mega-phylogeny
+approach for comparative biology: An alternative to supertree and
+supermatrix approaches. \emph{BMC evolutionary biology}, \textbf{9}, 37.
 
 \leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
-Stamatakis A. 2014. RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. Bioinformatics. 30:1312--1313.
-
-\leavevmode\hypertarget{ref-steppan2017muroid}{}%
-Steppan S.J., Schenk J.J. 2017. Muroid rodent phylogenetics: 900-species tree reveals increasing diversification rates. PLoS One. 12.
+Stamatakis, A., 2014: RAxML version 8: A tool for phylogenetic analysis
+and post-analysis of large phylogenies. \emph{Bioinformatics},
+\textbf{30}, 1312--1313,
+doi:\href{https://doi.org/10.1093/bioinformatics/btu033}{10.1093/bioinformatics/btu033}.
 
 \leavevmode\hypertarget{ref-stoltzfus2013phylotastic}{}%
-Stoltzfus A., Lapp H., Matasci N., Deus H., Sidlauskas B., Zmasek C.M., Vaidya G., Pontelli E., Cranston K., Vos R., others.
-2013.
-Phylotastic!
-Making tree-of-life knowledge accessible, reusable and convenient.
-BMC bioinformatics.
-14:158.
-
-\leavevmode\hypertarget{ref-strecker2014fish}{}%
-Strecker A.L., Olden J.D.
-2014.
-Fish species introductions provide novel insights into the patterns and drivers of phylogenetic structure in freshwaters.
-Proceedings of the Royal Society B: Biological Sciences.
-281:20133003.
-
-\leavevmode\hypertarget{ref-thomson2010sparse}{}%
-Thomson R.C., Shaffer H.B. 2010. Sparse supermatrices for phylogenetic inference: Taxonomy, alignment, rogue taxa, and the phylogeny of living turtles. Systematic biology. 59:42--58.
-
-\leavevmode\hypertarget{ref-tucker2012incorporating}{}%
-Tucker C.M., Cadotte M.W., Davies T.J., Rebelo T.G. 2012. Incorporating geographical and evolutionary rarity into conservation prioritization. Conservation Biology. 26:593--601.
-
-\leavevmode\hypertarget{ref-verbruggen2010data}{}%
-Verbruggen H., Maggs C.A., Saunders G.W., Le Gall L., Yoon H.S., De Clerck O.
-2010.
-Data mining approach identifies research priorities and data requirements for resolving the red algal tree of life.
-BMC evolutionary biology.
-10:16.
+Stoltzfus, A. and Coauthors, 2013: Phylotastic! Making tree-of-life
+knowledge accessible, reusable and convenient. \emph{BMC
+bioinformatics}, \textbf{14}, 158,
+doi:\href{https://doi.org/10.1186/1471-2105-14-158}{10.1186/1471-2105-14-158}.
 
 \leavevmode\hypertarget{ref-vos2012nexml}{}%
-Vos R.A., Balhoff J.P., Caravas J.A., Holder M.T., Lapp H., Maddison W.P., Midford P.E., Priyam A., Sukumaran J., Xia X., others.
-2012.
-NeXML: Rich, extensible, and verifiable representation of comparative data and metadata.
-Systematic biology.
-61:675--689.
+Vos, R. A. and Coauthors, 2012: NeXML: Rich, extensible, and verifiable
+representation of comparative data and metadata. \emph{Systematic
+biology}, \textbf{61}, 675--689,
+doi:\href{https://doi.org/10.1093/sysbio/sys025}{10.1093/sysbio/sys025}.
 
 \leavevmode\hypertarget{ref-webb2010biodiversity}{}%
-Webb C.O., Slik J.F., Triono T. 2010. Biodiversity inventory and informatics in southeast asia. Biodiversity and Conservation. 19:955--972.
-
-\leavevmode\hypertarget{ref-weber2014defense}{}%
-Weber M.G., Agrawal A.A. 2014. Defense mutualisms enhance plant diversification. Proceedings of the National Academy of Sciences. 111:16442--16447.
+Webb, C. O., J. F. Slik, and T. Triono, 2010: Biodiversity inventory and
+informatics in southeast asia. \emph{Biodiversity and Conservation},
+\textbf{19}, 955--972,
+doi:\href{https://doi.org/10.1007/s10531-010-9817-x}{10.1007/s10531-010-9817-x}.
 
 \leavevmode\hypertarget{ref-wheeler2000database}{}%
-Wheeler D.L., Chappey C., Lash A.E., Leipe D.D., Madden T.L., Schuler G.D., Tatusova T.A., Rapp B.A.
-2000.
-Database resources of the national center for biotechnology information.
-Nucleic acids research.
-28:10--14.
+Wheeler, D. L., C. Chappey, A. E. Lash, D. D. Leipe, T. L. Madden, G. D.
+Schuler, T. A. Tatusova, and B. A. Rapp, 2000: Database resources of the
+national center for biotechnology information. \emph{Nucleic acids
+research}, \textbf{28}, 10--14,
+doi:\href{https://doi.org/doi.org/10.1093/nar/28.1.10}{doi.org/10.1093/nar/28.1.10}.
 
 \leavevmode\hypertarget{ref-wu2008simple}{}%
-Wu M., Eisen J.A. 2008. A simple, fast, and accurate method of phylogenomic inference. Genome biology. 9:R151.
+Wu, M., and J. A. Eisen, 2008: A simple, fast, and accurate method of
+phylogenomic inference. \emph{Genome biology}, \textbf{9}, R151,
+doi:\href{https://doi.org/10.1186/gb-2008-9-10-r151}{10.1186/gb-2008-9-10-r151}.
 
 \leavevmode\hypertarget{ref-xu2015ncbiminer}{}%
-Xu X., Dimitrov D., Rahbek C., Wang Z. 2015. NCBIminer: Sequences harvest from genbank. Ecography. 38:426--430.
-
-\leavevmode\hypertarget{ref-yong2010screening}{}%
-Yong L., Zhen M., Qi L., Yanping G., Yuanchun Z., Jianhui L.
-2010.
-Screening data for phylogenetic analysis of land plants: A parallel approach.
-2010 first international conference on networking and distributed computing.:305--308.
-
+Xu, X., D. Dimitrov, C. Rahbek, and Z. Wang, 2015: NCBIminer: Sequences
+harvest from genbank. \emph{Ecography}, \textbf{38}, 426--430,
+doi:\href{https://doi.org/10.1111/ecog.01055}{10.1111/ecog.01055}.
 \end{document}

### 2020-06-24 submission_joss/joss_submission.Rmd
refactoring

--- 2020-05-11 joss_submission.Rmd
@@
 citation_author: Sanchez-Reyes et. al.
-date: 24 April 2020
+date: '`r format(Sys.time(), "%d %B, %Y")`'
 year: 2020

### 2020-06-24 submission_sysbio/submission-sysbio.Rmd
refactoring

---
title: "Systematic Biology regular submission"
subtitle: "Submission and Formatting Instructions"
author:
  - name: First Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
  - name: Second Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
  - name: Third Author
    affiliation:
    - Affiliation 1
    - Affiliation 2
    email:  email@domain
abstract: |
  This document provides a basic template for the 2-page extended abstract and some submission guidelines. Including an abstract is optional, as the document itself is an extended abstract. If included, an abstract should be a single paragraph.
keywords: "Optional, Yet Potentially, Helpful"
output:
  bookdown::pdf_document2:
    template: sysbio-regular-template.tex
    keep_tex: yes
bibliography: paper.bib # Replace with your .bib file
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r body, child="2-body.Rmd"}
```

### 2020-06-24 submission_sysbio/sysbio-regular-template.tex
refactoring

\documentclass[oupdraft]{sysbio}
% \usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
\usepackage{url}

\usepackage{indentfirst}

% Add history information for the article if required
%\history{Received Month X, 20XX;
%revised Month X, 20XX}

\begin{document}

% Title of paper
\title{Systematic Biology \LaTeX{} Template}
% Each important word in the title should begin with a capital letter

% List of authors, with corresponding author marked by asterisk
\author{First author name$^{1,\ast}$, Second author name$^{2}$, and
Third author name,$^{3}$\\[4pt]
% Author addresses
\textit{$^{1}$~Department, Institution, City, Post Code, Country}
\\
\textit{$^{2}$~Department, Institution, City, Post Code, Country}
\\
\textit{$^{3}$~Department, Institution, City, Post Code, Country}
\\[2pt]
% E-mail address for correspondence
\textit{*Corresponding author details here}}
% Identify the name, address, telephone/fax numbers, and e-mail address for the author who will receive proofs and be designated the "corresponding author" in text.

% Running headers of paper:
\markboth%
% First field is the short list of authors
{Author short list}
% Second field is the short title of the paper
{Running head}
% This should be shortened version of the title and no greater than 50 characters

\maketitle

\begin{abstract}
{Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text.}
{Insert key words here.}
\end{abstract}
\newline

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text \citet{Author20XX}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
\bigskip
% Each important word in the heading level 1 should begin with a capital letter; no heading for the introduction
\section{Materials and Methods}
\label{sec2}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text \citet{Author220XX}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

% Each important word in the heading level 2 should begin with a capital letter
\subsection{Second Level Heading}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
\bigskip
\section{Results}
\label{sec3}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

\subsection{Second Level Heading}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

% First word and proper nouns only should begin with a capital letter in heading level 3
\subsubsection{Third level heading.---}Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

\subsubsection{Third level heading.---}Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

\bigskip

\section{Discussion}
\label{sec4}

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text

%If you have any acknowledgements, please include them here.
%\section{Acknowledgements}

%If your paper has accompanying supplementary data, please include the below statement in your PDF.
%\section{Supplementary Material}
%Data available from the Dryad Digital Repository:
% \href{http://dx.doi.org/10.5061/dryad.[NNNN]}%
% {http://dx.doi.org/10.5061/dryad.[NNNN]}.
%\url{http://dx.doi.org/10.5061/dryad.[NNNN]}.

\bigskip\bigskip

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX.

% Alternatively you could enter them by hand, like this:
\begin{thebibliography}{39}
\bibitem[\protect\citeauthoryear{Author}{20XX}]{Author20XX}
Author. 20XX. Article title. (Abbreviated) journal name. 1:1--2.
\bibitem[\protect\citeauthoryear{Author2}{20XX}]{Author220XX}
Author 2. 20XX. Article title. (Abbreviated) journal name. 1:1--2.
\end{thebibliography}

% 1. All authors should be listed i.e. no use of et al.
% 2. Dashes should not be used to replace author names in repeat entries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please include any figure captions on a separate page after the references. Figures themselves should be embedded in the text.

%\begin{figure}[!p]
%\centering\includegraphics{fig1}
%\caption{Figure caption}
%\label{Fig1}
%\end{figure}

%\begin{table}[!p]
% 1. Table titles should be in caps and lowercase
% 2. Footnotes can be used in Tables (a,b,c)}
%\tblcaption{Table title
%\label{Table1}}
%{\tabcolsep=4.25pt
%\begin{tabular}{@{}cccccccccc@{}}
%\tblhead{Heading & Heading & Heading & Heading & Heading}
%Value & Value & Value & Value & Value
%\lastline
%\end{tabular}}
%\end{table}

%If you have any print appendices, please include them at the end of the document.

\end{document}

### 2020-06-24 docs/submission-mee-2020-06-24.tex
mee

--- 2020-06-24 docs/submission-sysbio-2020-06-24.tex
@@
   Obtention of molecular or morphological character data -- get DNA from some organisms
-  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
+  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
 \item
@@
 In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}),
+pipelines that mine already available molecular data from the GenBank repository (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}),
 to obtain homologous characters that can be used for phylogenetic reconstruction.
@@
 
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
+Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
 
@@
 allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000})).
+using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000})).
 
@@
 \item
-  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish et al. \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
+  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
   reason for this is that trees in phylesystem have a set of user defined characteristics
@@
   The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}; Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}),
+  alignments are usually stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
   DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
@@
 \item
-  The Basic Local Alignment Search Tool, BLAST {[}Altschul et al. (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
+  The Basic Local Alignment Search Tool, BLAST {[}Altschul \emph{et al.} (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide
   database, and the remaining sequences on the alignment.
-  The BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
+  The BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
 \item
@@
   with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
-  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine et al.
-(\protect\hyperlink{ref-lemoine2018renewing}{2018}).
+  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine \emph{et al.} (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
 \item
@@
 it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}).
+This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
 Updating the oldest tree, we should get something very similar to the newest tree.
@@
 Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by Schoch et al. (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
+The best tree currently available in OToL was published by Schoch \emph{et al.} (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
 The first step, is to get the Open Tree of Life study id. There are some options to do this:
@@
 
-Phylogeny of the Testudines 6 tips from Crawford et al. (\protect\hyperlink{ref-crawford2012more}{2012})
+Phylogeny of the Testudines 6 tips from Crawford \emph{et al.} (\protect\hyperlink{ref-crawford2012more}{2012})
 There is just one tree in OToL.
@@
 Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
+Mention statistics provided by other tools: PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
 Compare and discuss.
@@
 How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in Webb et al. (\protect\hyperlink{ref-webb2010biodiversity}{2010})
-- obtain trees for ecophylogenetic studies, as mentioned in Helmus and Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
-- one day could be used to sistematize nucleotide databases, such as Genbank (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro and Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus et al. (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})
+- rapid phylogenetic placing of newly discovered species, as mentioned in Webb \emph{et al.} (\protect\hyperlink{ref-webb2010biodiversity}{2010})
+- obtain trees for ecophylogenetic studies, as mentioned in Helmus \& Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
+- one day could be used to sistematize nucleotide databases, such as Genbank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro \& Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
+- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus \emph{et al.} (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})
 
 Things that physcraper does not do:
-- analyse the whole GenBank database (Benson et al.
-\protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al.
-\protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli et al.
-(\protect\hyperlink{ref-antonelli2017toward}{2017}).
+- analyse the whole GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017}).
 There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014})
+- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez \emph{et al.} (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
+- it is not a tree repo, as phylota is, mentioned in Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2014evominer}{2014})
 
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Sanderson et al. (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut
+Sanderson \emph{et al.} (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Wu and Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut
+Wu \& Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith et al. (\protect\hyperlink{ref-smith2009mega}{2009})\strut
+Smith \emph{et al.} (\protect\hyperlink{ref-smith2009mega}{2009})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Peters et al. (\protect\hyperlink{ref-peters2011taming}{2011})\strut
+Peters \emph{et al.} (\protect\hyperlink{ref-peters2011taming}{2011})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Grant and Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut
+Grant \& Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut
+Chesters \& Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Izquierdo-Carrasco et al. (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut
+Izquierdo-Carrasco \emph{et al.} (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Meng et al. (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut
+Meng \emph{et al.} (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Xu et al. (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut
+Xu \emph{et al.} (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-McMahon et al. (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut
+McMahon \emph{et al.} (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Papadopoulou et al. (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut
+Papadopoulou \emph{et al.} (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Kumar et al. (\protect\hyperlink{ref-kumar2015bir}{2015})\strut
+Kumar \emph{et al.} (\protect\hyperlink{ref-kumar2015bir}{2015})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut
+Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut
+Drori \emph{et al.} (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith and Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut
+Smith \& Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut
+Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Bennett et al. (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut
+Bennett \emph{et al.} (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut
 \end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
@@
 
-According to Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
+According to Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
 of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
@@
 
-PhyloFinder (Chen et al. \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
+PhyloFinder (Chen \emph{et al.} \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
 trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
@@
 
-PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
+PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
 assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
@@
 
-PhySpeTre (Fang et al. \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
+PhySpeTre (Fang \emph{et al.} \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
 pipeline.
@@
 
-Datataxa Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.
+Datataxa Ruiz-Sanchez \emph{et al.} (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.
 
@@
 Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
-database (Sanderson et al. \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+database (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 
@@
   \item
-    original publication of PHLAWD (Smith et al. \protect\hyperlink{ref-smith2009mega}{2009})
+    original publication of PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009})
   \item
     an analysis identifying research priorities and data requirements for resolving
-    the red algal tree of life (Verbruggen et al. \protect\hyperlink{ref-verbruggen2010data}{2010})
-  \item
-    Beaulieu et al. (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive
+    the red algal tree of life (Verbruggen \emph{et al.} \protect\hyperlink{ref-verbruggen2010data}{2010})
+  \item
+    Beaulieu \emph{et al.} (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive
     phylogeny from mined DNA sequence data, (even if no phylogeny was really published
@@
   \item
-    a review for ecologists about phylogenetic tools (Roquet et al. \protect\hyperlink{ref-roquet2013building}{2013})
-  \item
-    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith and Brown \protect\hyperlink{ref-smith2018constructing}{2018})
+    a review for ecologists about phylogenetic tools (Roquet \emph{et al.} \protect\hyperlink{ref-roquet2013building}{2013})
+  \item
+    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith \& Brown \protect\hyperlink{ref-smith2018constructing}{2018})
   \item
     a study presenting an ``assembly and alignment free'' method for phylogenetic reconstruction
-    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan et al. \protect\hyperlink{ref-fan2015assembly}{2015}).
-  \item
-    nexml format presentation (Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses
+    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan \emph{et al.} \protect\hyperlink{ref-fan2015assembly}{2015}).
+  \item
+    nexml format presentation (Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses
     stored phyloinformatic data that could benefit from adopting nexml, to increase
@@
     a study of fruit evolution, analysing a previously published phylogeny of 8911
-    tips of the Campanulidae, constructed with PHLAWD (Beaulieu and Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})
-  \item
-    a study of Southeast Asia plant biodiversity inventory (Webb et al. \protect\hyperlink{ref-webb2010biodiversity}{2010}) -
+    tips of the Campanulidae, constructed with PHLAWD (Beaulieu \& Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})
+  \item
+    a study of Southeast Asia plant biodiversity inventory (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}) -
     cites phylota as a tool that would allow rapid phylogenetic placing of newly
@@
   \item
-    a study of wood density for carbon stock assessments (Flores and Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),
+    a study of wood density for carbon stock assessments (Flores \& Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),
     cites phylota as an initiative to ``get supertrees resolved up to species level''.
   \item
-    a study proposing something similar to Open tree but applied only to land plants (Beaulieu et al.
-\protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})
-  \item
-    an analysis of the phylogenetic diversity-area curve (Helmus and Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
+    a study proposing something similar to Open tree but applied only to land plants (Beaulieu \emph{et al.} \protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})
+  \item
+    an analysis of the phylogenetic diversity-area curve (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
     cited phylota as a method alternative to phylomatic to ``obtain plant phylogenetic
@@
     a study generating a phylogeny of 6,098 species of vascular plants from China
-    (Chen et al.
-\protect\hyperlink{ref-chen2016tree}{2016}) - uses DarwinTree (Meng et al.
-\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}) and generates sequence
+    (Chen \emph{et al.} \protect\hyperlink{ref-chen2016tree}{2016}) - uses DarwinTree (Meng \emph{et al.} \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}) and generates sequence
     data \emph{de novo} for 781 genera.
@@
     a review of the state of methods and knowledge generated by molecular systematics
-    (San Mauro and Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
+    (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
     for large-scale molecular phylogenetics analysis''.
   \item
-    the first phylotastic paper (Stoltzfus et al. \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny
+    the first phylotastic paper (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny
     related resource that provides ways to generate custom species trees for downstream use''.
   \item
-    Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017}) cites phylota as a ``pipeline that pre-processes entire GenBank
+    Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017}) cites phylota as a ``pipeline that pre-processes entire GenBank
     releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
@@
   \item
-    Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014}) present an algorithm for mining of frequent subtrees (common patterns)
+    Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2014evominer}{2014}) present an algorithm for mining of frequent subtrees (common patterns)
     in collections of phylogenetic trees, as a way to extract meaningful phylogenetic
@@
     and majority-rule trees. They cite phylota as one of such tree collections available
-    along with TreeBASE (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}).
-  \item
-    Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) cites phylota as a ``program providing basic statistics
+    along with TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
+  \item
+    Ranwez \emph{et al.} (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) cites phylota as a ``program providing basic statistics
     on data availability for molecular datasets''. They propose a tool to upload and
@@
   \item
-    Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) cites phylota as a data mining tool that compiles metadata
+    Chesters \& Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) cites phylota as a data mining tool that compiles metadata
     from mining of public DNA databases ``for construction of large phylogenetic trees
@@
     in public databases are insufficient and that careful partitioning of orthologous
-    sequences is needed for supermatrix construction. Chesters and Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) present
+    sequences is needed for supermatrix construction. Chesters \& Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) present
     a procedure that minimizes the problem of forming multilocus species units in
@@
   \item
-    Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014}) present an algorithm to delineate species form GenBank
+    Chesters \& Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014}) present an algorithm to delineate species form GenBank
     DNA data, and cites phylota as a tool that partitions ``the contents of a database
@@
     the paper presenting phylotaR, a pipeline that recreates the phylota output
-    but uses the most updated GenBank release, and is available in R (Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}),
+    but uses the most updated GenBank release, and is available in R (Bennett \emph{et al.} \protect\hyperlink{ref-bennett2018phylotar}{2018}),
     cites phylota as its predecessor and inspiration. The authors mention that phylotaR
@@
     The paper presenting STBase, a database of one million precomputed species
-    trees (Deepak \protect\hyperlink{ref-deepak2013managing}{2013}; McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as a databse of gene trees or mul-trees,
+    trees (Deepak \protect\hyperlink{ref-deepak2013managing}{2013}; McMahon \emph{et al.} \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as a databse of gene trees or mul-trees,
     ``trees having multiple sequences with the same taxon name''.
   \item
-    Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018}) present a Web‐based, user-friendly, online tool for species-tree
+    Drori \emph{et al.} (\protect\hyperlink{ref-drori2018onetwotree}{2018}) present a Web‐based, user-friendly, online tool for species-tree
     reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available
@@
   \item
-    the study that present DarwinTree (Meng et al. \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}), and all derived studies: the study
+    the study that present DarwinTree (Meng \emph{et al.} \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}), and all derived studies: the study
     presenting an approach to screen sequence data for The Platform
     for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST (Yong et al.
-\protect\hyperlink{ref-yong2010screening}{2010}), as well as Gao et al.
-(\protect\hyperlink{ref-gao2011solution}{2011}), Li et al.
-(\protect\hyperlink{ref-li2013partfasttree}{2013}),
-    Meng et al.
-(\protect\hyperlink{ref-meng2014rapidtree}{2014}), Meng et al.
-(\protect\hyperlink{ref-meng2015sotree}{2015}\protect\hyperlink{ref-meng2015sotree}{c}), and Meng et al.
-(\protect\hyperlink{ref-meng2015solution}{2015}\protect\hyperlink{ref-meng2015solution}{b}), all cite phylota using the exact same
+    to parallelize BLAST (Yong \emph{et al.} \protect\hyperlink{ref-yong2010screening}{2010}), as well as Gao \emph{et al.} (\protect\hyperlink{ref-gao2011solution}{2011}), Li \emph{et al.} (\protect\hyperlink{ref-li2013partfasttree}{2013}),
+    Meng \emph{et al.} (\protect\hyperlink{ref-meng2014rapidtree}{2014}), Meng \emph{et al.} (\protect\hyperlink{ref-meng2015sotree}{2015}\protect\hyperlink{ref-meng2015sotree}{c}), and Meng \emph{et al.} (\protect\hyperlink{ref-meng2015solution}{2015}\protect\hyperlink{ref-meng2015solution}{b}), all cite phylota using the exact same
     introduction and sentence: as one among other ``studies based on data mining large numbers of taxa or loci''.
@@
     A study presenting a tool to asses gene sequence quality for automatic
-    construction of databases (Meng et al. \protect\hyperlink{ref-meng2012gsqct}{2012}\protect\hyperlink{ref-meng2012gsqct}{a}), as well as their parallelized version using MapReduce
-    (Meng et al.
-\protect\hyperlink{ref-meng2012cloud}{2012}\protect\hyperlink{ref-meng2012cloud}{b}), cite phylota (along with Yong et al.
-(\protect\hyperlink{ref-yong2010screening}{2010})) as a tool that relies on sequence
+    construction of databases (Meng \emph{et al.} \protect\hyperlink{ref-meng2012gsqct}{2012}\protect\hyperlink{ref-meng2012gsqct}{a}), as well as their parallelized version using MapReduce
+    (Meng \emph{et al.} \protect\hyperlink{ref-meng2012cloud}{2012}\protect\hyperlink{ref-meng2012cloud}{b}), cite phylota (along with Yong \emph{et al.} (\protect\hyperlink{ref-yong2010screening}{2010})) as a tool that relies on sequence
     similarity (BLAST) and not taxon name annotations in the database, for mining
@@
     researchers ``to acces equally and globally, without travel, a {[}phylogenetic?{]} model
-    of plants at the kingdom level'' (Jones et al. \protect\hyperlink{ref-jones2014trends}{2014}).
+    of plants at the kingdom level'' (Jones \emph{et al.} \protect\hyperlink{ref-jones2014trends}{2014}).
   \item
@@
     and ultimately facilitate cross-analyses and collaboration within the legume-systematics
-    community and with other stakeholders'' (Bruneau et al. \protect\hyperlink{ref-bruneau2019towards}{2019}), cites phylota (along with supersmart and pyphlawd) as a
+    community and with other stakeholders'' (Bruneau \emph{et al.} \protect\hyperlink{ref-bruneau2019towards}{2019}), cites phylota (along with supersmart and pyphlawd) as a
     ``pipeline for large-scale retrieval of GenBank data of particular taxa or clades''.
@@
     A study on morphological evolution of electric fish skull, that uses phylotaR
-    to retrieve sequences of the family Apteronotidae, order Gymnotiformes (Evans et al. \protect\hyperlink{ref-evans2019bony}{2019}),
+    to retrieve sequences of the family Apteronotidae, order Gymnotiformes (Evans \emph{et al.} \protect\hyperlink{ref-evans2019bony}{2019}),
     cites phylota as the inspiration and fundament of phylotaR.
@@
     sequence data in GenBank'' (Deepak \protect\hyperlink{ref-deepak2010searchtree}{2010}), that seems to be the preliminary
-    work for STBase (McMahon et al. \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as ``a standard strategy, to assemble sets of homologous sequences
+    work for STBase (McMahon \emph{et al.} \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as ``a standard strategy, to assemble sets of homologous sequences
     (clusters) from a database of all-against-all BLAST searches, {[}in which{]} clusters
@@
     In opposition, SearchTree uses true agglomerative hierarchical clustering
-    (AHC: Day and Edelsbrunner (\protect\hyperlink{ref-day1984efficient}{1984})) based on the BLAST estimates of sequence dissimilarity
+    (AHC: Day \& Edelsbrunner (\protect\hyperlink{ref-day1984efficient}{1984})) based on the BLAST estimates of sequence dissimilarity
     rather than the NCBI tree".
@@
   \item
-    the study presenting phylotol (Cerón-Romero et al. \protect\hyperlink{ref-ceron2019phylotol}{2019}), cites phylota as a tool
+    the study presenting phylotol (Cerón-Romero \emph{et al.} \protect\hyperlink{ref-ceron2019phylotol}{2019}), cites phylota as a tool
     that ``focus on the identification and collection of homologous genes from public
@@
   \item
-    Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) develops datataxa and cite phylota as ``software that has
+    Ruiz-Sanchez \emph{et al.} (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) develops datataxa and cite phylota as ``software that has
     been developed to mine the massive amount of information stored in GenBank'',
-    along with its R version (phylotaR; Bennett et al. \protect\hyperlink{ref-bennett2018phylotar}{2018}) and restez \url{https://www.rdocumen-tation.org/packages/restez/versions/1.0.0}.
-  \item
-    The phylotastic project (Stoltzfus et al.
-\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny-related resource providing ways to generate custom species trees \emph{de novo} for downstream use'' along with CIPRES.
+    along with its R version (phylotaR; Bennett \emph{et al.} \protect\hyperlink{ref-bennett2018phylotar}{2018}) and restez \url{https://www.rdocumen-tation.org/packages/restez/versions/1.0.0}.
+  \item
+    The phylotastic project (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny-related resource providing ways to generate custom species trees \emph{de novo} for downstream use'' along with CIPRES.
   \end{itemize}
@@
   \item
-    A 1000 tip phylogeny of the family of the nightshades (Särkinen et al. \protect\hyperlink{ref-sarkinen2013solanaceae}{2013})
-  \item
-    A 56 tip phylogeny of crustacean zooplancton (Helmus et al. \protect\hyperlink{ref-helmus2010communities}{2010}) -- ecological study
-  \item
-    A 63 tip phylogeny of the Salmonidae family (Crête-Lafrenière et al. \protect\hyperlink{ref-crete2012salmonidae}{2012})
-  \item
-    A 321 tip phylogeny of Testudines (Thomson and Shaffer \protect\hyperlink{ref-thomson2010sparse}{2010})
-  \item
-    A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish (Martin and Wainwright \protect\hyperlink{ref-martin2011trophic}{2011})
+    A 1000 tip phylogeny of the family of the nightshades (Särkinen \emph{et al.} \protect\hyperlink{ref-sarkinen2013solanaceae}{2013})
+  \item
+    A 56 tip phylogeny of crustacean zooplancton (Helmus \emph{et al.} \protect\hyperlink{ref-helmus2010communities}{2010}) -- ecological study
+  \item
+    A 63 tip phylogeny of the Salmonidae family (Crête-Lafrenière \emph{et al.} \protect\hyperlink{ref-crete2012salmonidae}{2012})
+  \item
+    A 321 tip phylogeny of Testudines (Thomson \& Shaffer \protect\hyperlink{ref-thomson2010sparse}{2010})
+  \item
+    A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish (Martin \& Wainwright \protect\hyperlink{ref-martin2011trophic}{2011})
   \item
@@
   \item
-    A 2,573 species phylogeny of the Papilionoidea (Hardy and Otto \protect\hyperlink{ref-hardy2014specialization}{2014})
-  \item
-    A 23 taxa phylogeny of the California flora (Anacker et al. \protect\hyperlink{ref-anacker2011origins}{2011})
+    A 2,573 species phylogeny of the Papilionoidea (Hardy \& Otto \protect\hyperlink{ref-hardy2014specialization}{2014})
+  \item
+    A 23 taxa phylogeny of the California flora (Anacker \emph{et al.} \protect\hyperlink{ref-anacker2011origins}{2011})
   \item
@@
     \emph{Polygoneae} (Polygoneaceae), \emph{Senna} (Fabaceae), \emph{Turnera} (Passifloraceae), and \emph{Viburnum}
-    (Adoxaceae) (Weber and Agrawal \protect\hyperlink{ref-weber2014defense}{2014}).
+    (Adoxaceae) (Weber \& Agrawal \protect\hyperlink{ref-weber2014defense}{2014}).
   \item
     To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
-    sampled at different taxonomic levels (Faurby and Svenning \protect\hyperlink{ref-faurby2015species}{2015})
-  \item
-    A 900 species tree of muroid rodents, Muroidea (Steppan and Schenk \protect\hyperlink{ref-steppan2017muroid}{2017}), where 300
+    sampled at different taxonomic levels (Faurby \& Svenning \protect\hyperlink{ref-faurby2015species}{2015})
+  \item
+    A 900 species tree of muroid rodents, Muroidea (Steppan \& Schenk \protect\hyperlink{ref-steppan2017muroid}{2017}), where 300
     species were newly added by the study and the rest obtained using phylota.
   \item
-    A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales (Ickert-Bond et al. \protect\hyperlink{ref-ickert2009fossil}{2009})
-  \item
-    A 1061 genera phylogeny of the Oscine birds (Selvatti et al. \protect\hyperlink{ref-selvatti2015paleogene}{2015})
-  \item
-    A 268 species phylogeny of sharks, representing all 8 orders and 32 families (Sorenson et al.
-\protect\hyperlink{ref-sorenson2014effect}{2014}; Sorenson \protect\hyperlink{ref-sorenson2014evolution}{2014})
-  \item
-    A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region (Tucker et al. \protect\hyperlink{ref-tucker2012incorporating}{2012}).
-  \item
-    A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects (Hardy and Cook \protect\hyperlink{ref-hardy2010gall}{2010}).
-  \item
-    A 196 species phylogeny of the family Boraginaceae (Nazaire and Hufford \protect\hyperlink{ref-nazaire2012broad}{2012}). The authors
+    A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales (Ickert-Bond \emph{et al.} \protect\hyperlink{ref-ickert2009fossil}{2009})
+  \item
+    A 1061 genera phylogeny of the Oscine birds (Selvatti \emph{et al.} \protect\hyperlink{ref-selvatti2015paleogene}{2015})
+  \item
+    A 268 species phylogeny of sharks, representing all 8 orders and 32 families (Sorenson \protect\hyperlink{ref-sorenson2014evolution}{2014}; Sorenson \emph{et al.} \protect\hyperlink{ref-sorenson2014effect}{2014})
+  \item
+    A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region (Tucker \emph{et al.} \protect\hyperlink{ref-tucker2012incorporating}{2012}).
+  \item
+    A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects (Hardy \& Cook \protect\hyperlink{ref-hardy2010gall}{2010}).
+  \item
+    A 196 species phylogeny of the family Boraginaceae (Nazaire \& Hufford \protect\hyperlink{ref-nazaire2012broad}{2012}). The authors
     actually found data for 318 Boraginaceae spp using phylota, but decided to reduce
@@
   \item
-    A phylogeny of 401 species of scale insects Coccoidea, Hemiptera (Ross et al. \protect\hyperlink{ref-ross2013large}{2013}),
+    A phylogeny of 401 species of scale insects Coccoidea, Hemiptera (Ross \emph{et al.} \protect\hyperlink{ref-ross2013large}{2013}),
     with some sequences generated \emph{de novo}.
@@
     lizards, agamids and diplodactyline geckos, groups considered to be radiating
-    in the Australia's Great Victoria Desert (Rabosky et al. \protect\hyperlink{ref-rabosky2011species}{2011})
+    in the Australia's Great Victoria Desert (Rabosky \emph{et al.} \protect\hyperlink{ref-rabosky2011species}{2011})
   \item
     A phylogeny of 91 species of sparid and centracanthid fishes, Sparidae, Percomorpha,
-    plus 2 outgroups, a lethrinid and a nemipterid exemplar (Santini et al. \protect\hyperlink{ref-santini2014first}{2014}).
+    plus 2 outgroups, a lethrinid and a nemipterid exemplar (Santini \emph{et al.} \protect\hyperlink{ref-santini2014first}{2014}).
   \item
@@
     tribes; Euterpeae, Leopoldinieae, Pelagodoxeae, Manicarieae, Geonomateae and Areceae
-    (Faurby et al. \protect\hyperlink{ref-faurby2016all}{2016}).
+    (Faurby \emph{et al.} \protect\hyperlink{ref-faurby2016all}{2016}).
   \end{enumerate}
@@
     A phylogeny of 768 Gesneriaceae species and 58 outgroups for a total species
-    sampling of 826 taxa (Roalson and Roberts \protect\hyperlink{ref-roalson2016distinct}{2016}) some sequence were generated \emph{de novo}.
+    sampling of 826 taxa (Roalson \& Roberts \protect\hyperlink{ref-roalson2016distinct}{2016}) some sequence were generated \emph{de novo}.
   \item
     A phylogeny of 47 species of scombrid fishes, with 2 outgroups, a gempylid and
-    a trichiurid (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013}).
-  \item
-    to update a dataset underlying a large-scale fern phylogeny (Lehtonen et al. \protect\hyperlink{ref-lehtonen2017environmentally}{2017}),
+    a trichiurid (Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}).
+  \item
+    to update a dataset underlying a large-scale fern phylogeny (Lehtonen \emph{et al.} \protect\hyperlink{ref-lehtonen2017environmentally}{2017}),
     data set in \url{https://zenodo.org/record/345670\#.Xr9QFRPYqqg}, also in TreeBASE,
@@
     A phylogeny of 13 species of billfishes, order Istiophoriformes: Acanthomorpha,
-    and four outgroups (Santini and Sorenson \protect\hyperlink{ref-santini2013first}{2013})
-  \item
-    A phylogeny of 765 aphid species, family Aphididae (Hardy et al. \protect\hyperlink{ref-hardy2015evolution}{2015})
-  \item
-    A phylogeny of less than 100 taxa of the family Ranunculaceae (Lehtonen et al. \protect\hyperlink{ref-lehtonen2016sensitive}{2016}),
+    and four outgroups (Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013})
+  \item
+    A phylogeny of 765 aphid species, family Aphididae (Hardy \emph{et al.} \protect\hyperlink{ref-hardy2015evolution}{2015})
+  \item
+    A phylogeny of less than 100 taxa of the family Ranunculaceae (Lehtonen \emph{et al.} \protect\hyperlink{ref-lehtonen2016sensitive}{2016}),
     even though they retrieved info from phylota for 194 taxa within the family, they
@@
     A phylogeny of 144 neobatrachian genera, assuming the monophyletic status of
-    genera to increase matrix-filling levels (Frazao et al. \protect\hyperlink{ref-frazao2015gondwana}{2015}).
+    genera to increase matrix-filling levels (Frazao \emph{et al.} \protect\hyperlink{ref-frazao2015gondwana}{2015}).
   \item
@@
   \item
-    A phylogeny of species of freshwater fish endemic to NorthAmerica (Strecker and Olden \protect\hyperlink{ref-strecker2014fish}{2014}),
+    A phylogeny of species of freshwater fish endemic to NorthAmerica (Strecker \& Olden \protect\hyperlink{ref-strecker2014fish}{2014}),
     phylota found data for 54 out of 66 spp.
   \item
-    A phylogeny of 520 species of the order Ericales (Hardy and Cook \protect\hyperlink{ref-hardy2012testing}{2012})
+    A phylogeny of 520 species of the order Ericales (Hardy \& Cook \protect\hyperlink{ref-hardy2012testing}{2012})
   \item
     A phylgeny of 16 fish species of the family Sphyraenidae (Percomorpha), as well
-    as two outgroup species of the Centropomidae (barracudas) (Santini et al. \protect\hyperlink{ref-santini2015first}{2015})
-  \item
-    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas et al. \protect\hyperlink{ref-garcia2016role}{2016})
-  \item
-    Kolmann et al. (\protect\hyperlink{ref-kolmann2017dna}{2017}) uses phylota to download all 1691 co1 sequences belonging to
+    as two outgroup species of the Centropomidae (barracudas) (Santini \emph{et al.} \protect\hyperlink{ref-santini2015first}{2015})
+  \item
+    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas \emph{et al.} \protect\hyperlink{ref-garcia2016role}{2016})
+  \item
+    Kolmann \emph{et al.} (\protect\hyperlink{ref-kolmann2017dna}{2017}) uses phylota to download all 1691 co1 sequences belonging to
     the order Carchariniformes, to place phylogenetically DNA samples obtained from
@@
     A phylogeny of 329 bird species in the Tyrannidae (77\% of the species in the
-    family) (Gómez Bahamón and others \protect\hyperlink{ref-gomez2015behavioral}{2015}; Gómez-Bahamón et al. \protect\hyperlink{ref-gomez2020speciation}{2020})
+    family) (Gómez Bahamón \& others \protect\hyperlink{ref-gomez2015behavioral}{2015}; Gómez-Bahamón \emph{et al.} \protect\hyperlink{ref-gomez2020speciation}{2020})
   \item
@@
     plus 4 outgroup sequences from \emph{Stichopus ocellatus}, all belonging to the order
-    Apodida of sea cucumbers (Kamarudin et al. \protect\hyperlink{ref-kamarudin2016phylogenetic}{2016})
+    Apodida of sea cucumbers (Kamarudin \emph{et al.} \protect\hyperlink{ref-kamarudin2016phylogenetic}{2016})
   \item
@@
     and \emph{Zea mays}, (Salariato \protect\hyperlink{ref-salariato2010filogenia}{2010}). Interestingly, phylota was not used
-    in the published study of the thesis (Salariato et al. \protect\hyperlink{ref-salariato2010molecular}{2010}). Ingroup sequences were generated \emph{de novo}.
+    in the published study of the thesis (Salariato \emph{et al.} \protect\hyperlink{ref-salariato2010molecular}{2010}). Ingroup sequences were generated \emph{de novo}.
   \item
@@
   \item
-    The 10k trees project (Arnold et al. \protect\hyperlink{ref-arnold201010ktrees}{2010}) uses phylota to construct a tree of 301 primate species
+    The 10k trees project (Arnold \emph{et al.} \protect\hyperlink{ref-arnold201010ktrees}{2010}) uses phylota to construct a tree of 301 primate species
     and the outgroup species \emph{Galeopterus variegates}, a tree of 17 extant odd-toed
@@
   \item
-    A study of the effect of poliploidy on niche evolution (Baniaga et al. \protect\hyperlink{ref-baniaga2018polyploid}{2018}),
+    A study of the effect of poliploidy on niche evolution (Baniaga \emph{et al.} \protect\hyperlink{ref-baniaga2018polyploid}{2018}),
     uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
@@
   \item
-    A 812 tips phylogeny of the Order Chiroptera (Shi and Rabosky \protect\hyperlink{ref-shi2015speciation}{2015}) -- dataset
+    A 812 tips phylogeny of the Order Chiroptera (Shi \& Rabosky \protect\hyperlink{ref-shi2015speciation}{2015}) -- dataset
     constructed with PHLAWD
   \item
-    A 1276 tips phylogeny of the Fabaceae (Group et al. \protect\hyperlink{ref-legume2013legume}{2013}) -- dataset constructed
+    A 1276 tips phylogeny of the Fabaceae (Group \emph{et al.} \protect\hyperlink{ref-legume2013legume}{2013}) -- dataset constructed
     by hand (I think??)
@@
     A review of dated phylogenies of fire-prone tropical savanna species from Brazil
-    (Simon and Pennington \protect\hyperlink{ref-simon2012cerrado}{2012}) -- just for discussion of the lack of markers available for
+    (Simon \& Pennington \protect\hyperlink{ref-simon2012cerrado}{2012}) -- just for discussion of the lack of markers available for
     these species on GenBank
@@
     the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
-    to demonstrate and exemplify phylogenetic decisiveness (Sanderson et al. \protect\hyperlink{ref-sanderson2010phylogenomics}{2010})
-  \item
-    Mentioned in a PHD thesis (Gagnon and others \protect\hyperlink{ref-gagnon2016systematique}{2016}), but not on the final publication (Gagnon et al. \protect\hyperlink{ref-gagnon2016new}{2016}),
+    to demonstrate and exemplify phylogenetic decisiveness (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2010phylogenomics}{2010})
+  \item
+    Mentioned in a PHD thesis (Gagnon \& others \protect\hyperlink{ref-gagnon2016systematique}{2016}), but not on the final publication (Gagnon \emph{et al.} \protect\hyperlink{ref-gagnon2016new}{2016}),
     phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
@@
   \item
-    In this 630 tip phylogeny of the Caryophyllaceae study (Greenberg and Donoghue \protect\hyperlink{ref-greenberg2011caryophyllaceae}{2011}) it might have been originally
+    In this 630 tip phylogeny of the Caryophyllaceae study (Greenberg \& Donoghue \protect\hyperlink{ref-greenberg2011caryophyllaceae}{2011}) it might have been originally
     cited as an example of large phylogenies that reflect well supported relationships
@@
     genera, and 760 families within the order Insecta (Chesters \protect\hyperlink{ref-chesters2017construction}{2017}),
-    uses its own algorithm (SOPHI) to mine public DNA databases (Chesters and Zhu \protect\hyperlink{ref-chesters2014protocol}{2014}).
+    uses its own algorithm (SOPHI) to mine public DNA databases (Chesters \& Zhu \protect\hyperlink{ref-chesters2014protocol}{2014}).
     It does not cite phylota as it should, but includes it in their references.
@@
   \item
-    Deepak et al. (\protect\hyperlink{ref-deepak2013extracting}{2013}) uses a sample of phylota trees to test their method to
+    Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2013extracting}{2013}) uses a sample of phylota trees to test their method to
     remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
@@
   \item
-    A review by Sanderson et al. (\protect\hyperlink{ref-sanderson2016perspective}{2016}), takes 134 595 gene trees from phylota
+    A review by Sanderson \emph{et al.} (\protect\hyperlink{ref-sanderson2016perspective}{2016}), takes 134 595 gene trees from phylota
     GenBank rel. 176 and estimates its degree of resolutin, calculating that less
@@
 \leavevmode\hypertarget{ref-altschul1990basic}{}%
-Altschul S.F., Gish W., Miller W., Myers E.W., Lipman D.J. 1990. Basic local alignment search tool. Journal of molecular biology. 215:403--410.
+Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of molecular biology}, \textbf{215}, 403--410.
 
 \leavevmode\hypertarget{ref-anacker2011origins}{}%
-Anacker B.L., Whittall J.B., Goldberg E.E., Harrison S.P.
-2011.
+Anacker, B.L., Whittall, J.B., Goldberg, E.E.
+\& Harrison, S.P.
+(2011).
 Origins and consequences of serpentine endemism in the california flora.
-Evolution: International Journal of Organic Evolution.
-65:365--376.
+\emph{Evolution: International Journal of Organic Evolution}, \textbf{65}, 365--376.
 
 \leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli A., Hettling H., Condamine F.L., Vos K., Nilsson R.H., Sanderson M.J., Sauquet H., Scharn R., Silvestro D., Töpel M., others.
-2017.
+Antonelli, A., Hettling, H., Condamine, F.L., Vos, K., Nilsson, R.H., Sanderson, M.J., Sauquet, H., Scharn, R., Silvestro, D., Töpel, M.
+\& others.
+(2017).
 Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
-Systematic Biology.
-66:152--166.
+\emph{Systematic Biology}, \textbf{66}, 152--166.
 
 \leavevmode\hypertarget{ref-arnold201010ktrees}{}%
-Arnold C., Matthews L.J., Nunn C.L. 2010. The 10kTrees website: A new online resource for primate phylogeny. Evolutionary Anthropology: Issues, News, and Reviews. 19:114--118.
+Arnold, C., Matthews, L.J. \& Nunn, C.L. (2010). The 10kTrees website: A new online resource for primate phylogeny. \emph{Evolutionary Anthropology: Issues, News, and Reviews}, \textbf{19}, 114--118.
 
 \leavevmode\hypertarget{ref-baniaga2018polyploid}{}%
-Baniaga A.E., Marx H.E., Arrigo N., Barker M.S. 2018. Polyploid plants have faster rates of multivariate climatic niche evolution than their diploid relatives. BioRxiv.:406314.
+Baniaga, A.E., Marx, H.E., Arrigo, N. \& Barker, M.S. (2018). Polyploid plants have faster rates of multivariate climatic niche evolution than their diploid relatives. \emph{BioRxiv}, 406314.
 
 \leavevmode\hypertarget{ref-beaulieu2013fruit}{}%
-Beaulieu J.M., Donoghue M.J. 2013. Fruit evolution and diversification in campanulid angiosperms. Evolution. 67:3132--3144.
+Beaulieu, J.M. \& Donoghue, M.J. (2013). Fruit evolution and diversification in campanulid angiosperms. \emph{Evolution}, \textbf{67}, 3132--3144.
 
 \leavevmode\hypertarget{ref-beaulieu2012modeling}{}%
-Beaulieu J.M., Jhwueng D.-C., Boettiger C., O'Meara B.C.
-2012a.
+Beaulieu, J.M., Jhwueng, D.-C., Boettiger, C.
+\& O'Meara, B.C.
+(2012a).
 Modeling stabilizing selection: Expanding the ornstein--uhlenbeck model of adaptive evolution.
-Evolution: International Journal of Organic Evolution.
-66:2369--2383.
+\emph{Evolution: International Journal of Organic Evolution}, \textbf{66}, 2369--2383.
 
 \leavevmode\hypertarget{ref-beaulieu2012synthesizing}{}%
-Beaulieu J.M., Ree R.H., Cavender-Bares J., Weiblen G.D., Donoghue M.J. 2012b. Synthesizing phylogenetic knowledge for ecological research. Ecology. 93:S4--S13.
+Beaulieu, J.M., Ree, R.H., Cavender-Bares, J., Weiblen, G.D. \& Donoghue, M.J. (2012b). Synthesizing phylogenetic knowledge for ecological research. \emph{Ecology}, \textbf{93}, S4--S13.
 
 \leavevmode\hypertarget{ref-bennett2018phylotar}{}%
-Bennett D.J., Hettling H., Silvestro D., Zizka A., Bacon C.D., Faurby S., Vos R.A., Antonelli A.
-2018.
+Bennett, D.J., Hettling, H., Silvestro, D., Zizka, A., Bacon, C.D., Faurby, S., Vos, R.A.
+\& Antonelli, A.
+(2018).
 PhylotaR: An automated pipeline for retrieving orthologous dna sequences from genbank in r.
-Life.
-8:20.
+\emph{Life}, \textbf{8}, 20.
 
 \leavevmode\hypertarget{ref-benson2000genbank}{}%
-Benson D.A., Karsch-Mizrachi I., Lipman D.J., Ostell J., Rapp B.A., Wheeler D.L. 2000. GenBank. Nucleic acids research. 28:15--18.
+Benson, D.A., Karsch-Mizrachi, I., Lipman, D.J., Ostell, J., Rapp, B.A. \& Wheeler, D.L. (2000). GenBank. \emph{Nucleic acids research}, \textbf{28}, 15--18.
 
 \leavevmode\hypertarget{ref-blackmon2017synthesis}{}%
-Blackmon H. 2017. Synthesis and phylogenetic comparative analyses of the causes and consequences of karyotype evolution in arthropods..
+Blackmon, H.
+(2017).
+\emph{Synthesis and phylogenetic comparative analyses of the causes and consequences of karyotype evolution in arthropods}.
+PhD thesis thesis, University of Texas, Arlington.
+Retrieved from \url{http://hdl.handle.net/10106/26711}
 
 \leavevmode\hypertarget{ref-bruneau2019towards}{}%
-Bruneau A., Borges L.M., Allkin R., Egan A.N., De La Estrella M., Javadi F., Klitgaard B., Miller J.T., Murphy D.J., Sinou C., others.
-2019.
+Bruneau, A., Borges, L.M., Allkin, R., Egan, A.N., De La Estrella, M., Javadi, F., Klitgaard, B., Miller, J.T., Murphy, D.J., Sinou, C.
+\& others.
+(2019).
 Towards a new online species-information system for legumes.
-Australian Systematic Botany.
-32:495--518.
+\emph{Australian Systematic Botany}, \textbf{32}, 495--518.
 
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
-Camacho C., George C., Vahram A., Ning M., Jason P., Kevin B., Thomas L. 2009. BLAST+: Architecture and applications. BMC bioinformatics. 10:421.
+Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC bioinformatics}, \textbf{10}, 421.
 
 \leavevmode\hypertarget{ref-ceron2019phylotol}{}%
-Cerón-Romero M.A., Maurer-Alcalá X.X., Grattepanche J.-D., Yan Y., Fonseca M.M., Katz L.
-2019.
+Cerón-Romero, M.A., Maurer-Alcalá, X.X., Grattepanche, J.-D., Yan, Y., Fonseca, M.M.
+\& Katz, L.
+(2019).
 PhyloToL: A taxon/gene-rich phylogenomic pipeline to explore genome evolution of diverse eukaryotes.
-Molecular biology and evolution.
-36:1831--1842.
+\emph{Molecular biology and evolution}, \textbf{36}, 1831--1842.
 
 \leavevmode\hypertarget{ref-chen2008phylofinder}{}%
-Chen D., Burleigh J.G., Bansal M.S., Fernández-Baca D. 2008. PhyloFinder: An intelligent search engine for phylogenetic tree databases. BMC Evolutionary Biology. 8:90.
+Chen, D., Burleigh, J.G., Bansal, M.S. \& Fernández-Baca, D. (2008). PhyloFinder: An intelligent search engine for phylogenetic tree databases. \emph{BMC Evolutionary Biology}, \textbf{8}, 90.
 
 \leavevmode\hypertarget{ref-chen2016tree}{}%
-Chen Z.-D., Yang T., Lin L., Lu L.-M., Li H.-L., Sun M., Liu B., Chen M., Niu Y.-T., Ye J.-F., others.
-2016.
+Chen, Z.-D., Yang, T., Lin, L., Lu, L.-M., Li, H.-L., Sun, M., Liu, B., Chen, M., Niu, Y.-T., Ye, J.-F.
+\& others.
+(2016).
 Tree of life for the genera of chinese vascular plants.
-Journal of Systematics and Evolution.
-54:277--306.
+\emph{Journal of Systematics and Evolution}, \textbf{54}, 277--306.
 
 \leavevmode\hypertarget{ref-chesters2017construction}{}%
-Chesters D. 2017. Construction of a species-level tree of life for the insects and utility in taxonomic profiling. Systematic biology. 66:426--439.
+Chesters, D. (2017). Construction of a species-level tree of life for the insects and utility in taxonomic profiling. \emph{Systematic biology}, \textbf{66}, 426--439.
 
 \leavevmode\hypertarget{ref-chesters2019phylogeny}{}%
-Chesters D. 2019. The phylogeny of insects in the data-driven era. Systematic Entomology.
+Chesters, D. (2019). The phylogeny of insects in the data-driven era. \emph{Systematic Entomology}.
 
 \leavevmode\hypertarget{ref-chesters2013resolving}{}%
-Chesters D., Vogler A.P. 2013. Resolving ambiguity of species limits and concatenation in multilocus sequence data for the construction of phylogenetic supermatrices. Systematic Biology. 62:456--466.
+Chesters, D.
+\& Vogler, A.P.
+(2013).
+Resolving ambiguity of species limits and concatenation in multilocus sequence data for the construction of phylogenetic supermatrices.
+\emph{Systematic Biology}, \textbf{62}, 456--466.
 
 \leavevmode\hypertarget{ref-chesters2014protocol}{}%
-Chesters D., Zhu C.-D. 2014. A protocol for species delineation of public dna databases, applied to the insecta. Systematic biology. 63:712--725.
+Chesters, D. \& Zhu, C.-D. (2014). A protocol for species delineation of public dna databases, applied to the insecta. \emph{Systematic biology}, \textbf{63}, 712--725.
 
 \leavevmode\hypertarget{ref-crawford2012more}{}%
-Crawford N.G., Faircloth B.C., McCormack J.E., Brumfield R.T., Winker K., Glenn T.C.
-2012.
+Crawford, N.G., Faircloth, B.C., McCormack, J.E., Brumfield, R.T., Winker, K.
+\& Glenn, T.C.
+(2012).
 More than 1000 ultraconserved elements provide evidence that turtles are the sister group of archosaurs.
-Biology letters.
-8:783--786.
+\emph{Biology letters}, \textbf{8}, 783--786.
 
 \leavevmode\hypertarget{ref-crete2012salmonidae}{}%
-Crête-Lafrenière A., Weir L.K., Bernatchez L. 2012. Framing the salmonidae family phylogenetic portrait: A more complete picture from increased taxon sampling. PloS one. 7.
+Crête-Lafrenière, A., Weir, L.K. \& Bernatchez, L. (2012). Framing the salmonidae family phylogenetic portrait: A more complete picture from increased taxon sampling. \emph{PloS one}, \textbf{7}.
 
 \leavevmode\hypertarget{ref-day1984efficient}{}%
-Day W.H., Edelsbrunner H. 1984. Efficient algorithms for agglomerative hierarchical clustering methods. Journal of classification. 1:7--24.
+Day, W.H. \& Edelsbrunner, H. (1984). Efficient algorithms for agglomerative hierarchical clustering methods. \emph{Journal of classification}, \textbf{1}, 7--24.
+
+\leavevmode\hypertarget{ref-deepak2013managing}{}%
+Deepak, A. (2013). \emph{Managing and analyzing phylogenetic databases}. PhD thesis thesis, Retrieved from \url{https://lib.dr.iastate.edu/etd/12995}
 
 \leavevmode\hypertarget{ref-deepak2010searchtree}{}%
-Deepak A. 2010. SearchTree: Mining robust phylogenetic trees..
-
-\leavevmode\hypertarget{ref-deepak2013managing}{}%
-Deepak A. 2013. Managing and analyzing phylogenetic databases..
+Deepak, A. (2010). SearchTree: Mining robust phylogenetic trees.
 
 \leavevmode\hypertarget{ref-deepak2013extracting}{}%
-Deepak A., Fernández-Baca D., McMahon M.M. 2013. Extracting conflict-free information from multi-labeled trees. Algorithms for Molecular Biology. 8:18.
+Deepak, A., Fernández-Baca, D. \& McMahon, M.M. (2013). Extracting conflict-free information from multi-labeled trees. \emph{Algorithms for Molecular Biology}, \textbf{8}, 18.
 
 \leavevmode\hypertarget{ref-deepak2014evominer}{}%
-Deepak A., Fernández-Baca D., Tirthapura S., Sanderson M.J., McMahon M.M. 2014. EvoMiner: Frequent subtree mining in phylogenetic databases. Knowledge and Information Systems. 41:559--590.
+Deepak, A., Fernández-Baca, D., Tirthapura, S., Sanderson, M.J.
+\& McMahon, M.M.
+(2014).
+EvoMiner: Frequent subtree mining in phylogenetic databases.
+\emph{Knowledge and Information Systems}, \textbf{41}, 559--590.
 
 \leavevmode\hypertarget{ref-drori2018onetwotree}{}%
-Drori M., Rice A., Einhorn M., Chay O., Glick L., Mayrose I. 2018. OneTwoTree: An online tool for phylogeny reconstruction. Molecular ecology resources. 18:1492--1499.
+Drori, M., Rice, A., Einhorn, M., Chay, O., Glick, L. \& Mayrose, I. (2018). OneTwoTree: An online tool for phylogeny reconstruction. \emph{Molecular ecology resources}, \textbf{18}, 1492--1499.
+
+\leavevmode\hypertarget{ref-dufort2016augmented}{}%
+Dufort, M.J. (2016). An augmented supermatrix phylogeny of the avian family picidae reveals uncertainty deep in the family tree. \emph{Molecular phylogenetics and evolution}, \textbf{94}, 313--326.
 
 \leavevmode\hypertarget{ref-dufort2015coexistence}{}%
-Dufort M. 2015. Coexistence, ecomorphology, and diversification in the avian family picidae (woodpeckers and allies)..
-
-\leavevmode\hypertarget{ref-dufort2016augmented}{}%
-Dufort M.J. 2016. An augmented supermatrix phylogeny of the avian family picidae reveals uncertainty deep in the family tree. Molecular phylogenetics and evolution. 94:313--326.
+Dufort, M.
+(2015).
+\emph{Coexistence, ecomorphology, and diversification in the avian family picidae (woodpeckers and allies)}.
+PhD thesis thesis, University of Minnesota.
+Retrieved from \url{http://hdl.handle.net/11299/175702}
 
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
-Edgar R.C. 2004. MUSCLE: Multiple sequence alignment with high accuracy and high throughput. Nucleic acids research. 32:1792--1797.
+Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic acids research}, \textbf{32}, 1792--1797.
 
 \leavevmode\hypertarget{ref-evans2019bony}{}%
-Evans K.M., Vidal-García M., Tagliacollo V.A., Taylor S.J., Fenolio D.B.
-2019.
+Evans, K.M., Vidal-García, M., Tagliacollo, V.A., Taylor, S.J.
+\& Fenolio, D.B.
+(2019).
 Bony patchwork: Mosaic patterns of evolution in the skull of electric fishes (apteronotidae: Gymnotiformes).
-Integrative and comparative biology.
-59:420--431.
+\emph{Integrative and comparative biology}, \textbf{59}, 420--431.
+
+\leavevmode\hypertarget{ref-fang2019physpetree}{}%
+Fang, Y., Liu, C., Lin, J., Li, X., Alavian, K.N., Yang, Y.
+\& Niu, Y.
+(2019).
+PhySpeTree: An automated pipeline for reconstructing phylogenetic species trees.
+\emph{BMC evolutionary biology}, \textbf{19}, 1--8.
 
 \leavevmode\hypertarget{ref-fan2015assembly}{}%
-Fan H., Ives A.R., Surget-Groba Y., Cannon C.H. 2015. An assembly and alignment-free method of phylogeny reconstruction from next-generation sequencing data. BMC genomics. 16:522.
-
-\leavevmode\hypertarget{ref-fang2019physpetree}{}%
-Fang Y., Liu C., Lin J., Li X., Alavian K.N., Yang Y., Niu Y. 2019. PhySpeTree: An automated pipeline for reconstructing phylogenetic species trees. BMC evolutionary biology. 19:1--8.
+Fan, H., Ives, A.R., Surget-Groba, Y.
+\& Cannon, C.H.
+(2015).
+An assembly and alignment-free method of phylogeny reconstruction from next-generation sequencing data.
+\emph{BMC genomics}, \textbf{16}, 522.
 
 \leavevmode\hypertarget{ref-faurby2016all}{}%
-Faurby S., Eiserhardt W.L., Baker W.J., Svenning J.-C. 2016. An all-evidence species-level supertree for the palms (arecaceae). Molecular Phylogenetics and Evolution. 100:57--69.
+Faurby, S., Eiserhardt, W.L., Baker, W.J.
+\& Svenning, J.-C.
+(2016).
+An all-evidence species-level supertree for the palms (arecaceae).
+\emph{Molecular Phylogenetics and Evolution}, \textbf{100}, 57--69.
 
 \leavevmode\hypertarget{ref-faurby2015species}{}%
-Faurby S., Svenning J.-C.
-2015.
+Faurby, S.
+\& Svenning, J.-C.
+(2015).
 A species-level phylogeny of all extant and late quaternary extinct mammals using a novel heuristic-hierarchical bayesian approach.
-Molecular phylogenetics and evolution.
-84:14--26.
+\emph{Molecular phylogenetics and evolution}, \textbf{84}, 14--26.
 
 \leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
-Felsenstein J. 1985. Confidence intervals on phylogenetics: An approach using bootstrap. Evolution. 39:783--791.
+Felsenstein, J. (1985). Confidence intervals on phylogenetics: An approach using bootstrap. \emph{Evolution}, \textbf{39}, 783--791.
 
 \leavevmode\hypertarget{ref-flores2011estimating}{}%
-Flores O., Coomes D.A. 2011. Estimating the wood density of species for carbon stock assessments. Methods in Ecology and Evolution. 2:214--220.
+Flores, O. \& Coomes, D.A. (2011). Estimating the wood density of species for carbon stock assessments. \emph{Methods in Ecology and Evolution}, \textbf{2}, 214--220.
 
 \leavevmode\hypertarget{ref-frazao2015gondwana}{}%
-Frazao A., Silva H.R.
-da, Moraes Russo C.A.
+Frazao, A., Silva, H.R.
+da \& Moraes Russo, C.A.
 de.
-2015.
+(2015).
 The gondwana breakup and the history of the atlantic and indian oceans unveils two new clades for early neobatrachian diversification.
-PloS one.
-10.
+\emph{PloS one}, \textbf{10}.
+
+\leavevmode\hypertarget{ref-freyman2017phylogenetic}{}%
+Freyman, W.A.
+(2017).
+\emph{Phylogenetic models linking speciation and extinction to chromosome and mating system evolution}.
+PhD thesis thesis, UC Berkeley.
+Retrieved from \url{https://escholarship.org/uc/item/29n8r0nm}
 
 \leavevmode\hypertarget{ref-freyman2015sumac}{}%
-Freyman W.A. 2015. SUMAC: Constructing phylogenetic supermatrices and assessing partially decisive taxon coverage. Evolutionary Bioinformatics. 11:EBO--S35384.
-
-\leavevmode\hypertarget{ref-freyman2017phylogenetic}{}%
-Freyman W.A. 2017. Phylogenetic models linking speciation and extinction to chromosome and mating system evolution..
+Freyman, W.A. (2015). SUMAC: Constructing phylogenetic supermatrices and assessing partially decisive taxon coverage. \emph{Evolutionary Bioinformatics}, \textbf{11}, EBO--S35384.
 
 \leavevmode\hypertarget{ref-gagnon2016new}{}%
-Gagnon E., Bruneau A., Hughes C.E., Queiroz L.P. de, Lewis G.P. 2016. A new generic system for the pantropical caesalpinia group (leguminosae). PhytoKeys.:1.
+Gagnon, E., Bruneau, A., Hughes, C.E., Queiroz, L.P. de \& Lewis, G.P. (2016). A new generic system for the pantropical caesalpinia group (leguminosae). \emph{PhytoKeys}, 1.
 
 \leavevmode\hypertarget{ref-gagnon2016systematique}{}%
-Gagnon E., others. 2016. Systématique et biogéographie du groupe caesalpinia (famille leguminosae)..
+Gagnon, E. \& others. (2016). Systématique et biogéographie du groupe caesalpinia (famille leguminosae). Retrieved from \url{http://hdl.handle.net/1866/13587}
 
 \leavevmode\hypertarget{ref-gao2011solution}{}%
-Gao Y., Meng Z., He X., Liu Y., Zhou Y., Li J. 2011. A solution to integrate data for phylogenetic research. 2011 5th international conference on bioinformatics and biomedical engineering.:1--4.
+Gao, Y., Meng, Z., He, X., Liu, Y., Zhou, Y.
+\& Li, J.
+(2011).
+A solution to integrate data for phylogenetic research.
+\emph{2011 5th international conference on bioinformatics and biomedical engineering} pp.
+1--4.
+IEEE.
 
 \leavevmode\hypertarget{ref-garcia2016role}{}%
-García-Navas V., Bonnet T., Bonal R., Postma E.
-2016.
+García-Navas, V., Bonnet, T., Bonal, R.
+\& Postma, E.
+(2016).
 The role of fecundity and sexual selection in the evolution of size and sexual size dimorphism in new world and old world voles (rodentia: Arvicolinae).
-Oikos.
-125:1250--1260.
+\emph{Oikos}, \textbf{125}, 1250--1260.
 
 \leavevmode\hypertarget{ref-gomez2020speciation}{}%
-Gómez-Bahamón V., Márquez R., Jahn A.E., Miyaki C.Y., Tuero D.T., Laverde-R O., Restrepo S., Cadena C.D.
-2020.
+Gómez-Bahamón, V., Márquez, R., Jahn, A.E., Miyaki, C.Y., Tuero, D.T., Laverde-R, O., Restrepo, S.
+\& Cadena, C.D.
+(2020).
 Speciation associated with shifts in migratory behavior in an avian radiation.
-Current Biology.
+\emph{Current Biology}.
 
 \leavevmode\hypertarget{ref-gomez2015behavioral}{}%
-Gómez Bahamón V., others. 2015. A behavioral polymorphism as an intermediate stage in the evolution of divergent forms-partial migration in new world flycatchers (aves, tyrannidae)..
+Gómez Bahamón, V.
+\& others.
+(2015).
+\emph{A behavioral polymorphism as an intermediate stage in the evolution of divergent forms-partial migration in new world flycatchers (aves, tyrannidae)}.
+Master's thesis thesis, Bogotá-Uniandes.
+Retrieved from \url{https://repositorio.uniandes.edu.co/bitstream/handle/1992/12859/u703694.pdf?sequence=1}
 
 \leavevmode\hypertarget{ref-grant2014building}{}%
-Grant J.R., Katz L.A. 2014. Building a phylogenomic pipeline for the eukaryotic tree of life-addressing deep phylogenies with genome-scale data. PLoS currents. 6.
+Grant, J.R. \& Katz, L.A. (2014). Building a phylogenomic pipeline for the eukaryotic tree of life-addressing deep phylogenies with genome-scale data. \emph{PLoS currents}, \textbf{6}.
 
 \leavevmode\hypertarget{ref-greenberg2011caryophyllaceae}{}%
-Greenberg A.K., Donoghue M.J. 2011. Molecular systematics and character evolution in caryophyllaceae. Taxon. 60:1637--1652.
+Greenberg, A.K. \& Donoghue, M.J. (2011). Molecular systematics and character evolution in caryophyllaceae. \emph{Taxon}, \textbf{60}, 1637--1652.
 
 \leavevmode\hypertarget{ref-legume2013legume}{}%
-Group L.P.W., Bruneau A., Doyle J.J., Herendeen P., Hughes C., Kenicer G., Lewis G., Mackinder B., Pennington R.T., Sanderson M.J., others.
-2013.
+Group, L.P.W., Bruneau, A., Doyle, J.J., Herendeen, P., Hughes, C., Kenicer, G., Lewis, G., Mackinder, B., Pennington, R.T., Sanderson, M.J.
+\& others.
+(2013).
 Legume phylogeny and classification in the 21st century: Progress, prospects and lessons for other species--rich clades.
-Taxon.
-62:217--248.
+\emph{Taxon}, \textbf{62}, 217--248.
 
 \leavevmode\hypertarget{ref-guy2017phyloskeleton}{}%
-Guy L. 2017. PhyloSkeleton: Taxon selection, data retrieval and marker identification for phylogenomics. Bioinformatics. 33:1230--1232.
+Guy, L. (2017). PhyloSkeleton: Taxon selection, data retrieval and marker identification for phylogenomics. \emph{Bioinformatics}, \textbf{33}, 1230--1232.
 
 \leavevmode\hypertarget{ref-hardy2010gall}{}%
-Hardy N.B., Cook L.G. 2010. Gall-induction in insects: Evolutionary dead-end or speciation driver? BMC evolutionary biology. 10:257.
+Hardy, N.B. \& Cook, L.G. (2010). Gall-induction in insects: Evolutionary dead-end or speciation driver? \emph{BMC evolutionary biology}, \textbf{10}, 257.
 
 \leavevmode\hypertarget{ref-hardy2012testing}{}%
-Hardy N.B., Cook L.G. 2012. Testing for ecological limitation of diversification: A case study using parasitic plants. The American Naturalist. 180:438--449.
+Hardy, N.B. \& Cook, L.G. (2012). Testing for ecological limitation of diversification: A case study using parasitic plants. \emph{The American Naturalist}, \textbf{180}, 438--449.
 
 \leavevmode\hypertarget{ref-hardy2014specialization}{}%
-Hardy N.B., Otto S.P.
-2014.
+Hardy, N.B.
+\& Otto, S.P.
+(2014).
 Specialization and generalization in the diversification of phytophagous insects: Tests of the musical chairs and oscillation hypotheses.
-Proceedings of the Royal Society B: Biological Sciences.
-281:20132960.
+\emph{Proceedings of the Royal Society B: Biological Sciences}, \textbf{281}, 20132960.
 
 \leavevmode\hypertarget{ref-hardy2015evolution}{}%
-Hardy N.B., Peterson D.A., Dohlen C.D. von. 2015. The evolution of life cycle complexity in aphids: Ecological optimization or historical constraint? Evolution. 69:1423--1432.
+Hardy, N.B., Peterson, D.A. \& Dohlen, C.D. von. (2015). The evolution of life cycle complexity in aphids: Ecological optimization or historical constraint? \emph{Evolution}, \textbf{69}, 1423--1432.
 
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
-Helmus M.R., Ives A.R. 2012. Phylogenetic diversity--area curves. Ecology. 93:S31--S43.
+Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
 
 \leavevmode\hypertarget{ref-helmus2010communities}{}%
-Helmus M.R., Keller W., Paterson M.J., Yan N.D., Cannon C.H., Rusak J.A. 2010. Communities contain closely related species during ecosystem disturbance. Ecology letters. 13:162--174.
+Helmus, M.R., Keller, W., Paterson, M.J., Yan, N.D., Cannon, C.H.
+\& Rusak, J.A.
+(2010).
+Communities contain closely related species during ecosystem disturbance.
+\emph{Ecology letters}, \textbf{13}, 162--174.
 
 \leavevmode\hypertarget{ref-ickert2009fossil}{}%
-Ickert-Bond S.M., Rydin C., Renner S.S.
-2009.
+Ickert-Bond, S.M., Rydin, C.
+\& Renner, S.S.
+(2009).
 A fossil-calibrated relaxed clock for ephedra indicates an oligocene age for the divergence of asian and new world clades and miocene dispersal into south america.
-Journal of Systematics and Evolution.
-47:444--456.
+\emph{Journal of Systematics and Evolution}, \textbf{47}, 444--456.
 
 \leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco F., Cazes J., Smith S.A., Stamatakis A. 2014. PUmPER: Phylogenies updated perpetually. Bioinformatics. 30:1476--1477.
+Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
 
 \leavevmode\hypertarget{ref-jamil2016visual}{}%
-Jamil H.M. 2016. A visual interface for querying heterogeneous phylogenetic databases. IEEE/ACM transactions on computational biology and bioinformatics. 14:131--144.
+Jamil, H.M. (2016). A visual interface for querying heterogeneous phylogenetic databases. \emph{IEEE/ACM transactions on computational biology and bioinformatics}, \textbf{14}, 131--144.
 
 \leavevmode\hypertarget{ref-jones2014trends}{}%
-Jones T.M., Baxter D.G., Hagedorn G., Legler B., Gilbert E., Thiele K., Vargas-Rodriguez Y., Urbatsch L.E.
-2014.
+Jones, T.M., Baxter, D.G., Hagedorn, G., Legler, B., Gilbert, E., Thiele, K., Vargas-Rodriguez, Y.
+\& Urbatsch, L.E.
+(2014).
 Trends in access of plant biodiversity data revealed by google analytics.
-Biodiversity data journal.
+\emph{Biodiversity data journal}.
 
 \leavevmode\hypertarget{ref-kamarudin2016phylogenetic}{}%
-Kamarudin K.R., Rehan A.M., Hashim R., Usup G., Rehan M.M.
-2016.
+Kamarudin, K.R., Rehan, A.M., Hashim, R., Usup, G.
+\& Rehan, M.M.
+(2016).
 Phylogenetic relationships within the genus holothuria inferred from 16S mitochondiral rRNA gene sequences.
-Sains Malaysiana.
-45:1079--1087.
+\emph{Sains Malaysiana}, \textbf{45}, 1079--1087.
+Retrieved from \url{http://journalarticle.ukm.my/9982/1/10\%20Kamarul\%20Rahim\%20Kamarudin.pdf}
 
 \leavevmode\hypertarget{ref-kolmann2017dna}{}%
-Kolmann M.A., Elbassiouny A.A., Liverpool E.A., Lovejoy N.R. 2017. DNA barcoding reveals the diversity of sharks in guyana coastal markets. Neotropical Ichthyology. 15.
+Kolmann, M.A., Elbassiouny, A.A., Liverpool, E.A. \& Lovejoy, N.R. (2017). DNA barcoding reveals the diversity of sharks in guyana coastal markets. \emph{Neotropical Ichthyology}, \textbf{15}.
 
 \leavevmode\hypertarget{ref-kumar2015bir}{}%
-Kumar S., Krabberød A.K., Neumann R.S., Michalickova K., Zhao S., Zhang X., Shalchian-Tabrizi K. 2015. BIR pipeline for preparation of phylogenomic data. Evolutionary Bioinformatics. 11:EBO--S10189.
+Kumar, S., Krabberød, A.K., Neumann, R.S., Michalickova, K., Zhao, S., Zhang, X.
+\& Shalchian-Tabrizi, K.
+(2015).
+BIR pipeline for preparation of phylogenomic data.
+\emph{Evolutionary Bioinformatics}, \textbf{11}, EBO--S10189.
 
 \leavevmode\hypertarget{ref-lehtonen2011towards}{}%
-Lehtonen S. 2011. Towards resolving the complete fern tree of life. PLoS One. 6.
+Lehtonen, S. (2011). Towards resolving the complete fern tree of life. \emph{PLoS One}, \textbf{6}.
 
 \leavevmode\hypertarget{ref-lehtonen2016sensitive}{}%
-Lehtonen S., Christenhusz M.J., Falck D. 2016. Sensitive phylogenetics of clematis and its position in ranunculaceae. Botanical Journal of the Linnean Society. 182:825--867.
+Lehtonen, S., Christenhusz, M.J. \& Falck, D. (2016). Sensitive phylogenetics of clematis and its position in ranunculaceae. \emph{Botanical Journal of the Linnean Society}, \textbf{182}, 825--867.
 
 \leavevmode\hypertarget{ref-lehtonen2017environmentally}{}%
-Lehtonen S., Silvestro D., Karger D.N., Scotese C., Tuomisto H., Kessler M., Peña C., Wahlberg N., Antonelli A.
-2017.
+Lehtonen, S., Silvestro, D., Karger, D.N., Scotese, C., Tuomisto, H., Kessler, M., Peña, C., Wahlberg, N.
+\& Antonelli, A.
+(2017).
 Environmentally driven extinction and opportunistic origination explain fern diversification patterns.
-Scientific Reports.
-7:1--12.
+\emph{Scientific Reports}, \textbf{7}, 1--12.
 
 \leavevmode\hypertarget{ref-lemoine2018renewing}{}%
-Lemoine F., Entfellner J.-B.D., Wilkinson E., Correia D., Felipe M.D., De Oliveira T., Gascuel O. 2018. Renewing felsenstein's phylogenetic bootstrap in the era of big data. Nature. 556:452--456.
+Lemoine, F., Entfellner, J.-B.D., Wilkinson, E., Correia, D., Felipe, M.D., De Oliveira, T.
+\& Gascuel, O.
+(2018).
+Renewing felsenstein's phylogenetic bootstrap in the era of big data.
+\emph{Nature}, \textbf{556}, 452--456.
 
 \leavevmode\hypertarget{ref-li2013partfasttree}{}%
-Li J., Meng Z., Hou Y., Zhou Y., Gao Y.
-2013.
+Li, J., Meng, Z., Hou, Y., Zhou, Y.
+\& Gao, Y.
+(2013).
 PartFastTree: Constructing large phylogenetic trees and estimating their reliability.
-2013 ninth international conference on natural computation (icnc).:1052--1056.
+\emph{2013 ninth international conference on natural computation (icnc)} pp.
+1052--1056.
+IEEE.
 
 \leavevmode\hypertarget{ref-mahmood2015avian}{}%
-Mahmood M.T. 2015. Avian raptor evolution..
+Mahmood, M.T.
+(2015).
+\emph{Avian raptor evolution}.
+PhD thesis thesis, Institute of Fundamental Sciences, Massey University, New Zealand.
+Retrieved from \url{https://mro.massey.ac.nz/bitstream/handle/10179/7198/02_whole.pdf}
 
 \leavevmode\hypertarget{ref-martin2011trophic}{}%
-Martin C.H., Wainwright P.C.
-2011.
+Martin, C.H.
+\& Wainwright, P.C.
+(2011).
 Trophic novelty is linked to exceptional rates of morphological diversification in two adaptive radiations of cyprinodon pupfish.
-Evolution: International Journal of Organic Evolution.
-65:2197--2212.
+\emph{Evolution: International Journal of Organic Evolution}, \textbf{65}, 2197--2212.
 
 \leavevmode\hypertarget{ref-mcmahon2015stbase}{}%
-McMahon M.M., Deepak A., Fernández-Baca D., Boss D., Sanderson M.J. 2015. STBase: One million species trees for comparative biology. PloS one. 10.
+McMahon, M.M., Deepak, A., Fernández-Baca, D., Boss, D. \& Sanderson, M.J. (2015). STBase: One million species trees for comparative biology. \emph{PloS one}, \textbf{10}.
 
 \leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
-McTavish E.J., Hinchliff C.E., Allman J.F., Brown J.W., Cranston K.A., Holder M.T., Rees J.A., Smith S.A.
-2015.
+McTavish, E.J., Hinchliff, C.E., Allman, J.F., Brown, J.W., Cranston, K.A., Holder, M.T., Rees, J.A.
+\& Smith, S.A.
+(2015).
 Phylesystem: A git-based data store for community-curated phylogenetic estimates.
-Bioinformatics.
-31:2794--2800.
+\emph{Bioinformatics}, \textbf{31}, 2794--2800.
 
 \leavevmode\hypertarget{ref-meng2015darwintree}{}%
-Meng Z., Dong H., Li J., Chen Z., Zhou Y., Wang X., Zhang S. 2015a. Darwintree: A molecular data analysis and application environment for phylogenetic study. Data Science Journal. 14.
+Meng, Z., Dong, H., Li, J., Chen, Z., Zhou, Y., Wang, X.
+\& Zhang, S.
+(2015a).
+Darwintree: A molecular data analysis and application environment for phylogenetic study.
+\emph{Data Science Journal}, \textbf{14}.
 
 \leavevmode\hypertarget{ref-meng2015solution}{}%
-Meng Z., Li J., Chen Z. 2015b. A solution to phylogeny assembly for ecologists. 2015 12th international conference on fuzzy systems and knowledge discovery (fskd).:1103--1107.
+Meng, Z., Li, J. \& Chen, Z. (2015b). A solution to phylogeny assembly for ecologists. \emph{2015 12th international conference on fuzzy systems and knowledge discovery (fskd)} pp. 1103--1107. IEEE.
 
 \leavevmode\hypertarget{ref-meng2015sotree}{}%
-Meng Z., Li J., Yang T., Lin L., Chen Z.
-2015c.
+Meng, Z., Li, J., Yang, T., Lin, L.
+\& Chen, Z.
+(2015c).
 SoTree: An automated phylogeny assembly tool for ecologists from big tree.
-2015 ieee international conference on smart city/socialcom/sustaincom (smartcity).:792--797.
+\emph{2015 ieee international conference on smart city/socialcom/sustaincom (smartcity)} pp.
+792--797.
+IEEE.
 
 \leavevmode\hypertarget{ref-meng2012gsqct}{}%
-Meng Z., Li J., Zhou Y., Cao W., Xiao X., Zhao J., Dong H., Zhang S.
-2012a.
+Meng, Z., Li, J., Zhou, Y., Cao, W., Xiao, X., Zhao, J., Dong, H.
+\& Zhang, S.
+(2012a).
 GSQCT: A solution to screening gene sequences for phylogenetics analysis.
-2012 9th international conference on fuzzy systems and knowledge discovery.:2929--2933.
+\emph{2012 9th international conference on fuzzy systems and knowledge discovery} pp.
+2929--2933.
+IEEE.
 
 \leavevmode\hypertarget{ref-meng2014rapidtree}{}%
-Meng Z., Shao J., Cao W., Li J., Zhou Y., Wang X.
-2014.
+Meng, Z., Shao, J., Cao, W., Li, J., Zhou, Y.
+\& Wang, X.
+(2014).
 RapidTree: A solution to rapid reconstruction phylogenetic tree.
-2014 11th international conference on fuzzy systems and knowledge discovery (fskd).:513--517.
+\emph{2014 11th international conference on fuzzy systems and knowledge discovery (fskd)} pp.
+513--517.
+IEEE.
 
 \leavevmode\hypertarget{ref-meng2012cloud}{}%
-Meng Z., Xiao X., Li J., Zhou Y., Cao W., Shen G.
-2012b.
+Meng, Z., Xiao, X., Li, J., Zhou, Y., Cao, W.
+\& Shen, G.
+(2012b).
 Cloud-gsqct: A parallel approach to screen gene sequences for phylogenetics analysis.
-2012 international conference on computer science and information processing (csip).:660--663.
+\emph{2012 international conference on computer science and information processing (csip)} pp.
+660--663.
+IEEE.
 
 \leavevmode\hypertarget{ref-morrison2009apicomplexa}{}%
-Morrison D.A. 2009. Evolution of the apicomplexa: Where are we now? Trends in parasitology. 25:375--382.
+Morrison, D.A. (2009). Evolution of the apicomplexa: Where are we now? \emph{Trends in parasitology}, \textbf{25}, 375--382.
 
 \leavevmode\hypertarget{ref-nazaire2012broad}{}%
-Nazaire M., Hufford L. 2012. A broad phylogenetic analysis of boraginaceae: Implications for the relationships of mertensia. Systematic Botany. 37:758--783.
+Nazaire, M. \& Hufford, L. (2012). A broad phylogenetic analysis of boraginaceae: Implications for the relationships of mertensia. \emph{Systematic Botany}, \textbf{37}, 758--783.
+
+\leavevmode\hypertarget{ref-page2013bionames}{}%
+Page, R.D. (2013). BioNames: Linking taxonomy, texts, and trees. \emph{PeerJ}, \textbf{1}, e190.
 
 \leavevmode\hypertarget{ref-page2011linking}{}%
-Page R.D. 2011. Linking ncbi to wikipedia: A wiki-based approach. PLoS currents. 3.
-
-\leavevmode\hypertarget{ref-page2013bionames}{}%
-Page R.D. 2013. BioNames: Linking taxonomy, texts, and trees. PeerJ. 1:e190.
+Page, R.D. (2011). Linking ncbi to wikipedia: A wiki-based approach. \emph{PLoS currents}, \textbf{3}.
 
 \leavevmode\hypertarget{ref-papadopoulou2015automated}{}%
-Papadopoulou A., Chesters D., Coronado I., De la Cadena G., Cardoso A., Reyes J.C., Maes J.-M., Rueda R.M., Gómez-Zurita J.
-2015.
+Papadopoulou, A., Chesters, D., Coronado, I., De la Cadena, G., Cardoso, A., Reyes, J.C., Maes, J.-M., Rueda, R.M.
+\& Gómez-Zurita, J.
+(2015).
 Automated dna-based plant identification for large-scale biodiversity assessment.
-Molecular ecology resources.
-15:136--152.
+\emph{Molecular ecology resources}, \textbf{15}, 136--152.
 
 \leavevmode\hypertarget{ref-pereira2013padroes}{}%
-Pereira J.E.S. 2013. Padrões e processos na evolução de primatas neotropicais (platyrrhini, primates)..
+Pereira, J.E.S.
+(2013).
+\emph{Padrões e processos na evolução de primatas neotropicais (platyrrhini, primates)}.
+PhD thesis thesis, Tese de doutorado.
+Setor de Ciências Biológicas, Universidade Federal do~\ldots{}.
+Retrieved from \url{https://www.acervodigital.ufpr.br/handle/1884/33775}
 
 \leavevmode\hypertarget{ref-peters2011taming}{}%
-Peters R.S., Meyer B., Krogmann L., Borner J., Meusemann K., Schütte K., Niehuis O., Misof B.
-2011.
+Peters, R.S., Meyer, B., Krogmann, L., Borner, J., Meusemann, K., Schütte, K., Niehuis, O.
+\& Misof, B.
+(2011).
 The taming of an impossible child: A standardized all-in approach to the phylogeny of hymenoptera using public database sequences.
-BMC biology.
-9:55.
+\emph{BMC biology}, \textbf{9}, 55.
+Retrieved from \url{https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55\#Sec21}
 
 \leavevmode\hypertarget{ref-piel2009treebase}{}%
-Piel W., Chan L., Dominus M., Ruan J., Vos R., Tannen V. 2009. Treebase v. 2: A database of phylogenetic knowledge. E-biosphere..
+Piel, W., Chan, L., Dominus, M., Ruan, J., Vos, R. \& Tannen, V. (2009). Treebase v. 2: A database of phylogenetic knowledge. E-biosphere.
 
 \leavevmode\hypertarget{ref-rabosky2011species}{}%
-Rabosky D.L., Cowan M.A., Talaba A.L., Lovette I.J.
-2011.
+Rabosky, D.L., Cowan, M.A., Talaba, A.L.
+\& Lovette, I.J.
+(2011).
 Species interactions mediate phylogenetic community structure in a hyperdiverse lizard assemblage from arid australia.
-The American Naturalist.
-178:579--595.
+\emph{The American Naturalist}, \textbf{178}, 579--595.
 
 \leavevmode\hypertarget{ref-ranwez2009phyloexplorer}{}%
-Ranwez V., Clairon N., Delsuc F., Pourali S., Auberval N., Diser S., Berry V. 2009. PhyloExplorer: A web server to validate, explore and query phylogenetic trees. BMC evolutionary biology. 9:108.
+Ranwez, V., Clairon, N., Delsuc, F., Pourali, S., Auberval, N., Diser, S.
+\& Berry, V.
+(2009).
+PhyloExplorer: A web server to validate, explore and query phylogenetic trees.
+\emph{BMC evolutionary biology}, \textbf{9}, 108.
 
 \leavevmode\hypertarget{ref-roalson2016distinct}{}%
-Roalson E.H., Roberts W.R. 2016. Distinct processes drive diversification in different clades of gesneriaceae. Systematic Biology. 65:662--684.
+Roalson, E.H. \& Roberts, W.R. (2016). Distinct processes drive diversification in different clades of gesneriaceae. \emph{Systematic Biology}, \textbf{65}, 662--684.
 
 \leavevmode\hypertarget{ref-roquet2013building}{}%
-Roquet C., Thuiller W., Lavergne S. 2013. Building megaphylogenies for macroecology: Taking up the challenge. Ecography. 36:13--26.
+Roquet, C., Thuiller, W. \& Lavergne, S. (2013). Building megaphylogenies for macroecology: Taking up the challenge. \emph{Ecography}, \textbf{36}, 13--26.
 
 \leavevmode\hypertarget{ref-ross2013large}{}%
-Ross L., Hardy N.B., Okusu A., Normark B.B. 2013. Large population size predicts the distribution of asexuality in scale insects. Evolution: International Journal of Organic Evolution. 67:196--206.
+Ross, L., Hardy, N.B., Okusu, A.
+\& Normark, B.B.
+(2013).
+Large population size predicts the distribution of asexuality in scale insects.
+\emph{Evolution: International Journal of Organic Evolution}, \textbf{67}, 196--206.
 
 \leavevmode\hypertarget{ref-ruiz2019datataxa}{}%
-Ruiz-Sanchez E., Maya-Lastra C.A., Steinmann V.W., Zamudio S., Carranza E., Murillo R.M., Rzedowski J.
-2019.
+Ruiz-Sanchez, E., Maya-Lastra, C.A., Steinmann, V.W., Zamudio, S., Carranza, E., Murillo, R.M.
+\& Rzedowski, J.
+(2019).
 Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajío as a case study.
-Botanical Sciences.
-97:754--760.
+\emph{Botanical Sciences}, \textbf{97}, 754--760.
 
 \leavevmode\hypertarget{ref-salariato2010filogenia}{}%
-Salariato D.L. 2010. Filogenia y evolución de la subtribu melinidinae (paniceae: Panicoideae: Poaceae)..
+Salariato, D.L.
+(2010).
+\emph{Filogenia y evolución de la subtribu melinidinae (paniceae: Panicoideae: Poaceae)}.
+PhD thesis thesis, Universidad de Buenos Aires.
+Facultad de Ciencias Exactas y Naturales.
+Retrieved from \url{http://hdl.handle.net/20.500.12110/tesis_n4771_Salariato}
 
 \leavevmode\hypertarget{ref-salariato2010molecular}{}%
-Salariato D.L., Zuloaga F.O., Giussani L.M., Morrone O.
-2010.
+Salariato, D.L., Zuloaga, F.O., Giussani, L.M.
+\& Morrone, O.
+(2010).
 Molecular phylogeny of the subtribe melinidinae (poaceae: Panicoideae: Paniceae) and evolutionary trends in the homogenization of inflorescences.
-Molecular Phylogenetics and Evolution.
-56:355--369.
+\emph{Molecular Phylogenetics and Evolution}, \textbf{56}, 355--369.
 
 \leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson M.J., Boss D., Chen D., Cranston K.A., Wehe A. 2008. The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. Systematic Biology. 57:335--346.
+Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A.
+\& Wehe, A.
+(2008).
+The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research.
+\emph{Systematic Biology}, \textbf{57}, 335--346.
+Retrieved from \url{https://doi.org/10.1080/10635150802158688}
 
 \leavevmode\hypertarget{ref-sanderson2010phylogenomics}{}%
-Sanderson M.J., McMahon M.M., Steel M. 2010. Phylogenomics with incomplete taxon coverage: The limits to inference. BMC Evolutionary Biology. 10:155.
+Sanderson, M.J., McMahon, M.M. \& Steel, M. (2010). Phylogenomics with incomplete taxon coverage: The limits to inference. \emph{BMC Evolutionary Biology}, \textbf{10}, 155.
 
 \leavevmode\hypertarget{ref-sanderson2016perspective}{}%
-Sanderson M.J., Olson P., Hughes J., Cotton J. 2016. Perspective: Challenges in assembling the `next generation'Tree of life. Olson PD, Hughes J and Cotton JA.:13--27.
+Sanderson, M.J., Olson, P., Hughes, J.
+\& Cotton, J.
+(2016).
+Perspective: Challenges in assembling the `next generation'Tree of life.
+\emph{Olson PD, Hughes J and Cotton JA}, 13--27.
+Retrieved from \url{https://books.google.com/books?hl=es\&lr=\&id=hFU2DAAAQBAJ\&oi=fnd\&pg=PA13\&ots=2Oa_TVGCic\&sig=fdxyYmSQASm2XFoU7jv4JKLFaTQ\#v=onepage\&q\&f=false}
 
 \leavevmode\hypertarget{ref-san2010molecular}{}%
-San Mauro D., Agorreta A. 2010. Molecular systematics: A synthesis of the common methods and the state of knowledge. Cellular \& Molecular Biology Letters. 15:311.
+San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
 
 \leavevmode\hypertarget{ref-santini2014first}{}%
-Santini F., Carnevale G., Sorenson L. 2014. First multi-locus timetree of seabreams and porgies (percomorpha: Sparidae). Italian Journal of Zoology. 81:55--71.
+Santini, F., Carnevale, G. \& Sorenson, L. (2014). First multi-locus timetree of seabreams and porgies (percomorpha: Sparidae). \emph{Italian Journal of Zoology}, \textbf{81}, 55--71.
 
 \leavevmode\hypertarget{ref-santini2015first}{}%
-Santini F., Carnevale G., Sorenson L.
-2015.
+Santini, F., Carnevale, G.
+\& Sorenson, L.
+(2015).
 First timetree of sphyraenidae (percomorpha) reveals a middle eocene crown age and an oligo--miocene radiation of barracudas.
-Italian Journal of Zoology.
-82:133--142.
+\emph{Italian Journal of Zoology}, \textbf{82}, 133--142.
 
 \leavevmode\hypertarget{ref-santini2013first}{}%
-Santini F., Sorenson L. 2013. First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies. Italian journal of zoology. 80:481--489.
+Santini, F.
+\& Sorenson, L.
+(2013).
+First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies.
+\emph{Italian journal of zoology}, \textbf{80}, 481--489.
 
 \leavevmode\hypertarget{ref-sarkinen2013solanaceae}{}%
-Särkinen T., Bohs L., Olmstead R.G., Knapp S. 2013. A phylogenetic framework for evolutionary study of the nightshades (solanaceae): A dated 1000-tip tree. BMC evolutionary biology. 13:214.
+Särkinen, T., Bohs, L., Olmstead, R.G.
+\& Knapp, S.
+(2013).
+A phylogenetic framework for evolutionary study of the nightshades (solanaceae): A dated 1000-tip tree.
+\emph{BMC evolutionary biology}, \textbf{13}, 214.
 
 \leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
-Schoch C.L., Sung G.-H., López-Giráldez F., Townsend J.P., Miadlikowska J., Hofstetter V., Robbertse B., Matheny P.B., Kauff F., Wang Z., others.
-2009.
+Schoch, C.L., Sung, G.-H., López-Giráldez, F., Townsend, J.P., Miadlikowska, J., Hofstetter, V., Robbertse, B., Matheny, P.B., Kauff, F., Wang, Z.
+\& others.
+(2009).
 The ascomycota tree of life: A phylum-wide phylogeny clarifies the origin and evolution of fundamental reproductive and ecological traits.
-Systematic biology.
-58:224--239.
+\emph{Systematic biology}, \textbf{58}, 224--239.
 
 \leavevmode\hypertarget{ref-selvatti2015paleogene}{}%
-Selvatti A.P., Gonzaga L.P., Moraes Russo C.A.
+Selvatti, A.P., Gonzaga, L.P.
+\& Moraes Russo, C.A.
 de.
-2015.
+(2015).
 A paleogene origin for crown passerines and the diversification of the oscines in the new world.
-Molecular phylogenetics and evolution.
-88:1--15.
+\emph{Molecular phylogenetics and evolution}, \textbf{88}, 1--15.
 
 \leavevmode\hypertarget{ref-shi2015speciation}{}%
-Shi J.J., Rabosky D.L. 2015. Speciation dynamics during the global radiation of extant bats. Evolution. 69:1528--1545.
+Shi, J.J. \& Rabosky, D.L. (2015). Speciation dynamics during the global radiation of extant bats. \emph{Evolution}, \textbf{69}, 1528--1545.
 
 \leavevmode\hypertarget{ref-simon2012cerrado}{}%
-Simon M.F., Pennington T. 2012. Evidence for adaptation to fire regimes in the tropical savannas of the brazilian cerrado. International Journal of Plant Sciences. 173:711--723.
+Simon, M.F. \& Pennington, T. (2012). Evidence for adaptation to fire regimes in the tropical savannas of the brazilian cerrado. \emph{International Journal of Plant Sciences}, \textbf{173}, 711--723.
 
 \leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith S.A., Beaulieu J.M., Donoghue M.J. 2009. Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches. BMC evolutionary biology. 9:37.
+Smith, S.A., Beaulieu, J.M.
+\& Donoghue, M.J.
+(2009).
+Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
+\emph{BMC evolutionary biology}, \textbf{9}, 37.
 
 \leavevmode\hypertarget{ref-smith2018constructing}{}%
-Smith S.A., Brown J.W. 2018. Constructing a broadly inclusive seed plant phylogeny. American Journal of Botany. 105:302--314.
+Smith, S.A. \& Brown, J.W. (2018). Constructing a broadly inclusive seed plant phylogeny. \emph{American Journal of Botany}, \textbf{105}, 302--314.
 
 \leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
-Smith S.A., Walker J.F. 2019. PyPHLAWD: A python tool for phylogenetic dataset construction. Methods in Ecology and Evolution. 10:104--108.
+Smith, S.A. \& Walker, J.F. (2019). PyPHLAWD: A python tool for phylogenetic dataset construction. \emph{Methods in Ecology and Evolution}, \textbf{10}, 104--108.
 
 \leavevmode\hypertarget{ref-sorenson2014evolution}{}%
-Sorenson L. 2014. Evolution of marine fish biodiversity: Phylogenomics and ecological processes shaping diversification..
+Sorenson, L.
+(2014).
+\emph{Evolution of marine fish biodiversity: Phylogenomics and ecological processes shaping diversification}.
+PhD thesis thesis, UCLA.
+Retrieved from \url{https://escholarship.org/uc/item/31n0c9km}
 
 \leavevmode\hypertarget{ref-sorenson2014effect}{}%
-Sorenson L., Santini F., Alfaro M. 2014. The effect of habitat on modern shark diversification. Journal of Evolutionary Biology. 27:1536--1548.
+Sorenson, L., Santini, F. \& Alfaro, M. (2014). The effect of habitat on modern shark diversification. \emph{Journal of Evolutionary Biology}, \textbf{27}, 1536--1548.
 
 \leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
-Stamatakis A. 2014. RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. Bioinformatics. 30:1312--1313.
+Stamatakis, A. (2014). RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. \emph{Bioinformatics}, \textbf{30}, 1312--1313.
 
 \leavevmode\hypertarget{ref-steppan2017muroid}{}%
-Steppan S.J., Schenk J.J. 2017. Muroid rodent phylogenetics: 900-species tree reveals increasing diversification rates. PLoS One. 12.
+Steppan, S.J. \& Schenk, J.J. (2017). Muroid rodent phylogenetics: 900-species tree reveals increasing diversification rates. \emph{PLoS One}, \textbf{12}.
 
 \leavevmode\hypertarget{ref-stoltzfus2013phylotastic}{}%
-Stoltzfus A., Lapp H., Matasci N., Deus H., Sidlauskas B., Zmasek C.M., Vaidya G., Pontelli E., Cranston K., Vos R., others.
-2013.
+Stoltzfus, A., Lapp, H., Matasci, N., Deus, H., Sidlauskas, B., Zmasek, C.M., Vaidya, G., Pontelli, E., Cranston, K., Vos, R.
+\& others.
+(2013).
 Phylotastic!
 Making tree-of-life knowledge accessible, reusable and convenient.
-BMC bioinformatics.
-14:158.
+\emph{BMC bioinformatics}, \textbf{14}, 158.
 
 \leavevmode\hypertarget{ref-strecker2014fish}{}%
-Strecker A.L., Olden J.D.
-2014.
+Strecker, A.L.
+\& Olden, J.D.
+(2014).
 Fish species introductions provide novel insights into the patterns and drivers of phylogenetic structure in freshwaters.
-Proceedings of the Royal Society B: Biological Sciences.
-281:20133003.
+\emph{Proceedings of the Royal Society B: Biological Sciences}, \textbf{281}, 20133003.
 
 \leavevmode\hypertarget{ref-thomson2010sparse}{}%
-Thomson R.C., Shaffer H.B. 2010. Sparse supermatrices for phylogenetic inference: Taxonomy, alignment, rogue taxa, and the phylogeny of living turtles. Systematic biology. 59:42--58.
+Thomson, R.C.
+\& Shaffer, H.B.
+(2010).
+Sparse supermatrices for phylogenetic inference: Taxonomy, alignment, rogue taxa, and the phylogeny of living turtles.
+\emph{Systematic biology}, \textbf{59}, 42--58.
 
 \leavevmode\hypertarget{ref-tucker2012incorporating}{}%
-Tucker C.M., Cadotte M.W., Davies T.J., Rebelo T.G. 2012. Incorporating geographical and evolutionary rarity into conservation prioritization. Conservation Biology. 26:593--601.
+Tucker, C.M., Cadotte, M.W., Davies, T.J.
+\& Rebelo, T.G.
+(2012).
+Incorporating geographical and evolutionary rarity into conservation prioritization.
+\emph{Conservation Biology}, \textbf{26}, 593--601.
 
 \leavevmode\hypertarget{ref-verbruggen2010data}{}%
-Verbruggen H., Maggs C.A., Saunders G.W., Le Gall L., Yoon H.S., De Clerck O.
-2010.
+Verbruggen, H., Maggs, C.A., Saunders, G.W., Le Gall, L., Yoon, H.S.
+\& De Clerck, O.
+(2010).
 Data mining approach identifies research priorities and data requirements for resolving the red algal tree of life.
-BMC evolutionary biology.
-10:16.
+\emph{BMC evolutionary biology}, \textbf{10}, 16.
 
 \leavevmode\hypertarget{ref-vos2012nexml}{}%
-Vos R.A., Balhoff J.P., Caravas J.A., Holder M.T., Lapp H., Maddison W.P., Midford P.E., Priyam A., Sukumaran J., Xia X., others.
-2012.
+Vos, R.A., Balhoff, J.P., Caravas, J.A., Holder, M.T., Lapp, H., Maddison, W.P., Midford, P.E., Priyam, A., Sukumaran, J., Xia, X.
+\& others.
+(2012).
 NeXML: Rich, extensible, and verifiable representation of comparative data and metadata.
-Systematic biology.
-61:675--689.
+\emph{Systematic biology}, \textbf{61}, 675--689.
 
 \leavevmode\hypertarget{ref-webb2010biodiversity}{}%
-Webb C.O., Slik J.F., Triono T. 2010. Biodiversity inventory and informatics in southeast asia. Biodiversity and Conservation. 19:955--972.
+Webb, C.O., Slik, J.F. \& Triono, T. (2010). Biodiversity inventory and informatics in southeast asia. \emph{Biodiversity and Conservation}, \textbf{19}, 955--972.
 
 \leavevmode\hypertarget{ref-weber2014defense}{}%
-Weber M.G., Agrawal A.A. 2014. Defense mutualisms enhance plant diversification. Proceedings of the National Academy of Sciences. 111:16442--16447.
+Weber, M.G. \& Agrawal, A.A. (2014). Defense mutualisms enhance plant diversification. \emph{Proceedings of the National Academy of Sciences}, \textbf{111}, 16442--16447.
 
 \leavevmode\hypertarget{ref-wheeler2000database}{}%
-Wheeler D.L., Chappey C., Lash A.E., Leipe D.D., Madden T.L., Schuler G.D., Tatusova T.A., Rapp B.A.
-2000.
+Wheeler, D.L., Chappey, C., Lash, A.E., Leipe, D.D., Madden, T.L., Schuler, G.D., Tatusova, T.A.
+\& Rapp, B.A.
+(2000).
 Database resources of the national center for biotechnology information.
-Nucleic acids research.
-28:10--14.
+\emph{Nucleic acids research}, \textbf{28}, 10--14.
 
 \leavevmode\hypertarget{ref-wu2008simple}{}%
-Wu M., Eisen J.A. 2008. A simple, fast, and accurate method of phylogenomic inference. Genome biology. 9:R151.
+Wu, M. \& Eisen, J.A. (2008). A simple, fast, and accurate method of phylogenomic inference. \emph{Genome biology}, \textbf{9}, R151.
 
 \leavevmode\hypertarget{ref-xu2015ncbiminer}{}%
-Xu X., Dimitrov D., Rahbek C., Wang Z. 2015. NCBIminer: Sequences harvest from genbank. Ecography. 38:426--430.
+Xu, X., Dimitrov, D., Rahbek, C. \& Wang, Z. (2015). NCBIminer: Sequences harvest from genbank. \emph{Ecography}, \textbf{38}, 426--430.
 
 \leavevmode\hypertarget{ref-yong2010screening}{}%
-Yong L., Zhen M., Qi L., Yanping G., Yuanchun Z., Jianhui L.
-2010.
+Yong, L., Zhen, M., Qi, L., Yanping, G., Yuanchun, Z.
+\& Jianhui, L.
+(2010).
 Screening data for phylogenetic analysis of land plants: A parallel approach.
-2010 first international conference on networking and distributed computing.:305--308.
+\emph{2010 first international conference on networking and distributed computing} pp.
+305--308.
+IEEE.
 

### 2020-06-24 render2-submission-mee.Rmd
mee

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    includes:
      # in_header: preamble.tex
      # after_body: [biorxiv_2019.09.28figcaps.tex]
    # fig_caption: no
geometry: margin = 1in
bibliography: paper.bib
csl: submission-mee/methods-in-ecology-and-evolution.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    - \usepackage{caption}
    - \usepackage{blindtext}
    - \usepackage[left]{lineno}
    - \linenumbers
  # - \usepackage{xr}
    # - \usepackage[utf8]{inputenc}
    # - \usepackage[english]{babel}
    # - \usepackage[document]{ragged2e}
# nocite: |
#   @barker2012going, @barker2015new, @burns2014phylogenetics, # @claramunt2015new, @gibb2015new, @Hedges2015, # @hooper2017chromosomal, @Jetz2012, @price2014niche
---

<!-- ``` {r child = '0-header.Rmd'} -->
<!-- ``` -->
\newpage
``` {r child = '1-abstract.Rmd'}
```
\newpage

``` {r child = '2-body.Rmd'}
```
\newpage
\begin{center}
\textsc{References}
\end{center}

### 2020-06-24 2-body.Rmd
methods specifications

@@
   similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
+  database, and the accepted sequences on the alignment.
+  The blastn function from the BLAST command line tools [@camacho2009blast] is used for local-database searches.
+  A modified biopython blast function is used for web-based searches.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
+- The DNA sequence similarity search can be done on a local database that is easily
+  setup by the user. In this case, the blastn function is used to performs the similarity search [@camacho2009blast].
+- The search can also be performed remotely, on the NCBI database. In this case, the
+  bioPython BLAST function was modified to accepts is used to perform the similarity search.
 - A pairwise alignment-against-all BLAST search is performed. This means that each sequence
@@
 This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-- The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
 - Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
+  and maximum length threshold are discarded. ***REPORT THE DEFAULT VALUES AND DESCRIBE WHAT THEY MEAN*** This filtering leaves out genomic sequences.
   All acepted sequences are asigned an internal identifier, and are further filtered.
@@
 
-- A gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml]
-  with 100 classic bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
+- A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML [@stamatakis2014raxml]
+  with 100 classic rapid bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
   Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+- Bootstrap results are summarized with Dendropy ADD CITATION
 - The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
 - Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
-- Robinson Foulds metrics
-- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
+- Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
+- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to CITE REDELINGS AND HOLDER [@redelings and holder]
 - For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
+- A SUGGESTED DEFAULT ROOTING BASED ON THE OPEN TREE TAXONOMY is implemented for now.
+DESCRIBE HOW IT WORKS.
+SAY THAT IT IS A PROBLEM.
+Automatic rooting is not that smart yet.
+The best way right now is for users to define outgroups so trees are better rooted.
 - Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
@@
 
+# Authors' Contributions
+
+# Data Avilability
+
 # References

### 2020-06-24 _TODO.Rmd
refactoring
methods specifications

@@
 
+- [ ] Do not cite Felsenstein 1985 for bootstrap, just the call that we use for RAxML:
+rax_ex, "-T", "{}".format(self.config.num_threads),
+                                 "-m", "GTRCAT",
+                                 "-s", alignment,
+                                 "-p", "1", "-b", "1", "-#", str(num_reps),
+                                 "-n", "BOOT{}".format(self.date)
 - [ ] Finding a tree to update
@@
 
-- [ ] check the all-vs-all blast method, whre does it come from?
+- [ ] check the all-vs-all blast method, where does it come from?
 
@@
 
-- [ ] the root
-  - [ ] write something about the need to set the root to be able to do comparisons
-
 - [ ]  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT for that single taxon would be determined as ***HOW?***
+  - [ ] automatically, it will return the same taxon, even if its a species.
+  - [ ] all tips in original tree are mapped to ott ids
+  - [ ] sequences in alignment that are not mapped to ott ids will be kept in the alignment but not blast.
+Would only happen if somebody decides not to map a tip bc it is no longer the taxon they tought based on the new phylogenetic relationships.
 
 # Done!
+
+- [ ] determining the root
+  - [ ] write something about the need to set the root to be able to do comparisons
+  - [ ] say that utis hard to set it up automatically and that solutions are being sought
 

### 2020-06-24 docs/submission-mee-2020-06-24.tex
methods specifications

@@
 
+Manuscript prepared for submission to Methods in Ecology and Evolution
+Article type: Application
+
 \newpage
+
+\hypertarget{summary}{%
+\section{Summary}\label{summary}}
 
@@
 \end{enumerate}
+
+\textbf{Keywords}: phylogeny,
 
@@
   similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
+  database, and the accepted sequences on the alignment.
+  The blastn function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database searches.
+  A modified biopython blast function is used for web-based searches.
+\item
+  The DNA sequence similarity search can be done on a local database that is easily
+  setup by the user. In this case, the blastn function is used to performs the similarity search (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}).
+\item
+  The search can also be performed remotely, on the NCBI database. In this case, the
+  bioPython BLAST function was modified to accepts is used to perform the similarity search.
 \item
@@
 \item
-  The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-\item
-  The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
-\item
   Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
+  and maximum length threshold are discarded. \textbf{\emph{REPORT THE DEFAULT VALUES AND DESCRIBE WHAT THEY MEAN}} This filtering leaves out genomic sequences.
   All acepted sequences are asigned an internal identifier, and are further filtered.
@@
 \item
-  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
-  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
+  A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
+  with 100 classic rapid bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
   Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine \emph{et al.} (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
 \item
+  Bootstrap results are summarized with Dendropy ADD CITATION
+\item
   The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
@@
 \item
-  Robinson Foulds metrics
-\item
-  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
+  Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
+\item
+  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to CITE REDELINGS AND HOLDER ({\textbf{???}} and holder)
 \item
   For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
+\item
+  A SUGGESTED DEFAULT ROOTING BASED ON THE OPEN TREE TAXONOMY is implemented for now.
+DESCRIBE HOW IT WORKS.
+SAY THAT IT IS A PROBLEM.
+Automatic rooting is not that smart yet.
+The best way right now is for users to define outgroups so trees are better rooted.
 \item
@@
 
+\hypertarget{authors-contributions}{%
+\section{Authors' Contributions}\label{authors-contributions}}
+
+\hypertarget{data-avilability}{%
+\section{Data Avilability}\label{data-avilability}}
+
 \hypertarget{references}{%

### 2020-06-24 render2-submission-mee.Rmd
methods specifications

@@
     toc: false
-    includes:
-      # in_header: preamble.tex
-      # after_body: [biorxiv_2019.09.28figcaps.tex]
-    # fig_caption: no
 geometry: margin = 1in
@@
 link-citations: yes
-# indent: true # this works
-# header-includes:
-#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
-# nope, 4 spaces indents are not recognized, they do something else
 linestretch: 2
@@
     - \linenumbers
-  # - \usepackage{xr}
-    # - \usepackage[utf8]{inputenc}
-    # - \usepackage[english]{babel}
-    # - \usepackage[document]{ragged2e}
-# nocite: |
-#   @barker2012going, @barker2015new, @burns2014phylogenetics, # @claramunt2015new, @gibb2015new, @Hedges2015, # @hooper2017chromosomal, @Jetz2012, @price2014niche
+params:
+  author1: Luna L. Sanchez Reyes
+  affiliation: School of Natural Sciences, University of California, Merced
 ---
 
-<!-- ``` {r child = '0-header.Rmd'} -->
-<!-- ``` -->
+`r params$affiliations`
+Manuscript prepared for submission to Methods in Ecology and Evolution
+Article type: Application
+
 \newpage
+# Summary
+
 ``` {r child = '1-abstract.Rmd'}
 ```
+
+**Keywords**: phylogeny,
+
 \newpage
@@
 ```
+
 \newpage

### 2020-06-25 docs/submission-mee-2020-06-24.tex
final mee submission format for now

--- 2020-06-24 docs/submission-sysbio-2020-06-24.tex
@@
 \hypersetup{
+            pdftitle={Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life},
             pdfborder={0 0 0},
@@
 
+\usepackage{color}
+\usepackage{hyperref}
 \usepackage{caption}
 \usepackage{blindtext}
+\usepackage{url}
 \usepackage[left]{lineno}
@@
 
+\title{Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life}
 \author{}
@@
 \begin{document}
+\maketitle
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Luna L. Sanchez Reyes
+\end{enumerate}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Martha Kandziora
+\end{enumerate}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:martha.kandziora@mailbox.org}{\nolinkurl{martha.kandziora@mailbox.org}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Emily Jane McTavish
+\end{enumerate}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+
+Correspondence:
+
+Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
+
+email: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+
+Running title: Continually updated gene trees with Physcraper
+
+Word count: 4021
+
+Manuscript prepared for submission to Methods in Ecology and Evolution
+
+Article type: Application
 
@@
 \end{enumerate}
+
+\textbf{Keywords}: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, update alignment
 
@@
   similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
+  database, and the accepted sequences on the alignment.
+  The blastn function from the BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database searches.
+  A modified biopython blast function is used for web-based searches.
+\item
+  The DNA sequence similarity search can be done on a local database that is easily
+  setup by the user. In this case, the blastn function is used to performs the similarity search (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}).
+\item
+  The search can also be performed remotely, on the NCBI database. In this case, the
+  bioPython BLAST function was modified to accepts is used to perform the similarity search.
 \item
@@
 \item
-  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
-  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
+  A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
+  with 100 classic rapid bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
   Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine et al.
@@
 \item
+  Bootstrap results are summarized with Dendropy ADD CITATION
+\item
   The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
@@
 \NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property =} \StringTok{"treebaseId"}\NormalTok{, }\DataTypeTok{value =} \StringTok{"S2137"}\NormalTok{)}
+\CommentTok{##   study_ids n_trees tree_ids candidate study_year title}
+\CommentTok{## 1    pg_238       1  tree109                 2009      }
+\CommentTok{##                                 study_doi}
+\CommentTok{## 1 http://dx.doi.org/10.1093/sysbio/syp020}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-##   study_ids n_trees tree_ids candidate study_year title
-## 1    pg_238       1  tree109                 2009      
-##                                 study_doi
-## 1 http://dx.doi.org/10.1093/sysbio/syp020
-\end{verbatim}
 
@@
 \NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_tree_ids}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "tree109"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\CommentTok{## [1] "tree109"}
 \NormalTok{rotl}\OperatorTok{::}\KeywordTok{candidate_for_synth}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## NULL
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\CommentTok{## NULL}
 \NormalTok{my_trees <-}\StringTok{ }\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_study}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{)}

### 2020-06-25 render2-submission-mee.Rmd
final mee submission format for now

@@
 csl: submission-mee/methods-in-ecology-and-evolution.csl
+# csl: submission_sysbio/systematic-biology.csl
 link-citations: yes
@@
 header-includes:
+    - \usepackage{color}
+    - \usepackage{hyperref}
     - \usepackage{caption}
     - \usepackage{blindtext}
+    - \usepackage{url}
     - \usepackage[left]{lineno}
     - \linenumbers
+title: 'Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life'
 params:
+  running_title: Continually updated gene trees with Physcraper
   author1: Luna L. Sanchez Reyes
+  email1: sanchez.reyes.luna@gmail.com
+  author2: Martha Kandziora
+  email2: martha.kandziora@mailbox.org
+  author3: Emily Jane McTavish
+  email3: ejmctavish@gmail.com
   affiliation: School of Natural Sciences, University of California, Merced
+  correspondence_address: Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
 ---
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "##",
+  eval = TRUE
+)
+library(wordcountaddin)
+word_count_body <- wordcountaddin::word_count(filename="2-body.Rmd")
+```
 
-`r params$affiliations`
+1. `r params$author1`
+
+`r params$affiliation`
+
+email: <`r params$email1`>
+
+2. `r params$author2`
+
+`r params$affiliation`
+
+email: <`r params$email2`>
+
+3. `r params$author3`
+
+`r params$affiliation`
+
+email: <`r params$email3`>
+
+Correspondence:
+
+`r params$correspondence_address`
+
+email: <`r params$email1`>, <`r params$email3`>
+
+Running title: `r params$running_title`
+
+Word count: `r word_count_body`
+
 Manuscript prepared for submission to Methods in Ecology and Evolution
+
 Article type: Application
@@
 \newpage
-# Summary
 
@@
 
-**Keywords**: phylogeny,
+**Keywords**: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, update alignment
 

### 2020-06-25 2-body.Rmd
examples

@@
 
+We will address two use-case scenarios. One in which the user is interested in a particular group. Anothr one in which the user is interested in a particular phylogeny.
+
 ## The hollies
 
-The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-***Contact authors? They seem old school, probably do not wanna share their data.***
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
+A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
+It encompasses between 400-600 living species. A review of literature shows them that there are three published phylogenetic trees, showing relationships within the hollies.
+The first one has been made available both in the OToL phylesystem and is part of the synthetic tree, and on treeBASE. It samples 48 species.
+The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
+<!-- ***Contact authors? They seem old school, probably do not wanna share their data.*** -->
+The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
 This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
-Updating the oldest tree, we should get something very similar to the newest tree.
+We would expect that updating the oldest tree, we would get something very similar to the newest tree.
 
-## The Ascomycota
+## The Malvaceae
 
-Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by @schoch2009ascomycota.
-The first step, is to get the Open Tree of Life study id. There are some options to do this:
-- You can go to the Open Tree of Life website and browse until you find it, or
-- you can get the study id using R tools:
-  - By using the TreeBase ID of the study (which is not fully exposed on the
-    TreeBase website home page of the study, so you have to really look it up manually):
-```{r, treebase, eval = TRUE}
-rotl::studies_find_studies(property = "treebaseId", value = "S2137")
-```
-  - By using the name of the focal clade of study (but this behaved very differently):
-```{r, focal, eval = FALSE}
-rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
-```
-
-Once we have the study id, we can gather the trees published on that study:
-```{r, treeids, eval = TRUE}
-rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
-rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
-my_trees <- rotl::get_study("pg_238")
-```
-Both trees from this study have `r tryCatch(ape::Ntip(my_trees[[1]]), error = function(e){return(NA)})` tips.
-
-Let's check what one of the trees looks like:
-
-<!-- ```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.width = 2} -->
-<!-- ape::plot.phylo(ape::ladderize(my_trees[[1]]), type = "phylogram", cex = 0.1) -->
-<!-- ``` -->
-
-1. Download the alignment from TreeBase
-If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
-To make this task easier, you can use a command to download everything into your working folder:
-```
-physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
-```
-In this example, all alignments posted on TreeBase were used to reconstruct both trees.
-
-1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:
-
-```
-physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
-```
-
-## Testudines example
-
-Phylogeny of the Testudines 6 tips from @crawford2012more
-There is just one tree in OToL.
-There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.
-
-```
-physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
-```
 # Discussion
@@
 
-## Tools that automatize any part of the process of phylogenetic reconstruction:
-
-### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
-
-| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
-|-----|------|:------:|:------:|:------:|
-| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
-| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
-| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
-| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
-The authors present a nice comparison with the phylota algorithm | Species trees??
-|
-| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
-| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
-| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
-|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
-|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
-| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
-| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
-| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
-| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
-| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
-| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
-| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
-| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
-| phylotaR | @bennett2018phylotar | studies |  |  |
-
-According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
-"focus on the construction and refinement of robust alignments rather than the collection of homologs."
-
-### 2. Searching phylogenetic tree databases
-
-PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
-
-### 3. Mining phylogenetic tree databases
-
-PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database".
-
-### 4. Pipeline for phylogenetic reconstruction
-
-PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
-
-### 5. getting metadata and not sequences from GenBank.
-
-Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.
-
-## Phylota overview
-
-```{r phylota, child="100-phylota-overview.Rmd"}
-```
-
 # Acknowledgements

### 2020-06-25 2.3-extra examples.Rmd
examples

--- 2020-06-23 physcraper_ms.Rmd
@@
----
-title: 'Physcraper: A Python package for continually updated gene trees'
-author:
-  - Luna L. Sanchez Reyes
-  - Martha Kandziora
-  - Emily Jane McTavish
-  - University of California, Merced
-date: '`r format(Sys.time(), "%d %B, %Y")`'
-bibliography: paper.bib
-csl: apa.csl
-output:
-  redoc::redoc
----
-
-# Abstract
-
-1. Phylogenies are a key part of research in all areas of biology. Tools that automatize
-some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
-have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
-However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
-to one that is not a specialist either on phylogenetic methods or on a particular group of study.
-
-1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
-phylogenies by making use of public DNA sequence data and taxonomic information,
-providing a framework for comparison of published phylogenies with their updated versions.
-
-1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on already available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
-of alternative phylogenetic hypotheses (topologies).
-***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know much about a certain biological group.***
-
-1. Physcraper implements node by node/topology comparison of the the original and the updated
-trees using the conflict API of OToL, and summarizes differences.
-
-1. We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-
-1. Physcraper can be used with any OS. Detailed instructions for installation and
-use are available at <https://github.com/McTavishLab/physcraper>.
-
-# Introduction
-
-Phylogenies are important.
-
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-***paragraph better to end discussion???
-***
-
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
-
-1. Obtention of molecular or morphological character data -- get DNA from some organisms
-and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
-1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
-aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-1. Analyse this hypothesis of homology to infer phylogenetic relationships among
-the organisms you are studying -- Use different available programs to infer molecular
-evolution, trees and times of divergence.
-1. Discuss the inferred relationships in the context of previous hypothesis, the
-biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
-
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
-
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
-
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
-
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
-
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
-
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered "best" by experts
-not only in phylogenetics, but also experts on the focal group of study.
-
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control along the process
-- Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
-
-All these pipelines start tree construction from zero? Yes.
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-The original tree is by default used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
-
-# How does Physcraper work?
-
-## The input: a study tree and an alignment
-
-- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
-- The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, `physcraper` can download it directly,
-  either from the TreeBASE website (<https://treebase.org/>)
-  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
-- A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-- A ".csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from both the tree and alignment.
-  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.
-
-## DNA sequence search and cleaning
-
-- The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be provided by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
-
-  In the case that only one taxon from the ingroup is matched in both the tree and alignment, the MRCAT
-  for that single taxon would be determined as ***HOW?***
-  <!-- ***Physcraper does not take into account the focal clade from Open Tree.
-Because that is arbitrarily assigned or not by a curator.
-The focal clade is a search MRCA more than a taxonomic MRCA, physcraper's default is to get the ingroup from Otol, which is not the focal group.-->
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
-  <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-- The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
-- Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack database id numbers, a filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
-- Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-  <!--***How does pyphlawd, or phylota does the exemplar per species choosing?***-->
-- Reverse complement sequences are identified and translated.
-- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
-***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
-- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
-
-## DNA sequence alignment
-
-- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
-- First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the alignment of new sequences
-  is aligned against the original alignment, working as a template. This ensures that
-  the final alignment follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. We encourage users
-  to check it using any of the many tools for alignment processing, and eliminate
-  columns with no information.
-
-## Tree reconstruction
-
-- A gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with 100 bootstrap replicates.
-- The root of the tree is determined by finding the parent node of the sequences that do not belong to the ingroup.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- The final result is a tree for eac of the genes provided in the alignment.
-
-## Tree comparison
-
-- Gene trees are then coupled to conlict info.
-- Conflict information can only be generated in the context of the whole Open Tree of
-Life. Otherwise, it is not really possible to get conflict data.
-***- One way to compare two independent phylogenetic trees is to compare them both to
-the synthetic OToL and then measure how well they do against each other***
-
-# Examples
-
-## The hollies
-
-The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-***Contact authors? They seem old school, probably do not wanna share their data.***
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
-Updating the oldest tree, we should get something very similar to the newest tree.
-
 ## The Ascomycota
@@
 ```
-# Discussion
-
-Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
-
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
-exposing new comers to expert understanding about the data.
-
-There are many tools that are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
-Most focus on generating full trees *de novo*, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have been already studied and that have phylogenetic data already.
-
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
-
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.
-
-Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer [@ranwez2009phyloexplorer].
-Compare and discuss.
-
-How is physcraper already useful:
-- mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think.
--
-
-How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
-- obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
-- one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
-
-Things that physcraper does not do:
-- analyse the whole GenBank database [@benson2000genbank; @wheeler2000database] to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward.
-There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
-
-## Tools that automatize any part of the process of phylogenetic reconstruction:
-
-### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
-
-| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
-|-----|------|:------:|:------:|:------:|
-| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
-| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
-| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
-| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
-The authors present a nice comparison with the phylota algorithm | Species trees??
-|
-| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
-| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
-| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
-|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
-|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
-| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
-| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
-| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
-| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
-| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
-| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
-| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
-| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
-
-According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
-"focus on the construction and refinement of robust alignments rather than the collection of homologs."
-
-### 2. Searching phylogenetic tree databases
-
-PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
-
-### 3. Mining phylogenetic tree databases
-
-PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database".
-
-### 4. Pipeline for phylogenetic reconstruction
-
-PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
-
-### 5. getting metadata and not sequences from GenBank.
-
-Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.
-
-## Phylota overview
-
-```{r phylota, child="phylota-overview.Rmd"}
-```
-
-# Acknowledgements
-
-We acknowledge contributions from
-
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
-
-# References

### 2020-06-25 99-tools-review.Rmd
examples

---
output:
  redoc::redoc
---

## Tools that automatize any part of the process of phylogenetic reconstruction:

### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction

| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
|-----|------|:------:|:------:|:------:|
| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta. The authors present a nice comparison with the phylota algorithm | Species trees?? |
| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants. PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
| phylotaR | @bennett2018phylotar | studies |  |  |

According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
"focus on the construction and refinement of robust alignments rather than the collection of homologs."

### 2. Searching phylogenetic tree databases

PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies

### 3. Mining phylogenetic tree databases

PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database".

### 4. Pipeline for phylogenetic reconstruction

PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
pipeline.

### 5. getting metadata and not sequences from GenBank.

Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.

## Phylota overview

```{r phylota, child="100-phylota-overview.Rmd"}
```

### 2020-06-25 docs/submission-mee-2020-06-25.tex
examples

--- 2020-06-24 docs/submission-mee-2020-06-24.tex
@@
 \begin{document}
+\maketitle
+
+\textbf{1. Luna L. Sanchez Reyes}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}
+
+\textbf{2. Martha Kandziora}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:martha.kandziora@mailbox.org}{\nolinkurl{martha.kandziora@mailbox.org}}
+
+\textbf{3. Emily Jane McTavish}
+
+School of Natural Sciences, University of California, Merced
+
+email: \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+
+\textbf{Correspondence address}: Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
+
+\textbf{Correspondence email}: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+
+\textbf{Running title}: Continually updated gene trees with Physcraper
+
+\textbf{Word count}: 3246
 
 Manuscript prepared for submission to Methods in Ecology and Evolution
+
 Article type: Application
@@
 \newpage
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
 
@@
 
-\textbf{Keywords}: phylogeny,
+\textbf{Keywords}: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, update alignment
 
@@
 
-\hypertarget{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}{%
-\subsection{Tools that automatize any part of the process of phylogenetic reconstruction:}\label{tools-that-automatize-any-part-of-the-process-of-phylogenetic-reconstruction}}
-
-\hypertarget{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}{%
-\subsubsection{1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction}\label{mining-dna-databases-to-generate-datasets-suitable-for-phylogenetic-reconstruction}}
-
-\begin{longtable}[]{@{}llccc@{}}
-\toprule
-\begin{minipage}[b]{0.12\columnwidth}\raggedright
-Tool\strut
-\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright
-Citation\strut
-\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
-Cited by\strut
-\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
-Description\strut
-\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\centering
-Supermatrix/gene tree/species tree\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Phylota\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Sanderson \emph{et al.} (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-122 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-AMPHORA\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Wu \& Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-458 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-PHLAWD\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith \emph{et al.} (\protect\hyperlink{ref-smith2009mega}{2009})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-234 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Baited search of DNA markers on the GenBank database; phylogenetic reconstruction\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Unnamed \href{https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals}{ruby pipeline}, only available from \href{https://static-content.springer.com/esm/art\%3A10.1186\%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP}{supplementary data} of the journal\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Peters \emph{et al.} (\protect\hyperlink{ref-peters2011taming}{2011})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-64 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established ``criteria of compositional homogeneity and defined levels of density and overlap''\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Unnamed\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Grant \& Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-38 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Unnamed\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Chesters \& Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-10 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-algorithm that mines GenBank data to delineate species in the insecta. The authors present a nice comparison with the phylota algorithm\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Species trees??\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-PUmPER\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Izquierdo-Carrasco \emph{et al.} (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-14 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-perpetual updating with newly added sequences to GenBank\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure yet\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-DarwinTree\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Meng \emph{et al.} (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-6 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-NCBIminer\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Xu \emph{et al.} (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-4 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-part of darwintree\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-SUMAC\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-19 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-both ``baited'' analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-STBase\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-McMahon \emph{et al.} (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-7 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-pipeline for species tree construction and the public database of one million precomputed species trees\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-species trees\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Unnamed\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Papadopoulou \emph{et al.} (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-17 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Automated DNA-based plant identification for large-scale biodiversity assessment\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-BIR\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Kumar \emph{et al.} (\protect\hyperlink{ref-kumar2015bir}{2015})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-6 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-blast, align, identify homologs via constructed trees, curate and realign\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-SUPERSMART\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-35 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-baited analyses up to bayesian divergence time estimation\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-SOPHI\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-{[}Chesters (\protect\hyperlink{ref-chesters2017construction}{2017})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-17 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-not sure\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-phyloSkeleton\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Guy (\protect\hyperlink{ref-guy2017phyloskeleton}{2017})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-5 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-OneTwoTree\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Drori \emph{et al.} (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-7 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Web‐based, user-friendly, online tool for species-tree reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available sequence data from NCBI GenBank\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-pyPhlawd\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Smith \& Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-6 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-baited and clustering analyses\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-Supermatrix or gene tree\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-Phylotol\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-5 studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-``phylogenomic pipeline to allow easy incorporation of data from high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life'', i.e., bacteria and unicellular eukaryotes\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-supermatrix and gene trees\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\raggedright
-phylotaR\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-Bennett \emph{et al.} (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-studies\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-\strut
-\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\centering
-\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-According to Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
-``focus on the construction and refinement of robust alignments rather than the collection of homologs.''
-
-\hypertarget{searching-phylogenetic-tree-databases}{%
-\subsubsection{2. Searching phylogenetic tree databases}\label{searching-phylogenetic-tree-databases}}
-
-PhyloFinder (Chen \emph{et al.} \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
-
-\hypertarget{mining-phylogenetic-tree-databases}{%
-\subsubsection{3. Mining phylogenetic tree databases}\label{mining-phylogenetic-tree-databases}}
-
-PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database''.
-
-\hypertarget{pipeline-for-phylogenetic-reconstruction}{%
-\subsubsection{4. Pipeline for phylogenetic reconstruction}\label{pipeline-for-phylogenetic-reconstruction}}
-
-PhySpeTre (Fang \emph{et al.} \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
-
-\hypertarget{getting-metadata-and-not-sequences-from-genbank.}{%
-\subsubsection{5. getting metadata and not sequences from GenBank.}\label{getting-metadata-and-not-sequences-from-genbank.}}
-
-Datataxa Ruiz-Sanchez \emph{et al.} (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.
-
-\hypertarget{phylota-overview}{%
-\subsection{Phylota overview}\label{phylota-overview}}
-
-Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
-database (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-
-Since then, it has been cited 122 times for different reasons.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  As an example of a tool that mines GenBank data for phylogenetic reconstruction,
-  or that is useful in any way for phylogenetics:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    original publication of PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009})
-  \item
-    an analysis identifying research priorities and data requirements for resolving
-    the red algal tree of life (Verbruggen \emph{et al.} \protect\hyperlink{ref-verbruggen2010data}{2010})
-  \item
-    Beaulieu \emph{et al.} (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive
-    phylogeny from mined DNA sequence data, (even if no phylogeny was really published
-    there, only the method to do so)
-  \item
-    a review for ecologists about phylogenetic tools (Roquet \emph{et al.} \protect\hyperlink{ref-roquet2013building}{2013})
-  \item
-    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith \& Brown \protect\hyperlink{ref-smith2018constructing}{2018})
-  \item
-    a study presenting an ``assembly and alignment free'' method for phylogenetic reconstruction
-    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan \emph{et al.} \protect\hyperlink{ref-fan2015assembly}{2015}).
-  \item
-    nexml format presentation (Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses
-    stored phyloinformatic data that could benefit from adopting nexml, to increase
-    interoperability.
-  \item
-    a study of fruit evolution, analysing a previously published phylogeny of 8911
-    tips of the Campanulidae, constructed with PHLAWD (Beaulieu \& Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})
-  \item
-    a study of Southeast Asia plant biodiversity inventory (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}) -
-    cites phylota as a tool that would allow rapid phylogenetic placing of newly
-    discovered species, and generation of phylogenetically informed guides for field
-    identification.
-  \item
-    a study of wood density for carbon stock assessments (Flores \& Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),
-    cites phylota as an initiative to ``get supertrees resolved up to species level''.
-  \item
-    a study proposing something similar to Open tree but applied only to land plants (Beaulieu \emph{et al.} \protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})
-  \item
-    an analysis of the phylogenetic diversity-area curve (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-    cited phylota as a method alternative to phylomatic to ``obtain plant phylogenetic
-    trees for ecophylogenetic studies''.
-  \item
-    a study generating a phylogeny of 6,098 species of vascular plants from China
-    (Chen \emph{et al.} \protect\hyperlink{ref-chen2016tree}{2016}) - uses DarwinTree (Meng \emph{et al.} \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}) and generates sequence
-    data \emph{de novo} for 781 genera.
-  \item
-    a review of the state of methods and knowledge generated by molecular systematics
-    (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
-    for large-scale molecular phylogenetics analysis''.
-  \item
-    the first phylotastic paper (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny
-    related resource that provides ways to generate custom species trees for downstream use''.
-  \item
-    Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017}) cites phylota as a ``pipeline that pre-processes entire GenBank
-    releases in pursuit of sufficiently overlapping reciprocal BLAST hits, which are
-    then clustered into candidate data sets''. They also use the PHYLOTA database in its
-    own pipeline.
-  \item
-    Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2014evominer}{2014}) present an algorithm for mining of frequent subtrees (common patterns)
-    in collections of phylogenetic trees, as a way to extract meaningful phylogenetic
-    information from collections of trees when compared to maximum agreement subtrees
-    and majority-rule trees. They cite phylota as one of such tree collections available
-    along with TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
-  \item
-    Ranwez \emph{et al.} (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) cites phylota as a ``program providing basic statistics
-    on data availability for molecular datasets''. They propose a tool to upload and
-    explore user phylogenies to obtain detailed summary statistics on user tree collections.
-  \item
-    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}) cites phylota as a tool that ``provides a web interface to view
-    all GenBank sequences within taxonomic groups clustered into homologs'' but that
-    does not mine for targeted sequences, as opposed to NCBIminer or PHLAWD. They
-    compare the performance of SUMAC to Phylota. This is also presented in their PhD dissertation (Freyman \protect\hyperlink{ref-freyman2017phylogenetic}{2017}).
-  \item
-    Chesters \& Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) cites phylota as a data mining tool that compiles metadata
-    from mining of public DNA databases ``for construction of large phylogenetic trees
-    and multiple gene sets'' and that the authors have recognised that gene annotations
-    in public databases are insufficient and that careful partitioning of orthologous
-    sequences is needed for supermatrix construction. Chesters \& Vogler (\protect\hyperlink{ref-chesters2013resolving}{2013}) present
-    a procedure that minimizes the problem of forming multilocus species units in
-    a large phylogenetic data set using algorithms from graph theory.
-  \item
-    Chesters \& Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014}) present an algorithm to delineate species form GenBank
-    DNA data, and cites phylota as a tool that partitions ``the contents of a database
-    according to homology'', by ``grouping of database sequences according to internal
-    criteria'', searching ``from a standardized set of references {[}\ldots{}{]} patterns in
-    sequence similarity and overlap.''
-  \item
-    the paper presenting phylotaR, a pipeline that recreates the phylota output
-    but uses the most updated GenBank release, and is available in R (Bennett \emph{et al.} \protect\hyperlink{ref-bennett2018phylotar}{2018}),
-    cites phylota as its predecessor and inspiration. The authors mention that phylotaR
-    pipeline mimics phylota's pipeline but with improvements.
-  \item
-    The paper presenging PhyloBase (Jamil \protect\hyperlink{ref-jamil2016visual}{2016}), cites phylota as one of
-    its resources to get phylogenies, along with TreeBASE and others.
-  \item
-    The paper presenting STBase, a database of one million precomputed species
-    trees (Deepak \protect\hyperlink{ref-deepak2013managing}{2013}; McMahon \emph{et al.} \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as a databse of gene trees or mul-trees,
-    ``trees having multiple sequences with the same taxon name''.
-  \item
-    Drori \emph{et al.} (\protect\hyperlink{ref-drori2018onetwotree}{2018}) present a Web‐based, user-friendly, online tool for species-tree
-    reconstruction, based on the \emph{supermatrix paradigm} and retrieves all available
-    sequence data from NCBI GenBank. They cite phylota in the intro as a tool that is ``designed to provide
-    users with precomputed sets of clusters that were assembled through a single‐linkage
-    clustering approach and additionally provides precomputed gene trees that were
-    reconstructed for each cluster. In particular, the results obtained by PhyLoTa
-    are taxonomically constrained; that is, all sequences of the most recent common
-    ancestor are collected even if one specifies only part of a clade''.
-  \item
-    A study developing a tool to link wikipedia data to NCBI taxonomy (Page \protect\hyperlink{ref-page2011linking}{2011})
-    cites phylota as a phylogenetic resource that uses the NCBI taxonomy.
-  \item
-    the study that present DarwinTree (Meng \emph{et al.} \protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a}), and all derived studies: the study
-    presenting an approach to screen sequence data for The Platform
-    for Phylogenetic Analysis of Land Plants (PALPP), using the MapReduce paradigm
-    to parallelize BLAST (Yong \emph{et al.} \protect\hyperlink{ref-yong2010screening}{2010}), as well as Gao \emph{et al.} (\protect\hyperlink{ref-gao2011solution}{2011}), Li \emph{et al.} (\protect\hyperlink{ref-li2013partfasttree}{2013}),
-    Meng \emph{et al.} (\protect\hyperlink{ref-meng2014rapidtree}{2014}), Meng \emph{et al.} (\protect\hyperlink{ref-meng2015sotree}{2015}\protect\hyperlink{ref-meng2015sotree}{c}), and Meng \emph{et al.} (\protect\hyperlink{ref-meng2015solution}{2015}\protect\hyperlink{ref-meng2015solution}{b}), all cite phylota using the exact same
-    introduction and sentence: as one among other ``studies based on data mining large numbers of taxa or loci''.
-  \item
-    A study presenting a tool to asses gene sequence quality for automatic
-    construction of databases (Meng \emph{et al.} \protect\hyperlink{ref-meng2012gsqct}{2012}\protect\hyperlink{ref-meng2012gsqct}{a}), as well as their parallelized version using MapReduce
-    (Meng \emph{et al.} \protect\hyperlink{ref-meng2012cloud}{2012}\protect\hyperlink{ref-meng2012cloud}{b}), cite phylota (along with Yong \emph{et al.} (\protect\hyperlink{ref-yong2010screening}{2010})) as a tool that relies on sequence
-    similarity (BLAST) and not taxon name annotations in the database, for mining
-    large numbers of taxa or loci, without making any control on the quality of the
-    sequencing.
-  \item
-    A review on online plant databases aiming to ``provide recommendations for current
-    information managers and developers concerning the user interface and experience;
-    and to provide a picture about the possible directions to take for those in charge
-    of the creation of information at all levels''. They cite phylota as a tool allowing
-    researchers ``to acces equally and globally, without travel, a {[}phylogenetic?{]} model
-    of plants at the kingdom level'' (Jones \emph{et al.} \protect\hyperlink{ref-jones2014trends}{2014}).
-  \item
-    a paper aiming to establish an online information system for the legumes and
-    to outline ``best practices for development of a legume portal to enable data
-    sharing and a better understanding of what data are available, missing, or erroneous,
-    and ultimately facilitate cross-analyses and collaboration within the legume-systematics
-    community and with other stakeholders'' (Bruneau \emph{et al.} \protect\hyperlink{ref-bruneau2019towards}{2019}), cites phylota (along with supersmart and pyphlawd) as a
-    ``pipeline for large-scale retrieval of GenBank data of particular taxa or clades''.
-    In their Table 1, they also list phylota as a potential data source for developing a legume portal.
-  \item
-    A study on morphological evolution of electric fish skull, that uses phylotaR
-    to retrieve sequences of the family Apteronotidae, order Gymnotiformes (Evans \emph{et al.} \protect\hyperlink{ref-evans2019bony}{2019}),
-    cites phylota as the inspiration and fundament of phylotaR.
-  \item
-    A phylogenetic revision of the Gymnotidae fish (Teleostei: Gymnotiformes),
-    uses phylotaR to retrieve seqeunces, but cites phylota as ``a pipeline that implements
-    BLAST searches to both identify and download sequence clusters for listed taxonomic
-    groups to assemble a robust collection of sequences in a reproducible way based
-    on publicly-available gene sequences while avoiding selection bias on the part
-    of the assembler''.
-  \item
-    A master thesis on SearchTree, a ``software tool that allows users to query
-    efficiently on an arbitrary user taxon list and returns high scoring matches
-    from approximately one billion phylogenetic trees being constructed from molecular
-    sequence data in GenBank'' (Deepak \protect\hyperlink{ref-deepak2010searchtree}{2010}), that seems to be the preliminary
-    work for STBase (McMahon \emph{et al.} \protect\hyperlink{ref-mcmahon2015stbase}{2015}), cites phylota as ``a standard strategy, to assemble sets of homologous sequences
-    (clusters) from a database of all-against-all BLAST searches, {[}in which{]} clusters
-    are constructed in the context of the NCBI taxonomy tree for convenience of display,
-    thus child clusters are contained within parent clusters, following the NCBI hierarchy''.
-    In opposition, SearchTree uses true agglomerative hierarchical clustering
-    (AHC: Day \& Edelsbrunner (\protect\hyperlink{ref-day1984efficient}{1984})) based on the BLAST estimates of sequence dissimilarity
-    rather than the NCBI tree".
-  \item
-    a recent review on the state of large phylogeny (namely insects) generation
-    using tools of the data-driven era (Chesters \protect\hyperlink{ref-chesters2019phylogeny}{2019}) cites phylota as
-    a tool for homology inference and retrieval.
-  \item
-    the study presenting phylotol (Cerón-Romero \emph{et al.} \protect\hyperlink{ref-ceron2019phylotol}{2019}), cites phylota as a tool
-    that ``focus on the identification and collection of homologous genes from public
-    databases''.
-  \item
-    The \href{https://www.researchgate.net/profile/Douglas_Soltis/publication/228815637_Assembling_the_Tree_of_Life_for_the_Plant_Sciences_iPTOL/links/56a7c6bc08aeded22e3700dc.pdf}{iPTOL project}
-    cites phylota as a resource of phylogenetic trees.
-  \item
-    Mahmood (\protect\hyperlink{ref-mahmood2015avian}{2015}) PhD dissertation presents a database of avian Raptor sequences
-    (raptorbase), based on the phylota pipeline.
-  \item
-    Ruiz-Sanchez \emph{et al.} (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) develops datataxa and cite phylota as ``software that has
-    been developed to mine the massive amount of information stored in GenBank'',
-    along with its R version (phylotaR; Bennett \emph{et al.} \protect\hyperlink{ref-bennett2018phylotar}{2018}) and restez \url{https://www.rdocumen-tation.org/packages/restez/versions/1.0.0}.
-  \item
-    The phylotastic project (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}) cites phylota as a ``phylogeny-related resource providing ways to generate custom species trees \emph{de novo} for downstream use'' along with CIPRES.
-  \end{itemize}
-\item
-  When the software was actually used to construct (partially or in full) a DNA
-  data set to be used for phylogenetic reconstruction:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A 1000 tip phylogeny of the family of the nightshades (Särkinen \emph{et al.} \protect\hyperlink{ref-sarkinen2013solanaceae}{2013})
-  \item
-    A 56 tip phylogeny of crustacean zooplancton (Helmus \emph{et al.} \protect\hyperlink{ref-helmus2010communities}{2010}) -- ecological study
-  \item
-    A 63 tip phylogeny of the Salmonidae family (Crête-Lafrenière \emph{et al.} \protect\hyperlink{ref-crete2012salmonidae}{2012})
-  \item
-    A 321 tip phylogeny of Testudines (Thomson \& Shaffer \protect\hyperlink{ref-thomson2010sparse}{2010})
-  \item
-    A 69 taxa phylogeny of the family Cyprinodontidae of the pupfish (Martin \& Wainwright \protect\hyperlink{ref-martin2011trophic}{2011})
-  \item
-    A 2,957 taxa phylogeny of the class Moniloformopses of living ferns (Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011})
-  \item
-    A 2,573 species phylogeny of the Papilionoidea (Hardy \& Otto \protect\hyperlink{ref-hardy2014specialization}{2014})
-  \item
-    A 23 taxa phylogeny of the California flora (Anacker \emph{et al.} \protect\hyperlink{ref-anacker2011origins}{2011})
-  \item
-    Phylogenies of 6 different clades of flowering plants representing an independent
-    evolutionary origin of extrafloral nectaries: \emph{Byttneria} (Malvaceae), \emph{Pleopeltis} (Polypodiaceae),
-    \emph{Polygoneae} (Polygoneaceae), \emph{Senna} (Fabaceae), \emph{Turnera} (Passifloraceae), and \emph{Viburnum}
-    (Adoxaceae) (Weber \& Agrawal \protect\hyperlink{ref-weber2014defense}{2014}).
-  \item
-    To supplement DNA data sets of various pre-existing mammalian phylogenetic trees
-    sampled at different taxonomic levels (Faurby \& Svenning \protect\hyperlink{ref-faurby2015species}{2015})
-  \item
-    A 900 species tree of muroid rodents, Muroidea (Steppan \& Schenk \protect\hyperlink{ref-steppan2017muroid}{2017}), where 300
-    species were newly added by the study and the rest obtained using phylota.
-  \item
-    A 95 taxa phylogeny of Gymnosperms, focused on Ephedra, Gnetales (Ickert-Bond \emph{et al.} \protect\hyperlink{ref-ickert2009fossil}{2009})
-  \item
-    A 1061 genera phylogeny of the Oscine birds (Selvatti \emph{et al.} \protect\hyperlink{ref-selvatti2015paleogene}{2015})
-  \item
-    A 268 species phylogeny of sharks, representing all 8 orders and 32 families (Sorenson \protect\hyperlink{ref-sorenson2014evolution}{2014}; Sorenson \emph{et al.} \protect\hyperlink{ref-sorenson2014effect}{2014})
-  \item
-    A 466 species phylogeny of the Proteaceae, focusing on the species found in the Cape Floristic Region (Tucker \emph{et al.} \protect\hyperlink{ref-tucker2012incorporating}{2012}).
-  \item
-    A series of small phylogenies of unreported exact size, of sister groups of gall-forming insects (Hardy \& Cook \protect\hyperlink{ref-hardy2010gall}{2010}).
-  \item
-    A 196 species phylogeny of the family Boraginaceae (Nazaire \& Hufford \protect\hyperlink{ref-nazaire2012broad}{2012}). The authors
-    actually found data for 318 Boraginaceae spp using phylota, but decided to reduce
-    their data set to focus on the monophyly of genus \emph{Mertensia}.
-  \item
-    A phylogeny of 401 species of scale insects Coccoidea, Hemiptera (Ross \emph{et al.} \protect\hyperlink{ref-ross2013large}{2013}),
-    with some sequences generated \emph{de novo}.
-  \item
-    Two phylogenies sampling all species of two different clades of insectivorous
-    lizards, agamids and diplodactyline geckos, groups considered to be radiating
-    in the Australia's Great Victoria Desert (Rabosky \emph{et al.} \protect\hyperlink{ref-rabosky2011species}{2011})
-  \item
-    A phylogeny of 91 species of sparid and centracanthid fishes, Sparidae, Percomorpha,
-    plus 2 outgroups, a lethrinid and a nemipterid exemplar (Santini \emph{et al.} \protect\hyperlink{ref-santini2014first}{2014}).
-  \item
-    Updating a phylogeny of Arecaceae, constructing relationships in 6 cldes within
-    the group: subfamilies Calamoideae and Coryphoideae, the tribe Ceroxyleae within
-    subfamily Ceroxyloideae and three groups within subfamily Arecoideae: (1) Iriarteeae,
-  \end{itemize}
-
-  \begin{enumerate}
-  \def\labelenumii{(\arabic{enumii})}
-  \setcounter{enumii}{1}
-  \tightlist
-  \item
-    Cocoseae: Attaleinae except Beccariophoenix and (3) a group containing six
-    tribes; Euterpeae, Leopoldinieae, Pelagodoxeae, Manicarieae, Geonomateae and Areceae
-    (Faurby \emph{et al.} \protect\hyperlink{ref-faurby2016all}{2016}).
-  \end{enumerate}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A phylogeny of 768 Gesneriaceae species and 58 outgroups for a total species
-    sampling of 826 taxa (Roalson \& Roberts \protect\hyperlink{ref-roalson2016distinct}{2016}) some sequence were generated \emph{de novo}.
-  \item
-    A phylogeny of 47 species of scombrid fishes, with 2 outgroups, a gempylid and
-    a trichiurid (Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}).
-  \item
-    to update a dataset underlying a large-scale fern phylogeny (Lehtonen \emph{et al.} \protect\hyperlink{ref-lehtonen2017environmentally}{2017}),
-    data set in \url{https://zenodo.org/record/345670\#.Xr9QFRPYqqg}, also in TreeBASE,
-    but it is one of those studies that is broken.
-  \item
-    A phylogeny of 13 species of billfishes, order Istiophoriformes: Acanthomorpha,
-    and four outgroups (Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013})
-  \item
-    A phylogeny of 765 aphid species, family Aphididae (Hardy \emph{et al.} \protect\hyperlink{ref-hardy2015evolution}{2015})
-  \item
-    A phylogeny of less than 100 taxa of the family Ranunculaceae (Lehtonen \emph{et al.} \protect\hyperlink{ref-lehtonen2016sensitive}{2016}),
-    even though they retrieved info from phylota for 194 taxa within the family, they
-    reduced their data set because of low sampling of markers for some taxa.
-  \item
-    A phylogeny of 144 neobatrachian genera, assuming the monophyletic status of
-    genera to increase matrix-filling levels (Frazao \emph{et al.} \protect\hyperlink{ref-frazao2015gondwana}{2015}).
-  \item
-    A 179 species phylogeny of the bird family Picidae (woodpeckers, piculets,
-    and wrynecks) (Dufort \protect\hyperlink{ref-dufort2015coexistence}{2015}, \protect\hyperlink{ref-dufort2016augmented}{2016}), augmented with data from an updated GenBank
-    release and newly sequenced data.
-  \item
-    A phylogeny of species of freshwater fish endemic to NorthAmerica (Strecker \& Olden \protect\hyperlink{ref-strecker2014fish}{2014}),
-    phylota found data for 54 out of 66 spp.
-  \item
-    A phylogeny of 520 species of the order Ericales (Hardy \& Cook \protect\hyperlink{ref-hardy2012testing}{2012})
-  \item
-    A phylgeny of 16 fish species of the family Sphyraenidae (Percomorpha), as well
-    as two outgroup species of the Centropomidae (barracudas) (Santini \emph{et al.} \protect\hyperlink{ref-santini2015first}{2015})
-  \item
-    A phylogeny of 34 vole species, Arvicolinae, Rodentia (García-Navas \emph{et al.} \protect\hyperlink{ref-garcia2016role}{2016})
-  \item
-    Kolmann \emph{et al.} (\protect\hyperlink{ref-kolmann2017dna}{2017}) uses phylota to download all 1691 co1 sequences belonging to
-    the order Carchariniformes, to place phylogenetically DNA samples obtained from
-    fish markets.
-  \item
-    A phylogeny of 329 bird species in the Tyrannidae (77\% of the species in the
-    family) (Gómez Bahamón \& others \protect\hyperlink{ref-gomez2015behavioral}{2015}; Gómez-Bahamón \emph{et al.} \protect\hyperlink{ref-gomez2020speciation}{2020})
-  \item
-    Retrive 145 sequences registered as \emph{Holothuria} species, but kept 84 as ingroup,
-    plus 4 outgroup sequences from \emph{Stichopus ocellatus}, all belonging to the order
-    Apodida of sea cucumbers (Kamarudin \emph{et al.} \protect\hyperlink{ref-kamarudin2016phylogenetic}{2016})
-  \item
-    On a master thesis, to get the sequences of the outgroups of Melinidinae, family Poaceae, namely several spp of the
-    subfamily Panicoideae, plus \emph{Gynerium sagittatum}, \emph{Chasmanthium latifolium},
-    and \emph{Zea mays}, (Salariato \protect\hyperlink{ref-salariato2010filogenia}{2010}). Interestingly, phylota was not used
-    in the published study of the thesis (Salariato \emph{et al.} \protect\hyperlink{ref-salariato2010molecular}{2010}). Ingroup sequences were generated \emph{de novo}.
-  \item
-    On a PhD thesis, to construct a phylogeny of Platyrrhini (internal group),
-    Catarrhini (outgroup), and Tarsiiformes Pereira (\protect\hyperlink{ref-pereira2013padroes}{2013}). Have not found a published study.
-  \item
-    The 10k trees project (Arnold \emph{et al.} \protect\hyperlink{ref-arnold201010ktrees}{2010}) uses phylota to construct a tree of 301 primate species
-    and the outgroup species \emph{Galeopterus variegates}, a tree of 17 extant odd-toed
-    ungulates species and the outgroup species \emph{Bos taurus}, and a tree of 70 different
-    species of carnivorans and \emph{Equus caballus} as outgroup. However, the do not
-    cite it on the paper, but only on their documentation \url{http://www.academia.edu/download/49690788/10kTrees_Documentation.pdf}.
-  \item
-    Freyman (\protect\hyperlink{ref-freyman2015sumac}{2015}, also in \protect\hyperlink{ref-freyman2017phylogenetic}{2017}), use phylota to construct
-    a phylogeny (or maybe only mine genbank???) of the Onagraceae and Lythracea,
-    and compare it to the tool they propose, SUMAC.
-  \item
-    Blackmon (\protect\hyperlink{ref-blackmon2017synthesis}{2017}) PhD study applies phylota to reconstruct a 822 mite
-    species tree.
-  \item
-    A study of the effect of poliploidy on niche evolution (Baniaga \emph{et al.} \protect\hyperlink{ref-baniaga2018polyploid}{2018}),
-    uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
-    16 families and 25 genera,and a tree of 33 genera from 20 different families
-    comprising 1706 taxa.
-  \end{itemize}
-\item
-  When the website was used to identify sequences and markers available in
-  GenBank for a particular group. In this cases, the dataset mining was either performed
-  with other tools, or not performed at all and just used for discussion:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    A 812 tips phylogeny of the Order Chiroptera (Shi \& Rabosky \protect\hyperlink{ref-shi2015speciation}{2015}) -- dataset
-    constructed with PHLAWD
-  \item
-    A 1276 tips phylogeny of the Fabaceae (Group \emph{et al.} \protect\hyperlink{ref-legume2013legume}{2013}) -- dataset constructed
-    by hand (I think??)
-  \item
-    A review of dated phylogenies of fire-prone tropical savanna species from Brazil
-    (Simon \& Pennington \protect\hyperlink{ref-simon2012cerrado}{2012}) -- just for discussion of the lack of markers available for
-    these species on GenBank
-  \item
-    A review of the phylogeetic sof the Apicomplexa, a parasitic phylum on unicellular
-    protists (Morrison \protect\hyperlink{ref-morrison2009apicomplexa}{2009}).
-  \item
-    Three data sets from phylota (the suborder Pleurodira of side-necked turtles;
-    the family Cactaceae of cacti; and the Amorpheae, a clade of legumes) were used
-    to demonstrate and exemplify phylogenetic decisiveness (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2010phylogenomics}{2010})
-  \item
-    Mentioned in a PHD thesis (Gagnon \& others \protect\hyperlink{ref-gagnon2016systematique}{2016}), but not on the final publication (Gagnon \emph{et al.} \protect\hyperlink{ref-gagnon2016new}{2016}),
-    phylota was used to state that there are very few sequences available for the Legumes (7,482 out of 19,500 spp) on GenBank's release 194 (Feb2013).
-  \end{itemize}
-\item
-  Sometimes, it was cited by mistake:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    In this 630 tip phylogeny of the Caryophyllaceae study (Greenberg \& Donoghue \protect\hyperlink{ref-greenberg2011caryophyllaceae}{2011}) it might have been originally
-    cited as an example of large phylogenies that reflect well supported relationships
-    from previous smaller phylogenies. However, it was removed from the text but
-    not from the final list of references. The DNA data set was constructed by hand
-    most probably.
-  \item
-    a study reconstructing the insect tree of life with 49,358 species, 13,865
-    genera, and 760 families within the order Insecta (Chesters \protect\hyperlink{ref-chesters2017construction}{2017}),
-    uses its own algorithm (SOPHI) to mine public DNA databases (Chesters \& Zhu \protect\hyperlink{ref-chesters2014protocol}{2014}).
-    It does not cite phylota as it should, but includes it in their references.
-  \end{itemize}
-\item
-  When phylota was used to extract full trees (not only DNA data sets or markers):
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Page (\protect\hyperlink{ref-page2013bionames}{2013}) uses it to generate phylogenies for the \href{http://bionames.org}{bionames website},
-    a ``database linking taxonomic names to their original descriptions, to taxa, and
-    to phylogenies'' generated with phylota.
-  \item
-    Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2013extracting}{2013}) uses a sample of phylota trees to test their method to
-    remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
-    trees with two or more leaves sharing a label, e.g., a species name, which can
-    imply multiple conflicting phylogenetic relationships for the same set of taxa.
-  \item
-    A review by Sanderson \emph{et al.} (\protect\hyperlink{ref-sanderson2016perspective}{2016}), takes 134 595 gene trees from phylota
-    GenBank rel. 176 and estimates its degree of resolutin, calculating that less
-    than half of clades are supported with minilam statistical support (0.53 ± 0.32).
-  \end{itemize}
-\end{enumerate}
-
 \hypertarget{acknowledgements}{%
@@
 Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of molecular biology}, \textbf{215}, 403--410.
-
-\leavevmode\hypertarget{ref-anacker2011origins}{}%
-Anacker, B.L., Whittall, J.B., Goldberg, E.E.
-\& Harrison, S.P.
-(2011).
-Origins and consequences of serpentine endemism in the california flora.
-\emph{Evolution: International Journal of Organic Evolution}, \textbf{65}, 365--376.
 
@@
 
-\leavevmode\hypertarget{ref-arnold201010ktrees}{}%
-Arnold, C., Matthews, L.J. \& Nunn, C.L. (2010). The 10kTrees website: A new online resource for primate phylogeny. \emph{Evolutionary Anthropology: Issues, News, and Reviews}, \textbf{19}, 114--118.
-
-\leavevmode\hypertarget{ref-baniaga2018polyploid}{}%
-Baniaga, A.E., Marx, H.E., Arrigo, N. \& Barker, M.S. (2018). Polyploid plants have faster rates of multivariate climatic niche evolution than their diploid relatives. \emph{BioRxiv}, 406314.
-
-\leavevmode\hypertarget{ref-beaulieu2013fruit}{}%
-Beaulieu, J.M. \& Donoghue, M.J. (2013). Fruit evolution and diversification in campanulid angiosperms. \emph{Evolution}, \textbf{67}, 3132--3144.
-
-\leavevmode\hypertarget{ref-beaulieu2012modeling}{}%
-Beaulieu, J.M., Jhwueng, D.-C., Boettiger, C.
-\& O'Meara, B.C.
-(2012a).
-Modeling stabilizing selection: Expanding the ornstein--uhlenbeck model of adaptive evolution.
-\emph{Evolution: International Journal of Organic Evolution}, \textbf{66}, 2369--2383.
-
-\leavevmode\hypertarget{ref-beaulieu2012synthesizing}{}%
-Beaulieu, J.M., Ree, R.H., Cavender-Bares, J., Weiblen, G.D. \& Donoghue, M.J. (2012b). Synthesizing phylogenetic knowledge for ecological research. \emph{Ecology}, \textbf{93}, S4--S13.
-
-\leavevmode\hypertarget{ref-bennett2018phylotar}{}%
-Bennett, D.J., Hettling, H., Silvestro, D., Zizka, A., Bacon, C.D., Faurby, S., Vos, R.A.
-\& Antonelli, A.
-(2018).
-PhylotaR: An automated pipeline for retrieving orthologous dna sequences from genbank in r.
-\emph{Life}, \textbf{8}, 20.
-
 \leavevmode\hypertarget{ref-benson2000genbank}{}%
@@
 
-\leavevmode\hypertarget{ref-blackmon2017synthesis}{}%
-Blackmon, H.
-(2017).
-\emph{Synthesis and phylogenetic comparative analyses of the causes and consequences of karyotype evolution in arthropods}.
-PhD thesis thesis, University of Texas, Arlington.
-Retrieved from \url{http://hdl.handle.net/10106/26711}
-
-\leavevmode\hypertarget{ref-bruneau2019towards}{}%
-Bruneau, A., Borges, L.M., Allkin, R., Egan, A.N., De La Estrella, M., Javadi, F., Klitgaard, B., Miller, J.T., Murphy, D.J., Sinou, C.
-\& others.
-(2019).
-Towards a new online species-information system for legumes.
-\emph{Australian Systematic Botany}, \textbf{32}, 495--518.
-
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
 Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC bioinformatics}, \textbf{10}, 421.
-
-\leavevmode\hypertarget{ref-ceron2019phylotol}{}%
-Cerón-Romero, M.A., Maurer-Alcalá, X.X., Grattepanche, J.-D., Yan, Y., Fonseca, M.M.
-\& Katz, L.
-(2019).
-PhyloToL: A taxon/gene-rich phylogenomic pipeline to explore genome evolution of diverse eukaryotes.
-\emph{Molecular biology and evolution}, \textbf{36}, 1831--1842.
-
-\leavevmode\hypertarget{ref-chen2008phylofinder}{}%
-Chen, D., Burleigh, J.G., Bansal, M.S. \& Fernández-Baca, D. (2008). PhyloFinder: An intelligent search engine for phylogenetic tree databases. \emph{BMC Evolutionary Biology}, \textbf{8}, 90.
-
-\leavevmode\hypertarget{ref-chen2016tree}{}%
-Chen, Z.-D., Yang, T., Lin, L., Lu, L.-M., Li, H.-L., Sun, M., Liu, B., Chen, M., Niu, Y.-T., Ye, J.-F.
-\& others.
-(2016).
-Tree of life for the genera of chinese vascular plants.
-\emph{Journal of Systematics and Evolution}, \textbf{54}, 277--306.
-
-\leavevmode\hypertarget{ref-chesters2017construction}{}%
-Chesters, D. (2017). Construction of a species-level tree of life for the insects and utility in taxonomic profiling. \emph{Systematic biology}, \textbf{66}, 426--439.
-
-\leavevmode\hypertarget{ref-chesters2019phylogeny}{}%
-Chesters, D. (2019). The phylogeny of insects in the data-driven era. \emph{Systematic Entomology}.
-
-\leavevmode\hypertarget{ref-chesters2013resolving}{}%
-Chesters, D.
-\& Vogler, A.P.
-(2013).
-Resolving ambiguity of species limits and concatenation in multilocus sequence data for the construction of phylogenetic supermatrices.
-\emph{Systematic Biology}, \textbf{62}, 456--466.
-
-\leavevmode\hypertarget{ref-chesters2014protocol}{}%
-Chesters, D. \& Zhu, C.-D. (2014). A protocol for species delineation of public dna databases, applied to the insecta. \emph{Systematic biology}, \textbf{63}, 712--725.
 
@@
 
-\leavevmode\hypertarget{ref-crete2012salmonidae}{}%
-Crête-Lafrenière, A., Weir, L.K. \& Bernatchez, L. (2012). Framing the salmonidae family phylogenetic portrait: A more complete picture from increased taxon sampling. \emph{PloS one}, \textbf{7}.
-
-\leavevmode\hypertarget{ref-day1984efficient}{}%
-Day, W.H. \& Edelsbrunner, H. (1984). Efficient algorithms for agglomerative hierarchical clustering methods. \emph{Journal of classification}, \textbf{1}, 7--24.
-
-\leavevmode\hypertarget{ref-deepak2013managing}{}%
-Deepak, A. (2013). \emph{Managing and analyzing phylogenetic databases}. PhD thesis thesis, Retrieved from \url{https://lib.dr.iastate.edu/etd/12995}
-
-\leavevmode\hypertarget{ref-deepak2010searchtree}{}%
-Deepak, A. (2010). SearchTree: Mining robust phylogenetic trees.
-
-\leavevmode\hypertarget{ref-deepak2013extracting}{}%
-Deepak, A., Fernández-Baca, D. \& McMahon, M.M. (2013). Extracting conflict-free information from multi-labeled trees. \emph{Algorithms for Molecular Biology}, \textbf{8}, 18.
-
 \leavevmode\hypertarget{ref-deepak2014evominer}{}%
@@
 
-\leavevmode\hypertarget{ref-drori2018onetwotree}{}%
-Drori, M., Rice, A., Einhorn, M., Chay, O., Glick, L. \& Mayrose, I. (2018). OneTwoTree: An online tool for phylogeny reconstruction. \emph{Molecular ecology resources}, \textbf{18}, 1492--1499.
-
-\leavevmode\hypertarget{ref-dufort2016augmented}{}%
-Dufort, M.J. (2016). An augmented supermatrix phylogeny of the avian family picidae reveals uncertainty deep in the family tree. \emph{Molecular phylogenetics and evolution}, \textbf{94}, 313--326.
-
-\leavevmode\hypertarget{ref-dufort2015coexistence}{}%
-Dufort, M.
-(2015).
-\emph{Coexistence, ecomorphology, and diversification in the avian family picidae (woodpeckers and allies)}.
-PhD thesis thesis, University of Minnesota.
-Retrieved from \url{http://hdl.handle.net/11299/175702}
-
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
@@
 
-\leavevmode\hypertarget{ref-evans2019bony}{}%
-Evans, K.M., Vidal-García, M., Tagliacollo, V.A., Taylor, S.J.
-\& Fenolio, D.B.
-(2019).
-Bony patchwork: Mosaic patterns of evolution in the skull of electric fishes (apteronotidae: Gymnotiformes).
-\emph{Integrative and comparative biology}, \textbf{59}, 420--431.
-
-\leavevmode\hypertarget{ref-fang2019physpetree}{}%
-Fang, Y., Liu, C., Lin, J., Li, X., Alavian, K.N., Yang, Y.
-\& Niu, Y.
-(2019).
-PhySpeTree: An automated pipeline for reconstructing phylogenetic species trees.
-\emph{BMC evolutionary biology}, \textbf{19}, 1--8.
-
-\leavevmode\hypertarget{ref-fan2015assembly}{}%
-Fan, H., Ives, A.R., Surget-Groba, Y.
-\& Cannon, C.H.
-(2015).
-An assembly and alignment-free method of phylogeny reconstruction from next-generation sequencing data.
-\emph{BMC genomics}, \textbf{16}, 522.
-
-\leavevmode\hypertarget{ref-faurby2016all}{}%
-Faurby, S., Eiserhardt, W.L., Baker, W.J.
-\& Svenning, J.-C.
-(2016).
-An all-evidence species-level supertree for the palms (arecaceae).
-\emph{Molecular Phylogenetics and Evolution}, \textbf{100}, 57--69.
-
-\leavevmode\hypertarget{ref-faurby2015species}{}%
-Faurby, S.
-\& Svenning, J.-C.
-(2015).
-A species-level phylogeny of all extant and late quaternary extinct mammals using a novel heuristic-hierarchical bayesian approach.
-\emph{Molecular phylogenetics and evolution}, \textbf{84}, 14--26.
-
 \leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
@@
 
-\leavevmode\hypertarget{ref-flores2011estimating}{}%
-Flores, O. \& Coomes, D.A. (2011). Estimating the wood density of species for carbon stock assessments. \emph{Methods in Ecology and Evolution}, \textbf{2}, 214--220.
-
-\leavevmode\hypertarget{ref-frazao2015gondwana}{}%
-Frazao, A., Silva, H.R.
-da \& Moraes Russo, C.A.
-de.
-(2015).
-The gondwana breakup and the history of the atlantic and indian oceans unveils two new clades for early neobatrachian diversification.
-\emph{PloS one}, \textbf{10}.
-
-\leavevmode\hypertarget{ref-freyman2017phylogenetic}{}%
-Freyman, W.A.
-(2017).
-\emph{Phylogenetic models linking speciation and extinction to chromosome and mating system evolution}.
-PhD thesis thesis, UC Berkeley.
-Retrieved from \url{https://escholarship.org/uc/item/29n8r0nm}
-
-\leavevmode\hypertarget{ref-freyman2015sumac}{}%
-Freyman, W.A. (2015). SUMAC: Constructing phylogenetic supermatrices and assessing partially decisive taxon coverage. \emph{Evolutionary Bioinformatics}, \textbf{11}, EBO--S35384.
-
-\leavevmode\hypertarget{ref-gagnon2016new}{}%
-Gagnon, E., Bruneau, A., Hughes, C.E., Queiroz, L.P. de \& Lewis, G.P. (2016). A new generic system for the pantropical caesalpinia group (leguminosae). \emph{PhytoKeys}, 1.
-
-\leavevmode\hypertarget{ref-gagnon2016systematique}{}%
-Gagnon, E. \& others. (2016). Systématique et biogéographie du groupe caesalpinia (famille leguminosae). Retrieved from \url{http://hdl.handle.net/1866/13587}
-
-\leavevmode\hypertarget{ref-gao2011solution}{}%
-Gao, Y., Meng, Z., He, X., Liu, Y., Zhou, Y.
-\& Li, J.
-(2011).
-A solution to integrate data for phylogenetic research.
-\emph{2011 5th international conference on bioinformatics and biomedical engineering} pp.
-1--4.
-IEEE.
-
-\leavevmode\hypertarget{ref-garcia2016role}{}%
-García-Navas, V., Bonnet, T., Bonal, R.
-\& Postma, E.
-(2016).
-The role of fecundity and sexual selection in the evolution of size and sexual size dimorphism in new world and old world voles (rodentia: Arvicolinae).
-\emph{Oikos}, \textbf{125}, 1250--1260.
-
-\leavevmode\hypertarget{ref-gomez2020speciation}{}%
-Gómez-Bahamón, V., Márquez, R., Jahn, A.E., Miyaki, C.Y., Tuero, D.T., Laverde-R, O., Restrepo, S.
-\& Cadena, C.D.
-(2020).
-Speciation associated with shifts in migratory behavior in an avian radiation.
-\emph{Current Biology}.
-
-\leavevmode\hypertarget{ref-gomez2015behavioral}{}%
-Gómez Bahamón, V.
-\& others.
-(2015).
-\emph{A behavioral polymorphism as an intermediate stage in the evolution of divergent forms-partial migration in new world flycatchers (aves, tyrannidae)}.
-Master's thesis thesis, Bogotá-Uniandes.
-Retrieved from \url{https://repositorio.uniandes.edu.co/bitstream/handle/1992/12859/u703694.pdf?sequence=1}
-
-\leavevmode\hypertarget{ref-grant2014building}{}%
-Grant, J.R. \& Katz, L.A. (2014). Building a phylogenomic pipeline for the eukaryotic tree of life-addressing deep phylogenies with genome-scale data. \emph{PLoS currents}, \textbf{6}.
-
-\leavevmode\hypertarget{ref-greenberg2011caryophyllaceae}{}%
-Greenberg, A.K. \& Donoghue, M.J. (2011). Molecular systematics and character evolution in caryophyllaceae. \emph{Taxon}, \textbf{60}, 1637--1652.
-
-\leavevmode\hypertarget{ref-legume2013legume}{}%
-Group, L.P.W., Bruneau, A., Doyle, J.J., Herendeen, P., Hughes, C., Kenicer, G., Lewis, G., Mackinder, B., Pennington, R.T., Sanderson, M.J.
-\& others.
-(2013).
-Legume phylogeny and classification in the 21st century: Progress, prospects and lessons for other species--rich clades.
-\emph{Taxon}, \textbf{62}, 217--248.
-
-\leavevmode\hypertarget{ref-guy2017phyloskeleton}{}%
-Guy, L. (2017). PhyloSkeleton: Taxon selection, data retrieval and marker identification for phylogenomics. \emph{Bioinformatics}, \textbf{33}, 1230--1232.
-
-\leavevmode\hypertarget{ref-hardy2010gall}{}%
-Hardy, N.B. \& Cook, L.G. (2010). Gall-induction in insects: Evolutionary dead-end or speciation driver? \emph{BMC evolutionary biology}, \textbf{10}, 257.
-
-\leavevmode\hypertarget{ref-hardy2012testing}{}%
-Hardy, N.B. \& Cook, L.G. (2012). Testing for ecological limitation of diversification: A case study using parasitic plants. \emph{The American Naturalist}, \textbf{180}, 438--449.
-
-\leavevmode\hypertarget{ref-hardy2014specialization}{}%
-Hardy, N.B.
-\& Otto, S.P.
-(2014).
-Specialization and generalization in the diversification of phytophagous insects: Tests of the musical chairs and oscillation hypotheses.
-\emph{Proceedings of the Royal Society B: Biological Sciences}, \textbf{281}, 20132960.
-
-\leavevmode\hypertarget{ref-hardy2015evolution}{}%
-Hardy, N.B., Peterson, D.A. \& Dohlen, C.D. von. (2015). The evolution of life cycle complexity in aphids: Ecological optimization or historical constraint? \emph{Evolution}, \textbf{69}, 1423--1432.
-
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
 Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
-
-\leavevmode\hypertarget{ref-helmus2010communities}{}%
-Helmus, M.R., Keller, W., Paterson, M.J., Yan, N.D., Cannon, C.H.
-\& Rusak, J.A.
-(2010).
-Communities contain closely related species during ecosystem disturbance.
-\emph{Ecology letters}, \textbf{13}, 162--174.
-
-\leavevmode\hypertarget{ref-ickert2009fossil}{}%
-Ickert-Bond, S.M., Rydin, C.
-\& Renner, S.S.
-(2009).
-A fossil-calibrated relaxed clock for ephedra indicates an oligocene age for the divergence of asian and new world clades and miocene dispersal into south america.
-\emph{Journal of Systematics and Evolution}, \textbf{47}, 444--456.
-
-\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
-
-\leavevmode\hypertarget{ref-jamil2016visual}{}%
-Jamil, H.M. (2016). A visual interface for querying heterogeneous phylogenetic databases. \emph{IEEE/ACM transactions on computational biology and bioinformatics}, \textbf{14}, 131--144.
-
-\leavevmode\hypertarget{ref-jones2014trends}{}%
-Jones, T.M., Baxter, D.G., Hagedorn, G., Legler, B., Gilbert, E., Thiele, K., Vargas-Rodriguez, Y.
-\& Urbatsch, L.E.
-(2014).
-Trends in access of plant biodiversity data revealed by google analytics.
-\emph{Biodiversity data journal}.
-
-\leavevmode\hypertarget{ref-kamarudin2016phylogenetic}{}%
-Kamarudin, K.R., Rehan, A.M., Hashim, R., Usup, G.
-\& Rehan, M.M.
-(2016).
-Phylogenetic relationships within the genus holothuria inferred from 16S mitochondiral rRNA gene sequences.
-\emph{Sains Malaysiana}, \textbf{45}, 1079--1087.
-Retrieved from \url{http://journalarticle.ukm.my/9982/1/10\%20Kamarul\%20Rahim\%20Kamarudin.pdf}
-
-\leavevmode\hypertarget{ref-kolmann2017dna}{}%
-Kolmann, M.A., Elbassiouny, A.A., Liverpool, E.A. \& Lovejoy, N.R. (2017). DNA barcoding reveals the diversity of sharks in guyana coastal markets. \emph{Neotropical Ichthyology}, \textbf{15}.
-
-\leavevmode\hypertarget{ref-kumar2015bir}{}%
-Kumar, S., Krabberød, A.K., Neumann, R.S., Michalickova, K., Zhao, S., Zhang, X.
-\& Shalchian-Tabrizi, K.
-(2015).
-BIR pipeline for preparation of phylogenomic data.
-\emph{Evolutionary Bioinformatics}, \textbf{11}, EBO--S10189.
-
-\leavevmode\hypertarget{ref-lehtonen2011towards}{}%
-Lehtonen, S. (2011). Towards resolving the complete fern tree of life. \emph{PLoS One}, \textbf{6}.
-
-\leavevmode\hypertarget{ref-lehtonen2016sensitive}{}%
-Lehtonen, S., Christenhusz, M.J. \& Falck, D. (2016). Sensitive phylogenetics of clematis and its position in ranunculaceae. \emph{Botanical Journal of the Linnean Society}, \textbf{182}, 825--867.
-
-\leavevmode\hypertarget{ref-lehtonen2017environmentally}{}%
-Lehtonen, S., Silvestro, D., Karger, D.N., Scotese, C., Tuomisto, H., Kessler, M., Peña, C., Wahlberg, N.
-\& Antonelli, A.
-(2017).
-Environmentally driven extinction and opportunistic origination explain fern diversification patterns.
-\emph{Scientific Reports}, \textbf{7}, 1--12.
 
@@
 
-\leavevmode\hypertarget{ref-li2013partfasttree}{}%
-Li, J., Meng, Z., Hou, Y., Zhou, Y.
-\& Gao, Y.
-(2013).
-PartFastTree: Constructing large phylogenetic trees and estimating their reliability.
-\emph{2013 ninth international conference on natural computation (icnc)} pp.
-1052--1056.
-IEEE.
-
-\leavevmode\hypertarget{ref-mahmood2015avian}{}%
-Mahmood, M.T.
-(2015).
-\emph{Avian raptor evolution}.
-PhD thesis thesis, Institute of Fundamental Sciences, Massey University, New Zealand.
-Retrieved from \url{https://mro.massey.ac.nz/bitstream/handle/10179/7198/02_whole.pdf}
-
-\leavevmode\hypertarget{ref-martin2011trophic}{}%
-Martin, C.H.
-\& Wainwright, P.C.
-(2011).
-Trophic novelty is linked to exceptional rates of morphological diversification in two adaptive radiations of cyprinodon pupfish.
-\emph{Evolution: International Journal of Organic Evolution}, \textbf{65}, 2197--2212.
-
-\leavevmode\hypertarget{ref-mcmahon2015stbase}{}%
-McMahon, M.M., Deepak, A., Fernández-Baca, D., Boss, D. \& Sanderson, M.J. (2015). STBase: One million species trees for comparative biology. \emph{PloS one}, \textbf{10}.
-
 \leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
@@
 
-\leavevmode\hypertarget{ref-meng2015darwintree}{}%
-Meng, Z., Dong, H., Li, J., Chen, Z., Zhou, Y., Wang, X.
-\& Zhang, S.
-(2015a).
-Darwintree: A molecular data analysis and application environment for phylogenetic study.
-\emph{Data Science Journal}, \textbf{14}.
-
-\leavevmode\hypertarget{ref-meng2015solution}{}%
-Meng, Z., Li, J. \& Chen, Z. (2015b). A solution to phylogeny assembly for ecologists. \emph{2015 12th international conference on fuzzy systems and knowledge discovery (fskd)} pp. 1103--1107. IEEE.
-
-\leavevmode\hypertarget{ref-meng2015sotree}{}%
-Meng, Z., Li, J., Yang, T., Lin, L.
-\& Chen, Z.
-(2015c).
-SoTree: An automated phylogeny assembly tool for ecologists from big tree.
-\emph{2015 ieee international conference on smart city/socialcom/sustaincom (smartcity)} pp.
-792--797.
-IEEE.
-
-\leavevmode\hypertarget{ref-meng2012gsqct}{}%
-Meng, Z., Li, J., Zhou, Y., Cao, W., Xiao, X., Zhao, J., Dong, H.
-\& Zhang, S.
-(2012a).
-GSQCT: A solution to screening gene sequences for phylogenetics analysis.
-\emph{2012 9th international conference on fuzzy systems and knowledge discovery} pp.
-2929--2933.
-IEEE.
-
-\leavevmode\hypertarget{ref-meng2014rapidtree}{}%
-Meng, Z., Shao, J., Cao, W., Li, J., Zhou, Y.
-\& Wang, X.
-(2014).
-RapidTree: A solution to rapid reconstruction phylogenetic tree.
-\emph{2014 11th international conference on fuzzy systems and knowledge discovery (fskd)} pp.
-513--517.
-IEEE.
-
-\leavevmode\hypertarget{ref-meng2012cloud}{}%
-Meng, Z., Xiao, X., Li, J., Zhou, Y., Cao, W.
-\& Shen, G.
-(2012b).
-Cloud-gsqct: A parallel approach to screen gene sequences for phylogenetics analysis.
-\emph{2012 international conference on computer science and information processing (csip)} pp.
-660--663.
-IEEE.
-
-\leavevmode\hypertarget{ref-morrison2009apicomplexa}{}%
-Morrison, D.A. (2009). Evolution of the apicomplexa: Where are we now? \emph{Trends in parasitology}, \textbf{25}, 375--382.
-
-\leavevmode\hypertarget{ref-nazaire2012broad}{}%
-Nazaire, M. \& Hufford, L. (2012). A broad phylogenetic analysis of boraginaceae: Implications for the relationships of mertensia. \emph{Systematic Botany}, \textbf{37}, 758--783.
-
-\leavevmode\hypertarget{ref-page2013bionames}{}%
-Page, R.D. (2013). BioNames: Linking taxonomy, texts, and trees. \emph{PeerJ}, \textbf{1}, e190.
-
-\leavevmode\hypertarget{ref-page2011linking}{}%
-Page, R.D. (2011). Linking ncbi to wikipedia: A wiki-based approach. \emph{PLoS currents}, \textbf{3}.
-
-\leavevmode\hypertarget{ref-papadopoulou2015automated}{}%
-Papadopoulou, A., Chesters, D., Coronado, I., De la Cadena, G., Cardoso, A., Reyes, J.C., Maes, J.-M., Rueda, R.M.
-\& Gómez-Zurita, J.
-(2015).
-Automated dna-based plant identification for large-scale biodiversity assessment.
-\emph{Molecular ecology resources}, \textbf{15}, 136--152.
-
-\leavevmode\hypertarget{ref-pereira2013padroes}{}%
-Pereira, J.E.S.
-(2013).
-\emph{Padrões e processos na evolução de primatas neotropicais (platyrrhini, primates)}.
-PhD thesis thesis, Tese de doutorado.
-Setor de Ciências Biológicas, Universidade Federal do~\ldots{}.
-Retrieved from \url{https://www.acervodigital.ufpr.br/handle/1884/33775}
-
-\leavevmode\hypertarget{ref-peters2011taming}{}%
-Peters, R.S., Meyer, B., Krogmann, L., Borner, J., Meusemann, K., Schütte, K., Niehuis, O.
-\& Misof, B.
-(2011).
-The taming of an impossible child: A standardized all-in approach to the phylogeny of hymenoptera using public database sequences.
-\emph{BMC biology}, \textbf{9}, 55.
-Retrieved from \url{https://bmcbiol.biomedcentral.com/articles/10.1186/1741-7007-9-55\#Sec21}
-
 \leavevmode\hypertarget{ref-piel2009treebase}{}%
 Piel, W., Chan, L., Dominus, M., Ruan, J., Vos, R. \& Tannen, V. (2009). Treebase v. 2: A database of phylogenetic knowledge. E-biosphere.
-
-\leavevmode\hypertarget{ref-rabosky2011species}{}%
-Rabosky, D.L., Cowan, M.A., Talaba, A.L.
-\& Lovette, I.J.
-(2011).
-Species interactions mediate phylogenetic community structure in a hyperdiverse lizard assemblage from arid australia.
-\emph{The American Naturalist}, \textbf{178}, 579--595.
 
@@
 
-\leavevmode\hypertarget{ref-roalson2016distinct}{}%
-Roalson, E.H. \& Roberts, W.R. (2016). Distinct processes drive diversification in different clades of gesneriaceae. \emph{Systematic Biology}, \textbf{65}, 662--684.
-
-\leavevmode\hypertarget{ref-roquet2013building}{}%
-Roquet, C., Thuiller, W. \& Lavergne, S. (2013). Building megaphylogenies for macroecology: Taking up the challenge. \emph{Ecography}, \textbf{36}, 13--26.
-
-\leavevmode\hypertarget{ref-ross2013large}{}%
-Ross, L., Hardy, N.B., Okusu, A.
-\& Normark, B.B.
-(2013).
-Large population size predicts the distribution of asexuality in scale insects.
-\emph{Evolution: International Journal of Organic Evolution}, \textbf{67}, 196--206.
-
-\leavevmode\hypertarget{ref-ruiz2019datataxa}{}%
-Ruiz-Sanchez, E., Maya-Lastra, C.A., Steinmann, V.W., Zamudio, S., Carranza, E., Murillo, R.M.
-\& Rzedowski, J.
-(2019).
-Datataxa: A new script to extract metadata sequence information from genbank, the flora of bajío as a case study.
-\emph{Botanical Sciences}, \textbf{97}, 754--760.
-
-\leavevmode\hypertarget{ref-salariato2010filogenia}{}%
-Salariato, D.L.
-(2010).
-\emph{Filogenia y evolución de la subtribu melinidinae (paniceae: Panicoideae: Poaceae)}.
-PhD thesis thesis, Universidad de Buenos Aires.
-Facultad de Ciencias Exactas y Naturales.
-Retrieved from \url{http://hdl.handle.net/20.500.12110/tesis_n4771_Salariato}
-
-\leavevmode\hypertarget{ref-salariato2010molecular}{}%
-Salariato, D.L., Zuloaga, F.O., Giussani, L.M.
-\& Morrone, O.
-(2010).
-Molecular phylogeny of the subtribe melinidinae (poaceae: Panicoideae: Paniceae) and evolutionary trends in the homogenization of inflorescences.
-\emph{Molecular Phylogenetics and Evolution}, \textbf{56}, 355--369.
-
-\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A.
-\& Wehe, A.
-(2008).
-The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research.
-\emph{Systematic Biology}, \textbf{57}, 335--346.
-Retrieved from \url{https://doi.org/10.1080/10635150802158688}
-
-\leavevmode\hypertarget{ref-sanderson2010phylogenomics}{}%
-Sanderson, M.J., McMahon, M.M. \& Steel, M. (2010). Phylogenomics with incomplete taxon coverage: The limits to inference. \emph{BMC Evolutionary Biology}, \textbf{10}, 155.
-
-\leavevmode\hypertarget{ref-sanderson2016perspective}{}%
-Sanderson, M.J., Olson, P., Hughes, J.
-\& Cotton, J.
-(2016).
-Perspective: Challenges in assembling the `next generation'Tree of life.
-\emph{Olson PD, Hughes J and Cotton JA}, 13--27.
-Retrieved from \url{https://books.google.com/books?hl=es\&lr=\&id=hFU2DAAAQBAJ\&oi=fnd\&pg=PA13\&ots=2Oa_TVGCic\&sig=fdxyYmSQASm2XFoU7jv4JKLFaTQ\#v=onepage\&q\&f=false}
-
 \leavevmode\hypertarget{ref-san2010molecular}{}%
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
-
-\leavevmode\hypertarget{ref-santini2014first}{}%
-Santini, F., Carnevale, G. \& Sorenson, L. (2014). First multi-locus timetree of seabreams and porgies (percomorpha: Sparidae). \emph{Italian Journal of Zoology}, \textbf{81}, 55--71.
-
-\leavevmode\hypertarget{ref-santini2015first}{}%
-Santini, F., Carnevale, G.
-\& Sorenson, L.
-(2015).
-First timetree of sphyraenidae (percomorpha) reveals a middle eocene crown age and an oligo--miocene radiation of barracudas.
-\emph{Italian Journal of Zoology}, \textbf{82}, 133--142.
-
-\leavevmode\hypertarget{ref-santini2013first}{}%
-Santini, F.
-\& Sorenson, L.
-(2013).
-First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies.
-\emph{Italian journal of zoology}, \textbf{80}, 481--489.
-
-\leavevmode\hypertarget{ref-sarkinen2013solanaceae}{}%
-Särkinen, T., Bohs, L., Olmstead, R.G.
-\& Knapp, S.
-(2013).
-A phylogenetic framework for evolutionary study of the nightshades (solanaceae): A dated 1000-tip tree.
-\emph{BMC evolutionary biology}, \textbf{13}, 214.
 
@@
 
-\leavevmode\hypertarget{ref-selvatti2015paleogene}{}%
-Selvatti, A.P., Gonzaga, L.P.
-\& Moraes Russo, C.A.
-de.
-(2015).
-A paleogene origin for crown passerines and the diversification of the oscines in the new world.
-\emph{Molecular phylogenetics and evolution}, \textbf{88}, 1--15.
-
-\leavevmode\hypertarget{ref-shi2015speciation}{}%
-Shi, J.J. \& Rabosky, D.L. (2015). Speciation dynamics during the global radiation of extant bats. \emph{Evolution}, \textbf{69}, 1528--1545.
-
-\leavevmode\hypertarget{ref-simon2012cerrado}{}%
-Simon, M.F. \& Pennington, T. (2012). Evidence for adaptation to fire regimes in the tropical savannas of the brazilian cerrado. \emph{International Journal of Plant Sciences}, \textbf{173}, 711--723.
-
-\leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith, S.A., Beaulieu, J.M.
-\& Donoghue, M.J.
-(2009).
-Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
-\emph{BMC evolutionary biology}, \textbf{9}, 37.
-
-\leavevmode\hypertarget{ref-smith2018constructing}{}%
-Smith, S.A. \& Brown, J.W. (2018). Constructing a broadly inclusive seed plant phylogeny. \emph{American Journal of Botany}, \textbf{105}, 302--314.
-
-\leavevmode\hypertarget{ref-smith2019pyphlawd}{}%
-Smith, S.A. \& Walker, J.F. (2019). PyPHLAWD: A python tool for phylogenetic dataset construction. \emph{Methods in Ecology and Evolution}, \textbf{10}, 104--108.
-
-\leavevmode\hypertarget{ref-sorenson2014evolution}{}%
-Sorenson, L.
-(2014).
-\emph{Evolution of marine fish biodiversity: Phylogenomics and ecological processes shaping diversification}.
-PhD thesis thesis, UCLA.
-Retrieved from \url{https://escholarship.org/uc/item/31n0c9km}
-
-\leavevmode\hypertarget{ref-sorenson2014effect}{}%
-Sorenson, L., Santini, F. \& Alfaro, M. (2014). The effect of habitat on modern shark diversification. \emph{Journal of Evolutionary Biology}, \textbf{27}, 1536--1548.
-
 \leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
 Stamatakis, A. (2014). RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. \emph{Bioinformatics}, \textbf{30}, 1312--1313.
-
-\leavevmode\hypertarget{ref-steppan2017muroid}{}%
-Steppan, S.J. \& Schenk, J.J. (2017). Muroid rodent phylogenetics: 900-species tree reveals increasing diversification rates. \emph{PLoS One}, \textbf{12}.
 
@@
 
-\leavevmode\hypertarget{ref-strecker2014fish}{}%
-Strecker, A.L.
-\& Olden, J.D.
-(2014).
-Fish species introductions provide novel insights into the patterns and drivers of phylogenetic structure in freshwaters.
-\emph{Proceedings of the Royal Society B: Biological Sciences}, \textbf{281}, 20133003.
-
-\leavevmode\hypertarget{ref-thomson2010sparse}{}%
-Thomson, R.C.
-\& Shaffer, H.B.
-(2010).
-Sparse supermatrices for phylogenetic inference: Taxonomy, alignment, rogue taxa, and the phylogeny of living turtles.
-\emph{Systematic biology}, \textbf{59}, 42--58.
-
-\leavevmode\hypertarget{ref-tucker2012incorporating}{}%
-Tucker, C.M., Cadotte, M.W., Davies, T.J.
-\& Rebelo, T.G.
-(2012).
-Incorporating geographical and evolutionary rarity into conservation prioritization.
-\emph{Conservation Biology}, \textbf{26}, 593--601.
-
-\leavevmode\hypertarget{ref-verbruggen2010data}{}%
-Verbruggen, H., Maggs, C.A., Saunders, G.W., Le Gall, L., Yoon, H.S.
-\& De Clerck, O.
-(2010).
-Data mining approach identifies research priorities and data requirements for resolving the red algal tree of life.
-\emph{BMC evolutionary biology}, \textbf{10}, 16.
-
 \leavevmode\hypertarget{ref-vos2012nexml}{}%
@@
 
-\leavevmode\hypertarget{ref-weber2014defense}{}%
-Weber, M.G. \& Agrawal, A.A. (2014). Defense mutualisms enhance plant diversification. \emph{Proceedings of the National Academy of Sciences}, \textbf{111}, 16442--16447.
-
 \leavevmode\hypertarget{ref-wheeler2000database}{}%
@@
 
-\leavevmode\hypertarget{ref-wu2008simple}{}%
-Wu, M. \& Eisen, J.A. (2008). A simple, fast, and accurate method of phylogenomic inference. \emph{Genome biology}, \textbf{9}, R151.
-
-\leavevmode\hypertarget{ref-xu2015ncbiminer}{}%
-Xu, X., Dimitrov, D., Rahbek, C. \& Wang, Z. (2015). NCBIminer: Sequences harvest from genbank. \emph{Ecography}, \textbf{38}, 426--430.
-
-\leavevmode\hypertarget{ref-yong2010screening}{}%
-Yong, L., Zhen, M., Qi, L., Yanping, G., Yuanchun, Z.
-\& Jianhui, L.
-(2010).
-Screening data for phylogenetic analysis of land plants: A parallel approach.
-\emph{2010 first international conference on networking and distributed computing} pp.
-305--308.
-IEEE.
-
 \end{document}

### 2020-06-25 render2-submission-mee.Rmd
examples

@@
 ```
-
-1. `r params$author1`
+**1. `r params$author1`**
 
@@
 
-2. `r params$author2`
+**2. `r params$author2`**
 
@@
 
-3. `r params$author3`
+**3. `r params$author3`**
 
@@
 
-Correspondence:
+**Correspondence address**: `r params$correspondence_address`
 
-`r params$correspondence_address`
+**Correspondence email**: <`r params$email1`>, <`r params$email3`>
 
-email: <`r params$email1`>, <`r params$email3`>
+**Running title**: `r params$running_title`
 
-Running title: `r params$running_title`
+**Word count**: `r word_count_body`
 
-Word count: `r word_count_body`
+**Manuscript prepared for submission to Methods in Ecology and Evolution**
 
-Manuscript prepared for submission to Methods in Ecology and Evolution
-
-Article type: Application
+**Article type**: Application
 
@@
 
-**Keywords**: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, update alignment
+**Keywords**: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
 

### 2020-06-25 2-body.Rmd
intro clean up

@@
 ---
-title: 'Physcraper: A Python package for continually updated gene trees'
+title: 'Physcraper: Body'
 author:
@@
 
-Phylogenies are important.
+From molecular data to alignments and phylogenies, public databases such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life data deposition system, the phylesystem [@mctavish2015phylesystem] are still accumulating data.
 
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-***paragraph better to end discussion???
-***
+More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about `r round(236023/240708, digits=4)*100`% of taxa included in the NCBI taxonomy [@sanderson2008phylota; @federhen2003taxonomy] which represents about 10% of described diversity [@scott2011ncbi], taking a conservative estimate of extant diversity.
+In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences <https://ftp.ncbi.nih.gov/genbank/gbrel.txt>.
 
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
+Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public and private databases.
+Most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
+Tools that automatize the assembly of homology hypotheses for phylogenetics such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward] have been widely used to assemble genetic datasets to reconstruct phylogenetic relationships among different organisms.
 
-1. Obtention of molecular or morphological character data -- get DNA from some organisms
-and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
-1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
-aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-1. Analyse this hypothesis of homology to infer phylogenetic relationships among
-the organisms you are studying -- Use different available programs to infer molecular
-evolution, trees and times of divergence.
-1. Discuss the inferred relationships in the context of previous hypothesis, the
-biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
+Considering that 400 phylogenies are published each year, most published phylogenies have been infered from a homology hypothesis that has been constructed "by hand".
 
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
+On one hand, there are concerns about the use of automated tools.
+It is suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
 
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
+Indeed, public manually curated homology hypotheses have the potential to be enriched by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
 
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
+There are THIS MANY curated alignments publicly available on TreeBASE
 
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
+This suggests a lot of potential for phylogenetic update of different regions of th eopen tree of life.
 
@@
 
+A postdoc started working with a new reserach group. They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species.
+A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
+However, the head of the group wants to use a particular marker they beilieve to be the best one to be able to solve the relationships in the group.
+They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating.
+
+![]("https://github.com/McTavishLab/physcraperex/raw/master/vignettes/malvaceae-images/cotree-plot2-1.png")
+
 # Discussion

### 2020-06-25 2.1-intro-ideas.Rmd
intro clean up

@@
+The PHYLOTA [@sanderson2008phylota] tool has been used to construct homolog DNA datasets for about 42 studies reconstructing phylogenetic trees from various organisms.
+
+In face of the science reproducibility crisis, more and more researchers are contributing to making data available, contributing to a continual growth in information available from different databases.
+
 It is well known that GenBank [@benson2000genbank; @wheeler2000database] holds enormous amounts of genetic data, and it continues to grow. **Too cliché**
@@
 *** Introduce the Open Tree of Life***
+
+In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
+pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
+to obtain homologous characters that can be used for phylogenetic reconstruction.
+These tools have been presented as aid for the nonspecialist to decrease some of
+the difficulties in the generation of phylogenetic knowledge. However, they are
+not that often used as so, suggesting that there are still difficulties for the nonspecialist.
+
+The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
+However, automatizing the assembly of the character data set is a crucial step towards
+reproducibility for a task that was otherwise primarily artisanal and hence largely
+non-reproducible.
+
+Phylogenies are important.
+
+Generating phylogenies is not easy and it is largely artisanal.
+Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
+***paragraph better to end discussion???
+***
+
+The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
+
+1. Obtention of molecular or morphological character data -- get DNA from some organisms
+and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
+1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
+aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
+1. Analyse this hypothesis of homology to infer phylogenetic relationships among
+the organisms you are studying -- Use different available programs to infer molecular
+evolution, trees and times of divergence.
+1. Discuss the inferred relationships in the context of previous hypothesis, the
+biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
+
+Each of these steps require different types of specialized training: in the field,
+in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
+All of these steps also require considerable amounts of time for training and implementation.
+
+Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
+Phylogenies acan be used to solve issues such as food security, global warming, global health, all crucial for our well-being.
+There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
+In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
+
+Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].

### 2020-06-25 99-tools-review.Rmd
intro clean up

@@
 
-### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
+### 1. Mining databases to generate DNA datasets suitable for phylogenetic reconstruction
 
@@
 | Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
 | PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
@@
 
-### 2. Searching phylogenetic tree databases
+### 2. Mining databases to generate protein/aminoacid (AA) datasets suitable for phylogenetic reconstruction
+| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
+
+### 3. Searching phylogenetic tree databases
 
@@
 
-### 3. Mining phylogenetic tree databases
+### 4. Mining phylogenetic tree databases
 
@@
 
-### 4. Pipeline for phylogenetic reconstruction
+### 5. Pipeline for phylogenetic reconstruction
 

### 2020-06-25 docs/submission-mee-2020-06-25.tex
intro clean up

@@
 \usepackage[margin = 1in]{geometry}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs}
@@
 
-\textbf{Word count}: 3246
+\textbf{Word count}: 3056
 
-Manuscript prepared for submission to Methods in Ecology and Evolution
+\textbf{Manuscript prepared for submission to Methods in Ecology and Evolution}
 
-Article type: Application
+\textbf{Article type}: Application
 
@@
 
-\textbf{Keywords}: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, update alignment
+\textbf{Keywords}: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
 
@@
 
+We will address two use-case scenarios. One in which the user is interested in a particular group. Anothr one in which the user is interested in a particular phylogeny.
+
 \hypertarget{the-hollies}{%
@@
 
-The genus \emph{Ilex} is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-\textbf{\emph{Contact authors? They seem old school, probably do not wanna share their data.}}
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
+A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
+It encompasses between 400-600 living species. A review of literature shows them that there are three published phylogenetic trees, showing relationships within the hollies.
+The first one has been made available both in the OToL phylesystem and is part of the synthetic tree, and on treeBASE. It samples 48 species.
+The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
+The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
 This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
-Updating the oldest tree, we should get something very similar to the newest tree.
+We would expect that updating the oldest tree, we would get something very similar to the newest tree.
 
-\hypertarget{the-ascomycota}{%
-\subsection{The Ascomycota}\label{the-ascomycota}}
+\hypertarget{the-malvaceae}{%
+\subsection{The Malvaceae}\label{the-malvaceae}}
 
-Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by Schoch \emph{et al.} (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
-The first step, is to get the Open Tree of Life study id. There are some options to do this:
-- You can go to the Open Tree of Life website and browse until you find it, or
-- you can get the study id using R tools:
-- By using the TreeBase ID of the study (which is not fully exposed on the
-TreeBase website home page of the study, so you have to really look it up manually):
+A postdoc started working with a new reserach group. They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species.
+A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
+However, the head of the group wants to use a particular marker they beilieve to be the best one to be able to solve the relationships in the group.
+They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property =} \StringTok{"treebaseId"}\NormalTok{, }\DataTypeTok{value =} \StringTok{"S2137"}\NormalTok{)}
-\CommentTok{##   study_ids n_trees tree_ids candidate study_year title}
-\CommentTok{## 1    pg_238       1  tree109                 2009      }
-\CommentTok{##                                 study_doi}
-\CommentTok{## 1 http://dx.doi.org/10.1093/sysbio/syp020}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{itemize}
-\tightlist
-\item
-  By using the name of the focal clade of study (but this behaved very differently):
-\end{itemize}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{studies_find_studies}\NormalTok{(}\DataTypeTok{property=}\StringTok{"ot:focalCladeOTTTaxonName"}\NormalTok{, }\DataTypeTok{value=}\StringTok{"Ascomycota"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Once we have the study id, we can gather the trees published on that study:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_tree_ids}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\CommentTok{## [1] "tree109"}
-\NormalTok{rotl}\OperatorTok{::}\KeywordTok{candidate_for_synth}\NormalTok{(rotl}\OperatorTok{::}\KeywordTok{get_study_meta}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{))}
-\CommentTok{## NULL}
-\NormalTok{my_trees <-}\StringTok{ }\NormalTok{rotl}\OperatorTok{::}\KeywordTok{get_study}\NormalTok{(}\StringTok{"pg_238"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Both trees from this study have NA tips.
-
-Let's check what one of the trees looks like:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Download the alignment from TreeBase
-  If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
-  To make this task easier, you can use a command to download everything into your working folder:
-\end{enumerate}
-
-\begin{verbatim}
-physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
-\end{verbatim}
-
-In this example, all alignments posted on TreeBase were used to reconstruct both trees.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  With the study id and the alignment files saved locally, we can do a physcraper run with the command:
-\end{enumerate}
-
-\begin{verbatim}
-physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
-\end{verbatim}
-
-\hypertarget{testudines-example}{%
-\subsection{Testudines example}\label{testudines-example}}
-
-Phylogeny of the Testudines 6 tips from Crawford \emph{et al.} (\protect\hyperlink{ref-crawford2012more}{2012})
-There is just one tree in OToL.
-There is just one alignment on \href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase} with all the 1 145 loci.
-
-\begin{verbatim}
-physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
-\end{verbatim}
+\includegraphics{"https://github.com/McTavishLab/physcraperex/raw/master/vignettes/malvaceae-images/cotree-plot2-1.png"}
 
@@
 
-\leavevmode\hypertarget{ref-crawford2012more}{}%
-Crawford, N.G., Faircloth, B.C., McCormack, J.E., Brumfield, R.T., Winker, K.
-\& Glenn, T.C.
-(2012).
-More than 1000 ultraconserved elements provide evidence that turtles are the sister group of archosaurs.
-\emph{Biology letters}, \textbf{8}, 783--786.
-
 \leavevmode\hypertarget{ref-deepak2014evominer}{}%
@@
 
-\leavevmode\hypertarget{ref-schoch2009ascomycota}{}%
-Schoch, C.L., Sung, G.-H., López-Giráldez, F., Townsend, J.P., Miadlikowska, J., Hofstetter, V., Robbertse, B., Matheny, P.B., Kauff, F., Wang, Z.
-\& others.
-(2009).
-The ascomycota tree of life: A phylum-wide phylogeny clarifies the origin and evolution of fundamental reproductive and ecological traits.
-\emph{Systematic biology}, \textbf{58}, 224--239.
-
 \leavevmode\hypertarget{ref-stamatakis2014raxml}{}%

### 2020-06-26 2-body.Rmd
figures

@@
 On one hand, there are concerns about the use of automated tools.
-It is suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
+Concerns I think people have about these tools:
+- Errors in identification of sequences
+- Little control along the process
+- Too much of a black box?
+Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
+-
+It has also been suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
 
-Indeed, public manually curated homology hypotheses have the potential to be enriched by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
+Fortunately, public manually curated homology hypotheses have the potential to be enriched and updated by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
 
@@
 
-This suggests a lot of potential for phylogenetic update of different regions of th eopen tree of life.
+This suggests a lot of potential for update of phylogenetic relationships in different regions of the tree of life.
+
+OToL has integrated the phylogenies fro TreeBASe, and linked to the alignment repo but not the exact alignment that generated the phylogeny.
+
+Linkage of original alignment with corresponding phylogeny has t be done by a human curator.
 
@@
 
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control along the process
-- Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
 The pipelines are so powerful and they will give you an answer, but there
@@
 better because it used more data.
-
-All these pipelines start tree construction from zero? Yes.
 
@@
 are provided by the users.
-
-The original tree is by default used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
 
@@
   Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+- The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
 - Bootstrap results are summarized with Dendropy ADD CITATION
@@
 
-![]("https://github.com/McTavishLab/physcraperex/raw/master/vignettes/malvaceae-images/cotree-plot2-1.png")
+```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
+wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/vignettes/malvaceae-images/cotree-plot2-1.png"
+mv cotree-plot2-1.png docs/figs/cotree-plot2-1.png
+```
+```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and update tree of the Malvaceae.", out.width = '75%', fig.retina= 2}
+knitr::include_graphics("docs/figs/cotree-plot2-1.png")
+```
+\newpage
 

### 2020-06-26 docs/submission-mee-2020-06-25.tex
figures

@@
 
-\textbf{Word count}: 3056
+\textbf{Word count}: 2969
 
@@
 
-Phylogenies are important.
+From molecular data to alignments and phylogenies, public databases such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life data deposition system, the phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) are still accumulating data.
 
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-\textbf{\emph{paragraph better to end discussion???
-}}
+More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about 98.05\% of taxa included in the NCBI taxonomy (Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}) which represents about 10\% of described diversity (Scott \protect\hyperlink{ref-scott2011ncbi}{2011}), taking a conservative estimate of extant diversity.
+In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \url{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}.
 
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
+Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public and private databases.
+Most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
+Tools that automatize the assembly of homology hypotheses for phylogenetics such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}) have been widely used to assemble genetic datasets to reconstruct phylogenetic relationships among different organisms.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Obtention of molecular or morphological character data -- get DNA from some organisms
-  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
-\item
-  Assemble a hypothesis of homology -- Create a matrix of your character data, by
-  aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-\item
-  Analyse this hypothesis of homology to infer phylogenetic relationships among
-  the organisms you are studying -- Use different available programs to infer molecular
-  evolution, trees and times of divergence.
-\item
-  Discuss the inferred relationships in the context of previous hypothesis, the
-  biology and biogeography of the organisms, etc. -- Answer the question, \emph{is this phylogenetic solution fair/reasonable?}
-\end{enumerate}
+Considering that 400 phylogenies are published each year, most published phylogenies have been infered from a homology hypothesis that has been constructed ``by hand''.
 
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
+On one hand, there are concerns about the use of automated tools.
+Concerns I think people have about these tools:
+- Errors in identification of sequences
+- Little control along the process
+- Too much of a black box?
+Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
+-
+It has also been suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
 
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}),
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
+Fortunately, public manually curated homology hypotheses have the potential to be enriched and updated by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
 
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
+There are THIS MANY curated alignments publicly available on TreeBASE
 
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
+This suggests a lot of potential for update of phylogenetic relationships in different regions of the tree of life.
+
+OToL has integrated the phylogenies fro TreeBASe, and linked to the alignment repo but not the exact alignment that generated the phylogeny.
+
+Linkage of original alignment with corresponding phylogeny has t be done by a human curator.
 
@@
 \item
+  The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
+\item
   Bootstrap results are summarized with Dendropy ADD CITATION
@@
 
-\includegraphics{"https://github.com/McTavishLab/physcraperex/raw/master/vignettes/malvaceae-images/cotree-plot2-1.png"}
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{docs/figs/cotree-plot2-1} \caption{Comparison of original tree and update tree of the Malvaceae.}\label{fig:fig-2}
+\end{figure}
 
@@
 
+\leavevmode\hypertarget{ref-federhen2003taxonomy}{}%
+Federhen, S. (2003). The taxonomy project. \emph{The NCBI Handbook}.
+
 \leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
@@
 
+\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
+Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A.
+\& Wehe, A.
+(2008).
+The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research.
+\emph{Systematic Biology}, \textbf{57}, 335--346.
+Retrieved from \url{https://doi.org/10.1080/10635150802158688}
+
 \leavevmode\hypertarget{ref-san2010molecular}{}%
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
+
+\leavevmode\hypertarget{ref-scott2011ncbi}{}%
+Scott, F. (2011). The ncbi taxonomy database. \emph{Nucleic Acids Research}, \textbf{40}, D136--D14.
+
+\leavevmode\hypertarget{ref-smith2009mega}{}%
+Smith, S.A., Beaulieu, J.M.
+\& Donoghue, M.J.
+(2009).
+Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
+\emph{BMC evolutionary biology}, \textbf{9}, 37.
 

### 2020-06-26 2-body.Rmd
example

@@
 
-We will address two use-case scenarios. One in which the user is interested in a particular group. Anothr one in which the user is interested in a particular phylogeny.
+To exemplify the utility of physcraper we will address two use-case scenarios.
+One in which the user is interested in a particular group.
+Another one in which the user is interested in a particular phylogeny.
+A tutrial as well as illustrated examples of steps needed to perform a physcraper analyses are available elsewhere.
 
@@
 A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of literature shows them that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both in the OToL phylesystem and is part of the synthetic tree, and on treeBASE. It samples 48 species.
+The genus encompasses between 400-600 living species. A review of literature reveals that there are three phylogenetic trees showing relationships within the hollies published.
+The first one has been made available in TreeBASE as well as in the OToL phylesystem and is part of the synthetic tree. It samples 48 species.
 The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
@@
 The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
-We would expect that updating the oldest tree, we would get something very similar to the newest tree.
+This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
 
@@
 However, the head of the group wants to use a particular marker they beilieve to be the best one to be able to solve the relationships in the group.
-They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating.
+They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating and that they trust.
 

### 2020-06-26 2-body.Rmd
more intro

@@
 
-From molecular data to alignments and phylogenies, public databases such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life data deposition system, the phylesystem [@mctavish2015phylesystem] are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life (OToL) data deposition tool, the phylesystem [@mctavish2015phylesystem], are still accumulating data.
 
-More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about `r round(236023/240708, digits=4)*100`% of taxa included in the NCBI taxonomy [@sanderson2008phylota; @federhen2003taxonomy] which represents about 10% of described diversity [@scott2011ncbi], taking a conservative estimate of extant diversity.
+More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about `r round(236023/240708, digits=4)*100`% of taxa included in the NCBI taxonomy [@sanderson2008phylota] which represents about 10% of described diversity [@scott2011ncbi; @federhen2003taxonomy], taking a conservative estimate of extant diversity.
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences <https://ftp.ncbi.nih.gov/genbank/gbrel.txt>.
@@
 Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
-Tools that automatize the assembly of homology hypotheses for phylogenetics such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward] have been widely used to assemble genetic datasets to reconstruct phylogenetic relationships among different organisms.
+Tools that automatize the assembly of homology hypotheses from the GenBAnk database for phylogenetics such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward] have been widely used to reconstruct the evolutionary relationships among different organisms.
 
-Considering that 400 phylogenies are published each year, most published phylogenies have been infered from a homology hypothesis that has been constructed "by hand".
+This has been an important contribution to the field of phylogenetics, especially bc they represent a possible way for reproducibility in phylogenetics.
+Indeed, most published phylogenies have been infered from a homology hypothesis that has been curated "by hand" [@morrison2009would].
 

### 2020-06-26 2-body.Rmd
more intro

--- 2020-06-26 2-body.Rmd
@@
 
-From molecular data to alignments and phylogenies, public databases such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life data deposition system, the phylesystem [@mctavish2015phylesystem] are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data.
 
-More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about `r round(236023/240708, digits=4)*100`% of taxa included in the NCBI taxonomy [@sanderson2008phylota; @federhen2003taxonomy] which represents about 10% of described diversity [@scott2011ncbi], taking a conservative estimate of extant diversity.
+More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of most of the taxa represented [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota], covering about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences <https://ftp.ncbi.nih.gov/genbank/gbrel.txt>.
 
-Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public and private databases.
-Most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
-Tools that automatize the assembly of homology hypotheses for phylogenetics such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward] have been widely used to assemble genetic datasets to reconstruct phylogenetic relationships among different organisms.
+Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public databases such as GenBank, as well as private databases.
+Generally refered to as "pipelines", most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses (aka molecular alignments) to reconstruct phylogenetic relationships.
+Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (refered to as "phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely used to study the evolutionary relationships among different organisms (TABLE??
+maybe supplementary data), from a phylogeny of no more than 20 fish species of the family of barracudas [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 living fern species [@lehtonen2011towards].
 
-Considering that 400 phylogenies are published each year, most published phylogenies have been infered from a homology hypothesis that has been constructed "by hand".
-
-On one hand, there are concerns about the use of automated tools.
+Pipelines have been an important incorporation to the field of phylogenetics, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies have been infered using homology hypotheses that have been curated "by hand" [@morrison2009would].
+What explains this preference?
+aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
+On one hand, there are concerns about the use of automated pipelines in phylogenetics.
 Concerns I think people have about these tools:
 - Errors in identification of sequences
-- Little control along the process
+- Little control at different steps of the process
 - Too much of a black box?
@@
 -
-It has also been suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
+It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions [@fragoso2017pilot].
 
-Fortunately, public manually curated homology hypotheses have the potential to be enriched and updated by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
+An way to incorporate the best of two worlds (newly released molecular data with fine knowledge from human experts) would be to rely on published manually curated homology hypotheses that have been made publicly available.
+This expert-curated alignments can be enriched and updated by incorporating newly released data from public molecular databases.
 
@@
 ```
-```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and update tree of the Malvaceae.", out.width = '75%', fig.retina= 2}
+```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and update tree of the Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 knitr::include_graphics("docs/figs/cotree-plot2-1.png")

### 2020-06-26 docs/submission-mee-2020-06-25.tex
more intro

@@
 
-\textbf{Word count}: 2969
+\textbf{Word count}: 3016
 
@@
 
-From molecular data to alignments and phylogenies, public databases such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life data deposition system, the phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OToL) data deposition tool, the phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data.
 
-More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about 98.05\% of taxa included in the NCBI taxonomy (Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}) which represents about 10\% of described diversity (Scott \protect\hyperlink{ref-scott2011ncbi}{2011}), taking a conservative estimate of extant diversity.
+More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about 98.05\% of taxa included in the NCBI taxonomy (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}) which represents about 10\% of described diversity (Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}), taking a conservative estimate of extant diversity.
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \url{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}.
@@
 Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
-Tools that automatize the assembly of homology hypotheses for phylogenetics such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}) have been widely used to assemble genetic datasets to reconstruct phylogenetic relationships among different organisms.
+Tools that automatize the assembly of homology hypotheses from the GenBAnk database for phylogenetics such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}) have been widely used to reconstruct the evolutionary relationships among different organisms.
 
-Considering that 400 phylogenies are published each year, most published phylogenies have been infered from a homology hypothesis that has been constructed ``by hand''.
+This has been an important contribution to the field of phylogenetics, especially bc they represent a possible way for reproducibility in phylogenetics.
+Indeed, most published phylogenies have been infered from a homology hypothesis that has been curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
 
@@
 A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of literature shows them that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both in the OToL phylesystem and is part of the synthetic tree, and on treeBASE. It samples 48 species.
+The genus encompasses between 400-600 living species. A review of literature reveals that there are three phylogenetic trees showing relationships within the hollies published.
+The first one has been made available in TreeBASE as well as in the OToL phylesystem and is part of the synthetic tree. It samples 48 species.
 The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
 The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
-We would expect that updating the oldest tree, we would get something very similar to the newest tree.
+This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
 
@@
 However, the head of the group wants to use a particular marker they beilieve to be the best one to be able to solve the relationships in the group.
-They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating.
+They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating and that they trust.
 
 \begin{figure}
-\includegraphics[width=0.75\linewidth]{docs/figs/cotree-plot2-1} \caption{Comparison of original tree and update tree of the Malvaceae.}\label{fig:fig-2}
+
+{\centering \includegraphics[width=0.85\linewidth]{docs/figs/cotree-plot2-1}
+
+}
+
+\caption{Comparison of original tree and update tree of the Malvaceae.}\label{fig:fig-2}
 \end{figure}
+\newpage
 
@@
 
+\leavevmode\hypertarget{ref-morrison2009would}{}%
+Morrison, D.A. (2009). Why would phylogeneticists ignore computerized sequence alignment? \emph{Systematic biology}, \textbf{58}, 150--158.
+
 \leavevmode\hypertarget{ref-piel2009treebase}{}%

### 2020-06-26 2-body.Rmd
more intro

@@
 
-More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of most of the taxa represented [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota], covering about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], covering about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences <https://ftp.ncbi.nih.gov/genbank/gbrel.txt>.
 
-Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public databases such as GenBank, as well as private databases.
+Many useful tools have been developed in the past decades in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
 Generally refered to as "pipelines", most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
@@
 Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (refered to as "phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely used to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 fish species of the family of barracudas [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 living fern species [@lehtonen2011towards].
+maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [@lehtonen2011towards].
 
-Pipelines have been an important incorporation to the field of phylogenetics, particularly because they represent a clear step towards reproducibility in the field.
+Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
 In contrast, most published phylogenies have been infered using homology hypotheses that have been curated "by hand" [@morrison2009would].
@@
 
-An way to incorporate the best of two worlds (newly released molecular data with fine knowledge from human experts) would be to rely on published manually curated homology hypotheses that have been made publicly available.
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data VS fine curation from human experts) would be to rely on published manually curated homology hypotheses.
 This expert-curated alignments can be enriched and updated by incorporating newly released data from public molecular databases.
 
-There are THIS MANY curated alignments publicly available on TreeBASE
+As of April 2014, TreeBASE hosted a bit more than 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa [(see TreeBASE about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
+This provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in different regions of the tree of life.
 
-This suggests a lot of potential for update of phylogenetic relationships in different regions of the tree of life.
+The OToL tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates the phylogenies from TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+This provides a loose means of linking the tree with the exact alignment that generated it.
+This can be a difficult task when multiple alignments were originally deposited and no clear link between alignment and tree was provided in TreeBASE.
 
-OToL has integrated the phylogenies fro TreeBASe, and linked to the alignment repo but not the exact alignment that generated the phylogeny.
+Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
+Moreover, each one of these data repositories follow their own system for taxon identification.
+A real challenge is to find an effective way to link data from across databases that belongs to the same taxon.
+OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.
 
-Linkage of original alignment with corresponding phylogeny has t be done by a human curator.
+Physcraper is a python pipeline that uses the OToL metadata system to connect databases through taxon identification numbers.
 
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
+It's main goal is to connect public phylogenetic relationships in OToL, with alignments from TreeBASE and GenBank DNA data.
 
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered "best" by experts
-not only in phylogenetics, but also experts on the focal group of study.
+It also allows automatizing and standardizing the comparison of phylogenetic hypotheses.
 
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
+This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
+And an effort to more effectively connect the world of big data in biology.
 
@@
 This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
+  A user can choose from the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
 If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
@@
 - Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
-- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to CITE REDELINGS AND HOLDER [@redelings and holder]
+- Describe what a conflict analysis is: Node by node comparison between the original and updated tree, and the synthetic OToL using the conflict API of otol CITE REDELINGS AND HOLDER [@redelings and holder].
+
 - For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.

### 2020-06-26 2.1-intro-ideas.Rmd
more intro

@@
 Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
+
+The pipelines are so powerful and they will give you an answer, but there
+is no way to assess if it is better than previous answers, it just assumes it is
+better because it used more data.
+
+A less automated workflow is one in which the alignments that generated the published
+phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
+are provided by the users.
+
+A key aspect of the standard phylogenetic workflow is comparison with already existing
+phylogenetic hypotheses and with phylogenies that are considered "best" by experts
+not only in phylogenetics, but also experts on the focal group of study.
+
+Physcraper uses the Open Tree of Life phylesystem and connects it
+to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
+a phylogeny that was published and then made available in the OToL database, (2)
+use this DNA alignments as a starting point to get new genetic data belonging
+to the focal group of study, to (3) finally update the phylogenetic relationships in the group.

### 2020-06-26 _TODO.Rmd
more intro

@@
 
+- [ ] intro: challenge in connecting different parts of the database with other databases that might refer to the same thing.
+For example taxon ids represent onw challenge in this sense, but it has been solved by the otol taxonomy project.
+Othe rchallenge is the study ids.
+This is not fully resolved yet by otol, but only partially by linking to the we solve it in
 - [ ] Do not cite Felsenstein 1985 for bootstrap, just the call that we use for RAxML:

### 2020-06-26 docs/submission-mee-2020-06-25.tex
more intro

@@
 
-\textbf{Word count}: 3016
+\textbf{Word count}: 3073
 
@@
 
-From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OToL) data deposition tool, the phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data.
 
-More than a decade ago, the GenBank database release number 159 (April 15, 2007) had 72 million DNA sequences, and was estimated to have the potential to resolve phylogenetic relationships of about 98.05\% of taxa included in the NCBI taxonomy (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}) which represents about 10\% of described diversity (Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}), taking a conservative estimate of extant diversity.
+More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), covering about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \url{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}.
 
-Many useful tools have been developed in the past decades in an effort to make sense of the large amount of DNA data in public and private databases.
-Most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses to reconstruct phylogenetic relationships.
-Tools that automatize the assembly of homology hypotheses from the GenBAnk database for phylogenetics such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}) have been widely used to reconstruct the evolutionary relationships among different organisms.
+Many useful tools have been developed in the past decades in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
+Generally refered to as ``pipelines'', most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses (aka molecular alignments) to reconstruct phylogenetic relationships.
+Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (refered to as ``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}), have been widely used to study the evolutionary relationships among different organisms (TABLE??
+maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish (Sphyraenidae; Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}), to a mega-phylogeny of almost 3 000 species of living ferns (Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011}).
 
-This has been an important contribution to the field of phylogenetics, especially bc they represent a possible way for reproducibility in phylogenetics.
-Indeed, most published phylogenies have been infered from a homology hypothesis that has been curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-
-On one hand, there are concerns about the use of automated tools.
+Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies have been infered using homology hypotheses that have been curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
+What explains this preference?
+aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
+On one hand, there are concerns about the use of automated pipelines in phylogenetics.
 Concerns I think people have about these tools:
 - Errors in identification of sequences
-- Little control along the process
+- Little control at different steps of the process
 - Too much of a black box?
@@
 -
-It has also been suggested that manual curation of genetic datasets always produces better results in phylogenetic analyses.
+It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-Fortunately, public manually curated homology hypotheses have the potential to be enriched and updated by incorporating newly released data from public molecular databases as an alternative to automatic identification and clustering of DNA homologs for assembly of homology hypotheses to reconstruct phylogenetic relationships.
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data VS fine curation from human experts) would be to rely on published manually curated homology hypotheses.
+This expert-curated alignments can be enriched and updated by incorporating newly released data from public molecular databases.
 
-There are THIS MANY curated alignments publicly available on TreeBASE
+As of April 2014, TreeBASE hosted a bit more than 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE about)}.
+This provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in different regions of the tree of life.
 
-This suggests a lot of potential for update of phylogenetic relationships in different regions of the tree of life.
+The OToL tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates the phylogenies from TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+This provides a loose means of linking the tree with the exact alignment that generated it.
+This can be a difficult task when multiple alignments were originally deposited and no clear link between alignment and tree was provided in TreeBASE.
 
-OToL has integrated the phylogenies fro TreeBASe, and linked to the alignment repo but not the exact alignment that generated the phylogeny.
+Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
+Moreover, each one of these data repositories follow their own system for taxon identification.
+A real challenge is to find an effective way to link data from across databases that belongs to the same taxon.
+OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.
 
-Linkage of original alignment with corresponding phylogeny has t be done by a human curator.
+Physcraper is a python pipeline that uses the OToL metadata system to connect databases through taxon identification numbers.
 
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
+It's main goal is to connect public phylogenetic relationships in OToL, with alignments from TreeBASE and GenBank DNA data.
 
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered ``best'' by experts
-not only in phylogenetics, but also experts on the focal group of study.
+It also allows automatizing and standardizing the comparison of phylogenetic hypotheses.
 
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000})).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
+This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
+And an effort to more effectively connect the world of big data in biology.
 
@@
 This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the `r rotl::tol\_about()\$num\_source\_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}).
+  A user can choose from the 1216 published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}).
 If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
@@
 \begin{itemize}
-\tightlist
 \item
@@
 \item
-  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to CITE REDELINGS AND HOLDER ({\textbf{???}} and holder)
+  Describe what a conflict analysis is: Node by node comparison between the original and updated tree, and the synthetic OToL using the conflict API of otol CITE REDELINGS AND HOLDER ({\textbf{???}} and holder).
 \item
@@
 
+\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
+Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
+\& Mendoza, C.G.
+(2017).
+A pilot study applying the plant anchored hybrid enrichment method to new world sages (salvia subgenus calosphace; lamiaceae).
+\emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
+
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
 Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
+
+\leavevmode\hypertarget{ref-lehtonen2011towards}{}%
+Lehtonen, S. (2011). Towards resolving the complete fern tree of life. \emph{PLoS One}, \textbf{6}.
 
@@
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
+
+\leavevmode\hypertarget{ref-santini2013first}{}%
+Santini, F.
+\& Sorenson, L.
+(2013).
+First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies.
+\emph{Italian journal of zoology}, \textbf{80}, 481--489.
 

### 2020-06-26 2-body.Rmd
more intro

@@
 
-From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data and there is no straightforward way to automatically connect data belonging to the same taxon from these various resources.
 
@@
 
-Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility.
 In contrast, most published phylogenies have been infered using homology hypotheses that have been curated "by hand" [@morrison2009would].
-What explains this preference?
-aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
+There seems to be a preference of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers, potentially poorly aligned, but the amount of data will overcome errors in the alignment).
+
 On one hand, there are concerns about the use of automated pipelines in phylogenetics.
@@
 Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
--
-It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions [@fragoso2017pilot].
+
+It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and it has been demonstrated for genomic alignments [@fragoso2017pilot].
 
@@
 Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, each one of these data repositories follow their own system for taxon identification.
-A real challenge is to find an effective way to link data from across databases that belongs to the same taxon.
+Moreover, each one of these data repositories follow their own system for taxon identification, posing a real challenge to automatically link data from across databases that belong to the same taxon.
 OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.

### 2020-06-26 _TODO.Rmd
more intro

--- 2020-06-24 _TODO.Rmd
@@
 # next:
+
+- [ ] intro: challenge in connecting different parts of the database with other databases that might refer to the same thing.
+For example taxon ids represent onw challenge in this sense, but it has been solved by the otol taxonomy project.
+Othe rchallenge is the study ids.
+This is not fully resolved yet by otol, but only partially by linking to the alignment repo.
+- [ ] intro: programatic access.
 

### 2020-06-26 render2-sysbio.Rmd
more intro

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    includes:
      # in_header: preamble.tex
      # after_body: [biorxiv_2019.09.28figcaps.tex]
    # fig_caption: no
geometry: margin = 1in
bibliography: paper.bib
csl: submission_sysbio/systematic-biology.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    - \usepackage{caption}
    - \usepackage{blindtext}
    - \usepackage[left]{lineno}
    - \linenumbers
  # - \usepackage{xr}
    # - \usepackage[utf8]{inputenc}
    # - \usepackage[english]{babel}
    # - \usepackage[document]{ragged2e}
# nocite: |
#   @barker2012going, @barker2015new, @burns2014phylogenetics, # @claramunt2015new, @gibb2015new, @Hedges2015, # @hooper2017chromosomal, @Jetz2012, @price2014niche
---

<!-- ``` {r child = '0-header.Rmd'} -->
<!-- ``` -->
\newpage
``` {r child = '1-abstract.Rmd'}
```
\newpage

``` {r child = '2-body.Rmd'}
```
\newpage
\begin{center}
\textsc{References}
\end{center}

### 2020-06-26 2-body.Rmd
more intro

--- 2020-06-26 2-body.Rmd
@@
 
-From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data, and there is currently no straightforward way to automatically connect their data.
 
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], covering about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
-In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences <https://ftp.ncbi.nih.gov/genbank/gbrel.txt>.
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns coveres about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
+Yet, publicly avilable phylogenetic relationships still cover about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data avilable a decade ago.
+While it is true that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn], most recently published large trees have been made available in recent years, indicating a lag between the amount of new DNA data generated and its analysis in a phylogenetic context.
 
-Many useful tools have been developed in the past decades in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally refered to as "pipelines", most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses (aka molecular alignments) to reconstruct phylogenetic relationships.
-Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (refered to as "phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely used to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [@lehtonen2011towards].
+Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
+Generally refered to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
+Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely applied to study the evolutionary relationships among different organisms (TABLE??
+maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [Moniloformopses; @lehtonen2011towards].
 
 Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies have been infered using homology hypotheses that have been curated "by hand" [@morrison2009would].
-What explains this preference?
-aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
-On one hand, there are concerns about the use of automated pipelines in phylogenetics.
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control at different steps of the process
-- Too much of a black box?
-Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
--
-It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions [@fragoso2017pilot].
+In contrast, most published phylogenies to date have been infered using alignments that have been assembled and curated "by hand" [@morrison2009would].
+There seems to be a preference in the field of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data VS fine curation from human experts) would be to rely on published manually curated homology hypotheses.
-This expert-curated alignments can be enriched and updated by incorporating newly released data from public molecular databases.
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine curation from human experts) would be to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
+This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
 
-As of April 2014, TreeBASE hosted a bit more than 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa [(see TreeBASE about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
-This provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in different regions of the tree of life.
+As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
+This database provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
 
-The OToL tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates the phylogenies from TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+The OToL tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
 This provides a loose means of linking the tree with the exact alignment that generated it.
-This can be a difficult task when multiple alignments were originally deposited and no clear link between alignment and tree was provided in TreeBASE.
 
 Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, each one of these data repositories follow their own system for taxon identification.
-A real challenge is to find an effective way to link data from across databases that belongs to the same taxon.
+Moreover, each one of these data repositories follow their own system for taxon identification, posing a real challenge to automatically link data from across databases that belong to the same taxon.
 OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.
@@
 Physcraper is a python pipeline that uses the OToL metadata system to connect databases through taxon identification numbers.
-
-It's main goal is to connect public phylogenetic relationships in OToL, with alignments from TreeBASE and GenBank DNA data.
-
-It also allows automatizing and standardizing the comparison of phylogenetic hypotheses.
+It's main functionality is to connect phylogenies stored in the OToL phylesystem, with alignments from TreeBASE and newly realeased DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Because of its design, it allows taking advantage of the many resources provided by the OToL.
+For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships.
 
 This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
-And an effort to more effectively connect the world of big data in biology.
+And an effort to more effectively make big data connections for the advantage of phylogenetics and biology in general.
 
@@
 
-## The input: a study tree and an alignment
+## The input: a tree and an alignment
 
-- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
+- In order to take advantage of the OToL tools, the tree needs to be stored in the
+  OToL phylesystem [@mctavish2015phylesystem]. Currently, only trees connected to
+  a published study can be stored in the phylesystem (although there are plans to
+  allow storage of non published trees). The main
   reason for this is that trees in phylesystem have a set of user defined characteristics
   that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
+The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
 Outgroup and ingroup taxa in the original tree are identified and tagged.
 This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
+  A user can choose from the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OToL website ([See OToL's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+If the published tree you are interested in updating is not in there, you can upload it via OToL's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
 - The alignment should be a gene alignment that was used to generate the tree. The original
   alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
+  DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal were the tree was originally published.
   If the alignment is stored in TreeBase, `physcraper` can download it directly,
-  either from the TreeBASE website (<https://treebase.org/>)
-  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
+  either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
+  or through the TreeBASE GitHub repository (SuperTreeBASE; [github.com/TreeBASE/supertreebase](https://github.com/TreeBASE/supertreebase)).
+  If the alignment is on another repository, or provided personally by the owner, a path to a local copy of the alignment has to be provided.
 - A taxon name matching step is performed to verify that all taxon names on the tips
@@
 - Unmatched taxon names are dropped from both the tree and alignment.
-  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
   Technically, just one matching name is needed to perform the searches. Please, see next section.
@@
   the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
+- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide

### 2020-06-26 2.1-intro-ideas.Rmd
more intro

@@
 to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
+
+On one hand, there are concerns about the use of automated pipelines in phylogenetics.
+Concerns I think people have about these tools:
+- Errors in identification of sequences
+- Little control at different steps of the process
+- Too much of a black box?
+Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.

### 2020-06-26 docs/submission-mee-2020-06-25.tex
more intro

@@
 
-\textbf{Word count}: 3073
+\textbf{Word count}: 3169
 
@@
 
-From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data and there is currently no straightforward way to automatically connect data belonging to the same taxon from these various resources.
 
@@
 With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), covering about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
-In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \url{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}.
+In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(See GenBank release data)}.
+Yet, publicly avilable phylogenetic relationships cover about 90 000 taxa {[}Hinchliff \emph{et al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}); current synthesis citation{]}, covering less than one third of the taxonomic diversity with data avilable a decade ago.
+While it is true that many phylogenies are not publicly shared (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}), most recently published large trees have been made available, indicating a lag between the amount of new DNA data generated and its analysis in a phylogenetic context.
 
@@
 
-Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility.
 In contrast, most published phylogenies have been infered using homology hypotheses that have been curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-What explains this preference?
-aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
+There seems to be a preference of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers, potentially poorly aligned, but the amount of data will overcome errors in the alignment).
+
 On one hand, there are concerns about the use of automated pipelines in phylogenetics.
@@
 Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
--
-It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+
+It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and it has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
@@
 
+\leavevmode\hypertarget{ref-drew2013lost}{}%
+Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez, R., Katz, L.A., Crandall, K.A., Hibbett, D.S.
+\& Soltis, D.E.
+(2013).
+Lost branches on the tree of life.
+\emph{PLoS biology}, \textbf{11}.
+
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
@@
 
+\leavevmode\hypertarget{ref-hinchliff2015synthesis}{}%
+Hinchliff, C.E., Smith, S.A., Allman, J.F., Burleigh, J.G., Chaudhary, R., Coghill, L.M., Crandall, K.A., Deng, J., Drew, B.T., Gazis, R.
+\& others.
+(2015).
+Synthesis of phylogeny and taxonomy into a comprehensive tree of life.
+\emph{Proceedings of the National Academy of Sciences}, \textbf{112}, 12764--12769.
+
 \leavevmode\hypertarget{ref-lehtonen2011towards}{}%
@@
 \emph{Nature}, \textbf{556}, 452--456.
+
+\leavevmode\hypertarget{ref-magee2014dawn}{}%
+Magee, A.F., May, M.R. \& Moore, B.R. (2014). The dawn of open access to phylogenetic data. \emph{PLoS One}, \textbf{9}.
 

### 2020-06-26 2-body.Rmd
more intro

@@
 As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
-This database provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
+This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
 
@@
 Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, each one of these data repositories follow their own system for taxon identification, posing a real challenge to automatically link data from across databases that belong to the same taxon.
-OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.
+Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
+OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
 
-Physcraper is a python pipeline that uses the OToL metadata system to connect databases through taxon identification numbers.
-It's main functionality is to connect phylogenies stored in the OToL phylesystem, with alignments from TreeBASE and newly realeased DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Physcraper is a python pipeline that relies on the OToL metadata system to connect databases through taxon identification numbers.
+It's main functionality is to connect phylogenies stored in the OToL phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
 Because of its design, it allows taking advantage of the many resources provided by the OToL.
@@
 
-- In order to take advantage of the OToL tools, the tree needs to be stored in the
+- In order to take advantage of the OToL tools, the input tree needs to be stored in the
   OToL phylesystem [@mctavish2015phylesystem]. Currently, only trees connected to
@@
   or through the TreeBASE GitHub repository (SuperTreeBASE; [github.com/TreeBASE/supertreebase](https://github.com/TreeBASE/supertreebase)).
-  If the alignment is on another repository, or provided personally by the owner, a path to a local copy of the alignment has to be provided.
+  If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
 - A taxon name matching step is performed to verify that all taxon names on the tips

### 2020-06-26 2-body.Rmd
methods

@@
 - In order to take advantage of the OToL tools, the input tree needs to be stored in the
-  OToL phylesystem [@mctavish2015phylesystem]. Currently, only trees connected to
+  OToL phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem).
+  Currently, only trees connected to
   a published study can be stored in the phylesystem (although there are plans to
@@
 Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OToL website ([See OToL's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+This allows to automatically set the root for the updated tree on the final steps of the pipeline.
+  A user can choose from among the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OToL website ([See OToL's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
 If the published tree you are interested in updating is not in there, you can upload it via OToL's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
@@
 
+- Technically, any DNA molecular database can be used to search for new sequences.
+  However, we rely on the GenBank database.
+  The new sequence search can be performed on the remote database or in a local database.
 - The next step is to identify the search taxon within the reference taxonomy.
@@
 Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be provided by the user.
+The search taxon can be determined by the user.
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
@@
   To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
+I FORGOT TO ASK YOU ABOUT THIS CITATION!
 
@@
   similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the accepted sequences on the alignment.
-  The blastn function from the BLAST command line tools [@camacho2009blast] is used for local-database searches.
-  A modified biopython blast function is used for web-based searches.
+  database, and the sequences on the checked alignment.
+  The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local-database sequence searches.
+  For remote-database searches, we modified the BioPython [@cock2009biopython] BLAST function in the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternate BLAST URL.
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the blastn function is used to performs the similarity search [@camacho2009blast].
-- The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST function was modified to accepts is used to perform the similarity search.
-- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
+- A pairwise BLAST search is performed. This means that each sequence
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
@@
 This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. ***REPORT THE DEFAULT VALUES AND DESCRIBE WHAT THEY MEAN*** This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack database id numbers, a filtering
+- Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
+  This filtering guarantees that genomic sequences are not included.
+  All acepted sequences are assigned an internal identifier, and are further filtered.
+- Because the original alignments usually lack taxonomic identification numbers, an extra filtering
   step is needed. Accepted sequences that belong to the
@@
 - Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
+  Although it can be useful to keep some of them, for example, to investigate monophyly
   within species, there can be hundreds of exemplar sequences per taxon for some markers.
@@
 - Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
+This can be done iteratively, but by default only sequences in the checked alignment are blasted.
 ***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
+- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (default to physcraper/taxonomy), so they are accesible for further runs.
 - A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
@@
 
-- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
+- The software MUSCLE [@edgar2004muscle] is implemented to perform sequence alignments.
 - First, all new sequences are aligned using default MUSCLE options.
@@
   follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. We encourage users
+- The final alignment is not further processed automatically. So, we encourage users
   to check it either by eye and perform manual refinement or using any of the many
@@
 
-- A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML [@stamatakis2014raxml]
-  with 100 classic rapid bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
-  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+- A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML [@stamatakis2014raxml], using default settings such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with default method.
+Currently only the number of bootsrap replicates can be modified by the user.
 - The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
-- Bootstrap results are summarized with Dendropy ADD CITATION
+- Bootstrap results are summarized with Dendropy [@sukumaran2010dendropy]
 - The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
 - Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
-- Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
-- Describe what a conflict analysis is: Node by node comparison between the original and updated tree, and the synthetic OToL using the conflict API of otol CITE REDELINGS AND HOLDER [@redelings and holder].
-
+- Robinson Foulds weighted and unweighted metrics are calculated with Dendropy functions.
+- Finally a conflict analysis is performed.
+This is basically a node by node comparison between the the synthetic OToL and the original and updated tree individually.
+This is performed with OToL's conflict Application Programming Interface [@redelings2017supertree].
 - For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
-- A SUGGESTED DEFAULT ROOTING BASED ON THE OPEN TREE TAXONOMY is implemented for now.
-DESCRIBE HOW IT WORKS.
-SAY THAT IT IS A PROBLEM.
-Automatic rooting is not that smart yet.
-The best way right now is for users to define outgroups so trees are better rooted.
-- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- Conflict information can only be generated in the context of the whole Open Tree of
-  Life. Otherwise, it is not really possible to get conflict data.
-***- One way to compare two independent phylogenetic trees is to compare them both to
-the synthetic OToL and then measure how well they do against each other***
+- A suggested default rooting based on OToL's taxonomy is implemented for now.
+It takes one descendant from the corresponding induced subtree in OToL's synthetic tree and then finds the MRCA of those in th eupdated trees.
+However, if the updated tree changes expectations from taxonomy, the root will no longer be accurate.
+Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
+The best way right now is for users to define outgroups so trees are accurately rooted.
 
@@
 
-To exemplify the utility of physcraper we will address two use-case scenarios.
+We will illustrate the utility of physcraper in here with two use-case scenarios.
 One in which the user is interested in a particular group.
 Another one in which the user is interested in a particular phylogeny.
-A tutrial as well as illustrated examples of steps needed to perform a physcraper analyses are available elsewhere.
+A tutorial as well as illustrated examples of steps needed to perform a physcraper analysis are available elsewhere.
 
@@
 The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-<!-- ***Contact authors? They seem old school, probably do not wanna share their data.*** -->
 The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.

### 2020-06-26 2.1-intro-ideas.Rmd
methods

@@
 Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
+
+Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+
+***- One way to compare two independent phylogenetic trees is to compare them both to
+the synthetic OToL and then measure how well they do against each other***
+
+- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
+This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
+- Conflict information can only be generated in the context of the whole Open Tree of Life. Otherwise, it is not really possible to get conflict data.

### 2020-06-26 2-body.Rmd
discussion

@@
 This can be done iteratively, but by default only sequences in the checked alignment are blasted.
-***Is there an argument to control the number of cycles of blast searches with new sequences?***
+I ALSO FORGOT TO ASK: ***Is there an argument to control the number of cycles of blast searches with new sequences?***
 - Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (default to physcraper/taxonomy), so they are accesible for further runs.
@@
 Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of steps needed to perform a physcraper analysis are available elsewhere.
+A tutorial as well as illustrated examples of commands for every step needed to perform a physcraper analysis are available elsewhere.
 
@@
 A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three phylogenetic trees showing relationships within the hollies published.
+The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
 The first one has been made available in TreeBASE as well as in the OToL phylesystem and is part of the synthetic tree. It samples 48 species.
@@
 The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+This also makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
 Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
+
+DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OTOL
+RF DISTANCE INTERPRETATION
+HOW MUCH TIME THE BLAST RUN TOOK
+ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE
+
+FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OTOL.
 
@@
 
-A postdoc started working with a new reserach group. They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species.
+A postdoc started working with a new reserach group.
+They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species, containing the relatives of cacao, cotton, durian and okra.
 A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
-However, the head of the group wants to use a particular marker they beilieve to be the best one to be able to solve the relationships in the group.
-They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology they have been curating and that they trust.
+However, the head of the research group wants to use a particular marker they believe to be the best one to be able to solve the relationships in the group.
+They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
 
@@
 Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
+Besides the actual data, they are rich sources of information that can be used for the advantage of biology and science in general.
 
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
-exposing new comers to expert understanding about the data.
-
-There are many tools that are making use of DNA data repositories in different ways.
+Many pipelines are making use of DNA data repositories in different ways.
 Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
+Some focus on accurate ways of mining the data -- getting real and clean homologs.
+Others focus on refining the alignment.
 Most focus on generating full trees *de novo*, mainly for regions of the Tree of
@@
 regions that have been already studied and that have phylogenetic data already.
+However, expert phylogenetic knowledge is also an important source of data in public
+and open repositories that is potentiallly being diregarded.
 
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
+All these tools are key efforts for advancing towards reproducibility in phylogenetics,
 a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
+In here, we highlight the potential of other sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+community to build up our phylogenetic knowledge, using not only data accumulated in
+molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+While not generating full phylogenies *de novo*, pyscraper is still capable of generating new phylogenetic knowledge.
+In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??).
+It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
 
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.
-
-Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer [@ranwez2009phyloexplorer].
-Compare and discuss.
-
-How is physcraper already useful:
-- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
-- Finding
-
-How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
-- obtain trees for ecophylogenetic studies, as mentioned in @helmus2012phylogenetic
-- one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
-
-Things that physcraper does not do:
-- analyse the whole GenBank database [@benson2000genbank; @wheeler2000database] to find homolog regions suitable to reconstruct phylogenies, as mentioned in @antonelli2017toward.
-There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by @ranwez2009phyloexplorer. Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
+Physcraper has the potential to be applied for the advantage of the field to rapidly
+place newly discovered species phylogenetically [@webb2010biodiversity],
+obtain trees for ecophylogenetic studies [@helmus2012phylogenetic],
+help to sistematize molecular databases, i.e., curate taxonomic assignations [@san2010molecular],
+and rapidly generate custom species trees for downstream analyses [@stoltzfus2013phylotastic].
 
@@
 
-We acknowledge contributions from
+The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
 
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
+We acknowledge contributions from ...
 

### 2020-06-26 2.4-discussion-ideas.Rmd
discussion

--- 2020-06-24 physcraper_ms.Rmd
@@
----
-title: 'Physcraper: A Python package for continually updated gene trees'
-author:
-  - Luna L. Sanchez Reyes
-  - Martha Kandziora
-  - Emily Jane McTavish
-  - University of California, Merced
-date: '`r format(Sys.time(), "%d %B, %Y")`'
-bibliography: paper.bib
-csl: apa.csl
-output:
-  redoc::redoc
----
+Usually, initial ideas about the data are changed by analyses.
+We expect that this new ideas on the data can be registered on databases,
+exposing newcomers to expert understanding about the data.
 
-# Abstract
-
-1. Phylogenies are a key part of research in all areas of biology. Tools that automatize
-some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
-have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
-However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
-to one that is not a specialist either on phylogenetic methods or on a particular group of study.
-
-1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
-phylogenies by making use of public DNA sequence data and taxonomic information,
-providing a framework for comparison of published phylogenies with their updated versions.
-
-1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on already available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
-of alternative phylogenetic hypotheses (topologies).
-***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know much about a certain biological group.***
-
-1. Physcraper implements node by node/topology comparison of the the original and the updated
-trees using the conflict API of OToL, and summarizes differences.
-
-1. We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-
-1. Physcraper can be used with any OS. Detailed instructions for installation and
-use are available at <https://github.com/McTavishLab/physcraper>.
-
-# Introduction
-
-Phylogenies are important.
-
-Generating phylogenies is not easy and it is largely artisanal.
-Although many efforts to automatize the process have been done, and the community is using those more and more, automatization of phylogenetic reconstruction is still not a widespread practice and among other benefits, it might be key for adoption of better reproducibility practices in the phylogenetics community.
-***paragraph better to end discussion???
-***
-
-The process of phylogenetic reconstruction implies many steps (that I generalize to the following):
-
-1. Obtention of molecular or morphological character data -- get DNA from some organisms
-and sequence it, or get it from an online nucleotide data repository, such as GenBank [@benson2000genbank; @wheeler2000database].
-1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
-aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
-1. Analyse this hypothesis of homology to infer phylogenetic relationships among
-the organisms you are studying -- Use different available programs to infer molecular
-evolution, trees and times of divergence.
-1. Discuss the inferred relationships in the context of previous hypothesis, the
-biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*
-
-Each of these steps require different types of specialized training: in the field,
-in the lab, in front of a computer, discussions with experts in the methods, and/or in the biological group of study.
-All of these steps also require considerable amounts of time for training and implementation.
-
-In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
-pipelines that mine already available molecular data from the GenBank repository [@benson2000genbank; @wheeler2000database],
-to obtain homologous characters that can be used for phylogenetic reconstruction.
-These tools have been presented as aid for the nonspecialist to decrease some of
-the difficulties in the generation of phylogenetic knowledge. However, they are
-not that often used as so, suggesting that there are still difficulties for the nonspecialist.
-The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
-However, automatizing the assembly of the character data set is a crucial step towards
-reproducibility for a task that was otherwise primarily artisanal and hence largely
-non-reproducible.
-
-Even if it is hard to obtain phylogenies, we invest copious amounts of time and energy in generating them.
-Issues such as food security, global warming, global health are crucial to solve and phylogenies might help.
-There is a lot of phylogenetic knowledge already available in published peer-reviewed studies.
-In this sense, the non-specialists (and also the specialist) face a new problem: how do I choose the best phylogeny.
-
-Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank [@benson2000genbank; @wheeler2000database].
-
-We present a way to automatize and standardize the comparison of phylogenetic hypotheses and to allow reproducibility of this last step of the research process.
-
-A key aspect of the standard phylogenetic workflow is comparison with already existing
-phylogenetic hypotheses and with phylogenies that are considered "best" by experts
-not only in phylogenetics, but also experts on the focal group of study.
-
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control along the process
-- Too much of a black box?
-
-Most of these phylogenies are being constructed by people learning about the methods,
-so they want to know what is going on.
-
-The pipelines are so powerful and they will give you an answer, but there
-is no way to assess if it is better than previous answers, it just assumes it is
-better because it used more data.
-
-All these pipelines start tree construction from zero? Yes.
-
-The goal of Physcraper is to build upon previous phylogenetic knowledge,
-allowing a direct comparison between existing phylogenies and phylogenies that are constructed
-using new genetic data retrieved from a public nucleotide database (i.e., GenBank [@benson2000genbank; @wheeler2000database]).
-
-To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
-to the TreeBase database, to (1) get the original DNA data set matrices (alignments) that produced
-a phylogeny that was published and then made available in the OToL database, (2)
-use this DNA alignments as a starting point to get new genetic data belonging
-to the focal group of study, to (3) finally update the phylogenetic relationships in the group.
-
-A less automated workflow is one in which the alignments that generated the published
-phylogeny are stored in other public database (such as DRYAD) or elsewhere (the users computer), and
-are provided by the users.
-
-The original tree is by default used as starting tree for the phylogenetic searches, but it
-can also be set as a full topological constraint or not used at all, depending on
-the goals of the user.
-
-Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.
-
-# How does Physcraper work?
-
-## The input: a study tree and an alignment
-
-- The study tree is a published phylogenetic tree stored in the OToL database, phylesystem [@mctavish2015phylesystem]. The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the 'r rotl::tol_about()$num_source_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (<>).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (<https://tree.opentreeoflife.org/curator).
-- The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD (<http://datadryad.org/>), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, `physcraper` can download it directly,
-  either from the TreeBASE website (<https://treebase.org/>)
-  or through the TreeBASE GitHub repository (SuperTreeBASE; <https://github.com/TreeBASE/supertreebase>).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
-- A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-- A ".csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from both the tree and alignment.
-  <!--The minimum amount of tips necessary for a physcraper run is just one!!!-->
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.
-
-## DNA sequence search and cleaning
-
-- The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be provided by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
-
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; altschul1997gapped] is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the remaining sequences on the alignment.
-  The BLAST command line tools [@camacho2009blast] are used for both web-based and local-database searches.
-  <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise alignment-against-all BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
-- The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST algorithm is used to perform the similarity search.
-- Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack database id numbers, a filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
-- Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- Reverse complement sequences are identified and translated.
-- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
-***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
-- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
-
-## DNA sequence alignment
-
-- The software MUSCLE [@edgar2004muscle] is implemented to perform alignments.
-- First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align new sequences. This ensures that the final alignment
-  follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. We encourage users
-  to check it either by eye and perform manual refinement or using any of the many
-  tools for alignment processing, to eliminate columns with no information.
-
-## Tree reconstruction and comparison
-
-- A gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml]
-  with 100 classic bootstrap [@felsenstein1985confidence] replicates by default. The number of bootsrap replicates can be modified by the user.
-  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
-- The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
-- Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
-- Robinson Foulds metrics
-- Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
-- For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
-- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-- Conflict information can only be generated in the context of the whole Open Tree of
-  Life. Otherwise, it is not really possible to get conflict data.
-***- One way to compare two independent phylogenetic trees is to compare them both to
-the synthetic OToL and then measure how well they do against each other***
-
-# Examples
-
-## The hollies
-
-The genus *Ilex* is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
-The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
-The second has not been made available anywhere, not even in supplementary data of the journal.
-***Contact authors? They seem old school, probably do not wanna share their data.***
-The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
-it has not been added to the syntehtic tree yet.
-This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database [@benson2000genbank; @wheeler2000database].
-Updating the oldest tree, we should get something very similar to the newest tree.
-
-## The Ascomycota
-
-Let's be more specific now about our X group and say it is the Ascomycota.
-The best tree currently available in OToL was published by @schoch2009ascomycota.
-The first step, is to get the Open Tree of Life study id. There are some options to do this:
-- You can go to the Open Tree of Life website and browse until you find it, or
-- you can get the study id using R tools:
-  - By using the TreeBase ID of the study (which is not fully exposed on the
-    TreeBase website home page of the study, so you have to really look it up manually):
-```{r, treebase, eval = TRUE}
-rotl::studies_find_studies(property = "treebaseId", value = "S2137")
-```
-  - By using the name of the focal clade of study (but this behaved very differently):
-```{r, focal, eval = FALSE}
-rotl::studies_find_studies(property="ot:focalCladeOTTTaxonName", value="Ascomycota")
-```
-
-Once we have the study id, we can gather the trees published on that study:
-```{r, treeids, eval = TRUE}
-rotl::get_tree_ids(rotl::get_study_meta("pg_238"))
-rotl::candidate_for_synth(rotl::get_study_meta("pg_238"))
-my_trees <- rotl::get_study("pg_238")
-```
-Both trees from this study have `r ape::Ntip(my_trees[[1]])` tips.
-
-Let's check what one of the trees looks like:
-
-<!-- ```{r fungi-phylo, eval = TRUE, fig.retina = 2, fig.width = 2} -->
-<!-- ape::plot.phylo(ape::ladderize(my_trees[[1]]), type = "phylogram", cex = 0.1) -->
-<!-- ``` -->
-
-1. Download the alignment from TreeBase
-If you are on the TreeBase home page of the study, you can navigate to the matrix tab, and manually download the alignments that were used to reconstruct the trees reported on the study that were also uploaded to TreeBase and to the Open Tree of Life repository.
-To make this task easier, you can use a command to download everything into your working folder:
-```
-physcraper_run.py -s pg_238 -t tree109 -o ../physcraper_example/pg_238
-```
-In this example, all alignments posted on TreeBase were used to reconstruct both trees.
-
-1. With the study id and the alignment files saved locally, we can do a physcraper run with the command:
-
-```
-physcraper_run.py -s pg_238 -t tree109 -a treebase_alns/pg_238tree109.aln -as "nexus" -o pg_238
-```
-
-## Testudines example
-
-Phylogeny of the Testudines 6 tips from @crawford2012more
-There is just one tree in OToL.
-There is just one alignment on [treebase](https://treebase.org/treebase-web/search/study/matrices.html?id=12742) with all the 1 145 loci.
-
-```
-physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
-```
-# Discussion
-
-Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
-
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
-exposing new comers to expert understanding about the data.
-
-There are many tools that are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
-Most focus on generating full trees *de novo*, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have been already studied and that have phylogenetic data already.
-
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
-
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies *de novo*.
+Taking advantage of public DNA data bases have been the main focus.
 
@@
 - to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
--
+- Finding
 
-How can it be used for the advantage of the field:
+Physcraper potential:
 - rapid phylogenetic placing of newly discovered species, as mentioned in @webb2010biodiversity
@@
 - it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
-
-## Tools that automatize any part of the process of phylogenetic reconstruction:
-
-### 1. Mining DNA databases to generate datasets suitable for phylogenetic reconstruction
-
-| Tool | Citation | Cited by | Description | Supermatrix/gene tree/species tree |
-|-----|------|:------:|:------:|:------:|
-| Phylota | @sanderson2008phylota | 122 studies | finds sets of DNA homologs on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| AMPHORA | @wu2008simple | 458 studies | baited search; protein markers on phylogenomic data; personal database of genomes or metagenomic data, manually downloaded either from a public database or from private data; phylogenetic reconstruction | Supermatrix |
-| PHLAWD | @smith2009mega | 234 studies | Baited search of DNA markers on the GenBank database; phylogenetic reconstruction | Supermatrix |
-| Unnamed [ruby pipeline](https://www.zfmk.de/en/research/research-centres-and-groups/taming-of-an-impossible-child-pipeline-tools-and-manuals), only available from [supplementary data](https://static-content.springer.com/esm/art%3A10.1186%2F1741-7007-9-55/MediaObjects/12915_2011_480_MOESM1_ESM.ZIP) of the journal | @peters2011taming | 64 studies | mining public DNA databases, focuses on filtering massive amounts of mined sequences by using established "criteria of compositional homogeneity and defined levels of density and overlap" | Supermatrix |
-| Unnamed | @grant2014building | 38 studies | predecessor of phylotol; homolog clustering; public and/or personal DNA database; phylogenetic reconstruction; broad taxon analyses; remove contaminant sequences, based on similarity and on phylogenetic position | supermatrix |
-| Unnamed | @chesters2014protocol | 10 studies | algorithm that mines GenBank data to delineate species in the insecta.
-The authors present a nice comparison with the phylota algorithm | Species trees??
-|
-| PUmPER | @izquierdo2014pumper | 14 studies | perpetual updating with newly added sequences to GenBank | not sure yet |
-| DarwinTree | @meng2015darwintree | 6 studies | predecessor is Phylogenetic Analysis of Land Plants Platform (PALPP), takes data from GenBank, EMBL and DDBJ for land plants only| not sure |
-| NCBIminer | @xu2015ncbiminer | 4 studies | part of darwintree | not sure |
-|SUMAC | @freyman2015sumac | 19 studies | both “baited” analyses and single‐linkage clustering methods, as well as a novel means of determining when there are enough overlapping data in the DNA matrix | not sure |
-|STBase | @mcmahon2015stbase | 7 studies | pipeline for species tree construction and the public database of one million precomputed species trees | species trees |
-| Unnamed | @papadopoulou2015automated | 17 studies | Automated DNA-based plant identification for large-scale biodiversity assessment | not sure |
-| BIR | @kumar2015bir | 6 studies | blast, align, identify homologs via constructed trees, curate and realign | supermatrix |
-| SUPERSMART | @antonelli2017toward | 35 studies | baited analyses up to bayesian divergence time estimation | supermatrix |
-| SOPHI | [@chesters2017construction | 17 studies | Searches DNA sequence data from repos other than GenBank, such as transcriptomic and barcoding repos | not sure |
-| phyloSkeleton | @guy2017phyloskeleton | 5 studies | focuses on taxon sampling; baited genomic sequences; public database (NCBI and JGI); marker identification| supermatrix |
-| OneTwoTree | @drori2018onetwotree | 7 studies | Web‐based, user-friendly, online tool for species-tree reconstruction, based on the *supermatrix paradigm* and retrieves all available sequence data from NCBI GenBank| supermatrix |
-| pyPhlawd | @smith2019pyphlawd | 6 studies | baited and clustering analyses | Supermatrix or gene tree |
-| Phylotol | @ceron2019phylotol | 5 studies | "phylogenomic pipeline to allow easy incorporation of data from  high-throughput sequencing studies, to automate production of both multiple sequence alignments and gene trees, and to identify and remove contaminants.
-PhyloToL is designed for phylogenomic analyses of diverse lineages across the tree of life", i.e., bacteria and unicellular eukaryotes | supermatrix and gene trees |
-
-According to @ceron2019phylotol, PhyLoTA and BIR "focus on the identification and collection
-of homologous and paralog genes from public databases such as GenBank", while both AMPHORA and PHLAWD
-"focus on the construction and refinement of robust alignments rather than the collection of homologs."
-
-### 2. Searching phylogenetic tree databases
-
-PhyloFinder [@chen2008phylofinder] - cited by 18: a search engine for phylogenetic databases, using
-trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies
-
-### 3. Mining phylogenetic tree databases
-
-PhyloExplorer [@ranwez2009phyloexplorer] - cited by 21: a python and MySQL based website to facilitate
-assessment and management of phylogenetic tree collections. It provides "statistics describing the collection,
-correcting invalid taxon names, extracting taxonomically relevant parts of the collection
-using a dedicated query language, and identifying related trees in the TreeBASE database".
-
-### 4. Pipeline for phylogenetic reconstruction
-
-PhySpeTre [@fang2019physpetree] - no citations yet - no sequence retrieval, just phylogenetic reconstruction
-pipeline.
-
-### 5. getting metadata and not sequences from GenBank.
-
-Datataxa @ruiz2019datataxa - no citations yet - focus on extracting metadata from GenBank sequence information.
-
-## Phylota overview
-
-```{r phylota, child="phylota-overview.Rmd"}
-```
-
-# Acknowledgements
-
-We acknowledge contributions from
-
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
-
-# References

### 2020-06-26 docs/submission-mee-2020-06-26.tex
discussion

--- 2020-06-26 docs/submission-mee-2020-06-25.tex
@@
 
-\textbf{Word count}: 3073
+\textbf{Word count}: 3170
 
@@
 
-From molecular data to alignments and phylogenies, public biological data repositories such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data.
+From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data, and there is currently no straightforward way to automatically connect their data.
 
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), covering about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
-In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \url{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}.
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), which in turns coveres about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
+In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(See GenBank's release website)}.
+Yet, publicly avilable phylogenetic relationships still cover about 90 000 taxa {[}Hinchliff \emph{et al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}); CURRENT SYNTH TREE CITATION{]}, covering less than one third of the taxonomic diversity with data avilable a decade ago.
+While it is true that many phylogenies are not publicly shared (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}), most recently published large trees have been made available in recent years, indicating a lag between the amount of new DNA data generated and its analysis in a phylogenetic context.
 
-Many useful tools have been developed in the past decades in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally refered to as ``pipelines'', most of these tools were motivated by the genomics revolution, to identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses (aka molecular alignments) to reconstruct phylogenetic relationships.
-Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (refered to as ``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}), have been widely used to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish (Sphyraenidae; Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}), to a mega-phylogeny of almost 3 000 species of living ferns (Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011}).
+Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
+Generally refered to as ``pipelines'', most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
+Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}), have been widely applied to study the evolutionary relationships among different organisms (TABLE??
+maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish (Sphyraenidae; Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}), to a mega-phylogeny of almost 3 000 species of living ferns (Moniloformopses; Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011}).
 
 Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies have been infered using homology hypotheses that have been curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-What explains this preference?
-aka the discordance between the genomics approach (a gazillion markers poorly aligned, but the amount of data will overcome errors in alignment) and the classic phylogenetics approach (few markers thoughtfully curated).
-On one hand, there are concerns about the use of automated pipelines in phylogenetics.
-Concerns I think people have about these tools:
-- Errors in identification of sequences
-- Little control at different steps of the process
-- Too much of a black box?
-Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
--
-It has also been suggested that manual curation of classic and genomic alignments produces better phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+In contrast, most published phylogenies to date have been infered using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
+There seems to be a preference in the field of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data VS fine curation from human experts) would be to rely on published manually curated homology hypotheses.
-This expert-curated alignments can be enriched and updated by incorporating newly released data from public molecular databases.
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine curation from human experts) would be to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
 
-As of April 2014, TreeBASE hosted a bit more than 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE about)}.
-This provides an untapped source of valuable knowledge with the potential to update phylogenetic relationships in different regions of the tree of life.
+As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
+This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
 
-The OToL tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates the phylogenies from TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+The OToL tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
 This provides a loose means of linking the tree with the exact alignment that generated it.
-This can be a difficult task when multiple alignments were originally deposited and no clear link between alignment and tree was provided in TreeBASE.
 
 Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, each one of these data repositories follow their own system for taxon identification.
-A real challenge is to find an effective way to link data from across databases that belongs to the same taxon.
-OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc.
+Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
+OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
 
-Physcraper is a python pipeline that uses the OToL metadata system to connect databases through taxon identification numbers.
-
-It's main goal is to connect public phylogenetic relationships in OToL, with alignments from TreeBASE and GenBank DNA data.
-
-It also allows automatizing and standardizing the comparison of phylogenetic hypotheses.
+Physcraper is a python pipeline that relies on the OToL metadata system to connect databases through taxon identification numbers.
+It's main functionality is to connect phylogenies stored in the OToL phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Because of its design, it allows taking advantage of the many resources provided by the OToL.
+For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships.
 
 This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
-And an effort to more effectively connect the world of big data in biology.
+And an effort to more effectively make big data connections for the advantage of phylogenetics and biology in general.
 
@@
 
-\hypertarget{the-input-a-study-tree-and-an-alignment}{%
-\subsection{The input: a study tree and an alignment}\label{the-input-a-study-tree-and-an-alignment}}
+\hypertarget{the-input-a-tree-and-an-alignment}{%
+\subsection{The input: a tree and an alignment}\label{the-input-a-tree-and-an-alignment}}
 
@@
 \item
-  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
+  In order to take advantage of the OToL tools, the input tree needs to be stored in the
+  OToL phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}.
+  Currently, only trees connected to
+  a published study can be stored in the phylesystem (although there are plans to
+  allow storage of non published trees). The main
   reason for this is that trees in phylesystem have a set of user defined characteristics
   that are essential for automatizing the phylogeny update process.
-The most relevant of these being the definition of ingroup and outgroup.
+The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
 Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the next steps of the pipeline.
-  A user can choose from the 1216 published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}).
-If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
+This allows to automatically set the root for the updated tree on the final steps of the pipeline.
+  A user can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OToL website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OToL's website about}).
+If the published tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
 \item
@@
   alignments are usually stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
-  DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
+  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal were the tree was originally published.
   If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
-  either from the TreeBASE website (\url{https://treebase.org/})
-  or through the TreeBASE GitHub repository (SuperTreeBASE; \url{https://github.com/TreeBASE/supertreebase}).
-  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
-  downloaded by the user, and it's local path has to be provided as an argument.
+  either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
+  or through the TreeBASE GitHub repository (SuperTreeBASE; \href{https://github.com/TreeBASE/supertreebase}{github.com/TreeBASE/supertreebase}).
+  If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
 \item
@@
 \item
+  Technically, any DNA molecular database can be used to search for new sequences.
+  However, we rely on the GenBank database.
+  The new sequence search can be performed on the remote database or in a local database.
+\item
   The next step is to identify the search taxon within the reference taxonomy.
@@
   To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (\url{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}).
+I FORGOT TO ASK YOU ABOUT THIS CITATION!
 
@@
 \item
-  The Basic Local Alignment Search Tool, BLAST {[}Altschul \emph{et al.} (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
+  The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}, \protect\hyperlink{ref-altschul1997gapped}{1997}) is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the accepted sequences on the alignment.
-  The blastn function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database searches.
-  A modified biopython blast function is used for web-based searches.
+  database, and the sequences on the checked alignment.
+  The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database sequence searches.
+  For remote-database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function in the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternate BLAST URL.
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
 \item
-  The DNA sequence similarity search can be done on a local database that is easily
-  setup by the user. In this case, the blastn function is used to performs the similarity search (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}).
-\item
-  The search can also be performed remotely, on the NCBI database. In this case, the
-  bioPython BLAST function was modified to accepts is used to perform the similarity search.
-\item
-  A pairwise alignment-against-all BLAST search is performed. This means that each sequence
+  A pairwise BLAST search is performed. This means that each sequence
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
@@
 \item
-  Matched sequences below an e-value, percentage similarity, and outside a minimum
-  and maximum length threshold are discarded. \textbf{\emph{REPORT THE DEFAULT VALUES AND DESCRIBE WHAT THEY MEAN}} This filtering leaves out genomic sequences.
-  All acepted sequences are asigned an internal identifier, and are further filtered.
+  Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
+  This filtering guarantees that genomic sequences are not included.
+  All acepted sequences are assigned an internal identifier, and are further filtered.
 \item
-  Because the original alignments usually lack database id numbers, a filtering
+  Because the original alignments usually lack taxonomic identification numbers, an extra filtering
   step is needed. Accepted sequences that belong to the
@@
   Among the remaining filtered sequences, there are usually several exemplars per taxon.
-  Although it can be useful to keep some of them to, for example, investigate monophyly
+  Although it can be useful to keep some of them, for example, to investigate monophyly
   within species, there can be hundreds of exemplar sequences per taxon for some markers.
@@
   Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the alignment are blasted.
-\textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
+This can be done iteratively, but by default only sequences in the checked alignment are blasted.
+I ALSO FORGOT TO ASK: \textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
 \item
-  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
+  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (default to physcraper/taxonomy), so they are accesible for further runs.
 \item
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is implemented to perform alignments.
+  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is implemented to perform sequence alignments.
 \item
@@
 \item
-  The final alignment is not further processed automatically. We encourage users
+  The final alignment is not further processed automatically. So, we encourage users
   to check it either by eye and perform manual refinement or using any of the many
@@
 \begin{itemize}
+\tightlist
 \item
-  A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
-  with 100 classic rapid bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
-  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine \emph{et al.} (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
+  A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}), using default settings such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with default method.
+Currently only the number of bootsrap replicates can be modified by the user.
 \item
@@
 \item
-  Bootstrap results are summarized with Dendropy ADD CITATION
+  Bootstrap results are summarized with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010})
 \item
@@
 \item
-  Robinson Foulds weighted and unweighted metrics ARE CALCULATED WITH DENDROPY TOO.
+  Robinson Foulds weighted and unweighted metrics are calculated with Dendropy functions.
 \item
-  Describe what a conflict analysis is: Node by node comparison between the original and updated tree, and the synthetic OToL using the conflict API of otol CITE REDELINGS AND HOLDER ({\textbf{???}} and holder).
+  Finally a conflict analysis is performed.
+This is basically a node by node comparison between the the synthetic OToL and the original and updated tree individually.
+This is performed with OToL's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
 \item
@@
 \item
-  A SUGGESTED DEFAULT ROOTING BASED ON THE OPEN TREE TAXONOMY is implemented for now.
-DESCRIBE HOW IT WORKS.
-SAY THAT IT IS A PROBLEM.
-Automatic rooting is not that smart yet.
-The best way right now is for users to define outgroups so trees are better rooted.
-\item
-  Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
-This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
-\item
-  Conflict information can only be generated in the context of the whole Open Tree of
-  Life. Otherwise, it is not really possible to get conflict data.
-  \textbf{\emph{- One way to compare two independent phylogenetic trees is to compare them both to
-  the synthetic OToL and then measure how well they do against each other}}
+  A suggested default rooting based on OToL's taxonomy is implemented for now.
+It takes one descendant from the corresponding induced subtree in OToL's synthetic tree and then finds the MRCA of those in th eupdated trees.
+However, if the updated tree changes expectations from taxonomy, the root will no longer be accurate.
+Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
+The best way right now is for users to define outgroups so trees are accurately rooted.
 \end{itemize}
@@
 
-To exemplify the utility of physcraper we will address two use-case scenarios.
+We will illustrate the utility of physcraper in here with two use-case scenarios.
 One in which the user is interested in a particular group.
 Another one in which the user is interested in a particular phylogeny.
-A tutrial as well as illustrated examples of steps needed to perform a physcraper analyses are available elsewhere.
+A tutorial as well as illustrated examples of commands for every step needed to perform a physcraper analysis are available elsewhere.
 
@@
 A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three phylogenetic trees showing relationships within the hollies published.
+The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
 The first one has been made available in TreeBASE as well as in the OToL phylesystem and is part of the synthetic tree. It samples 48 species.
@@
 Data repositories hold more information than meets the eye.
-Besides the actual data, they have other types of information that can be used for the advantage of science.
+Besides the actual data, they are rich sources of information that can be used for the advantage of biology and science in general.
 
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on data bases,
-exposing new comers to expert understanding about the data.
-
-There are many tools that are making use of DNA data repositories in different ways.
+Many pipelines are making use of DNA data repositories in different ways.
 Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data - getting real and clean homologs.
-Others focus on refinement of the alignment.
+Some focus on accurate ways of mining the data -- getting real and clean homologs.
+Others focus on refining the alignment.
 Most focus on generating full trees \emph{de novo}, mainly for regions of the Tree of
@@
 regions that have been already studied and that have phylogenetic data already.
+However, expert phylogenetic knowledge is also an important source of data in public
+and open repositories that is potentiallly being diregarded.
 
-All these tools are great efforts for advancing towards reproducibility in phylogenetics,
+All these tools are key efforts for advancing towards reproducibility in phylogenetics,
 a field that has been largely recognised as somewhat artisanal.
-We propose adding focus to other sources of information available from data repositories.
-Taking advantage of public DNA data bases have been the main focus. However, phylogenetic knowledge is
-also accumulating fast in public and open repositories.
-In this way, the physcraper pipeline can be complemented with other tools that have
-been developed for other purposes.
+In here, we highlight the potential of other sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+community to build up our phylogenetic knowledge, using not only data accumulated in
+molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+While not generating full phylogenies \emph{de novo}, pyscraper is still capable of generating new phylogenetic knowledge.
+In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??).
+It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
 
-We emphasize that physcraper takes advantage of the knowledge and intuition of the expert
-community to build upon phylogenetic knowledge, using not only data accumulated in
-DNA repositories, but phylogenetic knowledge accumulated in tree repositories.
-This might help generate new phylogenetic data. But physcraper does not seek to generate full phylogenies \emph{de novo}.
-
-Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
-Mention statistics provided by other tools: PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
-Compare and discuss.
-
-How is physcraper already useful:
-- to mine targeted sequences, in this way it is similar to baited analyses from PHLAWD and pyPHLAWD. Phylota does not do baited analyses, I think, only clustered analyses.
-- Finding
-
-How can it be used for the advantage of the field:
-- rapid phylogenetic placing of newly discovered species, as mentioned in Webb \emph{et al.} (\protect\hyperlink{ref-webb2010biodiversity}{2010})
-- obtain trees for ecophylogenetic studies, as mentioned in Helmus \& Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
-- one day could be used to sistematize nucleotide databases, such as Genbank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro \& Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus \emph{et al.} (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})
-
-Things that physcraper does not do:
-- analyse the whole GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017}).
-There are already some very good tools that do that.
-- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez \emph{et al.} (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
-- it is not a tree repo, as phylota is, mentioned in Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2014evominer}{2014})
+Physcraper has the potential to be applied for the advantage of the field to rapidly
+place newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
+obtain trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
+help to sistematize molecular databases, i.e., curate taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
+and rapidly generate custom species trees for downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
 
@@
 
-We acknowledge contributions from
+The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No.~ACI-1429783).
 
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No.~ACI-1429783).
+We acknowledge contributions from \ldots{}
 
@@
 
+\leavevmode\hypertarget{ref-altschul1997gapped}{}%
+Altschul, S.F., Madden, T.L., Schäffer, A.A., Zhang, J., Zhang, Z., Miller, W.
+\& Lipman, D.J.
+(1997).
+Gapped blast and psi-blast: A new generation of protein database search programs.
+\emph{Nucleic acids research}, \textbf{25}, 3389--3402.
+
 \leavevmode\hypertarget{ref-antonelli2017toward}{}%
@@
 
-\leavevmode\hypertarget{ref-deepak2014evominer}{}%
-Deepak, A., Fernández-Baca, D., Tirthapura, S., Sanderson, M.J.
-\& McMahon, M.M.
-(2014).
-EvoMiner: Frequent subtree mining in phylogenetic databases.
-\emph{Knowledge and Information Systems}, \textbf{41}, 559--590.
+\leavevmode\hypertarget{ref-cock2009biopython}{}%
+Cock, P.J., Antao, T., Chang, J.T., Chapman, B.A., Cox, C.J., Dalke, A., Friedberg, I., Hamelryck, T., Kauff, F., Wilczynski, B.
+\& others.
+(2009).
+Biopython: Freely available python tools for computational molecular biology and bioinformatics.
+\emph{Bioinformatics}, \textbf{25}, 1422--1423.
+
+\leavevmode\hypertarget{ref-drew2013lost}{}%
+Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez, R., Katz, L.A., Crandall, K.A., Hibbett, D.S.
+\& Soltis, D.E.
+(2013).
+Lost branches on the tree of life.
+\emph{PLoS biology}, \textbf{11}.
 
@@
 Federhen, S. (2003). The taxonomy project. \emph{The NCBI Handbook}.
-
-\leavevmode\hypertarget{ref-felsenstein1985confidence}{}%
-Felsenstein, J. (1985). Confidence intervals on phylogenetics: An approach using bootstrap. \emph{Evolution}, \textbf{39}, 783--791.
 
@@
 
-\leavevmode\hypertarget{ref-lemoine2018renewing}{}%
-Lemoine, F., Entfellner, J.-B.D., Wilkinson, E., Correia, D., Felipe, M.D., De Oliveira, T.
-\& Gascuel, O.
-(2018).
-Renewing felsenstein's phylogenetic bootstrap in the era of big data.
-\emph{Nature}, \textbf{556}, 452--456.
+\leavevmode\hypertarget{ref-magee2014dawn}{}%
+Magee, A.F., May, M.R. \& Moore, B.R. (2014). The dawn of open access to phylogenetic data. \emph{PLoS One}, \textbf{9}.
 
@@
 
+\leavevmode\hypertarget{ref-morrison2006multiple}{}%
+Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
+
 \leavevmode\hypertarget{ref-morrison2009would}{}%
@@
 
-\leavevmode\hypertarget{ref-ranwez2009phyloexplorer}{}%
-Ranwez, V., Clairon, N., Delsuc, F., Pourali, S., Auberval, N., Diser, S.
-\& Berry, V.
-(2009).
-PhyloExplorer: A web server to validate, explore and query phylogenetic trees.
-\emph{BMC evolutionary biology}, \textbf{9}, 108.
+\leavevmode\hypertarget{ref-redelings2017supertree}{}%
+Redelings, B.D. \& Holder, M.T. (2017). A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species. \emph{PeerJ}, \textbf{5}, e3058.
 
 \leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A.
-\& Wehe, A.
-(2008).
-The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research.
-\emph{Systematic Biology}, \textbf{57}, 335--346.
-Retrieved from \url{https://doi.org/10.1080/10635150802158688}
+Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A. \& Wehe, A. (2008). The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. \emph{Systematic Biology}, \textbf{57}, 335--346.
 
@@
 
+\leavevmode\hypertarget{ref-sukumaran2010dendropy}{}%
+Sukumaran, J. \& Holder, M.T. (2010). DendroPy: A python library for phylogenetic computing. \emph{Bioinformatics}, \textbf{26}, 1569--1571.
+
 \leavevmode\hypertarget{ref-vos2012nexml}{}%

### 2020-06-26 2-body.Rmd
format

@@
 ```
-```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and update tree of the Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 knitr::include_graphics("docs/figs/cotree-plot2-1.png")
@@
 
+\newpage
 # References

### 2020-06-26 2-body.Rmd
notes pt 1

@@
 
-From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database [@benson2000genbank; @wheeler2000database], the TreeBASE repository [@piel2009treebase] and the Open Tree of Life curating system [@mctavish2015phylesystem], are still accumulating data, and there is currently no straightforward way to automatically connect their data.
+Public biological databases provide an amazing resource for evolutionary estimation.
+By linking molecular data, available from database such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
+the TreeBASE repository [@piel2009treebase] and the Open Tree of Life datastore [@mctavish2015phylesystem], we can place new biological data in evolutionary context.
+However, it has been challenging to connect new sequence data with existing estimates of evolutionary relationships.
 
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns coveres about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-Yet, publicly avilable phylogenetic relationships still cover about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data avilable a decade ago.
-While it is true that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn], most recently published large trees have been made available in recent years, indicating a lag between the amount of new DNA data generated and its analysis in a phylogenetic context.
+Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
+Some of this discrepance is because that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
+However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
 
 Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally refered to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
+Generally referred to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
 Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
@@
 
-Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been infered using alignments that have been assembled and curated "by hand" [@morrison2009would].
-There seems to be a preference in the field of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
+Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
+Automated approaches often to suffer from errors in homology inference that are rapidly caught by human oversight. ()
+Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships.
+\comment(Citations...
+I think there are some Scott Edwards papers...https://www.pnas.org/content/109/37/14942, maybe Roajs et al...)
+It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine curation from human experts) would be to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is be to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
 This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
@@
 
-The OToL tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+The OpenTree tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
 However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
@@
 
-Ultimately, linkage of original alignment with its corresponding phylogeny has to be done by a human curator.
+Often linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
 Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OToL's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
+OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
 
-Physcraper is a python pipeline that relies on the OToL metadata system to connect databases through taxon identification numbers.
-It's main functionality is to connect phylogenies stored in the OToL phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
-Because of its design, it allows taking advantage of the many resources provided by the OToL.
+Physcraper is a python pipeline that relies on the OpenTree metadata system to connect databases through taxon identification numbers.
+It's main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Because of its design, it allows taking advantage of the many resources provided by the OpenTree.
 For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships.
@@
 
-- In order to take advantage of the OToL tools, the input tree needs to be stored in the
-  OToL phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem).
+- In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
+  OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem).
   Currently, only trees connected to
@@
 This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  A user can choose from among the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OToL website ([See OToL's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
-If the published tree you are interested in updating is not in there, you can upload it via OToL's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
+  A user can choose from among the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
 - The alignment should be a gene alignment that was used to generate the tree. The original
@@
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
+  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
 I FORGOT TO ASK YOU ABOUT THIS CITATION!
@@
 - Finally a conflict analysis is performed.
-This is basically a node by node comparison between the the synthetic OToL and the original and updated tree individually.
-This is performed with OToL's conflict Application Programming Interface [@redelings2017supertree].
+This is basically a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
+This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
 - For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
-- A suggested default rooting based on OToL's taxonomy is implemented for now.
-It takes one descendant from the corresponding induced subtree in OToL's synthetic tree and then finds the MRCA of those in th eupdated trees.
+- A suggested default rooting based on OpenTree's taxonomy is implemented for now.
+It takes one descendant from the corresponding induced subtree in OpenTree's synthetic tree and then finds the MRCA of those in th eupdated trees.
 However, if the updated tree changes expectations from taxonomy, the root will no longer be accurate.
@@
 The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
-The first one has been made available in TreeBASE as well as in the OToL phylesystem and is part of the synthetic tree. It samples 48 species.
+The first one has been made available in TreeBASE as well as in the OpenTree phylesystem and is part of the synthetic tree. It samples 48 species.
 The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-The most recent one has been made available in the OToL Phylesystem and in the DRYAD repository. It is the best sampled yet, with 200 species. However, it has not been added to the syntehtic tree yet.
+The most recent one has been made available in the OpenTree Phylesystem and in the DRYAD repository.
+It is the best sampled yet, with 200 species.
+However, it has not been added to the syntehtic tree yet.
 This also makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
@@
 
-DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OTOL
+DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OpenTree
 RF DISTANCE INTERPRETATION
@@
 
-FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OTOL.
+FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OpenTree.
 

### 2020-06-26 2-body.Rmd
more notes

@@
   OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem).
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of non published trees). The main
+  The main
   reason for this is that trees in phylesystem have a set of user defined characteristics
@@
 This allows to automatically set the root for the updated tree on the final steps of the pipeline.
+  Currently, only trees connected to
+  a published study can be stored in the phylesystem (although there are plans to
+  allow storage of unpublished trees). 
   A user can choose from among the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
@@
 
-## DNA sequence search and cleaning
+## DNA sequence search and filtering
 
 - Technically, any DNA molecular database can be used to search for new sequences.
-  However, we rely on the GenBank database.
+  By default we rely on the GenBank database.
   The new sequence search can be performed on the remote database or in a local database.
@@
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
+  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy.
+\comment(hmm, a figure might be good for this.
+I can work on one!) This is known as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) \comment(prob need a cite here).
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
@@
 I FORGOT TO ASK YOU ABOUT THIS CITATION!
+\comment( it's here!
+https://bdj.pensoft.net/article/12581/instance/3545675/)
 
@@
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
+  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
 This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
 - Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
-  This filtering guarantees that genomic sequences are not included.
-  All acepted sequences are assigned an internal identifier, and are further filtered.
+These parameters can be configured for each run.
+  This filtering guarantees that whole genome sequences are not included, as they 
+  All accepted sequences are assigned an internal identifier, and are further filtered.
 - Because the original alignments usually lack taxonomic identification numbers, an extra filtering
@@
 This can be done iteratively, but by default only sequences in the checked alignment are blasted.
-I ALSO FORGOT TO ASK: ***Is there an argument to control the number of cycles of blast searches with new sequences?***
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (default to physcraper/taxonomy), so they are accesible for further runs.
+I ALSO FORGOT TO ASK: ***Is there an argument to control the number of cycles of blast searches with new sequences?*** \comment(there's only two options - one cycle or, -r where it runs until it doesn't find anything new)
+- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
 - A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.

### 2020-06-26 2-body.Rmd
more notes

@@
 
-Data repositories hold more information than meets the eye.
-Besides the actual data, they are rich sources of information that can be used for the advantage of biology and science in general.
+Data repositories hold even more information than meets the eye.
+Besides the actual data, they are rich sources of metadata that can be used for the advantage of biology and science in general.
 
@@
 However, expert phylogenetic knowledge is also an important source of data in public
-and open repositories that is potentiallly being diregarded.
+and open repositories that is not being used to its full potential.
 
 All these tools are key efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-In here, we highlight the potential of other sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
-community to build up our phylogenetic knowledge, using not only data accumulated in
+a field that has relied on processes which are somewhat artisanal.
+In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+community to build up our phylogenetic knowledge, we can use not only data accumulated in
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-While not generating full phylogenies *de novo*, pyscraper is still capable of generating new phylogenetic knowledge.
-In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??).
+While not generating full phylogenies *de novo*, physcraper is still capable of generating new phylogenetic knowledge.
+In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??). \comment(hmmmm mayeb cut bc I don't know!)
 It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
@@
 obtain trees for ecophylogenetic studies [@helmus2012phylogenetic],
-help to sistematize molecular databases, i.e., curate taxonomic assignations [@san2010molecular],
+help to systematize molecular databases, i.e., curate taxonomic assignations [@san2010molecular],
 and rapidly generate custom species trees for downstream analyses [@stoltzfus2013phylotastic].
@@
 The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
-
+\comment(need to add teh OPenTree ABI grant..)
 We acknowledge contributions from ...

### 2020-06-26 2-body.Rmd
comments on everythin except the xamples

@@
   All accepted sequences are assigned an internal identifier, and are further filtered.
-- Because the original alignments usually lack taxonomic identification numbers, an extra filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
-- Among the remaining filtered sequences, there are usually several exemplars per taxon.
+
+- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
+Longer sequences belonging to the
+  same taxon as the original sequence will be considered further for analysis.
+- Among the filtered sequences, there are often several representatives per taxon.
   Although it can be useful to keep some of them, for example, to investigate monophyly
@@
 - Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
-- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
+- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 
-- The software MUSCLE [@edgar2004muscle] is implemented to perform sequence alignments.
+- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments.
 - First, all new sequences are aligned using default MUSCLE options.
@@
   tools for alignment processing, to eliminate columns with no information.
+- Users may also use physcraper only to gather sequence matches, and apply their own preferred alignemnt and phylogenetic inference tools.
 
@@
 - The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
-- Bootstrap results are summarized with Dendropy [@sukumaran2010dendropy]
+- Bootstrap results are summarized with Dendropy [@sukumaran2010dendropy].
 - The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
-- Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
-- Robinson Foulds weighted and unweighted metrics are calculated with Dendropy functions.
+- Tips on all trees generated by physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
+- Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions [@sukumaran2010dendropy].
 - Finally a conflict analysis is performed.
@@
 This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
-- For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
+- For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
 - A suggested default rooting based on OpenTree's taxonomy is implemented for now.
-It takes one descendant from the corresponding induced subtree in OpenTree's synthetic tree and then finds the MRCA of those in th eupdated trees.
-However, if the updated tree changes expectations from taxonomy, the root will no longer be accurate.
+This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
+However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate.
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.

### 2020-06-29 2.2-methods-ideas.Rmd
methods extras

# What does physcraper do?

Takes an alignment of any number of taxa.

Finds the mrca from the opentree synthetic tree, so it is not a taxonomic mrca, but a phylogenetic one (whenever there is phylogenetic information).

Blast sequences from the alignment, within the mrca context.

Blast can be done directly on NCBI (very slow, as all blast searches are) or on a copy of the NCBI database installed on a server (way faster).

Blasted sequences come back with NCBI taxonomy ids, so we need to match them to ott ids (using an api call).

Why not use phlawd or supersmart?

They are not integrated to the open tree workflow
physcraper allows tagging

### 2020-06-30 _TODO.Rmd
more intro
cleaning todo's

@@
 
+- [ ] ilex remote example started on Sun Jun 28 at 20:07:32 PDT 2020 and finished on Mon Jun 29 at 06:33:57 PDT 2020
+
+- [ ] on methods, bootstrap summary part cite:
+ SumTrees: Phylogenetic Tree Summarization Version 4.4.0, from DendroPy 4.4.0 (main-bc20409)
+
 - [ ] intro: challenge in connecting different parts of the database with other databases that might refer to the same thing.
-For example taxon ids represent onw challenge in this sense, but it has been solved by the otol taxonomy project.
+For example taxon ids represent own challenge in this sense, but it has been solved by the otol taxonomy project.
 Othe rchallenge is the study ids.

### 2020-07-01 2-body.Rmd
celaning abstract

@@
 Public biological databases provide an amazing resource for evolutionary estimation.
-By linking molecular data, available from database such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
-the TreeBASE repository [@piel2009treebase] and the Open Tree of Life datastore [@mctavish2015phylesystem], we can place new biological data in evolutionary context.
-However, it has been challenging to connect new sequence data with existing estimates of evolutionary relationships.
+By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
+the TreeBASE repository [@piel2009treebase] and the Open Tree of Life datastore [@mctavish2015phylesystem], we can place new biological data in an evolutionary context.
+However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
 
@@
   allow storage of unpublished trees).
-  A user can choose from among the `r rotl::tol_about()$num_source_trees` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+  A user can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
 If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).

### 2020-07-01 2.4-discussion-ideas.Rmd
celaning abstract

@@
 - it is not a tree repo, as phylota is, mentioned in @deepak2014evominer
+
+***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
+the ones that do not know about phylogenetic methods and the ones that might know
+about phylogenetic methods but do not know much about a certain biological group.***
+\comment(hmm yeah, I think it is probably better for the latter - phylogeneticists who are tired of blast searching :P)

### 2020-07-01 2-body.Rmd
cleaning comments

@@
 In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
-Automated approaches often to suffer from errors in homology inference that are rapidly caught by human oversight. ()
+Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
 Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships.
-\comment(Citations...
-I think there are some Scott Edwards papers...https://www.pnas.org/content/109/37/14942, maybe Roajs et al...)
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.].
 It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
@@
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy.
-\comment(hmm, a figure might be good for this.
-I can work on one!) This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) \comment(prob need a cite here).
+  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomy tool (<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>).
-I FORGOT TO ASK YOU ABOUT THIS CITATION!
-\comment( it's here!
-https://bdj.pensoft.net/article/12581/instance/3545675/)
+  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool [Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
 
@@
 - Reverse complement sequences are identified and translated.
-- Users can choose to perform a more "cycles" of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the checked alignment are blasted.
-I ALSO FORGOT TO ASK: ***Is there an argument to control the number of cycles of blast searches with new sequences?*** \comment(there's only two options - one cycle or, -r where it runs until it doesn't find anything new)
-- Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
+- Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
+By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
+- Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
 - A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
@@
 While not generating full phylogenies *de novo*, physcraper is still capable of generating new phylogenetic knowledge.
-In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??). \comment(hmmmm mayeb cut bc I don't know!)
 It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
@@
 
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No. ACI-1429783).
-\comment(need to add teh OPenTree ABI grant..)
-We acknowledge contributions from ...
+Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838.
+Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
+ACI-1429783.
+
+We thank member of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
 

### 2020-07-01 1-abstract.Rmd
pdf

@@
 
-# Abstract
-
-1. Phylogenies are a key part of research in all areas of biology. Tools that automatize
+1. Phylogenies are a key part of research in many areas of biology. Tools that automatize
 some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
@@
 1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
-phylogenies by making use of public DNA sequence data and taxonomic information,
-providing a framework for comparison of published phylogenies with their updated versions.
+phylogenies by enriching underlying gene alignments with public DNA sequence data, and linking taxonomic information across databases.
+This provides a framework for comparison of published phylogenies with their updated versions, by using the conflict Application Programming Interface (API) function of the Open Tree of Life project.
 
@@
 hypothesis based on already available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
+Phylogeneticists and group specialists will find it useful as a tool to facilitate dataset gathering and comparison
 of alternative phylogenetic hypotheses (topologies).
-***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-the ones that do not know about phylogenetic methods and the ones that might know
-about phylogenetic methods but do not know much about a certain biological group.***
 
-1. Physcraper implements node by node/topology comparison of the the original and the updated
-trees using the conflict API of OToL, and summarizes differences.
-
-1. We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-
-1. Physcraper can be used with any OS. Detailed instructions for installation and
-use are available at <https://github.com/McTavishLab/physcraper>.
+1.
+We hope that the Physcraper workflow demonstrates the benefits of opening results in phylogenetics, encouraging more researchers to strive for better data sharing practices.
+Physcraper can be used with any OS.
+Detailed instructions for installation and
+use are available at [https://physcraper.readthedocs](https://physcraper.readthedocs.io/en/tutorial/index.html).

### 2020-07-01 2-body.Rmd
pdf

@@
 
-We thank member of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
+We thank members of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
 

### 2020-07-01 docs/submission-mee-2020-07-01.tex
pdf

--- 2020-06-26 docs/submission-mee-2020-06-26.tex
@@
 
-\textbf{Word count}: 3170
+\textbf{Word count}: 3340
 
@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\begingroup\Large
+
+\textbf{Abstract}
+\endgroup
 
@@
 \item
-  Phylogenies are a key part of research in all areas of biology. Tools that automatize
+  Phylogenies are a key part of research in many areas of biology. Tools that automatize
   some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
@@
   Physcraper is an open‐source, command-line Python program that automatizes the update of published
-  phylogenies by making use of public DNA sequence data and taxonomic information,
-  providing a framework for comparison of published phylogenies with their updated versions.
+  phylogenies by enriching underlying gene alignments with public DNA sequence data, and linking taxonomic information across databases.
+  This provides a framework for comparison of published phylogenies with their updated versions, by using the conflict Application Programming Interface (API) function of the Open Tree of Life project.
 \item
@@
   hypothesis based on already available expert phylogenetic knowledge.
-  Phylogeneticists and group specialists will find it useful as a tool to facilitate comparison
+  Phylogeneticists and group specialists will find it useful as a tool to facilitate dataset gathering and comparison
   of alternative phylogenetic hypotheses (topologies).
-  \textbf{\emph{Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
-  the ones that do not know about phylogenetic methods and the ones that might know
-  about phylogenetic methods but do not know much about a certain biological group.}}
 \item
-  Physcraper implements node by node/topology comparison of the the original and the updated
-  trees using the conflict API of OToL, and summarizes differences.
-\item
-  We hope the physcraper workflow demonstrates the benefits of opening results in phylogenetics and encourages researchers to strive for better data sharing practices.
-\item
-  Physcraper can be used with any OS. Detailed instructions for installation and
-  use are available at \url{https://github.com/McTavishLab/physcraper}.
+  We hope that the Physcraper workflow demonstrates the benefits of opening results in phylogenetics, encouraging more researchers to strive for better data sharing practices.
+Physcraper can be used with any OS.
+Detailed instructions for installation and
+  use are available at \href{https://physcraper.readthedocs.io/en/tutorial/index.html}{https://physcraper.readthedocs}.
 \end{enumerate}
@@
 
-From molecular data to alignments and phylogenies, public biological data resources such as the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life curating system (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), are still accumulating data, and there is currently no straightforward way to automatically connect their data.
+Public biological databases provide an amazing resource for evolutionary estimation.
+By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
+the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
+However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
 
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), which in turns coveres about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), which in turns covers about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(See GenBank's release website)}.
-Yet, publicly avilable phylogenetic relationships still cover about 90 000 taxa {[}Hinchliff \emph{et al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}); CURRENT SYNTH TREE CITATION{]}, covering less than one third of the taxonomic diversity with data avilable a decade ago.
-While it is true that many phylogenies are not publicly shared (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}), most recently published large trees have been made available in recent years, indicating a lag between the amount of new DNA data generated and its analysis in a phylogenetic context.
+Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa {[}Hinchliff \emph{et al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}); CURRENT SYNTH TREE CITATION{]}, covering less than one third of the taxonomic diversity with data available a decade ago.
+Some of this discrepance is because that many phylogenies are not publicly shared (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}, @mctavish2018bioessay).
+However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
 
 Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally refered to as ``pipelines'', most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
+Generally referred to as ``pipelines'', most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
 Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
@@
 
-Pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been infered using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-There seems to be a preference in the field of the classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-It has also been suggested that manual curation of classic alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
+Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
+Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.{]}.
+It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine curation from human experts) would be to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is be to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
 This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
@@
 
-The OToL tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
+The OpenTree tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
 However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
@@
 \item
-  In order to take advantage of the OToL tools, the input tree needs to be stored in the
-  OToL phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}.
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of non published trees). The main
+  In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
+  OpenTree phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}.
+  The main
   reason for this is that trees in phylesystem have a set of user defined characteristics
@@
 This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  A user can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OToL website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OToL's website about}).
-If the published tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
+  Currently, only trees connected to
+  a published study can be stored in the phylesystem (although there are plans to
+  allow storage of unpublished trees).
+  A user can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OpenTree website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OpenTree's website about}).
+If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
 \item
@@
 
-\hypertarget{dna-sequence-search-and-cleaning}{%
-\subsection{DNA sequence search and cleaning}\label{dna-sequence-search-and-cleaning}}
+\hypertarget{dna-sequence-search-and-filtering}{%
+\subsection{DNA sequence search and filtering}\label{dna-sequence-search-and-filtering}}
 
@@
   Technically, any DNA molecular database can be used to search for new sequences.
-  However, we rely on the GenBank database.
+  By default we rely on the GenBank database.
   The new sequence search can be performed on the remote database or in a local database.
@@
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the tree, that is also a named clade in the reference taxonomy. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA).
+  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OToL taxonomy tool (\url{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}).
-I FORGOT TO ASK YOU ABOUT THIS CITATION!
+  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{Taxonomy-API-v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
 
@@
   taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database).
+  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
 This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
@@
   Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
-  This filtering guarantees that genomic sequences are not included.
-  All acepted sequences are assigned an internal identifier, and are further filtered.
+  These parameters can be configured for each run.
+  This filtering guarantees that whole genome sequences are not included, as they
+  All accepted sequences are assigned an internal identifier, and are further filtered.
 \item
-  Because the original alignments usually lack taxonomic identification numbers, an extra filtering
-  step is needed. Accepted sequences that belong to the
-  same taxon of the query sequence, and that are either identical or shorter than
-  the original sequence are discarded. Only longer sequences belonging to the
-  same taxon as the orignal sequence will be considered further for analysis.
+  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
+  Longer sequences belonging to the
+  same taxon as the original sequence will be considered further for analysis.
 \item
-  Among the remaining filtered sequences, there are usually several exemplars per taxon.
+  Among the filtered sequences, there are often several representatives per taxon.
   Although it can be useful to keep some of them, for example, to investigate monophyly
@@
 \item
-  Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences.
-This can be done iteratively, but by default only sequences in the checked alignment are blasted.
-I ALSO FORGOT TO ASK: \textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
+  Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
+By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
 \item
-  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (default to physcraper/taxonomy), so they are accesible for further runs.
+  Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
 \item
-  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
+  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 \end{itemize}
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is implemented to perform sequence alignments.
+  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used by default to perform sequence alignments.
 \item
@@
   tools for alignment processing, to eliminate columns with no information.
+\item
+  Users may also use physcraper only to gather sequence matches, and apply their own preferred alignemnt and phylogenetic inference tools.
 \end{itemize}
@@
 \item
-  Bootstrap results are summarized with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010})
+  Bootstrap results are summarized with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 \item
@@
 \item
-  Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
+  Tips on all trees generated by physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
 \item
-  Robinson Foulds weighted and unweighted metrics are calculated with Dendropy functions.
+  Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 \item
   Finally a conflict analysis is performed.
-This is basically a node by node comparison between the the synthetic OToL and the original and updated tree individually.
-This is performed with OToL's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
+This is basically a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
+This is performed with OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
 \item
-  For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
+  For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
 \item
-  A suggested default rooting based on OToL's taxonomy is implemented for now.
-It takes one descendant from the corresponding induced subtree in OToL's synthetic tree and then finds the MRCA of those in th eupdated trees.
-However, if the updated tree changes expectations from taxonomy, the root will no longer be accurate.
+  A suggested default rooting based on OpenTree's taxonomy is implemented for now.
+This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
+However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate.
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
@@
 
-\caption{Comparison of original tree and update tree of the Malvaceae.}\label{fig:fig-2}
+\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}\label{fig:fig-2}
 \end{figure}
@@
 However, expert phylogenetic knowledge is also an important source of data in public
-and open repositories that is potentiallly being diregarded.
+and open repositories that is not being used to its full potential.
 
 All these tools are key efforts for advancing towards reproducibility in phylogenetics,
-a field that has been largely recognised as somewhat artisanal.
-In here, we highlight the potential of other sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
-community to build up our phylogenetic knowledge, using not only data accumulated in
+a field that has relied on processes which are somewhat artisanal.
+In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+community to build up our phylogenetic knowledge, we can use not only data accumulated in
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-While not generating full phylogenies \emph{de novo}, pyscraper is still capable of generating new phylogenetic knowledge.
-In this way, the physcraper workflow can be complemented with other pipelines, to potentially increase speed (IS THIS TRUE??).
+While not generating full phylogenies \emph{de novo}, physcraper is still capable of generating new phylogenetic knowledge.
 It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
@@
 obtain trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-help to sistematize molecular databases, i.e., curate taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
+help to systematize molecular databases, i.e., curate taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
 and rapidly generate custom species trees for downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
@@
 
-The University of California, Merced cluster, MERCED (Multi-Environment Research Computer for Exploration and Discovery) supported by the National Science Foundation (Grant No.~ACI-1429783).
+Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838.
+Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
 
-We acknowledge contributions from \ldots{}
+We thank members of the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
 
@@
 \section{Data Avilability}\label{data-avilability}}
+
+\newpage
 
@@
 
+\leavevmode\hypertarget{ref-mctavish2018bioessay}{}%
+McTavish, E.J., Drew, B.T., Redelings, B.
+\& Cranston, K.A.
+(2017).
+How and Why to Build a Unified Tree of Life.
+\emph{BioEssays}, \textbf{39}.
+Retrieved April 10, 2018, from \url{https://onlinelibrary.wiley.com/doi/abs/10.1002/bies.201700114}
+
 \leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
@@
 Redelings, B.D. \& Holder, M.T. (2017). A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species. \emph{PeerJ}, \textbf{5}, e3058.
+
+\leavevmode\hypertarget{ref-rees2017automated}{}%
+Rees, J.A. \& Cranston, K. (2017). Automated assembly of a reference taxonomy for phylogenetic data synthesis. \emph{Biodiversity Data Journal}.
 
@@
 \emph{BMC evolutionary biology}, \textbf{9}, 37.
+
+\leavevmode\hypertarget{ref-song2012resolving}{}%
+Song, S., Liu, L., Edwards, S.V.
+\& Wu, S.
+(2012).
+Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
+\emph{Proceedings of the National Academy of Sciences}, \textbf{109}, 14942--14947.
 

### 2020-07-01 2-body.Rmd
adding todos

@@
 Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.].
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.
+bats paper, take citations from there].
 It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].

### 2020-07-01 _TODO.Rmd
adding todos

@@
 # next:
+
+- [ ] hammer home the message that gene alignments are the gold standard for phylogenetics in many ways:
+  - [ ] genomic data, uces, radseqs are not comparable to other data
+  - [ ] homology in genomic data is not fully known (assembly based on similarity)
+  - [ ] genomic is incorporating tests to try to curate alignments better
+  - [ ] but data from gene sequences can only be used in a gene alignment context, not genomic data
 

### 2020-07-01 _TODO.Rmd
cleaning todos

@@
 
-- [ ] Do not cite Felsenstein 1985 for bootstrap, just the call that we use for RAxML:
-rax_ex, "-T", "{}".format(self.config.num_threads),
-                                 "-m", "GTRCAT",
-                                 "-s", alignment,
-                                 "-p", "1", "-b", "1", "-#", str(num_reps),
-                                 "-n", "BOOT{}".format(self.date)
 - [ ] Finding a tree to update
@@
 - [ ] How does pyphlawd, or phylota do the exemplar per species choosing for the baited analyses?
-
-- [ ] bootsrap: is it classic [@felsenstein1985confidence] or transfer bootstrap expectation [@lemoine2018renewing]
 
@@
   - [x] how is it relevant for blast results
+
+- [x] Do not cite Felsenstein 1985 for bootstrap, just explain it while explaining each argument from the call that we use for RAxML:
+  rax_ex, "-T", "{}".format(self.config.num_threads),
+                                   "-m", "GTRCAT",
+                                   "-s", alignment,
+                                   "-p", "1", "-b", "1", "-#", str(num_reps),
+                                   "-n", "BOOT{}".format(self.date)
+
+- [x] bootsrap: is it classic [@felsenstein1985confidence] or transfer bootstrap expectation [@lemoine2018renewing]
+  - probably classic, but we are not citing, just explaining, see point above.

### 2020-07-01 2-body.Rmd
finished ejm comments

@@
 By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
-the TreeBASE repository [@piel2009treebase] and the Open Tree of Life datastore [@mctavish2015phylesystem], we can place new biological data in an evolutionary context.
+the TreeBASE repository [@piel2009treebase] and the Open Tree of Life (OpenTree) datastore [@mctavish2015phylesystem], we can place new biological data in an evolutionary context.
 However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
@@
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is be to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
 This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
@@
 
-The OpenTree tree repository [phylesystem; @mctavish2015phylesystem] automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
-However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
+The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
+If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
 This provides a loose means of linking the tree with the exact alignment that generated it.
 
-Often linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
+Often, linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
 Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
@@
 
-Physcraper is a python pipeline that relies on the OpenTree metadata system to connect databases through taxon identification numbers.
+Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through taxon identification numbers.
 It's main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+EXPLIAN WHY PYTHON IS BEST FOR THIS.
 Because of its design, it allows taking advantage of the many resources provided by the OpenTree.
@@
 If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
+- THE INPUT TREE CAN ALSO BE A PERSONAL ONE. A TNRS MATCHING NEEDS TO BE PERFORMED (AND A FILE GENERATED) IN ORDER TO UPDATE A PERSONAL UNPUBLISHED TREE.
 - The alignment should be a gene alignment that was used to generate the tree. The original
@@
   DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, `physcraper` can download it directly,
+  If the alignment is stored in TreeBase, Physcraper can download it directly,
   either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
@@
   along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
-- Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
+This information will be used later to store the whole sequences in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
+- Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
 These parameters can be configured for each run.
-  This filtering guarantees that whole genome sequences are not included, as they
+  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
   All accepted sequences are assigned an internal identifier, and are further filtered.
@@
 By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
-- Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
+- Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to "physcraper/taxonomy" folder), so they are accessible for further runs.
 - A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
@@
   tools for alignment processing, to eliminate columns with no information.
-- Users may also use physcraper only to gather sequence matches, and apply their own preferred alignemnt and phylogenetic inference tools.
+- Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
 
@@
 - The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
-- Bootstrap results are summarized with Dendropy [@sukumaran2010dendropy].
+- Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
 - The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
-- Tips on all trees generated by physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
+- Tips on all trees generated by Physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
 - Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions [@sukumaran2010dendropy].
@@
 
-We will illustrate the utility of physcraper in here with two use-case scenarios.
+We will illustrate the utility of Physcraper in here with two use-case scenarios.
 One in which the user is interested in a particular group.
 Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of commands for every step needed to perform a physcraper analysis are available elsewhere.
+A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available elsewhere.
 
@@
 However, it has not been added to the syntehtic tree yet.
-This also makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+This also makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
 Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
@@
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-While not generating full phylogenies *de novo*, physcraper is still capable of generating new phylogenetic knowledge.
+While not generating full phylogenies *de novo*, Physcraper is still capable of generating new phylogenetic knowledge.
 It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.

### 2020-07-01 docs/submission-mee-2020-07-01.tex
finished ejm comments

@@
 
-\textbf{Word count}: 3340
+\textbf{Word count}: 3394
 
@@
 By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
-the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
+the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OpenTree) datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
 However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
@@
 Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.{]}.
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.
+bats paper, take citations from there{]}.
 It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is be to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
 This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
@@
 
-The OpenTree tree repository (phylesystem; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies uploaded into TreeBASE, and stores metadata linking the tree to its corresponding alignment repository in TreeBASE.
-However, if there are multiple alignments, TreeBASE does not indicate how they were used to generate the tree.
+The Phylesystem (OpenTree's datastore) (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
+If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
 This provides a loose means of linking the tree with the exact alignment that generated it.
 
-Often linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
+Often, linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
 Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
@@
 \item
+  THE INPUT TREE CAN ALSO BE A PERSONAL ONE. A TNRS MATCHING NEEDS TO BE PERFORMED (AND A FILE GENERATED) IN ORDER TO UPDATE A PERSONAL UNPUBLISHED TREE.
+\item
   The alignment should be a gene alignment that was used to generate the tree. The original
@@
   DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
+  If the alignment is stored in TreeBase, Physcraper can download it directly,
   either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
@@
   along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
+This information will be used later to store the whole sequences in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
 \item
-  Matched sequences will be discarded if the fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length of sequences in the checked alignment (gaps dropped).
+  Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
   These parameters can be configured for each run.
-  This filtering guarantees that whole genome sequences are not included, as they
+  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
   All accepted sequences are assigned an internal identifier, and are further filtered.
@@
 \item
-  Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to physcraper/taxonomy), so they are accessible for further runs.
+  Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to ``physcraper/taxonomy'' folder), so they are accessible for further runs.
 \item
@@
 \item
-  Users may also use physcraper only to gather sequence matches, and apply their own preferred alignemnt and phylogenetic inference tools.
+  Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
 \end{itemize}
@@
 \item
-  Bootstrap results are summarized with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+  Bootstrap results are summarized with the SumTrees module of DendroPy (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 \item
@@
 \item
-  Tips on all trees generated by physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
+  Tips on all trees generated by Physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
 \item
@@
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-While not generating full phylogenies \emph{de novo}, physcraper is still capable of generating new phylogenetic knowledge.
+While not generating full phylogenies \emph{de novo}, Physcraper is still capable of generating new phylogenetic knowledge.
 It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.

### 2020-07-02 1-abstract.Rmd
kinsey comments

@@
 1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypothesis based on already available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate dataset gathering and comparison
+hypotheses based on already available expert phylogenetic knowledge.
+Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
 of alternative phylogenetic hypotheses (topologies).
@@
 1.
-We hope that the Physcraper workflow demonstrates the benefits of opening results in phylogenetics, encouraging more researchers to strive for better data sharing practices.
+We hope that the Physcraper workflow demonstrates the benefits of open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS.

### 2020-07-02 _TODO.Rmd
finished ejm comments
kinsey comments

@@
 
+- [ ] add interactive table of csv metadata to physcraper examples
+
 - [ ] hammer home the message that gene alignments are the gold standard for phylogenetics in many ways:
@@
 
-- [ ] on methods, bootstrap summary part cite:
+- [x] on methods, bootstrap summary part cite:
  SumTrees: Phylogenetic Tree Summarization Version 4.4.0, from DendroPy 4.4.0 (main-bc20409)
@@
 This is not fully resolved yet by otol, but only partially by linking to the alignment repo.
+
 - [ ] intro: programatic access.

### 2020-07-02 2-body.Rmd
comments
kinsey comments

@@
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turn covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
-Some of this discrepance is because that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
+Yet, we only have data on phylogenetic relationships available for about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
+Some of this discrepancy is explained by the many phylogenies that are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
 However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 # How does Physcraper work?
+
+The general workflow is described in FIGURE. Next, we will describe the technical details of each step.
 
@@
 
-- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments.
+- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
 - First, all new sequences are aligned using default MUSCLE options.

### 2020-07-02 2-body.Rmd
jessica comments

@@
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turn covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa covered in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota] which in turn represent only about 10% of all extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-Yet, we only have data on phylogenetic relationships available for about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
+Yet, the phylogeneyic potential of the database has not been attained yet, with data on phylogenetic relationships available for about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], representing only one third of the taxonomic diversity with data available a decade ago (or 3% of the whole diversity) ALSO COMPARE WITH NUMBER OF TAXA REPRESENTED IN GENEBANK TODAY.
 Some of this discrepancy is explained by the many phylogenies that are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].

### 2020-07-02 2-body.Rmd
kinsey comment pipelines

@@
 
-Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally referred to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
-Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely applied to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [Moniloformopses; @lehtonen2011towards].
+Computational tools are crucial to making sense of the massive amount of data in molecular databases.
+"Pipelines" that automate similarity based identification of DNA homologs and assembly of molecular homology hypotheses, i.e., molecular alignments [@sanderson2008phylota; @antonelli2017toward; @smith2009mega] have been a very important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+They have been widely applied to study the evolutionary relationships among different organisms, from vertebrates, to plants and bacteria (TABLE OF PIPELINES AND STUDIES THAT HAVE USED THEM??
+maybe supplementary data).
 
-Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
-Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
+Still, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
+GETTING AN ESTIMATE OF TOTAL PUBLISHED PHYLOGENIES WOULD BE NICE.
+Automated approaches often suffer from errors in homology inference that could be rapidly caught by human oversight. CITATION
 Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).

### 2020-07-02 2.1-intro-ideas.Rmd
kinsey comment pipelines

--- 2020-06-26 2.1-intro-ideas.Rmd
@@
 Most of these phylogenies are being constructed by people learning about the methods (aka students), so they want to know what is going on at each step and end up doing it manually.
+
+Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
+Generally referred to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
+Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
+
+Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely applied to study the evolutionary relationships among different organisms (TABLE??
+maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [Moniloformopses; @lehtonen2011towards].

### 2020-07-02 2.2-methods-ideas.Rmd
kinsey comment pipelines

@@
 physcraper allows tagging
+
+Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in @lemoine2018renewing.
+
+***- One way to compare two independent phylogenetic trees is to compare them both to
+the synthetic OToL and then measure how well they do against each other***
+
+- Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon.
+This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
+- Conflict information can only be generated in the context of the whole Open Tree of Life. Otherwise, it is not really possible to get conflict data.

### 2020-07-02 _TODO.Rmd
kinsey comment pipelines

@@
 # next:
+
+- [ ] Kinsey comment: reduce paragraph about history of pipelines and how they were developed, maybe send to discussion?
+  She suggests deleting the whole paragraph and just citing the pipelines at the beginning of L72
+  Focus on importance of pipelines for the context of physcraper
+    - [x] they do database wide homology search
+    - [ ] they forget about original alignments
+    - [x] do they take into account homology?
+    - [x] how pipelines are crucial to fulfill the phylogenetically informative potential of GenBank.
+    - [ ] but why have they failed
 

### 2020-07-09 2-body.Rmd
intro re-structure

--- 2020-07-02 2-body.Rmd
@@
 More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turn covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
 In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-Yet, we only have data on phylogenetic relationships available for about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
-Some of this discrepancy is explained by the many phylogenies that are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
+Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
+Some of this discrepance is because that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
 However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 - In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem).
+  OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem), or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (https://tree.opentreeoflife.org/curator/tnrs/).
   The main
@@
 If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
-- THE INPUT TREE CAN ALSO BE A PERSONAL ONE. A TNRS MATCHING NEEDS TO BE PERFORMED (AND A FILE GENERATED) IN ORDER TO UPDATE A PERSONAL UNPUBLISHED TREE.
-- The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
+  If the user is not ready to make the input tree public, `physcraper` can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
+- The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
+  alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
   DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal were the tree was originally published.
@@
 
-Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838.
-Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
+Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838, and ABI No. 1759846.
+Compute time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
 ACI-1429783.

### 2020-07-10 2-body.Rmd
more re-struring and edits

@@
 
-Public biological databases provide an amazing resource for evolutionary estimation.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide context for biological observations.
+Public biological databases provide an amazing resource for evolutionary estimation, but there is a lot of sequence data available that has never been incorporated into any phylogenetic estimates.
+
+GenBank contains over 217 million DNA sequences [(release number 238; June 15, 2020)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
+These sequences are estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa in GenBank [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota].
+This in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+
+Yet, publicly available estimates of phylogenetic relationships still only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife_2019_synth].
+These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
+
+Some of this discrepancy between data availability and phylogenetic estimates is because many phylogenies which are generated and published are not publicly shared in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
+However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
+Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
+
+We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
+By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees for of loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
+These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
+
+Physcraper leverages phylogentic data stored in Open Tree of Life and in Treebase
+The Open Tree of Life (https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
+Open Tree of Life aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.  
+Currently the tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
+
+The Open Tree of Life data store, [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373), contains more than 4,500 phylogenetic trees from published studies.  
+The tips in these trees are mapped a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
+This provides a resource for finding existing estimates of phylogenetic relationships, 
+and assessing which regions of the tree of life which are lacking available phylogenetic estimates.
+
 By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
 the TreeBASE repository [@piel2009treebase] and the Open Tree of Life (OpenTree) datastore [@mctavish2015phylesystem], we can place new biological data in an evolutionary context.
-However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
 
-More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota], which in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
-In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences [(See GenBank's release website)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa [@hinchliff2015synthesis; CURRENT SYNTH TREE CITATION], covering less than one third of the taxonomic diversity with data available a decade ago.
-Some of this discrepance is because that many phylogenies are not publicly shared [@drew2013lost; @magee2014dawn, @mctavish2018bioessay].
-However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
+While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
+One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species. 
+Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq and UCE's.
+This variety of approaches results in non-overlapping data sets across taxa. 
+Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016]
+In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
 
-Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally referred to as "pipelines", most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
-Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward], have been widely applied to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish [Sphyraenidae; @santini2013first], to a mega-phylogeny of almost 3 000 species of living ferns [Moniloformopses; @lehtonen2011towards].
-
-Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
-Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
-Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+Indeed, some systematics support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
 Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.
@@
 
-As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
+In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward].
+Physcraper shares a similar conceptual framework to Pumper [@izquierdo-carrasco_pumper:_2014], but that software is not currently supported or developed (*or runnable at all honestly...*)
+
+Data input availability:
+As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of around 100 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
 This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
@@
 
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through taxon identification numbers.
-It's main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers.
+ main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
 EXPLIAN WHY PYTHON IS BEST FOR THIS.
-Because of its design, it allows taking advantage of the many resources provided by the OpenTree.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships.
+By design, this approach focusses on interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
+For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
 
@@
 # References
+
+Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
+
+Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION

### 2020-07-10 render2-mee.Rmd

--- 2020-06-25 render2-submission-mee.Rmd
@@
 )
-library(wordcountaddin)
-word_count_body <- wordcountaddin::word_count(filename="2-body.Rmd")
+
 ```
@@
 
-**Word count**: `r word_count_body`
-
 **Manuscript prepared for submission to Methods in Ecology and Evolution**
@@
 \newpage
+
+\begingroup\Large
+**Abstract**
+\endgroup
 

### 2020-07-10 2-body.Rmd
kinsey comment pipelines
fix refs

@@
 
-This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
+ELABORATE MORE ON THIS, USE LESS CASUAL LANGUAGE: This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
 And an effort to more effectively make big data connections for the advantage of phylogenetics and biology in general.
@@
 ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
-wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/vignettes/malvaceae-images/cotree-plot2-1.png"
+wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/data-raw/malvaceae-images/cotree-plot2-1.png"
 mv cotree-plot2-1.png docs/figs/cotree-plot2-1.png

### 2020-07-10 docs/submission-mee-2020-07-10.tex
fix refs

--- 2020-07-01 docs/submission-mee-2020-07-01.tex
@@
 
-\textbf{Word count}: 3394
+\textbf{Word count}: 3769
 
@@
   Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-  hypothesis based on already available expert phylogenetic knowledge.
-  Phylogeneticists and group specialists will find it useful as a tool to facilitate dataset gathering and comparison
+  hypotheses based on already available expert phylogenetic knowledge.
+  Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
   of alternative phylogenetic hypotheses (topologies).
 \item
-  We hope that the Physcraper workflow demonstrates the benefits of opening results in phylogenetics, encouraging more researchers to strive for better data sharing practices.
+  We hope that the Physcraper workflow demonstrates the benefits of open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS.
@@
 
-Public biological databases provide an amazing resource for evolutionary estimation.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide context for biological observations.
+Public biological databases provide an amazing resource for evolutionary estimation, but there is a lot of sequence data available that has never been incorporated into any phylogenetic estimates.
+
+GenBank contains over 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(release number 238; June 15, 2020)}.
+These sequences are estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa in GenBank (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+This in turns covers about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
+
+Yet, publicly available estimates of phylogenetic relationships still only about 100 000 taxa ({\textbf{???}}; Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}).
+These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
+
+Some of this discrepancy between data availability and phylogenetic estimates is because many phylogenies which are generated and published are not publicly shared in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
+However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
+Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
+
+We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
+By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees for of loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
+These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
+
+Physcraper leverages phylogentic data stored in Open Tree of Life and in Treebase
+The Open Tree of Life (\url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
+Open Tree of Life aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+Currently the tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
+
+The Open Tree of Life data store, \href{https://academic.oup.com/bioinformatics/article/31/17/2794/183373}{Phylesystem}, contains more than 4,500 phylogenetic trees from published studies.
+The tips in these trees are mapped a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
+This provides a resource for finding existing estimates of phylogenetic relationships,
+and assessing which regions of the tree of life which are lacking available phylogenetic estimates.
+
 By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
 the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OpenTree) datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
-However, connecting new sequence data with existing estimates of evolutionary relationships represents a challenge.
 
-More than a decade ago, the National Center for Biodiversity Information (NCBI) molecular database, GenBank, released its version number 159 (April 15, 2007).
-With 72 million DNA sequences, it was estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa represented in it (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), which in turns covers about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
-In comparison, the current GenBank release number 238 (June 15, 2020) has tripled in size, hosting data for more than 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(See GenBank's release website)}.
-Yet, publicly available estimates of phylogenetic relationships still cover only about 90 000 taxa {[}Hinchliff \emph{et al.} (\protect\hyperlink{ref-hinchliff2015synthesis}{2015}); CURRENT SYNTH TREE CITATION{]}, covering less than one third of the taxonomic diversity with data available a decade ago.
-Some of this discrepance is because that many phylogenies are not publicly shared (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}, @mctavish2018bioessay).
-However, most recently published large trees have been made available in recent years, indicating there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
+While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
+One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
+Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq and UCE's.
+This variety of approaches results in non-overlapping data sets across taxa.
+Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences ({\textbf{???}})
+In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
 
-Many useful computational tools have been developed in an effort to make sense of the large amount of data in public molecular databases, as well as private ones.
-Generally referred to as ``pipelines'', most of these tools were motivated by the genomics revolution, to help researchers to deal with the massive amount of data, and automatically identify clusters of homologs for genomic assembly.
-Notably, this homolog DNA clusters can be used as homology hypotheses (i.e., molecular alignments) to reconstruct phylogenetic relationships.
-Pipelines that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}), have been widely applied to study the evolutionary relationships among different organisms (TABLE??
-maybe supplementary data), from a phylogeny of no more than 20 species of the family of barracuda fish (Sphyraenidae; Santini \& Sorenson \protect\hyperlink{ref-santini2013first}{2013}), to a mega-phylogeny of almost 3 000 species of living ferns (Moniloformopses; Lehtonen \protect\hyperlink{ref-lehtonen2011towards}{2011}).
-
-Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
-Many systemacists support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
+Indeed, some systematics support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
 Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.
@@
 
-As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of almost 105 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
+In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+Physcraper shares a similar conceptual framework to Pumper (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo-carrasco_pumper:_2014}{2014}), but that software is not currently supported or developed (\emph{or runnable at all honestly\ldots{}})
+
+Data input availability:
+As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of around 100 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
 This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
@@
 
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through taxon identification numbers.
-It's main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
+Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers.
+main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
 EXPLIAN WHY PYTHON IS BEST FOR THIS.
-Because of its design, it allows taking advantage of the many resources provided by the OpenTree.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships.
+By design, this approach focusses on interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
+For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
 
-This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
+ELABORATE MORE ON THIS, USE LESS CASUAL LANGUAGE: This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
 And an effort to more effectively make big data connections for the advantage of phylogenetics and biology in general.
@@
 \section{How does Physcraper work?}\label{how-does-physcraper-work}}
+
+The general workflow is described in FIGURE. Next, we will describe the technical details of each step.
 
@@
   In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}.
+  OpenTree phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}, or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (\url{https://tree.opentreeoflife.org/curator/tnrs/}).
   The main
@@
 If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
+  If the user is not ready to make the input tree public, \texttt{physcraper} can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
 \item
-  THE INPUT TREE CAN ALSO BE A PERSONAL ONE. A TNRS MATCHING NEEDS TO BE PERFORMED (AND A FILE GENERATED) IN ORDER TO UPDATE A PERSONAL UNPUBLISHED TREE.
-\item
-  The alignment should be a gene alignment that was used to generate the tree. The original
-  alignments are usually stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
+  The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
+  alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
   DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal were the tree was originally published.
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used by default to perform sequence alignments.
+  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
 \item
@@
 
-Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838.
-Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
+Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838, and ABI No.~1759846.
+Compute time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
 
@@
 \section{References}\label{references}}
+
+Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
+
+Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
 
@@
 
-\leavevmode\hypertarget{ref-lehtonen2011towards}{}%
-Lehtonen, S. (2011). Towards resolving the complete fern tree of life. \emph{PLoS One}, \textbf{6}.
+\leavevmode\hypertarget{ref-izquierdo-carrasco_pumper:_2014}{}%
+Izquierdo-Carrasco, F., Cazes, J., Smith, S.A.
+\& Stamatakis, A.
+(2014).
+PUmPER: Phylogenies Updated Perpetually.
+\emph{Bioinformatics}, \textbf{30}, 1476--7.
+Retrieved February 1, 2014, from \url{http://bioinformatics.oxfordjournals.org/content/early/2014/01/28/bioinformatics.btu053}
 
@@
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
-
-\leavevmode\hypertarget{ref-santini2013first}{}%
-Santini, F.
-\& Sorenson, L.
-(2013).
-First molecular timetree of billfishes (istiophoriformes: Acanthomorpha) shows a late miocene radiation of marlins and allies.
-\emph{Italian journal of zoology}, \textbf{80}, 481--489.
 

### 2020-07-10 2-body.Rmd
fix refs

@@
 
-Yet, publicly available estimates of phylogenetic relationships still only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife_2019_synth].
+Yet, publicly available estimates of phylogenetic relationships still cover only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth].
 These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
@@
 This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016]
+Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016misconceptions]
 In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
@@
 
-```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
+<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
 wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/data-raw/malvaceae-images/cotree-plot2-1.png"
 mv cotree-plot2-1.png docs/figs/cotree-plot2-1.png
-```
+``` -->
 ```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}

### 2020-07-10 docs/submission-mee-2020-07-10.tex
fix refs

@@
 
-\textbf{Word count}: 3769
+\textbf{Word count}: 3768
 
@@
 
-Yet, publicly available estimates of phylogenetic relationships still only about 100 000 taxa ({\textbf{???}}; Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}).
+Yet, publicly available estimates of phylogenetic relationships still cover only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}).
 These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
@@
 This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences ({\textbf{???}})
+Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016})
 In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
@@
 
+\leavevmode\hypertarget{ref-eaton2016misconceptions}{}%
+Eaton, D.A.R., Spriggs, E.L., Park, B.
+\& Donoghue, M.J.
+(2016).
+Misconceptions on Missing Data in RAD-seq Phylogenetics with a Deep-scale Example from Flowering Plants.
+\emph{Systematic Biology}, \textbf{66}, 399--412.
+Retrieved from \url{https://doi.org/10.1093/sysbio/syw092}
+
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
@@
 Morrison, D.A. (2009). Why would phylogeneticists ignore computerized sequence alignment? \emph{Systematic biology}, \textbf{58}, 150--158.
+
+\leavevmode\hypertarget{ref-opentreeoflife2019synth}{}%
+OpenTreeOfLife, Redelings, B., Reyes, L.L.S., Cranston, K.A., Allman, J., Holder, M.T.
+\& McTavish, E.J.
+(2019).
+Open tree of life synthetic tree.
+Retrieved from \url{https://doi.org/10.5281/zenodo.3937742}
 

### 2020-07-10 render2-mee.Rmd
fix refs
mk new affiliation

--- 2020-06-25 render2-submission-mee.Rmd
@@
   author2: Martha Kandziora
-  email2: martha.kandziora@mailbox.org
+  email2: kandziom@natur.cuni.cz
   author3: Emily Jane McTavish
@@
   affiliation: School of Natural Sciences, University of California, Merced
+  affiliation2: Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
   correspondence_address: Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
@@
 
-`r params$affiliation`
+`r params$affiliation2`
 
@@
 
+\begingroup\Large
+**Abstract**
+\endgroup
+
 ``` {r child = '1-abstract.Rmd'}

### 2020-07-10 render2-mee.Rmd
keywords

@@
 
-**Keywords**: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
+**Keywords**: gene tree, interoperability, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
 

### 2020-07-10 _TODO.Rmd
updating todos

@@
 
+- [ ] from phylotar paper: "requires stand-alone BLAST. Instructions for installing the BLAST+ suite can be found via NCBI [27]"
+
 - [ ] Kinsey comment: reduce paragraph about history of pipelines and how they were developed, maybe send to discussion?
@@
 - [ ] ilex remote example started on Sun Jun 28 at 20:07:32 PDT 2020 and finished on Mon Jun 29 at 06:33:57 PDT 2020
-
-- [x] on methods, bootstrap summary part cite:
- SumTrees: Phylogenetic Tree Summarization Version 4.4.0, from DendroPy 4.4.0 (main-bc20409)
 
@@
 
-- [ ] determining the root
+- [x] on methods, bootstrap summary part cite:
+ SumTrees: Phylogenetic Tree Summarization Version 4.4.0, from DendroPy 4.4.0 (main-bc20409)
+
+- [x] determining the root
   - [ ] write something about the need to set the root to be able to do comparisons
-  - [ ] say that utis hard to set it up automatically and that solutions are being sought
+  - [ ] say that it is hard to set it up automatically and that solutions are being sought
 

### 2020-07-10 2-body.Rmd
intro edit first 3 paragraphs

@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide context for biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but there is a lot of sequence data available that has never been incorporated into any phylogenetic estimates.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
+Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
 
-GenBank contains over 217 million DNA sequences [(release number 238; June 15, 2020)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt).
-These sequences are estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa in GenBank [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy; @sanderson2008phylota].
-This in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
+Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-Yet, publicly available estimates of phylogenetic relationships still cover only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth].
-These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
-
-Some of this discrepancy between data availability and phylogenetic estimates is because many phylogenies which are generated and published are not publicly shared in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
+Some of the discrepancy between data availability and phylogenetic estimates can be explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 # References
-
-Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
-
-Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION

### 2020-07-10 2.1-intro-ideas.Rmd
intro edit first 3 paragraphs

@@
+over 217 million DNA sequences [(release number 238; June 15, 2020)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt)
+
+This in turns covers about 10% of extant described biodiversity [taking a conservative estimate of extant diversity; @scott2011ncbi; @federhen2003taxonomy].
+
+Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
+In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated "by hand" [@morrison2009would].
+
+Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
+
 The PHYLOTA [@sanderson2008phylota] tool has been used to construct homolog DNA datasets for about 42 studies reconstructing phylogenetic trees from various organisms.

### 2020-07-10 docs/submission-mee-2020-07-10.tex
mk new affiliation
intro edit first 3 paragraphs

@@
 
-School of Natural Sciences, University of California, Merced
+Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
 
-email: \href{mailto:martha.kandziora@mailbox.org}{\nolinkurl{martha.kandziora@mailbox.org}}
+email: \href{mailto:kandziom@natur.cuni.cz}{\nolinkurl{kandziom@natur.cuni.cz}}
 
@@
 
-\textbf{Word count}: 3768
+\textbf{Word count}: 3702
 
@@
 
-\textbf{Keywords}: cross-connectivity, gene tree, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
+\textbf{Keywords}: gene tree, interoperability, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
 
@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide context for biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but there is a lot of sequence data available that has never been incorporated into any phylogenetic estimates.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
+Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate DO WE NEED CITATIONS FOR THIS?.
 
-GenBank contains over 217 million DNA sequences \href{https://ftp.ncbi.nih.gov/genbank/gbrel.txt}{(release number 238; June 15, 2020)}.
-These sequences are estimated to have the potential to resolve evolutionary relationships of most of the 241 000 distinct taxa in GenBank (about 98.05\% of taxa in the NCBI taxonomy; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-This in turns covers about 10\% of extant described biodiversity (taking a conservative estimate of extant diversity; Scott \protect\hyperlink{ref-scott2011ncbi}{2011}; Federhen \protect\hyperlink{ref-federhen2003taxonomy}{2003}).
+GenBank, the National Center for Biodiversity Information (NCBI) molecular database, release number 159 (from April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 
-Yet, publicly available estimates of phylogenetic relationships still cover only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}).
-These data comprise less than one third of the taxonomic diversity for which sequence data was available a decade ago!
+Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogeentically informative sequence data available in GenBank more than a decade ago.
 
-Some of this discrepancy between data availability and phylogenetic estimates is because many phylogenies which are generated and published are not publicly shared in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
+Some of the discrepancy between data availability and phylogenetic estimates can be explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 
-Automated sequence gathering pipelines have been an important incorporation to the field of phylogenetics in many ways, particularly because they represent a clear step towards reproducibility in the field.
-In contrast, most published phylogenies to date have been inferred using alignments that have been assembled and curated ``by hand'' (Morrison \protect\hyperlink{ref-morrison2009would}{2009}).
-
-Automated approaches often suffer from errors in homology inference that could rapidly caught by human oversight. CITATION
-
 \newpage
@@
 
-\leavevmode\hypertarget{ref-federhen2003taxonomy}{}%
-Federhen, S. (2003). The taxonomy project. \emph{The NCBI Handbook}.
-
 \leavevmode\hypertarget{ref-fragoso2017pilot}{}%
@@
 
-\leavevmode\hypertarget{ref-morrison2009would}{}%
-Morrison, D.A. (2009). Why would phylogeneticists ignore computerized sequence alignment? \emph{Systematic biology}, \textbf{58}, 150--158.
-
 \leavevmode\hypertarget{ref-opentreeoflife2019synth}{}%
@@
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
-
-\leavevmode\hypertarget{ref-scott2011ncbi}{}%
-Scott, F. (2011). The ncbi taxonomy database. \emph{Nucleic Acids Research}, \textbf{40}, D136--D14.
 

### 2020-07-10 1-abstract.Rmd
Kinsey intro comments finished yay :tada:

@@
 1.
-We hope that the Physcraper workflow demonstrates the benefits of open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS.

### 2020-07-10 2-body.Rmd
Kinsey intro comments finished yay :tada:

@@
 Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
+WHY IS IT A CHALLENGE
 
 We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
-By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees for of loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees using loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
 The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
@@
 
-Physcraper leverages phylogentic data stored in Open Tree of Life and in Treebase
+Physcraper leverages phylogentic data stored in Open Tree of Life and in TreeBase.
 The Open Tree of Life (https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
@@
 
-ELABORATE MORE ON THIS, USE LESS CASUAL LANGUAGE: This is an effort to keep on directing ourselves towards a fully reproducible workflow in phylogenetics.
-And an effort to more effectively make big data connections for the advantage of phylogenetics and biology in general.
+We envision Physraper as an staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
 

### 2020-07-10 docs/submission-mee-2020-07-10.tex
Kinsey intro comments finished yay :tada:

@@
 
-\textbf{Word count}: 3702
+\textbf{Word count}: 3694
 
@@
 \item
-  We hope that the Physcraper workflow demonstrates the benefits of open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+  We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS.
@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate DO WE NEED CITATIONS FOR THIS?.
+Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the National Center for Biodiversity Information (NCBI) molecular database, release number 159 (from April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-
-Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogeentically informative sequence data available in GenBank more than a decade ago.
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
@@
 
-Physcraper leverages phylogentic data stored in Open Tree of Life and in Treebase
+Physcraper leverages phylogentic data stored in Open Tree of Life and in TreeBase.
 The Open Tree of Life (\url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.

### 2020-07-10 2-body.Rmd
typo

@@
 
-We envision Physraper as an staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
+We envision Physraper as a staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
 

### 2020-07-14 2-body.Rmd
figure 1

@@
 
-The general workflow is described in FIGURE. Next, we will describe the technical details of each step.
+```{r framework, echo=FALSE, fig.cap="Physcraper general framework.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
+# knitr::include_graphics("docs/figs/schematic.svg")
+```
+
+The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation, tutorials, examples and function usage.
+The general Physcraper workflow is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
 
@@
 ``` -->
-```{r fig-2, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r malvaceae, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 knitr::include_graphics("docs/figs/cotree-plot2-1.png")

### 2020-07-14 docs/submission-mee-2020-07-13.tex
figure 1

--- 2020-07-10 docs/submission-mee-2020-07-10.tex
@@
 
-\textbf{Word count}: 3694
+\textbf{Word count}: 3722
 
@@
 
-The general workflow is described in FIGURE. Next, we will describe the technical details of each step.
+\begin{figure}
+
+{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic}
+
+}
+
+\caption{Physcraper general framework.}(\#fig:framework)
+\end{figure}
+
+The Physcraper software is fully described on its documentation website at \url{https://physcraper.readthedocs.io/en/latest/}, along with installation, tutorials, examples and function usage.
+The general Physcraper workflow is shown in Figure @ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
 
@@
 
-\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}\label{fig:fig-2}
+\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}(\#fig:malvaceae)
 \end{figure}

### 2020-07-14 render2-mee.tex
figure 1

--- 2020-07-14 docs/submission-mee-2020-07-13.tex
@@
 
-\caption{Physcraper general framework.}(\#fig:framework)
+\caption{Physcraper general framework.}\label{fig:framework}
 \end{figure}
@@
 The Physcraper software is fully described on its documentation website at \url{https://physcraper.readthedocs.io/en/latest/}, along with installation, tutorials, examples and function usage.
-The general Physcraper workflow is shown in Figure @ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
+The general Physcraper workflow is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
 
@@
 
-\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}(\#fig:malvaceae)
+\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}\label{fig:malvaceae}
 \end{figure}

### 2020-07-15 2-body.Rmd
intro closing

@@
 
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers.
- main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
-EXPLIAN WHY PYTHON IS BEST FOR THIS.
-By design, this approach focusses on interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
+Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers, automatizing taxonomic name matching across biological data stores.
+It's main goal is to update previously known phylogenetic relationships in a continuous manner by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank.
+EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
+By design, this approach focuses on data interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
 For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
 
-We envision Physraper as a staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
+We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
 

### 2020-07-21 2-body.Rmd
documentation

@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological data stores constitute an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
 Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-Some of the discrepancy between data availability and phylogenetic estimates can be explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
+The discrepancy between data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 
-The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation, tutorials, examples and function usage.
+The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.
 The general Physcraper workflow is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.

### 2020-07-22 2-body.Rmd
rob's comments

@@
 
-Physcraper leverages phylogentic data stored in Open Tree of Life and in TreeBase.
+Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
 The Open Tree of Life (https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
@@
 
+ARGUMENT - GENES ARE STILL USEFUL IN THE GENOMICS ERA, AND HERE'S WHY:
+GENOMIC MARKERS SUCH AS RADSEQ, SNP, MISCROSATS AND UCES - ARE HOMOLOGY HYPOTHESIS THE SAME ON THESE MARKERS THAN FOR PROTEIN CODING AND NON CODING LOCI?
+WHAT ARE THE PROS AND CONS OF USING GENE ALIGNMENTS ONLY AND NOT GENOMIC MARKERS?
 While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
 One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq and UCE's.
+Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
 This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016misconceptions]
+Even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016misconceptions]
 In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
@@
 In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward].
-Physcraper shares a similar conceptual framework to Pumper [@izquierdo-carrasco_pumper:_2014], but that software is not currently supported or developed (*or runnable at all honestly...*)
+Physcraper shares a similar conceptual framework to Pumper [@izquierdo2014pumper], but that software is not currently supported or developed (*or runnable at all honestly...*)
 
@@
 By design, this approach focuses on data interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
+For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and the phylogenetic data store.
 
@@
 - Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the `inputs` folder to be used in the following steps.
+  Technically, just one matching name is needed to perform the searches. See below.
+- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the "inputs" folder to be used in the following steps.
 
@@
 
-- Technically, any DNA molecular database can be used to search for new sequences.
-  By default we rely on the GenBank database.
-  The new sequence search can be performed on the remote database or in a local database.
+- Physcraper's code design relies on the GenBank DNA database to search for new sequences.
+The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
+Using the latter speeds up the process.
+Detailed nstructions to setup a local database are provided on the software documenttaion.
+<!-- Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
+The code to do this with Physcraper is not yet implemented.
+-->
+
 - The next step is to identify the search taxon within the reference taxonomy.
@@
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
+  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
   Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
@@
 
+COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING
+WHY WE DID NOT MAKE A BENCHMARK COMPARISON
+
 Many pipelines are making use of DNA data repositories in different ways.
@@
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+
 While not generating full phylogenies *de novo*, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
+It can also be combined with data from repositories other than molecular data.
+For example geographic locations (GBIF), fossils (PBDB), etc.
+*from Robert: I think you can sell the program more here.
+Why is it better than the other methods?
+You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages*
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly
+Physcraper has the potential to be applied for the advantage of the field to rapidly HOW FAST IS "RAPID"
 place newly discovered species phylogenetically [@webb2010biodiversity],

### 2020-07-22 2-body.Rmd
kb's comments

@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- Reverse complement sequences are identified and translated.
+- Reverse complement sequences are identified and translated. HOW IS THIS DONE.
 - Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
@@
 
-## DNA sequence alignment
+## New DNA sequence alignment
 
-- The software MUSCLE [@edgar2004muscle] is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
-- First, all new sequences are aligned using default MUSCLE options.
-- Then, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align new sequences. This ensures that the final alignment
+- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform the sequence alignments.
+Instructions on how to install this nad other software dependencies used by Physcraper is provided in the documentation.
+- The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
+- Second, a MUSCLE profile alignment is performed, in which the original alignment
+  is used as a template to align the new sequences. This ensures that the final alignment
   follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. So, we encourage users
+- The final alignment is not further processed automatically. We encourage users
   to check it either by eye and perform manual refinement or using any of the many
   tools for alignment processing, to eliminate columns with no information.
+- Manual curation is a critical step in te process, it is not reproducible, bc the process is nottracked.
+A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
+Versioning alignments and adding them to a  global database is a next step for us!
 - Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.

### 2020-07-22 2-body.Rmd
laura's comments

@@
 
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers, automatizing taxonomic name matching across biological data stores.
-It's main goal is to update previously known phylogenetic relationships in a continuous manner by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank.
+Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
 EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
-By design, this approach focuses on data interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and the phylogenetic data store.
+By design, this approach focuses on data interoperability.
+By automating taxonomic name matching across NCBI, OpenTree, GBIF and virtually any biological database, users can perform downstream analyses straightforwardly.
+For example, it automatizes and standardizes comparison of phylogenetic hypotheses with currently known relationships from the synthetic Open Tree of Life, the Open Tree of Life taxonomy tree and any phylogeny that is stored in the
+Phylesystem.
 
-We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
+We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
 
-# How does Physcraper work?
+# The Physcraper framework
 
-```{r framework, echo=FALSE, fig.cap="Physcraper general framework.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
@@
 
-The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.
-The general Physcraper workflow is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
+The general Physcraper framework is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
 
-## The input: a tree and an alignment
+## The inputs: a tree and an alignment
 
@@
   alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal were the tree was originally published.
+  DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal where the tree was originally published.
   If the alignment is stored in TreeBase, Physcraper can download it directly,
@@
   of the tree are in the DNA character matrix and vice versa.
-- A ".csv" file with the summary of taxon name matching is produced for the user.
+- A "csv" file with the summary of taxon name matching is produced for the user.
 - Unmatched taxon names are dropped from both the tree and alignment.
   Technically, just one matching name is needed to perform the searches. See below.
-- A ".tre" file and a ".fas" file containing only the matched taxa are generated and saved in the "inputs" folder to be used in the following steps.
+- A "newick" file and a "fasta" file containing only the matched taxa are generated and saved in the "inputs" folder to be used in the following steps.
 
@@
   matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
@@
   The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython [@cock2009biopython] BLAST function in the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternate BLAST URL.
+  For remote-database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
 This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
@@
 - Reverse complement sequences are identified and translated. HOW IS THIS DONE.
-- Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
-- Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to "physcraper/taxonomy" folder), so they are accessible for further runs.
-- A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+- Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
+- To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
+- A "fasta" file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 
-- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform the sequence alignments.
-Instructions on how to install this nad other software dependencies used by Physcraper is provided in the documentation.
+- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
+Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
 - The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
@@
   follows the homology criteria established by the original alignment.
-- The final alignment is not further processed automatically. We encourage users
-  to check it either by eye and perform manual refinement or using any of the many
-  tools for alignment processing, to eliminate columns with no information.
-- Manual curation is a critical step in te process, it is not reproducible, bc the process is nottracked.
+- The final alignment is not further processed by Physcraper.
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
+- While curating the alignment is a critical step, it is not a reproducible one.
+The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
 A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
 Versioning alignments and adding them to a  global database is a next step for us!
-- Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
+- Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
 
@@
 
-- A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML [@stamatakis2014raxml], using default settings such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with default method.
-Currently only the number of bootsrap replicates can be modified by the user.
-- The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
+- A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml] with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
+Currently only the number of bootsrap replicates can be specified by the user.
+- By default, the original tree is used as a starting tree for the ML searches.
+Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
 - Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
-- The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
-- Tips on all trees generated by Physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
-- Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions [@sukumaran2010dendropy].
-- Finally a conflict analysis is performed.
-This is basically a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
+- Physcraper's final result is an updated phylogenetic hypothesis for each of the genes provided in the input alignment.
+- Tips on all trees generated by Physcraper are defined by a taxon "name space", which captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
+- Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
+- Second, a conflict analysis is performed.
+This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
 This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
@@
 However, the head of the research group wants to use a particular marker they believe to be the best one to be able to solve the relationships in the group.
-They have been working in the alignment for long and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
+They have been working on the alignment for a long time and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
 
@@
 Data repositories hold even more information than meets the eye.
-Besides the actual data, they are rich sources of metadata that can be used for the advantage of biology and science in general.
+Besides the actual data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
 
@@
 Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have been already studied and that have phylogenetic data already.
+regions that have already been studied and which have phylogenetic data.
 However, expert phylogenetic knowledge is also an important source of data in public
@@
 a field that has relied on processes which are somewhat artisanal.
-In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
 community to build up our phylogenetic knowledge, we can use not only data accumulated in

### 2020-07-22 2-body.Rmd
all comments added

@@
 ```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
-knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-# knitr::include_graphics("docs/figs/schematic.svg")
+# knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
+knitr::include_graphics("docs/figs/schematic.svg")
 ```

### 2020-07-22 docs/submission-mee-2020-07-22.tex
all comments added

--- 2020-07-14 render2-mee.tex
@@
 
-\textbf{Word count}: 3722
+\textbf{Word count}: 3966
 
@@
   We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS.
+Physcraper can be used with any OS and is released under an open source license.
 Detailed instructions for installation and
@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological data stores constitute an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-Some of the discrepancy between data availability and phylogenetic estimates can be explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
+The discrepancy between data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
@@
 
-Physcraper leverages phylogentic data stored in Open Tree of Life and in TreeBase.
+Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
 The Open Tree of Life (\url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
@@
 
+ARGUMENT - GENES ARE STILL USEFUL IN THE GENOMICS ERA, AND HERE'S WHY:
+GENOMIC MARKERS SUCH AS RADSEQ, SNP, MISCROSATS AND UCES - ARE HOMOLOGY HYPOTHESIS THE SAME ON THESE MARKERS THAN FOR PROTEIN CODING AND NON CODING LOCI?
+WHAT ARE THE PROS AND CONS OF USING GENE ALIGNMENTS ONLY AND NOT GENOMIC MARKERS?
 While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
 One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq and UCE's.
+Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
 This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016})
+Even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016})
 In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
@@
 In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-Physcraper shares a similar conceptual framework to Pumper (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo-carrasco_pumper:_2014}{2014}), but that software is not currently supported or developed (\emph{or runnable at all honestly\ldots{}})
+Physcraper shares a similar conceptual framework to Pumper (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}), but that software is not currently supported or developed (\emph{or runnable at all honestly\ldots{}})
 
@@
 
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers.
-main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
-EXPLIAN WHY PYTHON IS BEST FOR THIS.
-By design, this approach focusses on interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
+Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
+EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
+By design, this approach focuses on data interoperability.
+By automating taxonomic name matching across NCBI, OpenTree, GBIF and virtually any biological database, users can perform downstream analyses straightforwardly.
+For example, it automatizes and standardizes comparison of phylogenetic hypotheses with currently known relationships from the synthetic Open Tree of Life, the Open Tree of Life taxonomy tree and any phylogeny that is stored in the
+Phylesystem.
 
-We envision Physraper as a staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
+We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
 
-\hypertarget{how-does-physcraper-work}{%
-\section{How does Physcraper work?}\label{how-does-physcraper-work}}
+\hypertarget{the-physcraper-framework}{%
+\section{The Physcraper framework}\label{the-physcraper-framework}}
 
@@
 
-\caption{Physcraper general framework.}\label{fig:framework}
+\caption{The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
 \end{figure}
 
-The Physcraper software is fully described on its documentation website at \url{https://physcraper.readthedocs.io/en/latest/}, along with installation, tutorials, examples and function usage.
-The general Physcraper workflow is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
+The general Physcraper framework is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
 
-\hypertarget{the-input-a-tree-and-an-alignment}{%
-\subsection{The input: a tree and an alignment}\label{the-input-a-tree-and-an-alignment}}
+\hypertarget{the-inputs-a-tree-and-an-alignment}{%
+\subsection{The inputs: a tree and an alignment}\label{the-inputs-a-tree-and-an-alignment}}
 
@@
   alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
-  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal were the tree was originally published.
+  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal where the tree was originally published.
   If the alignment is stored in TreeBase, Physcraper can download it directly,
@@
 \item
-  A ``.csv'' file with the summary of taxon name matching is produced for the user.
+  A ``csv'' file with the summary of taxon name matching is produced for the user.
 \item
   Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
+  Technically, just one matching name is needed to perform the searches. See below.
 \item
-  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
+  A ``newick'' file and a ``fasta'' file containing only the matched taxa are generated and saved in the ``inputs'' folder to be used in the following steps.
 \end{itemize}
@@
 \item
-  Technically, any DNA molecular database can be used to search for new sequences.
-  By default we rely on the GenBank database.
-  The new sequence search can be performed on the remote database or in a local database.
+  Physcraper's code design relies on the GenBank DNA database to search for new sequences.
+The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
+Using the latter speeds up the process.
+Detailed nstructions to setup a local database are provided on the software documenttaion.
 \item
@@
 If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
+  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
@@
   The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function in the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternate BLAST URL.
+  For remote-database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
 This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
@@
 \item
-  Reverse complement sequences are identified and translated.
+  Reverse complement sequences are identified and translated. HOW IS THIS DONE.
 \item
-  Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
+  Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
 \item
-  Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to ``physcraper/taxonomy'' folder), so they are accessible for further runs.
+  To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to ``physcraper/taxonomy'' folder) that is globally accessible to users.
 \item
-  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+  A ``fasta'' file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 \end{itemize}
 
-\hypertarget{dna-sequence-alignment}{%
-\subsection{DNA sequence alignment}\label{dna-sequence-alignment}}
+\hypertarget{new-dna-sequence-alignment}{%
+\subsection{New DNA sequence alignment}\label{new-dna-sequence-alignment}}
 
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
+  Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
+Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
 \item
-  First, all new sequences are aligned using default MUSCLE options.
+  The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
 \item
-  Then, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align new sequences. This ensures that the final alignment
+  Second, a MUSCLE profile alignment is performed, in which the original alignment
+  is used as a template to align the new sequences. This ensures that the final alignment
   follows the homology criteria established by the original alignment.
 \item
-  The final alignment is not further processed automatically. So, we encourage users
-  to check it either by eye and perform manual refinement or using any of the many
-  tools for alignment processing, to eliminate columns with no information.
+  The final alignment is not further processed by Physcraper.
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
 \item
-  Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
+  While curating the alignment is a critical step, it is not a reproducible one.
+The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
+A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
+Versioning alignments and adding them to a global database is a next step for us!
+\item
+  Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
 \end{itemize}
@@
 \item
-  A gene tree is reconstructed for each alignment provided, using a Maximum Likelihood approach implemented with the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}), using default settings such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with default method.
-Currently only the number of bootsrap replicates can be modified by the user.
+  A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
+Currently only the number of bootsrap replicates can be specified by the user.
 \item
-  The original tree is used as starting tree for the ML searches. It can also be set as a full topological constraint or not used at all, depending on the goals of the user.
+  By default, the original tree is used as a starting tree for the ML searches.
+Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
 \item
@@
 \item
-  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
+  Physcraper's final result is an updated phylogenetic hypothesis for each of the genes provided in the input alignment.
 \item
-  Tips on all trees generated by Physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
+  Tips on all trees generated by Physcraper are defined by a taxon ``name space'', which captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
 \item
-  Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+  Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 \item
-  Finally a conflict analysis is performed.
-This is basically a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
+  Second, a conflict analysis is performed.
+This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
 This is performed with OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
@@
 Data repositories hold even more information than meets the eye.
-Besides the actual data, they are rich sources of metadata that can be used for the advantage of biology and science in general.
+Besides the actual data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+
+COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING
+WHY WE DID NOT MAKE A BENCHMARK COMPARISON
 
@@
 a field that has relied on processes which are somewhat artisanal.
-In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
 community to build up our phylogenetic knowledge, we can use not only data accumulated in
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+
 While not generating full phylogenies \emph{de novo}, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
+It can also be combined with data from repositories other than molecular data.
+For example geographic locations (GBIF), fossils (PBDB), etc.
+\emph{from Robert: I think you can sell the program more here.
+Why is it better than the other methods?
+You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages}
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly
+Physcraper has the potential to be applied for the advantage of the field to rapidly HOW FAST IS ``RAPID''
 place newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
@@
 
-\leavevmode\hypertarget{ref-izquierdo-carrasco_pumper:_2014}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A.
-\& Stamatakis, A.
-(2014).
-PUmPER: Phylogenies Updated Perpetually.
-\emph{Bioinformatics}, \textbf{30}, 1476--7.
-Retrieved February 1, 2014, from \url{http://bioinformatics.oxfordjournals.org/content/early/2014/01/28/bioinformatics.btu053}
+\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
+Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
 

### 2020-07-23 2-body.Rmd
some edits

@@
 Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
-The Open Tree of Life (https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
-Open Tree of Life aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently the tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
+The Open Tree of Life (OpenTree from now on https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
+OpenTree aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+This "synthetic" tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
 
-The Open Tree of Life data store, [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373), contains more than 4,500 phylogenetic trees from published studies.
-The tips in these trees are mapped a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
+The Open Tree of Life data store, the [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373), contains more than 4,500 phylogenetic trees from published studies.
+The tips in these trees are mapped to a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
 This provides a resource for finding existing estimates of phylogenetic relationships,
-and assessing which regions of the tree of life which are lacking available phylogenetic estimates.
+and assessing which regions of the tree of life are lacking available phylogenetic estimates.
 
 By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
-the TreeBASE repository [@piel2009treebase] and the Open Tree of Life (OpenTree) datastore [@mctavish2015phylesystem], we can place new biological data in an evolutionary context.
+the TreeBASE repository [@piel2009treebase] and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
 

### 2020-07-24 1-abstract.Rmd
license
mk's comments

@@
 1. Phylogenies are a key part of research in many areas of biology. Tools that automatize
-some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
+some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
 have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
@@
 We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS.
+Physcraper can be used with any OS and is released under an open source license.
 Detailed instructions for installation and

### 2020-07-24 2-alt-1-intro.Rmd
mk's comments

--- 2020-07-23 2-body.Rmd
@@
----
-title: 'Physcraper: Body'
-author:
-  - Luna L. Sanchez Reyes
-  - Martha Kandziora
-  - Emily Jane McTavish
-  - University of California, Merced
-date: '`r format(Sys.time(), "%d %B, %Y")`'
-bibliography: paper.bib
-csl: apa.csl
-output:
-  redoc::redoc
----
+Phylogenetic estimates of evolutionary relationships capture the shared history of
+living organisms, and provide key context for all our biological observations.
+Public biological databases constitute an amazing resource for evolutionary estimation,
+but a large portion of publicly available molecular data has never been incorporated into any public phylogenetic estimate.
 
-# Introduction
+Connecting new sequence data with existing estimates of evolutionary relationships is a challenge
+largely because taxonomy across databases is not consistent.
 
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological data stores constitute an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
-Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
+Currently, estimates of phylogenetic relationships are publicly available for about 100 000 taxa only [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-The discrepancy between data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
+The discrepancy between molecular data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
-WHY IS IT A CHALLENGE
 
@@
 WHAT ARE THE PROS AND CONS OF USING GENE ALIGNMENTS ONLY AND NOT GENOMIC MARKERS?
+Martha: Genomic data is not available for a large number of taxa.
 While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
@@
 Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
+OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., MORE EXAMPLES OF DATABASES providing a way to automatically link data from different databases.
 
 Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
-EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
+<!-- WE COULD EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
+Maybe not relevant for now, so we'll keep it out-->
 By design, this approach focuses on data interoperability.
@@
 We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
-
-# The Physcraper framework
-
-```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
-# knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-knitr::include_graphics("docs/figs/schematic.svg")
-```
-
-The general Physcraper framework is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
-
-## The inputs: a tree and an alignment
-
-- In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem), or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (https://tree.opentreeoflife.org/curator/tnrs/).
-  The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of unpublished trees).
-  A user can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
-If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
-  If the user is not ready to make the input tree public, `physcraper` can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
-- The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
-  alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper can download it directly,
-  either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
-  or through the TreeBASE GitHub repository (SuperTreeBASE; [github.com/TreeBASE/supertreebase](https://github.com/TreeBASE/supertreebase)).
-  If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-- A "csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. See below.
-- A "newick" file and a "fasta" file containing only the matched taxa are generated and saved in the "inputs" folder to be used in the following steps.
-
-## DNA sequence search and filtering
-
-- Physcraper's code design relies on the GenBank DNA database to search for new sequences.
-The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
-Using the latter speeds up the process.
-Detailed nstructions to setup a local database are provided on the software documenttaion.
-<!-- Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
-The code to do this with Physcraper is not yet implemented.
--->
-
-- The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be determined by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool [Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
-
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the sequences on the checked alignment.
-  The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
-  <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
-These parameters can be configured for each run.
-  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
-  All accepted sequences are assigned an internal identifier, and are further filtered.
-
-- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
-Longer sequences belonging to the
-  same taxon as the original sequence will be considered further for analysis.
-- Among the filtered sequences, there are often several representatives per taxon.
-  Although it can be useful to keep some of them, for example, to investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- Reverse complement sequences are identified and translated. HOW IS THIS DONE.
-- Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
-- To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
-- A "fasta" file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
-
-## New DNA sequence alignment
-
-- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
-Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
-- The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
-- Second, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align the new sequences. This ensures that the final alignment
-  follows the homology criteria established by the original alignment.
-- The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
-- While curating the alignment is a critical step, it is not a reproducible one.
-The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
-Versioning alignments and adding them to a  global database is a next step for us!
-- Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
-
-## Tree reconstruction and comparison
-
-- A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml] with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
-Currently only the number of bootsrap replicates can be specified by the user.
-- By default, the original tree is used as a starting tree for the ML searches.
-Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
-- Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
-- Physcraper's final result is an updated phylogenetic hypothesis for each of the genes provided in the input alignment.
-- Tips on all trees generated by Physcraper are defined by a taxon "name space", which captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
-- Two ways to compare the updated tree with the original tree are implemented in Physcraper.
-First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
-- Second, a conflict analysis is performed.
-This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
-This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
-- For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-- A suggested default rooting based on OpenTree's taxonomy is implemented for now.
-This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
-However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate.
-Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroups so trees are accurately rooted.
-
-# Examples
-
-We will illustrate the utility of Physcraper in here with two use-case scenarios.
-One in which the user is interested in a particular group.
-Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available elsewhere.
-
-## The hollies
-
-A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
-The first one has been made available in TreeBASE as well as in the OpenTree phylesystem and is part of the synthetic tree. It samples 48 species.
-The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-The most recent one has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet, with 200 species.
-However, it has not been added to the syntehtic tree yet.
-This also makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
-
-DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OpenTree
-RF DISTANCE INTERPRETATION
-HOW MUCH TIME THE BLAST RUN TOOK
-ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE
-
-FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OpenTree.
-
-## The Malvaceae
-
-A postdoc started working with a new reserach group.
-They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species, containing the relatives of cacao, cotton, durian and okra.
-A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
-However, the head of the research group wants to use a particular marker they believe to be the best one to be able to solve the relationships in the group.
-They have been working on the alignment for a long time and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
-
-<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
-wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/data-raw/malvaceae-images/cotree-plot2-1.png"
-mv cotree-plot2-1.png docs/figs/cotree-plot2-1.png
-``` -->
-```{r malvaceae, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
-knitr::include_graphics("docs/figs/cotree-plot2-1.png")
-```
-\newpage
-
-# Discussion
-
-Data repositories hold even more information than meets the eye.
-Besides the actual data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-
-COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING
-WHY WE DID NOT MAKE A BENCHMARK COMPARISON
-
-Many pipelines are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data -- getting real and clean homologs.
-Others focus on refining the alignment.
-Most focus on generating full trees *de novo*, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have already been studied and which have phylogenetic data.
-However, expert phylogenetic knowledge is also an important source of data in public
-and open repositories that is not being used to its full potential.
-
-All these tools are key efforts for advancing towards reproducibility in phylogenetics,
-a field that has relied on processes which are somewhat artisanal.
-Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
-community to build up our phylogenetic knowledge, we can use not only data accumulated in
-molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-
-While not generating full phylogenies *de novo*, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data.
-For example geographic locations (GBIF), fossils (PBDB), etc.
-*from Robert: I think you can sell the program more here.
-Why is it better than the other methods?
-You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages*
-
-Physcraper has the potential to be applied for the advantage of the field to rapidly HOW FAST IS "RAPID"
-place newly discovered species phylogenetically [@webb2010biodiversity],
-obtain trees for ecophylogenetic studies [@helmus2012phylogenetic],
-help to systematize molecular databases, i.e., curate taxonomic assignations [@san2010molecular],
-and rapidly generate custom species trees for downstream analyses [@stoltzfus2013phylotastic].
-
-# Acknowledgements
-
-Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838, and ABI No. 1759846.
-Compute time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
-ACI-1429783.
-
-We thank members of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
-
-# Authors' Contributions
-
-# Data Avilability
-
-\newpage
-# References

### 2020-07-24 2-body.Rmd
mk's comments

@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological data stores constitute an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
-Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
+Currently, estimates of phylogenetic relationships are publicly available for about 100 000 taxa only [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-The discrepancy between data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
+The discrepancy between molecular data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
-WHY IS IT A CHALLENGE
 
@@
   A user can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+WE SHOYLD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE???
 If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- A taxon name matching step is performed to verify that all taxon names on the tips
+- EXPLAIN CONCEPT OF UNIFIED TAXONOMY HERE A taxon name matching step is performed to verify that all taxon names on the tips
   of the tree are in the DNA character matrix and vice versa.
 - A "csv" file with the summary of taxon name matching is produced for the user.
-- Unmatched taxon names are dropped from both the tree and alignment.
+- EXPLAIN THIS BETTER, IT IS NOT THAT MUCH A DROP UNMATCHED AS DROP ABSENT!
+Unmatched taxon names are dropped from both the tree and alignment.
+Technically, all taxa in alignment should be in NCBI taxonomy and hence in the OpenTree taxonomy.
+Single locus alignments might contain less taxa than the tree constructed using a concatenated matrix.
   Technically, just one matching name is needed to perform the searches. See below.
-- A "newick" file and a "fasta" file containing only the matched taxa are generated and saved in the "inputs" folder to be used in the following steps.
+- A "newick" file and a "fasta" file containing the tree and alignment respectively with matched taxa only are generated and saved in the "inputs" folder to be used in the following steps.
 
@@
 
-- Physcraper's code design relies on the GenBank DNA database to search for new sequences.
+- Physcraper's uses the GenBank DNA database as source to search for new sequences.
 The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
 Using the latter speeds up the process.
-Detailed nstructions to setup a local database are provided on the software documenttaion.
+Detailed instructions to setup a local database are provided on the software documentation.
 <!-- Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
@@
 
-- The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be determined by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
+- The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
+The search taxon can be chosen by the user from the NCBI taxonomy. If none is provided, then the search taxon is automatically set as the Most Recent Common Ancestor (MRCA) of the
+  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
+  <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
+  This is known as the Most Recent Common
   Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool [Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
+  To automatically identify the MRCAT of a group of taxon names, we use the OpenTree taxonomic tool v3 [[Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>); @rees2017automated].
 
@@
   database, and the sequences on the checked alignment.
-  The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
+  The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local database sequence searches.
+  For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required wait times, which slow down the searches markedly.
+  *MK suggest explaining here what the nbi waiting times are, but I'm not sure it's needed*
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
@@
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
+  taxon. Results from each one of these BLAST runs are written down, and matched sequences are saved
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers.
 This information will be used later to store the whole sequences in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
@@
 
-- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
+- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different NCBI *or otol?* taxon than the existing sequence.
 Longer sequences belonging to the
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroups so trees are accurately rooted.
+The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
+*It would be a nice addiion to have users give the output of the inout tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon*
 
@@
 A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
-The first one has been made available in TreeBASE as well as in the OpenTree phylesystem and is part of the synthetic tree. It samples 48 species.
-The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-The most recent one has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet, with 200 species.
-However, it has not been added to the syntehtic tree yet.
-This also makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
+The genus encompasses between 400-600 living species.
+A review of literature (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
+@gottlieb2005molecular made tree and alignment data available in TreeBASE. The tree sampling 48 species was integrated to the OpenTree Phylesystem and is part of OpenTree's synthetic tree.
+<!-- The second tree has not been made available anywhere, not even in the supplementary data of the original publication. -->
+The most recent *Ilex* tree from @yao2020phylogeny has been made available in the OpenTree Phylesystem and in the DRYAD repository.
+It is the best sampled yet for the genus, with 200 species.
+However, it has not been added to OpenTree's synthtetic tree yet.
+This makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
 
@@
 While not generating full phylogenies *de novo*, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data.
-For example geographic locations (GBIF), fossils (PBDB), etc.
+Moreover, it can combine phylogenies with data from repositories other than molecular data.
+For example geographic locations (using GBIF ids), fossils (using PBDB ids), etc.
 *from Robert: I think you can sell the program more here.
@@
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly HOW FAST IS "RAPID"
+Physcraper has the potential to be applied for the advantage of the field to rapidly *HOW FAST IS "RAPID" mention it in results and then here again*
 place newly discovered species phylogenetically [@webb2010biodiversity],
@@
 
+EJM: Conceived study, wrote most of the code, documentation and tests.
+MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
+LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
+All authors contributed to the manuscript.
+
 # Data Avilability

### 2020-07-24 2.1-intro-ideas.Rmd
some edits
mk's comments

@@
+TELL THEM WHAT YOU ARE GONNA TELL THEM
+TELL TEHM
+THEN TELL THEM WHAT YOU JUST TOLD THEM
+
 over 217 million DNA sequences [(release number 238; June 15, 2020)](https://ftp.ncbi.nih.gov/genbank/gbrel.txt)
+Data of number of nodes in the NCBI taxonomy can be found in their [homepage](https://www.ncbi.nlm.nih.gov/Taxonomy/taxonomyhome.html/index.cgi?chapter=statistics&?&uncultured=hide&unspecified=hide&m=0)
+In 2020 it has 668 929 taxa below spp, 474 760 species.
 

### 2020-07-24 2-body.Rmd
edit methods

@@
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-knitr::include_graphics("docs/figs/schematic.svg")
+knitr::include_graphics("docs/figs/schematic.pdf")
 ```
@@
 
-- In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem [github.com/opentreeoflife/phylesystem](https://github.com/opentreeoflife/phylesystem), or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (https://tree.opentreeoflife.org/curator/tnrs/).
-  The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of unpublished trees).
-  A user can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
-WE SHOYLD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE???
-If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool ([www.opentreeoflife.org/curator](https://tree.opentreeoflife.org/curator).
-  If the user is not ready to make the input tree public, `physcraper` can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
-- The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
-  alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
+- In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+  OpenTree [Phylesystem] (https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator).
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
+Ix taxonomic names can't be standardized, they will be excluded for further analysis.
+  Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
+  that are essential for automatizing the phylogeny updating process.
+The most relevant of these is the standardization of taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+Currently, only trees connected to a published study can be stored in the Phylesystem.
+  <!-- Discussion: (although there are plans to allow storage of unpublished trees). -->
+  Users can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+*MK suggests that WE SHOULD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE, I think it's a good idea*
+- The input alignment should be a single locus alignment that was used to generate the tree. Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
   DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper can download it directly,
+  If the alignment is stored in TreeBase, Physcraper downloads it directly,
   either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- EXPLAIN CONCEPT OF UNIFIED TAXONOMY HERE A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
-- A "csv" file with the summary of taxon name matching is produced for the user.
-- EXPLAIN THIS BETTER, IT IS NOT THAT MUCH A DROP UNMATCHED AS DROP ABSENT!
-Unmatched taxon names are dropped from both the tree and alignment.
-Technically, all taxa in alignment should be in NCBI taxonomy and hence in the OpenTree taxonomy.
-Single locus alignments might contain less taxa than the tree constructed using a concatenated matrix.
-  Technically, just one matching name is needed to perform the searches. See below.
+- When dealing with single locus alignments, it is common for alignments to have less taxa than the tree, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
+Hence, a pruning step to reconcile taxon presence in the tree and alignment is performed.
+This verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name (and its coresonding sequence in the alignment) is needed to continue the algorithm.
+See next section.
+*MK mentioned that nothing shoul be dropped here bc the name standardization should have matched everything, so I explained a bit more what this step is.
+It is not an unmatched taxa pruning but more of an absent taxa pruning.
+Also, taxa that are not matched should be dropped before right?*
+- A "csv" file with the summary of taxon name standardization and pruning is produced for the user.
 - A "newick" file and a "fasta" file containing the tree and alignment respectively with matched taxa only are generated and saved in the "inputs" folder to be used in the following steps.
@@
 
-- Physcraper's uses the GenBank DNA database as source to search for new sequences.
+- Physcraper uses the GenBank DNA database as source to search for new sequences.
 The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
@@
   <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
-  This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
+  This is known in the OpenTree as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) *I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only*.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we use the OpenTree taxonomic tool v3 [[Taxonomy-API-v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>); @rees2017automated].
+  To identify the MRCAT of a group of taxon names, we use the OpenTree [taxonomic tool v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
 
@@
 This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required wait times, which slow down the searches markedly.
-  *MK suggest explaining here what the nbi waiting times are, but I'm not sure it's needed*
+  *MK suggest explaining here what the ncbi waiting times are, but I don;t think it's needed, what do you think?*
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
@@
 
-- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different NCBI *or otol?* taxon than the existing sequence.
+- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon than the existing sequence *MK mentioned it was unclear to her if this was an NCBI or OTT taxon.
+I think it is the unified taxonomy, hence an OTT taxon?
+I think it doesn;t harm to be super explicit about it*.
 Longer sequences belonging to the
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- Reverse complement sequences are identified and translated. HOW IS THIS DONE.
+- Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions.
 - Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
@@
 - To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
-- A "fasta" file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+- A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
-*It would be a nice addiion to have users give the output of the inout tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon*
+*It would be a nice addiion to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon*
 

### 2020-07-24 docs/submission-mee-2020-07-24.tex
edit methods

--- 2020-07-22 docs/submission-mee-2020-07-22.tex
@@
 
-\textbf{Word count}: 3966
+\textbf{Word count}: 4198
 
@@
   Phylogenies are a key part of research in many areas of biology. Tools that automatize
-  some parts of the process of phylogenetic reconstruction (mainly character matrix construction)
+  some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
   have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological data stores constitute an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
+GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+Currently, estimates of phylogenetic relationships are publicly available for about 100 000 taxa only (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
 
-The discrepancy between data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
+The discrepancy between molecular data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
 However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
-WHY IS IT A CHALLENGE
 
@@
 Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
-The Open Tree of Life (\url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
-Open Tree of Life aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently the tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
+The Open Tree of Life (OpenTree from now on \url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
+OpenTree aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
 
-The Open Tree of Life data store, \href{https://academic.oup.com/bioinformatics/article/31/17/2794/183373}{Phylesystem}, contains more than 4,500 phylogenetic trees from published studies.
-The tips in these trees are mapped a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
+The Open Tree of Life data store, the \href{https://academic.oup.com/bioinformatics/article/31/17/2794/183373}{Phylesystem}, contains more than 4,500 phylogenetic trees from published studies.
+The tips in these trees are mapped to a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
 This provides a resource for finding existing estimates of phylogenetic relationships,
-and assessing which regions of the tree of life which are lacking available phylogenetic estimates.
+and assessing which regions of the tree of life are lacking available phylogenetic estimates.
 
 By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
-the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OpenTree) datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
+the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
 
@@
 Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
+OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., MORE EXAMPLES OF DATABASES providing a way to automatically link data from different databases.
 
 Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
-EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
 By design, this approach focuses on data interoperability.
@@
 \item
-  In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}, or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (\url{https://tree.opentreeoflife.org/curator/tnrs/}).
-  The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of unpublished trees).
-  A user can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OpenTree website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OpenTree's website about}).
-If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
-  If the user is not ready to make the input tree public, \texttt{physcraper} can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
+  In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+  OpenTree {[}Phylesystem{]} (\url{https://github.com/opentreeoflife/phylesystem}), or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application}.
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+Ix taxonomic names can't be standardized, they will be excluded for further analysis.
+  Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
+  that are essential for automatizing the phylogeny updating process.
+The most relevant of these is the standardization of taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+Currently, only trees connected to a published study can be stored in the Phylesystem.
+  Users can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OpenTree website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OpenTree's website about}).
+\emph{MK suggests that WE SHOULD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE, I think it's a good idea}
 \item
-  The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
-  alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
+  The input alignment should be a single locus alignment that was used to generate the tree.
+Alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
   DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper can download it directly,
+  If the alignment is stored in TreeBase, Physcraper downloads it directly,
   either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
@@
 \item
-  A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
+  When dealing with single locus alignments, it is common for alignments to have less taxa than the tree, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
+Hence, a pruning step to reconcile taxon presence in the tree and alignment is performed.
+This verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name (and its coresonding sequence in the alignment) is needed to continue the algorithm.
+See next section.
+\emph{MK mentioned that nothing shoul be dropped here bc the name standardization should have matched everything, so I explained a bit more what this step is.
+It is not an unmatched taxa pruning but more of an absent taxa pruning.
+Also, taxa that are not matched should be dropped before right?}
 \item
-  A ``csv'' file with the summary of taxon name matching is produced for the user.
+  A ``csv'' file with the summary of taxon name standardization and pruning is produced for the user.
 \item
-  Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. See below.
-\item
-  A ``newick'' file and a ``fasta'' file containing only the matched taxa are generated and saved in the ``inputs'' folder to be used in the following steps.
+  A ``newick'' file and a ``fasta'' file containing the tree and alignment respectively with matched taxa only are generated and saved in the ``inputs'' folder to be used in the following steps.
 \end{itemize}
@@
 \item
-  Physcraper's code design relies on the GenBank DNA database to search for new sequences.
+  Physcraper uses the GenBank DNA database as source to search for new sequences.
 The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
 Using the latter speeds up the process.
-Detailed nstructions to setup a local database are provided on the software documenttaion.
+Detailed instructions to setup a local database are provided on the software documentation.
 \item
-  The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be determined by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral taxon - LICA) CITATION NEEDED.
+  The next step is to identify a ``search taxon'' to constrain the sequence search on the GenBank database within that taxonomic group.
+  The search taxon can be chosen by the user from the NCBI taxonomy. If none is provided, then the search taxon is automatically set as the Most Recent Common Ancestor (MRCA) of the
+  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
+  This is known in the OpenTree as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) \emph{I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only}.
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{Taxonomy-API-v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+  To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{taxonomic tool v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
 
@@
   database, and the sequences on the checked alignment.
-  The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
+  The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local database sequence searches.
+  For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required wait times, which slow down the searches markedly.
+  \emph{MK suggest explaining here what the ncbi waiting times are, but I don;t think it's needed, what do you think?}
 \item
@@
   in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
+  taxon. Results from each one of these BLAST runs are written down, and matched sequences are saved
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers.
 This information will be used later to store the whole sequences in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
@@
 \item
-  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
+  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon than the existing sequence \emph{MK mentioned it was unclear to her if this was an NCBI or OTT taxon.
+I think it is the unified taxonomy, hence an OTT taxon?
+I think it doesn;t harm to be super explicit about it}.
   Longer sequences belonging to the
@@
 \item
-  Reverse complement sequences are identified and translated. HOW IS THIS DONE.
+  Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions.
 \item
@@
 \item
-  A ``fasta'' file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+  A ``fasta'' file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 \end{itemize}
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroups so trees are accurately rooted.
+The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
+\emph{It would be a nice addiion to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon}
 \end{itemize}
@@
 A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
-The first one has been made available in TreeBASE as well as in the OpenTree phylesystem and is part of the synthetic tree. It samples 48 species.
-The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-The most recent one has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet, with 200 species.
-However, it has not been added to the syntehtic tree yet.
-This also makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
+The genus encompasses between 400-600 living species.
+A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (CUÉNOUD \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
+Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made tree and alignment data available in TreeBASE.
+The tree sampling 48 species was integrated to the OpenTree Phylesystem and is part of OpenTree's synthetic tree.
+The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) has been made available in the OpenTree Phylesystem and in the DRYAD repository.
+It is the best sampled yet for the genus, with 200 species.
+However, it has not been added to OpenTree's synthtetic tree yet.
+This makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
 
@@
 While not generating full phylogenies \emph{de novo}, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data.
-For example geographic locations (GBIF), fossils (PBDB), etc.
+Moreover, it can combine phylogenies with data from repositories other than molecular data.
+For example geographic locations (using GBIF ids), fossils (using PBDB ids), etc.
 \emph{from Robert: I think you can sell the program more here.
@@
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly HOW FAST IS ``RAPID''
+Physcraper has the potential to be applied for the advantage of the field to rapidly \emph{HOW FAST IS ``RAPID'' mention it in results and then here again}
 place newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
@@
 \section{Authors' Contributions}\label{authors-contributions}}
+
+EJM: Conceived study, wrote most of the code, documentation and tests.
+MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
+LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
+All authors contributed to the manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
+CUÉNOUD, P., MARTINEZ, M.A.D.P., LOIZEAU, P.-A., SPICHIGER, R., ANDREWS, S.
+\& MANEN, J.-F.
+(2000).
+Molecular phylogeny and biogeography of the genus ilex l.(Aquifoliaceae).
+\emph{Annals of Botany}, \textbf{85}, 111--122.
+
 \leavevmode\hypertarget{ref-drew2013lost}{}%
@@
 
+\leavevmode\hypertarget{ref-gottlieb2005molecular}{}%
+Gottlieb, A.M., Giberti, G.C.
+\& Poggio, L.
+(2005).
+Molecular analyses of the genus ilex (aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
+\emph{American Journal of Botany}, \textbf{92}, 352--369.
+
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
@@
 Magee, A.F., May, M.R. \& Moore, B.R. (2014). The dawn of open access to phylogenetic data. \emph{PLoS One}, \textbf{9}.
+
+\leavevmode\hypertarget{ref-manen2010history}{}%
+Manen, J.-F., Barriera, G., Loizeau, P.-A.
+\& Naciri, Y.
+(2010).
+The history of extant ilex species (aquifoliaceae): Evidence of hybridization within a miocene radiation.
+\emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
 
@@
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
+
+\leavevmode\hypertarget{ref-selbach2009new}{}%
+Selbach-Schnadelbach, A., Cavalli, S.S., Manen, J.-F., Coelho, G.C.
+\& De Souza-Chies, T.T.
+(2009).
+New information for ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (aquifoliaceae).
+\emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.
+
+\leavevmode\hypertarget{ref-setoguchi2000intersectional}{}%
+Setoguchi, H.
+\& Watanabe, I.
+(2000).
+Intersectional gene flow between insular endemics of ilex (aquifoliaceae) on the bonin islands and the ryukyu islands.
+\emph{American Journal of Botany}, \textbf{87}, 793--810.
 
@@
 
+\leavevmode\hypertarget{ref-yao2020phylogeny}{}%
+Yao, X., Song, Y., Yang, J.-B., Tan, Y.-H. \& Corlett, R.T. (2020). Phylogeny and biogeography of the hollies (ilex l., aquifoliaceae). \emph{Journal of Systematics and Evolution}.
+
 \end{document}

### 2020-07-24 2-body.Rmd
edit methods 2

@@
 
-```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at [physcraper.readthedocs.io](https://physcraper.readthedocs.io/en/latest/), along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
@@
 - In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
-  OpenTree [Phylesystem] (https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator).
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-Ix taxonomic names can't be standardized, they will be excluded for further analysis.
+  OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool *Is there a tnrs bulk citation?*.
+If taxonomic names can't be standardized, they will be excluded for further analysis.
   Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
@@
 Also, taxa that are not matched should be dropped before right?*
-- A "csv" file with the summary of taxon name standardization and pruning is produced for the user.
-- A "newick" file and a "fasta" file containing the tree and alignment respectively with matched taxa only are generated and saved in the "inputs" folder to be used in the following steps.
+- A summary "csv" file with the results from the taxon name standardization and pruning is produced for the user.
+- The standardized and pruned tree and alignment (checked tree and alignment from now on) are written down as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
 
@@
 - Physcraper uses the GenBank DNA database as source to search for new sequences.
-The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
-Using the latter speeds up the process.
-Detailed instructions to setup a local database are provided on the software documentation.
+The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
+Detailed instructions to setup a local database are provided on Physcraper's software documentation.
 <!-- Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
@@
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
-- A pairwise BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are written down, and matched sequences are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers.
-This information will be used later to store the whole sequences in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
-These parameters can be configured for each run.
-  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
-  All accepted sequences are assigned an internal identifier, and are further filtered.
-
+- A constrained BLAST search is performed, in which each sequence
+  in the alignment is BLASTed once against all database DNA sequences belonging to the search
+  taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
+- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.
+  This filtering guarantees the exclusion of whole genome sequences.
+- All accepted sequences up to this point are assigned an internal identifier, and are further filtered.
 - New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon than the existing sequence *MK mentioned it was unclear to her if this was an NCBI or OTT taxon.
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions.
+- All BLAST and filtering parameters can be customized by the user.
+- Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions [@cock2009biopython].
 - Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
@@
 - To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
+- New seqeunces that pass all filtering steps are added to the "csv" taxon summary file.
 - A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.

### 2020-07-24 docs/submission-mee-2020-07-24.tex
edit methods 2

@@
 
-\textbf{Word count}: 4198
+\textbf{Word count}: 4225
 
@@
   In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
-  OpenTree {[}Phylesystem{]} (\url{https://github.com/opentreeoflife/phylesystem}), or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application}.
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Ix taxonomic names can't be standardized, they will be excluded for further analysis.
+  OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool \emph{Is there a tnrs bulk citation?}.
+If taxonomic names can't be standardized, they will be excluded for further analysis.
   Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
@@
 \item
-  A ``csv'' file with the summary of taxon name standardization and pruning is produced for the user.
+  A summary ``csv'' file with the results from the taxon name standardization and pruning is produced for the user.
 \item
-  A ``newick'' file and a ``fasta'' file containing the tree and alignment respectively with matched taxa only are generated and saved in the ``inputs'' folder to be used in the following steps.
+  The standardized and pruned tree and alignment (checked tree and alignment from now on) are written down as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
 \end{itemize}
@@
   Physcraper uses the GenBank DNA database as source to search for new sequences.
-The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user.
-Using the latter speeds up the process.
-Detailed instructions to setup a local database are provided on the software documentation.
+The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
+Detailed instructions to setup a local database are provided on Physcraper's software documentation.
 \item
@@
 \item
-  A pairwise BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are written down, and matched sequences are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers.
-This information will be used later to store the whole sequences in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
+  A constrained BLAST search is performed, in which each sequence
+  in the alignment is BLASTed once against all database DNA sequences belonging to the search
+  taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
 \item
-  Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
-  These parameters can be configured for each run.
-  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
-  All accepted sequences are assigned an internal identifier, and are further filtered.
+  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.
+  This filtering guarantees the exclusion of whole genome sequences.
+\item
+  All accepted sequences up to this point are assigned an internal identifier, and are further filtered.
 \item
@@
 \item
-  Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions.
+  All BLAST and filtering parameters can be customized by the user.
+\item
+  Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
 \item
@@
   To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to ``physcraper/taxonomy'' folder) that is globally accessible to users.
+\item
+  New seqeunces that pass all filtering steps are added to the ``csv'' taxon summary file.
 \item

### 2020-07-24 2-body.Rmd
edit methods 3

@@
   taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
+- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length without gaps of sequences in the checked alignment.
   This filtering guarantees the exclusion of whole genome sequences.
@@
 - To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
-- New seqeunces that pass all filtering steps are added to the "csv" taxon summary file.
+- New sequences passing all filtering steps are added to the "csv" taxon summary file.
 - A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
@@
 - The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [@castresana2000selection; @castresana2002gblocks].
 - While curating the alignment is a critical step, it is not a reproducible one.
 The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
+A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
 Versioning alignments and adding them to a  global database is a next step for us!
-- Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
+- Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
 

### 2020-07-24 docs/submission-mee-2020-07-24.tex
edit methods 3

@@
 
-\caption{The Physcraper software is fully described on its documentation website at https://physcraper.readthedocs.io/en/latest/, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
+\caption{The Physcraper software is fully described on its documentation website at [physcraper.readthedocs.io](https://physcraper.readthedocs.io/en/latest/), along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
 \end{figure}
@@
   taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
+  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
 \item
-  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment.
+  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length without gaps of sequences in the checked alignment.
   This filtering guarantees the exclusion of whole genome sequences.
@@
 \item
-  New seqeunces that pass all filtering steps are added to the ``csv'' taxon summary file.
+  New sequences passing all filtering steps are added to the ``csv'' taxon summary file.
 \item
@@
   The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for alignment processing EXAMPLES OF SUCH TOOLS.
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
 \item
@@
 The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the siftware and the manually curated aligment will be ideal.
+A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
 Versioning alignments and adding them to a global database is a next step for us!
 \item
-  Users may also use Physcraper to gather matched sequences only, and apply their own preferred alignment and phylogenetic inference methods.
+  Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
 \end{itemize}
@@
 
+\leavevmode\hypertarget{ref-castresana2002gblocks}{}%
+Castresana, J. (2002). GBLOCKS: Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Version 0.91 b. Copyrighted by J. Castresana, EMBL}.
+
+\leavevmode\hypertarget{ref-castresana2000selection}{}%
+Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular biology and evolution}, \textbf{17}, 540--552.
+
 \leavevmode\hypertarget{ref-cock2009biopython}{}%

### 2020-07-24 2-body.Rmd
comments on methods

@@
   OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool *Is there a tnrs bulk citation?*.
-If taxonomic names can't be standardized, they will be excluded for further analysis.
-  Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool *Is there a tnrs bulk citation?* **EJM: Nope :/, just the link**.
+If taxonomic names can't be standardized, their taxonomic information will be excluded for further analysis.
+  Mapping tip names to this unified taxonomy saves a set of user defined characteristics
   that are essential for automatizing the phylogeny updating process.
@@
 *MK suggests that WE SHOULD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE, I think it's a good idea*
+  **EJM: Currently 2950 studies have treebase data links. It is possible that some/many of these have multiple trees in them, but that is harder to count.**
 - The input alignment should be a single locus alignment that was used to generate the tree. Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD ([www.datadryad.org](http://datadryad.org/)), or the journal where the tree was originally published.
+  DRYAD ([www.datadryad.org](http://datadryad.org/)), or a data repository associated with the journal where the tree was originally published.
   If the alignment is stored in TreeBase, Physcraper downloads it directly,
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- When dealing with single locus alignments, it is common for alignments to have less taxa than the tree, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
-Hence, a pruning step to reconcile taxon presence in the tree and alignment is performed.
-This verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name (and its coresonding sequence in the alignment) is needed to continue the algorithm.
+- When dealing with single locus alignments, it is sometimes alignments to have fewer taxa than a tree interfered from concatenated data, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
+Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
 See next section.
@@
 - A summary "csv" file with the results from the taxon name standardization and pruning is produced for the user.
-- The standardized and pruned tree and alignment (checked tree and alignment from now on) are written down as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
+- The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
 
@@
 - The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
-The search taxon can be chosen by the user from the NCBI taxonomy. If none is provided, then the search taxon is automatically set as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
+  The search taxon can be chosen by the user from the NCBI taxonomy.
+  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the 'ingroup'.
+  The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
   <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
@@
   Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) *I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only*.
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
+  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
   To identify the MRCAT of a group of taxon names, we use the OpenTree [taxonomic tool v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
@@
   <!--***How does the blast is setup? Is it an all-blast-all??***-->
+   **EJM: I think your next section answers the balst set up q nicely**
 - A constrained BLAST search is performed, in which each sequence
   in the alignment is BLASTed once against all database DNA sequences belonging to the search
-  taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80% and 120%, respectively, of the average length without gaps of sequences in the checked alignment.
-  This filtering guarantees the exclusion of whole genome sequences.
+  taxon. All results from each BLAST run are stored, and sequences with match scores better than the e-value cutoff (default to 0.00001) are saved
+  along with their corresponding metadata, i.e., their GenBank accession number.
+  The full sequence for each match is downloaded from NCBI into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
+- BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment(defaults values of 80% and 120%, respectively).
+  This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
+  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in outputfiles.
 - All accepted sequences up to this point are assigned an internal identifier, and are further filtered.
@@
 I think it is the unified taxonomy, hence an OTT taxon?
-I think it doesn;t harm to be super explicit about it*.
-Longer sequences belonging to the
-  same taxon as the original sequence will be considered further for analysis.
+I think it doesn't harm to be super explicit about it*.
+**EJM: It checks using the OTT taxonomy, but it practice OpenTree doesn't merge ncbi taxa, so it is both/either**
+Longer sequences belonging to the same taxon as the original sequence will be considered further for analysis.
 - Among the filtered sequences, there are often several representatives per taxon.
@@
 By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
-- To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to "physcraper/taxonomy" folder) that is globally accessible to users.
 - New sequences passing all filtering steps are added to the "csv" taxon summary file.
@@
 
-- Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
+- By default, Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
 Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
@@
 A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
-Versioning alignments and adding them to a  global database is a next step for us!
 - Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 - Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
-- Physcraper's final result is an updated phylogenetic hypothesis for each of the genes provided in the input alignment.
-- Tips on all trees generated by Physcraper are defined by a taxon "name space", which captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
+- Physcraper's final result is an updated phylogenetic hypothesis for the gene provided in the input alignment.
+- Tips on all trees generated by Physcraper are defined by a taxon "name space".
+The taxon metadat captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
 - Two ways to compare the updated tree with the original tree are implemented in Physcraper.

### 2020-07-24 2-body.Rmd
EJM methods comments

@@
   OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool *Is there a tnrs bulk citation?* **EJM: Nope :/, just the link**.
-If taxonomic names can't be standardized, their taxonomic information will be excluded for further analysis.
-  Mapping tip names to this unified taxonomy saves a set of user defined characteristics
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
+*Should we saythat it is based on Boyle's 2012 algorithm?
+or is it a TNRS algorithm independently developed by the OpenTree team?*
+  This step is referred to as taxon name mapping. If taxonomic names can't be mapped, their taxonomic information will be excluded from further analysis.
+  Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
   that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the standardization of taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+The most relevant of these is the taxonomic name mapping and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
 Currently, only trees connected to a published study can be stored in the Phylesystem.
   <!-- Discussion: (although there are plans to allow storage of unpublished trees). -->
-  Users can choose from among the `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
-*MK suggests that WE SHOULD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE, I think it's a good idea*
-  **EJM: Currently 2950 studies have treebase data links. It is possible that some/many of these have multiple trees in them, but that is harder to count.**
+  Users can choose from among the 2950 studies that have alignments on
+  TreeBASE that are also stored in OpenTree's Phylesystem.
+  <!--
+  **EJM: Currently 2950 studies have treebase data links. It is possible that some/many of these have multiple trees in them, but that is harder to count.** -->
+  <!-- there are `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website
+  ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).   -->
+
 - The input alignment should be a single locus alignment that was used to generate the tree. Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- When dealing with single locus alignments, it is sometimes alignments to have fewer taxa than a tree interfered from concatenated data, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
+- When dealing with single locus alignments, they sometimes have fewer taxa than the tree inferred from concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for phylogenetic analysis.
 Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
@@
 See next section.
-*MK mentioned that nothing shoul be dropped here bc the name standardization should have matched everything, so I explained a bit more what this step is.
-It is not an unmatched taxa pruning but more of an absent taxa pruning.
-Also, taxa that are not matched should be dropped before right?*
 - A summary "csv" file with the results from the taxon name standardization and pruning is produced for the user.
@@
   The search taxon can be chosen by the user from the NCBI taxonomy.
-  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the 'ingroup'.
+  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the "ingroup" (Fig. \@ref(fig:framework)).
   The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
@@
   For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required wait times, which slow down the searches markedly.
-  *MK suggest explaining here what the ncbi waiting times are, but I don;t think it's needed, what do you think?*
-  <!--***How does the blast is setup? Is it an all-blast-all??***-->
-   **EJM: I think your next section answers the balst set up q nicely**
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
 - A constrained BLAST search is performed, in which each sequence
@@
   The full sequence for each match is downloaded from NCBI into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment(defaults values of 80% and 120%, respectively).
+- BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80% and 120%, respectively).
   This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
-  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in outputfiles.
-- All accepted sequences up to this point are assigned an internal identifier, and are further filtered.
-- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon than the existing sequence *MK mentioned it was unclear to her if this was an NCBI or OTT taxon.
-I think it is the unified taxonomy, hence an OTT taxon?
-I think it doesn't harm to be super explicit about it*.
-**EJM: It checks using the OTT taxonomy, but it practice OpenTree doesn't merge ncbi taxa, so it is both/either**
-Longer sequences belonging to the same taxon as the original sequence will be considered further for analysis.
+  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in output files.
+- All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
+- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon in the unified taxonomy than the existing sequence or they are longer than the sequence in the original alignment.
+<!-- *MK mentioned it was unclear to her if this was an NCBI or OTT taxon. I think it is the unified taxonomy, hence an OTT taxon? I think it doesn't harm to be super explicit about it*
+**EJM: It checks using the OTT taxonomy, but it practice OpenTree doesn't merge ncbi taxa, so it is both/either** -->
 - Among the filtered sequences, there are often several representatives per taxon.
@@
 - Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
-- Physcraper's final result is an updated phylogenetic hypothesis for the gene provided in the input alignment.
+- Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
 - Tips on all trees generated by Physcraper are defined by a taxon "name space".
-The taxon metadat captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
+The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
 - Two ways to compare the updated tree with the original tree are implemented in Physcraper.

### 2020-07-24 docs/submission-mee-2020-07-24.tex
EJM methods comments

@@
 
-\textbf{Word count}: 4225
+\textbf{Word count}: 4114
 
@@
   OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool \emph{Is there a tnrs bulk citation?}.
-If taxonomic names can't be standardized, they will be excluded for further analysis.
-  Adding the tree to OpenTree's Phylesystem is recommended because it saves a set of user defined characteristics
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+\emph{Should we saythat it is based on Boyle's 2012 algorithm?
+or is it a TNRS algorithm independently developed by the OpenTree team?}
+  This step is referred to as taxon name mapping. If taxonomic names can't be mapped, their taxonomic information will be excluded from further analysis.
+  Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
   that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the standardization of taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+The most relevant of these is the taxonomic name mapping and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
 Currently, only trees connected to a published study can be stored in the Phylesystem.
-  Users can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OpenTree website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OpenTree's website about}).
-\emph{MK suggests that WE SHOULD ACTUALLY SAY HERE HOW MANY OPENTREE TREES HAVE ALIGNMENT DATA FROM TREEBASE, I think it's a good idea}
+  Users can choose from among the 2950 studies that have alignments on
+  TreeBASE that are also stored in OpenTree's Phylesystem.
 \item
@@
 Alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
-  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal where the tree was originally published.
+  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or a data repository associated with the journal where the tree was originally published.
   If the alignment is stored in TreeBase, Physcraper downloads it directly,
@@
 \item
-  When dealing with single locus alignments, it is common for alignments to have less taxa than the tree, simply because a single molecular marker usually does not cover all the taxa included in the phylogenetic analysis.
-Hence, a pruning step to reconcile taxon presence in the tree and alignment is performed.
-This verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name (and its coresonding sequence in the alignment) is needed to continue the algorithm.
+  When dealing with single locus alignments, they sometimes have fewer taxa than the tree inferred from concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for phylogenetic analysis.
+Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
 See next section.
-\emph{MK mentioned that nothing shoul be dropped here bc the name standardization should have matched everything, so I explained a bit more what this step is.
-It is not an unmatched taxa pruning but more of an absent taxa pruning.
-Also, taxa that are not matched should be dropped before right?}
 \item
@@
 \item
-  The standardized and pruned tree and alignment (checked tree and alignment from now on) are written down as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
+  The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
 \end{itemize}
@@
   The next step is to identify a ``search taxon'' to constrain the sequence search on the GenBank database within that taxonomic group.
-  The search taxon can be chosen by the user from the NCBI taxonomy. If none is provided, then the search taxon is automatically set as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
+  The search taxon can be chosen by the user from the NCBI taxonomy.
+  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the ``ingroup'' (Fig. \ref{fig:framework}).
+  The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
   This is known in the OpenTree as the Most Recent Common
   Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) \emph{I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only}.
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
+  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
   To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{taxonomic tool v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
@@
   For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required wait times, which slow down the searches markedly.
-  \emph{MK suggest explaining here what the ncbi waiting times are, but I don;t think it's needed, what do you think?}
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
 \item
@@
   in the alignment is BLASTed once against all database DNA sequences belonging to the search
-  taxon. All results from each BLAST run are written down, and sequences above the e-value (default to 0.00001) are saved
-  along with their corresponding identification numbers, i.e., their GenBank accession numbers, which will be used later to get the whole sequence from the GenBank database into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
+  taxon. All results from each BLAST run are stored, and sequences with match scores better than the e-value cutoff (default to 0.00001) are saved
+  along with their corresponding metadata, i.e., their GenBank accession number.
+  The full sequence for each match is downloaded from NCBI into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
 \item
-  BLAST result sequences will be discarded if they fall outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length without gaps of sequences in the checked alignment.
-  This filtering guarantees the exclusion of whole genome sequences.
+  BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80\% and 120\%, respectively).
+  This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
+  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in output files.
 \item
-  All accepted sequences up to this point are assigned an internal identifier, and are further filtered.
+  All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
 \item
-  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon than the existing sequence \emph{MK mentioned it was unclear to her if this was an NCBI or OTT taxon.
-I think it is the unified taxonomy, hence an OTT taxon?
-I think it doesn;t harm to be super explicit about it}.
-  Longer sequences belonging to the
-  same taxon as the original sequence will be considered further for analysis.
+  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon in the unified taxonomy than the existing sequence or they are longer than the sequence in the original alignment.
 \item
@@
 \item
-  To speed up future runs, accepted sequences are downloaded in full, and stored in a local directory (default to ``physcraper/taxonomy'' folder) that is globally accessible to users.
-\item
   New sequences passing all filtering steps are added to the ``csv'' taxon summary file.
@@
 \item
-  Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
+  By default, Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
 Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
@@
 A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
-Versioning alignments and adding them to a global database is a next step for us!
 \item
@@
 \item
-  Physcraper's final result is an updated phylogenetic hypothesis for each of the genes provided in the input alignment.
+  Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
 \item
-  Tips on all trees generated by Physcraper are defined by a taxon ``name space'', which captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
+  Tips on all trees generated by Physcraper are defined by a taxon ``name space''.
+The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
 \item

### 2020-07-24 2-alt-1-intro.Rmd
clean intro

@@
 By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees using loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
+<!-- The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree. -->
 These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
@@
 
-The Open Tree of Life data store, the [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373), contains more than 4,500 phylogenetic trees from published studies.
+The OpenTree database, the [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373) [@mctavish2015phylesystem], stores more than 4,500 phylogenetic trees from published studies.
 The tips in these trees are mapped to a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.

### 2020-07-24 2-body.Rmd
clean intro

@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
+Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+Automated trees can provide a quick inference of potential phylogenetic relationships, of problems in the taxonomic assignments of sequences, and flag areas of systematic interest.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
-Currently, estimates of phylogenetic relationships are publicly available for about 100 000 taxa only [@piel2009treebase; @hinchliff2015synthesis; @opentreeoflife2019synth], representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
-
-The discrepancy between molecular data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way [@drew2013lost; @magee2014dawn; @mctavish2018bioessay].
-However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-
-We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
-By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees using loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
-These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
-
-Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
-The Open Tree of Life (OpenTree from now on https://opentreeoflife.github.io/) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
-OpenTree aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This "synthetic" tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
-
-The Open Tree of Life data store, the [Phylesystem](https://academic.oup.com/bioinformatics/article/31/17/2794/183373), contains more than 4,500 phylogenetic trees from published studies.
-The tips in these trees are mapped to a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
-This provides a resource for finding existing estimates of phylogenetic relationships,
-and assessing which regions of the tree of life are lacking available phylogenetic estimates.
-
-By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
-the TreeBASE repository [@piel2009treebase] and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
-
-ARGUMENT - GENES ARE STILL USEFUL IN THE GENOMICS ERA, AND HERE'S WHY:
-GENOMIC MARKERS SUCH AS RADSEQ, SNP, MISCROSATS AND UCES - ARE HOMOLOGY HYPOTHESIS THE SAME ON THESE MARKERS THAN FOR PROTEIN CODING AND NON CODING LOCI?
-WHAT ARE THE PROS AND CONS OF USING GENE ALIGNMENTS ONLY AND NOT GENOMIC MARKERS?
-Martha: Genomic data is not available for a large number of taxa.
-While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
-One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
-This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016misconceptions]
-In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
-
-Indeed, some systematics support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving; ROJAS ET AL.
-bats paper, take citations from there].
-It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
-
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
-
-In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction ("phylogenetic pipelines") such as PHYLOTA [@sanderson2008phylota], PHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward].
-Physcraper shares a similar conceptual framework to Pumper [@izquierdo2014pumper], but that software is not currently supported or developed (*or runnable at all honestly...*)
-
-Data input availability:
-As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of around 100 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
-This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
-
-The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
-If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
-This provides a loose means of linking the tree with the exact alignment that generated it.
-
-Often, linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., MORE EXAMPLES OF DATABASES providing a way to automatically link data from different databases.
-
-Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
-<!-- WE COULD EXPLAIN WHY PYTHON IS BEST FOR THIS: does it help with interoperability? It is flexible enough?
-Maybe not relevant for now, so we'll keep it out-->
-By design, this approach focuses on data interoperability.
-By automating taxonomic name matching across NCBI, OpenTree, GBIF and virtually any biological database, users can perform downstream analyses straightforwardly.
-For example, it automatizes and standardizes comparison of phylogenetic hypotheses with currently known relationships from the synthetic Open Tree of Life, the Open Tree of Life taxonomy tree and any phylogeny that is stored in the
-Phylesystem.
-
-We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
+One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
+To standardize taxonomic names, a unified taxonomy is necessary.
+The main aim of the Open Tree of Life (OpenTree from now on) project
+<!-- [project](https://opentreeoflife.github.io/)  -->
+is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+This "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
+For this goal, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility (GBIF) [@secretariat2017gbif], the Encyclopedia of Life (EOL) [@wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- When dealing with single locus alignments, they sometimes have fewer taxa than the tree inferred from concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for phylogenetic analysis.
+- Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
 Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
@@
   This is known in the OpenTree as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) *I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only*.
+  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA).
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
@@
 - The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [@castresana2000selection; @castresana2002gblocks].
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection; @castresana2002gblocks].
 - While curating the alignment is a critical step, it is not a reproducible one.
@@
 The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
-*It would be a nice addiion to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon*
+<!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->
 

### 2020-07-24 2.4-discussion-ideas.Rmd
mk's comments
clean intro

@@
+databases designed to preserve and democratize access to biological data constitute an amazing resource for evolutionary estimation
+
 Usually, initial ideas about the data are changed by analyses.
@@
 \comment(hmm yeah, I think it is probably better for the latter - phylogeneticists who are tired of blast searching :P)
+
+There is danger in creating a unified standard, however, in this case it provides usefulness.
+Physcraper might be overlapping/partially solving the goal of the Encyclopedia of Life:
+"the exploration and analysis of biodiversity at a vastly accelerated pace.
+Its momentum will return systematics from its long sojourn at the margin and back into the mainstream of science." [@wilson2003encyclopedia]

### 2020-07-24 docs/submission-mee-2020-07-24.tex
clean intro

@@
 
-\textbf{Word count}: 4114
+\textbf{Word count}: 3107
 
@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
+Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+Automated trees can provide a quick inference of potential phylogenetic relationships, of problems in the taxonomic assignments of sequences, and flag areas of systematic interest.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Currently, estimates of phylogenetic relationships are publicly available for about 100 000 taxa only (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
-
-The discrepancy between molecular data availability and phylogenetic estimates can be partially explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
-However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-
-We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
-By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees using loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
-These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
-
-Physcraper leverages public phylogenetic data stored in Open Tree of Life and in TreeBase.
-The Open Tree of Life (OpenTree from now on \url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
-OpenTree aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
-
-The Open Tree of Life data store, the \href{https://academic.oup.com/bioinformatics/article/31/17/2794/183373}{Phylesystem}, contains more than 4,500 phylogenetic trees from published studies.
-The tips in these trees are mapped to a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
-This provides a resource for finding existing estimates of phylogenetic relationships,
-and assessing which regions of the tree of life are lacking available phylogenetic estimates.
-
-By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
-the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
-
-ARGUMENT - GENES ARE STILL USEFUL IN THE GENOMICS ERA, AND HERE'S WHY:
-GENOMIC MARKERS SUCH AS RADSEQ, SNP, MISCROSATS AND UCES - ARE HOMOLOGY HYPOTHESIS THE SAME ON THESE MARKERS THAN FOR PROTEIN CODING AND NON CODING LOCI?
-WHAT ARE THE PROS AND CONS OF USING GENE ALIGNMENTS ONLY AND NOT GENOMIC MARKERS?
-Martha: Genomic data is not available for a large number of taxa.
-While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
-One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
-This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016})
-In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
-
-Indeed, some systematics support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.
-bats paper, take citations from there{]}.
-It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
-
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
-
-In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-Physcraper shares a similar conceptual framework to Pumper (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}), but that software is not currently supported or developed (\emph{or runnable at all honestly\ldots{}})
-
-Data input availability:
-As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of around 100 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
-This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
-
-The Phylesystem (OpenTree's datastore) (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
-If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
-This provides a loose means of linking the tree with the exact alignment that generated it.
-
-Often, linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., MORE EXAMPLES OF DATABASES providing a way to automatically link data from different databases.
-
-Physcraper is a Python encoded pipeline designed to update previously known phylogenetic relationships in a continuous manner, by connecting phylogenies stored in the OpenTree Phylesystem with alignments from TreeBASE and newly released DNA data from GenBank, by using the OpenTree metadata system to connect independent databases through their unique taxon identifiers, automatizing taxonomic name matching across them.
-By design, this approach focuses on data interoperability.
-By automating taxonomic name matching across NCBI, OpenTree, GBIF and virtually any biological database, users can perform downstream analyses straightforwardly.
-For example, it automatizes and standardizes comparison of phylogenetic hypotheses with currently known relationships from the synthetic Open Tree of Life, the Open Tree of Life taxonomy tree and any phylogeny that is stored in the
-Phylesystem.
-
-We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
+One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
+To standardize taxonomic names, a unified taxonomy is necessary.
+The main aim of the Open Tree of Life (OpenTree from now on) project
+is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
+For this goal, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility (GBIF) (Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL) (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 \item
-  When dealing with single locus alignments, they sometimes have fewer taxa than the tree inferred from concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for phylogenetic analysis.
+  Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
 Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
@@
   This is known in the OpenTree as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA) \emph{I looked for a citation but it seems that it is a concept used on the open tree of life API wikis only}.
+  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA).
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
@@
   The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
 \item
@@
 The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
-\emph{It would be a nice addiion to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon}
 \end{itemize}
@@
 The genus encompasses between 400-600 living species.
-A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (CUÉNOUD \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
+A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
 Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made tree and alignment data available in TreeBASE.
@@
 
-\leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli, A., Hettling, H., Condamine, F.L., Vos, K., Nilsson, R.H., Sanderson, M.J., Sauquet, H., Scharn, R., Silvestro, D., Töpel, M.
-\& others.
-(2017).
-Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
-\emph{Systematic Biology}, \textbf{66}, 152--166.
-
 \leavevmode\hypertarget{ref-benson2000genbank}{}%
@@
 \leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
-CUÉNOUD, P., MARTINEZ, M.A.D.P., LOIZEAU, P.-A., SPICHIGER, R., ANDREWS, S.
-\& MANEN, J.-F.
+Cuénoud, P., Martinez, M.A.
+del P., Loizeay, P.-A., Spichiger, R., Andrews, S.
+\& Manen, J.-F.
 (2000).
@@
 
-\leavevmode\hypertarget{ref-drew2013lost}{}%
-Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez, R., Katz, L.A., Crandall, K.A., Hibbett, D.S.
-\& Soltis, D.E.
-(2013).
-Lost branches on the tree of life.
-\emph{PLoS biology}, \textbf{11}.
-
-\leavevmode\hypertarget{ref-eaton2016misconceptions}{}%
-Eaton, D.A.R., Spriggs, E.L., Park, B.
-\& Donoghue, M.J.
-(2016).
-Misconceptions on Missing Data in RAD-seq Phylogenetics with a Deep-scale Example from Flowering Plants.
-\emph{Systematic Biology}, \textbf{66}, 399--412.
-Retrieved from \url{https://doi.org/10.1093/sysbio/syw092}
-
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
 Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic acids research}, \textbf{32}, 1792--1797.
-
-\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
-Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
-\& Mendoza, C.G.
-(2017).
-A pilot study applying the plant anchored hybrid enrichment method to new world sages (salvia subgenus calosphace; lamiaceae).
-\emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
 
@@
 
-\leavevmode\hypertarget{ref-hinchliff2015synthesis}{}%
-Hinchliff, C.E., Smith, S.A., Allman, J.F., Burleigh, J.G., Chaudhary, R., Coghill, L.M., Crandall, K.A., Deng, J., Drew, B.T., Gazis, R.
-\& others.
-(2015).
-Synthesis of phylogeny and taxonomy into a comprehensive tree of life.
-\emph{Proceedings of the National Academy of Sciences}, \textbf{112}, 12764--12769.
-
-\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
-
-\leavevmode\hypertarget{ref-magee2014dawn}{}%
-Magee, A.F., May, M.R. \& Moore, B.R. (2014). The dawn of open access to phylogenetic data. \emph{PLoS One}, \textbf{9}.
-
 \leavevmode\hypertarget{ref-manen2010history}{}%
@@
 
-\leavevmode\hypertarget{ref-mctavish2018bioessay}{}%
-McTavish, E.J., Drew, B.T., Redelings, B.
-\& Cranston, K.A.
-(2017).
-How and Why to Build a Unified Tree of Life.
-\emph{BioEssays}, \textbf{39}.
-Retrieved April 10, 2018, from \url{https://onlinelibrary.wiley.com/doi/abs/10.1002/bies.201700114}
-
 \leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
@@
 \emph{Bioinformatics}, \textbf{31}, 2794--2800.
-
-\leavevmode\hypertarget{ref-morrison2006multiple}{}%
-Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
-
-\leavevmode\hypertarget{ref-opentreeoflife2019synth}{}%
-OpenTreeOfLife, Redelings, B., Reyes, L.L.S., Cranston, K.A., Allman, J., Holder, M.T.
-\& McTavish, E.J.
-(2019).
-Open tree of life synthetic tree.
-Retrieved from \url{https://doi.org/10.5281/zenodo.3937742}
 
@@
 
-\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A. \& Wehe, A. (2008). The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. \emph{Systematic Biology}, \textbf{57}, 335--346.
-
 \leavevmode\hypertarget{ref-san2010molecular}{}%
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
+
+\leavevmode\hypertarget{ref-secretariat2017gbif}{}%
+Secretariat, G. (2017). GBIF backbone taxonomy. \emph{Checklist Dataset {[}cited 2017 Nov 14{]}. doi}, \textbf{10}.
 
@@
 \emph{American Journal of Botany}, \textbf{87}, 793--810.
-
-\leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith, S.A., Beaulieu, J.M.
-\& Donoghue, M.J.
-(2009).
-Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
-\emph{BMC evolutionary biology}, \textbf{9}, 37.
-
-\leavevmode\hypertarget{ref-song2012resolving}{}%
-Song, S., Liu, L., Edwards, S.V.
-\& Wu, S.
-(2012).
-Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
-\emph{Proceedings of the National Academy of Sciences}, \textbf{109}, 14942--14947.
 
@@
 
+\leavevmode\hypertarget{ref-wilson2003encyclopedia}{}%
+Wilson, E.O. (2003). The encyclopedia of life. \emph{Trends in Ecology \& Evolution}, \textbf{18}, 77--80.
+
 \leavevmode\hypertarget{ref-yao2020phylogeny}{}%

### 2020-07-24 2-body.Rmd
data availability

@@
 Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-Automated trees can provide a quick inference of potential phylogenetic relationships, of problems in the taxonomic assignments of sequences, and flag areas of systematic interest.
 
@@
 This "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility (GBIF) [@secretariat2017gbif], the Encyclopedia of Life (EOL) [@wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+For this goal, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility[GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
+The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
+If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
+This provides a loose means of linking the tree with the exact alignment that generated it.
 # The Physcraper framework
@@
 
+Physcraper source code available at https://github.com/McTavishLab/physcraper
+Documentation available at https://physcraper.readthedocs.io/en/latest/index.html
+Illustrated examples available at https://github.com/McTavishLab/physcraperex
+This is a reproducible manuscript available at https://github.com/McTavishLab/physcraper_ms
+
 \newpage

### 2020-07-24 docs/submission-mee-2020-07-24.tex
data availability

@@
 
-\textbf{Word count}: 3107
+\textbf{Word count}: 3167
 
@@
 This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility (GBIF) (Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL) (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+For this goal, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility(GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL; Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
-\hypertarget{the-physcraper-framework}{%
-\section{The Physcraper framework}\label{the-physcraper-framework}}
+The Phylesystem (OpenTree's datastore) (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
+If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
+This provides a loose means of linking the tree with the exact alignment that generated it.
+\# The Physcraper framework
 
@@
 
+Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
+Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
+Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
+This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}
+
 \newpage

### 2020-07-24 2-alt-1-intro.Rmd
clean intro 2

@@
 One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
+Different research questions call for different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq, SNP, mirosats and UCE's.
 This variety of approaches results in non-overlapping data sets across taxa.

### 2020-07-24 2-body.Rmd
clean intro 2

@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
-Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+Extending existing phylogenetic estimates with DNA sequence data from molecular databases in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Here, we propose Physcraper, a tool that takes advantage of existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, leveraging on molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
+One of the main challenges to automatically connect data from disparate biological databases in a phylogenetic context is the prevalence of taxonomic idiosincracies across databases.
 To standardize taxonomic names, a unified taxonomy is necessary.
@@
 is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility[GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies/integrates taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility[GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
-The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
-If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
-This provides a loose means of linking the tree with the exact alignment that generated it.
+Another challenge to incorporating molecular data from public databases to update our phylogenetic knowledge is producing high-quality homology hypotheses.
+While the genomic era has allowed us to obtain large quantities of molecular data from a single species, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, RadSeq, SNP, mirosats and UCE's.
+Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
+Moreover, even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences [@eaton2016misconceptions].
+In contrast, decades of single locus sequencing have generated homologous DNA data sets across large phylogenetic scales.
+Indeed, GenBank release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241, 000 distinct taxa [about `r round(236023/240708, digits=4)*100`% of taxa in the NCBI taxonomy release 159; @sanderson2008phylota].
+Various phylogenetic pipelines have been developed to assemble homology hypothesis by mining molecular databases (*maybe a box of all pipelines?*)
+
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 curated *non-genomic?* alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.),
+constituting an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
+
+The OpenTree Phylesystem has incorporated all phylogenies from TreeBASE, while saving metadata linking trees to the corresponding alignment repository in TreeBASE.
+If there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
+Hence, this metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
+
+By linking molecular data, available from databases such as the GenBank [@benson2000genbank; @wheeler2000database], to alignments and phylogenies, available in
+the TreeBASE repository [@piel2009treebase] and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
+
 # The Physcraper framework
@@
 
-We thank members of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
+We thank members of the "short bar" Science and Engineering Building 1, University of California Merced, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
 

### 2020-07-24 docs/submission-mee-2020-07-24.tex
clean intro 2

--- 2020-07-24 docs/submission-mee-2020-07-24.tex
@@
 
-\textbf{Word count}: 3107
+\textbf{Word count}: 3431
 
@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
-Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-Automated trees can provide a quick inference of potential phylogenetic relationships, of problems in the taxonomic assignments of sequences, and flag areas of systematic interest.
+Extending existing phylogenetic estimates with DNA sequence data from molecular databases in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Here, we propose Physcraper, a tool that takes advantage of existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, leveraging on molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
+One of the main challenges to automatically connect data from disparate biological databases in a phylogenetic context is the prevalence of taxonomic idiosincracies across databases.
 To standardize taxonomic names, a unified taxonomy is necessary.
@@
 is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility (GBIF) (Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL) (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies/integrates taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility(GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL; Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+
+Another challenge to incorporate molecular data from public databases to update our phylogenetic knowledge is producing high-quality homology hypotheses.
+While the genomic era has allowed us to obtain large quantities of molecular data from a single species, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, RadSeq, SNP, mirosats and UCE's.
+Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
+Moreover, even when the same sequencing approach is applied, such as RadSeq, phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016}).
+In contrast, decades of single locus sequencing have generated homologous DNA data sets across large phylogenetic scales.
+Indeed, GenBank release number 159 (April 15, 2007) hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 curated \emph{non-genomic?} alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)},
+constituting an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
+
+The OpenTree Phylesystem has incorporated all phylogenies from TreeBASE, while saving metadata linking trees to the corresponding alignment repository in TreeBASE.
+If there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
+Hence, this metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
+
+By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
+the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and OpenTree's Phylesystem, we can place new biological data in an evolutionary context.
 
@@
 
-We thank members of the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
+We thank members of the ``short bar'' Science and Engineering Building 1, University of California Merced, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
 
@@
 \section{Data Avilability}\label{data-avilability}}
+
+Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
+Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
+Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
+This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}
 
@@
 Rees, J.A. \& Cranston, K. (2017). Automated assembly of a reference taxonomy for phylogenetic data synthesis. \emph{Biodiversity Data Journal}.
+
+\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
+Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A. \& Wehe, A. (2008). The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. \emph{Systematic Biology}, \textbf{57}, 335--346.
 

### 2020-07-27 1-abstract.Rmd
abstract

@@
 
-1. Physcraper is an open‐source, command-line Python program that automatizes the update of published
-phylogenies by enriching underlying gene alignments with public DNA sequence data, and linking taxonomic information across databases.
-This provides a framework for comparison of published phylogenies with their updated versions, by using the conflict Application Programming Interface (API) function of the Open Tree of Life project.
+1. Physcraper is a command-line Python program that automatizes the update of published
+phylogenies by enriching underlying gene alignments with public DNA sequence data.
+It builds upon tools from the Open Tree of Life project to link taxonomic information across databases, providing a framework for straightforward comparison of published phylogenies with their updated versions.
+<!-- the conflict Application Programming Interface (API) function of  the Open Tree of Life project-->
 
 1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-hypotheses based on already available expert phylogenetic knowledge.
+hypotheses based on publicly available expert phylogenetic knowledge.
 Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
@@
 We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS and is released under an open source license.
+Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and

### 2020-07-27 2-alt-1-intro.Rmd
intro clean 3

@@
 bats paper, take citations from there].
-It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments [@fragoso2017pilot].
+It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for automated assemblies [@fragoso2017pilot].
 

### 2020-07-27 2-body.Rmd
intro clean 3

--- 2020-07-24 2-body.Rmd
@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
-Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
+Extending existing phylogenetic estimates with new DNA sequence data, geographical location, fossil data, etc., in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
+The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
 To standardize taxonomic names, a unified taxonomy is necessary.
@@
 is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility[GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from all these databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies/integrates taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility [GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
-The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
-If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
-This provides a loose means of linking the tree with the exact alignment that generated it.
+Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
+<!-- ROJAS ET AL. bats paper, take citations from there -->
+While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites and ultra-conserved elements (UCEs).
+Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
+Moreover, some types of genomic data are not suitable for phylogenetic inference at any scale.
+For example, phylogenetic distance can cause allelic dropout at deeper divergences with RADSeq data sets [@eaton2016misconceptions].
+<!-- allelic dropout -->
+In contrast, decades of single locus sequencing have generated homologous DNA data
+<!-- across large phylogenetic scales  -->
+that can be used for phylogenetic reconstruction at many scales.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
+Assembling a DNA alignment from such a massive database can be done "by hand" but it requires huge amounts of time and it is a mostly non-reproducible method.
+Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
+<!-- *maybe a box of all pipelines?* -->
+Still, there is a large tradition in phylogenetics of building molecular character matrices of few, thoughtfully curated markers, as manual curation of alignments often improves phylogenetic reconstructions [@fragoso2017pilot].
+
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), as well as an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+
+The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] has integrated all phylogenies from TreeBASE, while saving metadata linking trees to their corresponding alignment repository in TreeBASE.
+If there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
+Hence, this metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
+
+Physcraper provides programmatic access to the GenBank, TreeBASE and OpenTree Phylesystem public databases, to link molecular data to alignments and phylogenies, as well as other biological databases to place new biological data in an evolutionary context.
+
 # The Physcraper framework

### 2020-07-27 2-alt-1-intro.Rmd
intro clean 4

@@
 We propose Physcraper as a tool to make data connections across biological databases in a phylogenetic context for the advantage of phylogenetics and comparative biology, as well as an effort towards establishing fully reproducible workflows in phylogenetics.
+
+Physcraper is a Python encoded pipeline that focuses on database interoperability and provides programmatic access to the GenBank, TreeBASE and OpenTree public databases, to automatize taxonomic name matching across virtually any  biological database, so users can perform any kind of downstream analysis straightforwardly.
+It's main goal is to link molecular data to alignments and phylogenies, as well as other biological databases to place new biological data in an evolutionary context.

### 2020-07-27 2-body.Rmd
intro clean 4

@@
 The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified taxonomy is necessary.
+To standardize taxonomic names, a unified taxonomy is needed.
 The main aim of the Open Tree of Life (OpenTree from now on) project
@@
 Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies/integrates taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global biodiversity Information Facility [GBIF; @secretariat2017gbif], the Encyclopedia of Life [EOL; @wilson2003encyclopedia], etc., providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the Encyclopedia of Life [EOL; @wilson2003encyclopedia], providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 <!-- ROJAS ET AL. bats paper, take citations from there -->
-While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites and ultra-conserved elements (UCEs).
+While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
 Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
@@
 that can be used for phylogenetic reconstruction at many scales.
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
-Assembling a DNA alignment from such a massive database can be done "by hand" but it requires huge amounts of time and it is a mostly non-reproducible method.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
+Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible method.
 Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
@@
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), as well as an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
+<!-- However, if there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
+Hence, Phylesystem's metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
+-->
 
-The Phylesystem (OpenTree's datastore) [@mctavish2015phylesystem] has integrated all phylogenies from TreeBASE, while saving metadata linking trees to their corresponding alignment repository in TreeBASE.
-If there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
-Hence, this metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
-
-Physcraper provides programmatic access to the GenBank, TreeBASE and OpenTree Phylesystem public databases, to link molecular data to alignments and phylogenies, as well as other biological databases to place new biological data in an evolutionary context.
+We propose Physcraper as a Python encoded pipeline that relies on programmatic access protocols to link molecular data from GenBank, to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
+It focuses on database interoperability, by automatizing taxonomic name matching across biological databases, so users can perform any downstream analysis straightforwardly, as well as placing new data from virtually any biological database in an evolutionary context.
 

### 2020-07-27 docs/submission-mee-2020-07-27.tex
intro clean 4

--- 2020-07-24 docs/submission-mee-2020-07-24.tex
@@
 
-\textbf{Word count}: 3107
+\textbf{Word count}: 3652
 
@@
 \item
-  Physcraper is an open‐source, command-line Python program that automatizes the update of published
-  phylogenies by enriching underlying gene alignments with public DNA sequence data, and linking taxonomic information across databases.
-  This provides a framework for comparison of published phylogenies with their updated versions, by using the conflict Application Programming Interface (API) function of the Open Tree of Life project.
+  Physcraper is a command-line Python program that automatizes the update of published
+  phylogenies by enriching underlying gene alignments with public DNA sequence data.
+It builds upon tools from the Open Tree of Life project to link taxonomic information across databases, providing a framework for straightforward comparison of published phylogenies with their updated versions.
 \item
   Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-  hypotheses based on already available expert phylogenetic knowledge.
+  hypotheses based on publicly available expert phylogenetic knowledge.
   Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
@@
   We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS and is released under an open source license.
+Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and
@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with DNA sequence data from molecular databases in an automatic and continuous manner is possible by generating a connection between biological databases.
-Here, we propose Physcraper, a tool that uses a starting tree and single locus alignment, to take advantage of existing research and extend trees using molecular data that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-Automated trees can provide a quick inference of potential phylogenetic relationships, of problems in the taxonomic assignments of sequences, and flag areas of systematic interest.
+Extending existing phylogenetic estimates with new DNA sequence data, geographical location, fossil data, etc., in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-One of the main challenges in connecting data from disparate biological databases is the prevalence of taxonomic idiosincracies across databases.
-To standardize taxonomic names, a unified taxonomy is necessary.
+The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
+To standardize taxonomic names, a unified taxonomy is needed.
 The main aim of the Open Tree of Life (OpenTree from now on) project
 is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-This ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the rest of taxa are placed in the tree based on their taxonomic assignment.
-For this goal, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global biodiversity Information Facility (GBIF) (Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the Encyclopedia of Life (EOL) (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), etc., providing a key resource that can be used to connect data from all the aforementioned databases to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the Encyclopedia of Life (EOL; Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+
+Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
+Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
+Moreover, some types of genomic data are not suitable for phylogenetic inference at any scale.
+For example, phylogenetic distance can cause allelic dropout at deeper divergences with RADSeq data sets (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016}).
+In contrast, decades of single locus sequencing have generated homologous DNA data
+that can be used for phylogenetic reconstruction at many scales.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time (about 98.05\% of taxa; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible method.
+Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+Still, there is a large tradition in phylogenetics of building molecular character matrices of few, thoughtfully curated markers, as manual curation of alignments often improves phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
+
+We propose Physcraper as a Python encoded pipeline that relies on programmatic access protocols to link molecular data from GenBank, to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
+It focuses on database interoperability, by automatizing taxonomic name matching across biological databases, so users can perform any downstream analysis straightforwardly, as well as placing new data from virtually any biological database in an evolutionary context.
 
@@
 \emph{Nucleic acids research}, \textbf{25}, 3389--3402.
+
+\leavevmode\hypertarget{ref-antonelli2017toward}{}%
+Antonelli, A., Hettling, H., Condamine, F.L., Vos, K., Nilsson, R.H., Sanderson, M.J., Sauquet, H., Scharn, R., Silvestro, D., Töpel, M.
+\& others.
+(2017).
+Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
+\emph{Systematic Biology}, \textbf{66}, 152--166.
 
@@
 
+\leavevmode\hypertarget{ref-eaton2016misconceptions}{}%
+Eaton, D.A.R., Spriggs, E.L., Park, B.
+\& Donoghue, M.J.
+(2016).
+Misconceptions on Missing Data in RAD-seq Phylogenetics with a Deep-scale Example from Flowering Plants.
+\emph{Systematic Biology}, \textbf{66}, 399--412.
+Retrieved from \url{https://doi.org/10.1093/sysbio/syw092}
+
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
 Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic acids research}, \textbf{32}, 1792--1797.
+
+\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
+Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
+\& Mendoza, C.G.
+(2017).
+A pilot study applying the plant anchored hybrid enrichment method to new world sages (salvia subgenus calosphace; lamiaceae).
+\emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
 
@@
 Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
+
+\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
+Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
 
@@
 
+\leavevmode\hypertarget{ref-morrison2006multiple}{}%
+Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
+
 \leavevmode\hypertarget{ref-piel2009treebase}{}%
@@
 \emph{American Journal of Botany}, \textbf{87}, 793--810.
+
+\leavevmode\hypertarget{ref-smith2009mega}{}%
+Smith, S.A., Beaulieu, J.M.
+\& Donoghue, M.J.
+(2009).
+Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
+\emph{BMC evolutionary biology}, \textbf{9}, 37.
+
+\leavevmode\hypertarget{ref-song2012resolving}{}%
+Song, S., Liu, L., Edwards, S.V.
+\& Wu, S.
+(2012).
+Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
+\emph{Proceedings of the National Academy of Sciences}, \textbf{109}, 14942--14947.
 

### 2020-07-28 2-body.Rmd
worms

@@
 Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the Encyclopedia of Life [EOL; @wilson2003encyclopedia], providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the World Register of Marine Species [WoRMS; www.marinespecies.org/], providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 
-```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at [physcraper.readthedocs.io](https://physcraper.readthedocs.io/en/latest/), along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
@@
   <!-- Discussion: (although there are plans to allow storage of unpublished trees). -->
-  Users can choose from among the 2950 studies that have alignments on
-  TreeBASE that are also stored in OpenTree's Phylesystem.
+  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on
+  TreeBASE.
   <!--
@@
 See next section.
-- A summary "csv" file with the results from the taxon name standardization and pruning is produced for the user.
+- A summary "csv" file with results from the taxon name standardization and pruning is produced for the user.
 - The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
@@
   This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
-  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in output files.
+  The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
 - All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
-- New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon in the unified taxonomy than the existing sequence or they are longer than the sequence in the original alignment.
+- New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the unified taxonomy, or they are longer than the sequence in the input alignment.
 <!-- *MK mentioned it was unclear to her if this was an NCBI or OTT taxon. I think it is the unified taxonomy, hence an OTT taxon? I think it doesn't harm to be super explicit about it*
-**EJM: It checks using the OTT taxonomy, but it practice OpenTree doesn't merge ncbi taxa, so it is both/either** -->
+**EJM: It checks using the OTT taxonomy, but in practice OpenTree doesn't merge ncbi taxa, so it is both/either** -->
 - Among the filtered sequences, there are often several representatives per taxon.
@@
 - Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
+By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
 - New sequences passing all filtering steps are added to the "csv" taxon summary file.

### 2020-07-28 docs/submission-mee-2020-07-27.tex
worms

@@
 
-\textbf{Word count}: 3652
+\textbf{Word count}: 3646
 
@@
 Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the Encyclopedia of Life (EOL; Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}), providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}, providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 
-\caption{The Physcraper software is fully described on its documentation website at [physcraper.readthedocs.io](https://physcraper.readthedocs.io/en/latest/), along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
+\caption{The Physcraper software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
 \end{figure}
@@
 Currently, only trees connected to a published study can be stored in the Phylesystem.
-  Users can choose from among the 2950 studies that have alignments on
-  TreeBASE that are also stored in OpenTree's Phylesystem.
+  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on
+  TreeBASE.
 \item
@@
 \item
-  A summary ``csv'' file with the results from the taxon name standardization and pruning is produced for the user.
+  A summary ``csv'' file with results from the taxon name standardization and pruning is produced for the user.
 \item
@@
   This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
-  The GenBank accession numbers of sequences removed to due to not meeting e-value or length cutoffs are stored in output files.
+  The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
 \item
@@
 \item
-  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they represent a different taxon in the unified taxonomy than the existing sequence or they are longer than the sequence in the original alignment.
+  New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the unified taxonomy, or they are longer than the sequence in the input alignment.
 \item
@@
   Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST search cycle is performed in which only sequences in the processed original alignment are blasted.
+By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
 \item
@@
 
-\leavevmode\hypertarget{ref-wilson2003encyclopedia}{}%
-Wilson, E.O. (2003). The encyclopedia of life. \emph{Trends in Ecology \& Evolution}, \textbf{18}, 77--80.
-
 \leavevmode\hypertarget{ref-yao2020phylogeny}{}%

### 2020-07-28 2-body.Rmd
looking fab

@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, fossil data, etc., in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
 Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified taxonomy is needed.
+The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
+To standardize taxonomic names, a unified system is needed.
 The main aim of the Open Tree of Life (OpenTree from now on) project
@@
 Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
-<!-- ROJAS ET AL. bats paper, take citations from there -->
-While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
-Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
-Moreover, some types of genomic data are not suitable for phylogenetic inference at any scale.
-For example, phylogenetic distance can cause allelic dropout at deeper divergences with RADSeq data sets [@eaton2016misconceptions].
-<!-- allelic dropout -->
-In contrast, decades of single locus sequencing have generated homologous DNA data
-<!-- across large phylogenetic scales  -->
-that can be used for phylogenetic reconstruction at many scales.
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
-Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible method.
-Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
+Genomics has, and will continue to, revolutionize phylogenetic inference.
+However, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
+Having a variety of approaches can result in largely non-overlapping genomic data sets across taxa.
+Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]."
+While targeted capture approaches are ameliorating the problem of non-overlapping genomic data sets [@andermann2020guide; @jones2016targeted], in the meantime decades of single locus sequencing have already generated homologous DNA data that can be used for phylogenetic reconstruction at many scales.
+
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
+These sequences were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
+Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible approach.
+Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
 <!-- *maybe a box of all pipelines?* -->
-Still, there is a large tradition in phylogenetics of building molecular character matrices of few, thoughtfully curated markers, as manual curation of alignments often improves phylogenetic reconstructions [@fragoso2017pilot].
+However, building molecular character matrices of few, thoughtfully curated markers, including manual curation of alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
+Linking these alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
 The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
@@
 
-We propose Physcraper as a Python encoded pipeline that relies on programmatic access protocols to link molecular data from GenBank, to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+We present Physcraper, a Python pipeline that relies on programmatic access protocols to link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
-It focuses on database interoperability, by automatizing taxonomic name matching across biological databases, so users can perform any downstream analysis straightforwardly, as well as placing new data from virtually any biological database in an evolutionary context.
+Physcraper focuses on database interoperability, by integrating taxonomic name matching across biological databases.
+This integration makes downstream analyses straightforward, and allows users to place new data from a range biological database in an evolutionary context.
 

### 2020-07-30 1-abstract.Rmd

@@
 
-1. Phylogenies are a key part of research in many areas of biology. Tools that automatize
+1. Phylogenies are a key part of research in many areas of biology. Tools that automate
 some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
@@
 
-1. Physcraper is a command-line Python program that automatizes the update of published
+1. Physcraper is a command-line Python program that automates the update of published
 phylogenies by enriching underlying gene alignments with public DNA sequence data.

### 2020-07-30 2-body.Rmd

@@
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
 Linking these alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.

### 2020-07-30 render2-mee.tex

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[]{article}
+\documentclass[
+]{article}
 \usepackage{lmodern}
 \usepackage{setspace}
-\setstretch{2}
 \usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
   pdftitle={Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin = 1in]{geometry}
 \usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
 \usepackage{color}
@@
 
+\setstretch{2}
 \textbf{1. Luna L. Sanchez Reyes}
@@
 
+School of Natural Sciences, University of California, Merced
 Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
@@
 
-\textbf{Word count}: 3722
-
 \textbf{Manuscript prepared for submission to Methods in Ecology and Evolution}
@@
   We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS.
+Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and
@@
 Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
-Public biological databases provide an amazing resource for evolutionary estimation, but a large portion of sequence data available has never been incorporated into any phylogenetic estimate.
+Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Extending existing phylogenetic estimates with new DNA sequence data, geographical location, fossil data, etc., in a reproducible and continuous manner is possible by automatizing connections between biological databases.
+Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-GenBank, the USA National Center for Biodiversity Information (NCBI) molecular database, release number 159 (April 15, 2007) hosts 72 million DNA sequences that are estimated to have the potential to resolve evolutionary relationships of most of its 241 000 distinct taxa (about 98.05\% of taxa in the NCBI taxonomy release 159; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Yet, current public estimates of phylogenetic relationships are available for only about 100 000 taxa (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Hinchliff \emph{et al.} \protect\hyperlink{ref-hinchliff2015synthesis}{2015}; OpenTreeOfLife \emph{et al.} \protect\hyperlink{ref-opentreeoflife2019synth}{2019}), representing less than half of the taxonomic diversity with phylogenetically informative sequence data available in GenBank more than a decade ago.
+The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
+To standardize taxonomic names, a unified taxonomy is needed.
+The main aim of the Open Tree of Life (OpenTree from now on) project
+is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
+Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}, providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
-Some of the discrepancy between data availability and phylogenetic estimates can be explained by the many phylogenies that are generated and published and not shared publicly in an accesible way (Drew \emph{et al.} \protect\hyperlink{ref-drew2013lost}{2013}; Magee \emph{et al.} \protect\hyperlink{ref-magee2014dawn}{2014}; McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2018bioessay}{2017}).
-However, there is also a lag between the amount of new DNA data generated and the analysis of these data in a phylogenetic context.
-Connecting new sequence data with existing estimates of evolutionary relationships is a challenge.
-WHY IS IT A CHALLENGE
+Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
+Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
+Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
+Moreover, some types of genomic data are not suitable for phylogenetic inference at any scale.
+For example, phylogenetic distance can cause allelic dropout at deeper divergences with RADSeq data sets (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016}).
+In contrast, decades of single locus sequencing have generated homologous DNA data
+that can be used for phylogenetic reconstruction at many scales.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time (about 98.05\% of taxa; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible method.
+Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+Still, there is a large tradition in phylogenetics of building molecular character matrices of few, thoughtfully curated markers, as manual curation of alignments often improves phylogenetic reconstructions (Fragoso-Martı́nez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-We address this gap by extending existing phylogenetic estimates with publicly available sequence data.
-By using a starting tree and single locus alignment, Physcraper, takes advantage of existing research, and extends trees using loci that taxon specialists have assessed and deemed appropriate for the phylogenetic scope.
-The sequences added in the search are limited to a user specified taxon or monophyletic group, or within the taxonomic scope of the in-group of the starting tree.
-These automated trees can provide a quick inference of potential relationships, of problems in the taxonomic assignments of sequences, and flag areas of potential systematic interest.
+A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
+The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
 
-Physcraper leverages phylogentic data stored in Open Tree of Life and in TreeBase.
-The Open Tree of Life (\url{https://opentreeoflife.github.io/}) is a project that unites phylogenetic inferences and taxonomy to provide a synthetic estimate of species relationships across the entire tree of life.
-Open Tree of Life aims to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently the tree comprises 2.3 million tips, of which around 90,000 of those taxa are represented by phylogenetic estimates - the rest are placed in the tree based on their taxonomic names.
+We propose Physcraper as a Python encoded pipeline that relies on programmatic access protocols to link molecular data from GenBank, to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
+It focuses on database interoperability, by automatizing taxonomic name matching across biological databases, so users can perform any downstream analysis straightforwardly, as well as placing new data from virtually any biological database in an evolutionary context.
 
-The Open Tree of Life data store, \href{https://academic.oup.com/bioinformatics/article/31/17/2794/183373}{Phylesystem}, contains more than 4,500 phylogenetic trees from published studies.
-The tips in these trees are mapped a unified taxonomy, which makes these data searchable in a phylogenetically explicit way.
-This provides a resource for finding existing estimates of phylogenetic relationships,
-and assessing which regions of the tree of life which are lacking available phylogenetic estimates.
-
-By linking molecular data, available from databases such as the GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), to alignments and phylogenies, available in
-the TreeBASE repository (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) and the Open Tree of Life (OpenTree) datastore (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), we can place new biological data in an evolutionary context.
-
-While the focus on single locus and gene sequence alignments could appear backwards-looking, in the age of genomics, single locus data has a lot to offer phylogenetics.
-One major challenge of inferring phylogenies from genome scale data is inference of homology, and acquiring homologous data across divergent species.
-Different research questions call for on different approaches to genomic sequencing, from whole genomes, to transcriptomes, to RadSeq and UCE's.
-This variety of approaches results in non-overlapping data sets across taxa.
-Even when the same sequencing approach is applied, such as RadSeq phylogenetic distance can cause allelic dropout at deeper divergences (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016})
-In contrast, single locus sequencing generates homologous data across large phylogenetic scales.
-
-Indeed, some systematics support a classic phylogenetics approach (few markers thoughtfully curated) over the genomics approach (a massive amount of DNA markers that will overcome potential errors in the alignment coming from a lack of human curation).
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships {[}Song \emph{et al.} (\protect\hyperlink{ref-song2012resolving}{2012}); ROJAS ET AL.
-bats paper, take citations from there{]}.
-It has also been suggested that manual curation of locus alignments produces better phylogenetic reconstructions and this has been demonstrated for genomic alignments (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
-
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data AND fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-This expert-curated alignments can be continuously enriched and updated by incorporating newly released data from public molecular databases.
-
-In leveraging existing homology statements in the form of alignments, this approach differs from existing approaches that automatize the assembly of DNA alignments from the GenBank database for phylogenetic reconstruction (``phylogenetic pipelines'') such as PHYLOTA (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-Physcraper shares a similar conceptual framework to Pumper (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo-carrasco_pumper:_2014}{2014}), but that software is not currently supported or developed (\emph{or runnable at all honestly\ldots{}})
-
-Data input availability:
-As of April 2014, the TreeBASE repository hosted about 8 200 curated alignments, providing information on evolutionary relationships of around 100 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
-This database provides an untapped source of valuable expert knowledge with the potential to update phylogenetic relationships in several different regions of the tree of life.
-
-The Phylesystem (OpenTree's datastore) (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}) automatically incorporates phylogenies from TreeBASE, and saves metadata linking the original tree to its corresponding alignment repository in TreeBASE.
-If there are multiple alignments, TreeBASE does not always indicate how they were used to generate the tree.
-This provides a loose means of linking the tree with the exact alignment that generated it.
-
-Often, linking data in an original alignment with its corresponding phylogeny has to be done by a human curator.
-Moreover, different data repositories follow different systems for taxon and study identification, posing a real challenge to automatically link data from across databases that belong to the same taxon and study.
-OpenTree's metadata system incorporates taxon identifiers from a variety of taxonomies and repositories, including the NCBI taxonomy, GBIF, etc., providing a way to automatically link data from different databases.
-
-Physcraper is a Python encoded pipeline that relies on the OpenTree metadata system to connect databases through unique taxon identifiers.
-main functionality is to connect phylogenies stored in the OpenTree phylesystem, with alignments from TreeBASE and newly released DNA data from GenBank, in order to update previously known phylogenetic relationships in a continuous manner.
-EXPLIAN WHY PYTHON IS BEST FOR THIS.
-By design, this approach focusses on interoperability. By automating taxonomic name matching across NCBI, OpenTree, GBIF and other data stores, users can easily perform downstream analyses.
-For example, it allows automatizing and standardizing the comparison of phylogenetic hypotheses with currently known relationships in the Open Tree of Life taxonomy and phylogenetic data store.
-
-We envision Physraper as a staple tool to make data connections across databases in a phylogenetic context for the advantage of phylogenetics and biology in general, as well as an effort towards fully reproducible workflows in phylogenetics.
-
-\hypertarget{how-does-physcraper-work}{%
-\section{How does Physcraper work?}\label{how-does-physcraper-work}}
+\hypertarget{the-physcraper-framework}{%
+\section{The Physcraper framework}\label{the-physcraper-framework}}
 
@@
 
-\caption{Physcraper general framework.}\label{fig:framework}
+\caption{The Physcraper software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
 \end{figure}
 
-The Physcraper software is fully described on its documentation website at \url{https://physcraper.readthedocs.io/en/latest/}, along with installation, tutorials, examples and function usage.
-The general Physcraper workflow is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
+The general Physcraper framework is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
 
-\hypertarget{the-input-a-tree-and-an-alignment}{%
-\subsection{The input: a tree and an alignment}\label{the-input-a-tree-and-an-alignment}}
+\hypertarget{the-inputs-a-tree-and-an-alignment}{%
+\subsection{The inputs: a tree and an alignment}\label{the-inputs-a-tree-and-an-alignment}}
 
 \begin{itemize}
-\tightlist
 \item
-  In order to take advantage of the OpenTree tools, the input tree needs to be stored in the
-  OpenTree phylesystem \href{https://github.com/opentreeoflife/phylesystem}{github.com/opentreeoflife/phylesystem}, or need to have the tip labels mapped to the unified OpenTree taxonomy via bulk taxonomic name resolution (\url{https://tree.opentreeoflife.org/curator/tnrs/}).
-  The main
-  reason for this is that trees in phylesystem have a set of user defined characteristics
-  that are essential for automatizing the phylogeny update process.
-The most relevant of these being the standardization of taxonomic names and the definition of ingroup and outgroup.
-Outgroup and ingroup taxa in the original tree are identified and tagged.
-This allows to automatically set the root for the updated tree on the final steps of the pipeline.
-  Currently, only trees connected to
-  a published study can be stored in the phylesystem (although there are plans to
-  allow storage of unpublished trees).
-  A user can choose from among the 1216 published trees supporting the resolved nodes of the synthetic tree in the OpenTree website (\href{https://tree.opentreeoflife.org/about/synthesis-release/v12.3}{See OpenTree's website about}).
-If the published tree you are interested in updating is not in there, you can upload it via OpenTree's curator tool (\href{https://tree.opentreeoflife.org/curator}{www.opentreeoflife.org/curator}.
-  If the user is not ready to make the input tree public, \texttt{physcraper} can use as input a tree, and aliognment, and a file of mappings between the tip labels and taxonomic names.
+  In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+  OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
+If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+\emph{Should we saythat it is based on Boyle's 2012 algorithm?
+or is it a TNRS algorithm independently developed by the OpenTree team?}
+  This step is referred to as taxon name mapping. If taxonomic names can't be mapped, their taxonomic information will be excluded from further analysis.
+  Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
+  that are essential for automatizing the phylogeny updating process.
+The most relevant of these is the taxonomic name mapping and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+Currently, only trees connected to a published study can be stored in the Phylesystem.
+  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on
+  TreeBASE.
 \item
-  The input alignment should be a single locus alignment that was used to generate the tree. For public data, the original
-  alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
-  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or the journal were the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper can download it directly,
+  The input alignment should be a single locus alignment that was used to generate the tree.
+Alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
+  DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or a data repository associated with the journal where the tree was originally published.
+  If the alignment is stored in TreeBase, Physcraper downloads it directly,
   either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
@@
 \item
-  A taxon name matching step is performed to verify that all taxon names on the tips
-  of the tree are in the DNA character matrix and vice versa.
+  Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
+Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
+See next section.
 \item
-  A ``.csv'' file with the summary of taxon name matching is produced for the user.
+  A summary ``csv'' file with results from the taxon name standardization and pruning is produced for the user.
 \item
-  Unmatched taxon names are dropped from both the tree and alignment.
-  Technically, just one matching name is needed to perform the searches. Please, see next section.
-\item
-  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
+  The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
 \end{itemize}
@@
 \item
-  Technically, any DNA molecular database can be used to search for new sequences.
-  By default we rely on the GenBank database.
-  The new sequence search can be performed on the remote database or in a local database.
+  Physcraper uses the GenBank DNA database as source to search for new sequences.
+The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
+Detailed instructions to setup a local database are provided on Physcraper's software documentation.
 \item
-  The next step is to identify the search taxon within the reference taxonomy.
-The search taxon will be used to constraint the DNA sequence search on the nucleotide database within that taxonomic group.
-Because we are using the NCBI nucleotide database, by default the reference taxonomy is the NCBI taxonomy.
-The search taxon can be determined by the user.
-If none is provided, then the search taxon is identified as the Most Recent Common Ancestor (MRCA) of the
-  matched taxa belonging to the ingroup in the OPenTree synthetic tree, that is also a named clade in the reference taxonomy. FIGURE RECOMMENDED. This is known as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred in the literature as the Least Inclusive Common Ancestral Taxon - LICA) CITATION NEEDED.
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the reference taxonomy.
-  To automatically identify the MRCAT of a group of taxon names, we make use of the OpenTree taxonomic tool \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{Taxonomy-API-v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+  The next step is to identify a ``search taxon'' to constrain the sequence search on the GenBank database within that taxonomic group.
+  The search taxon can be chosen by the user from the NCBI taxonomy.
+  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the ``ingroup'' (Fig. \ref{fig:framework}).
+  The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
+  This is known in the OpenTree as the Most Recent Common
+  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA).
+  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
+  To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{taxonomic tool v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
 
@@
   database, and the sequences on the checked alignment.
-  The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local-database sequence searches.
-  For remote-database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function in the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternate BLAST URL.
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST without NCBI's required wait times.
+  The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local database sequence searches.
+  For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
+This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
 \item
-  A pairwise BLAST search is performed. This means that each sequence
-  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
-  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
-  along with their corresponding identification numbers (accession numbers in the case of the GenBank database).
-This information will be used later to store the whole sequences in a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
+  A constrained BLAST search is performed, in which each sequence
+  in the alignment is BLASTed once against all database DNA sequences belonging to the search
+  taxon. All results from each BLAST run are stored, and sequences with match scores better than the e-value cutoff (default to 0.00001) are saved
+  along with their corresponding metadata, i.e., their GenBank accession number.
+  The full sequence for each match is downloaded from NCBI into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
 \item
-  Matched sequences will be discarded if they fall below a default e-value of 0.00001, and outside a default minimum and maximum length of 80\% and 120\%, respectively, of the average length (gaps dropped) of sequences in the checked alignment .
-  These parameters can be configured for each run.
-  This filtering guarantees the exclusion of whole genome sequences. EXPLAIN WHY THIS IS IMPORTANT.
-  All accepted sequences are assigned an internal identifier, and are further filtered.
+  BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80\% and 120\%, respectively).
+  This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
+  The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
 \item
-  New sequences that are identical to existing sequences, or to subsets of an existing sequence are discarded, unless they reperesent a different taxon than the existing sequence.
-  Longer sequences belonging to the
-  same taxon as the original sequence will be considered further for analysis.
+  All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
+\item
+  New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the unified taxonomy, or they are longer than the sequence in the input alignment.
 \item
@@
 \item
-  Reverse complement sequences are identified and translated.
+  All BLAST and filtering parameters can be customized by the user.
 \item
-  Users can choose to perform cycles of sequence similarity search iteratively, by blasting the newly found sequences until no new sequences are found.
-By default only one BLAST cycle is performed and only sequences in the checked alignment are blasted.
+  Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
 \item
-  Accepted sequences are then downloaded in full, and stored as a local database in a directory that is globally accessible (default to ``physcraper/taxonomy'' folder), so they are accessible for further runs.
+  Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
 \item
-  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+  New sequences passing all filtering steps are added to the ``csv'' taxon summary file.
+\item
+  A ``fasta'' file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 \end{itemize}
 
-\hypertarget{dna-sequence-alignment}{%
-\subsection{DNA sequence alignment}\label{dna-sequence-alignment}}
+\hypertarget{new-dna-sequence-alignment}{%
+\subsection{New DNA sequence alignment}\label{new-dna-sequence-alignment}}
 
@@
 \item
-  The software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used by default to perform sequence alignments. It has to be installed by the user, previous to running physcraper.
+  By default, Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
+Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
 \item
-  First, all new sequences are aligned using default MUSCLE options.
+  The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
 \item
-  Then, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align new sequences. This ensures that the final alignment
+  Second, a MUSCLE profile alignment is performed, in which the original alignment
+  is used as a template to align the new sequences. This ensures that the final alignment
   follows the homology criteria established by the original alignment.
 \item
-  The final alignment is not further processed automatically. So, we encourage users
-  to check it either by eye and perform manual refinement or using any of the many
-  tools for alignment processing, to eliminate columns with no information.
+  The final alignment is not further processed by Physcraper.
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
 \item
-  Users may also use Physcraper to only gather matched sequences, and apply their own preferred alignment and phylogenetic inference methods.
+  While curating the alignment is a critical step, it is not a reproducible one.
+The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
+A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
+\item
+  Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
 \end{itemize}
@@
 \item
-  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
+  Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
 \item
-  Tips on all trees generated by Physcraper are defined by a taxon name space, capturing the NCBI accession information, as well as the taxon identifiers, allowing the user perform comparisons and conflict analyses.
+  Tips on all trees generated by Physcraper are defined by a taxon ``name space''.
+The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
 \item
-  Robinson Foulds weighted and unweighted metrics between the tips in the input tree and those tips in the updated tree are calculated with Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+  Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 \item
-  Finally a conflict analysis is performed.
-This is basically a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
+  Second, a conflict analysis is performed.
+This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
 This is performed with OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroups so trees are accurately rooted.
+The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
 \end{itemize}
@@
 A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species. A review of literature reveals that there are three published phylogenetic trees showing relationships within the hollies.
-The first one has been made available in TreeBASE as well as in the OpenTree phylesystem and is part of the synthetic tree. It samples 48 species.
-The second tree has not been made available anywhere, not even in the supplementary data of the original publication.
-The most recent one has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet, with 200 species.
-However, it has not been added to the syntehtic tree yet.
-This also makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree will produce something very similar to the newest tree.
+The genus encompasses between 400-600 living species.
+A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
+Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made tree and alignment data available in TreeBASE.
+The tree sampling 48 species was integrated to the OpenTree Phylesystem and is part of OpenTree's synthetic tree.
+The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) has been made available in the OpenTree Phylesystem and in the DRYAD repository.
+It is the best sampled yet for the genus, with 200 species.
+However, it has not been added to OpenTree's synthtetic tree yet.
+This makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
+Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
 
@@
 a field that has relied on processes which are somewhat artisanal.
-In here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging on the knowledge and intuition of the expert
+Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
 community to build up our phylogenetic knowledge, we can use not only data accumulated in
 molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+
 While not generating full phylogenies \emph{de novo}, Physcraper is still capable of generating new phylogenetic knowledge.
-It can also be combined with data from repositories other than molecular data. For example geographic locations (GBIF), fossils (PBDB), etc.
+Moreover, it can combine phylogenies with data from repositories other than molecular data.
+For example geographic locations (using GBIF ids), fossils (using PBDB ids), etc.
+\emph{from Robert: I think you can sell the program more here.
+Why is it better than the other methods?
+You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages}
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly
+Physcraper has the potential to be applied for the advantage of the field to rapidly \emph{HOW FAST IS ``RAPID'' mention it in results and then here again}
 place newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
@@
 
+EJM: Conceived study, wrote most of the code, documentation and tests.
+MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
+LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
+All authors contributed to the manuscript.
+
 \hypertarget{data-avilability}{%
 \section{Data Avilability}\label{data-avilability}}
+
+Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
+Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
+Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
+This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}
 
@@
 
-\leavevmode\hypertarget{ref-drew2013lost}{}%
-Drew, B.T., Gazis, R., Cabezas, P., Swithers, K.S., Deng, J., Rodriguez, R., Katz, L.A., Crandall, K.A., Hibbett, D.S.
-\& Soltis, D.E.
-(2013).
-Lost branches on the tree of life.
-\emph{PLoS biology}, \textbf{11}.
+\leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
+Cuénoud, P., Martinez, M.A.
+del P., Loizeay, P.-A., Spichiger, R., Andrews, S.
+\& Manen, J.-F.
+(2000).
+Molecular phylogeny and biogeography of the genus ilex l.(Aquifoliaceae).
+\emph{Annals of Botany}, \textbf{85}, 111--122.
 
@@
 \leavevmode\hypertarget{ref-fragoso2017pilot}{}%
-Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
+Fragoso-Martı́nez, I., Salazar, G.A., Martı́nez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
 \& Mendoza, C.G.
@@
 
-\leavevmode\hypertarget{ref-hinchliff2015synthesis}{}%
-Hinchliff, C.E., Smith, S.A., Allman, J.F., Burleigh, J.G., Chaudhary, R., Coghill, L.M., Crandall, K.A., Deng, J., Drew, B.T., Gazis, R.
-\& others.
-(2015).
-Synthesis of phylogeny and taxonomy into a comprehensive tree of life.
-\emph{Proceedings of the National Academy of Sciences}, \textbf{112}, 12764--12769.
+\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
+Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
 
-\leavevmode\hypertarget{ref-izquierdo-carrasco_pumper:_2014}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A.
-\& Stamatakis, A.
-(2014).
-PUmPER: Phylogenies Updated Perpetually.
-\emph{Bioinformatics}, \textbf{30}, 1476--7.
-Retrieved February 1, 2014, from \url{http://bioinformatics.oxfordjournals.org/content/early/2014/01/28/bioinformatics.btu053}
-
-\leavevmode\hypertarget{ref-magee2014dawn}{}%
-Magee, A.F., May, M.R. \& Moore, B.R. (2014). The dawn of open access to phylogenetic data. \emph{PLoS One}, \textbf{9}.
-
-\leavevmode\hypertarget{ref-mctavish2018bioessay}{}%
-McTavish, E.J., Drew, B.T., Redelings, B.
-\& Cranston, K.A.
-(2017).
-How and Why to Build a Unified Tree of Life.
-\emph{BioEssays}, \textbf{39}.
-Retrieved April 10, 2018, from \url{https://onlinelibrary.wiley.com/doi/abs/10.1002/bies.201700114}
+\leavevmode\hypertarget{ref-manen2010history}{}%
+Manen, J.-F., Barriera, G., Loizeau, P.-A.
+\& Naciri, Y.
+(2010).
+The history of extant ilex species (aquifoliaceae): Evidence of hybridization within a miocene radiation.
+\emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
 
@@
 Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
-
-\leavevmode\hypertarget{ref-opentreeoflife2019synth}{}%
-OpenTreeOfLife, Redelings, B., Reyes, L.L.S., Cranston, K.A., Allman, J., Holder, M.T.
-\& McTavish, E.J.
-(2019).
-Open tree of life synthetic tree.
-Retrieved from \url{https://doi.org/10.5281/zenodo.3937742}
 
@@
 San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
+
+\leavevmode\hypertarget{ref-secretariat2017gbif}{}%
+Secretariat, G. (2017). GBIF backbone taxonomy. \emph{Checklist Dataset {[}cited 2017 Nov 14{]}. doi}, \textbf{10}.
+
+\leavevmode\hypertarget{ref-selbach2009new}{}%
+Selbach-Schnadelbach, A., Cavalli, S.S., Manen, J.-F., Coelho, G.C.
+\& De Souza-Chies, T.T.
+(2009).
+New information for ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (aquifoliaceae).
+\emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.
+
+\leavevmode\hypertarget{ref-setoguchi2000intersectional}{}%
+Setoguchi, H.
+\& Watanabe, I.
+(2000).
+Intersectional gene flow between insular endemics of ilex (aquifoliaceae) on the bonin islands and the ryukyu islands.
+\emph{American Journal of Botany}, \textbf{87}, 793--810.
 

### 2020-07-30 2-body.Rmd
ejm comments intro

@@
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
@@
 Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
+Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-However, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
-Having a variety of approaches can result in largely non-overlapping genomic data sets across taxa.
-Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]."
-While targeted capture approaches are ameliorating the problem of non-overlapping genomic data sets [@andermann2020guide; @jones2016targeted], in the meantime decades of single locus sequencing have already generated homologous DNA data that can be used for phylogenetic reconstruction at many scales.
+Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+<!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 
@@
 <!-- *maybe a box of all pipelines?* -->
-However, building molecular character matrices of few, thoughtfully curated markers, including manual curation of alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
+However, building molecular character matrices of few, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
@@
 The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
-Linking these alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
+Linking published alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
+The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
 <!-- However, if there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
@@
 
-We present Physcraper, a Python pipeline that relies on programmatic access protocols to link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
-Physcraper focuses on database interoperability, by integrating taxonomic name matching across biological databases.
+Physcraper is coded as a Python pipeline that focuses on database interoperability, by integrating taxonomic name matching across biological databases.
 This integration makes downstream analyses straightforward, and allows users to place new data from a range biological database in an evolutionary context.
@@
 If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-*Should we saythat it is based on Boyle's 2012 algorithm?
+*Should we say that it is based on Boyle's 2012 algorithm?
 or is it a TNRS algorithm independently developed by the OpenTree team?*

### 2020-07-30 docs/submission-mee-2020-07-30.tex
pdf

--- 2020-07-28 docs/submission-mee-2020-07-27.tex
@@
 
+School of Natural Sciences, University of California, Merced
+
 Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
@@
 
-\textbf{Word count}: 3646
+\textbf{Word count}: 3633
 
@@
 \item
-  Phylogenies are a key part of research in many areas of biology. Tools that automatize
+  Phylogenies are a key part of research in many areas of biology. Tools that automate
   some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
@@
 \item
-  Physcraper is a command-line Python program that automatizes the update of published
+  Physcraper is a command-line Python program that automates the update of published
   phylogenies by enriching underlying gene alignments with public DNA sequence data.
@@
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, fossil data, etc., in a reproducible and continuous manner is possible by automatizing connections between biological databases.
-Here, we propose Physcraper, a tool that uses existing phylogenetic research and data from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
+Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
-The prevalence of taxonomic idiosincracies across databases represent a big challenge to automatically connect data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified taxonomy is needed.
+The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
+To standardize taxonomic names, a unified system is needed.
 The main aim of the Open Tree of Life (OpenTree from now on) project
@@
 Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multi-gene data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-While genomics have allowed us to obtain large quantities of molecular data from one organism, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing (RADSeq), single nucleotide polymorphisms (SNPs), microsatellites, and ultra-conserved elements (UCEs).
-Having a variety of approaches has resulted in largely non-overlapping genomic data sets across taxa.
-Moreover, some types of genomic data are not suitable for phylogenetic inference at any scale.
-For example, phylogenetic distance can cause allelic dropout at deeper divergences with RADSeq data sets (Eaton \emph{et al.} \protect\hyperlink{ref-eaton2016misconceptions}{2016}).
-In contrast, decades of single locus sequencing have generated homologous DNA data
-that can be used for phylogenetic reconstruction at many scales.
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time (about 98.05\% of taxa; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible method.
-Computer pipelines that make DNA sequence search faster and more efficient, as well as more reproducible have been utilized to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-Still, there is a large tradition in phylogenetics of building molecular character matrices of few, thoughtfully curated markers, as manual curation of alignments often improves phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+Genomics has, and will continue to, revolutionize phylogenetic inference.
+Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
+These sequences were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time (about 98.05\% of taxa; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible approach.
+Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+However, building molecular character matrices of few, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accesible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing an untapped source of valuable expert knowledge with the potential to continuously enrich and update phylogenetic relationships in several different regions of the tree of life, by incorporating data from molecular databases that has not been included in any public phylogenetic estimate.
-The phylogenies associated to all these alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking these trees to their corresponding alignment repository in TreeBASE is available.
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a source of valuable expert knowledge.
+Linking published alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
+The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
 
-We propose Physcraper as a Python encoded pipeline that relies on programmatic access protocols to link molecular data from GenBank, to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to automatically update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
-It focuses on database interoperability, by automatizing taxonomic name matching across biological databases, so users can perform any downstream analysis straightforwardly, as well as placing new data from virtually any biological database in an evolutionary context.
+Physcraper is coded as a Python pipeline that focuses on database interoperability, by integrating taxonomic name matching across biological databases.
+This integration makes downstream analyses straightforward, and allows users to place new data from a range biological database in an evolutionary context.
 
@@
 If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-\emph{Should we saythat it is based on Boyle's 2012 algorithm?
+\emph{Should we say that it is based on Boyle's 2012 algorithm?
 or is it a TNRS algorithm independently developed by the OpenTree team?}
@@
 
+\leavevmode\hypertarget{ref-andermann2020guide}{}%
+Andermann, T., Torres Jiménez, M.F., Matos-Maraví, P., Batista, R., Blanco-Pastor, J.L., Gustafsson, A.L.S., Kistler, L., Liberal, I.M., Oxelman, B., Bacon, C.D.
+\& Antonelli, A.
+(2020).
+A Guide to Carrying Out a Phylogenomic Target Sequence Capture Project.
+\emph{Frontiers in Genetics}, \textbf{10}.
+Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7047930/}
+
 \leavevmode\hypertarget{ref-antonelli2017toward}{}%
@@
 
-\leavevmode\hypertarget{ref-eaton2016misconceptions}{}%
-Eaton, D.A.R., Spriggs, E.L., Park, B.
-\& Donoghue, M.J.
-(2016).
-Misconceptions on Missing Data in RAD-seq Phylogenetics with a Deep-scale Example from Flowering Plants.
-\emph{Systematic Biology}, \textbf{66}, 399--412.
-Retrieved from \url{https://doi.org/10.1093/sysbio/syw092}
-
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
@@
 Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
+
+\leavevmode\hypertarget{ref-jones2016targeted}{}%
+Jones, M.R.
+\& Good, J.M.
+(2016).
+TARGETED CAPTURE IN EVOLUTIONARY AND ECOLOGICAL GENOMICS.
+\emph{Molecular ecology}, \textbf{25}, 185--202.
+Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4823023/}
 

### 2020-08-13 2-body.Rmd
updating results

@@
 Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available elsewhere.
+A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available in the documentation website.
 
@@
 Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
+
+<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
+wget -O docs/figs/ilex-plot-tips-and-branches-1.png https://raw.githubusercontent.com/McTavishLab/physcraperex/master/docs/articles/ilex_files/figure-html/plot-tips-and-branches-1.png
+``` -->
+
+```{r ilex-updated-tree, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.
+In this example, Physcraper updated an alignment of 48 samples of the genus Ilex from Gottlieb et al.
+2005 (black tips) and added 222 new samples from the GenBank database (red tips), corresponding to 78 new taxa, aligned the new seqeunces to the original alignment using MUSCLE and performed the phylogenetic reconstruction using RAxML.", out.width = '100%', fig.retina= 2}
+
+knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
+
+```
+
+The Physcraper updated tree has 57 original tips, representing 51 unique OTT taxa, and 222 new tips representing 78 new OTT taxa added to the original tree.
+
+It has THIS many resolved nodes, and THIS MANY nodes with bootsrap values above 95%
+
+The original tree has THIS MANY resolved nodes
+
+The alignments have THIS MANY columns and THIS AMOUNT of missing data.
 
@@
 Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838, and ABI No. 1759846.
-Compute time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
+Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
 ACI-1429783.

### 2020-08-13 docs/submission-mee-2020-08-12.tex
updating results

--- 2020-07-30 docs/submission-mee-2020-07-30.tex
@@
 
-\textbf{Word count}: 3633
+\textbf{Word count}: 3698
 
@@
 Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{docs/figs/ilex-plot-tips-and-branches-1} \caption{Physcraper outpus tree obtained by updating the alignment from Gottlieb *et al.* 2005}\label{fig:ilex-updated-tree}
+\end{figure}
+
+The Physcraper updated tree has 57 original tips, representing 51 unique OTT taxa, and 222 new tips representing 78 new OTT taxa added to the original tree.
+
+It has THIS many resolved nodes, and THIS MANY nodes with bootsrap values above 95\%
+
+The original tree has THIS MANY resolved nodes
+
+The alignments have THIS MANY columns and THIS AMOUNT of missing data.
 
@@
 Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838, and ABI No.~1759846.
-Compute time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
+Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
 

### 2020-08-13 docs/submission-mee-2020-08-12.tex
updating pdf

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-plot-tips-and-branches-1} \caption{Physcraper outpus tree obtained by updating the alignment from Gottlieb *et al.* 2005}\label{fig:ilex-updated-tree}
+\includegraphics[width=1\linewidth]{docs/figs/ilex-plot-tips-and-branches-1} \caption{A phylogenetic tree is the main result of a Physcraper analysis.
+In this example, Physcraper updated an alignment of 48 samples of the genus Ilex from Gottlieb et al.
+2005 (black tips) and added 222 new samples from the GenBank database (red tips), corresponding to 78 new taxa, aligned the new seqeunces to the original alignment using MUSCLE and performed the phylogenetic reconstruction using RAxML.}\label{fig:ilex-updated-tree}
 \end{figure}

### 2020-08-13 2-body.Rmd
a round to intro and methods

@@
 Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time [about `r round(236023/240708, digits=4)*100`% of taxa; @sanderson2008phylota].
+These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
 Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible approach.
@@
 <!-- *maybe a box of all pipelines?* -->
-However, building molecular character matrices of few, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
+However, even during phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
@@
 The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
-Linking published alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
+Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
 The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
@@
 Physcraper relies on programmatic access protocols to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various general OpenTree programmatic tools for custom comparison of phylogenetic hypotheses.
-Physcraper is coded as a Python pipeline that focuses on database interoperability, by integrating taxonomic name matching across biological databases.
-This integration makes downstream analyses straightforward, and allows users to place new data from a range biological database in an evolutionary context.
+Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
+Physcraper is coded as a Python pipeline that focuses on data interoperability, by integrating taxonomic name matching across biological databases.
+This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, making a variety of downstream analyses straightforward.
 
@@
 
-```{r framework, echo=FALSE, fig.cap="The Physcraper software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, tutorials, examples and function usage documentation.", out.width = '85%', fig.retina= 2, fig.align = "center"}
+```{r framework, echo=FALSE, fig.cap="The Physcraper framework consists of 4 steps (see text).
+The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.", out.width = '85%', fig.retina= 2, fig.align = "center"}
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
@@
 
-The general Physcraper framework is shown in Figure \@ref(fig:framework). Next, we will describe the technical details of each step of the workflow.
+The general Physcraper framework is depicted in Figure \@ref(fig:framework).
+Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
+Next, we will describe technical details for each step.
 
@@
   OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
+  Currently, only trees connected to a published study can be stored in the Phylesystem.
+  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
+  If the user is not ready to make the input tree public, tree tip labels must be first standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
 *Should we say that it is based on Boyle's 2012 algorithm?
 or is it a TNRS algorithm independently developed by the OpenTree team?*
-  This step is referred to as taxon name mapping. If taxonomic names can't be mapped, their taxonomic information will be excluded from further analysis.
+  This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
+  Physcraper saves a summary "csv" file with results from the taxon name standardization for advantage of the user.
+  If taxon names can't be mapped, their taxonomic information can't be used in comparison analysis. They will still be used in the sequence search and phylogenetic reconstruction steps.
   Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
   that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the taxonomic name mapping and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
-Currently, only trees connected to a published study can be stored in the Phylesystem.
+The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
   <!-- Discussion: (although there are plans to allow storage of unpublished trees). -->
-  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on
-  TreeBASE.
+
   <!--
@@
 See next section.
-- A summary "csv" file with results from the taxon name standardization and pruning is produced for the user.
 - The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
@@
 -->
-
 - The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
@@
 
-## Tree reconstruction and comparison
+## Tree reconstruction
 
@@
 
-# Examples
+# Example: The hollies
 
-We will illustrate the utility of Physcraper in here with two use-case scenarios.
-One in which the user is interested in a particular group.
-Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available in the documentation website.
+We will illustrate the utility of Physcraper in here with one use-case scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, encompassing between 400-600 living species, is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
+A tutorial as well as illustrated examples of commands implementation on every step of the Physcraper workflow are available in the documentation website.
 
-## The hollies
-
-A student is interested in the genus *Ilex*, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species.
 A review of literature (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].

### 2020-08-13 docs/submission-mee-2020-08-13.tex
a round to intro and methods

--- 2020-08-13 docs/submission-mee-2020-08-12.tex
@@
 
-\textbf{Word count}: 3698
+\textbf{Word count}: 3784
 
@@
 Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most of the 241, 000 distinct taxa in the NCBI taxonomy at the time (about 98.05\% of taxa; Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible approach.
 Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, building molecular character matrices of few, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+However, even during phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
@@
 The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a source of valuable expert knowledge.
-Linking published alignments with data from molecular databases that has not yet been included in any public phylogenetic estimate, has the potential to enrich and update phylogenetic relationships in many regions of the tree of life.
+Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
 The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
@@
 
-\caption{The Physcraper software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, tutorials, examples and function usage documentation.}\label{fig:framework}
+\caption{The Physcraper framework consists of 4 steps (see text).
+The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.}\label{fig:framework}
 \end{figure}
 
-The general Physcraper framework is shown in Figure \ref{fig:framework}. Next, we will describe the technical details of each step of the workflow.
+The general Physcraper framework is depicted in Figure \ref{fig:framework}.
+Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
+Next, we will describe technical details for each step.
 
@@
   OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
-If the user is not ready to make the input tree public, tree tip labels must be standardized to the unified OpenTree taxonomy using the bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+  Currently, only trees connected to a published study can be stored in the Phylesystem.
+  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
+  If the user is not ready to make the input tree public, tree tip labels must be first standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
 \emph{Should we say that it is based on Boyle's 2012 algorithm?
 or is it a TNRS algorithm independently developed by the OpenTree team?}
-  This step is referred to as taxon name mapping. If taxonomic names can't be mapped, their taxonomic information will be excluded from further analysis.
+  This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
+  Physcraper saves a summary ``csv'' file with results from the taxon name standardization for advantage of the user.
+  If taxon names can't be mapped, their taxonomic information can't be used in comparison analysis. They will still be used in the sequence search and phylogenetic reconstruction steps.
   Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
   that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the taxonomic name mapping and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
-Currently, only trees connected to a published study can be stored in the Phylesystem.
-  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on
-  TreeBASE.
+The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
 \item
@@
 See next section.
-\item
-  A summary ``csv'' file with results from the taxon name standardization and pruning is produced for the user.
 \item
@@
 
-\hypertarget{tree-reconstruction-and-comparison}{%
-\subsection{Tree reconstruction and comparison}\label{tree-reconstruction-and-comparison}}
+\hypertarget{tree-reconstruction}{%
+\subsection{Tree reconstruction}\label{tree-reconstruction}}
 
@@
 
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+\hypertarget{example-the-hollies}{%
+\section{Example: The hollies}\label{example-the-hollies}}
 
-We will illustrate the utility of Physcraper in here with two use-case scenarios.
-One in which the user is interested in a particular group.
-Another one in which the user is interested in a particular phylogeny.
-A tutorial as well as illustrated examples of commands for every step needed to perform a Physcraper analysis are available in the documentation website.
+We will illustrate the utility of Physcraper in here with one use-case scenario in which a user is interested in a particular group, e.g., the genus \emph{Ilex}, encompassing between 400-600 living species, is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as ``hollies''.
+A tutorial as well as illustrated examples of commands implementation on every step of the Physcraper workflow are available in the documentation website.
 
-\hypertarget{the-hollies}{%
-\subsection{The hollies}\label{the-hollies}}
-
-A student is interested in the genus \emph{Ilex}, the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-The genus encompasses between 400-600 living species.
 A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).

### 2020-08-14 2-body.Rmd
cleaning up results

@@
 
-We will illustrate the utility of Physcraper in here with one use-case scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, encompassing between 400-600 living species, is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
-A tutorial as well as illustrated examples of commands implementation on every step of the Physcraper workflow are available in the documentation website.
+To illustrate the utility of Physcraper we will use a scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
 
-A review of literature (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
-@gottlieb2005molecular made tree and alignment data available in TreeBASE. The tree sampling 48 species was integrated to the OpenTree Phylesystem and is part of OpenTree's synthetic tree.
+An online literature review in June 2020 (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
+@gottlieb2005molecular made original tree and alignment data available in TreeBASE.
+The "Gottlieb2005" tree sampling 48 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).
 <!-- The second tree has not been made available anywhere, not even in the supplementary data of the original publication. -->
-The most recent *Ilex* tree from @yao2020phylogeny has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet for the genus, with 200 species.
-However, it has not been added to OpenTree's synthtetic tree yet.
-This makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
+The most recent *Ilex* tree from @yao2020phylogeny, has been made available in the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/view/ot_1984) and in the [DRYAD repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x).
+The "Yao2020" tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
+<!-- However, it has not been added to OpenTree's synthtetic tree yet. -->
+This makes it a great case to highlight the basic functionalities of Physcraper: we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
+Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
+A tutorial as well as illustrated examples of functions implementated on each step of the Physcraper framework are available in the documentation website.
 
@@
 
-## The Malvaceae
-
-A postdoc started working with a new reserach group.
-They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species, containing the relatives of cacao, cotton, durian and okra.
-A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
-However, the head of the research group wants to use a particular marker they believe to be the best one to be able to solve the relationships in the group.
-They have been working on the alignment for a long time and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
-
-<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
-wget "https://raw.githubusercontent.com/McTavishLab/physcraperex/master/data-raw/malvaceae-images/cotree-plot2-1.png"
-mv cotree-plot2-1.png docs/figs/cotree-plot2-1.png
-``` -->
-```{r malvaceae, echo=FALSE, fig.cap="Comparison of original tree and tree updated with Physcraper, family Malvaceae.", out.width = '85%', fig.retina= 2, fig.align = "center"}
-knitr::include_graphics("docs/figs/cotree-plot2-1.png")
-```
 \newpage

### 2020-08-14 docs/submission-mee-2020-08-13.tex
cleaning up results

@@
 
-\textbf{Word count}: 3784
+\textbf{Word count}: 3658
 
@@
 
-We will illustrate the utility of Physcraper in here with one use-case scenario in which a user is interested in a particular group, e.g., the genus \emph{Ilex}, encompassing between 400-600 living species, is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as ``hollies''.
-A tutorial as well as illustrated examples of commands implementation on every step of the Physcraper workflow are available in the documentation website.
+To illustrate the utility of Physcraper we will use a scenario in which a user is interested in a particular group, e.g., the genus \emph{Ilex}, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as ``hollies''.
 
-A review of literature (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
-Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made tree and alignment data available in TreeBASE.
-The tree sampling 48 species was integrated to the OpenTree Phylesystem and is part of OpenTree's synthetic tree.
-The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) has been made available in the OpenTree Phylesystem and in the DRYAD repository.
-It is the best sampled yet for the genus, with 200 species.
-However, it has not been added to OpenTree's synthtetic tree yet.
-This makes it a perfect case to test the basic functionalities of Physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database.
-Hence, we expect that updating the oldest tree should at least contain the same species sampled in the largest tree.
+An online literature review in June 2020 (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
+Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment data available in TreeBASE.
+The ``Gottlieb2005'' tree sampling 48 species was added to the \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and its information has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
+The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}), has been made available in the \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
+The ``Yao2020'' tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
+This makes it a great case to highlight the basic functionalities of Physcraper: we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
+Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
+A tutorial as well as illustrated examples of functions implementated on each step of the Physcraper framework are available in the documentation website.
 
@@
 
-\hypertarget{the-malvaceae}{%
-\subsection{The Malvaceae}\label{the-malvaceae}}
-
-A postdoc started working with a new reserach group.
-They are interested in solving relationships among lineages of the Malvaceae, a family of flowering plants with almost 6 000 known species, containing the relatives of cacao, cotton, durian and okra.
-A review of the literature shows them that there are many phylogenetic trees encompassing some of the linegaes in the group.
-However, the head of the research group wants to use a particular marker they believe to be the best one to be able to solve the relationships in the group.
-They have been working on the alignment for a long time and they want to incorporate new data into the hypothesis of homology that they have been curating and that they trust.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{docs/figs/cotree-plot2-1}
-
-}
-
-\caption{Comparison of original tree and tree updated with Physcraper, family Malvaceae.}\label{fig:malvaceae}
-\end{figure}
 \newpage

### 2020-08-17 2-body.Rmd
missing title
tiny edits

@@
 
-## Tree reconstruction
+## Tree reconstruction and comparison
 
@@
 
-To illustrate the utility of Physcraper we will use a scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
+To illustrate the utility of Physcraper we propose a scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
 
@@
 Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
-A tutorial as well as illustrated examples of functions implementated on each step of the Physcraper framework are available in the documentation website.
+A tutorial as well as illustrated examples of functions implemented on each step of the framework are available in Physcraper's documentation website.
 
@@
 ```{r ilex-updated-tree, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.
-In this example, Physcraper updated an alignment of 48 samples of the genus Ilex from Gottlieb et al.
-2005 (black tips) and added 222 new samples from the GenBank database (red tips), corresponding to 78 new taxa, aligned the new seqeunces to the original alignment using MUSCLE and performed the phylogenetic reconstruction using RAxML.", out.width = '100%', fig.retina= 2}
+The tree results from updating an input tree of 48 tips belonging to the genus Ilex published in Gottlieb et al.
+2005 (black tips) with 222 new sequences (red tips) corresponding to 78 distinct taxa that were not sampled in the original alignment.
+Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML as well as a conflict analysis", out.width = '100%', fig.retina= 2}
 
@@
 
-The Physcraper updated tree has 57 original tips, representing 51 unique OTT taxa, and 222 new tips representing 78 new OTT taxa added to the original tree.
+The Gottlieb2005 updated tree has 48 original tips, representing 45 distinct taxa, and 231 new tips representing 84 distinct taxa.
 
@@
 
-COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING
-WHY WE DID NOT MAKE A BENCHMARK COMPARISON
+*COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING*
+*WHY WE DID NOT MAKE A BENCHMARK COMPARISON?*
 
@@
 
-We thank members of the "short bar" Science and Engineering Building 1, University of California Merced, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
+We thank members of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on a first version of this manuscript.
 

### 2020-08-17 2-body.Rmd
resolved nodes

@@
 
-To illustrate the utility of Physcraper we propose a scenario in which a user is interested in a particular group, e.g., the genus *Ilex*, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as "hollies".
+To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus *Ilex*.
+Commonly known as "hollies", the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
 
@@
 The "Gottlieb2005" tree sampling 48 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).
-<!-- The second tree has not been made available anywhere, not even in the supplementary data of the original publication. -->
+
 The most recent *Ilex* tree from @yao2020phylogeny, has been made available in the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/view/ot_1984) and in the [DRYAD repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x).
@@
 <!-- However, it has not been added to OpenTree's synthtetic tree yet. -->
-This makes it a great case to highlight the basic functionalities of Physcraper: we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
+This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
+<!-- we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
 Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
-A tutorial as well as illustrated examples of functions implemented on each step of the framework are available in Physcraper's documentation website.
+-->
+A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
+Figure \@ref(fig:results) shows results from updating one alignment from the Gottlieb2005 tree.
+It contains all 41 distinct taxa from the original Gottlieb2005 tree plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
+The tree is 99% resolved, however, a large portion of branches are negligibly small (133 branches < 0.00001 substitution rate units), appearing as unresolved nodes in the figure.
 
@@
 
-```{r ilex-updated-tree, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.
-The tree results from updating an input tree of 48 tips belonging to the genus Ilex published in Gottlieb et al.
-2005 (black tips) with 222 new sequences (red tips) corresponding to 78 distinct taxa that were not sampled in the original alignment.
-Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML as well as a conflict analysis", out.width = '100%', fig.retina= 2}
+```{r results, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.
+The tree results from updating an alignment that was used to generate a tree of 48 tips belonging to the genus Ilex published in Gottlieb et al.
+2005 (black tips).
+Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
+Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML as well as a conflict analysis.", out.width = '100%', fig.retina= 2}
 
@@
 ```
-
-The Gottlieb2005 updated tree has 48 original tips, representing 45 distinct taxa, and 231 new tips representing 84 distinct taxa.
 

### 2020-08-18 2-body.Rmd
results fig

@@
 An online literature review in June 2020 (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
-@gottlieb2005molecular made original tree and alignment data available in TreeBASE.
-The "Gottlieb2005" tree sampling 48 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).
+@gottlieb2005molecular made original tree and alignment data available in [TreeBASE](https://treebase.org/treebase-web/search/study/summary.html?id=1091).
+The "Gottlieb2005" tree sampling 41 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).
 
@@
 A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
-Figure \@ref(fig:results) shows results from updating one alignment from the Gottlieb2005 tree.
-It contains all 41 distinct taxa from the original Gottlieb2005 tree plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
-The tree is 99% resolved, however, a large portion of branches are negligibly small (133 branches < 0.00001 substitution rate units), appearing as unresolved nodes in the figure.
+Figure \@ref(fig:results) shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from @gottlieb2005molecular.
+Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to finish BLAST and RAxML analyses.
+Bootstrap analyses took an additional 13hrs.
+The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
+The tree is 99% resolved, however, a large portion of branches are negligibly small (133 branches < 0.00001 substitution rate units), appearing as unresolved nodes (Fig.
+\@ref(fig:results)).
+<!-- ADD percentage of nodes with bootsrap values above 95%  and below 50%?-->
+For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
+<!-- has less new tips (134), but  -->
+and contributes phylogenetic data to 134 additional taxa within *Ilex*, from which
+`r 135-68` are also in updated Gottlieb2005.
+<!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
+<!-- RF DISTANCE INTERPRETATION -->
+<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE -->
 
@@
 
-```{r results, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.
-The tree results from updating an alignment that was used to generate a tree of 48 tips belonging to the genus Ilex published in Gottlieb et al.
-2005 (black tips).
-Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
-Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML as well as a conflict analysis.", out.width = '100%', fig.retina= 2}
+```{r results, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.", out.width = '100%', fig.retina= 2}
 
-knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
+# knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
+knitr::include_graphics("docs/figs/ilex-results.png")
 
@@
 
-It has THIS many resolved nodes, and THIS MANY nodes with bootsrap values above 95%
+\newpage
 
-The original tree has THIS MANY resolved nodes
-
-The alignments have THIS MANY columns and THIS AMOUNT of missing data.
-
-DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OpenTree
-RF DISTANCE INTERPRETATION
-HOW MUCH TIME THE BLAST RUN TOOK
-ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE
-
-FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OpenTree.
-
-\newpage
+Figure \@ref(fig:results) caption continued: The tree results from updating an alignment that was used to generate a tree of 48 tips of the genus Ilex, published in Gottlieb et al.
+2005 (black tips).
+Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
+Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML, as well as a conflict analysis using the OpenTree tools.
 

### 2020-08-18 docs/submission-mee-2020-08-17.tex
results fig

--- 2020-08-14 docs/submission-mee-2020-08-13.tex
@@
 
-\textbf{Word count}: 3658
+\textbf{Word count}: 3745
 
@@
 
-\hypertarget{tree-reconstruction}{%
-\subsection{Tree reconstruction}\label{tree-reconstruction}}
+\hypertarget{tree-reconstruction-and-comparison}{%
+\subsection{Tree reconstruction and comparison}\label{tree-reconstruction-and-comparison}}
 
@@
 
-To illustrate the utility of Physcraper we will use a scenario in which a user is interested in a particular group, e.g., the genus \emph{Ilex}, which encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants, commonly known as ``hollies''.
+To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus \emph{Ilex}.
+Commonly known as ``hollies'', the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
 
 An online literature review in June 2020 (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
-Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment data available in TreeBASE.
-The ``Gottlieb2005'' tree sampling 48 species was added to the \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and its information has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
+Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment data available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
+The ``Gottlieb2005'' tree sampling 41 species was added to the \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and its information has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
+
 The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}), has been made available in the \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
 The ``Yao2020'' tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-This makes it a great case to highlight the basic functionalities of Physcraper: we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
-Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
-A tutorial as well as illustrated examples of functions implementated on each step of the Physcraper framework are available in the documentation website.
+This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
+A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
+Figure \ref{fig:results} shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
+Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to finish BLAST and RAxML analyses.
+Bootstrap analyses took an additional 13hrs.
+The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
+The tree is 99\% resolved, however, a large portion of branches are negligibly small (133 branches \textless{} 0.00001 substitution rate units), appearing as unresolved nodes (Fig.
+\ref{fig:results}).
+For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
+and contributes phylogenetic data to 134 additional taxa within \emph{Ilex}, from which
+67 are also in updated Gottlieb2005.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-plot-tips-and-branches-1} \caption{A phylogenetic tree is the main result of a Physcraper analysis.
-In this example, Physcraper updated an alignment of 48 samples of the genus Ilex from Gottlieb et al.
-2005 (black tips) and added 222 new samples from the GenBank database (red tips), corresponding to 78 new taxa, aligned the new seqeunces to the original alignment using MUSCLE and performed the phylogenetic reconstruction using RAxML.}\label{fig:ilex-updated-tree}
+\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A phylogenetic tree is the main result of a Physcraper analysis.}\label{fig:results}
 \end{figure}
 
-The Physcraper updated tree has 57 original tips, representing 51 unique OTT taxa, and 222 new tips representing 78 new OTT taxa added to the original tree.
+\newpage
 
-It has THIS many resolved nodes, and THIS MANY nodes with bootsrap values above 95\%
-
-The original tree has THIS MANY resolved nodes
-
-The alignments have THIS MANY columns and THIS AMOUNT of missing data.
-
-DESCRIBE RESULTS: SUMMARY OF NEW TAXA FOUND RELATIVE TO ORIGINAL TREE AND RELATIVE TO OpenTree
-RF DISTANCE INTERPRETATION
-HOW MUCH TIME THE BLAST RUN TOOK
-ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE
-
-FIGURE: FACE TO FACE ORIGINAL VS UPDATED PHYLOGENY, IN RED NEW TAXA NOT IN OpenTree.
-
-\newpage
+Figure \ref{fig:results} caption continued: The tree results from updating an alignment that was used to generate a tree of 48 tips of the genus Ilex, published in Gottlieb et al.~2005 (black tips).
+Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
+Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML, as well as a conflict analysis using the OpenTree tools.
 
@@
 
-COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING
-WHY WE DID NOT MAKE A BENCHMARK COMPARISON
+\emph{COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING}
+\emph{WHY WE DID NOT MAKE A BENCHMARK COMPARISON?}
 
@@
 
-We thank members of the ``short bar'' Science and Engineering Building 1, University of California Merced, joint lab paper discussion meeting for valuable comments on a first version of this manuscript.
+We thank members of the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on a first version of this manuscript.
 

### 2020-08-18 2-body.Rmd
discussion

@@
 
-Physcraper relies on programmatic access protocols to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
@@
 Figure \@ref(fig:results) shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from @gottlieb2005molecular.
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to finish BLAST and RAxML analyses.
-Bootstrap analyses took an additional 13hrs.
+Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
@@
 `r 135-68` are also in updated Gottlieb2005.
+@yao2020phylogeny also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+The Yao2020 tree also misses 18 taxa that were added by Physcraper.
+This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.
 <!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
 <!-- RF DISTANCE INTERPRETATION -->
-<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE -->
+<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE ?-->
 
@@
 
-```{r results, echo=FALSE, fig.cap="A phylogenetic tree is the main result of a Physcraper analysis.", out.width = '100%', fig.retina= 2}
+```{r results, echo=FALSE, fig.cap="A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.", out.width = '100%', fig.retina= 2}
 
@@
 
-Figure \@ref(fig:results) caption continued: The tree results from updating an alignment that was used to generate a tree of 48 tips of the genus Ilex, published in Gottlieb et al.
-2005 (black tips).
-Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
-Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML, as well as a conflict analysis using the OpenTree tools.
+*Figure \@ref(fig:results) caption continued*: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
+Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section "DNA sequence search and filtering", aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
+B-D show results of the conflict analysis performed with OpenTree tools.
 
@@
 
-Data repositories hold even more information than meets the eye.
-Besides the actual data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Data repositories hold more information than meets the eye.
+Besides the main data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
 
@@
 Others focus on refining the alignment.
-Most focus on generating full trees *de novo*, mainly for regions of the Tree of
+Most focus on generating full trees *de novo* for regions of the Tree of
 Life that have no phylogenetic assessment yet in published studies, but also for
@@
 
-Physcraper has the potential to be applied for the advantage of the field to rapidly *HOW FAST IS "RAPID" mention it in results and then here again*
+Physcraper has the potential to rapidly (in a matter of hours)
+solve  in various 
 place newly discovered species phylogenetically [@webb2010biodiversity],

### 2020-08-18 docs/submission-mee-2020-08-18.tex
discussion

--- 2020-08-18 docs/submission-mee-2020-08-17.tex
@@
 
-\textbf{Word count}: 3745
+\textbf{Word count}: 3803
 
@@
 67 are also in updated Gottlieb2005.
+Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+The Yao2020 tree also misses 18 taxa that were added by Physcraper.
+This might be related to the method they used to download existing ITS \emph{Ilex} sequences from GenBank, but they do not clarify this in their publication.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A phylogenetic tree is the main result of a Physcraper analysis.}\label{fig:results}
+\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.}\label{fig:results}
 \end{figure}
@@
 
-Figure \ref{fig:results} caption continued: The tree results from updating an alignment that was used to generate a tree of 48 tips of the genus Ilex, published in Gottlieb et al.~2005 (black tips).
-Physcraper added 231 new sequences (red tips) corresponding to 84 distinct taxa that were not sampled in the original alignment.
-Physcraper obtained sequences from the GenBank database via BLAST, filtered them following specified criteria, aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML, as well as a conflict analysis using the OpenTree tools.
+Figure \ref{fig:results} caption continued: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
+Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
+B-D show results of the conflict analysis performed with OpenTree tools.
 

### 2020-08-18 2-body.Rmd
discussion

@@
 
+Data repositories designed to preserve and democratize access to biological data
+constitute an essential resource for evolutionary estimation. While data keep accumulating,
+our ability to reanalyze and incorporate new data to available knowledge is being challenged.
+
+Various phylogenetic pipelines have been designed to make sense of the vast amount of public molecular data available in different ways.
+Pipelines like Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward]
+focus mainly on generating full trees *de novo* for regions of the Tree of
+Life that have no phylogenetic assessment yet in published studies.
+While Physcraper does not generate phylogenies *de novo* in a traditional sense,
+it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
+phylogenetic knowledge available in tree repostories to facilitate phylogenetic placement of public molecular data.
+The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
+pre-existing alignments to incorporate public molecular data into phylogenies.
+Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+
+As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is allowing us to reach decade-long dreams of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+
+As such, Physcraper has the potential to rapidly (in a matter of hours)
+solve challenges overarching both fields of ecology and evolution, such as
+placing newly discovered species phylogenetically [@webb2010biodiversity],
+obtaining trees for ecophylogenetic studies [@helmus2012phylogenetic],
+sistematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
+and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
 Data repositories hold more information than meets the eye.
-Besides the main data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
 
-*COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING*
-*WHY WE DID NOT MAKE A BENCHMARK COMPARISON?*
+Usually, initial ideas about the data are changed by new analyses.
+Ideally, new understanding of the data can be registered on databases,
+exposing newcomers to the most up to date knowledge about the data.
 
-Many pipelines are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
+<!-- Most of them focus on efficient ways to mine the data -- getting the most homologs.
 Some focus on accurate ways of mining the data -- getting real and clean homologs.
-Others focus on refining the alignment.
-Most focus on generating full trees *de novo* for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have already been studied and which have phylogenetic data.
-However, expert phylogenetic knowledge is also an important source of data in public
-and open repositories that is not being used to its full potential.
+Others focus on refining the alignment. -->
 
-All these tools are key efforts for advancing towards reproducibility in phylogenetics,
-a field that has relied on processes which are somewhat artisanal.
-Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
-community to build up our phylogenetic knowledge, we can use not only data accumulated in
-molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
-
-While not generating full phylogenies *de novo*, Physcraper is still capable of generating new phylogenetic knowledge.
-Moreover, it can combine phylogenies with data from repositories other than molecular data.
-For example geographic locations (using GBIF ids), fossils (using PBDB ids), etc.
-*from Robert: I think you can sell the program more here.
+ <!-- *from Robert: I think you can sell the program more here.
 Why is it better than the other methods?
-You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages*
-
-Physcraper has the potential to rapidly (in a matter of hours)
-solve  in various 
-place newly discovered species phylogenetically [@webb2010biodiversity],
-obtain trees for ecophylogenetic studies [@helmus2012phylogenetic],
-help to systematize molecular databases, i.e., curate taxonomic assignations [@san2010molecular],
-and rapidly generate custom species trees for downstream analyses [@stoltzfus2013phylotastic].
+You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages* -->
 
@@
 Physcraper source code available at https://github.com/McTavishLab/physcraper
+
 Documentation available at https://physcraper.readthedocs.io/en/latest/index.html
+
 Illustrated examples available at https://github.com/McTavishLab/physcraperex
+
 This is a reproducible manuscript available at https://github.com/McTavishLab/physcraper_ms

### 2020-08-18 2.4-discussion-ideas.Rmd
discussion

--- 2020-07-01 2.4-discussion-ideas.Rmd
@@
-Usually, initial ideas about the data are changed by analyses.
-We expect that this new ideas on the data can be registered on databases,
-exposing newcomers to expert understanding about the data.
+<!-- databases designed to preserve and democratize access to biological data constitute an amazing resource for evolutionary estimation -->
 
-Taking advantage of public DNA data bases have been the main focus.
+*COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING*
+*WHY WE DID NOT MAKE A BENCHMARK COMPARISON?*
+
+<!-- Usually, initial ideas about the data are changed by analyses.
+Ideally, new understanding of the data can be registered on databases,
+exposing newcomers to the most up to date knowledge about the data. -->
+
+<!-- Taking advantage of public DNA data bases have been the main focus. -->
 
@@
 - Finding
-
+-
+<!--
 Physcraper potential:
@@
 - one day could be used to sistematize nucleotide databases, such as Genbank [@benson2000genbank; @wheeler2000database], as mentioned in @san2010molecular, i.e., curate ncbi taxonomic assignations.
-- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic
+- allows to generate custom species trees for downstream analyses, as mentioned in @stoltzfus2013phylotastic -->
 
@@
 \comment(hmm yeah, I think it is probably better for the latter - phylogeneticists who are tired of blast searching :P)
+
+All these tools are key efforts for advancing towards reproducibility in phylogenetics,
+a field that has relied on processes which are somewhat artisanal.
+Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
+
+<!-- Creating unified standards usually ends up generating an extra standard.
+However, the OpenTree Taxonomy is a powerful tool that is allowing us to reach decade-long dreams of "exploring and analysing biodiversity at an accelerated pace, to return systematics into the mianstream of science" [@wilson2003encyclopedia].
+-->
+
+<!-- Physcraper might be overlapping/partially solving the goal of the Encyclopedia of Life:
+"the exploration and analysis of biodiversity at a vastly accelerated pace.
+Its momentum will return systematics from its long sojourn at the margin and back into the mainstream of science." [@wilson2003encyclopedia] -->

### 2020-08-18 docs/submission-mee-2020-08-18.tex
discussion

@@
 
-\textbf{Word count}: 3803
+\textbf{Word count}: 3769
 
@@
 Figure \ref{fig:results} shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to finish BLAST and RAxML analyses.
-Bootstrap analyses took an additional 13hrs.
+Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
@@
 The Yao2020 tree also misses 18 taxa that were added by Physcraper.
-This might be related to the method they used to download existing ITS \emph{Ilex} sequences from GenBank, but they do not clarify this in their publication.
+This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication.
 
@@
 
-Figure \ref{fig:results} caption continued: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
+\emph{Figure \ref{fig:results} caption continued}: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
 Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
@@
 
-Data repositories hold even more information than meets the eye.
-Besides the actual data, they are rich sources of metadata that can be used for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Data repositories designed to preserve and democratize access to biological data
+constitute an essential resource for evolutionary estimation. While data keep accumulating,
+our ability to reanalyze and incorporate new data to available knowledge is being challenged.
 
-\emph{COMPARE WITH PERFORMANCE OF OTHER PIPELINES FOR SEQUENCE SCRAPING}
-\emph{WHY WE DID NOT MAKE A BENCHMARK COMPARISON?}
+Various phylogenetic pipelines have been designed to make sense of the vast amount of public molecular data available in different ways.
+Pipelines like Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})
+focus mainly on generating full trees \emph{de novo} for regions of the Tree of
+Life that have no phylogenetic assessment yet in published studies.
+While Physcraper does not generate phylogenies \emph{de novo} in a traditional sense,
+it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
+phylogenetic knowledge available in tree repostories to facilitate phylogenetic placement of public molecular data.
+The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
+pre-existing alignments to incorporate public molecular data into phylogenies.
+Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
 
-Many pipelines are making use of DNA data repositories in different ways.
-Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data -- getting real and clean homologs.
-Others focus on refining the alignment.
-Most focus on generating full trees \emph{de novo}, mainly for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies, but also for
-regions that have already been studied and which have phylogenetic data.
-However, expert phylogenetic knowledge is also an important source of data in public
-and open repositories that is not being used to its full potential.
+As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is allowing us to reach decade-long dreams of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
 
-All these tools are key efforts for advancing towards reproducibility in phylogenetics,
-a field that has relied on processes which are somewhat artisanal.
-Here, we highlight the potential of taking advantage of this careful curation work in previous phylogenetic estimates.
-By taking sources of information available from data repositories and present a method to link data from different repositories, while leveraging the knowledge and intuition of the expert
-community to build up our phylogenetic knowledge, we can use not only data accumulated in
-molecular data repositories, but phylogenetic knowledge accumulated in phylogenetic tree repositories.
+As such, Physcraper has the potential to rapidly (in a matter of hours)
+solve challenges overarching both fields of ecology and evolution, such as
+placing newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
+obtaining trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
+sistematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
+and generating custom species trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+Data repositories hold more information than meets the eye.
+Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
 
-While not generating full phylogenies \emph{de novo}, Physcraper is still capable of generating new phylogenetic knowledge.
-Moreover, it can combine phylogenies with data from repositories other than molecular data.
-For example geographic locations (using GBIF ids), fossils (using PBDB ids), etc.
-\emph{from Robert: I think you can sell the program more here.
-Why is it better than the other methods?
-You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages}
-
-Physcraper has the potential to be applied for the advantage of the field to rapidly \emph{HOW FAST IS ``RAPID'' mention it in results and then here again}
-place newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
-obtain trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-help to systematize molecular databases, i.e., curate taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
-and rapidly generate custom species trees for downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+Usually, initial ideas about the data are changed by new analyses.
+Ideally, new understanding of the data can be registered on databases,
+exposing newcomers to the most up to date knowledge about the data.
 
@@
 Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
+
 Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
+
 Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
+
 This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}

### 2020-08-18 2-body.Rmd
discussion

@@
 The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
-<!-- However, if there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
-Hence, Phylesystem's metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
--->
 
@@
 The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
-  <!-- Discussion: (although there are plans to allow storage of unpublished trees). -->
-
-  <!--
-  **EJM: Currently 2950 studies have treebase data links. It is possible that some/many of these have multiple trees in them, but that is harder to count.** -->
-  <!-- there are `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website
-  ([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).   -->
 
@@
 Detailed instructions to setup a local database are provided on Physcraper's software documentation.
-<!-- Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
-The code to do this with Physcraper is not yet implemented.
--->
+
 - The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
@@
 - All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
-- New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the unified taxonomy, or they are longer than the sequence in the input alignment.
-<!-- *MK mentioned it was unclear to her if this was an NCBI or OTT taxon. I think it is the unified taxonomy, hence an OTT taxon? I think it doesn't harm to be super explicit about it*
-**EJM: It checks using the OTT taxonomy, but in practice OpenTree doesn't merge ncbi taxa, so it is both/either** -->
+- New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
 - Among the filtered sequences, there are often several representatives per taxon.
@@
 The "Yao2020" tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-<!-- However, it has not been added to OpenTree's synthtetic tree yet. -->
 This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
-<!-- we know that the sequences of the Yao2020 tree have been made available on the GenBank database.
-Hence, we expect that updating the Gottlieb2005 tree should at least recover the same species sampled in the Yao2020 tree.
--->
 A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
@@
 it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repostories to facilitate phylogenetic placement of public molecular data.
+phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
 The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
@@
 As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is allowing us to reach decade-long dreams of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
 
-As such, Physcraper has the potential to rapidly (in a matter of hours)
+As such, Physcraper can be used to rapidly (in a matter of hours)
 solve challenges overarching both fields of ecology and evolution, such as
@@
 and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
+
 Data repositories hold more information than meets the eye.
 Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-
 Usually, initial ideas about the data are changed by new analyses.
@@
 exposing newcomers to the most up to date knowledge about the data.
-
-<!-- Most of them focus on efficient ways to mine the data -- getting the most homologs.
-Some focus on accurate ways of mining the data -- getting real and clean homologs.
-Others focus on refining the alignment. -->
-
- <!-- *from Robert: I think you can sell the program more here.
-Why is it better than the other methods?
-You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages* -->
 

### 2020-08-18 2.2-methods-ideas.Rmd
some edits
discussion

@@
 - Conflict information can only be generated in the context of the whole Open Tree of Life. Otherwise, it is not really possible to get conflict data.
+
+figure of unified taxonomy
+
+a table that is a snipet of the csv to show the unified taxonoy for some taxa
+also, a screen shoy of the diffeent taxon ides on opentree website works!
+
+**EJM: Currently 2950 studies have treebase data links. It is possible that some/many of these have multiple trees in them, but that is harder to count.**
+there are `r tryCatch({rotl::tol_about()$num_source_trees}, error = function(e) {1216})` published trees supporting the resolved nodes of the synthetic tree in the OpenTree website
+([See OpenTree's website about](https://tree.opentreeoflife.org/about/synthesis-release/v12.3)).
+
+Alternative DNA molecular databases could be used for this purpose, either personal or public databases, such as the European Nucleotide Archive or the DNA DataBank of Japan.
+The code to do this with Physcraper is not yet implemented.

### 2020-08-18 2.4-discussion-ideas.Rmd
discussion

@@
 <!-- Taking advantage of public DNA data bases have been the main focus. -->
+
+Discussion: (although there are plans to allow storage of unpublished trees).
 
@@
 Its momentum will return systematics from its long sojourn at the margin and back into the mainstream of science." [@wilson2003encyclopedia] -->
+
+<!-- Most of them focus on efficient ways to mine the data -- getting the most homologs.
+Some focus on accurate ways of mining the data -- getting real and clean homologs.
+Others focus on refining the alignment. -->
+
+ <!-- *from Robert: I think you can sell the program more here.
+Why is it better than the other methods?
+You mentioned in lab meeting that its difficult to run other programs, talk about that here, talk about the speed and other advantages* -->

### 2020-08-18 docs/submission-mee-2020-08-18.tex
discussion

@@
 
-\textbf{Word count}: 3769
+\textbf{Word count}: 3775
 
@@
 \item
-  New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the unified taxonomy, or they are longer than the sequence in the input alignment.
+  New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
 \item
@@
 it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repostories to facilitate phylogenetic placement of public molecular data.
+phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
 The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
@@
 As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is allowing us to reach decade-long dreams of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is contributing to the Encyclopedia of Life main goal of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
 
-As such, Physcraper has the potential to rapidly (in a matter of hours)
+As such, Physcraper can be used to rapidly (in a matter of hours)
 solve challenges overarching both fields of ecology and evolution, such as
@@
 and generating custom species trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+
 Data repositories hold more information than meets the eye.
 Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-
 Usually, initial ideas about the data are changed by new analyses.

### 2020-08-18 2-body.Rmd

@@
 To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life (OpenTree from now on) project
+The main aim of the Open Tree of Life project (OpenTree)
 <!-- [project](https://opentreeoflife.github.io/)  -->
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
@@
 
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
@@
 <!-- *maybe a box of all pipelines?* -->
-However, even during phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
+However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
@@
 Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
+The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
 
-Physcraper relies on programmatic access protocols available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
@@
   Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
-  If the user is not ready to make the input tree public, tree tip labels must be first standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-*Should we say that it is based on Boyle's 2012 algorithm?
-or is it a TNRS algorithm independently developed by the OpenTree team?*
+  If the user is not ready to make the input tree public, tree tip labels can be standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
   This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
-  Physcraper saves a summary "csv" file with results from the taxon name standardization for advantage of the user.
-  If taxon names can't be mapped, their taxonomic information can't be used in comparison analysis. They will still be used in the sequence search and phylogenetic reconstruction steps.
+  Physcraper saves a summary "csv" file with results from the taxon name standardization for advantage of the user, including the mappings to unique identifiers in the OpenTree and NCBI taxonomies.
+  If taxon names can't be mapped, their taxonomic information isn't be used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
   Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
@@
 
-- The input alignment should be a single locus alignment that was used to generate the tree. Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
+- The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
+Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
   DRYAD ([www.datadryad.org](http://datadryad.org/)), or a data repository associated with the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper downloads it directly,
+  If the alignment is stored in TreeBase, Physcraper can download it directly,
   either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
@@
 Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
-See next section.
 - The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
@@
   The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
-- All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
+- All sequences accepted up to this point are assigned an internal identifier.
 - New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
@@
 The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
+A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment would ideal.
 - Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 - For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-- A suggested default rooting based on OpenTree's taxonomy is implemented for now.
+A suggested default rooting based on OpenTree's taxonomy is implemented for now.
 This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
@@
 \@ref(fig:results)).
+**Is this in the consensus tree?
+I'm surprised they survive bootstrapping!**
 <!-- ADD percentage of nodes with bootsrap values above 95%  and below 50%?-->
@@
 @yao2020phylogeny also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-The Yao2020 tree also misses 18 taxa that were added by Physcraper.
+When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
+The Yao2020 tree also lacks 18 taxa that were added by Physcraper. 
+**Any statements about relationships or anything? Are there multiple seqs for these taxa?**
 This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.
+  Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
 <!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
@@
 constitute an essential resource for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data to available knowledge is being challenged.
+our ability to reanalyze and incorporate new data with available knowledge is being challenged.
 
@@
 obtaining trees for ecophylogenetic studies [@helmus2012phylogenetic],
-sistematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
 and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
+
+Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
+However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
+ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
 
@@
 
-We thank members of the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on a first version of this manuscript.
+We thank the members of the OpenTree development team and the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.
 

### 2020-08-19 2-body.Rmd
ejm comments

@@
   Physcraper saves a summary "csv" file with results from the taxon name standardization for advantage of the user, including the mappings to unique identifiers in the OpenTree and NCBI taxonomies.
-  If taxon names can't be mapped, their taxonomic information isn't be used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
+  If taxon names can't be mapped, their taxonomic information is not used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
   Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
@@
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
-The tree is 99% resolved, however, a large portion of branches are negligibly small (133 branches < 0.00001 substitution rate units), appearing as unresolved nodes (Fig.
-\@ref(fig:results)).
-**Is this in the consensus tree?
-I'm surprised they survive bootstrapping!**
-<!-- ADD percentage of nodes with bootsrap values above 95%  and below 50%?-->
+The best RaxML tree is 99% resolved, while the consensus tree is 75% resolved, meaning that 25% of nodes have a bootstrap < 0.1, while 48% nodes have bootstrap support > 0.75.
+A large portion of internal branches are negligibly small, with 30 branches < 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75 (Fig. \@ref(fig:results)).
 For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
@@
 and contributes phylogenetic data to 134 additional taxa within *Ilex*, from which
-`r 135-68` are also in updated Gottlieb2005.
+`r 135-68` are also in the updated Gottlieb2005 tree.
 @yao2020phylogeny also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
@@
 The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
-**Any statements about relationships or anything? Are there multiple seqs for these taxa?**
+<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?** -->
 This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.
@@
 
-Various phylogenetic pipelines have been designed to make sense of the vast amount of public molecular data available in different ways.
+Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
 Pipelines like Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward]
-focus mainly on generating full trees *de novo* for regions of the Tree of
+focus mainly on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
 Life that have no phylogenetic assessment yet in published studies.
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
+However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
+ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
 
 As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
 
@@
 
-Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
-However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
-ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
-
 Data repositories hold more information than meets the eye.
@@
 Usually, initial ideas about the data are changed by new analyses.
-Ideally, new understanding of the data can be registered on databases,
-exposing newcomers to the most up to date knowledge about the data.
+Ideally, new understanding of the data can be continually registered on databases
+to consistently expose newcomers to the most up to date knowledge about the data.
 

### 2020-08-19 2-body.Rmd
bullets

@@
 
-- In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
   OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
@@
 
-- The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
+The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
 Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
@@
   If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-- Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
+
+Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
 Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
 Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
-- The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
+The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
 
@@
 
-- Physcraper uses the GenBank DNA database as source to search for new sequences.
+Physcraper uses the GenBank DNA database as source to search for new sequences.
 The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
@@
 
-- The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
+The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
   The search taxon can be chosen by the user from the NCBI taxonomy.
@@
   the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-- The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
+The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
   similarity between DNA sequences within the search taxon in a nucleotide
@@
 This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
-- A constrained BLAST search is performed, in which each sequence
+A constrained BLAST search is performed, in which each sequence
   in the alignment is BLASTed once against all database DNA sequences belonging to the search
@@
   The full sequence for each match is downloaded from NCBI into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-- BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80% and 120%, respectively).
+BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80% and 120%, respectively).
   This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
   The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
-- All sequences accepted up to this point are assigned an internal identifier.
-- New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
-- Among the filtered sequences, there are often several representatives per taxon.
+All sequences accepted up to this point are assigned an internal identifier.
+New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
+Among the filtered sequences, there are often several representatives per taxon.
   Although it can be useful to keep some of them, for example, to investigate monophyly
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-- All BLAST and filtering parameters can be customized by the user.
-- Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions [@cock2009biopython].
-- Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+All BLAST and filtering parameters can be customized by the user.
+Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions [@cock2009biopython].
+Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
 By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
-- New sequences passing all filtering steps are added to the "csv" taxon summary file.
-- A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
+New sequences passing all filtering steps are added to the "csv" taxon summary file.
+A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 
-- By default, Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
+By default, Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
 Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
-- The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
-- Second, a MUSCLE profile alignment is performed, in which the original alignment
+The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
+Second, a MUSCLE profile alignment is performed, in which the original alignment
   is used as a template to align the new sequences. This ensures that the final alignment
   follows the homology criteria established by the original alignment.
-- The final alignment is not further processed by Physcraper.
+The final alignment is not further processed by Physcraper.
 It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection; @castresana2002gblocks].
-- While curating the alignment is a critical step, it is not a reproducible one.
+While curating the alignment is a critical step, it is not a reproducible one.
 The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
 A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment would ideal.
-- Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
+Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
 
@@
 
-- A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml] with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
+A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml] with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
 Currently only the number of bootsrap replicates can be specified by the user.
-- By default, the original tree is used as a starting tree for the ML searches.
-Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
-- Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
-- Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
-- Tips on all trees generated by Physcraper are defined by a taxon "name space".
+By default, the original tree is used as a starting tree for the ML searches. Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
+Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
+Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
+Tips on all trees generated by Physcraper are defined by a taxon "name space".
 The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
-- Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+Two ways to compare the updated tree with the original tree are implemented in Physcraper.
 First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
-- Second, a conflict analysis is performed.
+Second, a conflict analysis is performed.
 This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
 This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
-- For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
+For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
 A suggested default rooting based on OpenTree's taxonomy is implemented for now.

### 2020-08-19 2-body.Rmd
second draft final :tada:

@@
   the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
+
 The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
@@
   The full sequence for each match is downloaded from NCBI into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
+
 BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80% and 120%, respectively).
@@
   5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
+
 All BLAST and filtering parameters can be customized by the user.
@@
 The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
+
 Second, a MUSCLE profile alignment is performed, in which the original alignment
@@
 Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
+
 Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
@@
 The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
-<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?** -->
+<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?**
+add 1 sentence saying that it is odd that we are getting differnet relationships given what is likely to be the same data
+but we can't compare the exact seqs or alignemnt, bc they are not available.
+ -->
 This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.

### 2020-08-19 docs/submission-mee-2020-08-19.tex
second draft final :tada:

--- 2020-08-18 docs/submission-mee-2020-08-18.tex
@@
 
-\textbf{Word count}: 3775
+\textbf{Word count}: 3911
 
@@
 To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life (OpenTree from now on) project
+The main aim of the Open Tree of Life project (OpenTree)
 is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
 While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million non-genomic DNA sequences.
+Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
@@
 Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, even during phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
@@
 Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated to TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
+The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
 
-Physcraper relies on programmatic access protocols available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
@@
 
-\begin{itemize}
-\item
 In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
@@
 Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
-  If the user is not ready to make the input tree public, tree tip labels must be first standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-\emph{Should we say that it is based on Boyle's 2012 algorithm?
-or is it a TNRS algorithm independently developed by the OpenTree team?}
+If the user is not ready to make the input tree public, tree tip labels can be standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
 This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
-  Physcraper saves a summary ``csv'' file with results from the taxon name standardization for advantage of the user.
-  If taxon names can't be mapped, their taxonomic information can't be used in comparison analysis. They will still be used in the sequence search and phylogenetic reconstruction steps.
+Physcraper saves a summary ``csv'' file with results from the taxon name standardization for advantage of the user, including the mappings to unique identifiers in the OpenTree and NCBI taxonomies.
+If taxon names can't be mapped, their taxonomic information is not used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
 Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
@@
 The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
-\item
-  The input alignment should be a single locus alignment that was used to generate the tree.
+
+The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
 Alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
 DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or a data repository associated with the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper downloads it directly,
+If the alignment is stored in TreeBase, Physcraper can download it directly,
 either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
@@
 If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
-\item
+
 Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
@@
 Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
-See next section.
-\item
 The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
-\end{itemize}
 
@@
 
-\begin{itemize}
-\item
 Physcraper uses the GenBank DNA database as source to search for new sequences.
@@
 Detailed instructions to setup a local database are provided on Physcraper's software documentation.
-\item
+
 The next step is to identify a ``search taxon'' to constrain the sequence search on the GenBank database within that taxonomic group.
@@
 the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
-\item
+
 The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}, \protect\hyperlink{ref-altschul1997gapped}{1997}) is used to identify
@@
 This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
-\item
 A constrained BLAST search is performed, in which each sequence
@@
 The full sequence for each match is downloaded from NCBI into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
-\item
+
 BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80\% and 120\%, respectively).
@@
 The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
-\item
-  All sequences accepted up to this point are assigned an internal identifier, and are further filtered.
-\item
+All sequences accepted up to this point are assigned an internal identifier.
 New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
-\item
 Among the filtered sequences, there are often several representatives per taxon.
@@
 5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-\item
+
 All BLAST and filtering parameters can be customized by the user.
-\item
 Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
-\item
 Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
 By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
-\item
 New sequences passing all filtering steps are added to the ``csv'' taxon summary file.
-\item
 A ``fasta'' file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
-\end{itemize}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
 By default, Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
 Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
-\item
 The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
-\item
+
 Second, a MUSCLE profile alignment is performed, in which the original alignment
@@
 follows the homology criteria established by the original alignment.
-\item
 The final alignment is not further processed by Physcraper.
 It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
-\item
 While curating the alignment is a critical step, it is not a reproducible one.
 The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment will be ideal.
-\item
+A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment would ideal.
 Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
-\end{itemize}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
 A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
 Currently only the number of bootsrap replicates can be specified by the user.
-\item
-  By default, the original tree is used as a starting tree for the ML searches.
-Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
-\item
+By default, the original tree is used as a starting tree for the ML searches. Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
 Bootstrap results are summarized with the SumTrees module of DendroPy (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
-\item
+
 Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
-\item
 Tips on all trees generated by Physcraper are defined by a taxon ``name space''.
 The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
-\item
 Two ways to compare the updated tree with the original tree are implemented in Physcraper.
 First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
-\item
 Second, a conflict analysis is performed.
@@
 This is performed with OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
-\item
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-\item
 A suggested default rooting based on OpenTree's taxonomy is implemented for now.
@@
 The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
-\end{itemize}
 
@@
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
-The tree is 99\% resolved, however, a large portion of branches are negligibly small (133 branches \textless{} 0.00001 substitution rate units), appearing as unresolved nodes (Fig.
+The best RaxML tree is 99\% resolved, while the consensus tree is 75\% resolved, meaning that 25\% of nodes have a bootstrap \textless{} 0.1, while 48\% nodes have bootstrap support \textgreater{} 0.75.
+A large portion of internal branches are negligibly small, with 30 branches \textless{} 0.00001 substitution rate units, from which only 9 have a bootstrap support \textgreater{} 0.75 (Fig.
 \ref{fig:results}).
@@
 and contributes phylogenetic data to 134 additional taxa within \emph{Ilex}, from which
-67 are also in updated Gottlieb2005.
+67 are also in the updated Gottlieb2005 tree.
 Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-The Yao2020 tree also misses 18 taxa that were added by Physcraper.
+When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
+The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
 This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication.
+Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
 
@@
 constitute an essential resource for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data to available knowledge is being challenged.
+our ability to reanalyze and incorporate new data with available knowledge is being challenged.
 
-Various phylogenetic pipelines have been designed to make sense of the vast amount of public molecular data available in different ways.
+Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
 Pipelines like Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})
-focus mainly on generating full trees \emph{de novo} for regions of the Tree of
+focus mainly on generating full trees \emph{de novo} (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
 Life that have no phylogenetic assessment yet in published studies.
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.~ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
 
 As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, etc, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is contributing to the Encyclopedia of Life main goal of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
+It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is contributing to the Encyclopedia of Life main goal of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
 
@@
 obtaining trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-sistematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
 and generating custom species trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
@@
 
-We thank members of the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on a first version of this manuscript.
+We thank the members of the OpenTree development team and the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-bouckaert2019beast}{}%
+Bouckaert, R., Vaughan, T.G., Barido-Sottani, J., Duchêne, S., Fourment, M., Gavryushkina, A., Heled, J., Jones, G., Kühnert, D., Maio, N.D., Matschiner, M., Mendes, F.K., Müller, N.F., Ogilvie, H.A., Plessis, L.
+du, Popinga, A., Rambaut, A., Rasmussen, D., Siveroni, I., Suchard, M.A., Wu, C.-H., Xie, D., Zhang, C., Stadler, T.
+\& Drummond, A.J.
+(2019).
+BEAST 2.5: An advanced software platform for Bayesian evolutionary analysis.
+\emph{PLOS Computational Biology}, \textbf{15}, e1006650.
+Retrieved August 18, 2020, from \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006650}
+
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
@@
 Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular biology and evolution}, \textbf{17}, 540--552.
+
+\leavevmode\hypertarget{ref-chifman2014quartet}{}%
+Chifman, J.
+\& Kubatko, L.
+(2014).
+Quartet Inference from SNP Data Under the Coalescent Model.
+\emph{Bioinformatics}, \textbf{30}, 3317--3324.
+Retrieved August 18, 2020, from \url{https://academic.oup.com/bioinformatics/article/30/23/3317/206559}
 
@@
 \emph{Bioinformatics}, \textbf{31}, 2794--2800.
+
+\leavevmode\hypertarget{ref-mirarab2014astral}{}%
+Mirarab, S., Reaz, R., Bayzid, M.S., Zimmermann, T., Swenson, M.S.
+\& Warnow, T.
+(2014).
+ASTRAL: Genome-scale coalescent-based species tree estimation.
+\emph{Bioinformatics}, \textbf{30}, i541--i548.
+Retrieved July 25, 2015, from \url{http://bioinformatics.oxfordjournals.org/content/30/17/i541}
 

### 2020-08-21 comments/submission-mee-2020-08-19_MH.tex
MH comments

--- 2020-08-19 docs/submission-mee-2020-08-19.tex
@@
 
+\usepackage{ulem}
+\newcommand{\mthcomment}[1]{{\color{red} #1}}
+\newcommand{\mthsuggest}[2]{{\color{red} \sout{#1} #2}}
+
 \title{Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life}
@@
   However, interpretation of results, comparison with previously available phylogenetic
-  hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
+  hypotheses, and \mthsuggest{choosing}{choice} of one phylogeny for downstream analyses and discussion still impose difficulties
   to one that is not a specialist either on phylogenetic methods or on a particular group of study.
@@
   Physcraper is a command-line Python program that automates the update of published
-  phylogenies by enriching underlying gene alignments with public DNA sequence data.
+  phylogenies by \mthsuggest{enriching underlying gene alignments with public DNA sequence data}{adding public DNA sequences to underlying alignments}.
 It builds upon tools from the Open Tree of Life project to link taxonomic information across databases, providing a framework for straightforward comparison of published phylogenies with their updated versions.
+\mthcomment{Rephrase.
+I'm not sure what is ``providing'' here; is it OToL that is providing, or the taxonomic dbs, or physcraper?}
 \item
@@
   hypotheses based on publicly available expert phylogenetic knowledge.
-  Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
+  Phylogeneticists and \mthsuggest{}{taxonomic} group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
   of alternative phylogenetic hypotheses (topologies).
 \item
-  We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+  \mthsuggest{We hope that t}{T}he Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and
-  use are available at \href{https://physcraper.readthedocs.io/en/tutorial/index.html}{https://physcraper.readthedocs}.
+  use are available at \href{https://physcraper.readthedocs.io/en/tutorial/index.html}{https://physcraper.readthedocs}. \mthcomment{I think you need to chop the "index.html" from the end of the URL} 
 \end{enumerate}
@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms\mthsuggest{,}{} and provide key context for all our biological observations.
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments\mthsuggest{, to}{.
+Physcraper lets researchers} build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
@@
 Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}, providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}\mthsuggest{, providing}{.
+This provides} a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{taxonomic tool v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+\mthsuggest{}{Perhaps, OpenTree API rather than ``taxonomic tool v3''}
 
@@
 Figure \ref{fig:results} shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
+Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs \mthcomment{you probably need some details about the hardware, given the fact that you are discussing running times}.
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.

### 2020-08-24 1-abstract.Rmd
MH abstract comments

@@
 However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
+hypotheses, and choice of one phylogeny for downstream analyses and discussion still impose difficulties
 to one that is not a specialist either on phylogenetic methods or on a particular group of study.
@@
 1. Physcraper is a command-line Python program that automates the update of published
-phylogenies by enriching underlying gene alignments with public DNA sequence data.
-It builds upon tools from the Open Tree of Life project to link taxonomic information across databases, providing a framework for straightforward comparison of published phylogenies with their updated versions.
+phylogenies by adding public DNA sequences to underlying alignments of previously published phylogenies.
+It also provides a framework for straightforward comparison of published phylogenies with their updated versions, by leveraging upon tools from the Open Tree of Life project to link taxonomic information across databases.
 <!-- the conflict Application Programming Interface (API) function of  the Open Tree of Life project-->
@@
 hypotheses based on publicly available expert phylogenetic knowledge.
-Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
+Phylogeneticists and taxonomic group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
 of alternative phylogenetic hypotheses (topologies).
@@
 1.
-We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and
-use are available at [https://physcraper.readthedocs](https://physcraper.readthedocs.io/en/tutorial/index.html).
+use are available at [https://physcraper.readthedocs](https://physcraper.readthedocs.io).

### 2020-08-24 2-body.Rmd
MH all comments

@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms, and provide key context for all our biological observations.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms and provide key context for all our biological observations.
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments, to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments.
+Physcraper lets researchers build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 
@@
 Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the World Register of Marine Species [WoRMS; www.marinespecies.org/], providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the World Register of Marine Species [WoRMS; www.marinespecies.org/].
+This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
   The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
-  To identify the MRCAT of a group of taxon names, we use the OpenTree [taxonomic tool v3](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
+  To identify the MRCAT of a group of taxon names, we use the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
 
@@
 Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
+**MTH: you probably need some details
+252 about the hardware, given the fact that you are discussing running times**
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+**MTH: add some more details about what aspect of the PUMPER install failed**
 Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].

### 2020-08-24 docs/submission-mee-2020-08-24.tex
MH all comments

--- 2020-08-19 docs/submission-mee-2020-08-19.tex
@@
 
-\textbf{Word count}: 3911
+\textbf{Word count}: 3948
 
@@
   However, interpretation of results, comparison with previously available phylogenetic
-  hypotheses, and choosing of one phylogeny for downstream analyses and discussion still impose difficulties
+  hypotheses, and choice of one phylogeny for downstream analyses and discussion still impose difficulties
   to one that is not a specialist either on phylogenetic methods or on a particular group of study.
@@
   Physcraper is a command-line Python program that automates the update of published
-  phylogenies by enriching underlying gene alignments with public DNA sequence data.
-It builds upon tools from the Open Tree of Life project to link taxonomic information across databases, providing a framework for straightforward comparison of published phylogenies with their updated versions.
+  phylogenies by adding public DNA sequences to underlying alignments of previously published phylogenies.
+It also provides a framework for straightforward comparison of published phylogenies with their updated versions, by leveraging upon tools from the Open Tree of Life project to link taxonomic information across databases.
 \item
@@
   hypotheses based on publicly available expert phylogenetic knowledge.
-  Phylogeneticists and group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
+  Phylogeneticists and taxonomic group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
   of alternative phylogenetic hypotheses (topologies).
 \item
-  We hope that the Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+  The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
 Physcraper can be used with any OS and is released under an open-source license.
 Detailed instructions for installation and
-  use are available at \href{https://physcraper.readthedocs.io/en/tutorial/index.html}{https://physcraper.readthedocs}.
+  use are available at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs}.
 \end{enumerate}
@@
 Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}, providing a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}.
+This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
@@
 The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
-To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{taxonomic tool v3} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
 
@@
 Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
+\textbf{MTH: you probably need some details
+252 about the hardware, given the fact that you are discussing running times}
 The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+\textbf{MTH: add some more details about what aspect of the PUMPER install failed}
 Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).

### 2020-08-25 2-body.Rmd
declaration of conflict

@@
 
+The authors have no conflict of interest to declare.
+
 # Authors' Contributions
 
-EJM: Conceived study, wrote most of the code, documentation and tests.
-MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
-LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
-All authors contributed to the manuscript.
+LLSR wrote the manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
+All authors contributed to the manuscript and gave final approval for publication.
 
-# Data Avilability
+# Data Archiving
 

### 2020-08-25 docs/submission-mee-2020-08-25.tex
twitter abstract

--- 2020-08-24 docs/submission-mee-2020-08-24.tex
@@
 
-\textbf{Running title}: Continually updated gene trees with Physcraper
+\textbf{Running title}: Updating gene trees with the Open Tree of Life
 
-\textbf{Word count}: 3948
+\textbf{Word count}: 3963
 
@@
 
+The authors have no conflict of interest to declare.
+
 \hypertarget{authors-contributions}{%
@@
 
-EJM: Conceived study, wrote most of the code, documentation and tests.
-MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
-LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
-All authors contributed to the manuscript.
+LLSR wrote the manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
+All authors contributed to the manuscript and gave final approval for publication.
 
-\hypertarget{data-avilability}{%
-\section{Data Avilability}\label{data-avilability}}
+\hypertarget{data-archiving}{%
+\section{Data Archiving}\label{data-archiving}}
 

### 2020-08-25 render2-mee.Rmd
ejm comments intro
twitter abstract

@@
 params:
-  running_title: Continually updated gene trees with Physcraper
+  running_title: Updating gene trees with the Open Tree of Life
+  twitter: 'Physcraper: a tool to update published phylogenies and link taxa to biological databases using the Open Tree of Life'
   author1: Luna L. Sanchez Reyes
@@
 **2. `r params$author2`**
+
+`r params$affiliation`
 

### 2020-08-28 methods-extended.Rmd
extended methods

--- 2020-08-24 2-body.Rmd
@@
 ---
-title: 'Physcraper: Body'
-author:
-  - Luna L. Sanchez Reyes
-  - Martha Kandziora
-  - Emily Jane McTavish
-  - University of California, Merced
-date: '`r format(Sys.time(), "%d %B, %Y")`'
+output:
+  html_document:
+    keep_md: yes
+    toc: false
+geometry: margin = 1in
 bibliography: paper.bib
-csl: apa.csl
-output:
-  redoc::redoc
+csl: submission-mee/methods-in-ecology-and-evolution.csl
+# csl: submission_sysbio/systematic-biology.csl
+link-citations: yes
+linestretch: 2
 ---
 
-# Introduction
+# The Physcraper Framework
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms and provide key context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments.
-Physcraper lets researchers build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
-
-The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life project (OpenTree)
-<!-- [project](https://opentreeoflife.github.io/)  -->
-is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the World Register of Marine Species [WoRMS; www.marinespecies.org/].
-This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
-
-Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
-Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
-<!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
-
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
-Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible approach.
-Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
-<!-- *maybe a box of all pipelines?* -->
-However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
-
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
-Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
-
-Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by integrating taxonomic name matching across biological databases.
-This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, making a variety of downstream analyses straightforward.
-
-# The Physcraper framework
-
-```{r framework, echo=FALSE, fig.cap="The Physcraper framework consists of 4 steps (see text).
-The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.", out.width = '85%', fig.retina= 2, fig.align = "center"}
-# knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-knitr::include_graphics("docs/figs/schematic.pdf")
-```
-
-The general Physcraper framework is depicted in Figure \@ref(fig:framework).
-Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
-Next, we will describe technical details for each step.
+The general Physcraper framework consists of 4 general steps: **1) identifying and processing** a phylogenetic **tree to update and its underlying alignment**; **2) performing a constrained BLAST search** of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; **3) profile-aligning new sequences** that passed the filtering to the original alignment; **4) performing a phylogenetic analysis and comparing the updated tree** to previous phylogenetic estimates in the focus group.
+Next, we will describe each step with some technical detail.
 
@@
 
-In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+In order to take advantage of the OpenTree tools, it is recommended that the input tree is either stored in the
   OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
@@
 
-# Example: The hollies
-
-To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus *Ilex*.
-Commonly known as "hollies", the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-
-An online literature review in June 2020 (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
-@gottlieb2005molecular made original tree and alignment data available in [TreeBASE](https://treebase.org/treebase-web/search/study/summary.html?id=1091).
-The "Gottlieb2005" tree sampling 41 species was added to the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and its information has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).
-
-The most recent *Ilex* tree from @yao2020phylogeny, has been made available in the [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/view/ot_1984) and in the [DRYAD repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x).
-The "Yao2020" tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
-A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
-Figure \@ref(fig:results) shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from @gottlieb2005molecular.
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
-**MTH: you probably need some details
-252 about the hardware, given the fact that you are discussing running times**
-The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
-The best RaxML tree is 99% resolved, while the consensus tree is 75% resolved, meaning that 25% of nodes have a bootstrap < 0.1, while 48% nodes have bootstrap support > 0.75.
-A large portion of internal branches are negligibly small, with 30 branches < 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75 (Fig. \@ref(fig:results)).
-For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
-<!-- has less new tips (134), but  -->
-and contributes phylogenetic data to 134 additional taxa within *Ilex*, from which
-`r 135-68` are also in the updated Gottlieb2005 tree.
-@yao2020phylogeny also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
-The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
-<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?**
-add 1 sentence saying that it is odd that we are getting differnet relationships given what is likely to be the same data
-but we can't compare the exact seqs or alignemnt, bc they are not available.
- -->
-This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.
-  Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
-<!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
-<!-- RF DISTANCE INTERPRETATION -->
-<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE ?-->
-
-<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
-wget -O docs/figs/ilex-plot-tips-and-branches-1.png https://raw.githubusercontent.com/McTavishLab/physcraperex/master/docs/articles/ilex_files/figure-html/plot-tips-and-branches-1.png
-``` -->
-
-```{r results, echo=FALSE, fig.cap="A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.", out.width = '100%', fig.retina= 2}
-
-# knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
-knitr::include_graphics("docs/figs/ilex-results.png")
-
-```
-
-\newpage
-
-*Figure \@ref(fig:results) caption continued*: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
-Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section "DNA sequence search and filtering", aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis performed with OpenTree tools.
-
-# Discussion
-
-Data repositories designed to preserve and democratize access to biological data
-constitute an essential resource for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data with available knowledge is being challenged.
-
-Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
-Pipelines like Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward]
-focus mainly on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies.
-While Physcraper does not generate phylogenies *de novo* in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
-The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
-pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
-**MTH: add some more details about what aspect of the PUMPER install failed**
-Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
-However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
-ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
-
-As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
-
-As such, Physcraper can be used to rapidly (in a matter of hours)
-solve challenges overarching both fields of ecology and evolution, such as
-placing newly discovered species phylogenetically [@webb2010biodiversity],
-obtaining trees for ecophylogenetic studies [@helmus2012phylogenetic],
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
-and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
-
-Data repositories hold more information than meets the eye.
-Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-Usually, initial ideas about the data are changed by new analyses.
-Ideally, new understanding of the data can be continually registered on databases
-to consistently expose newcomers to the most up to date knowledge about the data.
-
-# Acknowledgements
-
-Research was supported by the grant "Sustaining the Open Tree of Life", National Science Foundation ABI No. 1759838, and ABI No. 1759846.
-Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.
-ACI-1429783.
-
-We thank the members of the OpenTree development team and the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.
-
-# Authors' Contributions
-
-EJM: Conceived study, wrote most of the code, documentation and tests.
-MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
-LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
-All authors contributed to the manuscript.
-
-# Data Avilability
-
-Physcraper source code available at https://github.com/McTavishLab/physcraper
-
-Documentation available at https://physcraper.readthedocs.io/en/latest/index.html
-
-Illustrated examples available at https://github.com/McTavishLab/physcraperex
-
-This is a reproducible manuscript available at https://github.com/McTavishLab/physcraper_ms
-
-\newpage
 # References

### 2020-08-28 methods-extended.Rmd
extended methods

@@
 output:
+  bookdown::pdf_document2:
+    keep_tex: yes
+    toc: false
   html_document:
     keep_md: yes
-    toc: false
 geometry: margin = 1in

### 2020-08-28 methods-extended.tex
extended methods

--- 2020-08-24 docs/submission-mee-2020-08-24.tex
@@
 \hypersetup{
-            pdftitle={Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life},
             pdfborder={0 0 0},
@@
 
-\usepackage{color}
-\usepackage{hyperref}
-\usepackage{caption}
-\usepackage{blindtext}
-\usepackage{url}
-\usepackage[left]{lineno}
-\linenumbers
-
-\title{Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life}
 \author{}
@@
 \begin{document}
-\maketitle
-
-\textbf{1. Luna L. Sanchez Reyes}
-
-School of Natural Sciences, University of California, Merced
-
-email: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}
-
-\textbf{2. Martha Kandziora}
-
-School of Natural Sciences, University of California, Merced
-
-Department of Botany, Faculty of Science, Charles University, Prague, Czech Republic
-
-email: \href{mailto:kandziom@natur.cuni.cz}{\nolinkurl{kandziom@natur.cuni.cz}}
-
-\textbf{3. Emily Jane McTavish}
-
-School of Natural Sciences, University of California, Merced
-
-email: \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
-
-\textbf{Correspondence address}: Science and Engineering Building 1, University of California, Merced, 5200 N. Lake Rd, Merced CA 95343
-
-\textbf{Correspondence email}: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
-
-\textbf{Running title}: Continually updated gene trees with Physcraper
-
-\textbf{Word count}: 3948
-
-\textbf{Manuscript prepared for submission to Methods in Ecology and Evolution}
-
-\textbf{Article type}: Application
-
-\newpage
-
-\begingroup\Large
-
-\textbf{Abstract}
-\endgroup
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Phylogenies are a key part of research in many areas of biology. Tools that automate
-  some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
-  have been developed for the advantage of both specialists in the field of phylogenetics and nonspecialists.
-  However, interpretation of results, comparison with previously available phylogenetic
-  hypotheses, and choice of one phylogeny for downstream analyses and discussion still impose difficulties
-  to one that is not a specialist either on phylogenetic methods or on a particular group of study.
-\item
-  Physcraper is a command-line Python program that automates the update of published
-  phylogenies by adding public DNA sequences to underlying alignments of previously published phylogenies.
-It also provides a framework for straightforward comparison of published phylogenies with their updated versions, by leveraging upon tools from the Open Tree of Life project to link taxonomic information across databases.
-\item
-  Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
-  hypotheses based on publicly available expert phylogenetic knowledge.
-  Phylogeneticists and taxonomic group specialists will find it useful as a tool to facilitate molecular dataset gathering and comparison
-  of alternative phylogenetic hypotheses (topologies).
-\item
-  The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
-Physcraper can be used with any OS and is released under an open-source license.
-Detailed instructions for installation and
-  use are available at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs}.
-\end{enumerate}
-
-\textbf{Keywords}: gene tree, interoperability, open science, open tree of life, phylogeny, public database, python, reproducibility, taxonomy, updated alignment
-
-\newpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms and provide key context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments.
-Physcraper lets researchers build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
-
-The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life project (OpenTree)
-is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}.
-This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
-
-Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
-
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible approach.
-Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
-
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a source of valuable expert knowledge.
-Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
-
-Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by integrating taxonomic name matching across biological databases.
-This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, making a variety of downstream analyses straightforward.
 
 \hypertarget{the-physcraper-framework}{%
-\section{The Physcraper framework}\label{the-physcraper-framework}}
+\section{The Physcraper Framework}\label{the-physcraper-framework}}
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic}
-
-}
-
-\caption{The Physcraper framework consists of 4 steps (see text).
-The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.}\label{fig:framework}
-\end{figure}
-
-The general Physcraper framework is depicted in Figure \ref{fig:framework}.
-Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
-Next, we will describe technical details for each step.
+The general Physcraper framework consists of 4 general steps: \textbf{1) identifying and processing} a phylogenetic \textbf{tree to update and its underlying alignment}; \textbf{2) performing a constrained BLAST search} of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; \textbf{3) profile-aligning new sequences} that passed the filtering to the original alignment; \textbf{4) performing a phylogenetic analysis and comparing the updated tree} to previous phylogenetic estimates in the focus group.
+Next, we will describe each step with some technical detail.
 
@@
 
-In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
+In order to take advantage of the OpenTree tools, it is recommended that the input tree is either stored in the
 OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
@@
 
-\hypertarget{example-the-hollies}{%
-\section{Example: The hollies}\label{example-the-hollies}}
-
-To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus \emph{Ilex}.
-Commonly known as ``hollies'', the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
-
-An online literature review in June 2020 (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
-Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment data available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
-The ``Gottlieb2005'' tree sampling 41 species was added to the \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and its information has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
-
-The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}), has been made available in the \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
-The ``Yao2020'' tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
-A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
-Figure \ref{fig:results} shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
-\textbf{MTH: you probably need some details
-252 about the hardware, given the fact that you are discussing running times}
-The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
-The best RaxML tree is 99\% resolved, while the consensus tree is 75\% resolved, meaning that 25\% of nodes have a bootstrap \textless{} 0.1, while 48\% nodes have bootstrap support \textgreater{} 0.75.
-A large portion of internal branches are negligibly small, with 30 branches \textless{} 0.00001 substitution rate units, from which only 9 have a bootstrap support \textgreater{} 0.75 (Fig.
-\ref{fig:results}).
-For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
-and contributes phylogenetic data to 134 additional taxa within \emph{Ilex}, from which
-67 are also in the updated Gottlieb2005 tree.
-Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
-The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
-This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication.
-Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.}\label{fig:results}
-\end{figure}
-
-\newpage
-
-\emph{Figure \ref{fig:results} caption continued}: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
-Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis performed with OpenTree tools.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Data repositories designed to preserve and democratize access to biological data
-constitute an essential resource for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data with available knowledge is being challenged.
-
-Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
-Pipelines like Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})
-focus mainly on generating full trees \emph{de novo} (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies.
-While Physcraper does not generate phylogenies \emph{de novo} in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
-The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
-pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
-\textbf{MTH: add some more details about what aspect of the PUMPER install failed}
-Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.~ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
-
-As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is contributing to the Encyclopedia of Life main goal of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
-
-As such, Physcraper can be used to rapidly (in a matter of hours)
-solve challenges overarching both fields of ecology and evolution, such as
-placing newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
-obtaining trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
-and generating custom species trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
-
-Data repositories hold more information than meets the eye.
-Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-Usually, initial ideas about the data are changed by new analyses.
-Ideally, new understanding of the data can be continually registered on databases
-to consistently expose newcomers to the most up to date knowledge about the data.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838, and ABI No.~1759846.
-Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
-
-We thank the members of the OpenTree development team and the ``short bar'' Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.
-
-\hypertarget{authors-contributions}{%
-\section{Authors' Contributions}\label{authors-contributions}}
-
-EJM: Conceived study, wrote most of the code, documentation and tests.
-MK: Wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests.
-LLSR: Wrote the manuscript, alignment code, documentation, performed analyses and developed examples.
-All authors contributed to the manuscript.
-
-\hypertarget{data-avilability}{%
-\section{Data Avilability}\label{data-avilability}}
-
-Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
-
-Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
-
-Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
-
-This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}
-
-\newpage
-
 \hypertarget{references}{%
-\section{References}\label{references}}
-
-\newpage
-\begin{center}
-\textsc{References}
-\end{center}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 
-\leavevmode\hypertarget{ref-andermann2020guide}{}%
-Andermann, T., Torres Jiménez, M.F., Matos-Maraví, P., Batista, R., Blanco-Pastor, J.L., Gustafsson, A.L.S., Kistler, L., Liberal, I.M., Oxelman, B., Bacon, C.D.
-\& Antonelli, A.
-(2020).
-A Guide to Carrying Out a Phylogenomic Target Sequence Capture Project.
-\emph{Frontiers in Genetics}, \textbf{10}.
-Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7047930/}
-
-\leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli, A., Hettling, H., Condamine, F.L., Vos, K., Nilsson, R.H., Sanderson, M.J., Sauquet, H., Scharn, R., Silvestro, D., Töpel, M.
-\& others.
-(2017).
-Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
-\emph{Systematic Biology}, \textbf{66}, 152--166.
-
-\leavevmode\hypertarget{ref-benson2000genbank}{}%
-Benson, D.A., Karsch-Mizrachi, I., Lipman, D.J., Ostell, J., Rapp, B.A. \& Wheeler, D.L. (2000). GenBank. \emph{Nucleic acids research}, \textbf{28}, 15--18.
-
-\leavevmode\hypertarget{ref-bouckaert2019beast}{}%
-Bouckaert, R., Vaughan, T.G., Barido-Sottani, J., Duchêne, S., Fourment, M., Gavryushkina, A., Heled, J., Jones, G., Kühnert, D., Maio, N.D., Matschiner, M., Mendes, F.K., Müller, N.F., Ogilvie, H.A., Plessis, L.
-du, Popinga, A., Rambaut, A., Rasmussen, D., Siveroni, I., Suchard, M.A., Wu, C.-H., Xie, D., Zhang, C., Stadler, T.
-\& Drummond, A.J.
-(2019).
-BEAST 2.5: An advanced software platform for Bayesian evolutionary analysis.
-\emph{PLOS Computational Biology}, \textbf{15}, e1006650.
-Retrieved August 18, 2020, from \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006650}
-
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
@@
 
-\leavevmode\hypertarget{ref-chifman2014quartet}{}%
-Chifman, J.
-\& Kubatko, L.
-(2014).
-Quartet Inference from SNP Data Under the Coalescent Model.
-\emph{Bioinformatics}, \textbf{30}, 3317--3324.
-Retrieved August 18, 2020, from \url{https://academic.oup.com/bioinformatics/article/30/23/3317/206559}
-
 \leavevmode\hypertarget{ref-cock2009biopython}{}%
@@
 
-\leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
-Cuénoud, P., Martinez, M.A.
-del P., Loizeay, P.-A., Spichiger, R., Andrews, S.
-\& Manen, J.-F.
-(2000).
-Molecular phylogeny and biogeography of the genus ilex l.(Aquifoliaceae).
-\emph{Annals of Botany}, \textbf{85}, 111--122.
-
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
 Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic acids research}, \textbf{32}, 1792--1797.
-
-\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
-Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
-\& Mendoza, C.G.
-(2017).
-A pilot study applying the plant anchored hybrid enrichment method to new world sages (salvia subgenus calosphace; lamiaceae).
-\emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
-
-\leavevmode\hypertarget{ref-gottlieb2005molecular}{}%
-Gottlieb, A.M., Giberti, G.C.
-\& Poggio, L.
-(2005).
-Molecular analyses of the genus ilex (aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
-\emph{American Journal of Botany}, \textbf{92}, 352--369.
-
-\leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
-Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
-
-\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
-
-\leavevmode\hypertarget{ref-jones2016targeted}{}%
-Jones, M.R.
-\& Good, J.M.
-(2016).
-TARGETED CAPTURE IN EVOLUTIONARY AND ECOLOGICAL GENOMICS.
-\emph{Molecular ecology}, \textbf{25}, 185--202.
-Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4823023/}
-
-\leavevmode\hypertarget{ref-manen2010history}{}%
-Manen, J.-F., Barriera, G., Loizeau, P.-A.
-\& Naciri, Y.
-(2010).
-The history of extant ilex species (aquifoliaceae): Evidence of hybridization within a miocene radiation.
-\emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
 
@@
 \emph{Bioinformatics}, \textbf{31}, 2794--2800.
-
-\leavevmode\hypertarget{ref-mirarab2014astral}{}%
-Mirarab, S., Reaz, R., Bayzid, M.S., Zimmermann, T., Swenson, M.S.
-\& Warnow, T.
-(2014).
-ASTRAL: Genome-scale coalescent-based species tree estimation.
-\emph{Bioinformatics}, \textbf{30}, i541--i548.
-Retrieved July 25, 2015, from \url{http://bioinformatics.oxfordjournals.org/content/30/17/i541}
-
-\leavevmode\hypertarget{ref-morrison2006multiple}{}%
-Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
 
@@
 
-\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A. \& Wehe, A. (2008). The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. \emph{Systematic Biology}, \textbf{57}, 335--346.
-
-\leavevmode\hypertarget{ref-san2010molecular}{}%
-San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
-
-\leavevmode\hypertarget{ref-secretariat2017gbif}{}%
-Secretariat, G. (2017). GBIF backbone taxonomy. \emph{Checklist Dataset {[}cited 2017 Nov 14{]}. doi}, \textbf{10}.
-
-\leavevmode\hypertarget{ref-selbach2009new}{}%
-Selbach-Schnadelbach, A., Cavalli, S.S., Manen, J.-F., Coelho, G.C.
-\& De Souza-Chies, T.T.
-(2009).
-New information for ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (aquifoliaceae).
-\emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.
-
-\leavevmode\hypertarget{ref-setoguchi2000intersectional}{}%
-Setoguchi, H.
-\& Watanabe, I.
-(2000).
-Intersectional gene flow between insular endemics of ilex (aquifoliaceae) on the bonin islands and the ryukyu islands.
-\emph{American Journal of Botany}, \textbf{87}, 793--810.
-
-\leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith, S.A., Beaulieu, J.M.
-\& Donoghue, M.J.
-(2009).
-Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
-\emph{BMC evolutionary biology}, \textbf{9}, 37.
-
-\leavevmode\hypertarget{ref-song2012resolving}{}%
-Song, S., Liu, L., Edwards, S.V.
-\& Wu, S.
-(2012).
-Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
-\emph{Proceedings of the National Academy of Sciences}, \textbf{109}, 14942--14947.
-
 \leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
 Stamatakis, A. (2014). RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. \emph{Bioinformatics}, \textbf{30}, 1312--1313.
-
-\leavevmode\hypertarget{ref-stoltzfus2013phylotastic}{}%
-Stoltzfus, A., Lapp, H., Matasci, N., Deus, H., Sidlauskas, B., Zmasek, C.M., Vaidya, G., Pontelli, E., Cranston, K., Vos, R.
-\& others.
-(2013).
-Phylotastic!
-Making tree-of-life knowledge accessible, reusable and convenient.
-\emph{BMC bioinformatics}, \textbf{14}, 158.
 
@@
 
-\leavevmode\hypertarget{ref-webb2010biodiversity}{}%
-Webb, C.O., Slik, J.F. \& Triono, T. (2010). Biodiversity inventory and informatics in southeast asia. \emph{Biodiversity and Conservation}, \textbf{19}, 955--972.
-
-\leavevmode\hypertarget{ref-wheeler2000database}{}%
-Wheeler, D.L., Chappey, C., Lash, A.E., Leipe, D.D., Madden, T.L., Schuler, G.D., Tatusova, T.A.
-\& Rapp, B.A.
-(2000).
-Database resources of the national center for biotechnology information.
-\emph{Nucleic acids research}, \textbf{28}, 10--14.
-
-\leavevmode\hypertarget{ref-wilson2003encyclopedia}{}%
-Wilson, E.O. (2003). The encyclopedia of life. \emph{Trends in Ecology \& Evolution}, \textbf{18}, 77--80.
-
-\leavevmode\hypertarget{ref-yao2020phylogeny}{}%
-Yao, X., Song, Y., Yang, J.-B., Tan, Y.-H. \& Corlett, R.T. (2020). Phylogeny and biogeography of the hollies (ilex l., aquifoliaceae). \emph{Journal of Systematics and Evolution}.
-
 \end{document}

### 2020-08-31 2-body.Rmd
word limit

@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms and provide key context for all our biological observations.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living and extinct organisms and provide key context for all our biological observations.
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments.
-Physcraper lets researchers build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using phylogenetic knowledge from public biological databases to update a starting tree and single locus alignments.
 
-The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life project (OpenTree)
+Automatic integration of data in a phylogenetic context is challenged by the prevalence of taxonomic idiosyncrasies across biological databases.
+Taxonomic name standardization represents a step toward tackling this issue, but requires a unified system.
+The Open Tree of Life project (OpenTree)
 <!-- [project](https://opentreeoflife.github.io/)  -->
-is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated] such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], and the World Register of Marine Species [WoRMS; www.marinespecies.org/].
-This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+has constructed a comprehensive and digitally-available tree of life by synthesizing published phylogenies along with taxonomic data.
+Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), etc.
+The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
 Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships [@song2012resolving].
+Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving].
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
+More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
@@
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a source of valuable expert knowledge.
+The TreeBASE database [@piel2009treebase] hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a public source of valuable expert knowledge.
 Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
+The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated framework linking trees with the alignment that generated them.
 
-Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by integrating taxonomic name matching across biological databases.
-This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, making a variety of downstream analyses straightforward.
+Physcraper relies on programmatic access protocols (API's) to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
+Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the standardized taxonomy as a way to link taxon data from different databases.
+This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.
 
@@
 The general Physcraper framework is depicted in Figure \@ref(fig:framework).
-Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
-Next, we will describe technical details for each step.
+It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
 
@@
 
-In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
-  OpenTree [Phylesystem](https://github.com/opentreeoflife/phylesystem), or submitted via OpenTree's curator [application](https://tree.opentreeoflife.org/curator) [@mctavish2015phylesystem].
-  Currently, only trees connected to a published study can be stored in the Phylesystem.
-  Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
-  If the user is not ready to make the input tree public, tree tip labels can be standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-  This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
-  Physcraper saves a summary "csv" file with results from the taxon name standardization for advantage of the user, including the mappings to unique identifiers in the OpenTree and NCBI taxonomies.
-  If taxon names can't be mapped, their taxonomic information is not used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
-  Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
-  that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+Taxon names in the input tree must be standardized or "mapped" to the unified OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
+Users can choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase; @vos2012nexml].
 
-The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
-Alignments are often stored in a public repository such as TreeBase [@piel2009treebase; @vos2012nexml],
-  DRYAD ([www.datadryad.org](http://datadryad.org/)), or a data repository associated with the journal where the tree was originally published.
-  If the alignment is stored in TreeBase, Physcraper can download it directly,
-  either from the TreeBASE website ([www.treebase.org](https://treebase.org/))
-  or through the TreeBASE GitHub repository (SuperTreeBASE; [github.com/TreeBASE/supertreebase](https://github.com/TreeBASE/supertreebase)).
-  If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
+The input alignment should be a single locus alignment that was used in part or in whole, to generate the input tree.
+If the alignment associated to the input tree is stred in TreeBASE, Physcraper retrieves it automatically.
+Alignments stored in any other repository, or constituting personal data have to be downloaded by the user.
 
-Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
-Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
-The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as "newick" and "fasta" respectively in the "inputs" folder to be used in the following steps.
+Physcraper processes the input tree to match taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name and its corresponding sequence in the alignment are needed to continue the algorithm.
 
@@
 
-Physcraper uses the GenBank DNA database as source to search for new sequences.
-The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
-Detailed instructions to setup a local database are provided on Physcraper's software documentation.
+The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped], either on the GenBank remote database or in a GenBank local database set up by the user.
+It is constrained to a taxonomic group in the NCBI taxonomy, defined as the "search taxon".
+Users can arbitrarily define a search taxon that is either a more or a less inclusive
+clade relative to the ingroup of the input tree.
+Otherwise, the search taxon is automatically identified using the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated], as the Most Recent Common Ancestral Taxon (MRCAT) of the ingroup taxa in the input tree, i.e., the MRCA that is also a named clade in the NCBI taxonomy (Fig.
+\@ref(fig:framework)).
+  <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
+The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
 
-The next step is to identify a "search taxon" to constrain the sequence search on the GenBank database within that taxonomic group.
-  The search taxon can be chosen by the user from the NCBI taxonomy.
-  If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the "ingroup" (Fig. \@ref(fig:framework)).
-  The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
-  <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
-  This is known in the OpenTree as the Most Recent Common
-  Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA).
-  The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
-  To identify the MRCAT of a group of taxon names, we use the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated].
+BLAST command line tools `blastn` function [@camacho2009blast] is used for local database sequence searches.
+For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
+This is useful when a user lacks access to the computer capacity needed to setup a local database, but has access to an institutional server.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank within the search taxon.
+GenBank sequences with match scores smaller than the e-value cutoff (default to 0.00001) are downloaded into a local library.
 
-  Users can provide a search taxon that is either a more or a less inclusive
-  clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
+Downlodaded sequences are not further considered in the analysis if they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OTT taxonomy or the NCBI taxonomy.
+If there are several sequences per taxon, an arbitrary number of 5 (can be modified by the user) sequences per taxon are chosen at random.
 
-The Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped] is used to identify
-  similarity between DNA sequences within the search taxon in a nucleotide
-  database, and the sequences on the checked alignment.
-  The `blastn` function from the BLAST command line tools [@camacho2009blast] is used for local database sequence searches.
-  For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
-A constrained BLAST search is performed, in which each sequence
-  in the alignment is BLASTed once against all database DNA sequences belonging to the search
-  taxon. All results from each BLAST run are stored, and sequences with match scores better than the e-value cutoff (default to 0.00001) are saved
-  along with their corresponding metadata, i.e., their GenBank accession number.
-  The full sequence for each match is downloaded from NCBI into a dedicated library within the "physcraper" folder, allowing for secondary analyses to run significantly faster.
-
-BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80% and 120%, respectively).
-  This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
-  The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
-All sequences accepted up to this point are assigned an internal identifier.
-New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
-Among the filtered sequences, there are often several representatives per taxon.
-  Although it can be useful to keep some of them, for example, to investigate monophyly
-  within species, there can be hundreds of exemplar sequences per taxon for some markers.
-  To control the number of sequences per taxon in downstream analyses,
-  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-
-All BLAST and filtering parameters can be customized by the user.
-Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions [@cock2009biopython].
-Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions [@cock2009biopython].
+Iterative cycles of BLAST searches can be performed, by blasting the new sequences until no new ones are found.
 By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
-New sequences passing all filtering steps are added to the "csv" taxon summary file.
-A "fasta" file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 
-By default, Physcraper uses the software MUSCLE [@edgar2004muscle] to perform DNA sequence alignments.
-Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
-The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
+By default, Physcraper uses MUSCLE [@edgar2004muscle] to perform DNA sequence alignments in a two step process. First, all new sequences are aligned using the default MUSCLE options.
 
-Second, a MUSCLE profile alignment is performed, in which the original alignment
-  is used as a template to align the new sequences. This ensures that the final alignment
+Second, a MUSCLE profile alignment is performed, in which the original alignment is used as a template to align the new sequences. This ensures that the final alignment
 follows the homology criteria established by the original alignment.
@@
 It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection; @castresana2002gblocks].
-While curating the alignment is a critical step, it is not a reproducible one.
-The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment would ideal.
 Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 
-A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML [@stamatakis2014raxml] with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
-Currently only the number of bootsrap replicates can be specified by the user.
-By default, the original tree is used as a starting tree for the ML searches. Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
+A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with default settings (GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm).
+Only the number of bootsrap replicates can be defined by the user.
+By default, the original tree is used as a starting tree for the ML searches. Alternatively, the original tree can be set as full topological constraint, or simply be ignored for the ML searches.
 Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
 
-Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
-Tips on all trees generated by Physcraper are defined by a taxon "name space".
-The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
-Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
+There are two ways to compare the updated tree with the original tree implemented in Physcraper.
 First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
 Second, a conflict analysis is performed.
-This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
-This is performed with OpenTree's conflict Application Programming Interface [@redelings2017supertree].
+This is a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict Application Programming Interface [@redelings2017supertree].
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
+The best way right now is for users to define the outgroup directly on the updated tree, so trees are accurately rooted.
 <!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->
@@
 
-To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus *Ilex*.
+A user is interested in the phylogenetic relationships within the genus *Ilex*.
 Commonly known as "hollies", the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
@@
 The "Yao2020" tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
+
 A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
-Figure \@ref(fig:results) shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from @gottlieb2005molecular.
+Figure \@ref(fig:results) shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from @gottlieb2005molecular.
 Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
-**MTH: you probably need some details
-252 about the hardware, given the fact that you are discussing running times**
-The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within *Ilex*.
-The best RaxML tree is 99% resolved, while the consensus tree is 75% resolved, meaning that 25% of nodes have a bootstrap < 0.1, while 48% nodes have bootstrap support > 0.75.
+<!-- **MTH: you probably need some details about the hardware, given the fact that you are discussing running times** -->
+The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional *Ilex* taxa.
+The best RaxML tree is 99% resolved, with 25% of nodes with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
 A large portion of internal branches are negligibly small, with 30 branches < 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75 (Fig. \@ref(fig:results)).
@@
 <!-- has less new tips (134), but  -->
-and contributes phylogenetic data to 134 additional taxa within *Ilex*, from which
-`r 135-68` are also in the updated Gottlieb2005 tree.
-@yao2020phylogeny also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
-The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
+and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
+`r 135-68` are also in the Physcraper updated Gottlieb2005 tree.
+While @yao2020phylogeny also used ITS as a marker, their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
 <!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?**
@@
  -->
-This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication.
-  Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
+This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication, but seems to be a "manual" process.
+<!-- More Physcraper examples at a range of taxonomic scales are available in the Physcraper documentation online. -->
 <!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
-**MTH: add some more details about what aspect of the PUMPER install failed**
+<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
 Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].

### 2020-08-31 docs/submission-mee-2020-08-31.tex
word limit

--- 2020-08-25 docs/submission-mee-2020-08-25.tex
@@
 
-\textbf{Word count}: 3963
+\textbf{Word count}: 2992
 
@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living organisms and provide key context for all our biological observations.
+Phylogenetic estimates of evolutionary relationships capture the shared history of living and extinct organisms and provide key context for all our biological observations.
 Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool that uses existing phylogenetic research from public biological databases to update a starting tree and single locus alignments.
-Physcraper lets researchers build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using phylogenetic knowledge from public biological databases to update a starting tree and single locus alignments.
 
-The prevalence of taxonomic idiosyncrasies across databases represent a key challenge to automatically connecting data from disparate biological databases in a phylogenetic context.
-To standardize taxonomic names, a unified system is needed.
-The main aim of the Open Tree of Life project (OpenTree)
-is to construct a comprehensive, dynamic and digitally-available tree of life by synthesizing published phylogenetic trees along with taxonomic data.
-Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are represented by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) such as the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), and the World Register of Marine Species {[}WoRMS; www.marinespecies.org/{]}.
-This provideslinks taxonomic information across databases a key resource that can be used to connect data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+Automatic integration of data in a phylogenetic context is challenged by the prevalence of taxonomic idiosyncrasies across biological databases.
+Taxonomic name standardization represents a step toward tackling this issue, but requires a unified system.
+The Open Tree of Life project (OpenTree)
+has constructed a comprehensive and digitally-available tree of life by synthesizing published phylogenies along with taxonomic data.
+Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the World Register of Marine Species (WoRMS), etc.
+The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
 
 Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are seen as the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches, from whole genomes, to transcriptomes, restriction-site associated DNA sequencing, single nucleotide polymorphisms, microsatellites, and ultra-conserved elements, which has lead to largely non-overlapping genomic data sets across taxa, creating difficulty in wide scale phylogenetic reconstructions.
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 
-Indeed, more than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
+More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
@@
 A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 publicly accessible alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a source of valuable expert knowledge.
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a public source of valuable expert knowledge.
 Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated way of linking trees with the alignment that generated them.
+The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated framework linking trees with the alignment that generated them.
 
-Physcraper relies on programmatic access protocols (API's) available to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various general OpenTree programmatic tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by integrating taxonomic name matching across biological databases.
-This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, making a variety of downstream analyses straightforward.
+Physcraper relies on programmatic access protocols (API's) to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
+Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the standardized taxonomy as a way to link taxon data from different databases.
+This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.
 
@@
 The general Physcraper framework is depicted in Figure \ref{fig:framework}.
-Briefly, it consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of sequences from the original alignment on the GenBank DNA database, and filtering of newly found sequences; 3) profile-aligning new sequences that passed the filtering to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates in the focus group.
-Next, we will describe technical details for each step.
+It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
 
@@
 
-In order to take advantage of the OpenTree tools, it is reccommended that the input tree is either stored in the
-OpenTree \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem}, or submitted via OpenTree's curator \href{https://tree.opentreeoflife.org/curator}{application} (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}).
-Currently, only trees connected to a published study can be stored in the Phylesystem.
-Users can choose from among the 2, 950 studies in OpenTree's Phylesystem that have alignments on TreeBASE.
-If the user is not ready to make the input tree public, tree tip labels can be standardized to the unified OpenTree taxonomy using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-This step is referred to as taxonomic name mapping and Phylesystem stored trees are processed in this way upon submission.
-Physcraper saves a summary ``csv'' file with results from the taxon name standardization for advantage of the user, including the mappings to unique identifiers in the OpenTree and NCBI taxonomies.
-If taxon names can't be mapped, their taxonomic information is not used in comparison analysis. These taxa will still be used in the sequence search and phylogenetic reconstruction steps.
-Mapping tip names to OpenTree's unified taxonomy saves a set of user defined characteristics
-that are essential for automatizing the phylogeny updating process.
-The most relevant of these is the standardized taxonomic names and the definition of ingroup and outgroup taxa, allowing to automatically set the root for the updated tree on the final steps of the pipeline.
+Taxon names in the input tree must be standardized or ``mapped'' to the unified OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+Users can choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}).
 
-The input alignment should be a single locus alignment that was used in part or in whole, to generate the tree.
-Alignments are often stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
-DRYAD (\href{http://datadryad.org/}{www.datadryad.org}), or a data repository associated with the journal where the tree was originally published.
-If the alignment is stored in TreeBase, Physcraper can download it directly,
-either from the TreeBASE website (\href{https://treebase.org/}{www.treebase.org})
-or through the TreeBASE GitHub repository (SuperTreeBASE; \href{https://github.com/TreeBASE/supertreebase}{github.com/TreeBASE/supertreebase}).
-If the alignment is on another repository, or constitutes personal data, a path to a local copy of the alignment has to be provided.
+The input alignment should be a single locus alignment that was used in part or in whole, to generate the input tree.
+If the alignment associated to the input tree is stred in TreeBASE, Physcraper retrieves it automatically.
+Alignments stored in any other repository, or constituting personal data have to be downloaded by the user.
 
-Single locus alignments sometimes have fewer taxa than the tree inferred from the full concatenated data, simply because a single molecular marker usually does not cover all the taxa sampled for the full phylogenetic analysis.
-Physcraper prunes the input tree to taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name (and its corresponding sequence in the alignment) is needed to continue the algorithm.
-The standardized and pruned tree and alignment (checked tree and alignment from now on) are output as ``newick'' and ``fasta'' respectively in the ``inputs'' folder to be used in the following steps.
+Physcraper processes the input tree to match taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
+Technically, just one taxon name and its corresponding sequence in the alignment are needed to continue the algorithm.
 
@@
 
-Physcraper uses the GenBank DNA database as source to search for new sequences.
-The DNA sequence search can be performed on the GenBank remote database or in a GenBank local database set up by the user, which can speed up the search process.
-Detailed instructions to setup a local database are provided on Physcraper's software documentation.
+The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}, \protect\hyperlink{ref-altschul1997gapped}{1997}), either on the GenBank remote database or in a GenBank local database set up by the user.
+It is constrained to a taxonomic group in the NCBI taxonomy, defined as the ``search taxon''.
+Users can arbitrarily define a search taxon that is either a more or a less inclusive
+clade relative to the ingroup of the input tree.
+Otherwise, the search taxon is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the Most Recent Common Ancestral Taxon (MRCAT) of the ingroup taxa in the input tree, i.e., the MRCA that is also a named clade in the NCBI taxonomy (Fig.
+\ref{fig:framework}).
+The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
 
-The next step is to identify a ``search taxon'' to constrain the sequence search on the GenBank database within that taxonomic group.
-The search taxon can be chosen by the user from the NCBI taxonomy.
-If none is provided, then the search taxon is automatically set using the taxa in the input tree labeled as the ``ingroup'' (Fig. \ref{fig:framework}).
-The search taxon is The Most Recent Common Ancestor (MRCA) of the ingroup taxa in the OpenTree synthetic tree, that is also a named clade in the NCBI taxonomy.
-This is known in the OpenTree as the Most Recent Common
-Ancestral Taxon (MRCAT; also referred as the Least Inclusive Common Ancestral taxon - LICA).
-The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
-To identify the MRCAT of a group of taxon names, we use the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+BLAST command line tools \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local database sequence searches.
+For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
+This is useful when a user lacks access to the computer capacity needed to setup a local database, but has access to an institutional server.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank within the search taxon.
+GenBank sequences with match scores smaller than the e-value cutoff (default to 0.00001) are downloaded into a local library.
 
-Users can provide a search taxon that is either a more or a less inclusive
-clade relative to the ingroup of the original phylogeny.
-If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
-the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
+Downlodaded sequences are not further considered in the analysis if they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80\% and 120\%, respectively); or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OTT taxonomy or the NCBI taxonomy.
+If there are several sequences per taxon, an arbitrary number of 5 (can be modified by the user) sequences per taxon are chosen at random.
 
-The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}, \protect\hyperlink{ref-altschul1997gapped}{1997}) is used to identify
-similarity between DNA sequences within the search taxon in a nucleotide
-database, and the sequences on the checked alignment.
-The \texttt{blastn} function from the BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local database sequence searches.
-For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
-This is useful when a user has no access to the computer capacity needed to setup a local database, and a local blast database can be set up on a remote machine to BLAST avoiding NCBI's required waiting times, which slow down the searches markedly.
-A constrained BLAST search is performed, in which each sequence
-in the alignment is BLASTed once against all database DNA sequences belonging to the search
-taxon. All results from each BLAST run are stored, and sequences with match scores better than the e-value cutoff (default to 0.00001) are saved
-along with their corresponding metadata, i.e., their GenBank accession number.
-The full sequence for each match is downloaded from NCBI into a dedicated library within the ``physcraper'' folder, allowing for secondary analyses to run significantly faster.
-
-BLAST result sequences will be discarded if they fall outside the user set min and max length cutoffs, set as proportions of the average length without gaps of sequences in the input alignment (defaults values of 80\% and 120\%, respectively).
-This filtering guarantees the exclusion of whole genome sequences, which create problems in multiple sequence alignment.
-The GenBank accession numbers of sequences removed due to not meeting e-value or length cutoffs are stored in output files.
-All sequences accepted up to this point are assigned an internal identifier.
-New sequences that are either identical or a subset of any existing sequence in the input alignment are discarded, unless they represent a different taxon in the OTT taxonomy or the NCBI taxonomy, or they are longer than the sequence in the input alignment.
-Among the filtered sequences, there are often several representatives per taxon.
-Although it can be useful to keep some of them, for example, to investigate monophyly
-within species, there can be hundreds of exemplar sequences per taxon for some markers.
-To control the number of sequences per taxon in downstream analyses,
-5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
-
-All BLAST and filtering parameters can be customized by the user.
-Reverse, complement, and reverse-complement BLAST result sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
-Iterative cycles of sequence similarity search can be performed, by blasting the newly found sequences until no new sequences are found.
+Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
+Iterative cycles of BLAST searches can be performed, by blasting the new sequences until no new ones are found.
 By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
-New sequences passing all filtering steps are added to the ``csv'' taxon summary file.
-A ``fasta'' file containing all new filtered and processed sequences resulting from the BLAST search is generated for the user, and is used as an input for alignment.
 
@@
 
-By default, Physcraper uses the software MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments.
-Instructions on how to install all software dependencies used by Physcraper are provided in the documentation.
-The process to align new sequences consists of two steps. First, all new sequences are aligned using the default MUSCLE options.
+By default, Physcraper uses MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments in a two step process.
+First, all new sequences are aligned using the default MUSCLE options.
 
-Second, a MUSCLE profile alignment is performed, in which the original alignment
-is used as a template to align the new sequences. This ensures that the final alignment
+Second, a MUSCLE profile alignment is performed, in which the original alignment is used as a template to align the new sequences. This ensures that the final alignment
 follows the homology criteria established by the original alignment.
@@
 It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
-While curating the alignment is a critical step, it is not a reproducible one.
-The main reason for its lack of reproducibility might be that it is hard to track changes made on the alignment.
-A form of version control, to register the differences between the alignment that was produced by the software and the manually curated alignment would ideal.
 Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 
-A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings, such as a GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm.
-Currently only the number of bootsrap replicates can be specified by the user.
-By default, the original tree is used as a starting tree for the ML searches. Alternatively, users can set the original tree as a full topological constraint, or ignore it completely for the searches.
+A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings (GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm).
+Only the number of bootsrap replicates can be defined by the user.
+By default, the original tree is used as a starting tree for the ML searches. Alternatively, the original tree can be set as full topological constraint, or simply be ignored for the ML searches.
 Bootstrap results are summarized with the SumTrees module of DendroPy (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 
-Physcraper's final result is an updated phylogenetic hypothesis for the locus provided in the input alignment.
-Tips on all trees generated by Physcraper are defined by a taxon ``name space''.
-The taxon metadata captures the NCBI accession information, as well as the taxon identifiers, allowing the user to perform comparisons and conflict analyses.
-Two ways to compare the updated tree with the original tree are implemented in Physcraper.
+Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
+There are two ways to compare the updated tree with the original tree implemented in Physcraper.
 First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 Second, a conflict analysis is performed.
-This is a node by node comparison between the the synthetic OpenTree and the original and updated tree individually.
-This is performed with OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
+This is a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define outgroup directly on the updated tree, so trees are accurately rooted.
+The best way right now is for users to define the outgroup directly on the updated tree, so trees are accurately rooted.
 
@@
 
-To illustrate the utility of Physcraper we propose a scenario in which a user is interested in phylogenetic relationships within the genus \emph{Ilex}.
+A user is interested in the phylogenetic relationships within the genus \emph{Ilex}.
 Commonly known as ``hollies'', the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
@@
 The ``Yao2020'' tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
-This makes it a great case to highlight the functionalities of Physcraper and assess its limitations.
+
 A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
-Figure \ref{fig:results} shows results from a Physcraper analysis applied to an alignment of internal transcribed spacer DNA regions (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
+Figure \ref{fig:results} shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
 Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
-\textbf{MTH: you probably need some details
-252 about the hardware, given the fact that you are discussing running times}
-The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional taxa within \emph{Ilex}.
-The best RaxML tree is 99\% resolved, while the consensus tree is 75\% resolved, meaning that 25\% of nodes have a bootstrap \textless{} 0.1, while 48\% nodes have bootstrap support \textgreater{} 0.75.
+The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional \emph{Ilex} taxa.
+The best RaxML tree is 99\% resolved, with 25\% of nodes with bootstrap support \textless{} 0.1 and 48\% nodes with bootstrap support \textgreater{} 0.75.
 A large portion of internal branches are negligibly small, with 30 branches \textless{} 0.00001 substitution rate units, from which only 9 have a bootstrap support \textgreater{} 0.75 (Fig.
@@
 For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
-and contributes phylogenetic data to 134 additional taxa within \emph{Ilex}, from which
-67 are also in the updated Gottlieb2005 tree.
-Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, but their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
-When those sequences are available online, repeating the physcraper run will easily incorporate them into the expanded analysis.
-The Yao2020 tree also lacks 18 taxa that were added by Physcraper.
-This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication.
-Other examples of Physcraper runs on empirical data at a range of taxonomic scales are available in the Physcraper documentation online.
+and contributes phylogenetic data to 134 additional \emph{Ilex} taxa, from which
+67 are also in the Physcraper updated Gottlieb2005 tree.
+While Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
+This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication, but seems to be a ``manual'' process.
 
@@
 Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
-\textbf{MTH: add some more details about what aspect of the PUMPER install failed}
 Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).

### 2020-08-31 2-body.Rmd

@@
 Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using phylogenetic knowledge from public biological databases to update a starting tree and single locus alignments.
+Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using sequence data from public biological databases to update a starting tree and single locus alignments.
 
@@
 Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), etc.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), and other resources.
 The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
 While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
@@
 <!-- *maybe a box of all pipelines?* -->
-However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
 
@@
 The TreeBASE database [@piel2009treebase] hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a public source of valuable expert knowledge.
-Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a non-automated framework linking trees with the alignment that generated them.
+Linking published alignments with molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
+The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a framework linking trees with the alignment that generated them.
 
@@
 The general Physcraper framework is depicted in Figure \@ref(fig:framework).
-It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
+It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
 
@@
 Taxon names in the input tree must be standardized or "mapped" to the unified OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-Users can choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase; @vos2012nexml].
+Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase; @vos2012nexml].
 
@@
 Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
-There are two ways to compare the updated tree with the original tree implemented in Physcraper.
-First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
-Second, a conflict analysis is performed.
-This is a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict Application Programming Interface [@redelings2017supertree].
+The updated tree can be compared with the original tree using an automated conlfict analysis which calculates Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
+Conflict analysis also performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
@@
 Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define the outgroup directly on the updated tree, so trees are accurately rooted.
+The best way right now is for users to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
 <!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->
 
-# Example: The hollies
+# Case Study: The hollies
 
@@
 Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section "DNA sequence search and filtering", aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis performed with OpenTree tools.
+B-D show results of the conflict analysis comparing estimated relationships to taxonomy, performed with OpenTree tools.
 
@@
 <!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
-Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
+
+Physcraper on its own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
 However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
@@
 
-As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the Encyclopedia of Life main goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+Physcraper is novel in that it links molecular and tree data repositories.
+It can also link inferences to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the goal described by [@wilson2003encyclopedia] of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science".
 
-As such, Physcraper can be used to rapidly (in a matter of hours)
-solve challenges overarching both fields of ecology and evolution, such as
+Physcraper in conjuction with OpenTree can be used to rapidly (in a matter of hours)
+to address challenges overarching both fields of ecology and evolution, such as
 placing newly discovered species phylogenetically [@webb2010biodiversity],
@@
 Usually, initial ideas about the data are changed by new analyses.
-Ideally, new understanding of the data can be continually registered on databases
-to consistently expose newcomers to the most up to date knowledge about the data.
+Physcraper can provide context for these ideas by streamlining inferences integrating new data with existing knowledge.
 

### 2020-09-01 2-body.Rmd
word limit includes references :sob:

@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living and extinct organisms and provide key context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
+Phylogenies capture the shared history of living and extinct organisms and provide key evolutionary context for all our biological observations.
+Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Extending existing phylogenies with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
 Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using sequence data from public biological databases to update a starting tree and single locus alignments.
@@
 <!-- [project](https://opentreeoflife.github.io/)  -->
-has constructed a comprehensive and digitally-available tree of life by synthesizing published phylogenies along with taxonomic data.
-Currently, OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
+has constructed a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
+OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
 To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), and other resources.
@@
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
@@
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
+A way to incorporate the benefits from massive amounts of newly released molecular data and fine-grained curation from human experts, is to rely on manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
 The TreeBASE database [@piel2009treebase] hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a public source of valuable expert knowledge.
 Linking published alignments with molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a framework linking trees with the alignment that generated them.
+<!-- The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a framework linking trees with the alignment that generated them.
+-->
 
@@
 
-The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST [@altschul1990basic; @altschul1997gapped], either on the GenBank remote database or in a GenBank local database set up by the user.
+The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST [@altschul1990basic], either on the GenBank remote database or in a GenBank local database set up by the user.
 It is constrained to a taxonomic group in the NCBI taxonomy, defined as the "search taxon".
@@
 The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection; @castresana2002gblocks].
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection].
 Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 
-Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
-The updated tree can be compared with the original tree using an automated conlfict analysis which calculates Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions [@sukumaran2010dendropy].
-Conflict analysis also performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
+Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
+The updated tree can be compared with the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy  [@sukumaran2010dendropy], and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-A suggested default rooting based on OpenTree's taxonomy is implemented for now.
-This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
-However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate.
-Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
+A default rooting based on OpenTree's taxonomy is implemented for now.
+It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
+However, if the updated tree changes expectations from taxonomy, the rooting may no longer be appropriate.
+Automatic identification of a phylogenetic tree root is a difficult problem that has not been solved yet.
 The best way right now is for users to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
@@
 Data repositories designed to preserve and democratize access to biological data
-constitute an essential resource for evolutionary estimation. While data keep accumulating,
+are essential resources for evolutionary estimation. While data keep accumulating,
 our ability to reanalyze and incorporate new data with available knowledge is being challenged.
 
-Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
-Pipelines like Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], and SUPERSMART [@antonelli2017toward]
-focus mainly on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies.
+While phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data are available (e.g., Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]), they focus on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates).
 While Physcraper does not generate phylogenies *de novo* in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
+it successfully generates new phylogenetic knowledge, revealing the potential of
+phylogenetic knowledge published in databases to facilitate phylogenetic placement of public molecular data.
 The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
 pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
+Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to benchmark a comparison.
 <!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
 
-Physcraper on its own generates individual gene trees, which do not capture the full complexity of species' evolutionary history [@song2012resolving].
-However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.
-ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
+Physcraper generates individual gene trees, which fail to capture the complexity of species' evolutionary history [@song2012resolving].
+Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used by any tool that performs coalescent analyses to generate species trees, e.g., ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
 
-Physcraper is novel in that it links molecular and tree data repositories.
-It can also link inferences to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated], a resource that is contributing to the goal described by [@wilson2003encyclopedia] of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science".
+Physcraper not only links molecular and tree databases, it can also link phylogenies to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated].
+<!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+-->
 
 Physcraper in conjuction with OpenTree can be used to rapidly (in a matter of hours)
-to address challenges overarching both fields of ecology and evolution, such as
+address challenges overarching both fields of ecology and evolution, such as
 placing newly discovered species phylogenetically [@webb2010biodiversity],
@@
 
-\newpage
 # References

### 2020-09-01 docs/submission-mee-2020-08-31.tex
word limit includes references :sob:

@@
 
-\textbf{Word count}: 2992
+\textbf{Word count}: 2849
 
@@
 
-Phylogenetic estimates of evolutionary relationships capture the shared history of living and extinct organisms and provide key context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary estimation, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenetic estimates with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using phylogenetic knowledge from public biological databases to update a starting tree and single locus alignments.
+Phylogenies capture the shared history of living and extinct organisms and provide key evolutionary context for all our biological observations.
+Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Extending existing phylogenies with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
+Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using sequence data from public biological databases to update a starting tree and single locus alignments.
 
@@
 The Open Tree of Life project (OpenTree)
-has constructed a comprehensive and digitally-available tree of life by synthesizing published phylogenies along with taxonomic data.
-Currently, OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on their taxonomic assignment.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the World Register of Marine Species (WoRMS), etc.
+has constructed a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
+OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the World Register of Marine Species (WoRMS), and other resources.
 The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, difficulting wide scale phylogenetic reconstructions.
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have already generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
@@
 Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-A way to incorporate the best of two worlds (massive amounts of newly released molecular data and fine-grained curation from human experts) is to rely on published manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
+A way to incorporate the benefits from massive amounts of newly released molecular data and fine-grained curation from human experts, is to rely on manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
 The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a public source of valuable expert knowledge.
-Linking published alignments with public molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
-The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, the Phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}), and metadata linking them to their corresponding alignment repository is available, providing a non-automated framework linking trees with the alignment that generated them.
+Linking published alignments with molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
 
@@
 The general Physcraper framework is depicted in Figure \ref{fig:framework}.
-It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update and its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
+It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
 
@@
 Taxon names in the input tree must be standardized or ``mapped'' to the unified OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Users can choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}).
+Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}).
 
@@
 
-The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}, \protect\hyperlink{ref-altschul1997gapped}{1997}), either on the GenBank remote database or in a GenBank local database set up by the user.
+The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}), either on the GenBank remote database or in a GenBank local database set up by the user.
 It is constrained to a taxonomic group in the NCBI taxonomy, defined as the ``search taxon''.
@@
 The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}, \protect\hyperlink{ref-castresana2002gblocks}{2002}).
+It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}).
 Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
@@
 
-Physcraper's final result is an updated phylogenetic hypothesis for the search taxon based on the data in the input alignment.
-There are two ways to compare the updated tree with the original tree implemented in Physcraper.
-First, Robinson Foulds weighted and unweighted metrics are estimated using Dendropy functions (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
-Second, a conflict analysis is performed.
-This is a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict Application Programming Interface (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
+Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
+The updated tree can be compared with the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-A suggested default rooting based on OpenTree's taxonomy is implemented for now.
-This approach uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
-However, if the updated tree changes expectations from taxonomy, the root may no longer be appropriate.
-Automatic identification of a phylogenetic tree root is indeed a difficult problem that has not been solved yet.
-The best way right now is for users to define the outgroup directly on the updated tree, so trees are accurately rooted.
+A default rooting based on OpenTree's taxonomy is implemented for now.
+It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
+However, if the updated tree changes expectations from taxonomy, the rooting may no longer be appropriate.
+Automatic identification of a phylogenetic tree root is a difficult problem that has not been solved yet.
+The best way right now is for users to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
 
-\hypertarget{example-the-hollies}{%
-\section{Example: The hollies}\label{example-the-hollies}}
+\hypertarget{case-study-the-hollies}{%
+\section{Case Study: The hollies}\label{case-study-the-hollies}}
 
@@
 Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis performed with OpenTree tools.
+B-D show results of the conflict analysis comparing estimated relationships to taxonomy, performed with OpenTree tools.
 
@@
 Data repositories designed to preserve and democratize access to biological data
-constitute an essential resource for evolutionary estimation. While data keep accumulating,
+are essential resources for evolutionary estimation. While data keep accumulating,
 our ability to reanalyze and incorporate new data with available knowledge is being challenged.
 
-Various phylogenetic pipelines have been designed to make evolutionary sense of the vast amount of public molecular data available in different ways.
-Pipelines like Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), and SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})
-focus mainly on generating full trees \emph{de novo} (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates) for regions of the Tree of
-Life that have no phylogenetic assessment yet in published studies.
+While phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data are available (e.g., Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})), they focus on generating full trees \emph{de novo} (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates).
 While Physcraper does not generate phylogenies \emph{de novo} in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of pre-existing
-phylogenetic knowledge available in tree repositories to facilitate phylogenetic placement of public molecular data.
+it successfully generates new phylogenetic knowledge, revealing the potential of
+phylogenetic knowledge published in databases to facilitate phylogenetic placement of public molecular data.
 The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
 pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to conduct a comparison analysis to Physcraper.
-Physcraper on it's own generates individual gene trees, which do not capture the full complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-However, using Physcraper it is straightforward to gather alignments and gene trees for multiple loci for a taxonomic group of interest, which can be used as inputs for downstream tools to perform coalescent analyses and assess species tree relationships, (e.g.~ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
+Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to benchmark a comparison.
 
-As far as we know, Physcraper is the only phylogenetic pipeline that links molecular and tree data repositories.
-It can also link to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), a resource that is contributing to the Encyclopedia of Life main goal of ``exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science'' (Wilson \protect\hyperlink{ref-wilson2003encyclopedia}{2003}).
+Physcraper generates individual gene trees, which fail to capture the complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used by any tool that performs coalescent analyses to generate species trees, e.g., ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
 
-As such, Physcraper can be used to rapidly (in a matter of hours)
-solve challenges overarching both fields of ecology and evolution, such as
+Physcraper not only links molecular and tree databases, it can also link phylogenies to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
+
+Physcraper in conjuction with OpenTree can be used to rapidly (in a matter of hours)
+address challenges overarching both fields of ecology and evolution, such as
 placing newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
@@
 
-\newpage
-
 \hypertarget{references}{%
-\section{References}\label{references}}
-
-\newpage
-\begin{center}
-\textsc{References}
-\end{center}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of molecular biology}, \textbf{215}, 403--410.
-
-\leavevmode\hypertarget{ref-altschul1997gapped}{}%
-Altschul, S.F., Madden, T.L., Schäffer, A.A., Zhang, J., Zhang, Z., Miller, W.
-\& Lipman, D.J.
-(1997).
-Gapped blast and psi-blast: A new generation of protein database search programs.
-\emph{Nucleic acids research}, \textbf{25}, 3389--3402.
 
@@
 \emph{Frontiers in Genetics}, \textbf{10}.
-Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7047930/}
 
@@
 \emph{PLOS Computational Biology}, \textbf{15}, e1006650.
-Retrieved August 18, 2020, from \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1006650}
 
@@
 Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC bioinformatics}, \textbf{10}, 421.
-
-\leavevmode\hypertarget{ref-castresana2002gblocks}{}%
-Castresana, J. (2002). GBLOCKS: Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Version 0.91 b. Copyrighted by J. Castresana, EMBL}.
 
@@
 \leavevmode\hypertarget{ref-chifman2014quartet}{}%
-Chifman, J.
-\& Kubatko, L.
-(2014).
-Quartet Inference from SNP Data Under the Coalescent Model.
-\emph{Bioinformatics}, \textbf{30}, 3317--3324.
-Retrieved August 18, 2020, from \url{https://academic.oup.com/bioinformatics/article/30/23/3317/206559}
+Chifman, J. \& Kubatko, L. (2014). Quartet Inference from SNP Data Under the Coalescent Model. \emph{Bioinformatics}, \textbf{30}, 3317--3324.
 
@@
 (2000).
-Molecular phylogeny and biogeography of the genus ilex l.(Aquifoliaceae).
+Molecular phylogeny and biogeography of the genus Ilex L.(Aquifoliaceae).
 \emph{Annals of Botany}, \textbf{85}, 111--122.
@@
 (2005).
-Molecular analyses of the genus ilex (aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
+Molecular analyses of the genus Ilex (Aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
 \emph{American Journal of Botany}, \textbf{92}, 352--369.
@@
 \leavevmode\hypertarget{ref-jones2016targeted}{}%
-Jones, M.R.
-\& Good, J.M.
-(2016).
-TARGETED CAPTURE IN EVOLUTIONARY AND ECOLOGICAL GENOMICS.
-\emph{Molecular ecology}, \textbf{25}, 185--202.
-Retrieved July 28, 2020, from \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4823023/}
+Jones, M.R. \& Good, J.M. (2016). TARGETED capture in evolutionary and ecological genomics. \emph{Molecular ecology}, \textbf{25}, 185--202.
 
@@
 (2010).
-The history of extant ilex species (aquifoliaceae): Evidence of hybridization within a miocene radiation.
+The history of extant Ilex species (Aquifoliaceae): Evidence of hybridization within a miocene radiation.
 \emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
 
-\leavevmode\hypertarget{ref-mctavish2015phylesystem}{}%
-McTavish, E.J., Hinchliff, C.E., Allman, J.F., Brown, J.W., Cranston, K.A., Holder, M.T., Rees, J.A.
-\& Smith, S.A.
-(2015).
-Phylesystem: A git-based data store for community-curated phylogenetic estimates.
-\emph{Bioinformatics}, \textbf{31}, 2794--2800.
-
 \leavevmode\hypertarget{ref-mirarab2014astral}{}%
-Mirarab, S., Reaz, R., Bayzid, M.S., Zimmermann, T., Swenson, M.S.
-\& Warnow, T.
-(2014).
-ASTRAL: Genome-scale coalescent-based species tree estimation.
-\emph{Bioinformatics}, \textbf{30}, i541--i548.
-Retrieved July 25, 2015, from \url{http://bioinformatics.oxfordjournals.org/content/30/17/i541}
+Mirarab, S., Reaz, R., Bayzid, M.S., Zimmermann, T., Swenson, M.S. \& Warnow, T. (2014). ASTRAL: Genome-scale coalescent-based species tree estimation. \emph{Bioinformatics}, \textbf{30}, i541--i548.
 
@@
 (2009).
-New information for ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (aquifoliaceae).
+New information for Ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (Aquifoliaceae).
 \emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.

### 2020-09-01 2-body.Rmd
reducing references

@@
 OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank; @wheeler2000database], the Global Biodiversity Information Facility [GBIF; @secretariat2017gbif], the World Register of Marine Species (WoRMS), and other resources.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank], the Global Biodiversity Information Facility (GBIF), the World Register of Marine Species (WoRMS), and other resources.
 The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes vs microsatellites), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide; @jones2016targeted], decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide], decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
@@
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
-Assembling a DNA alignment from such a massive database can be done "by hand", but it requires huge amounts of time and it is mostly a non-reproducible approach.
+Assembling a DNA alignment from such a massive database can be done "by hand", but it is a largely time consuming and mostly non-reproducible approach.
 Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
@@
 Taxon names in the input tree must be standardized or "mapped" to the unified OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase; @vos2012nexml].
+Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase].
 

### 2020-09-01 docs/submission-mee-2020-09-01.tex
reducing references

--- 2020-09-01 docs/submission-mee-2020-08-31.tex
@@
 
-\textbf{Word count}: 2849
+\textbf{Word count}: 2841
 
@@
 OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), the Global Biodiversity Information Facility (GBIF; Secretariat \protect\hyperlink{ref-secretariat2017gbif}{2017}), the World Register of Marine Species (WoRMS), and other resources.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), the Global Biodiversity Information Facility (GBIF), the World Register of Marine Species (WoRMS), and other resources.
 The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
@@
 Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes, microsatellites, ultra-conserved elements), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Jones \& Good \protect\hyperlink{ref-jones2016targeted}{2016}; Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
+Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes vs microsatellites), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
 Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
@@
 These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
-Assembling a DNA alignment from such a massive database can be done ``by hand'', but it requires huge amounts of time and it is mostly a non-reproducible approach.
+Assembling a DNA alignment from such a massive database can be done ``by hand'', but it is a largely time consuming and mostly non-reproducible approach.
 Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
@@
 Taxon names in the input tree must be standardized or ``mapped'' to the unified OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}).
+Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
 
@@
 \leavevmode\hypertarget{ref-altschul1990basic}{}%
-Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of molecular biology}, \textbf{215}, 403--410.
+Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of Molecular Biology}, \textbf{215}, 403--410.
 
@@
 (2020).
-A Guide to Carrying Out a Phylogenomic Target Sequence Capture Project.
-\emph{Frontiers in Genetics}, \textbf{10}.
+A guide to carrying out a phylogenomic target sequence capture project.
+\emph{Frontiers in Genetics}, \textbf{10}, 1--20.
 
@@
 \leavevmode\hypertarget{ref-benson2000genbank}{}%
-Benson, D.A., Karsch-Mizrachi, I., Lipman, D.J., Ostell, J., Rapp, B.A. \& Wheeler, D.L. (2000). GenBank. \emph{Nucleic acids research}, \textbf{28}, 15--18.
+Benson, D.A., Karsch-Mizrachi, I., Lipman, D.J., Ostell, J., Rapp, B.A. \& Wheeler, D.L. (2000). GenBank. \emph{Nucleic Acids Research}, \textbf{28}, 15--18.
 
@@
 \leavevmode\hypertarget{ref-camacho2009blast}{}%
-Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC bioinformatics}, \textbf{10}, 421.
+Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC Bioinformatics}, \textbf{10}, 421.
 
 \leavevmode\hypertarget{ref-castresana2000selection}{}%
-Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular biology and evolution}, \textbf{17}, 540--552.
+Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular Biology and Evolution}, \textbf{17}, 540--552.
 
 \leavevmode\hypertarget{ref-chifman2014quartet}{}%
-Chifman, J. \& Kubatko, L. (2014). Quartet Inference from SNP Data Under the Coalescent Model. \emph{Bioinformatics}, \textbf{30}, 3317--3324.
+Chifman, J. \& Kubatko, L. (2014). Quartet inference from SNP data under the coalescent model. \emph{Bioinformatics}, \textbf{30}, 3317--3324.
 
@@
 (2009).
-Biopython: Freely available python tools for computational molecular biology and bioinformatics.
+Biopython: freely available Python tools for computational molecular biology and bioinformatics.
 \emph{Bioinformatics}, \textbf{25}, 1422--1423.
@@
 (2000).
-Molecular phylogeny and biogeography of the genus Ilex L.(Aquifoliaceae).
+Molecular phylogeny and biogeography of the genus \emph{Ilex} L.(Aquifoliaceae).
 \emph{Annals of Botany}, \textbf{85}, 111--122.
@@
 \leavevmode\hypertarget{ref-edgar2004muscle}{}%
-Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic acids research}, \textbf{32}, 1792--1797.
+Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic Acids Research}, \textbf{32}, 1792--1797.
 
@@
 (2017).
-A pilot study applying the plant anchored hybrid enrichment method to new world sages (salvia subgenus calosphace; lamiaceae).
+A pilot study applying the plant Anchored Hybrid Enrichment method to New World sages (\emph{Salvia} subgenus Calosphace; Lamiaceae).
 \emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
@@
 (2005).
-Molecular analyses of the genus Ilex (Aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
+Molecular analyses of the genus \emph{Ilex} (Aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
 \emph{American Journal of Botany}, \textbf{92}, 352--369.
@@
 
-\leavevmode\hypertarget{ref-jones2016targeted}{}%
-Jones, M.R. \& Good, J.M. (2016). TARGETED capture in evolutionary and ecological genomics. \emph{Molecular ecology}, \textbf{25}, 185--202.
-
 \leavevmode\hypertarget{ref-manen2010history}{}%
@@
 (2010).
-The history of extant Ilex species (Aquifoliaceae): Evidence of hybridization within a miocene radiation.
+The history of extant \emph{Ilex} species (Aquifoliaceae): evidence of hybridization within a Miocene radiation.
 \emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
@@
 
-\leavevmode\hypertarget{ref-secretariat2017gbif}{}%
-Secretariat, G. (2017). GBIF backbone taxonomy. \emph{Checklist Dataset {[}cited 2017 Nov 14{]}. doi}, \textbf{10}.
-
 \leavevmode\hypertarget{ref-selbach2009new}{}%
@@
 (2009).
-New information for Ilex phylogenetics based on the plastid psbA-trnH intergenic spacer (Aquifoliaceae).
+New information for \emph{Ilex} phylogenetics based on the plastid psbA-trnH intergenic spacer (Aquifoliaceae).
 \emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.
@@
 (2000).
-Intersectional gene flow between insular endemics of ilex (aquifoliaceae) on the bonin islands and the ryukyu islands.
+Intersectional gene flow between insular endemics of \emph{Ilex} (Aquifoliaceae) on the Bonin Islands and the Ryukyu Islands.
 \emph{American Journal of Botany}, \textbf{87}, 793--810.
@@
 Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
-\emph{BMC evolutionary biology}, \textbf{9}, 37.
+\emph{BMC Evolutionary Biology}, \textbf{9}, 37.
 
@@
 Making tree-of-life knowledge accessible, reusable and convenient.
-\emph{BMC bioinformatics}, \textbf{14}, 158.
+\emph{BMC Bioinformatics}, \textbf{14}, 158.
 
 \leavevmode\hypertarget{ref-sukumaran2010dendropy}{}%
-Sukumaran, J. \& Holder, M.T. (2010). DendroPy: A python library for phylogenetic computing. \emph{Bioinformatics}, \textbf{26}, 1569--1571.
-
-\leavevmode\hypertarget{ref-vos2012nexml}{}%
-Vos, R.A., Balhoff, J.P., Caravas, J.A., Holder, M.T., Lapp, H., Maddison, W.P., Midford, P.E., Priyam, A., Sukumaran, J., Xia, X.
-\& others.
-(2012).
-NeXML: Rich, extensible, and verifiable representation of comparative data and metadata.
-\emph{Systematic biology}, \textbf{61}, 675--689.
+Sukumaran, J. \& Holder, M.T. (2010). DendroPy: a Python library for phylogenetic computing. \emph{Bioinformatics}, \textbf{26}, 1569--1571.
 
 \leavevmode\hypertarget{ref-webb2010biodiversity}{}%
-Webb, C.O., Slik, J.F. \& Triono, T. (2010). Biodiversity inventory and informatics in southeast asia. \emph{Biodiversity and Conservation}, \textbf{19}, 955--972.
-
-\leavevmode\hypertarget{ref-wheeler2000database}{}%
-Wheeler, D.L., Chappey, C., Lash, A.E., Leipe, D.D., Madden, T.L., Schuler, G.D., Tatusova, T.A.
-\& Rapp, B.A.
-(2000).
-Database resources of the national center for biotechnology information.
-\emph{Nucleic acids research}, \textbf{28}, 10--14.
+Webb, C.O., Slik, J.F. \& Triono, T. (2010). Biodiversity inventory and informatics in Southeast Asia. \emph{Biodiversity and Conservation}, \textbf{19}, 955--972.
 
 \leavevmode\hypertarget{ref-yao2020phylogeny}{}%
-Yao, X., Song, Y., Yang, J.-B., Tan, Y.-H. \& Corlett, R.T. (2020). Phylogeny and biogeography of the hollies (ilex l., aquifoliaceae). \emph{Journal of Systematics and Evolution}.
+Yao, X., Song, Y., Yang, J.-B., Tan, Y.-H.
+\& Corlett, R.T.
+(2020).
+Phylogeny and biogeography of the hollies (\emph{Ilex} L., Aquifoliaceae).
+\emph{Journal of Systematics and Evolution}, \textbf{58}, 1--10.
 

### 2020-09-02 2-body.Rmd
comments to figure 1

@@
 
-Phylogenies capture the shared history of living and extinct organisms and provide key evolutionary context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Phylogenies capture the shared history of organisms and provide key evolutionary context for our biological observations.
+Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of public molecular data has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenies with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using sequence data from public biological databases to update a starting tree and single locus alignments.
+Here, we introduce Physcraper, a tool to build upon DNA datasets that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using data from public DNA databases to update a starting tree and single locus alignments.
 
-Automatic integration of data in a phylogenetic context is challenged by the prevalence of taxonomic idiosyncrasies across biological databases.
-Taxonomic name standardization represents a step toward tackling this issue, but requires a unified system.
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic integration of data into phylogenies; challenge that can be addressed with a unified taxonomy for name standardization.
 The Open Tree of Life project (OpenTree)
 <!-- [project](https://opentreeoflife.github.io/)  -->
-has constructed a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
-OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank], the Global Biodiversity Information Facility (GBIF), the World Register of Marine Species (WoRMS), and other resources.
-The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+constructs a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
+OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank], among others.
+This OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
+<!-- to phylogenetic data that has been standardized to it. -->
 
-Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving].
-Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes vs microsatellites), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+Another challenhe for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
 <!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide], decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
-Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions [@fragoso2017pilot].
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative characters from independent and single-copy genetic markers [@andermann2020guide].
+Yet, decades of single locus sequencing have generated massive amounts of homologous DNA datasets that can be used for phylogenetic reconstruction at many scales.
+Moreover, species tree reconstructions from multiple single locus datasets using the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving].
 
-More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most (`r round(236023/240708, digits=4)*100`%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
+More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota].
 Assembling a DNA alignment from such a massive database can be done "by hand", but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
+Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to infer phylogenetic relationships in a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
 <!-- *maybe a box of all pipelines?* -->
+However, thoughtfully curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic reconstructions [@fragoso2017pilot].
 
 A way to incorporate the benefits from massive amounts of newly released molecular data and fine-grained curation from human experts, is to rely on manually curated homology hypotheses as "jump-start" alignments [@morrison2006multiple].
-The TreeBASE database [@piel2009treebase] hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.), representing a public source of valuable expert knowledge.
-Linking published alignments with molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
+The TreeBASE database [@piel2009treebase] hosts about 8,200 alignments from published peer-reviewed studies, providing information on evolutionary relationships of around 100,000 distinct taxa [(see TreeBASE's website about)](https://www.treebase.org/treebase-web/home.html#:~:text=TreeBASE%20is%20produced%20and%20governed,mapped%20to%20104%2C593%20distinct%20taxa.).
+Linking published alignments generated by experts with molecular data that has not yet been included in any public phylogenetic estimate has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
 <!-- The phylogenies associated with TreeBASE alignments have been integrated to the OpenTree's datastore, Phylesystem [@mctavish2015phylesystem], and metadata linking them to their corresponding alignment repository is available, providing a framework linking trees with the alignment that generated them.
@@
 
-Physcraper relies on programmatic access protocols (API's) to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols (API's) and the OpenTree taxonomy to automatically link phylogenetic data to molecular data from GenBank, alignments from TreeBASE, and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the standardized taxonomy as a way to link taxon data from different databases.
+Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the OpenTree taxonomy as a way to link taxon data from different databases.
 This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.
@@
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-knitr::include_graphics("docs/figs/schematic.pdf")
+knitr::include_graphics("docs/figs/schematic-3.png")
 ```
@@
 The general Physcraper framework is depicted in Figure \@ref(fig:framework).
-It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
+It consists of 4 steps: 1) identifying and processing a tree and its underlying alignment; 2) performing a BLAST search of DNA sequences from the original alignment on GenBank, and filtering of new sequences; 3) profile-aligning new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates.
 
@@
 
-Taxon names in the input tree must be standardized or "mapped" to the unified OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase].
+Taxon names in the input tree must be standardized to the OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
+Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase].
 
-The input alignment should be a single locus alignment that was used in part or in whole, to generate the input tree.
-If the alignment associated to the input tree is stred in TreeBASE, Physcraper retrieves it automatically.
-Alignments stored in any other repository, or constituting personal data have to be downloaded by the user.
-
-Physcraper processes the input tree to match taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name and its corresponding sequence in the alignment are needed to continue the algorithm.
+The input alignment is a single locus alignment that was used in part or in whole to generate the input tree.
+If it is stored in TreeBASE, Physcraper retrieves it automatically.
+Otherwise, the user must provide the path to a local copy of the alignment.
+<!-- Alignments stored in any other repository, or constituting personal data have to be stored locally. -->
+Only taxa that are both in the sequence alignment and in the tree are considered further for analysis; at least one taxon and its corresponding sequence are required.
 
@@
 
-The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST [@altschul1990basic], either on the GenBank remote database or in a GenBank local database set up by the user.
-It is constrained to a taxonomic group in the NCBI taxonomy, defined as the "search taxon".
-Users can arbitrarily define a search taxon that is either a more or a less inclusive
-clade relative to the ingroup of the input tree.
-Otherwise, the search taxon is automatically identified using the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated], as the Most Recent Common Ancestral Taxon (MRCAT) of the ingroup taxa in the input tree, i.e., the MRCA that is also a named clade in the NCBI taxonomy (Fig.
+The Basic Local Alignment Search Tool, BLAST [@altschul1990basic] is used for DNA sequence search on a remote or local GenBank database.
+It is constrained to a "search taxon", a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated], as the most recent common ancestor of the ingroup taxa in the tree that is also a named clade in the NCBI taxonomy (Fig.
 \@ref(fig:framework)).
   <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
-The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
+<!-- The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.  -->
+Alternatively, users can arbitrarily define a search taxon that is either a more or less inclusive clade relative to the ingroup of the input tree.
 
-BLAST command line tools `blastn` function [@camacho2009blast] is used for local database sequence searches.
-For remote database searches, we modified the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) to accept an alternative BLAST address (URL).
-This is useful when a user lacks access to the computer capacity needed to setup a local database, but has access to an institutional server.
-Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank within the search taxon.
-GenBank sequences with match scores smaller than the e-value cutoff (default to 0.00001) are downloaded into a local library.
-
-Downlodaded sequences are not further considered in the analysis if they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OTT taxonomy or the NCBI taxonomy.
-If there are several sequences per taxon, an arbitrary number of 5 (can be modified by the user) sequences per taxon are chosen at random.
+BLAST is implemented with the BLAST command line tools `blastn` function [@camacho2009blast] and the BioPython [@cock2009biopython] BLAST function from the [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) modified to accept an alternative BLAST address.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
+New sequences are excuded for analysis if 1) they do not belong to the search taxon; 2) they have an e-value above the cutoff (default to 0.00001); 3) they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OpenTree taxonomy or the NCBI taxonomy.
+An arbitrary maximum number of randomly chosen sequences per taxon are allowed (default to 5).
 
 Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions [@cock2009biopython].
-Iterative cycles of BLAST searches can be performed, by blasting the new sequences until no new ones are found.
-By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
+Iterative cycles of BLAST searches can be performed, by blasting all new sequences until no new ones are found. By default only one BLAST cycle is performed.
+<!-- in which only sequences in the input alignment are BLASTed. -->
 
@@
 
-By default, Physcraper uses MUSCLE [@edgar2004muscle] to perform DNA sequence alignments in a two step process. First, all new sequences are aligned using the default MUSCLE options.
-
-Second, a MUSCLE profile alignment is performed, in which the original alignment is used as a template to align the new sequences. This ensures that the final alignment
-follows the homology criteria established by the original alignment.
-The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing [e.g., GBlocks; @castresana2000selection].
-Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
+MUSCLE [@edgar2004muscle] is used to perform a profile alignment in which the original alignment is used as a template of homology criteria to align new sequences.
+The final alignment is not further automatically checked, and additional inspection and refinement are recommended.
 
@@
 
-A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with default settings (GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm).
-Only the number of bootsrap replicates can be defined by the user.
-By default, the original tree is used as a starting tree for the ML searches. Alternatively, the original tree can be set as full topological constraint, or simply be ignored for the ML searches.
-Bootstrap results are summarized with the SumTrees module of DendroPy [current version 4.4.0; @sukumaran2010dendropy].
+A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML [@stamatakis2014raxml] with default settings (GTRCAT model and 100 bootstrap replicates with the default algorithm) and using the input tree as a starting tree for the ML searches.
+Bootstrap results are summarized with DendroPy's SumTrees module [current version 4.4.0; @sukumaran2010dendropy].
 
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-The updated tree can be compared with the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy  [@sukumaran2010dendropy], and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
-For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-A default rooting based on OpenTree's taxonomy is implemented for now.
-It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
+The updated tree is compared to the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy  [@sukumaran2010dendropy], and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
+For the conflict analysis to be meaningful, the root of the tree needs to be accurately identified.
+The best way to currenty do this is by hand, although a default rooting based on OpenTree's taxonomy is also available.
+<!-- It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
 However, if the updated tree changes expectations from taxonomy, the rooting may no longer be appropriate.
 Automatic identification of a phylogenetic tree root is a difficult problem that has not been solved yet.
-The best way right now is for users to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
+The best way is to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
+-->
 <!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->
@@
 
-A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
 Figure \@ref(fig:results) shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from @gottlieb2005molecular.
@@
 Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section "DNA sequence search and filtering", aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis comparing estimated relationships to taxonomy, performed with OpenTree tools.
+B-D show results of the conflict analysis performed with OpenTree tools.
 
@@
 
-Data repositories designed to preserve and democratize access to biological data
-are essential resources for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data with available knowledge is being challenged.
+Databases that preserve and democratize access to biological data
+are essential resources for our scientific endeavor.
+New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are essential.
 
-While phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data are available (e.g., Phylota [@sanderson2008phylota], PyPHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]), they focus on generating full trees *de novo* (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates).
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota [@sanderson2008phylota], PHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]) focus on generating full phylogenies *de novo*, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
 While Physcraper does not generate phylogenies *de novo* in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of
-phylogenetic knowledge published in databases to facilitate phylogenetic placement of public molecular data.
+it successfully generates new phylogenetic knowledge, revealing the importance of
+open science in facilitating phylogenetic placement of public molecular data and accelerating the enrichment and updating of phylogenetic relationships in any region of the tree of life.
 The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
@@
 Physcraper generates individual gene trees, which fail to capture the complexity of species' evolutionary history [@song2012resolving].
-Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used by any tool that performs coalescent analyses to generate species trees, e.g., ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], SVD Quartets [@chifman2014quartet]).
+Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used to perform coalescent analysis to generate species trees with e.g., ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], or SVD Quartets [@chifman2014quartet]).
 
-Physcraper not only links molecular and tree databases, it can also link phylogenies to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree [@rees2017automated].
+Physcraper has the potential to link phylogenies to data available in any of the taxonomies integrated by the OpenTree taxonomy [@rees2017automated], such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
 <!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
 -->
-
-Physcraper in conjuction with OpenTree can be used to rapidly (in a matter of hours)
+The Physcraper workflow can be used to rapidly (in a matter of hours)
 address challenges overarching both fields of ecology and evolution, such as
@@
 systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
-and generating custom species trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
+and generating custom trees for ecological and evolutionary downstream analyses [@stoltzfus2013phylotastic].
 

### 2020-09-02 docs/submission-mee-2020-09-02.tex
comments to figure 1

--- 2020-09-01 docs/submission-mee-2020-09-01.tex
@@
 
-\textbf{Word count}: 2841
+\textbf{Word count}: 2419
 
@@
 
-Phylogenies capture the shared history of living and extinct organisms and provide key evolutionary context for all our biological observations.
-Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of molecular data publicly available has never been incorporated into any phylogenetic estimate.
+Phylogenies capture the shared history of organisms and provide key evolutionary context for our biological observations.
+Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of public molecular data has never been incorporated into any phylogenetic estimate.
 Extending existing phylogenies with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool to build upon molecular data that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using sequence data from public biological databases to update a starting tree and single locus alignments.
+Here, we introduce Physcraper, a tool to build upon DNA datasets that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using data from public DNA databases to update a starting tree and single locus alignments.
 
-Automatic integration of data in a phylogenetic context is challenged by the prevalence of taxonomic idiosyncrasies across biological databases.
-Taxonomic name standardization represents a step toward tackling this issue, but requires a unified system.
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic integration of data into phylogenies; challenge that can be addressed with a unified taxonomy for name standardization.
 The Open Tree of Life project (OpenTree)
-has constructed a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
-OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenetic estimates - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), the Global Biodiversity Information Facility (GBIF), the World Register of Marine Species (WoRMS), and other resources.
-The OpenTree unified taxonomy represents a key resource for connecting data from virtually any biological database to phylogenetic data that has been standardized to OpenTree's unified taxonomy.
+constructs a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
+OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
+To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), among others.
+This OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
 
-Another challenge to incorporating molecular data from public databases to update phylogenetic knowledge is assembling high-quality homology hypotheses.
-Species tree reconstructions from multiple single locus data sets taking into account the multispecies coalescent model are considered the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-Genomics has, and will continue to, revolutionize phylogenetic inference.
-Yet, different research questions call for different genomic sequencing approaches (e.g., whole genomes vs microsatellites), which produce largely non-overlapping genomic data sets across taxa, creating challenges in wide scale phylogenetic reconstruction.
-While phylogenomics ameliorates the problem of non-overlapping genomic data sets by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}), decades of single locus sequencing have generated massive amounts of homologous DNA data sets that can be used for phylogenetic reconstruction at many scales.
-Even in phylogenomic reconstructions, thoughtfully curated markers and alignments can improve phylogenetic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+Another challenhe for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}).
+Yet, decades of single locus sequencing have generated massive amounts of homologous DNA datasets that can be used for phylogenetic reconstruction at many scales.
+Moreover, species tree reconstructions from multiple single locus datasets using the multispecies coalescent model are considered the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
 
-More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences.
-These sequences were gauged to have the potential to resolve phylogenetic relationships of most (98.05\%) of the almost 241, 000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of 98.05\% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 Assembling a DNA alignment from such a massive database can be done ``by hand'', but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to study evolutionary relationships among a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to infer phylogenetic relationships in a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+However, thoughtfully curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
 A way to incorporate the benefits from massive amounts of newly released molecular data and fine-grained curation from human experts, is to rely on manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8, 200 alignments, providing information on evolutionary relationships of around 100, 000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}, representing a public source of valuable expert knowledge.
-Linking published alignments with molecular data that has not yet been included in any public phylogenetic estimate, has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
+The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8,200 alignments from published peer-reviewed studies, providing information on evolutionary relationships of around 100,000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
+Linking published alignments generated by experts with molecular data that has not yet been included in any public phylogenetic estimate has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
 
-Physcraper relies on programmatic access protocols (API's) to automatically link molecular data from GenBank to alignments from TreeBASE and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
+Physcraper relies on programmatic access protocols (API's) and the OpenTree taxonomy to automatically link phylogenetic data to molecular data from GenBank, alignments from TreeBASE, and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
 Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the standardized taxonomy as a way to link taxon data from different databases.
+Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the OpenTree taxonomy as a way to link taxon data from different databases.
 This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic}
+{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic-3}
 
@@
 The general Physcraper framework is depicted in Figure \ref{fig:framework}.
-It consists of 4 steps: 1) identifying and processing a phylogenetic tree to update its underlying alignment; 2) performing a constrained BLAST search of DNA sequences in the original alignment on the GenBank database, and filtering of new sequences; 3) profile-aligning filtered new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates within the focus group.
+It consists of 4 steps: 1) identifying and processing a tree and its underlying alignment; 2) performing a BLAST search of DNA sequences from the original alignment on GenBank, and filtering of new sequences; 3) profile-aligning new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates.
 
@@
 
-Taxon names in the input tree must be standardized or ``mapped'' to the unified OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Users can upload their own tree, or choose from among the 2, 950 mapped trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
+Taxon names in the input tree must be standardized to the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
 
-The input alignment should be a single locus alignment that was used in part or in whole, to generate the input tree.
-If the alignment associated to the input tree is stred in TreeBASE, Physcraper retrieves it automatically.
-Alignments stored in any other repository, or constituting personal data have to be downloaded by the user.
-
-Physcraper processes the input tree to match taxa found in the alignment, and verifies that all taxon names on the tips of the tree are in the DNA character matrix and vice versa.
-Technically, just one taxon name and its corresponding sequence in the alignment are needed to continue the algorithm.
+The input alignment is a single locus alignment that was used in part or in whole to generate the input tree.
+If it is stored in TreeBASE, Physcraper retrieves it automatically.
+Otherwise, the user must provide the path to a local copy of the alignment.
+Only taxa that are both in the sequence alignment and in the tree are considered further for analysis; at least one taxon and its corresponding sequence are required.
 
@@
 
-The DNA sequence search is performed with the Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}), either on the GenBank remote database or in a GenBank local database set up by the user.
-It is constrained to a taxonomic group in the NCBI taxonomy, defined as the ``search taxon''.
-Users can arbitrarily define a search taxon that is either a more or a less inclusive
-clade relative to the ingroup of the input tree.
-Otherwise, the search taxon is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the Most Recent Common Ancestral Taxon (MRCAT) of the ingroup taxa in the input tree, i.e., the MRCA that is also a named clade in the NCBI taxonomy (Fig.
+The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}) is used for DNA sequence search on a remote or local GenBank database.
+It is constrained to a ``search taxon'', a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the most recent common ancestor of the ingroup taxa in the tree that is also a named clade in the NCBI taxonomy (Fig.
 \ref{fig:framework}).
-The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.
+Alternatively, users can arbitrarily define a search taxon that is either a more or less inclusive clade relative to the ingroup of the input tree.
 
-BLAST command line tools \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) is used for local database sequence searches.
-For remote database searches, we modified the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} to accept an alternative BLAST address (URL).
-This is useful when a user lacks access to the computer capacity needed to setup a local database, but has access to an institutional server.
-Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank within the search taxon.
-GenBank sequences with match scores smaller than the e-value cutoff (default to 0.00001) are downloaded into a local library.
-
-Downlodaded sequences are not further considered in the analysis if they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80\% and 120\%, respectively); or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OTT taxonomy or the NCBI taxonomy.
-If there are several sequences per taxon, an arbitrary number of 5 (can be modified by the user) sequences per taxon are chosen at random.
+BLAST is implemented with the BLAST command line tools \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) and the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} modified to accept an alternative BLAST address.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
+New sequences are excuded for analysis if 1) they do not belong to the search taxon; 2) they have an e-value above the cutoff (default to 0.00001); 3) they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80\% and 120\%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OpenTree taxonomy or the NCBI taxonomy.
+An arbitrary maximum number of randomly chosen sequences per taxon are allowed (default to 5).
 
 Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
-Iterative cycles of BLAST searches can be performed, by blasting the new sequences until no new ones are found.
-By default only one BLAST search cycle is performed in which only sequences in the input alignment are blasted.
+Iterative cycles of BLAST searches can be performed, by blasting all new sequences until no new ones are found. By default only one BLAST cycle is performed.
 
@@
 
-By default, Physcraper uses MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) to perform DNA sequence alignments in a two step process.
-First, all new sequences are aligned using the default MUSCLE options.
-
-Second, a MUSCLE profile alignment is performed, in which the original alignment is used as a template to align the new sequences. This ensures that the final alignment
-follows the homology criteria established by the original alignment.
-The final alignment is not further processed by Physcraper.
-It is recommended that the alignment is checked by the user, by eye followed by manual refinement, or using a tool for automatic alignment processing (e.g., GBlocks; Castresana \protect\hyperlink{ref-castresana2000selection}{2000}).
-Users may also use Physcraper to only gather new GenBank sequences, to then apply their own preferred alignment and phylogenetic inference methods.
+MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used to perform a profile alignment in which the original alignment is used as a template of homology criteria to align new sequences.
+The final alignment is not further automatically checked, and additional inspection and refinement are recommended.
 
@@
 
-A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings (GTRCAT model of molecular evolution and 100 bootstrap replicates with the default algorithm).
-Only the number of bootsrap replicates can be defined by the user.
-By default, the original tree is used as a starting tree for the ML searches. Alternatively, the original tree can be set as full topological constraint, or simply be ignored for the ML searches.
-Bootstrap results are summarized with the SumTrees module of DendroPy (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings (GTRCAT model and 100 bootstrap replicates with the default algorithm) and using the input tree as a starting tree for the ML searches.
+Bootstrap results are summarized with DendroPy's SumTrees module (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-The updated tree can be compared with the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
-For the conflict analysis to be meaningful, the root of the tree needs to be accurately defined.
-A default rooting based on OpenTree's taxonomy is implemented for now.
-It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
-However, if the updated tree changes expectations from taxonomy, the rooting may no longer be appropriate.
-Automatic identification of a phylogenetic tree root is a difficult problem that has not been solved yet.
-The best way right now is for users to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted.
+The updated tree is compared to the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
+For the conflict analysis to be meaningful, the root of the tree needs to be accurately identified.
+The best way to currenty do this is by hand, although a default rooting based on OpenTree's taxonomy is also available.
 
@@
 
-A tutorial as well as illustrated examples of functions implemented on each step of the analysis are available in Physcraper's documentation website.
 Figure \ref{fig:results} shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
@@
 Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis comparing estimated relationships to taxonomy, performed with OpenTree tools.
+B-D show results of the conflict analysis performed with OpenTree tools.
 
@@
 
-Data repositories designed to preserve and democratize access to biological data
-are essential resources for evolutionary estimation. While data keep accumulating,
-our ability to reanalyze and incorporate new data with available knowledge is being challenged.
+Databases that preserve and democratize access to biological data
+are essential resources for our scientific endeavor.
+New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are essential.
 
-While phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data are available (e.g., Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PyPHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})), they focus on generating full trees \emph{de novo} (i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by assembling previous phylogenetic estimates).
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})) focus on generating full phylogenies \emph{de novo}, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
 While Physcraper does not generate phylogenies \emph{de novo} in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the potential of
-phylogenetic knowledge published in databases to facilitate phylogenetic placement of public molecular data.
+it successfully generates new phylogenetic knowledge, revealing the importance of
+open science in facilitating phylogenetic placement of public molecular data and accelerating the enrichment and updating of phylogenetic relationships in any region of the tree of life.
 The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
@@
 Physcraper generates individual gene trees, which fail to capture the complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used by any tool that performs coalescent analyses to generate species trees, e.g., ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
+Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used to perform coalescent analysis to generate species trees with e.g., ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), or SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
 
-Physcraper not only links molecular and tree databases, it can also link phylogenies to various biological databases, such as GBIF, the Paleobiology Database, and others, by leveraging the integrated taxonomy from OpenTree (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}).
-
-Physcraper in conjuction with OpenTree can be used to rapidly (in a matter of hours)
+Physcraper has the potential to link phylogenies to data available in any of the taxonomies integrated by the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
+The Physcraper workflow can be used to rapidly (in a matter of hours)
 address challenges overarching both fields of ecology and evolution, such as
@@
 systematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
-and generating custom species trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+and generating custom trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
 
@@
 Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC Bioinformatics}, \textbf{10}, 421.
-
-\leavevmode\hypertarget{ref-castresana2000selection}{}%
-Castresana, J. (2000). Selection of conserved blocks from multiple alignments for their use in phylogenetic analysis. \emph{Molecular Biology and Evolution}, \textbf{17}, 540--552.
 

### 2020-09-02 2-body.Rmd
fixing typos

@@
 
-Another challenhe for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
+Another challenge for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
 While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
@@
 Taxon names in the input tree must be standardized to the OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool.
-Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments avilable on TreeBASE [@piel2009treebase].
+Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments available on TreeBASE [@piel2009treebase].
 
@@
 Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
-New sequences are excuded for analysis if 1) they do not belong to the search taxon; 2) they have an e-value above the cutoff (default to 0.00001); 3) they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OpenTree taxonomy or the NCBI taxonomy.
+New sequences are excluded for analysis if 1) they do not belong to the search taxon; 2) they have an e-value above the cutoff (default to 0.00001); 3) they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80% and 120%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OpenTree taxonomy or the NCBI taxonomy.
 An arbitrary maximum number of randomly chosen sequences per taxon are allowed (default to 5).
@@
 For the conflict analysis to be meaningful, the root of the tree needs to be accurately identified.
-The best way to currenty do this is by hand, although a default rooting based on OpenTree's taxonomy is also available.
+The best way to currently do this is by hand, although a default rooting based on OpenTree's taxonomy is also available.
 <!-- It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree.
@@
 
-An online literature review in June 2020 (google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
+An online literature review in June 2020 (Google scholar search for "ilex phylogeny") reveals that there are several published phylogenetic trees showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have their data available publicly [@gottlieb2005molecular; @yao2020phylogeny].
 @gottlieb2005molecular made original tree and alignment data available in [TreeBASE](https://treebase.org/treebase-web/search/study/summary.html?id=1091).
@@
 pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to benchmark a comparison.
+Unfortunately, installation of the tool was unsuccessful following instructions from the author, and we were unable to benchmark a comparison.
 <!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->

### 2020-11-13 render2-mee.Rmd
word limit includes references :sob:
last mee render rmd

@@
   author3: Emily Jane McTavish
-  email3: ejmctavish@gmail.com
+  email3: ejmctavish@ucmerced.edu
   affiliation: School of Natural Sciences, University of California, Merced
@@
 word_count_body <- wordcountaddin::word_count(filename="2-body.Rmd")
+word_count_references <- 873
 ```
@@
 
-**Word count**: `r word_count_body`
+**Word count**: `r word_count_body + word_count_references`
 
@@
 ```
-
-\newpage
-\begin{center}
-\textsc{References}
-\end{center}

### 2020-11-13 docs/submission-mee-2020-09-02.tex
last mee version

@@
 
-email: \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+email: \href{mailto:ejmctavish@ucmerced.edu}{\nolinkurl{ejmctavish@ucmerced.edu}}
 
@@
 
-\textbf{Correspondence email}: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:ejmctavish@gmail.com}{\nolinkurl{ejmctavish@gmail.com}}
+\textbf{Correspondence email}: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:ejmctavish@ucmerced.edu}{\nolinkurl{ejmctavish@ucmerced.edu}}
 
@@
 
-\textbf{Word count}: 2419
+\textbf{Word count}: 2969
 
@@
   However, interpretation of results, comparison with previously available phylogenetic
-  hypotheses, and choice of one phylogeny for downstream analyses and discussion still impose difficulties
+  hypotheses, and selection of one phylogeny for downstream analyses and discussion still impose difficulties
   to one that is not a specialist either on phylogenetic methods or on a particular group of study.
@@
 \item
-  The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+  The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging researchers to strive for better sharing practices.
 Physcraper can be used with any OS and is released under an open-source license.
@@
 Phylogenies capture the shared history of organisms and provide key evolutionary context for our biological observations.
-Public biological databases constitute an amazing resource for evolutionary studies, but a large portion of public molecular data has never been incorporated into any phylogenetic estimate.
-Extending existing phylogenies with new DNA sequence data, geographical location, and other metadata in a reproducible and continuous manner is possible by automating connections between biological databases.
-Here, we introduce Physcraper, a tool to build upon DNA datasets that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope, by using data from public DNA databases to update a starting tree and single locus alignments.
+Public biological databases constitute an amazing resource for evolutionary studies.
+Updating existing phylogenies with molecular data that has never been incorporated into any phylogenetic estimate, geographical location, fossils, and other data in a reproducible and continuous manner is possible by establishing a data interoperability framework for biological databases.
+Here, we introduce Physcraper, a tool that automates database connections to build upon homology hypotheses that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope to update a starting tree and single locus alignments with public DNA data.
 
-Taxonomic idiosyncrasies across databases represent a huge challenge for automatic integration of data into phylogenies; challenge that can be addressed with a unified taxonomy for name standardization.
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic integration of data into phylogenies, which can be addressed with a unified taxonomy for name standardization.
 The Open Tree of Life project (OpenTree)
-constructs a comprehensive tree of life by synthesizing published phylogenies along with taxonomic data.
+constructs a comprehensive tree of life by synthesizing published phylogenies and taxonomy.
 OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomy from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), among others.
-This OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
+To achieve this, OpenTree unifies taxonomic data from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), among others.
+The OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
 
-Another challenhe for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
-While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
-Phylogenomics ameliorate this problem by focusing on targeted capture of informative characters from independent and single-copy genetic markers (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}).
+Another challenge for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches it uses produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative loci (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}).
 Yet, decades of single locus sequencing have generated massive amounts of homologous DNA datasets that can be used for phylogenetic reconstruction at many scales.
-Moreover, species tree reconstructions from multiple single locus datasets using the multispecies coalescent model are considered the gold standard for inferring species relationships (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
 
-More than a decade ago, GenBank release number 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of 98.05\% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of 98.05\% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
 Assembling a DNA alignment from such a massive database can be done ``by hand'', but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that make DNA sequence search faster and more efficient, as well as more reproducible, have been applied to infer phylogenetic relationships in a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, thoughtfully curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic reconstructions (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+Computational pipelines that mine DNA databases fast, efficiently, and reproducibly, have been applied to infer phylogenetic relationships in a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
+However, fine-grained curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic analyses (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
 
-A way to incorporate the benefits from massive amounts of newly released molecular data and fine-grained curation from human experts, is to rely on manually curated homology hypotheses as ``jump-start'' alignments (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}).
-The TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}) hosts about 8,200 alignments from published peer-reviewed studies, providing information on evolutionary relationships of around 100,000 distinct taxa \href{https://www.treebase.org/treebase-web/home.html\#:~:text=TreeBASE\%20is\%20produced\%20and\%20governed,mapped\%20to\%20104\%2C593\%20distinct\%20taxa.}{(see TreeBASE's website about)}.
-Linking published alignments generated by experts with molecular data that has not yet been included in any public phylogenetic estimate has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life.
+There are almost 8,200 publicly available, peer-reviewed alignments, covering around 100,000 distinct taxa in the TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}), which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments for phylogenetic reconstructions (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}) to continually enrich, update and compare existing phylogenetic knowledge.
 
-Physcraper relies on programmatic access protocols (API's) and the OpenTree taxonomy to automatically link phylogenetic data to molecular data from GenBank, alignments from TreeBASE, and phylogenies from OpenTree's Phylesystem, to continually update and enrich phylogenetic knowledge based on expertly-curated homology hypotheses.
-Physcraper also provides new types of access to various OpenTree tools for comparison of existing phylogenetic hypotheses with newly generated ones.
-Physcraper is coded as a Python pipeline that focuses on data interoperability, by using the OpenTree taxonomy as a way to link taxon data from different databases.
-This integration also allows users to rapidly place new data from a diverse range of biological databases in an evolutionary context, opening the possibility for a variety of comparative downstream analyses.
+Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access protocols (API's) to implement a database interoperability framework that automatically links phylogenies that have been standardized to OpenTree taxonomy, to alignments from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
+Physcraper aims to demonstrate the benefits of reproducible workflows and open science in phylogenetics, and encourage better data sharing practices in the community.
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic-3}
+{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic-4}
 
@@
 
-The general Physcraper framework is depicted in Figure \ref{fig:framework}.
-It consists of 4 steps: 1) identifying and processing a tree and its underlying alignment; 2) performing a BLAST search of DNA sequences from the original alignment on GenBank, and filtering of new sequences; 3) profile-aligning new sequences to the original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to previous phylogenetic estimates.
+The general Physcraper framework consists of 4 steps (Fig.
+\ref{fig:framework}): 1) identifying and processing a tree and its underlying alignment; 2) performing a BLAST search of DNA sequences from original alignment on GenBank, and filtering of new sequences; 3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to existing phylogenies.
 
@@
 
-Taxon names in the input tree must be standardized to the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments avilable on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
+Taxon names in the input tree must be standardized to OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments available on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
 
-The input alignment is a single locus alignment that was used in part or in whole to generate the input tree.
-If it is stored in TreeBASE, Physcraper retrieves it automatically.
-Otherwise, the user must provide the path to a local copy of the alignment.
+The input alignment is a single locus DNA dataset that was used in part or in whole to generate the input tree.
+Physcraper retrieves TreeBASE alignments automatically.
+Alternatively, users must provide the path to a local copy of the alignment.
 Only taxa that are both in the sequence alignment and in the tree are considered further for analysis; at least one taxon and its corresponding sequence are required.
@@
 The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}) is used for DNA sequence search on a remote or local GenBank database.
-It is constrained to a ``search taxon'', a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the most recent common ancestor of the ingroup taxa in the tree that is also a named clade in the NCBI taxonomy (Fig.
+It is constrained to a ``search taxon'', a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the most recent common ancestor of ingroup taxa that is also a named clade in the NCBI taxonomy (Fig.
 \ref{fig:framework}).
-Alternatively, users can arbitrarily define a search taxon that is either a more or less inclusive clade relative to the ingroup of the input tree.
+Alternatively, users can arbitrarily define a search taxon that is either a more or less inclusive clade relative to the ingroup taxa.
 
-BLAST is implemented with the BLAST command line tools \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) and the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from the \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} modified to accept an alternative BLAST address.
+BLAST is implemented with the \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) and the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} modified to accept an alternative BLAST address.
 Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
-New sequences are excuded for analysis if 1) they do not belong to the search taxon; 2) they have an e-value above the cutoff (default to 0.00001); 3) they fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in the input alignment (default values of 80\% and 120\%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in the OpenTree taxonomy or the NCBI taxonomy.
+New sequences are excluded for analysis if they 1) are not in the search taxon; 2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in input alignment (default values of 80\% and 120\%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in OpenTree or NCBI taxonomy.
 An arbitrary maximum number of randomly chosen sequences per taxon are allowed (default to 5).
@@
 
-A Maximum Likelihood (ML) gene tree is reconstructed for each alignment provided, using RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) with default settings (GTRCAT model and 100 bootstrap replicates with the default algorithm) and using the input tree as a starting tree for the ML searches.
-Bootstrap results are summarized with DendroPy's SumTrees module (current version 4.4.0; Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) is implemented to reconstruct a Maximum Likelihood (ML) gene tree for each input alignment with default settings (GTRCAT model and 100 bootstrap replicates with default algorithm), using input tree as starting tree for ML searches.
+Bootstrap results are summarized using DendroPy's SumTrees module (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
 
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-The updated tree is compared to the original tree using an automated conflict analysis which calculates Robinson Foulds weighted and unweighted metrics using Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and performs a node by node comparison between the synthetic OpenTree and the original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
-For the conflict analysis to be meaningful, the root of the tree needs to be accurately identified.
-The best way to currenty do this is by hand, although a default rooting based on OpenTree's taxonomy is also available.
+Updated and original tree are compared with Robinson-Foulds weighted and unweighted metrics estimated with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and with a node by node comparison between the synthetic OpenTree and original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
 
@@
 
-A user is interested in the phylogenetic relationships within the genus \emph{Ilex}.
+A user is interested in phylogenetic relationships within the genus \emph{Ilex}.
 Commonly known as ``hollies'', the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
 
-An online literature review in June 2020 (google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenetic trees showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have their data available publicly (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
-Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment data available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
-The ``Gottlieb2005'' tree sampling 41 species was added to the \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and its information has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
+An online literature review in June 2020 (Google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenies showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have data publicly available (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
+Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
+The ``Gottlieb2005'' tree sampling 41 species was added to \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and it has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
 
-The most recent \emph{Ilex} tree from Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}), has been made available in the \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
-The ``Yao2020'' tree, is the best sampled phylogenetic tree yet available for the hollies, with 175 tips.
+The most recent \emph{Ilex} tree (Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}) is available in \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
+With 175 tips, the ``Yao2020'' tree is the best sampled phylogeny yet available for the hollies.
 
-Figure \ref{fig:results} shows results from the Physcraper analysis of an alignment of the internal transcribed spacer DNA region (ITS) from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}).
-Physcraper ran on a local BLAST database, on a laptop Linux computer for 19hrs 45min to perform BLAST and RAxML analyses, with bootstrap analyses taking an additional 13hrs.
-The updated Gottlieb2005 tree contains all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional \emph{Ilex} taxa.
+We ran Physcraper on a laptop Linux computer to update an internal transcribed spacer DNA region (ITS) alignment from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}), using a local GenBank database.
+BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap analyses taking an additional 13hrs.
+The updated Gottlieb2005 tree (Fig.
+\ref{fig:results}) displays all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional \emph{Ilex} taxa.
 The best RaxML tree is 99\% resolved, with 25\% of nodes with bootstrap support \textless{} 0.1 and 48\% nodes with bootstrap support \textgreater{} 0.75.
@@
 \ref{fig:results}).
-For comparison, the Yao2020 tree also contains all 41 distinct taxa from the original Gottlieb2005 study,
+For comparison, Yao2020 also contains all 41 distinct taxa from the original Gottlieb2005 study,
 and contributes phylogenetic data to 134 additional \emph{Ilex} taxa, from which
-67 are also in the Physcraper updated Gottlieb2005 tree.
-While Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, their data in GenBank is not public yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+67 are also in updated Gottlieb2005.
+While Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, their GenBank data is not released yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
 However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
-This might be caused by the method they used to download existing ITS \emph{Ilex} sequences from GenBank, which is not fully explained in the publication, but seems to be a ``manual'' process.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A) Phylogenetic tree obtained by updating the Gottlieb et al. 2005 tree in B) using Physcraper.}\label{fig:results}
+\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A) Phylogeny updated with Physcraper from Gottlieb et al. 2005 tree in B.}\label{fig:results}
 \end{figure}
@@
 \emph{Figure \ref{fig:results} caption continued}: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively.
-Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria specified in section ``DNA sequence search and filtering'', aligned them to the original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
-B-D show results of the conflict analysis performed with OpenTree tools.
+Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria from section ``DNA sequence search and filtering'', aligned them to original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps.
+B-D conflict analyses performed with OpenTree tools.
 
@@
 
-Databases that preserve and democratize access to biological data
-are essential resources for our scientific endeavor.
-New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are essential.
+Databases preserving and democratizing access to biological data
+have become essential resources for science.
+New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are needed.
 
@@
 it successfully generates new phylogenetic knowledge, revealing the importance of
-open science in facilitating phylogenetic placement of public molecular data and accelerating the enrichment and updating of phylogenetic relationships in any region of the tree of life.
+open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
 The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
 pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation of the tool was unsuccesful following instructions from the author, and we were unable to benchmark a comparison.
+Unfortunately, installation was unsuccessful following instructions from the author, and a comparison analysis was unfeasible.
 
-Physcraper generates individual gene trees, which fail to capture the complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-Yet, Physcraper makes it straightforward to gather alignments and gene trees for multiple loci from a taxonomic group of interest, which can then be used to perform coalescent analysis to generate species trees with e.g., ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), or SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
+Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
+Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), or SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
 
-Physcraper has the potential to link phylogenies to data available in any of the taxonomies integrated by the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
+Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
 The Physcraper workflow can be used to rapidly (in a matter of hours)
@@
 placing newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
-obtaining trees for ecophylogenetic studies (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),
 systematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
-and generating custom trees for ecological and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+and generating custom trees for ecological (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}) and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
 
 Data repositories hold more information than meets the eye.
-Besides the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-Usually, initial ideas about the data are changed by new analyses.
-Physcraper can provide context for these ideas by streamlining inferences integrating new data with existing knowledge.
+Beyond the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Initial ideas about the data are constantly changed by results from new analyses.
+Physcraper provides a framework for reproducible phylogenetics that has the potential to consistently provide context for these ideas, highlighting the importance of data sharing and open science in the field, biology and science.
 
@@
 
-Research was supported by the grant ``Sustaining the Open Tree of Life'', National Science Foundation ABI No.~1759838, and ABI No.~1759846.
+Research was supported by the grant ``Sustaining the Open Tree of Life'', NSF ABI No.~1759838, and ABI No.~1759846.
 Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No.~ACI-1429783.
@@
 
-LLSR wrote the manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
+LLSR wrote manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
 All authors contributed to the manuscript and gave final approval for publication.
@@
 
-Physcraper source code available at \url{https://github.com/McTavishLab/physcraper}
+Physcraper source code: \url{https://github.com/McTavishLab/physcraper}
 
-Documentation available at \url{https://physcraper.readthedocs.io/en/latest/index.html}
+Documentation: \url{https://physcraper.readthedocs.io/en/latest/index.html}
 
-Illustrated examples available at \url{https://github.com/McTavishLab/physcraperex}
+Examples: \url{https://github.com/McTavishLab/physcraperex}
 
-This is a reproducible manuscript available at \url{https://github.com/McTavishLab/physcraper_ms}
+Reproducible manuscript: \url{https://github.com/McTavishLab/physcraper_ms}
 

### 2020-11-16 sources/bmc_template_submission/bmc_article.tex
bmc template working duplicate

 %% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% https://www.biomedcentral.com/getpublished                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% https://miktex.org/                                             %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
%\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A sample article title}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
  addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
  corref={aff1},                       % id of corresponding address, if any
% noteref={n1},                        % id's of article notes, if any
  email={jane.e.doe@cambridge.co.uk}   % email address
]{\inits{J.E.}\fnm{Jane E.} \snm{Doe}}
\author[
  addressref={aff1,aff2},
  email={john.RS.Smith@cambridge.co.uk}
]{\inits{J.R.S.}\fnm{John R.S.} \snm{Smith}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgdiv{Department of Science},             % department, if any
  \orgname{University of Cambridge},          % university, etc
  \city{London},                              % city
  \cny{UK}                                    % country
}
\address[id=aff2]{%
  \orgdiv{Institute of Biology},
  \orgname{National University of Sciences},
  %\street{},
  %\postcode{}
  \city{Kiel},
  \cny{Germany}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{artnotes}
%%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                           %%
%% The Abstract begins here                  %%
%%                                           %%
%% Please refer to the Instructions for      %%
%% authors on https://www.biomedcentral.com/ %%
%% and include the section headings          %%
%% accordingly for your article type.        %%
%%                                           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{sample}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for two column layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                            %%
%% The Main Body begins here                  %%
%%                                            %%
%% Please refer to the instructions for       %%
%% authors on:                                %%
%% https://www.biomedcentral.com/getpublished %%
%% and include the section headings           %%
%% accordingly for your article type.         %%
%%                                            %%
%% See the Results and Discussion section     %%
%% for details on how to create sub-sections  %%
%%                                            %%
%% use \cite{...} to cite references          %%
%%  \cite{koon} and                           %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}      %%
%%                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Content}
Text and results for this section, as per the individual journal's instructions for authors.

\section*{Section title}
Text for this section\ldots
\subsection*{Sub-heading for section}
Text for this sub-heading\ldots
\subsubsection*{Sub-sub heading for section}
Text for this sub-sub-heading\ldots
\paragraph*{Sub-sub-sub heading for section}
Text for this sub-sub-sub-heading\ldots

In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
addition, we examine a common modeling assumption and note the
importance of considering the tails of the extinction time $T_x$ in
studies of escape dynamics.
We will first consider the expected resistant population at $vT_x$ for
some $v>0$, (and temporarily assume $\alpha=0$)
%
\[
E \bigl[Z_1(vT_x) \bigr]=
\int_0^{v\wedge
1}Z_0(uT_x)
\exp (\lambda_1)\,du .
\]
%
If we assume that sensitive cells follow a deterministic decay
$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
estimate the expected value as
%
\begin{equation}\label{eqexpmuts}
\begin{aligned}[b]
&      E\bigl[Z_1(vT_x)\bigr]\\
&\quad      = \frac{\mu}{r}\log x
\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du .
\end{aligned}
\end{equation}
%
Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,xjon,marg,schn,koha,issnic}).

\section*{Appendix}
Text for this section\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Acknowledgements}%% if any
Text for this section\ldots

\section*{Funding}%% if any
Text for this section\ldots

\section*{Abbreviations}%% if any
Text for this section\ldots

\section*{Availability of data and materials}%% if any
Text for this section\ldots

\section*{Ethics approval and consent to participate}%% if any
Text for this section\ldots

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Consent for publication}%% if any
Text for this section\ldots

\section*{Authors' contributions}
Text for this section \ldots

\section*{Authors' information}%% if any
Text for this section\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}
  \begin{figure}[h!]
  \caption{Sample figure title}
\end{figure}

\begin{figure}[h!]
  \caption{Sample figure title}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go}
  \begin{tabular}{cccc}
    \hline
    & B1  &B2   & B3\\ \hline
    A1 & 0.1 & 0.2 & 0.3\\
    A2 & ... & ..  & .\\
    A3 & ..  & .   & .\\ \hline
  \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2020-11-18 docs/try-submission-bmc-2020-11-18.tex
try render bmc format pdf from rmd

--- 2020-11-13 docs/submission-mee-2020-09-02.tex
@@
 
-\textbf{2. Martha Kandziora}
+\textbf{2. Martha Kandziora}116*0.57
 
@@
 OpenTree's ``synthetic'' tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
-To achieve this, OpenTree unifies taxonomic data from various databases (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}), among others.
+To achieve this, OpenTree unifies taxonomic data from various databases {[}\protect\hyperlink{ref-rees2017automated}{1}{]}, including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank {[}\protect\hyperlink{ref-benson2000genbank}{2}{]}, among others.
 The OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
@@
 While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches it uses produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
-Phylogenomics ameliorate this problem by focusing on targeted capture of informative loci (Andermann \emph{et al.} \protect\hyperlink{ref-andermann2020guide}{2020}).
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative loci {[}\protect\hyperlink{ref-andermann2020guide}{3}{]}.
 Yet, decades of single locus sequencing have generated massive amounts of homologous DNA datasets that can be used for phylogenetic reconstruction at many scales.
 
-More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of 98.05\% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).
+More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of 98.05\% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time {[}\protect\hyperlink{ref-sanderson2008phylota}{4}{]}.
 Assembling a DNA alignment from such a massive database can be done ``by hand'', but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that mine DNA databases fast, efficiently, and reproducibly, have been applied to infer phylogenetic relationships in a variety of organisms (e.g., Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}; Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}; Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017}).
-However, fine-grained curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic analyses (Fragoso-Martínez \emph{et al.} \protect\hyperlink{ref-fragoso2017pilot}{2017}).
+Computational pipelines that mine DNA databases fast, efficiently, and reproducibly, have been applied to infer phylogenetic relationships in a variety of organisms {[}\protect\hyperlink{ref-smith2009mega}{5}--\protect\hyperlink{ref-izquierdo2014pumper}{7}{]}.
+However, fine-grained curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic analyses {[}\protect\hyperlink{ref-fragoso2017pilot}{8}{]}.
 
-There are almost 8,200 publicly available, peer-reviewed alignments, covering around 100,000 distinct taxa in the TreeBASE database (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}), which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments for phylogenetic reconstructions (Morrison \protect\hyperlink{ref-morrison2006multiple}{2006}) to continually enrich, update and compare existing phylogenetic knowledge.
+There are almost 8,200 publicly available, peer-reviewed alignments, covering around 100,000 distinct taxa in the TreeBASE database {[}\protect\hyperlink{ref-piel2009treebase}{9}{]}, which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments for phylogenetic reconstructions {[}\protect\hyperlink{ref-morrison2006multiple}{10}{]} to continually enrich, update and compare existing phylogenetic knowledge.
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{docs/figs/schematic-4}
+{\centering \includegraphics[width=0.85\linewidth]{../docs/figs/schematic-4}
 
@@
 
-Taxon names in the input tree must be standardized to OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}) using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
-Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments available on TreeBASE (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}).
+Taxon names in the input tree must be standardized to OpenTree taxonomy {[}\protect\hyperlink{ref-rees2017automated}{1}{]} using OpenTree's bulk Taxonomic Name Resolution Service \href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool.
+Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's \href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have alignments available on TreeBASE {[}\protect\hyperlink{ref-piel2009treebase}{9}{]}.
 
@@
 
-The Basic Local Alignment Search Tool, BLAST (Altschul \emph{et al.} \protect\hyperlink{ref-altschul1990basic}{1990}) is used for DNA sequence search on a remote or local GenBank database.
-It is constrained to a ``search taxon'', a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), as the most recent common ancestor of ingroup taxa that is also a named clade in the NCBI taxonomy (Fig.
+The Basic Local Alignment Search Tool, BLAST {[}\protect\hyperlink{ref-altschul1990basic}{11}{]} is used for DNA sequence search on a remote or local GenBank database.
+It is constrained to a ``search taxon'', a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree \href{https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3\#mrca}{API} {[}\protect\hyperlink{ref-rees2017automated}{1}{]}, as the most recent common ancestor of ingroup taxa that is also a named clade in the NCBI taxonomy (Fig.
 \ref{fig:framework}).
@@
 
-BLAST is implemented with the \texttt{blastn} function (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) and the BioPython (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}) BLAST function from \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} modified to accept an alternative BLAST address.
+BLAST is implemented with the \texttt{blastn} function {[}\protect\hyperlink{ref-camacho2009blast}{12}{]} and the BioPython {[}\protect\hyperlink{ref-cock2009biopython}{13}{]} BLAST function from \href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module} modified to accept an alternative BLAST address.
 Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
@@
 
-Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions (Cock \emph{et al.} \protect\hyperlink{ref-cock2009biopython}{2009}).
+Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions {[}\protect\hyperlink{ref-cock2009biopython}{13}{]}.
 Iterative cycles of BLAST searches can be performed, by blasting all new sequences until no new ones are found. By default only one BLAST cycle is performed.
@@
 
-MUSCLE (Edgar \protect\hyperlink{ref-edgar2004muscle}{2004}) is used to perform a profile alignment in which the original alignment is used as a template of homology criteria to align new sequences.
+MUSCLE {[}\protect\hyperlink{ref-edgar2004muscle}{14}{]} is used to perform a profile alignment in which the original alignment is used as a template of homology criteria to align new sequences.
 The final alignment is not further automatically checked, and additional inspection and refinement are recommended.
@@
 
-RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014}) is implemented to reconstruct a Maximum Likelihood (ML) gene tree for each input alignment with default settings (GTRCAT model and 100 bootstrap replicates with default algorithm), using input tree as starting tree for ML searches.
-Bootstrap results are summarized using DendroPy's SumTrees module (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}).
+RAxML {[}\protect\hyperlink{ref-stamatakis2014raxml}{15}{]} is implemented to reconstruct a Maximum Likelihood (ML) gene tree for each input alignment with default settings (GTRCAT model and 100 bootstrap replicates with default algorithm), using input tree as starting tree for ML searches.
+Bootstrap results are summarized using DendroPy's SumTrees module {[}\protect\hyperlink{ref-sukumaran2010dendropy}{16}{]}.
 
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original tree are compared with Robinson-Foulds weighted and unweighted metrics estimated with Dendropy (Sukumaran \& Holder \protect\hyperlink{ref-sukumaran2010dendropy}{2010}), and with a node by node comparison between the synthetic OpenTree and original and updated tree individually, using OpenTree's conflict API (Redelings \& Holder \protect\hyperlink{ref-redelings2017supertree}{2017}).
+Updated and original tree are compared with Robinson-Foulds weighted and unweighted metrics estimated with Dendropy {[}\protect\hyperlink{ref-sukumaran2010dendropy}{16}{]}, and with a node by node comparison between the synthetic OpenTree and original and updated tree individually, using OpenTree's conflict API {[}\protect\hyperlink{ref-redelings2017supertree}{17}{]}.
 
@@
 
-An online literature review in June 2020 (Google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenies showing relationships within the hollies (Cuénoud \emph{et al.} \protect\hyperlink{ref-cuenoud2000molecular}{2000}; Setoguchi \& Watanabe \protect\hyperlink{ref-setoguchi2000intersectional}{2000}; Selbach-Schnadelbach \emph{et al.} \protect\hyperlink{ref-selbach2009new}{2009}; Manen \emph{et al.} \protect\hyperlink{ref-manen2010history}{2010}), but only two have data publicly available (Gottlieb \emph{et al.} \protect\hyperlink{ref-gottlieb2005molecular}{2005}; Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}).
-Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}) made original tree and alignment available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
+An online literature review in June 2020 (Google scholar search for ``ilex phylogeny'') reveals that there are several published phylogenies showing relationships within the hollies {[}\protect\hyperlink{ref-cuenoud2000molecular}{18}--\protect\hyperlink{ref-selbach2009new}{21}{]}, but only two have data publicly available {[}\protect\hyperlink{ref-gottlieb2005molecular}{22}, \protect\hyperlink{ref-yao2020phylogeny}{23}{]}.
+{[}\protect\hyperlink{ref-gottlieb2005molecular}{22}{]} made original tree and alignment available in \href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE}.
 The ``Gottlieb2005'' tree sampling 41 species was added to \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{OpenTree Phylesystem} and it has been integrated into \href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{OpenTree's synthetic tree}.
 
-The most recent \emph{Ilex} tree (Yao \emph{et al.} \protect\hyperlink{ref-yao2020phylogeny}{2020}) is available in \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
+The most recent \emph{Ilex} tree {[}\protect\hyperlink{ref-yao2020phylogeny}{23}{]} is available in \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{OpenTree Phylesystem} and in the \href{https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x}{DRYAD repository}.
 With 175 tips, the ``Yao2020'' tree is the best sampled phylogeny yet available for the hollies.
 
-We ran Physcraper on a laptop Linux computer to update an internal transcribed spacer DNA region (ITS) alignment from Gottlieb \emph{et al.} (\protect\hyperlink{ref-gottlieb2005molecular}{2005}), using a local GenBank database.
+We ran Physcraper on a laptop Linux computer to update an internal transcribed spacer DNA region (ITS) alignment from {[}\protect\hyperlink{ref-gottlieb2005molecular}{22}{]}, using a local GenBank database.
 BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap analyses taking an additional 13hrs.
@@
 67 are also in updated Gottlieb2005.
-While Yao \emph{et al.} (\protect\hyperlink{ref-yao2020phylogeny}{2020}) also used ITS as a marker, their GenBank data is not released yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
+While {[}\protect\hyperlink{ref-yao2020phylogeny}{23}{]} also used ITS as a marker, their GenBank data is not released yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis.
 However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{docs/figs/ilex-results} \caption{A) Phylogeny updated with Physcraper from Gottlieb et al. 2005 tree in B.}\label{fig:results}
+\includegraphics[width=1\linewidth]{../docs/figs/ilex-results} \caption{A) Phylogeny updated with Physcraper from Gottlieb et al. 2005 tree in B.}\label{fig:results}
 \end{figure}
@@
 
-Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}), PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009}), SUPERSMART (Antonelli \emph{et al.} \protect\hyperlink{ref-antonelli2017toward}{2017})) focus on generating full phylogenies \emph{de novo}, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota {[}\protect\hyperlink{ref-sanderson2008phylota}{4}{]}, PHLAWD {[}\protect\hyperlink{ref-smith2009mega}{5}{]}, SUPERSMART {[}\protect\hyperlink{ref-antonelli2017toward}{6}{]}) focus on generating full phylogenies \emph{de novo}, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
 While Physcraper does not generate phylogenies \emph{de novo} in a traditional sense,
@@
 open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
-The PUMPER pipeline (Izquierdo-Carrasco \emph{et al.} \protect\hyperlink{ref-izquierdo2014pumper}{2014}) also uses the concept of updating
+The PUMPER pipeline {[}\protect\hyperlink{ref-izquierdo2014pumper}{7}{]} also uses the concept of updating
 pre-existing alignments to incorporate public molecular data into phylogenies.
@@
 
-Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history (Song \emph{et al.} \protect\hyperlink{ref-song2012resolving}{2012}).
-Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL (Mirarab \emph{et al.} \protect\hyperlink{ref-mirarab2014astral}{2014}), BEAST2 (Bouckaert \emph{et al.} \protect\hyperlink{ref-bouckaert2019beast}{2019}), or SVD Quartets (Chifman \& Kubatko \protect\hyperlink{ref-chifman2014quartet}{2014})).
+Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history {[}\protect\hyperlink{ref-song2012resolving}{24}{]}.
+Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL {[}\protect\hyperlink{ref-mirarab2014astral}{25}{]}, BEAST2 {[}\protect\hyperlink{ref-bouckaert2019beast}{26}{]}, or SVD Quartets {[}\protect\hyperlink{ref-chifman2014quartet}{27}{]}).
 
-Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy (Rees \& Cranston \protect\hyperlink{ref-rees2017automated}{2017}), such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
+Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy {[}\protect\hyperlink{ref-rees2017automated}{1}{]}, such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
 The Physcraper workflow can be used to rapidly (in a matter of hours)
 address challenges overarching both fields of ecology and evolution, such as
-placing newly discovered species phylogenetically (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}),
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}),
-and generating custom trees for ecological (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}) and evolutionary downstream analyses (Stoltzfus \emph{et al.} \protect\hyperlink{ref-stoltzfus2013phylotastic}{2013}).
+placing newly discovered species phylogenetically {[}\protect\hyperlink{ref-webb2010biodiversity}{28}{]},
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations {[}\protect\hyperlink{ref-san2010molecular}{29}{]},
+and generating custom trees for ecological {[}\protect\hyperlink{ref-helmus2012phylogenetic}{30}{]} and evolutionary downstream analyses {[}\protect\hyperlink{ref-stoltzfus2013phylotastic}{31}{]}.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-altschul1990basic}{}%
-Altschul, S.F., Gish, W., Miller, W., Myers, E.W. \& Lipman, D.J. (1990). Basic local alignment search tool. \emph{Journal of Molecular Biology}, \textbf{215}, 403--410.
+\leavevmode\hypertarget{ref-rees2017automated}{}%
+1. Rees JA, Cranston K. Automated assembly of a reference taxonomy for phylogenetic data synthesis. Biodiversity Data Journal. 2017.
+
+\leavevmode\hypertarget{ref-benson2000genbank}{}%
+2. Benson DA, Karsch-Mizrachi I, Lipman DJ, Ostell J, Rapp BA, Wheeler DL. GenBank. Nucleic Acids Research. 2000;28:15--8.
 
 \leavevmode\hypertarget{ref-andermann2020guide}{}%
-Andermann, T., Torres Jiménez, M.F., Matos-Maraví, P., Batista, R., Blanco-Pastor, J.L., Gustafsson, A.L.S., Kistler, L., Liberal, I.M., Oxelman, B., Bacon, C.D.
-\& Antonelli, A.
-(2020).
+3.
+Andermann T, Torres Jiménez MF, Matos-Maraví P, Batista R, Blanco-Pastor JL, Gustafsson ALS, et al.
 A guide to carrying out a phylogenomic target sequence capture project.
-\emph{Frontiers in Genetics}, \textbf{10}, 1--20.
+Frontiers in Genetics.
+2020;10:1--20.
+
+\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
+4. Sanderson MJ, Boss D, Chen D, Cranston KA, Wehe A. The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. Systematic Biology. 2008;57:335--46.
+
+\leavevmode\hypertarget{ref-smith2009mega}{}%
+5. Smith SA, Beaulieu JM, Donoghue MJ. Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches. BMC Evolutionary Biology. 2009;9:37.
 
 \leavevmode\hypertarget{ref-antonelli2017toward}{}%
-Antonelli, A., Hettling, H., Condamine, F.L., Vos, K., Nilsson, R.H., Sanderson, M.J., Sauquet, H., Scharn, R., Silvestro, D., Töpel, M.
-\& others.
-(2017).
+6.
+Antonelli A, Hettling H, Condamine FL, Vos K, Nilsson RH, Sanderson MJ, et al.
 Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of taxa.
-\emph{Systematic Biology}, \textbf{66}, 152--166.
+Systematic Biology.
+2017;66:152--66.
 
-\leavevmode\hypertarget{ref-benson2000genbank}{}%
-Benson, D.A., Karsch-Mizrachi, I., Lipman, D.J., Ostell, J., Rapp, B.A. \& Wheeler, D.L. (2000). GenBank. \emph{Nucleic Acids Research}, \textbf{28}, 15--18.
+\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
+7. Izquierdo-Carrasco F, Cazes J, Smith SA, Stamatakis A. PUmPER: Phylogenies updated perpetually. Bioinformatics. 2014;30:1476--7.
+
+\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
+8.
+Fragoso-Martínez I, Salazar GA, Martínez-Gordillo M, Magallón S, Sánchez-Reyes L, Lemmon EM, et al.
+A pilot study applying the plant Anchored Hybrid Enrichment method to New World sages (\emph{Salvia} subgenus Calosphace; Lamiaceae).
+Molecular Phylogenetics and Evolution.
+2017;117:124--34.
+
+\leavevmode\hypertarget{ref-piel2009treebase}{}%
+9. Piel W, Chan L, Dominus M, Ruan J, Vos R, Tannen V. Treebase v. 2: A database of phylogenetic knowledge. E-biosphere. 2009.
+
+\leavevmode\hypertarget{ref-morrison2006multiple}{}%
+10. Morrison DA. Multiple sequence alignment for phylogenetic purposes. Australian Systematic Botany. 2006;19:479--539.
+
+\leavevmode\hypertarget{ref-altschul1990basic}{}%
+11. Altschul SF, Gish W, Miller W, Myers EW, Lipman DJ. Basic local alignment search tool. Journal of Molecular Biology. 1990;215:403--10.
+
+\leavevmode\hypertarget{ref-camacho2009blast}{}%
+12. Camacho C, George C, Vahram A, Ning M, Jason P, Kevin B, et al. BLAST+: Architecture and applications. BMC Bioinformatics. 2009;10:421.
+
+\leavevmode\hypertarget{ref-cock2009biopython}{}%
+13. Cock PJ, Antao T, Chang JT, Chapman BA, Cox CJ, Dalke A, et al. Biopython: freely available Python tools for computational molecular biology and bioinformatics. Bioinformatics. 2009;25:1422--3.
+
+\leavevmode\hypertarget{ref-edgar2004muscle}{}%
+14. Edgar RC. MUSCLE: Multiple sequence alignment with high accuracy and high throughput. Nucleic Acids Research. 2004;32:1792--7.
+
+\leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
+15. Stamatakis A. RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. Bioinformatics. 2014;30:1312--3.
+
+\leavevmode\hypertarget{ref-sukumaran2010dendropy}{}%
+16. Sukumaran J, Holder MT. DendroPy: a Python library for phylogenetic computing. Bioinformatics. 2010;26:1569--71.
+
+\leavevmode\hypertarget{ref-redelings2017supertree}{}%
+17. Redelings BD, Holder MT. A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species. PeerJ. 2017;5:e3058.
+
+\leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
+18. Cuénoud P, Martinez MA del P, Loizeay P-A, Spichiger R, Andrews S, Manen J-F. Molecular phylogeny and biogeography of the genus \emph{Ilex} L.(Aquifoliaceae). Annals of Botany. 2000;85:111--22.
+
+\leavevmode\hypertarget{ref-manen2010history}{}%
+19.
+Manen J-F, Barriera G, Loizeau P-A, Naciri Y.
+The history of extant \emph{Ilex} species (Aquifoliaceae): evidence of hybridization within a Miocene radiation.
+Molecular Phylogenetics and Evolution.
+2010;57:961--77.
+
+\leavevmode\hypertarget{ref-setoguchi2000intersectional}{}%
+20. Setoguchi H, Watanabe I. Intersectional gene flow between insular endemics of \emph{Ilex} (Aquifoliaceae) on the Bonin Islands and the Ryukyu Islands. American Journal of Botany. 2000;87:793--810.
+
+\leavevmode\hypertarget{ref-selbach2009new}{}%
+21.
+Selbach-Schnadelbach A, Cavalli SS, Manen J-F, Coelho GC, De Souza-Chies TT.
+New information for \emph{Ilex} phylogenetics based on the plastid psbA-trnH intergenic spacer (Aquifoliaceae).
+Botanical Journal of the Linnean Society.
+2009;159:182--93.
+
+\leavevmode\hypertarget{ref-gottlieb2005molecular}{}%
+22.
+Gottlieb AM, Giberti GC, Poggio L.
+Molecular analyses of the genus \emph{Ilex} (Aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
+American Journal of Botany.
+2005;92:352--69.
+
+\leavevmode\hypertarget{ref-yao2020phylogeny}{}%
+23. Yao X, Song Y, Yang J-B, Tan Y-H, Corlett RT. Phylogeny and biogeography of the hollies (\emph{Ilex} L., Aquifoliaceae). Journal of Systematics and Evolution. 2020;58:1--10.
+
+\leavevmode\hypertarget{ref-song2012resolving}{}%
+24.
+Song S, Liu L, Edwards SV, Wu S.
+Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
+Proceedings of the National Academy of Sciences.
+2012;109:14942--7.
+
+\leavevmode\hypertarget{ref-mirarab2014astral}{}%
+25. Mirarab S, Reaz R, Bayzid MS, Zimmermann T, Swenson MS, Warnow T. ASTRAL: Genome-scale coalescent-based species tree estimation. Bioinformatics. 2014;30:i541--8.
 
 \leavevmode\hypertarget{ref-bouckaert2019beast}{}%
-Bouckaert, R., Vaughan, T.G., Barido-Sottani, J., Duchêne, S., Fourment, M., Gavryushkina, A., Heled, J., Jones, G., Kühnert, D., Maio, N.D., Matschiner, M., Mendes, F.K., Müller, N.F., Ogilvie, H.A., Plessis, L.
-du, Popinga, A., Rambaut, A., Rasmussen, D., Siveroni, I., Suchard, M.A., Wu, C.-H., Xie, D., Zhang, C., Stadler, T.
-\& Drummond, A.J.
-(2019).
+26.
+Bouckaert R, Vaughan TG, Barido-Sottani J, Duchêne S, Fourment M, Gavryushkina A, et al.
 BEAST 2.5: An advanced software platform for Bayesian evolutionary analysis.
-\emph{PLOS Computational Biology}, \textbf{15}, e1006650.
-
-\leavevmode\hypertarget{ref-camacho2009blast}{}%
-Camacho, C., George, C., Vahram, A., Ning, M., Jason, P., Kevin, B. \& Thomas, L. (2009). BLAST+: Architecture and applications. \emph{BMC Bioinformatics}, \textbf{10}, 421.
+PLOS Computational Biology.
+2019;15:e1006650.
 
 \leavevmode\hypertarget{ref-chifman2014quartet}{}%
-Chifman, J. \& Kubatko, L. (2014). Quartet inference from SNP data under the coalescent model. \emph{Bioinformatics}, \textbf{30}, 3317--3324.
+27. Chifman J, Kubatko L. Quartet inference from SNP data under the coalescent model. Bioinformatics. 2014;30:3317--24.
 
-\leavevmode\hypertarget{ref-cock2009biopython}{}%
-Cock, P.J., Antao, T., Chang, J.T., Chapman, B.A., Cox, C.J., Dalke, A., Friedberg, I., Hamelryck, T., Kauff, F., Wilczynski, B.
-\& others.
-(2009).
-Biopython: freely available Python tools for computational molecular biology and bioinformatics.
-\emph{Bioinformatics}, \textbf{25}, 1422--1423.
+\leavevmode\hypertarget{ref-webb2010biodiversity}{}%
+28. Webb CO, Slik JF, Triono T. Biodiversity inventory and informatics in Southeast Asia. Biodiversity and Conservation. 2010;19:955--72.
 
-\leavevmode\hypertarget{ref-cuenoud2000molecular}{}%
-Cuénoud, P., Martinez, M.A.
-del P., Loizeay, P.-A., Spichiger, R., Andrews, S.
-\& Manen, J.-F.
-(2000).
-Molecular phylogeny and biogeography of the genus \emph{Ilex} L.(Aquifoliaceae).
-\emph{Annals of Botany}, \textbf{85}, 111--122.
-
-\leavevmode\hypertarget{ref-edgar2004muscle}{}%
-Edgar, R.C. (2004). MUSCLE: Multiple sequence alignment with high accuracy and high throughput. \emph{Nucleic Acids Research}, \textbf{32}, 1792--1797.
-
-\leavevmode\hypertarget{ref-fragoso2017pilot}{}%
-Fragoso-Martínez, I., Salazar, G.A., Martínez-Gordillo, M., Magallón, S., Sánchez-Reyes, L., Lemmon, E.M., Lemmon, A.R., Sazatornil, F.
-\& Mendoza, C.G.
-(2017).
-A pilot study applying the plant Anchored Hybrid Enrichment method to New World sages (\emph{Salvia} subgenus Calosphace; Lamiaceae).
-\emph{Molecular Phylogenetics and Evolution}, \textbf{117}, 124--134.
-
-\leavevmode\hypertarget{ref-gottlieb2005molecular}{}%
-Gottlieb, A.M., Giberti, G.C.
-\& Poggio, L.
-(2005).
-Molecular analyses of the genus \emph{Ilex} (Aquifoliaceae) in southern south america, evidence from aflp and its sequence data.
-\emph{American Journal of Botany}, \textbf{92}, 352--369.
+\leavevmode\hypertarget{ref-san2010molecular}{}%
+29. San Mauro D, Agorreta A. Molecular systematics: A synthesis of the common methods and the state of knowledge. Cellular \& Molecular Biology Letters. 2010;15:311.
 
 \leavevmode\hypertarget{ref-helmus2012phylogenetic}{}%
-Helmus, M.R. \& Ives, A.R. (2012). Phylogenetic diversity--area curves. \emph{Ecology}, \textbf{93}, S31--S43.
-
-\leavevmode\hypertarget{ref-izquierdo2014pumper}{}%
-Izquierdo-Carrasco, F., Cazes, J., Smith, S.A. \& Stamatakis, A. (2014). PUmPER: Phylogenies updated perpetually. \emph{Bioinformatics}, \textbf{30}, 1476--1477.
-
-\leavevmode\hypertarget{ref-manen2010history}{}%
-Manen, J.-F., Barriera, G., Loizeau, P.-A.
-\& Naciri, Y.
-(2010).
-The history of extant \emph{Ilex} species (Aquifoliaceae): evidence of hybridization within a Miocene radiation.
-\emph{Molecular Phylogenetics and Evolution}, \textbf{57}, 961--977.
-
-\leavevmode\hypertarget{ref-mirarab2014astral}{}%
-Mirarab, S., Reaz, R., Bayzid, M.S., Zimmermann, T., Swenson, M.S. \& Warnow, T. (2014). ASTRAL: Genome-scale coalescent-based species tree estimation. \emph{Bioinformatics}, \textbf{30}, i541--i548.
-
-\leavevmode\hypertarget{ref-morrison2006multiple}{}%
-Morrison, D.A. (2006). Multiple sequence alignment for phylogenetic purposes. \emph{Australian Systematic Botany}, \textbf{19}, 479--539.
-
-\leavevmode\hypertarget{ref-piel2009treebase}{}%
-Piel, W., Chan, L., Dominus, M., Ruan, J., Vos, R. \& Tannen, V. (2009). Treebase v. 2: A database of phylogenetic knowledge. E-biosphere.
-
-\leavevmode\hypertarget{ref-redelings2017supertree}{}%
-Redelings, B.D. \& Holder, M.T. (2017). A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species. \emph{PeerJ}, \textbf{5}, e3058.
-
-\leavevmode\hypertarget{ref-rees2017automated}{}%
-Rees, J.A. \& Cranston, K. (2017). Automated assembly of a reference taxonomy for phylogenetic data synthesis. \emph{Biodiversity Data Journal}.
-
-\leavevmode\hypertarget{ref-sanderson2008phylota}{}%
-Sanderson, M.J., Boss, D., Chen, D., Cranston, K.A. \& Wehe, A. (2008). The PhyLoTA Browser: Processing GenBank for Molecular Phylogenetics Research. \emph{Systematic Biology}, \textbf{57}, 335--346.
-
-\leavevmode\hypertarget{ref-san2010molecular}{}%
-San Mauro, D. \& Agorreta, A. (2010). Molecular systematics: A synthesis of the common methods and the state of knowledge. \emph{Cellular \& Molecular Biology Letters}, \textbf{15}, 311.
-
-\leavevmode\hypertarget{ref-selbach2009new}{}%
-Selbach-Schnadelbach, A., Cavalli, S.S., Manen, J.-F., Coelho, G.C.
-\& De Souza-Chies, T.T.
-(2009).
-New information for \emph{Ilex} phylogenetics based on the plastid psbA-trnH intergenic spacer (Aquifoliaceae).
-\emph{Botanical Journal of the Linnean Society}, \textbf{159}, 182--193.
-
-\leavevmode\hypertarget{ref-setoguchi2000intersectional}{}%
-Setoguchi, H.
-\& Watanabe, I.
-(2000).
-Intersectional gene flow between insular endemics of \emph{Ilex} (Aquifoliaceae) on the Bonin Islands and the Ryukyu Islands.
-\emph{American Journal of Botany}, \textbf{87}, 793--810.
-
-\leavevmode\hypertarget{ref-smith2009mega}{}%
-Smith, S.A., Beaulieu, J.M.
-\& Donoghue, M.J.
-(2009).
-Mega-phylogeny approach for comparative biology: An alternative to supertree and supermatrix approaches.
-\emph{BMC Evolutionary Biology}, \textbf{9}, 37.
-
-\leavevmode\hypertarget{ref-song2012resolving}{}%
-Song, S., Liu, L., Edwards, S.V.
-\& Wu, S.
-(2012).
-Resolving conflict in eutherian mammal phylogeny using phylogenomics and the multispecies coalescent model.
-\emph{Proceedings of the National Academy of Sciences}, \textbf{109}, 14942--14947.
-
-\leavevmode\hypertarget{ref-stamatakis2014raxml}{}%
-Stamatakis, A. (2014). RAxML version 8: A tool for phylogenetic analysis and post-analysis of large phylogenies. \emph{Bioinformatics}, \textbf{30}, 1312--1313.
+30. Helmus MR, Ives AR. Phylogenetic diversity--area curves. Ecology. 2012;93:S31--43.
 
 \leavevmode\hypertarget{ref-stoltzfus2013phylotastic}{}%
-Stoltzfus, A., Lapp, H., Matasci, N., Deus, H., Sidlauskas, B., Zmasek, C.M., Vaidya, G., Pontelli, E., Cranston, K., Vos, R.
-\& others.
-(2013).
-Phylotastic!
-Making tree-of-life knowledge accessible, reusable and convenient.
-\emph{BMC Bioinformatics}, \textbf{14}, 158.
-
-\leavevmode\hypertarget{ref-sukumaran2010dendropy}{}%
-Sukumaran, J. \& Holder, M.T. (2010). DendroPy: a Python library for phylogenetic computing. \emph{Bioinformatics}, \textbf{26}, 1569--1571.
-
-\leavevmode\hypertarget{ref-webb2010biodiversity}{}%
-Webb, C.O., Slik, J.F. \& Triono, T. (2010). Biodiversity inventory and informatics in Southeast Asia. \emph{Biodiversity and Conservation}, \textbf{19}, 955--972.
-
-\leavevmode\hypertarget{ref-yao2020phylogeny}{}%
-Yao, X., Song, Y., Yang, J.-B., Tan, Y.-H.
-\& Corlett, R.T.
-(2020).
-Phylogeny and biogeography of the hollies (\emph{Ilex} L., Aquifoliaceae).
-\emph{Journal of Systematics and Evolution}, \textbf{58}, 1--10.
+31. Stoltzfus A, Lapp H, Matasci N, Deus H, Sidlauskas B, Zmasek CM, et al. Phylotastic! Making tree-of-life knowledge accessible, reusable and convenient. BMC Bioinformatics. 2013;14:158.
 

### 2020-11-18 sources/2pdf-bmc.Rmd
try render bmc format pdf from rmd

--- 2020-11-13 render2-mee.Rmd
@@
 bibliography: paper.bib
-csl: submission-mee/methods-in-ecology-and-evolution.csl
-# csl: submission_sysbio/systematic-biology.csl
+csl: bmc-bioinformatics.csl
 link-citations: yes
@@
 
-**2. `r params$author2`**
+**2. `r params$author2`**116*0.57
 

### 2020-11-19 sources/2-body.Rmd
bmc format failed try

---
title: 'Physcraper: Body'
author:
  - Luna L. Sanchez Reyes
  - Martha Kandziora
  - Emily Jane McTavish
  - University of California, Merced
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: paper.bib
csl: apa.csl
output:
  redoc::redoc
---

# Introduction

Phylogenies capture the shared history of organisms and provide key evolutionary context for our biological observations.
Public biological databases constitute an amazing resource for evolutionary studies. Updating existing phylogenies with molecular data that has never been incorporated into any phylogenetic estimate, geographical location, fossils, and other data in a reproducible and continuous manner is possible by establishing a data interoperability framework for biological databases. Here, we introduce Physcraper, a tool that automates database connections to build upon homology hypotheses that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope to update a starting tree and single locus alignments with public DNA data.

Taxonomic idiosyncrasies across databases represent a huge challenge for automatic integration of data into phylogenies, which can be addressed with a unified taxonomy for name standardization. The Open Tree of Life project (OpenTree)
<!-- [project](https://opentreeoflife.github.io/)  -->
constructs a comprehensive tree of life by synthesizing published phylogenies and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic data from various databases [@rees2017automated], including the USA National Center for Biodiversity Information (NCBI) molecular database GenBank [@benson2000genbank], among others. The OpenTree taxonomy represents a key resource for connecting data from any biological database that has been integrated to it.
<!-- to phylogenetic data that has been standardized to it. -->

Another challenge for incorporating public molecular data into existing phylogenies is assembling high-quality homology hypotheses.
While genomics has, and will continue to, revolutionize phylogenetic inference, the variety of alternative genomic sequencing approaches it uses produce largely non-overlapping genomic datasets across taxa, creating challenges in wide scale phylogenetic reconstruction.
<!-- Also, as @andermann2020guide point out, "assembling full genomes is often unnecessary for phylogenomic studies if the main goal is to retrieve an appropriate number of phylogenetically informative characters from several independent and single-copy genetic markers [@jones2016targeted]." -->
Phylogenomics ameliorate this problem by focusing on targeted capture of informative loci [@andermann2020guide]. Yet, decades of single locus sequencing have generated massive amounts of homologous DNA datasets that can be used for phylogenetic reconstruction at many scales.
<!-- Moreover, species tree reconstructions from multiple single locus datasets using the multispecies coalescent model are considered the gold standard for inferring species relationships [@song2012resolving]. -->

More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72 million DNA sequences that were gauged to have the potential to resolve phylogenetic relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000 distinct taxa in the NCBI taxonomy at the time [@sanderson2008phylota]. Assembling a DNA alignment from such a massive database can be done "by hand", but it is a largely time consuming and mostly non-reproducible approach. Computational pipelines that mine DNA databases fast, efficiently, and reproducibly, have been applied to infer phylogenetic relationships in a variety of organisms [e.g., @smith2009mega; @antonelli2017toward; @izquierdo2014pumper].
<!-- *maybe a box of all pipelines?* -->
However, fine-grained curated markers and alignments can improve phylogenetic reconstructions, even in phylogenomic analyses [@fragoso2017pilot].

There are almost 8,200 publicly available, peer-reviewed alignments, covering around 100,000 distinct taxa in the TreeBASE database [@piel2009treebase], which can be used as seeds to mine molecular databases, and as "jump-start" alignments for phylogenetic reconstructions [@morrison2006multiple] to continually enrich, update and compare existing phylogenetic knowledge.

<!-- Linking expert alignments with molecular data that has not yet been included in any public phylogenetic estimate has the potential to accelerate the enrichment and updating of phylogenetic relationships in many regions of the tree of life. -->

Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access protocols (API's) to implement a database interoperability framework that automatically links phylogenies that have been standardized to OpenTree taxonomy, to alignments from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
Physcraper aims to demonstrate the benefits of reproducible workflows and open science in phylogenetics, and encourage better data sharing practices in the community.

# The Physcraper framework

```{r framework, echo=FALSE, fig.cap="The Physcraper framework consists of 4 steps (see text). The software is fully described on its documentation website at physcraper.readthedocs.io, along with installation instructions, function usage descriptions, examples and tutorials.", out.width = '85%', fig.retina= 2, fig.align = "center"}
# knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
knitr::include_graphics("../docs/figs/schematic-4.png")
```

The general Physcraper framework consists of 4 steps (Fig. \@ref(fig:framework)): 1) identifying and processing a tree and its underlying alignment; 2) performing a BLAST search of DNA sequences from original alignment on GenBank, and filtering of new sequences; 3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic analysis and comparing the updated tree to existing phylogenies.

## The inputs: a tree and an alignment

Taxon names in the input tree must be standardized to OpenTree taxonomy [@rees2017automated] using OpenTree's bulk Taxonomic Name Resolution Service [TNRS](https://tree.opentreeoflife.org/curator/tnrs/) tool. Users can upload their own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's [Phylesystem](https://github.com/opentreeoflife/phylesystem) that also have alignments available on TreeBASE [@piel2009treebase].

The input alignment is a single locus DNA dataset that was used in part or in whole to generate the input tree. Physcraper retrieves TreeBASE alignments automatically. Alternatively, users must provide the path to a local copy of the alignment.
<!-- Alignments stored in any other repository, or constituting personal data have to be stored locally. -->
Only taxa that are both in the sequence alignment and in the tree are considered further for analysis; at least one taxon and its corresponding sequence are required.

## DNA sequence search and filtering

The Basic Local Alignment Search Tool, BLAST [@altschul1990basic] is used for DNA sequence search on a remote or local GenBank database. It is constrained to a "search taxon", a taxonomic group in the NCBI taxonomy that is automatically identified using the OpenTree [API](<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>) [@rees2017automated], as the most recent common ancestor of ingroup taxa that is also a named clade in the NCBI taxonomy (Fig. \@ref(fig:framework)).
  <!-- FIGURE RECOMMENDED: Figure \@ref(fig:search)  -->
<!-- The MRCAT can be different from the phylogenetic MRCA when the latter is an unnamed clade in the synthetic tree.  -->
Alternatively, users can arbitrarily define a search taxon that is either a more or less inclusive clade relative to the ingroup taxa.

BLAST is implemented with the `blastn` function [@camacho2009blast] and the BioPython [@cock2009biopython] BLAST function from [NCBIWWW module](https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html) modified to accept an alternative BLAST address.
Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
New sequences are excluded for analysis if they 1) are not in the search taxon; 2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min and max length threshold, defined as the proportion of the average length without gaps of all sequences in input alignment (default values of 80% and 120%, respectively); 4) or if they are either identical to or shorter than an existing sequence in the input alignment and they represent the same taxon in OpenTree or NCBI taxonomy.
An arbitrary maximum number of randomly chosen sequences per taxon are allowed (default to 5).

Reverse, complement, and reverse-complement sequences are identified and translated using BioPython internal functions [@cock2009biopython].
Iterative cycles of BLAST searches can be performed, by blasting all new sequences until no new ones are found. By default only one BLAST cycle is performed.
<!-- in which only sequences in the input alignment are BLASTed. -->

## New DNA sequence alignment

MUSCLE [@edgar2004muscle] is used to perform a profile alignment in which the original alignment is used as a template of homology criteria to align new sequences.
The final alignment is not further automatically checked, and additional inspection and refinement are recommended.

## Tree reconstruction and comparison

RAxML [@stamatakis2014raxml] is implemented to reconstruct a Maximum Likelihood (ML) gene tree for each input alignment with default settings (GTRCAT model and 100 bootstrap replicates with default algorithm), using input tree as starting tree for ML searches.
Bootstrap results are summarized using DendroPy's SumTrees module [@sukumaran2010dendropy].

Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
Updated and original tree are compared with Robinson-Foulds weighted and unweighted metrics estimated with Dendropy [@sukumaran2010dendropy], and with a node by node comparison between the synthetic OpenTree and original and updated tree individually, using OpenTree's conflict API [@redelings2017supertree].
<!-- For the conflict analysis to be meaningful, the root of the tree needs to be accurately identified. The best way to currently do this is by hand, although a default rooting based on OpenTree's taxonomy is also available. -->
<!-- It uses the taxon labels for all the tips in the updated tree, pulls an inferred subtree from OpenTree's taxonomy and then applies the same rooting to the inferred updated tree. However, if the updated tree changes expectations from taxonomy, the rooting may no longer be appropriate. Automatic identification of a phylogenetic tree root is a difficult problem that has not been solved yet. The best way is to define the outgroup directly on the updated tree as part of the conflict analysis, so trees are accurately rooted. -->
<!-- *It would be a nice addition to have users give the output of the input tree as an argument at some point, or maybe we could add a super outgroup at random based on the search taxon* -->

# Case Study: The hollies

A user is interested in phylogenetic relationships within the genus *Ilex*. Commonly known as "hollies", the genus encompasses between 400-700 living species, and is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.

An online literature review in June 2020 (Google scholar search for "ilex phylogeny") reveals that there are several published phylogenies showing relationships within the hollies [@cuenoud2000molecular; @manen2010history; @setoguchi2000intersectional; @selbach2009new], but only two have data publicly available [@gottlieb2005molecular; @yao2020phylogeny].
@gottlieb2005molecular made original tree and alignment available in [TreeBASE](https://treebase.org/treebase-web/search/study/summary.html?id=1091). The "Gottlieb2005" tree sampling 41 species was added to [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home) and it has been integrated into [OpenTree's synthetic tree](https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa).

The most recent *Ilex* tree [@yao2020phylogeny] is available in [OpenTree Phylesystem](https://tree.opentreeoflife.org/curator/study/view/ot_1984) and in the [DRYAD repository](https://datadryad.org/stash/dataset/doi:10.5061/dryad.k0p2ngf4x). With 175 tips, the "Yao2020" tree is the best sampled phylogeny yet available for the hollies.

We ran Physcraper on a laptop Linux computer to update an internal transcribed spacer DNA region (ITS) alignment from @gottlieb2005molecular, using a local GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap analyses taking an additional 13hrs.
<!-- **MTH: you probably need some details about the hardware, given the fact that you are discussing running times** -->
The updated Gottlieb2005 tree (Fig. \@ref(fig:results)) displays all 41 distinct taxa from the original study plus 231 new tips, contributing phylogenetic data to 84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
A large portion of internal branches are negligibly small, with 30 branches < 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75 (Fig. \@ref(fig:results)).
For comparison, Yao2020 also contains all 41 distinct taxa from the original Gottlieb2005 study,
<!-- has less new tips (134), but  -->
and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
`r 135-68` are also in updated Gottlieb2005. While @yao2020phylogeny also used ITS as a marker, their GenBank data is not released yet, so Physcraper was unable to incorporate 68 additional taxa into the analysis. However, Physcraper was able to incorporate 18 taxa that were not in Yao2020.
<!-- **Any statements about relationships or anything? Are there multiple seqs for these taxa?**
add 1 sentence saying that it is odd that we are getting differnet relationships given what is likely to be the same data
but we can't compare the exact seqs or alignemnt, bc they are not available.
 -->
<!-- This might be caused by the method they used to download existing ITS *Ilex* sequences from GenBank, which is not fully explained in the publication, but seems to be a "manual" process. -->
<!-- More Physcraper examples at a range of taxonomic scales are available in the Physcraper documentation online. -->
<!-- The alignments have THIS MANY columns and THIS AMOUNT of missing data. -->
<!-- RF DISTANCE INTERPRETATION -->
<!-- ADD ML ESTIMATES OF UPDATED TREE VS ORIGINAL TREE ?-->

<!-- ```{bash echo= FALSE, include=FALSE, message= FALSE, results='hide'}
wget -O docs/figs/ilex-plot-tips-and-branches-1.png https://raw.githubusercontent.com/McTavishLab/physcraperex/master/docs/articles/ilex_files/figure-html/plot-tips-and-branches-1.png
``` -->

```{r results, echo=FALSE, fig.cap="A) Phylogeny updated with Physcraper from Gottlieb et al. 2005 tree in B.", out.width = '100%', fig.retina= 2}

# knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
knitr::include_graphics("../docs/figs/ilex-results.png")

```

\newpage

*Figure \@ref(fig:results) caption continued*: Tips in original alignment and new tips added with Physcraper are depicted in black and red, respectively. Physcraper obtained sequences from the GenBank database via local BLAST of all sequences in the original alignment that generated tree in B), filtered them following criteria from section "DNA sequence search and filtering", aligned them to original alignment using MUSCLE and performed a phylogenetic reconstruction using RAxML with 100 bootstraps. B-D conflict analyses performed with OpenTree tools.

# Discussion

Databases preserving and democratizing access to biological data
have become essential resources for science.
New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are needed.

Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota [@sanderson2008phylota], PHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]) focus on generating full phylogenies *de novo*, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
While Physcraper does not generate phylogenies *de novo* in a traditional sense,
it successfully generates new phylogenetic knowledge, revealing the importance of
open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
pre-existing alignments to incorporate public molecular data into phylogenies. Unfortunately, installation was unsuccessful following instructions from the author, and a comparison analysis was unfeasible.
<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->

Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history [@song2012resolving]. Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], or SVD Quartets [@chifman2014quartet]).

Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy [@rees2017automated], such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
<!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia]. -->
The Physcraper workflow can be used to rapidly (in a matter of hours)
address challenges overarching both fields of ecology and evolution, such as
placing newly discovered species phylogenetically [@webb2010biodiversity],
systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
and generating custom trees for ecological [@helmus2012phylogenetic] and evolutionary downstream analyses [@stoltzfus2013phylotastic].

Data repositories hold more information than meets the eye.
Beyond the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
Initial ideas about the data are constantly changed by results from new analyses.
Physcraper provides a framework for reproducible phylogenetics that has the potential to consistently provide context for these ideas, highlighting the importance of data sharing and open science in the field, biology and science.

# Acknowledgements

Research was supported by the grant "Sustaining the Open Tree of Life", NSF ABI No. 1759838, and ABI No. 1759846.
Computer time was provided by the Multi-Environment Research Computer for Exploration and Discovery (MERCED) cluster from the University of California, Merced (UCM), supported by the NSF Grant No. ACI-1429783.

We thank the members of the OpenTree development team and the "short bar" Science and Engineering Building 1, UCM, joint lab paper discussion group for valuable comments on this manuscript.

The authors have no conflict of interest to declare.

# Authors' Contributions

LLSR wrote manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
All authors contributed to the manuscript and gave final approval for publication.

# Data Archiving

Physcraper source code: https://github.com/McTavishLab/physcraper

Documentation: https://physcraper.readthedocs.io/en/latest/index.html

Examples: https://github.com/McTavishLab/physcraperex

Reproducible manuscript: https://github.com/McTavishLab/physcraper_ms

# References

### 2020-11-24 sources/bmc_template_submission/bmc_article.tex
bmc discussion

@@
 
+\usepackage{hyperref}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{A sample article title}
+\title{Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life}
 
@@
 % noteref={n1},                        % id's of article notes, if any
-  email={jane.e.doe@cambridge.co.uk}   % email address
-]{\inits{J.E.}\fnm{Jane E.} \snm{Doe}}
+  email={sanchez.reyes.luna@gmail.com}   % email address
+]{\inits{L.L.}\fnm{Luna L.} \snm{Sanchez Reyes}}
 \author[
   addressref={aff1,aff2},
-  email={john.RS.Smith@cambridge.co.uk}
-]{\inits{J.R.S.}\fnm{John R.S.} \snm{Smith}}
-
+  email={kandziom@natur.cuni.cz}
+]{\inits{M.}\fnm{Martha} \snm{Kandziora}}
+\author[
+  addressref={aff1},
+  corref={aff1},
+  email={ejmctavish@ucmerced.edu}
+]{\inits{E.J.}\fnm{Emily Jane} \snm{McTavish}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \address[id=aff1]{%                           % unique id
-  \orgdiv{Department of Science},             % department, if any
-  \orgname{University of Cambridge},          % university, etc
-  \city{London},                              % city
-  \cny{UK}                                    % country
+  \orgdiv{School of Natural Sciences},             % department, if any
+  \orgname{University of California},          % university, etc
+  \city{Merced},                              % city
+  \cny{USA}                                    % country
 }
 \address[id=aff2]{%
-  \orgdiv{Institute of Biology},
-  \orgname{National University of Sciences},
+  \orgdiv{Department of Botany, Faculty of Science},
+  \orgname{Charles University},
   %\street{},
   %\postcode{}
-  \city{Kiel},
-  \cny{Germany}
+  \city{Prague },
+  \cny{Czech Republic}
 }
@@
 %%                                           %%
+%% I followed sections from here:            %%
+%% https://bmcbioinformatics.biomedcentral.com/submission-guidelines/preparing-your-manuscript/software-article
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
+\begin{abstract}
+\parttitle{Background}
+Phylogenies are a key part of research in many areas of biology. Tools that automate
+some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
+have been developed for the advantage of both specialists in the field of phylogenetics and non-specialists.
+However, interpretation of results, comparison with previously available phylogenetic
+hypotheses, and selection of one phylogeny for downstream analyses and discussion still impose difficulties
+to one that is not a specialist either on phylogenetic methods or on a particular group of study.
 
-\parttitle{Second part title} %if any
-Text for this section.
+\parttitle{Results}
+Physcraper is a command-line Python program that automates the update of published
+phylogenies by adding public DNA sequences to underlying alignments of previously
+published phylogenies. It also provides a framework for straightforward comparison
+of published phylogenies with their updated versions, by leveraging upon tools from
+the Open Tree of Life project to link taxonomic information across databases.
+The program can be used by the nonspecialist, as a tool to generate phylogenetic
+hypotheses based on publicly available expert phylogenetic knowledge.
+Phylogeneticists and taxonomic group specialists will find it useful as a tool
+to facilitate molecular dataset gathering and comparison
+of alternative phylogenetic hypotheses (topologies).
+
+\parttitle{Conclusions}
+The Physcraper workflow showcases the benefits of doing open science for
+phylogenetics, encouraging researchers to strive for better sharing practices.
+Physcraper can be used with any OS and is released under an open-source license.
+Detailed instructions for installation and
+usage are available at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io}.
+
 \end{abstract}
@@
 \begin{keyword}
-\kwd{sample}
-\kwd{article}
-\kwd{author}
+\kwd{gene tree}
+\kwd{interoperability}
+\kwd{open science}
+\kwd{reproducibility}
+\kwd{public database}
+\kwd{DNA alignment}
 \end{keyword}
@@
 %%
-\section*{Content}
-Text and results for this section, as per the individual journal's instructions for authors.
+\section*{Background}
+Phylogenies capture the shared history of organisms and provide key evolutionary
+context for our biological observations.
+Public biological databases constitute an amazing resource for evolutionary
+studies. Updating existing phylogenies with molecular data that has never been
+incorporated into any phylogenetic estimate, geographical location, fossils, and
+other data in a reproducible and continuous manner is possible by establishing a
+data interoperability framework for biological databases. Here, we introduce
+Physcraper, a tool that automates database connections to build upon homology
+hypotheses that taxon specialists have assessed and deemed appropriate for a
+specific phylogenetic scope to update a starting tree and single locus alignments
+with public DNA data.
 
-\section*{Section title}
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
+integration of data into phylogenies, which can be addressed with a unified
+taxonomy for name standardization. The Open Tree of Life project (OpenTree)
+constructs a comprehensive tree of life by synthesizing published phylogenies
+and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
+around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are
+placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic
+data from various databases \cite{rees2017automated}, including the USA National
+Center for Biodiversity Information (NCBI) molecular database GenBank
+\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
+for connecting data from any biological database that has been integrated to it.
+
+Another challenge for incorporating public molecular data into existing phylogenies
+is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference,
+the variety of alternative genomic sequencing approaches it uses produce largely
+non-overlapping genomic datasets across taxa, creating challenges in wide scale
+phylogenetic reconstruction.
+
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative
+loci \cite{andermann2020guide}. Yet, decades of single locus sequencing have generated
+massive amounts of homologous DNA datasets that can be used for phylogenetic
+reconstruction at many scales.
+
+More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
+million DNA sequences that were gauged to have the potential to resolve phylogenetic
+relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000
+distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
+Assembling a DNA alignment from such a massive database can be done "by hand",
+but it is a largely time consuming and mostly non-reproducible approach.
+Computational pipelines that mine DNA databases fast, efficiently, and reproducibly,
+have been applied to infer phylogenetic relationships in a variety of organisms
+(e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
+However, fine-grained curated markers and alignments can improve phylogenetic
+reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
+
+There are almost 8,200 publicly available, peer-reviewed alignments, covering
+around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
+can be used as seeds to mine molecular databases, and as "jump-start" alignments
+for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
+update and compare existing phylogenetic knowledge.
+
+Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access
+protocols (API's) to implement a database interoperability framework that automatically
+links phylogenies that have been standardized to OpenTree taxonomy, to alignments
+from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
+Physcraper aims to demonstrate the benefits of reproducible workflows and open
+science in phylogenetics, and encourage better data sharing practices in the community.
+
+\section*{Implementation}
+
+The general Physcraper framework (Figure 1) consists of 4 steps: 1)identifying and
+processing a tree and its underlying alignment; 2) performing a BLAST search of
+DNA sequences from original alignment on GenBank, and filtering of new sequences;
+3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic
+analysis and comparing the updated tree to existing phylogenies.
+
+\subsection*{The inputs: a tree and an alignment}
+
+Taxon names in the input tree must be standardized to OpenTree taxonomy
+\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
+\href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool. Users can upload their
+own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
+\href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have
+alignments available on TreeBASE \cite{piel2009treebase}.
+
+The input alignment is a single locus DNA dataset that was used in part or in
+whole to generate the input tree. Physcraper retrieves TreeBASE alignments
+automatically. Alternatively, users must provide the path to a local copy of the
+alignment.
+Only taxa that are both in the sequence alignment and in the tree are considered
+further for analysis; at least one taxon and its corresponding sequence are required.
+
+\subsection*{DNA sequence search and filtering}
+
+The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
+sequence search on a remote or local GenBank database. It is constrained to a
+"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
+identified using the OpenTree
+\href{<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>}{API}
+\cite{rees2017automated}, as the most recent common ancestor of ingroup taxa that is
+also a named clade in the NCBI taxonomy (Figure 1).
+
+BLAST is implemented with the `blastn` function \cite{camacho2009blast} and the BioPython
+\cite{cock2009biopython} BLAST function from
+\href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module}
+modified to accept an alternative BLAST address.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
+New sequences are excluded for analysis if they 1) are not in the search taxon;
+2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
+and max length threshold, defined as the proportion of the average length without
+gaps of all sequences in input alignment (default values of 80% and 120%,
+respectively); 4) or if they are either identical to or shorter than an existing
+sequence in the input alignment and they represent the same taxon in OpenTree or
+NCBI taxonomy.
+An arbitrary maximum number of randomly chosen sequences per taxon are allowed
+(default to 5).
+
+Reverse, complement, and reverse-complement sequences are identified and translated
+using BioPython internal functions \cite{cock2009biopython}.
+Iterative cycles of BLAST searches can be performed, by blasting all new sequences
+until no new ones are found. By default only one BLAST cycle is performed.
+
+\subsection*{New DNA sequence alignment}
+
+MUSCLE \cite{edgar2004muscle} is used to perform a profile alignment in which the
+original alignment is used as a template of homology criteria to align new sequences.
+The final alignment is not further automatically checked, and additional inspection
+and refinement are recommended.
+
+\subsection*{Tree reconstruction and comparison}
+
+RAxML \cite{stamatakis2014raxml} is implemented to reconstruct a Maximum Likelihood
+(ML) gene tree for each input alignment with default settings (GTRCAT model and
+100 bootstrap replicates with default algorithm), using input tree as starting
+tree for ML searches.
+Bootstrap results are summarized using DendroPy's SumTrees module
+\cite{sukumaran2010dendropy}.
+
+Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
+Updated and original tree are compared with Robinson-Foulds weighted and unweighted
+metrics estimated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
+comparison between the synthetic OpenTree and original and updated tree individually,
+using OpenTree's conflict API \cite{redelings2017supertree}.
+
+\section*{Results}
+\subsection*{Case Study: The hollies}
+
+A user is interested in phylogenetic relationships within the genus *Ilex*. Commonly
+known as "hollies", the genus encompasses between 400-700 living species, and is
+the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
+plants.
+
+An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
+reveals that there are several published phylogenies showing relationships within
+the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
+selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
+\cite{gottlieb2005molecular} made original tree and alignment available in
+\href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE study 1091}.
+The tree sampling 41 species was also added to
+OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{study pg\_2827}
+and it has been integrated into OpenTree's synthetic tree
+\href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{here}.
+
+The most recent *Ilex* tree \cite{yao2020phylogeny} is available in
+OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
+and in the DRYAD repository \cite{yao2020dryad}.
+With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
+the hollies.
+
+We ran Physcraper on a laptop Linux computer to update an internal transcribed
+spacer DNA region (ITS) alignment from \cite{gottlieb2005molecular}, using a local
+GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
+analyses taking an additional 13hrs.
+<!-- **MTH: you probably need some details about the hardware, given the fact
+hat you are discussing running times** -->
+The updated \cite{gottlieb2005molecular} tree (Fig. \@ref(fig:results)) displays
+all 41 distinct
+taxa from the original study plus 231 new tips, contributing phylogenetic data to
+84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes
+with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
+A large portion of internal branches are negligibly small, with 30 branches
+< 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75
+(Figure 2).
+For comparison, \cite{yao2020phylogeny} also contains all 41 distinct taxa from
+the original
+\cite{gottlieb2005molecular} study,
+and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
+`r 135-68` are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
+ITS as a marker, their GenBank data is not released yet, so Physcraper was unable
+to incorporate 68 additional taxa into the analysis. However, Physcraper was able
+to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
+
+\section*{Discussion}
+
+Databases preserving and democratizing access to biological data
+have become essential resources for science.
+New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are needed.
+
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota [@sanderson2008phylota], PHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]) focus on generating full phylogenies *de novo*, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
+While Physcraper does not generate phylogenies *de novo* in a traditional sense,
+it successfully generates new phylogenetic knowledge, revealing the importance of
+open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
+The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
+pre-existing alignments to incorporate public molecular data into phylogenies.
+Unfortunately, installation was unsuccessful following instructions from the author, and a comparison analysis was unfeasible.
+<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
+
+Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history [@song2012resolving].
+Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], or SVD Quartets [@chifman2014quartet]).
+
+Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy [@rees2017automated], such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
+<!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
+-->
+The Physcraper workflow can be used to rapidly (in a matter of hours)
+address challenges overarching both fields of ecology and evolution, such as
+placing newly discovered species phylogenetically [@webb2010biodiversity],
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
+and generating custom trees for ecological [@helmus2012phylogenetic] and evolutionary downstream analyses [@stoltzfus2013phylotastic].
+
+Data repositories hold more information than meets the eye.
+Beyond the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Initial ideas about the data are constantly changed by results from new analyses.
+Physcraper provides a framework for reproducible phylogenetics that has the potential to consistently provide context for these ideas, highlighting the importance of data sharing and open science in the field, biology and science.
+\section*{Conclusions}
 Text for this section\ldots
-\subsection*{Sub-heading for section}
-Text for this sub-heading\ldots
-\subsubsection*{Sub-sub heading for section}
-Text for this sub-sub-heading\ldots
-\paragraph*{Sub-sub-sub heading for section}
-Text for this sub-sub-sub-heading\ldots
 
-In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
-addition, we examine a common modeling assumption and note the
-importance of considering the tails of the extinction time $T_x$ in
-studies of escape dynamics.
-We will first consider the expected resistant population at $vT_x$ for
-some $v>0$, (and temporarily assume $\alpha=0$)
-%
-\[
-E \bigl[Z_1(vT_x) \bigr]=
-\int_0^{v\wedge
-1}Z_0(uT_x)
-\exp (\lambda_1)\,du .
-\]
-%
-If we assume that sensitive cells follow a deterministic decay
-$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
-$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
-estimate the expected value as
-%
-\begin{equation}\label{eqexpmuts}
-\begin{aligned}[b]
-&      E\bigl[Z_1(vT_x)\bigr]\\
-&\quad      = \frac{\mu}{r}\log x
-\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du .
-\end{aligned}
-\end{equation}
-%
-Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,xjon,marg,schn,koha,issnic}).
+\section*{Availability and requirements}
 
-\section*{Appendix}
-Text for this section\ldots
+\textbf{Project name:} Physcraper
+
+\textbf{Project home page:} \href{https://physcraper.readthedocs.io/en/latest/index.html}{https://physcraper.readthedocs.io/en/latest/index.html}
+
+\textbf{Operating System:} Linux, Mac, Windows
+
+\textbf{Programming Language:} Python
+
+\textbf{Other requirements:} Dependencies
+
+\textbf{License:} GNU
+
+\textbf{Any restrictions to use by non-academics:} As specified by the License
+
+\section*{Abbreviations}
+
+\textbf{OpenTree:} The Open Tree of Life project
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Declarations}
 
@@
 
-\section*{Acknowledgements}%% if any
-Text for this section\ldots
+\section*{Ethics approval and consent to participate}
+Not applicable.
 
-\section*{Funding}%% if any
-Text for this section\ldots
+\section*{Consent for publication}
+Not applicable.
 
-\section*{Abbreviations}%% if any
-Text for this section\ldots
+\section*{Availability of data and materials}
 
-\section*{Availability of data and materials}%% if any
-Text for this section\ldots
-
-\section*{Ethics approval and consent to participate}%% if any
-Text for this section\ldots
+The datasets generated and analysed during the current study are available in
+the repositories "physcraper" containing the source code,
+\href{https://github.com/McTavishLab/physcraper}{https://github.com/McTavishLab/physcraper};
+"physcraperex" containing the examples,
+\href{https://github.com/McTavishLab/physcraperex}{https://github.com/McTavishLab/physcraperex};
+and, "physcraper\_ms" containing this reproducible manuscript,
+\href{https://github.com/McTavishLab/physcraper\_ms}{https://github.com/McTavishLab/physcraper\_ms}.
 
@@
 
-\section*{Consent for publication}%% if any
-Text for this section\ldots
+\section*{Funding}
+This research was supported by the grant "Sustaining the Open Tree of Life", NSF ABI
+No. 1759838, and ABI No. 1759846.
+Computer time was provided by the Multi-Environment Research Computer for
+Exploration and Discovery (MERCED) cluster from the University of California,
+Merced (UCM), supported by the NSF Grant No. ACI-1429783.
 
 \section*{Authors' contributions}
-Text for this section \ldots
+LLSR wrote manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
+All authors contributed to the manuscript and gave final approval for publication.
+\ldots
 
-\section*{Authors' information}%% if any
-Text for this section\ldots
+\section*{Acknowledgements}
+We thank the members of the OpenTree development team and the "short bar" Science
+and Engineering Building 1, UCM, joint lab paper discussion group for valuable
+comments on this manuscript.
 
@@
   \begin{figure}[h!]
-  \caption{Sample figure title}
+  \caption{The Physcraper framework consists of 4 general steps (see text). The software is fully
+  described on its documentation website at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io},
+  along with installation instructions, function usage descriptions, examples and tutorials.
+}
 \end{figure}
@@
 \begin{figure}[h!]
-  \caption{Sample figure title}
+  \caption{A) Phylogeny updated with Physcraper from original
+  \cite{gottlieb2005molecular} tree in B.
+  Tips in original alignment and new tips added with Physcraper are depicted in
+  black and red, respectively. Physcraper obtained sequences from the GenBank
+  database via local BLAST of all sequences in the original alignment that generated
+  tree in B), filtered them following criteria from section "DNA sequence search and
+  filtering", aligned them to original alignment using MUSCLE and performed a
+  phylogenetic reconstruction using RAxML with 100 bootstraps. B-D conflict analyses
+  performed with OpenTree tools.}
 \end{figure}

### 2020-11-24 sources/bmc_template_submission/bmc_article.tex
bmc discussion citations

@@
 have become essential resources for science.
-New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are needed.
+New molecular data keep accumulating and tools facilitating its integration into
+existent evolutionary knowledge are needed.
 
-Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota [@sanderson2008phylota], PHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]) focus on generating full phylogenies *de novo*, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of
+public molecular data (e.g., Phylota \cite{sanderson2008phylota},
+PHLAWD \cite{smith2009mega},
+SUPERSMART \cite{antonelli2017toward}) focus on generating full phylogenies *de novo*,
+i.e., inferring phylogenetic relationships from a newly generated homology hypothesis,
+as opposed to e.g., supertrees, that are generated by summarizing previous
+phylogenetic estimates.
 While Physcraper does not generate phylogenies *de novo* in a traditional sense,
 it successfully generates new phylogenetic knowledge, revealing the importance of
-open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
-The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
+open science in facilitating phylogenetic placement of public molecular data and
+accelerating enrichment and updating of phylogenetic relationships in any region
+of the tree of life.
+The PUMPER pipeline \cite{izquierdo2014pumper} also uses the concept of updating
 pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation was unsuccessful following instructions from the author, and a comparison analysis was unfeasible.
-<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
+Unfortunately, installation was unsuccessful following instructions from the author,
+and a comparison analysis between Physcraper and PUMPER is unfeasible at present.
+%% **MTH: add some more details about what aspect of the PUMPER install failed**
 
-Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history [@song2012resolving].
-Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], or SVD Quartets [@chifman2014quartet]).
+Physcraper generates individual gene trees, failing to capture the complexity of
+species' evolutionary history \cite{song2012resolving}. Yet, Physcraper facilitates
+gathering alignments and gene trees for multiple loci from a group of interest,
+that can be used to reconstruct species trees with ASTRAL \cite{mirarab2014astral},
+BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
 
-Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy [@rees2017automated], such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
-<!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
--->
+Physcraper can potentially link phylogenies to data available in any of the
+taxonomies integrated in the OpenTree taxonomy \cite{rees2017automated}, such as
+geographical locations from the Global Biodiversity Information Facility, or
+fossils from the Paleobiology Database.
+%% a resource that is contributing to achieve the goal of "exploring and
+%% analysing biodiversity at an accelerated pace, and returning systematics into
+%% the mainstream of science" \cite{wilson2003encyclopedia}.
 The Physcraper workflow can be used to rapidly (in a matter of hours)
 address challenges overarching both fields of ecology and evolution, such as
-placing newly discovered species phylogenetically [@webb2010biodiversity],
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
-and generating custom trees for ecological [@helmus2012phylogenetic] and evolutionary downstream analyses [@stoltzfus2013phylotastic].
+phylogenetically placing newly discovered species \cite{webb2010biodiversity},
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations
+\cite{san2010molecular},
+and generating custom trees for ecological \cite{helmus2012phylogenetic} and
+evolutionary downstream analyses \cite{stoltzfus2013phylotastic}.
 
 Data repositories hold more information than meets the eye.
-Beyond the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
+Beyond the main data, they are rich sources of metadata that can be leveraged
+for the advantage of all areas of biology as well as the advancement of scientific
+policy and applications.
 Initial ideas about the data are constantly changed by results from new analyses.
-Physcraper provides a framework for reproducible phylogenetics that has the potential to consistently provide context for these ideas, highlighting the importance of data sharing and open science in the field, biology and science.
+Physcraper provides a framework for reproducible phylogenetics that has the
+potential to consistently provide context for these ideas, highlighting the
+importance of data sharing and open science in the field, biology and science.
+
 \section*{Conclusions}

### 2020-11-25 sources/bmc_template_submission/bmc_article.tex
italics

@@
 million DNA sequences that were gauged to have the potential to resolve phylogenetic
-relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000
+relationships of 98.05\% of the almost 241,000
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
@@
 
-A user is interested in phylogenetic relationships within the genus *Ilex*. Commonly
+A user is interested in phylogenetic relationships within the genus \textit{Ilex}. Commonly
 known as "hollies", the genus encompasses between 400-700 living species, and is
@@
 
-The most recent *Ilex* tree \cite{yao2020phylogeny} is available in
+The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
 OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
@@
 analyses taking an additional 13hrs.
-<!-- **MTH: you probably need some details about the hardware, given the fact
-hat you are discussing running times** -->
-The updated \cite{gottlieb2005molecular} tree (Fig. \@ref(fig:results)) displays
+%% <!-- MTH: you probably need some details about the hardware, given the fact
+hat you are discussing running times -->
+The updated \cite{gottlieb2005molecular} tree (Figure 2) displays
 all 41 distinct
 taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes
+84 additional \textit{Ilex} taxa. The best RaxML tree is 99% resolved, with 25% of nodes
 with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
@@
 \cite{gottlieb2005molecular} study,
-and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
+and contributes phylogenetic data to 134 additional \textit{Ilex} taxa, from which
 `r 135-68` are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
@@
 PHLAWD \cite{smith2009mega},
-SUPERSMART \cite{antonelli2017toward}) focus on generating full phylogenies *de novo*,
+SUPERSMART \cite{antonelli2017toward}) focus on generating full phylogenies \textit{de novo},
 i.e., inferring phylogenetic relationships from a newly generated homology hypothesis,
@@
 phylogenetic estimates.
-While Physcraper does not generate phylogenies *de novo* in a traditional sense,
+While Physcraper does not generate phylogenies \textit{de novo} in a traditional sense,
 it successfully generates new phylogenetic knowledge, revealing the importance of

### 2020-11-25 sources/1-abstract.Rmd
last mee body

--- 2020-08-24 1-abstract.Rmd
@@
 However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and choice of one phylogeny for downstream analyses and discussion still impose difficulties
+hypotheses, and selection of one phylogeny for downstream analyses and discussion still impose difficulties
 to one that is not a specialist either on phylogenetic methods or on a particular group of study.
@@
 It also provides a framework for straightforward comparison of published phylogenies with their updated versions, by leveraging upon tools from the Open Tree of Life project to link taxonomic information across databases.
-<!-- the conflict Application Programming Interface (API) function of  the Open Tree of Life project-->
 
@@
 1.
-The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging more researchers to strive for better sharing practices.
+The Physcraper workflow demonstrates the benefits of doing open science for phylogenetics, encouraging researchers to strive for better sharing practices.
 Physcraper can be used with any OS and is released under an open-source license.

### 2020-11-25 sources/2-body.Rmd
last mee body

@@
 # knitr::include_graphics("https://raw.githubusercontent.com/McTavishLab/physcraper/pyopensci/docs/img/schematic.svg")
-knitr::include_graphics("../docs/figs/schematic-4.png")
+knitr::include_graphics("docs/figs/schematic-4.png")
 ```
@@
 # knitr::include_graphics("docs/figs/ilex-plot-tips-and-branches-1.png")
-knitr::include_graphics("../docs/figs/ilex-results.png")
+knitr::include_graphics("docs/figs/ilex-results.png")
 

### 2020-12-28 sources/bmc_template_submission/bmc_article.tex
title and authors

--- 2020-11-24 sources/bmc_template_submission/bmc_article.tex
@@
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-
-\usepackage{hyperref}
 
@@
 %%                                           %%
-%% I followed sections from here:            %%
-%% https://bmcbioinformatics.biomedcentral.com/submission-guidelines/preparing-your-manuscript/software-article
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{abstract}
-\parttitle{Background}
-Phylogenies are a key part of research in many areas of biology. Tools that automate
-some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
-have been developed for the advantage of both specialists in the field of phylogenetics and non-specialists.
-However, interpretation of results, comparison with previously available phylogenetic
-hypotheses, and selection of one phylogeny for downstream analyses and discussion still impose difficulties
-to one that is not a specialist either on phylogenetic methods or on a particular group of study.
+\begin{abstract} % abstract
+\parttitle{First part title} %if any
+Text for this section.
 
-\parttitle{Results}
-Physcraper is a command-line Python program that automates the update of published
-phylogenies by adding public DNA sequences to underlying alignments of previously
-published phylogenies. It also provides a framework for straightforward comparison
-of published phylogenies with their updated versions, by leveraging upon tools from
-the Open Tree of Life project to link taxonomic information across databases.
-The program can be used by the nonspecialist, as a tool to generate phylogenetic
-hypotheses based on publicly available expert phylogenetic knowledge.
-Phylogeneticists and taxonomic group specialists will find it useful as a tool
-to facilitate molecular dataset gathering and comparison
-of alternative phylogenetic hypotheses (topologies).
-
-\parttitle{Conclusions}
-The Physcraper workflow showcases the benefits of doing open science for
-phylogenetics, encouraging researchers to strive for better sharing practices.
-Physcraper can be used with any OS and is released under an open-source license.
-Detailed instructions for installation and
-usage are available at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io}.
-
+\parttitle{Second part title} %if any
+Text for this section.
 \end{abstract}
@@
 \begin{keyword}
-\kwd{gene tree}
-\kwd{interoperability}
-\kwd{open science}
-\kwd{reproducibility}
-\kwd{public database}
-\kwd{DNA alignment}
+\kwd{sample}
+\kwd{article}
+\kwd{author}
 \end{keyword}
@@
 %%
-\section*{Background}
-Phylogenies capture the shared history of organisms and provide key evolutionary
-context for our biological observations.
-Public biological databases constitute an amazing resource for evolutionary
-studies. Updating existing phylogenies with molecular data that has never been
-incorporated into any phylogenetic estimate, geographical location, fossils, and
-other data in a reproducible and continuous manner is possible by establishing a
-data interoperability framework for biological databases. Here, we introduce
-Physcraper, a tool that automates database connections to build upon homology
-hypotheses that taxon specialists have assessed and deemed appropriate for a
-specific phylogenetic scope to update a starting tree and single locus alignments
-with public DNA data.
+\section*{Content}
+Text and results for this section, as per the individual journal's instructions for authors.
 
-Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
-integration of data into phylogenies, which can be addressed with a unified
-taxonomy for name standardization. The Open Tree of Life project (OpenTree)
-constructs a comprehensive tree of life by synthesizing published phylogenies
-and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
-around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are
-placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic
-data from various databases \cite{rees2017automated}, including the USA National
-Center for Biodiversity Information (NCBI) molecular database GenBank
-\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
-for connecting data from any biological database that has been integrated to it.
+\section*{Section title}
+Text for this section\ldots
+\subsection*{Sub-heading for section}
+Text for this sub-heading\ldots
+\subsubsection*{Sub-sub heading for section}
+Text for this sub-sub-heading\ldots
+\paragraph*{Sub-sub-sub heading for section}
+Text for this sub-sub-sub-heading\ldots
 
-Another challenge for incorporating public molecular data into existing phylogenies
-is assembling high-quality homology hypotheses.
-While genomics has, and will continue to, revolutionize phylogenetic inference,
-the variety of alternative genomic sequencing approaches it uses produce largely
-non-overlapping genomic datasets across taxa, creating challenges in wide scale
-phylogenetic reconstruction.
+In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
+addition, we examine a common modeling assumption and note the
+importance of considering the tails of the extinction time $T_x$ in
+studies of escape dynamics.
+We will first consider the expected resistant population at $vT_x$ for
+some $v>0$, (and temporarily assume $\alpha=0$)
+%
+\[
+E \bigl[Z_1(vT_x) \bigr]=
+\int_0^{v\wedge
+1}Z_0(uT_x)
+\exp (\lambda_1)\,du .
+\]
+%
+If we assume that sensitive cells follow a deterministic decay
+$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
+$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
+estimate the expected value as
+%
+\begin{equation}\label{eqexpmuts}
+\begin{aligned}[b]
+&      E\bigl[Z_1(vT_x)\bigr]\\
+&\quad      = \frac{\mu}{r}\log x
+\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du .
+\end{aligned}
+\end{equation}
+%
+Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,xjon,marg,schn,koha,issnic}).
 
-Phylogenomics ameliorate this problem by focusing on targeted capture of informative
-loci \cite{andermann2020guide}. Yet, decades of single locus sequencing have generated
-massive amounts of homologous DNA datasets that can be used for phylogenetic
-reconstruction at many scales.
-
-More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
-million DNA sequences that were gauged to have the potential to resolve phylogenetic
-relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000
-distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-Assembling a DNA alignment from such a massive database can be done "by hand",
-but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that mine DNA databases fast, efficiently, and reproducibly,
-have been applied to infer phylogenetic relationships in a variety of organisms
-(e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
-However, fine-grained curated markers and alignments can improve phylogenetic
-reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
-
-There are almost 8,200 publicly available, peer-reviewed alignments, covering
-around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be used as seeds to mine molecular databases, and as "jump-start" alignments
-for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
-update and compare existing phylogenetic knowledge.
-
-Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access
-protocols (API's) to implement a database interoperability framework that automatically
-links phylogenies that have been standardized to OpenTree taxonomy, to alignments
-from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
-Physcraper aims to demonstrate the benefits of reproducible workflows and open
-science in phylogenetics, and encourage better data sharing practices in the community.
-
-\section*{Implementation}
-
-The general Physcraper framework (Figure 1) consists of 4 steps: 1)identifying and
-processing a tree and its underlying alignment; 2) performing a BLAST search of
-DNA sequences from original alignment on GenBank, and filtering of new sequences;
-3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic
-analysis and comparing the updated tree to existing phylogenies.
-
-\subsection*{The inputs: a tree and an alignment}
-
-Taxon names in the input tree must be standardized to OpenTree taxonomy
-\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
-\href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool. Users can upload their
-own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
-\href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have
-alignments available on TreeBASE \cite{piel2009treebase}.
-
-The input alignment is a single locus DNA dataset that was used in part or in
-whole to generate the input tree. Physcraper retrieves TreeBASE alignments
-automatically. Alternatively, users must provide the path to a local copy of the
-alignment.
-Only taxa that are both in the sequence alignment and in the tree are considered
-further for analysis; at least one taxon and its corresponding sequence are required.
-
-\subsection*{DNA sequence search and filtering}
-
-The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
-sequence search on a remote or local GenBank database. It is constrained to a
-"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
-identified using the OpenTree
-\href{<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>}{API}
-\cite{rees2017automated}, as the most recent common ancestor of ingroup taxa that is
-also a named clade in the NCBI taxonomy (Figure 1).
-
-BLAST is implemented with the `blastn` function \cite{camacho2009blast} and the BioPython
-\cite{cock2009biopython} BLAST function from
-\href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module}
-modified to accept an alternative BLAST address.
-Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
-New sequences are excluded for analysis if they 1) are not in the search taxon;
-2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
-and max length threshold, defined as the proportion of the average length without
-gaps of all sequences in input alignment (default values of 80% and 120%,
-respectively); 4) or if they are either identical to or shorter than an existing
-sequence in the input alignment and they represent the same taxon in OpenTree or
-NCBI taxonomy.
-An arbitrary maximum number of randomly chosen sequences per taxon are allowed
-(default to 5).
-
-Reverse, complement, and reverse-complement sequences are identified and translated
-using BioPython internal functions \cite{cock2009biopython}.
-Iterative cycles of BLAST searches can be performed, by blasting all new sequences
-until no new ones are found. By default only one BLAST cycle is performed.
-
-\subsection*{New DNA sequence alignment}
-
-MUSCLE \cite{edgar2004muscle} is used to perform a profile alignment in which the
-original alignment is used as a template of homology criteria to align new sequences.
-The final alignment is not further automatically checked, and additional inspection
-and refinement are recommended.
-
-\subsection*{Tree reconstruction and comparison}
-
-RAxML \cite{stamatakis2014raxml} is implemented to reconstruct a Maximum Likelihood
-(ML) gene tree for each input alignment with default settings (GTRCAT model and
-100 bootstrap replicates with default algorithm), using input tree as starting
-tree for ML searches.
-Bootstrap results are summarized using DendroPy's SumTrees module
-\cite{sukumaran2010dendropy}.
-
-Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original tree are compared with Robinson-Foulds weighted and unweighted
-metrics estimated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
-comparison between the synthetic OpenTree and original and updated tree individually,
-using OpenTree's conflict API \cite{redelings2017supertree}.
-
-\section*{Results}
-\subsection*{Case Study: The hollies}
-
-A user is interested in phylogenetic relationships within the genus *Ilex*. Commonly
-known as "hollies", the genus encompasses between 400-700 living species, and is
-the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
-plants.
-
-An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
-reveals that there are several published phylogenies showing relationships within
-the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
-selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
-\cite{gottlieb2005molecular} made original tree and alignment available in
-\href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE study 1091}.
-The tree sampling 41 species was also added to
-OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{study pg\_2827}
-and it has been integrated into OpenTree's synthetic tree
-\href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{here}.
-
-The most recent *Ilex* tree \cite{yao2020phylogeny} is available in
-OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
-and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
-the hollies.
-
-We ran Physcraper on a laptop Linux computer to update an internal transcribed
-spacer DNA region (ITS) alignment from \cite{gottlieb2005molecular}, using a local
-GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
-analyses taking an additional 13hrs.
-<!-- **MTH: you probably need some details about the hardware, given the fact
-hat you are discussing running times** -->
-The updated \cite{gottlieb2005molecular} tree (Fig. \@ref(fig:results)) displays
-all 41 distinct
-taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes
-with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
-A large portion of internal branches are negligibly small, with 30 branches
-< 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75
-(Figure 2).
-For comparison, \cite{yao2020phylogeny} also contains all 41 distinct taxa from
-the original
-\cite{gottlieb2005molecular} study,
-and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
-`r 135-68` are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
-ITS as a marker, their GenBank data is not released yet, so Physcraper was unable
-to incorporate 68 additional taxa into the analysis. However, Physcraper was able
-to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
-
-\section*{Discussion}
-
-Databases preserving and democratizing access to biological data
-have become essential resources for science.
-New molecular data keep accumulating and tools facilitating its integration into existent evolutionary knowledge are needed.
-
-Phylogenetic pipelines designed to make evolutionary sense of the vast amount of public molecular data (e.g., Phylota [@sanderson2008phylota], PHLAWD [@smith2009mega], SUPERSMART [@antonelli2017toward]) focus on generating full phylogenies *de novo*, i.e., inferring phylogenetic relationships from a newly generated homology hypothesis, as opposed to e.g., supertrees, that are generated by summarizing previous phylogenetic estimates.
-While Physcraper does not generate phylogenies *de novo* in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the importance of
-open science in facilitating phylogenetic placement of public molecular data and accelerating enrichment and updating of phylogenetic relationships in any region of the tree of life.
-The PUMPER pipeline [@izquierdo2014pumper] also uses the concept of updating
-pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation was unsuccessful following instructions from the author, and a comparison analysis was unfeasible.
-<!-- **MTH: add some more details about what aspect of the PUMPER install failed** -->
-
-Physcraper generates individual gene trees, failing to capture the complexity of species' evolutionary history [@song2012resolving].
-Yet, Physcraper facilitates gathering alignments and gene trees for multiple loci from a group of interest, that can be used to reconstruct species trees with ASTRAL [@mirarab2014astral], BEAST2 [@bouckaert2019beast], or SVD Quartets [@chifman2014quartet]).
-
-Physcraper can potentially link phylogenies to data available in any of the taxonomies integrated in the OpenTree taxonomy [@rees2017automated], such as geographical locations from the Global Biodiversity Information Facility, or fossils from the Paleobiology Database.
-<!-- a resource that is contributing to achieve the goal of "exploring and analysing biodiversity at an accelerated pace, and returning systematics into the mainstream of science" [@wilson2003encyclopedia].
--->
-The Physcraper workflow can be used to rapidly (in a matter of hours)
-address challenges overarching both fields of ecology and evolution, such as
-placing newly discovered species phylogenetically [@webb2010biodiversity],
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations [@san2010molecular],
-and generating custom trees for ecological [@helmus2012phylogenetic] and evolutionary downstream analyses [@stoltzfus2013phylotastic].
-
-Data repositories hold more information than meets the eye.
-Beyond the main data, they are rich sources of metadata that can be leveraged for the advantage of all areas of biology as well as the advancement of scientific policy and applications.
-Initial ideas about the data are constantly changed by results from new analyses.
-Physcraper provides a framework for reproducible phylogenetics that has the potential to consistently provide context for these ideas, highlighting the importance of data sharing and open science in the field, biology and science.
-\section*{Conclusions}
+\section*{Appendix}
 Text for this section\ldots
-
-\section*{Availability and requirements}
-
-\textbf{Project name:} Physcraper
-
-\textbf{Project home page:} \href{https://physcraper.readthedocs.io/en/latest/index.html}{https://physcraper.readthedocs.io/en/latest/index.html}
-
-\textbf{Operating System:} Linux, Mac, Windows
-
-\textbf{Programming Language:} Python
-
-\textbf{Other requirements:} Dependencies
-
-\textbf{License:} GNU
-
-\textbf{Any restrictions to use by non-academics:} As specified by the License
-
-\section*{Abbreviations}
-
-\textbf{OpenTree:} The Open Tree of Life project
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Declarations}
 
@@
 
-\section*{Ethics approval and consent to participate}
-Not applicable.
+\section*{Acknowledgements}%% if any
+Text for this section\ldots
 
-\section*{Consent for publication}
-Not applicable.
+\section*{Funding}%% if any
+Text for this section\ldots
 
-\section*{Availability of data and materials}
+\section*{Abbreviations}%% if any
+Text for this section\ldots
 
-The datasets generated and analysed during the current study are available in
-the repositories "physcraper" containing the source code,
-\href{https://github.com/McTavishLab/physcraper}{https://github.com/McTavishLab/physcraper};
-"physcraperex" containing the examples,
-\href{https://github.com/McTavishLab/physcraperex}{https://github.com/McTavishLab/physcraperex};
-and, "physcraper\_ms" containing this reproducible manuscript,
-\href{https://github.com/McTavishLab/physcraper\_ms}{https://github.com/McTavishLab/physcraper\_ms}.
+\section*{Availability of data and materials}%% if any
+Text for this section\ldots
+
+\section*{Ethics approval and consent to participate}%% if any
+Text for this section\ldots
 
@@
 
-\section*{Funding}
-This research was supported by the grant "Sustaining the Open Tree of Life", NSF ABI
-No. 1759838, and ABI No. 1759846.
-Computer time was provided by the Multi-Environment Research Computer for
-Exploration and Discovery (MERCED) cluster from the University of California,
-Merced (UCM), supported by the NSF Grant No. ACI-1429783.
+\section*{Consent for publication}%% if any
+Text for this section\ldots
 
 \section*{Authors' contributions}
-LLSR wrote manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
-All authors contributed to the manuscript and gave final approval for publication.
-\ldots
+Text for this section \ldots
 
-\section*{Acknowledgements}
-We thank the members of the OpenTree development team and the "short bar" Science
-and Engineering Building 1, UCM, joint lab paper discussion group for valuable
-comments on this manuscript.
+\section*{Authors' information}%% if any
+Text for this section\ldots
 
@@
   \begin{figure}[h!]
-  \caption{The Physcraper framework consists of 4 general steps (see text). The software is fully
-  described on its documentation website at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io},
-  along with installation instructions, function usage descriptions, examples and tutorials.
-}
+  \caption{Sample figure title}
 \end{figure}
@@
 \begin{figure}[h!]
-  \caption{A) Phylogeny updated with Physcraper from original
-  \cite{gottlieb2005molecular} tree in B.
-  Tips in original alignment and new tips added with Physcraper are depicted in
-  black and red, respectively. Physcraper obtained sequences from the GenBank
-  database via local BLAST of all sequences in the original alignment that generated
-  tree in B), filtered them following criteria from section "DNA sequence search and
-  filtering", aligned them to original alignment using MUSCLE and performed a
-  phylogenetic reconstruction using RAxML with 100 bootstraps. B-D conflict analyses
-  performed with OpenTree tools.}
+  \caption{Sample figure title}
 \end{figure}

### 2020-12-28 sources/bmc_template_submission/bmc_article.tex
keywords
abstract

@@
 
-\begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
+\begin{abstract}
+\parttitle{Background}
+Phylogenies are a key part of research in many areas of biology. Tools that automate
+some parts of the process of phylogenetic reconstruction, mainly molecular character matrix assembly,
+have been developed for the advantage of both specialists in the field of phylogenetics and non-specialists.
+However, interpretation of results, comparison with previously available phylogenetic
+hypotheses, and selection of one phylogeny for downstream analyses and discussion still impose difficulties
+to one that is not a specialist either on phylogenetic methods or on a particular group of study.
 
-\parttitle{Second part title} %if any
-Text for this section.
+\parttitle{Results}
+Physcraper is a command-line Python program that automates the update of published
+phylogenies by adding public DNA sequences to underlying alignments of previously
+published phylogenies. It also provides a framework for straightforward comparison
+of published phylogenies with their updated versions, by leveraging upon tools from
+the Open Tree of Life project to link taxonomic information across databases.
+
+Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
+hypotheses based on publicly available expert phylogenetic knowledge.
+Phylogeneticists and taxonomic group specialists will find it useful as a tool
+to facilitate molecular dataset gathering and comparison
+of alternative phylogenetic hypotheses (topologies).
+
+\parttitle{Conclusions}
+The Physcraper workflow demonstrates the benefits of doing open science for
+phylogenetics, encouraging researchers to strive for better sharing practices.
+Physcraper can be used with any OS and is released under an open-source license.
+Detailed instructions for installation and
+use are available at [https://physcraper.readthedocs](https://physcraper.readthedocs.io).
+
 \end{abstract}
@@
 \begin{keyword}
-\kwd{sample}
-\kwd{article}
-\kwd{author}
+\kwd{gene tree}
+\kwd{interoperability}
+\kwd{open science}
+\kwd{reproducibility}
+\kwd{public database}
+\kwd{dna alignment}
 \end{keyword}

### 2020-12-28 sources/bmc_template_submission/bmc_article.tex
abstract

@@
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+
+\usepackage{hyperref}
 
@@
 the Open Tree of Life project to link taxonomic information across databases.
-
-Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
+The program can be used by the nonspecialist, as a tool to generate phylogenetic
 hypotheses based on publicly available expert phylogenetic knowledge.
@@
 Detailed instructions for installation and
-use are available at [https://physcraper.readthedocs](https://physcraper.readthedocs.io).
+usage are available at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io}.
 

### 2020-12-30 sources/bmc_template_submission/bmc_article.tex
sections
declarations

@@
 %%
-\section*{Content}
+\section*{Background}
 Text and results for this section, as per the individual journal's instructions for authors.
 
-\section*{Section title}
+\section*{Implementation}
 Text for this section\ldots
@@
 
-\section*{Appendix}
+\section*{Results}
 Text for this section\ldots
+
+\section*{Discussion}
+Text for this section\ldots
+
+\section*{Conclusions}
+Text for this section\ldots
+
+\section*{Availability and requirements}
+
+\textbf{Project name:} Physcraper
+
+\textbf{Project home page:} \href{https://physcraper.readthedocs.io/en/latest/index.html}{https://physcraper.readthedocs.io/en/latest/index.html}
+
+\textbf{Operating System:} Linux, Mac, Windows
+
+\textbf{Programming Language:} Python
+
+\textbf{Other requirements:} Dependencies
+
+\textbf{License:} GNU
+
+\textbf{Any restrictions to use by non-academics:} As specified by the License
+
+\section*{Abbreviations}
+
+\textbf{OpenTree:} The Open Tree of Life project
 

### 2020-12-30 sources/bmc_template_submission/bmc_article.tex
bib file

@@
 
-In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
-addition, we examine a common modeling assumption and note the
-importance of considering the tails of the extinction time $T_x$ in
-studies of escape dynamics.
-We will first consider the expected resistant population at $vT_x$ for
-some $v>0$, (and temporarily assume $\alpha=0$)
-%
-\[
-E \bigl[Z_1(vT_x) \bigr]=
-\int_0^{v\wedge
-1}Z_0(uT_x)
-\exp (\lambda_1)\,du .
-\]
-%
-If we assume that sensitive cells follow a deterministic decay
-$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
-$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
-estimate the expected value as
-%
-\begin{equation}\label{eqexpmuts}
-\begin{aligned}[b]
-&      E\bigl[Z_1(vT_x)\bigr]\\
-&\quad      = \frac{\mu}{r}\log x
-\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du .
-\end{aligned}
-\end{equation}
-%
-Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,xjon,marg,schn,koha,issnic}).
+(also see \cite{koon,xjon,marg,schn,koha,issnic}).
 

### 2020-12-30 sources/bmc_template_submission/bmc_article.tex
bib file
implementation sections

@@
 \parttitle{Conclusions}
-The Physcraper workflow demonstrates the benefits of doing open science for
+The Physcraper workflow showcases the benefits of doing open science for
 phylogenetics, encouraging researchers to strive for better sharing practices.
@@
 \kwd{public database}
-\kwd{dna alignment}
+\kwd{DNA alignment}
 \end{keyword}
@@
 \section*{Background}
-Text and results for this section, as per the individual journal's instructions for authors.
+Phylogenies capture the shared history of organisms and provide key evolutionary
+context for our biological observations.
+Public biological databases constitute an amazing resource for evolutionary
+studies. Updating existing phylogenies with molecular data that has never been
+incorporated into any phylogenetic estimate, geographical location, fossils, and
+other data in a reproducible and continuous manner is possible by establishing a
+data interoperability framework for biological databases. Here, we introduce
+Physcraper, a tool that automates database connections to build upon homology
+hypotheses that taxon specialists have assessed and deemed appropriate for a
+specific phylogenetic scope to update a starting tree and single locus alignments
+with public DNA data.
+
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
+integration of data into phylogenies, which can be addressed with a unified
+taxonomy for name standardization. The Open Tree of Life project (OpenTree)
+constructs a comprehensive tree of life by synthesizing published phylogenies
+and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
+around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are
+placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic
+data from various databases \cite{rees2017automated}, including the USA National
+Center for Biodiversity Information (NCBI) molecular database GenBank
+\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
+for connecting data from any biological database that has been integrated to it.
+
+Another challenge for incorporating public molecular data into existing phylogenies
+is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference,
+the variety of alternative genomic sequencing approaches it uses produce largely
+non-overlapping genomic datasets across taxa, creating challenges in wide scale
+phylogenetic reconstruction.
+
+Phylogenomics ameliorate this problem by focusing on targeted capture of informative
+loci \cite{andermann2020guide}. Yet, decades of single locus sequencing have generated
+massive amounts of homologous DNA datasets that can be used for phylogenetic
+reconstruction at many scales.
+
+More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
+million DNA sequences that were gauged to have the potential to resolve phylogenetic
+relationships of `r round(236023/240708, digits=4)*100`% of the almost 241,000
+distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
+Assembling a DNA alignment from such a massive database can be done "by hand",
+but it is a largely time consuming and mostly non-reproducible approach.
+Computational pipelines that mine DNA databases fast, efficiently, and reproducibly,
+have been applied to infer phylogenetic relationships in a variety of organisms
+(e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
+However, fine-grained curated markers and alignments can improve phylogenetic
+reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
+
+There are almost 8,200 publicly available, peer-reviewed alignments, covering
+around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
+can be used as seeds to mine molecular databases, and as "jump-start" alignments
+for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
+update and compare existing phylogenetic knowledge.
+
+Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access
+protocols (API's) to implement a database interoperability framework that automatically
+links phylogenies that have been standardized to OpenTree taxonomy, to alignments
+from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
+Physcraper aims to demonstrate the benefits of reproducible workflows and open
+science in phylogenetics, and encourage better data sharing practices in the community.
 
 \section*{Implementation}
-Text for this section\ldots
-\subsection*{Sub-heading for section}
-Text for this sub-heading\ldots
-\subsubsection*{Sub-sub heading for section}
-Text for this sub-sub-heading\ldots
-\paragraph*{Sub-sub-sub heading for section}
-Text for this sub-sub-sub-heading\ldots
+
+The general Physcraper framework (Figure 1) consists of 4 steps: 1)identifying and
+processing a tree and its underlying alignment; 2) performing a BLAST search of
+DNA sequences from original alignment on GenBank, and filtering of new sequences;
+3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic
+analysis and comparing the updated tree to existing phylogenies.
+
+\subsection*{The inputs: a tree and an alignment}
+
+Taxon names in the input tree must be standardized to OpenTree taxonomy
+\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
+\href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool. Users can upload their
+own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
+\href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have
+alignments available on TreeBASE \cite{piel2009treebase}.
+
+The input alignment is a single locus DNA dataset that was used in part or in
+whole to generate the input tree. Physcraper retrieves TreeBASE alignments
+automatically. Alternatively, users must provide the path to a local copy of the
+alignment.
+Only taxa that are both in the sequence alignment and in the tree are considered
+further for analysis; at least one taxon and its corresponding sequence are required.
+
+\subsection*{DNA sequence search and filtering}
+
+The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
+sequence search on a remote or local GenBank database. It is constrained to a
+"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
+identified using the OpenTree
+\href{<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>}{API}
+\cite{rees2017automated}, as the most recent common ancestor of ingroup taxa that is
+also a named clade in the NCBI taxonomy (Figure 1).
+
+BLAST is implemented with the `blastn` function \cite{camacho2009blast} and the BioPython
+\cite{cock2009biopython} BLAST function from
+\href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module}
+modified to accept an alternative BLAST address.
+Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
+New sequences are excluded for analysis if they 1) are not in the search taxon;
+2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
+and max length threshold, defined as the proportion of the average length without
+gaps of all sequences in input alignment (default values of 80% and 120%,
+respectively); 4) or if they are either identical to or shorter than an existing
+sequence in the input alignment and they represent the same taxon in OpenTree or
+NCBI taxonomy.
+An arbitrary maximum number of randomly chosen sequences per taxon are allowed
+(default to 5).
+
+Reverse, complement, and reverse-complement sequences are identified and translated
+using BioPython internal functions \cite{cock2009biopython}.
+Iterative cycles of BLAST searches can be performed, by blasting all new sequences
+until no new ones are found. By default only one BLAST cycle is performed.
+
+\subsection*{New DNA sequence alignment}
+
+MUSCLE \cite{edgar2004muscle} is used to perform a profile alignment in which the
+original alignment is used as a template of homology criteria to align new sequences.
+The final alignment is not further automatically checked, and additional inspection
+and refinement are recommended.
+
+\subsection*{Tree reconstruction and comparison}
+
+RAxML \cite{stamatakis2014raxml} is implemented to reconstruct a Maximum Likelihood
+(ML) gene tree for each input alignment with default settings (GTRCAT model and
+100 bootstrap replicates with default algorithm), using input tree as starting
+tree for ML searches.
+Bootstrap results are summarized using DendroPy's SumTrees module
+\cite{sukumaran2010dendropy}.
+
+Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
+Updated and original tree are compared with Robinson-Foulds weighted and unweighted
+metrics estimated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
+comparison between the synthetic OpenTree and original and updated tree individually,
+using OpenTree's conflict API \cite{redelings2017supertree}.
 

### 2020-12-31 sources/bmc_template_submission/bmc_article.tex
example section

@@
 
-(also see \cite{koon,xjon,marg,schn,koha,issnic}).
+\section*{Results}
+\subsection*{Case Study: The hollies}
 
-\section*{Results}
-Text for this section\ldots
+A user is interested in phylogenetic relationships within the genus *Ilex*. Commonly
+known as "hollies", the genus encompasses between 400-700 living species, and is
+the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
+plants.
+
+An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
+reveals that there are several published phylogenies showing relationships within
+the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
+selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
+\cite{gottlieb2005molecular} made original tree and alignment available in
+\href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE study 1091}.
+The tree sampling 41 species was also added to
+OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{study pg\_2827}
+and it has been integrated into OpenTree's synthetic tree
+\href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{here}.
+
+The most recent *Ilex* tree \cite{yao2020phylogeny} is available in
+OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
+and in the DRYAD repository \cite{yao2020dryad}.
+With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
+the hollies.
+
+We ran Physcraper on a laptop Linux computer to update an internal transcribed
+spacer DNA region (ITS) alignment from \cite{gottlieb2005molecular}, using a local
+GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
+analyses taking an additional 13hrs.
+<!-- **MTH: you probably need some details about the hardware, given the fact
+hat you are discussing running times** -->
+The updated \cite{gottlieb2005molecular} tree (Fig. \@ref(fig:results)) displays
+all 41 distinct
+taxa from the original study plus 231 new tips, contributing phylogenetic data to
+84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes
+with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
+A large portion of internal branches are negligibly small, with 30 branches
+< 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75
+(Figure 2).
+For comparison, \cite{yao2020phylogeny} also contains all 41 distinct taxa from
+the original
+\cite{gottlieb2005molecular} study,
+and contributes phylogenetic data to 134 additional *Ilex* taxa, from which
+`r 135-68` are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
+ITS as a marker, their GenBank data is not released yet, so Physcraper was unable
+to incorporate 68 additional taxa into the analysis. However, Physcraper was able
+to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
 

### 2021-01-05 sources/bmc_template_submission/bmc_article.tex
math formatting

@@
 %\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-%\RequirePackage{hyperref}
+\RequirePackage{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
@@
 analyses taking an additional 13hrs.
-<!-- **MTH: you probably need some details about the hardware, given the fact
-hat you are discussing running times** -->
-The updated \cite{gottlieb2005molecular} tree (Fig. \@ref(fig:results)) displays
+%% <!-- MTH: you probably need some details about the hardware, given the fact hat you are discussing running times -->
+The updated \cite{gottlieb2005molecular} tree (Figure 2) displays
 all 41 distinct
 taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional *Ilex* taxa. The best RaxML tree is 99% resolved, with 25% of nodes
-with bootstrap support < 0.1 and 48% nodes with bootstrap support > 0.75.
+84 additional \textit{Ilex} taxa. The best RaxML tree is 99\% resolved, with 25\% of nodes
+with bootstrap support $<$ 0.1 and 48\% nodes with bootstrap support $>$ 0.75.
 A large portion of internal branches are negligibly small, with 30 branches
-< 0.00001 substitution rate units, from which only 9 have a bootstrap support > 0.75
+< 0.00001 substitution rate units, from which only 9 have a bootstrap support $>$ 0.75
 (Figure 2).
@@
 \section*{Discussion}
-Text for this section\ldots
+
+Databases preserving and democratizing access to biological data
+have become essential resources for science.
+New molecular data keep accumulating and tools facilitating its integration into
+existent evolutionary knowledge are needed.
+
+Phylogenetic pipelines designed to make evolutionary sense of the vast amount of
+public molecular data (e.g., Phylota \cite{sanderson2008phylota},
+PHLAWD \cite{smith2009mega},
+SUPERSMART \cite{antonelli2017toward}) focus on generating full phylogenies \textit{de novo},
+i.e., inferring phylogenetic relationships from a newly generated homology hypothesis,
+as opposed to e.g., supertrees, that are generated by summarizing previous
+phylogenetic estimates.
+While Physcraper does not generate phylogenies \textit{de novo} in a traditional sense,
+it successfully generates new phylogenetic knowledge, revealing the importance of
+open science in facilitating phylogenetic placement of public molecular data and
+accelerating enrichment and updating of phylogenetic relationships in any region
+of the tree of life.
+The PUMPER pipeline \cite{izquierdo2014pumper} also uses the concept of updating
+pre-existing alignments to incorporate public molecular data into phylogenies.
+Unfortunately, installation was unsuccessful following instructions from the author,
+and a comparison analysis between Physcraper and PUMPER is unfeasible at present.
+%% **MTH: add some more details about what aspect of the PUMPER install failed**
+
+Physcraper generates individual gene trees, failing to capture the complexity of
+species' evolutionary history \cite{song2012resolving}. Yet, Physcraper facilitates
+gathering alignments and gene trees for multiple loci from a group of interest,
+that can be used to reconstruct species trees with ASTRAL \cite{mirarab2014astral},
+BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
+
+Physcraper can potentially link phylogenies to data available in any of the
+taxonomies integrated in the OpenTree taxonomy \cite{rees2017automated}, such as
+geographical locations from the Global Biodiversity Information Facility, or
+fossils from the Paleobiology Database.
+%% a resource that is contributing to achieve the goal of "exploring and
+%% analysing biodiversity at an accelerated pace, and returning systematics into
+%% the mainstream of science" \cite{wilson2003encyclopedia}.
+The Physcraper workflow can be used to rapidly (in a matter of hours)
+address challenges overarching both fields of ecology and evolution, such as
+phylogenetically placing newly discovered species \cite{webb2010biodiversity},
+systematizing molecular (and other) databases, i.e., curating taxonomic assignations
+\cite{san2010molecular},
+and generating custom trees for ecological \cite{helmus2012phylogenetic} and
+evolutionary downstream analyses \cite{stoltzfus2013phylotastic}.
+
+Data repositories hold more information than meets the eye.
+Beyond the main data, they are rich sources of metadata that can be leveraged
+for the advantage of all areas of biology as well as the advancement of scientific
+policy and applications.
+Initial ideas about the data are constantly changed by results from new analyses.
+Physcraper provides a framework for reproducible phylogenetics that has the
+potential to consistently provide context for these ideas, highlighting the
+importance of data sharing and open science in the field, biology and science.
 

### 2021-01-05 sources/bmc_template_submission/bmc_article.tex
intro again

@@
 Phylogenies capture the shared history of organisms and provide key evolutionary
-context for our biological observations.
+context for our biological observations \cite{Dobzhansky maybe}.
+A challenge for phylogenetic reconstruction
+is assembling high-quality homology hypotheses.
+While genomics has, and will continue to, revolutionize phylogenetic inference,
+the diversity of alternative genomic sequencing approaches implemented produce
+largely non-overlapping homology hypotheses across taxa \cite{BLA}, creating challenges for
+reconstructing the shared evolutionary history of taxa separated phylogenetically
+at deep scales \cite{BLA}.
+
+While phylogenomics tackles this problem by focusing on targeted capture of informative
+regions \cite{andermann2020guide, it is expensive. Few labs around the world can afford
+to implement it for all their study groups, or at all. It does not contribute to
+democratization of research yet.
+
+Decades of single locus sequencing have generated
+massive amounts of homologous DNA datasets that can be used for phylogenetic
+reconstruction at many scales.
+More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
+million DNA sequences that were gauged to have the potential to resolve phylogenetic
+relationships of 98.05\% of the almost 241,000
+distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
+How much of this is available on OpenTree??
+
+Is genomic data on genbank available for a smaller percentage of all life, compared to
+single loci in the same database?
+
+Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
+Magallon, Soltis, examples with fish, bats,
+
+These studies are widely cited and immensely relevant for the biological community.
+
+Some of these multilocus phylogenetic studies have implemented pipelines such as PHLAWD,
+other studies have curated their matrices by hand, which is largely non-reproducible.
+
 Public biological databases constitute an amazing resource for evolutionary
@@
 
-Another challenge for incorporating public molecular data into existing phylogenies
-is assembling high-quality homology hypotheses.
-While genomics has, and will continue to, revolutionize phylogenetic inference,
-the variety of alternative genomic sequencing approaches it uses produce largely
-non-overlapping genomic datasets across taxa, creating challenges in wide scale
-phylogenetic reconstruction.
+----------
 
-Phylogenomics ameliorate this problem by focusing on targeted capture of informative
-loci \cite{andermann2020guide}. Yet, decades of single locus sequencing have generated
-massive amounts of homologous DNA datasets that can be used for phylogenetic
-reconstruction at many scales.
-
-More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
-million DNA sequences that were gauged to have the potential to resolve phylogenetic
-relationships of 98.05\% of the almost 241,000
-distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
 Assembling a DNA alignment from such a massive database can be done "by hand",

### 2021-01-06 sources/bmc_template_submission/bmc_article.tex
intro

@@
 Phylogenies capture the shared history of organisms and provide key evolutionary
-context for our biological observations \cite{Dobzhansky maybe}.
+context for our biological observations \cite{dobzhansky1973nothing}.
 A challenge for phylogenetic reconstruction
@@
 the diversity of alternative genomic sequencing approaches implemented produce
-largely non-overlapping homology hypotheses across taxa \cite{BLA}, creating challenges for
-reconstructing the shared evolutionary history of taxa separated phylogenetically
-at deep scales \cite{BLA}.
+largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
+creating challenges for phylogenetic reconstruction.
 
 While phylogenomics tackles this problem by focusing on targeted capture of informative
-regions \cite{andermann2020guide, it is expensive. Few labs around the world can afford
+regions \cite{andermann2020guide}, it is expensive. Few labs around the world can afford
 to implement it for all their study groups, or at all. It does not contribute to
-democratization of research yet.
+democratization of research yet \cite{li2017enabling}.
 
@@
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-How much of this is available on OpenTree??
+The Open Tree of Life project (OpenTree)
+constructs a comprehensive tree of life by synthesizing published phylogenies
+and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
+around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
+placed in the tree based on taxonomy. Meaning that there is a lot of information
 
@@
 
-Public biological databases constitute an amazing resource for evolutionary
-studies. Updating existing phylogenies with molecular data that has never been
-incorporated into any phylogenetic estimate, geographical location, fossils, and
-other data in a reproducible and continuous manner is possible by establishing a
-data interoperability framework for biological databases. Here, we introduce
-Physcraper, a tool that automates database connections to build upon homology
-hypotheses that taxon specialists have assessed and deemed appropriate for a
-specific phylogenetic scope to update a starting tree and single locus alignments
-with public DNA data.
-
-Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
-integration of data into phylogenies, which can be addressed with a unified
-taxonomy for name standardization. The Open Tree of Life project (OpenTree)
-constructs a comprehensive tree of life by synthesizing published phylogenies
-and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
-around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are
-placed in the tree based on taxonomy. To achieve this, OpenTree unifies taxonomic
-data from various databases \cite{rees2017automated}, including the USA National
-Center for Biodiversity Information (NCBI) molecular database GenBank
-\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
-for connecting data from any biological database that has been integrated to it.
-
-----------
+Make emphasis on these really great studies (and pipelines) that do not rely on
+genomic data, and that there is a lot of single locus data that can be analysed
 
@@
 update and compare existing phylogenetic knowledge.
+
+Public biological databases constitute an amazing resource for evolutionary
+studies. Updating existing phylogenies with molecular data that has never been
+incorporated into any phylogenetic estimate, geographical location, fossils, and
+other data in a reproducible and continuous manner is possible by establishing a
+data interoperability framework for biological databases. Here, we introduce
+Physcraper, a tool that automates database connections to build upon homology
+hypotheses that taxon specialists have assessed and deemed appropriate for a
+specific phylogenetic scope to update a starting tree and single locus alignments
+with public DNA data.
+
+Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
+integration of data into phylogenies, which can be addressed with a unified
+taxonomy for name standardization.
+
+To achieve this, OpenTree unifies taxonomic
+data from various databases \cite{rees2017automated}, including the USA National
+Center for Biodiversity Information (NCBI) molecular database GenBank
+\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
+for connecting data from any biological database that has been integrated to it.
+
+----------
 

### 2021-01-06 100-phylota-overview.Rmd
intro clean up
intro

@@
     uses phylota to get a DNA data set for 132 unique taxa of vascular plants from
-    16 families and 25 genera,and a tree of 33 genera from 20 different families
+    16 families and 25 genera, and a tree of 33 genera from 20 different families
     comprising 1706 taxa.
@@
     uses its own algorithm (SOPHI) to mine public DNA databases [@chesters2014protocol].
-    It does not cite phylota as it should, but includes it in their references.
+    It does not reference to phylota in the main text (they should bc they used it), but includes it in their references.
 5. When phylota was used to extract full trees (not only DNA data sets or markers):

### 2021-01-06 sources/bmc_template_submission/bmc_article.tex
intro

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-A challenge for phylogenetic reconstruction
-is assembling high-quality homology hypotheses.
-While genomics has, and will continue to, revolutionize phylogenetic inference,
-the diversity of alternative genomic sequencing approaches implemented produce
-largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
-creating challenges for phylogenetic reconstruction.
-
-While phylogenomics tackles this problem by focusing on targeted capture of informative
-regions \cite{andermann2020guide}, it is expensive. Few labs around the world can afford
-to implement it for all their study groups, or at all. It does not contribute to
-democratization of research yet \cite{li2017enabling}.
+The Open Tree of Life project (OpenTree)
+constructs a comprehensive tree of life by synthesizing published phylogenies
+and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
+around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
+placed in the tree based on taxonomy. Meaning that there is still a lot of work
+to do to complete a full phylogenetic tree of all life.
 
 Decades of single locus sequencing have generated
-massive amounts of homologous DNA datasets that can be used for phylogenetic
-reconstruction at many scales.
+massive amounts of homologous DNA datasets that have the potential to be used for phylogenetic
+reconstruction at many scales \cite{sanderson2008phylota}.
 More than a decade ago, GenBank release 159 (April 15, 2007) already hosted 72
@@
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-The Open Tree of Life project (OpenTree)
-constructs a comprehensive tree of life by synthesizing published phylogenies
-and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
-around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
-placed in the tree based on taxonomy. Meaning that there is a lot of information
 
-Is genomic data on genbank available for a smaller percentage of all life, compared to
-single loci in the same database?
+With more than XXXX million single locus
+sequences in GenBank today, there is a considerable amount of phylogenetically
+informative data in GenBank that has not been analysed and/or made publicly available
+with the potential to give phylogenetic light to the tree of life.
+
+Phylogenies constructed from single locus super mayrix datasets are still widely
+generated and relevant to the field. Notable examples include the
 
@@
 Magallon, Soltis, examples with fish, bats,
+\cite{magallon2015metacalibrated} mining by hand - flowering plant families - has been cited 561 times
+\cite{smith2018constructing} 38k clades of plants with seeds - cited 176 times
+\cite{jetz2012global} 6k bird species - cited
+\cite{rabosky2018inverse} mining with phlawd - cited 202 times
 
@@
 Make emphasis on these really great studies (and pipelines) that do not rely on
-genomic data, and that there is a lot of single locus data that can be analysed
+genomic data, and that there is a lot of single locus data that can be analysed.
 
@@
 update and compare existing phylogenetic knowledge.
-
-Public biological databases constitute an amazing resource for evolutionary
-studies. Updating existing phylogenies with molecular data that has never been
-incorporated into any phylogenetic estimate, geographical location, fossils, and
-other data in a reproducible and continuous manner is possible by establishing a
-data interoperability framework for biological databases. Here, we introduce
-Physcraper, a tool that automates database connections to build upon homology
-hypotheses that taxon specialists have assessed and deemed appropriate for a
-specific phylogenetic scope to update a starting tree and single locus alignments
-with public DNA data.
 

### 2021-01-06 temp.Rmd
intro

While genomics has, and will continue to, revolutionize phylogenetic inference,
the diversity of alternative genomic sequencing approaches implemented produce
largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
creating challenges for phylogenetic reconstruction.
Phylogenomics addresses this problem by focusing on targeted capture of informative
regions \cite{andermann2020guide}. Yet, it is expensive. Few labs around the world can afford
to implement it for all their study groups, or at all. Creating a wealth gap for
science, while free information is still available to provide new knowledge.
Genomics does not contribute yet to democratization of phylogenetic research \cite{li2017enabling}.

%% Is genomic data on genbank available for a smaller percentage of all life, compared to
%% single loci in the same database?

Public biological databases constitute an amazing resource for evolutionary
studies. Connecting existing phylogenies to molecular data that has never been
incorporated into any phylogenetic estimate, geographical location, fossils, and
other data in a reproducible and continuous manner is possible by establishing a
data interoperability framework for biological databases. Here, we introduce
Physcraper, a tool that automates biological database connections
with the main goal of building upon homology
hypotheses that taxon specialists have assessed and deemed appropriate for a
specific phylogenetic scope to update a starting tree and single locus alignments
with public DNA data.

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
intro

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
+Public biological databases constitute an amazing resource for evolutionary
+studies. Connecting existing phylogenies to molecular data that has never been
+incorporated into any phylogenetic estimate, geographical location, fossils, and
+other data in a reproducible and continuous manner is possible by establishing a
+data interoperability framework for biological databases. Here, we introduce
+Physcraper, a tool that automates biological database connections
+with the main goal of building upon homology
+hypotheses that taxon specialists have assessed and deemed appropriate for a
+specific phylogenetic scope to update a starting tree and single locus alignments
+with public DNA data, with the added capability to connect the lineages represented
+in the phylogeny to biological data in public databases.
+
+One of the main challenges for automatic
+integration of biological data into phylogenies are varying taxonomic idiosyncrasies
+across databases. To address this challenge, the Open Tree of Life project (OpenTree)
+unifies taxonomic
+data from various databases \cite{rees2017automated}, including the USA National
+Center for Biodiversity Information (NCBI) molecular database GenBank
+\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
+that remains to be leveraged for connecting data from any biological database that
+has been integrated to it.
+
 The Open Tree of Life project (OpenTree)
-constructs a comprehensive tree of life by synthesizing published phylogenies
+also constructs a comprehensive tree of life by synthesizing published phylogenies
 and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
@@
 update and compare existing phylogenetic knowledge.
-
-Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
-integration of data into phylogenies, which can be addressed with a unified
-taxonomy for name standardization.
-
-To achieve this, OpenTree unifies taxonomic
-data from various databases \cite{rees2017automated}, including the USA National
-Center for Biodiversity Information (NCBI) molecular database GenBank
-\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
-for connecting data from any biological database that has been integrated to it.
 

### 2021-01-07 temp.Rmd
intro

@@
 
-Public biological databases constitute an amazing resource for evolutionary
-studies. Connecting existing phylogenies to molecular data that has never been
-incorporated into any phylogenetic estimate, geographical location, fossils, and
-other data in a reproducible and continuous manner is possible by establishing a
-data interoperability framework for biological databases. Here, we introduce
-Physcraper, a tool that automates biological database connections
-with the main goal of building upon homology
-hypotheses that taxon specialists have assessed and deemed appropriate for a
-specific phylogenetic scope to update a starting tree and single locus alignments
-with public DNA data.
+\cite{pyron2011phylogeny} PHLAWD, Colubroidea, 334 citations
+
+Large genomic dataset limit the complexity of evolutionary models that can be applied.
+Single locus allows for more complex evolutionary models, and therefore might
+provide better evolutionary estimates \cite{}.
+
+Single locus super-matrix have provided very good phylogenegic inferences, sometimes even better than genomic based estimates. SL also, allow for construction of large phylogenies and
+
+Line 78- Add extra sentence about advantages of Physcraper's implementation
+Physcraper has speciifc advantages over existing pipelines.

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
links to studys and data packages as citations

@@
 \cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
-\href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool. Users can upload their
+(TNRS) tool \cite{TNRStool}. Users can upload their
 own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
-\href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have
+Phylesystem \cite{phylesystemGithub, mctavish2015phylesystem} that also have
 alignments available on TreeBASE \cite{piel2009treebase}.
@@
 "search taxon", a taxonomic group in the NCBI taxonomy that is automatically
-identified using the OpenTree
-\href{<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>}{API}
-\cite{rees2017automated}, as the most recent common ancestor of ingroup taxa that is
+identified using the OpenTree's taxonomic Most Recent Common Ancestor (MRCA) API
+\cite{mrcaAPI, rees2017automated}, as the MRCA of all ingroup taxa that is
 also a named clade in the NCBI taxonomy (Figure 1).
 
-BLAST is implemented with the `blastn` function \cite{camacho2009blast} and the BioPython
-\cite{cock2009biopython} BLAST function from
-\href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module}
+BLAST is performed using the `blastn` algorithm \cite{camacho2009blast}
+implemented in BioPython's \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
 modified to accept an alternative BLAST address.
@@
 selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
-\cite{gottlieb2005molecular} made original tree and alignment available in
-\href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE study 1091}.
-The tree sampling 41 species was also added to
-OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{study pg\_2827}
-and it has been integrated into OpenTree's synthetic tree
-\href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{here}.
-
+\cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
+(study 1091 \cite{treebase1091}). The tree sampling 41 species is also available from
+OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
+and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
 The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
-OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
+OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
 and in the DRYAD repository \cite{yao2020dryad}.
@@
 \textbf{OpenTree:} The Open Tree of Life project
+
+\textbf{TNRS:} Taxonomic Name Resolution Service
+
+\textbf{MRCA:} Most Recent Common Ancestor
+
+\textbf{BLAST:} Basic Local Alignment Search Tool
+
+\textbf{NCBI:} USA National Center for Biodiversity Information
 

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
some more math

@@
 analyses taking an additional 13hrs.
-%% <!-- MTH: you probably need some details about the hardware, given the fact hat you are discussing running times -->
+%% <!-- MTH: you probably need some details about the hardware, given the fact that you are discussing running times -->
 The updated \cite{gottlieb2005molecular} tree (Figure 2) displays
@@
 A large portion of internal branches are negligibly small, with 30 branches
-< 0.00001 substitution rate units, from which only 9 have a bootstrap support $>$ 0.75
+$<$ 0.00001 substitution rate units, from which only 9 have a bootstrap support $>$ 0.75
 (Figure 2).
@@
 and contributes phylogenetic data to 134 additional \textit{Ilex} taxa, from which
-`r 135-68` are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
+67 are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
 ITS as a marker, their GenBank data is not released yet, so Physcraper was unable

### 2021-01-07 2.1-intro-ideas.Rmd
discussion
intro again
intro and discussion

@@
 THEN TELL THEM WHAT YOU JUST TOLD THEM
+
+Connecting existing phylogenies with data such as
+geographical location, fossil ranges, and genetic and phenotypic information increasingly
+available in public databases constitute would be a great progress for evolutionary studies.
+
+However, if there are multiple alignments and trees, TreeBASE does not always indicate how they are connected.
+Hence, Phylesystem's metadata only provides a loose means of linking the tree with the exact alignment that generated it and a human curator must make the link.
 
@@
 
-Various efforts to automatize parts of the process in order to be able to take advantage of data in public databases.
+Various efforts to automatize parts of the process in order to be able to take
+advantage of data in public databases.
 

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
conclusion
intro and discussion

@@
 Phylogenies capture the shared history of organisms and provide key evolutionary
-context for our biological observations \cite{dobzhansky1973nothing}.
-Public biological databases constitute an amazing resource for evolutionary
-studies. Connecting existing phylogenies to molecular data that has never been
-incorporated into any phylogenetic estimate, geographical location, fossils, and
-other data in a reproducible and continuous manner is possible by establishing a
-data interoperability framework for biological databases. Here, we introduce
-Physcraper, a tool that automates biological database connections
+context for our biological observations \cite{dobzhansky1973nothing}. Data such as
+geographical location, fossil ranges, and genetic and phenotypic information increasingly
+available in public databases constitute an amazing
+resource for biological discovery \cite{baxevanis2015importance}.
+Connecting existing phylogenies with public biological data in a reproducible and
+continuous manner would represent a key tool for evolutionary studies.
+
+Here, we introduce Physcraper, a tool that automates biological database connections
 with the main goal of building upon homology
@@
 with public DNA data, with the added capability to connect the lineages represented
-in the phylogeny to biological data in public databases.
+in the phylogeny to biological data from public databases.
 
@@
 across databases. To address this challenge, the Open Tree of Life project (OpenTree)
-unifies taxonomic
-data from various databases \cite{rees2017automated}, including the USA National
+created a unified taxonomy for name standardization, by integrating taxonomic
+data from several databases \cite{rees2017automated}, including the USA National
 Center for Biodiversity Information (NCBI) molecular database GenBank
-\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
-that remains to be leveraged for connecting data from any biological database that
-has been integrated to it.
+\cite{benson2000genbank}, among others.
+Existing OpenTree taxonomy programmatic tools
+are a key resource that can be used to establish a framework for connecting data from any
+biological database that has been integrated to OpenTree's unified taxonomy.
 
-The Open Tree of Life project (OpenTree)
-also constructs a comprehensive tree of life by synthesizing published phylogenies
-and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
+OpenTree's unified taxonomy is used to construct OpenTree's comprehensive tree of
+life that synthesizes published phylogenies and taxonomy. It comprises 2.3 million tips, of which
 around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
-placed in the tree based on taxonomy. Meaning that there is still a lot of work
-to do to complete a full phylogenetic tree of all life.
+placed in the tree based on taxonomy.
+This large gap in public phylogenetic knowledge is due to several factors:
+- lack of data sharing, most analyses essentially lost forever;
+- lack of data, we still lack genetic data for a large portion of organisms; and
+- lack of analyticla power of existing data.
 
@@
 
-With more than XXXX million single locus
-sequences in GenBank today, there is a considerable amount of phylogenetically
+With more than 221 million single locus
+sequences in GenBank today \cite{genbankstats}, there is a considerable amount of phylogenetically
 informative data in GenBank that has not been analysed and/or made publicly available
-with the potential to give phylogenetic light to the tree of life.
-
-Phylogenies constructed from single locus super mayrix datasets are still widely
-generated and relevant to the field. Notable examples include the
-
-Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
-Magallon, Soltis, examples with fish, bats,
-\cite{magallon2015metacalibrated} mining by hand - flowering plant families - has been cited 561 times
-\cite{smith2018constructing} 38k clades of plants with seeds - cited 176 times
-\cite{jetz2012global} 6k bird species - cited
-\cite{rabosky2018inverse} mining with phlawd - cited 202 times
-
-These studies are widely cited and immensely relevant for the biological community.
-
-Some of these multilocus phylogenetic studies have implemented pipelines such as PHLAWD,
-other studies have curated their matrices by hand, which is largely non-reproducible.
-
-Make emphasis on these really great studies (and pipelines) that do not rely on
-genomic data, and that there is a lot of single locus data that can be analysed.
+with the potential to significantly fill phylogenetic gaps in the tree of life.
 
@@
 but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that mine DNA databases fast, efficiently, and reproducibly,
+Computational pipelines that mine DNA databases fast, efficiently, and reproducibly
 have been applied to infer phylogenetic relationships in a variety of organisms
 (e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
-However, fine-grained curated markers and alignments can improve phylogenetic
+While genomics has, and will continue to, revolutionize phylogenetic inference,
+the diversity of alternative genomic sequencing approaches implemented produce
+largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
+creating challenges for phylogenetic reconstruction.
+Phylogenomics addresses this problem by focusing on targeted capture of informative
+regions \cite{andermann2020guide}.
+However, fine-grained curated markers and alignments can significantly improve phylogenetic
 reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
 
-There are almost 8,200 publicly available, peer-reviewed alignments, covering
+There are almost 8,200 publicly available, peer-reviewed curated alignments, covering
 around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be used as seeds to mine molecular databases, and as "jump-start" alignments
+can be leveraged as seeds to mine molecular databases, and as "jump-start" alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 
-----------
+Such an approach was proposed in PUMPER but the community has not picked it up,
+probably due to installation difficulties.
+Physcraper extends this approach by allowing for interoperability.
 
-Physcraper is a Python pipeline using OpenTree's taxonomy and programmatic access
+Physcraper is implemented as a Python pipeline using OpenTree's taxonomy and programmatic access
 protocols (API's) to implement a database interoperability framework that automatically
@@
 New molecular data keep accumulating and tools facilitating its integration into
-existent evolutionary knowledge are needed.
+existent evolutionary knowledge contribute to the acceleration of scientific discovery.
 
-Phylogenetic pipelines designed to make evolutionary sense of the vast amount of
-public molecular data (e.g., Phylota \cite{sanderson2008phylota},
-PHLAWD \cite{smith2009mega},
-SUPERSMART \cite{antonelli2017toward}) focus on generating full phylogenies \textit{de novo},
+Physcraper is the first tool for interoperability in phylogenetic analyses (???)
+As a tool for automatization of phylogenetic reconstruction from molecular databases,
+we believe it presents several advantages over existing phylogenetic pipelines designed to
+make evolutionary sense of the vast amount of
+public molecular data. Phylota \cite{sanderson2008phylota},
+and PHLAWD \cite{smith2009mega} are the first ones and have been cited and used
+the most for actual phylogenetic reconstruction and mining of databases. These are
+strong tools that will probably still have a great potential in their recent updated versions
+(phylotar \cite{bennett2018phylotar} and PyPHLAWD \cite{smith2019pyphlawd}).
+They focus on generating full phylogenies \textit{de novo},
 i.e., inferring phylogenetic relationships from a newly generated homology hypothesis,
 as opposed to e.g., supertrees, that are generated by summarizing previous
-phylogenetic estimates.
-While Physcraper does not generate phylogenies \textit{de novo} in a traditional sense,
-it successfully generates new phylogenetic knowledge, revealing the importance of
-open science in facilitating phylogenetic placement of public molecular data and
-accelerating enrichment and updating of phylogenetic relationships in any region
-of the tree of life.
+phylogenetic estimates, which requires more manual adjustment later and difficults
+comparison with previous phylogenetic information.
+
+SUPERSMART \cite{antonelli2017toward}) and PUMPER \cite{izquierdo2014pumper}
+
 The PUMPER pipeline \cite{izquierdo2014pumper} also uses the concept of updating
@@
 
+\section*{Conclusions}
 Data repositories hold more information than meets the eye.
@@
 importance of data sharing and open science in the field, biology and science.
-
-\section*{Conclusions}
-Text for this section\ldots
 

### 2021-01-07 temp.Rmd
intro and discussion

@@
-While genomics has, and will continue to, revolutionize phylogenetic inference,
-the diversity of alternative genomic sequencing approaches implemented produce
-largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
-creating challenges for phylogenetic reconstruction.
-Phylogenomics addresses this problem by focusing on targeted capture of informative
-regions \cite{andermann2020guide}. Yet, it is expensive. Few labs around the world can afford
+While Physcraper does not generate phylogenies \textit{de novo} in a traditional sense,
+it successfully generates new phylogenetic knowledge, revealing the importance of
+open science in facilitating phylogenetic placement of public molecular data and
+accelerating enrichment and updating of phylogenetic relationships in any region
+of the tree of life.
+
+Genomics is expensive. Few labs around the world can afford
 to implement it for all their study groups, or at all. Creating a wealth gap for
@@
 %% single loci in the same database?
+
+%% While taxonomic trees are useful, trees inferred from evolutionary estimates are
+preferred.
+
+Notable examples of phylogenies constructed using multi-locus, database mined, super matrix approach:
+Magallon, Soltis, examples with fish, bats,
+\cite{magallon2015metacalibrated} mining by hand - flowering plant families - has been cited 561 times
+\cite{smith2018constructing} 38k clades of plants with seeds - cited 176 times
+\cite{jetz2012global} 6k bird species - cited
+\cite{rabosky2018inverse} mining with phlawd - cited 202 times
+
+These studies are widely cited and immensely relevant for the biological community.
+
+Some of these multilocus phylogenetic studies have implemented pipelines such as PHLAWD,
+other studies have curated their matrices by hand, which is largely non-reproducible.
+
+Make emphasis on these really great studies (and pipelines) that do not rely on
+genomic data, and that there is a lot of single locus data that can be analysed.
 

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
discussion and bib

@@
 
-Physcraper is the first tool for interoperability in phylogenetic analyses (???)
-As a tool for automatization of phylogenetic reconstruction from molecular databases,
+To our knowledge, Physcraper is the first tool establishing an interoperability
+framework for phylogenetic analyses (should we say this???).
+We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics,
+facilitating phylogenetic knowledge for ecologists and other non-specialists, effectively
+democratizing phylogenetic studies.
+
+As a tool for automatizing phylogenetic reconstruction from molecular databases,
 we believe it presents several advantages over existing phylogenetic pipelines designed to
 make evolutionary sense of the vast amount of
-public molecular data. Phylota \cite{sanderson2008phylota},
-and PHLAWD \cite{smith2009mega} are the first ones and have been cited and used
-the most for actual phylogenetic reconstruction and mining of databases. These are
-strong tools that will probably still have a great potential in their recent updated versions
-(phylotar \cite{bennett2018phylotar} and PyPHLAWD \cite{smith2019pyphlawd}).
-They focus on generating full phylogenies \textit{de novo},
-i.e., inferring phylogenetic relationships from a newly generated homology hypothesis,
-as opposed to e.g., supertrees, that are generated by summarizing previous
-phylogenetic estimates, which requires more manual adjustment later and difficults
-comparison with previous phylogenetic information.
+public molecular data available. Phylota \cite{sanderson2008phylota},
+and PHLAWD \cite{smith2009mega} were the first of the kind and have been cited and used
+abundantly for phylogenetic reconstruction and mining of molecular databases. While these are
+strong tools that will probably still be widely used with great potential in their
+more recent updated versions (phylotar \cite{bennett2018phylotar} and PyPHLAWD \cite{smith2019pyphlawd}),
+they do not take into account prior work and knowledge on phylogenetic relationships
+and alignment construction, which makes them prone to error (requiring more
+manual downstream processing) and difficults comparison with previous phylogenetic
+knowledge.
 
 SUPERSMART \cite{antonelli2017toward}) and PUMPER \cite{izquierdo2014pumper}
+share conceptual strengths, but present many challenges for installation and user
+experience, which have resulted in a low level of adoption of these tools
+in the community, and have been applied in very few phylogenetic analyses since their
+publication (put number of citations as of today or to harsh??).
 
-The PUMPER pipeline \cite{izquierdo2014pumper} also uses the concept of updating
-pre-existing alignments to incorporate public molecular data into phylogenies.
-Unfortunately, installation was unsuccessful following instructions from the author,
-and a comparison analysis between Physcraper and PUMPER is unfeasible at present.
-%% **MTH: add some more details about what aspect of the PUMPER install failed**
+Phylogenerator \cite{pearse2013phylogenerator} gathers genetic data by relying only on
+genetic marker names and descriptions in the GenBank database, which can be prone
+to error due to misslabelling. Any form of sequence similarity analysis, such as BLAST
+\cite{camacho2009blast} greatly improves correct homologous identification and should
+be preferred for obtaining high-quality phylogenetic hypotheses.
 
-Physcraper generates individual gene trees, failing to capture the complexity of
-species' evolutionary history \cite{song2012resolving}. Yet, Physcraper facilitates
+While Physcraper generates individual gene trees, which generally fail to capture the complexity of
+species' evolutionary history \cite{song2012resolving}, it facilitates
 gathering alignments and gene trees for multiple loci from a group of interest,
@@
 
-Physcraper can potentially link phylogenies to data available in any of the
+Physcraper has the unique advantage to link phylogenies to data available in any of the
 taxonomies integrated in the OpenTree taxonomy \cite{rees2017automated}, such as

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
update intro

--- 2021-01-07 sources/bmc_template_submission/bmc_article.tex
@@
 Phylogenies capture the shared history of organisms and provide key evolutionary
-context for our biological observations \cite{dobzhansky1973nothing}. Data such as
+context for our biological observations \cite{dobzhansky1973nothing}.
+Updating existing phylogenies with with molecular sequence data that has never been 
+incorporated into any phylogenetic estimate provides the opportunity to capture the evolutionary history of many taxa.
+Here, we introduce Physcraper, a tool that automates database connections
+to build on published alignments to extend phylogenetic inferences with more taxa.
+Physcraper can be used to to update a starting tree and single locus alignments
+with public DNA data, and link the tips in these trees to a unified, interoperable taxonomic resource \cite{rees2017automated}.
+
+Data such as
 geographical location, fossil ranges, and genetic and phenotypic information increasingly
-available in public databases constitute an amazing
+available in public databases which constitute an amazing
 resource for biological discovery \cite{baxevanis2015importance}.
-Connecting existing phylogenies with public biological data in a reproducible and
-continuous manner would represent a key tool for evolutionary studies.
-
-Here, we introduce Physcraper, a tool that automates biological database connections
-with the main goal of building upon homology
-hypotheses that taxon specialists have assessed and deemed appropriate for a
-specific phylogenetic scope to update a starting tree and single locus alignments
-with public DNA data, with the added capability to connect the lineages represented
-in the phylogeny to biological data from public databases.
-
 One of the main challenges for automatic
-integration of biological data into phylogenies are varying taxonomic idiosyncrasies
-across databases. To address this challenge, the Open Tree of Life project (OpenTree)
+integration of biological data across databases are varying taxonomic idiosyncrasies. 
+To address this challenge, the Open Tree of Life project (OpenTree)
 created a unified taxonomy for name standardization, by integrating taxonomic
 data from several databases \cite{rees2017automated}, including the USA National
-Center for Biodiversity Information (NCBI) molecular database GenBank
-\cite{benson2000genbank}, among others.
-Existing OpenTree taxonomy programmatic tools
-are a key resource that can be used to establish a framework for connecting data from any
-biological database that has been integrated to OpenTree's unified taxonomy.
-
-OpenTree's unified taxonomy is used to construct OpenTree's comprehensive tree of
-life that synthesizes published phylogenies and taxonomy. It comprises 2.3 million tips, of which
-around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
-placed in the tree based on taxonomy.
-This large gap in public phylogenetic knowledge is due to several factors:
-- lack of data sharing, most analyses essentially lost forever;
-- lack of data, we still lack genetic data for a large portion of organisms; and
-- lack of analyticla power of existing data.
+Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020}, and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019} among others.
+By using the existing OpenTree taxonomy programmatic tools to map tip names,
+Physcraper has a framework for connecting updated phylogenies data from any
+biological database.
 
@@
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-
-With more than 221 million single locus
-sequences in GenBank today \cite{genbankstats}, there is a considerable amount of phylogenetically
-informative data in GenBank that has not been analysed and/or made publicly available
-with the potential to significantly fill phylogenetic gaps in the tree of life.
+However, even thirteen years later, phylogenetic estimates for many of these taxa are not available \cite{mctavish_how_2017}.
+OpenTree's comprehensive tree of life comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
+There is a considerable amount of phylogenetically
+informative data in GenBank with the potential to fill these phylogenetic gaps in the tree of life that, but this data either has not been analysed or the analyses have not been made publicly available \cite{mctavish_how_2017}.
 
 Assembling a DNA alignment from such a massive database can be done "by hand",
-but it is a largely time consuming and mostly non-reproducible approach.
-Computational pipelines that mine DNA databases fast, efficiently, and reproducibly
+but that is a time consuming approach which is not highly reproducible.
+A variety of computational pipelines that mine DNA databases fast, efficiently, and reproducibly
 have been applied to infer phylogenetic relationships in a variety of organisms
@@
 
+Physcraper improves on previous work in automating phylogenetics by leveraging the power of existing homology hypotheses that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
 There are almost 8,200 publicly available, peer-reviewed curated alignments, covering
 around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be leveraged as seeds to mine molecular databases, and as "jump-start" alignments
+can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 
-Such an approach was proposed in PUMPER but the community has not picked it up,
-probably due to installation difficulties.
-Physcraper extends this approach by allowing for interoperability.
-
-Physcraper is implemented as a Python pipeline using OpenTree's taxonomy and programmatic access
-protocols (API's) to implement a database interoperability framework that automatically
-links phylogenies that have been standardized to OpenTree taxonomy, to alignments
-from TreeBASE, data from GenBank, and phylogenies from OpenTree's Phylesystem.
-Physcraper aims to demonstrate the benefits of reproducible workflows and open
-science in phylogenetics, and encourage better data sharing practices in the community.
+Physcraper is a Python pipeline using OpenTree's programmatic access
+protocols (API's) to automatically
+links phylogenies mapped to a standardized taxonomy \cite{mctavish2015phylesystem}, to alignments
+from TreeBASE \cite{sanderson1994treebase}, data from GenBank \cite{benson2000genbank}.
+Physcraper provides automated and reproducible workflows for updating existing phylogenetic estimates.
 

### 2021-01-07 temp.Rmd

@@
 of the tree of life.
+
+The PUMPER pipeline \cite{izquierdo2014pumper} also uses the concept of updating
+pre-existing alignments to incorporate public molecular data into phylogenies.
+Unfortunately, installation was unsuccessful following instructions from the author,
+and a comparison analysis between Physcraper and PUMPER is unfeasible at present.
+%% **MTH: add some more details about what aspect of the PUMPER install failed**
 

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
no tables

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
-\section*{Tables}
-\begin{table}[h!]
-\caption{Sample table title. This is where the description of the table should go}
-  \begin{tabular}{cccc}
-    \hline
-    & B1  &B2   & B3\\ \hline
-    A1 & 0.1 & 0.2 & 0.3\\
-    A2 & ... & ..  & .\\
-    A3 & ..  & .   & .\\ \hline
-  \end{tabular}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Additional Files}
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
-
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
-
 \end{backmatter}

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
caption
tighten up discussion

@@
 \section*{Discussion}
-
 Databases preserving and democratizing access to biological data
@@
 
-To our knowledge, Physcraper is the first tool establishing an interoperability
-framework for phylogenetic analyses (should we say this???).
 We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics,
@@
 As a tool for automatizing phylogenetic reconstruction from molecular databases,
-we believe it presents several advantages over existing phylogenetic pipelines designed to
+Physcraper presents several advantages over existing phylogenetic pipelines designed to
 make evolutionary sense of the vast amount of
-public molecular data available. Phylota \cite{sanderson2008phylota},
-and PHLAWD \cite{smith2009mega} were the first of the kind and have been cited and used
-abundantly for phylogenetic reconstruction and mining of molecular databases. While these are
-strong tools that will probably still be widely used with great potential in their
-more recent updated versions (phylotar \cite{bennett2018phylotar} and PyPHLAWD \cite{smith2019pyphlawd}),
-they do not take into account prior work and knowledge on phylogenetic relationships
-and alignment construction, which makes them prone to error (requiring more
-manual downstream processing) and difficults comparison with previous phylogenetic
+public molecular data available.
+
+Several analysis tools create full phylogenies \textit{de novo} by mining of molecular databases \cite{antonelli2017toward, sanderson2008phylota, smith2019pyphlawd, bennett2018phylotar, pearse2013phylogenerator}.
+In particular, Phylota \cite{sanderson2008phylota},
+and PHLAWD \cite{smith2009mega}, have been cited and used abundantly.
+
+Physcraper builds on this automated database mining concept by incorporating prior phylogenetic work and existing taxonomic domain knowledge on appropriate markers and alignment construction.
+This decreases error (requiring more manual downstream processing) and eases comparison with previous phylogenetic
 knowledge. 
+Unlike placement approaches \cite{berger_performance_2011, matsen_pplacer_2010}, which add new taxa without modifying the input tree, Physcraper estimates all the relationships in the context of the new data.
+PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no longer under active development, is challenging to install and run, and has resulted in very few phylogenetic analyses since its publication.
 
-SUPERSMART \cite{antonelli2017toward}) and PUMPER \cite{izquierdo2014pumper}
-share conceptual strengths, but present many challenges for installation and user
-experience, which have resulted in a low level of adoption of these tools
-in the community, and have been applied in very few phylogenetic analyses since their
-publication (put number of citations as of today or to harsh??).
+Physcraper generates gene trees, which individually do not capture the full complexity of species' evolutionary history \cite{song2012resolving}.
+However, physcraper facilitates
+gathering alignments and gene trees for multiple loci from a group of interest,
+that together can be used to reconstruct species trees taking into account coalescent processes with ASTRAL \cite{mirarab2014astral},
+BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
+Rigorous analyses of multiple loci allows for more complex evolutionary models than analyses of large genomic data sets, and can provide better evolutionary estimates.
 
-Phylogenerator \cite{pearse2013phylogenerator} gathers genetic data by relying only on
-genetic marker names and descriptions in the GenBank database, which can be prone
-to error due to misslabelling. Any form of sequence similarity analysis, such as BLAST
-\cite{camacho2009blast} greatly improves correct homologous identification and should
-be preferred for obtaining high-quality phylogenetic hypotheses.
+Physcraper has the added advantage of linking taxonomic information about tips in the output phylogenies to data available in a variety of biological databases \cite{rees2017automated}, such as
+geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
 
-While Physcraper generates individual gene trees, which generally fail to capture the complexity of
-species' evolutionary history \cite{song2012resolving}, it facilitates
-gathering alignments and gene trees for multiple loci from a group of interest,
-that can be used to reconstruct species trees with ASTRAL \cite{mirarab2014astral},
-BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
-
-Physcraper has the unique advantage to link phylogenies to data available in any of the
-taxonomies integrated in the OpenTree taxonomy \cite{rees2017automated}, such as
-geographical locations from the Global Biodiversity Information Facility, or
-fossils from the Paleobiology Database.
-%% a resource that is contributing to achieve the goal of "exploring and
-%% analysing biodiversity at an accelerated pace, and returning systematics into
-%% the mainstream of science" \cite{wilson2003encyclopedia}.
-The Physcraper workflow can be used to rapidly (in a matter of hours)
-address challenges overarching both fields of ecology and evolution, such as
+The Physcraper workflow can be used to rapidly (in a matter of hours) to create phylogenies which can address challenges overarching both fields of ecology and evolution, such as
 phylogenetically placing newly discovered species \cite{webb2010biodiversity},
-systematizing molecular (and other) databases, i.e., curating taxonomic assignations
+curating taxonomic assignments
 \cite{san2010molecular},
@@
 \textbf{NCBI:} USA National Center for Biodiversity Information
+
+\textbf{GBIF:} Global Biodiversity Information Facility
 
@@
   \begin{figure}[h!]
-  \caption{The Physcraper framework consists of 4 general steps (see text). The software is fully
+  \caption{The Physcraper framework consists of four general steps. The software is fully
   described on its documentation website at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io},

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
intro small edits

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-Updating existing phylogenies with with molecular sequence data that has never been 
-incorporated into any phylogenetic estimate provides the opportunity to capture the evolutionary history of many taxa.
+Updating existing phylogenies with molecular sequence data that has never been
+incorporated into any phylogenetic estimate provides the opportunity to capture
+the evolutionary history of many taxa.
 Here, we introduce Physcraper, a tool that automates database connections
 to build on published alignments to extend phylogenetic inferences with more taxa.
-Physcraper can be used to to update a starting tree and single locus alignments
-with public DNA data, and link the tips in these trees to a unified, interoperable taxonomic resource \cite{rees2017automated}.
+Physcraper can be used to update a starting tree and single locus alignments
+with public DNA data, and link the tips in these trees to a unified, interoperable
+taxonomic resource \cite{rees2017automated}.
 
@@
 geographical location, fossil ranges, and genetic and phenotypic information increasingly
-available in public databases which constitute an amazing
+available in public databases constitute an amazing
 resource for biological discovery \cite{baxevanis2015importance}.
@@
 data from several databases \cite{rees2017automated}, including the USA National
-Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020}, and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019} among others.
+Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
+and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019} among others.
 By using the existing OpenTree taxonomy programmatic tools to map tip names,
@@
 There is a considerable amount of phylogenetically
-informative data in GenBank with the potential to fill these phylogenetic gaps in the tree of life that, but this data either has not been analysed or the analyses have not been made publicly available \cite{mctavish_how_2017}.
+informative data in GenBank with the potential to fill these phylogenetic gaps in
+the tree of life that, but this data either has not been analysed or the analyses
+have not been made publicly available \cite{mctavish_how_2017}.
 
@@
 A variety of computational pipelines that mine DNA databases fast, efficiently, and reproducibly
-have been applied to infer phylogenetic relationships in a variety of organisms
+have been developed and used to infer phylogenetic relationships in a variety of organisms
 (e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
@@
 
-Physcraper improves on previous work in automating phylogenetics by leveraging the power of existing homology hypotheses that taxon specialists have assessed and deemed appropriate for a specific phylogenetic scope.
+Physcraper improves on previous work in automating phylogenetics by leveraging
+the power of existing homology hypotheses that taxon specialists have assessed
+and deemed appropriate for a specific phylogenetic scope.
 There are almost 8,200 publicly available, peer-reviewed curated alignments, covering
@@
 protocols (API's) to automatically
-links phylogenies mapped to a standardized taxonomy \cite{mctavish2015phylesystem}, to alignments
+link phylogenies mapped to a standardized taxonomy \cite{mctavish2015phylesystem}, to alignments
 from TreeBASE \cite{sanderson1994treebase}, data from GenBank \cite{benson2000genbank}.

### 2021-01-08 sources/bmc_template_submission/bmc_article.tex
last edits

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-Updating existing phylogenies with molecular sequence data that has never been
+Updating existing phylogenies with publicly available molecular sequence data that
+has never been
 incorporated into any phylogenetic estimate provides the opportunity to capture
@@
 link phylogenies mapped to a standardized taxonomy \cite{mctavish2015phylesystem}, to alignments
-from TreeBASE \cite{sanderson1994treebase}, data from GenBank \cite{benson2000genbank}.
+from TreeBASE \cite{sanderson1994treebase}, and data from GenBank \cite{benson2000genbank}.
 Physcraper provides automated and reproducible workflows for updating existing phylogenetic estimates.
@@
 
-Several analysis tools create full phylogenies \textit{de novo} by mining of molecular databases \cite{antonelli2017toward, sanderson2008phylota, smith2019pyphlawd, bennett2018phylotar, pearse2013phylogenerator}.
+Several analysis tools create full phylogenies \textit{de novo} by mining of molecular
+databases \cite{antonelli2017toward, sanderson2008phylota, smith2019pyphlawd,
+bennett2018phylotar, pearse2013phylogenerator}.
 In particular, Phylota \cite{sanderson2008phylota},
@@
 
-Physcraper builds on this automated database mining concept by incorporating prior phylogenetic work and existing taxonomic domain knowledge on appropriate markers and alignment construction.
-This decreases error (requiring more manual downstream processing) and eases comparison with previous phylogenetic
+Physcraper builds on this automated database mining concept by incorporating prior
+phylogenetic work and existing taxonomic domain knowledge on appropriate markers
+and alignment construction.
+This decreases error (requiring less manual downstream processing) and
+eases comparison with previous phylogenetic
 knowledge.
-Unlike placement approaches \cite{berger_performance_2011, matsen_pplacer_2010}, which add new taxa without modifying the input tree, Physcraper estimates all the relationships in the context of the new data.
-PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no longer under active development, is challenging to install and run, and has resulted in very few phylogenetic analyses since its publication.
+Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
+which add new taxa without modifying the input tree, Physcraper estimates all
+the relationships anew in the context of the new data.
+PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no
+longer under active development, is challenging to install and run, and has resulted
+in very few phylogenetic analyses since its publication.
 
-Physcraper generates gene trees, which individually do not capture the full complexity of species' evolutionary history \cite{song2012resolving}.
-However, physcraper facilitates
+Physcraper generates gene trees, which individually do not capture the full complexity
+of species' evolutionary history \cite{song2012resolving}.
+However, Physcraper facilitates
 gathering alignments and gene trees for multiple loci from a group of interest,
-that together can be used to reconstruct species trees taking into account coalescent processes with ASTRAL \cite{mirarab2014astral},
+that together can be used to reconstruct species trees taking into account coalescent
+processes with ASTRAL \cite{mirarab2014astral},
 BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
-Rigorous analyses of multiple loci allows for more complex evolutionary models than analyses of large genomic data sets, and can provide better evolutionary estimates.
+Rigorous analyses of multiple loci allows for more complex evolutionary models
+than analyses of large genomic data sets, and can provide better evolutionary estimates.
 
-Physcraper has the added advantage of linking taxonomic information about tips in the output phylogenies to data available in a variety of biological databases \cite{rees2017automated}, such as
+Physcraper has the added advantage of linking taxonomic information about tips
+in the output phylogenies to data available in a variety of biological databases
+\cite{rees2017automated}, such as
 geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
 
-The Physcraper workflow can be used to rapidly (in a matter of hours) to create phylogenies which can address challenges overarching both fields of ecology and evolution, such as
+The Physcraper workflow can be used to rapidly (in a matter of hours) create 
+phylogenies which can address challenges overarching both fields of ecology and evolution, such as
 phylogenetically placing newly discovered species \cite{webb2010biodiversity},

### 2021-01-08 sources/bmc_template_submission/bmc_article.tex
first submission final

@@
 has never been
-incorporated into any phylogenetic estimate provides the opportunity to capture
-the evolutionary history of many taxa.
-Here, we introduce Physcraper, a tool that automates database connections
-to build on published alignments to extend phylogenetic inferences with more taxa.
-Physcraper can be used to update a starting tree and single locus alignments
-with public DNA data, and link the tips in these trees to a unified, interoperable
+incorporated into any phylogenetic estimate provides the opportunity to study
+the evolutionary history of many taxa in a reproducible and continuous manner.
+Here, we introduce Physcraper, a tool that establishes a data interoperability
+framework for biological databases to automate data connections across databases,
+with the main goal of building on published alignments and extending existing
+phylogenetic inferences with more data and taxa.
+Physcraper updates a starting tree and single locus alignments
+with public DNA data, and links the tips in these trees to a unified, interoperable
 taxonomic resource \cite{rees2017automated}.
@@
 By using the existing OpenTree taxonomy programmatic tools to map tip names,
-Physcraper has a framework for connecting updated phylogenies data from any
+Physcraper has a framework for connecting updated phylogenies with data from any
 biological database.
@@
 However, even thirteen years later, phylogenetic estimates for many of these taxa are not available \cite{mctavish_how_2017}.
-OpenTree's comprehensive tree of life comprises 2.3 million tips, of which around 90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed in the tree based on taxonomy.
+OpenTree's comprehensive tree of life comprises 2.3 million tips, of which around
+90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed
+in the tree based on taxonomy.
 There is a considerable amount of phylogenetically
 informative data in GenBank with the potential to fill these phylogenetic gaps in
-the tree of life that, but this data either has not been analysed or the analyses
+the tree of life, but this data either has not been analysed or the analyses
 have not been made publicly available \cite{mctavish_how_2017}.
 
-Assembling a DNA alignment from such a massive database can be done "by hand",
+Assembling a DNA alignment from such a massive database as GenBank can be done "by hand",
 but that is a time consuming approach which is not highly reproducible.
@@
 have been developed and used to infer phylogenetic relationships in a variety of organisms
-(e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper}).
+(e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper, pearse2013phylogenerator}).
 While genomics has, and will continue to, revolutionize phylogenetic inference,
@@
 
-Physcraper improves on previous work in automating phylogenetics by leveraging
+Physcraper improves on previous work in automating phylogenetic reconstruction by leveraging
 the power of existing homology hypotheses that taxon specialists have assessed
@@
 
-Physcraper is a Python pipeline using OpenTree's programmatic access
+Physcraper is implemented as a Python pipeline that uses OpenTree's programmatic access
 protocols (API's) to automatically
-link phylogenies mapped to a standardized taxonomy \cite{mctavish2015phylesystem}, to alignments
+link any phylogeny mapped to OpenTree's standardized taxonomy \cite{mctavish2015phylesystem},
+to alignments
 from TreeBASE \cite{sanderson1994treebase}, and data from GenBank \cite{benson2000genbank}.
-Physcraper provides automated and reproducible workflows for updating existing phylogenetic estimates.
+Its utility and functionalities are presented with a case-study analysis of a
+group of flowering plants, the hollies.
 
@@
 
-The general Physcraper framework (Figure 1) consists of 4 steps: 1)identifying and
+The general Physcraper framework (Figure 1) consists of 4 steps: 1) identifying and
 processing a tree and its underlying alignment; 2) performing a BLAST search of
@@
 
-Physcraper has the added advantage of linking taxonomic information about tips
+Physcraper has the added advantage of facilitating the linkage of taxonomic information about tips
 in the output phylogenies to data available in a variety of biological databases

### 2021-03-03 sources/bmc_template_submission/bmc_article.tex
adding verificaion test

@@
 framework for biological databases to automate data connections across databases,
-with the main goal of building on published alignments and extending existing
+with the main goal of building upon published alignments to extend existing
 phylogenetic inferences with more data and taxa.
@@
 \cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
-(study 1091 \cite{treebase1091}). The tree sampling 41 species is also available from
+(study 1091 \cite{treebase1091}). The tree samples 41 species, is available from
 OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
@@
 and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
+With 175 tips, the tree from \cite{yao2020phylogeny} is the best sampled phylogeny yet available for
 the hollies.
@@
 to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
+
+\subsection*{Verification test}
+
+We pruned 9 out of the 41 original tips of the Gottlieb et al. tree, corresponding
+to a 20\% trim, excluding the outgroups. We then performed a Physcraper run to test
+if we would recover the dropped tips.
+We successfully recovered 6 out of 9 pruned tips in the updated tree.
+Closer examination of results revealed that the 3 missing tips were correctly retrieved
+with BLAST along with the other 6 tips that were included in the updated tree. Examining
+the ITS sequences belonging to the tips missed by the test, by following the GenBank
+accession numbers reported in the original publication, showed that these sequences
+contain a 100 bp long gap of unidentified nucleotides (Ns). These Ns were absent
+in the 3 sequences published in the original alignment.
 

### 2021-03-04 sources/bmc_template_submission/bmc_article.tex
adding to results and discussion

@@
 \cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
-(study 1091 \cite{treebase1091}). The tree samples 41 species, is available from
+(study 1091 \cite{treebase1091}). This tree (Gottlieb2005 tree from now on) samples 41 species, is available from
 OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
 and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
-The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
+The most recent \textit{Ilex} tree \cite{yao2020phylogeny} (Yao2020 tree from now on) is available in
 OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
 and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the tree from \cite{yao2020phylogeny} is the best sampled phylogeny yet available for
-the hollies.
+With 175 tips, the Yao2020 tree is the best sampled phylogeny yet available for
+the genus \textiy{Ilex}.
 
@@
 %% <!-- MTH: you probably need some details about the hardware, given the fact that you are discussing running times -->
-The updated \cite{gottlieb2005molecular} tree (Figure 2) displays
+The updated Gottlieb2005 tree (Figure 2) displays
 all 41 distinct
@@
 (Figure 2).
-For comparison, \cite{yao2020phylogeny} also contains all 41 distinct taxa from
-the original
-\cite{gottlieb2005molecular} study,
-and contributes phylogenetic data to 134 additional \textit{Ilex} taxa, from which
-67 are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
-ITS as a marker, their GenBank data is not released yet, so Physcraper was unable
-to incorporate 68 additional taxa into the analysis. However, Physcraper was able
-to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
+The Yao2020 tree \cite{yao2020phylogeny} also contains all 41 distinct taxa
+sampled in the Gottlieb2005 tree \cite{gottlieb2005molecular},
+while contributing phylogenetic data to 134 additional \textit{Ilex} taxa. From these,
+67 taxa are also in the Physcraper updated Gottlieb2005 tree \cite{gottlieb2005molecular}.
+While the Yao2020 tree \cite{yao2020phylogeny} was also constructed using
+ITS as a marker, their GenBank data is not released yet. Hence, Physcraper was unable
+to incorporate 68 taxa that are only on the Yao2020 tree because the DNA data is unavailable.
+We also note that Physcraper incorporates 18 \textit{Ilex} taxa that are not in the Yao2020
+tree \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex} species
+and visual inspection of the DNA sequences suggests they are correctly assigned as \textit{Ilex}.
+The ITS alignment that underlies the Yao2020 tree was constructed without any tool to
+scrape GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
+incorporate these 18 additional \textit{Ilex} taxa.
 
@@
 existent evolutionary knowledge contribute to the acceleration of scientific discovery.
+
+Physcraper is a tool that builds upon previous knowledge stored in published
+alignments, with the main goal to extend our knowledge of phylogenetic relationships
+across the tree of life.
 
@@
 knowledge.
+
+Results from the verification test highlight the importance of incorporating existing expertly
+curated homology statements to automatically update phylogenetic relationships, instead
+of ignoring the information they contain and building homology statements fully \textit{de novo}.
+The three sequences that were in the original tree but were not included in
+the updated tree by Physcraper in the verification test were excluded because
+they contain a 100 bp long gap of unidentified nucleotides (Ns) in GenBank that
+was manually curated and trimmed in the original alignment.
+
 Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},

### 2021-03-05 sources/bmc_template_submission/bmc_article.tex
more results

@@
 and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the Yao2020 tree is the best sampled phylogeny yet available for
-the genus \textiy{Ilex}.
+With 175 tips, the Yao2020 tree \cite{yao2020phylogeny} is the best sampled phylogeny available for
+the genus \textit{Ilex}.
 
 We ran Physcraper on a laptop Linux computer to update an internal transcribed
-spacer DNA region (ITS) alignment from \cite{gottlieb2005molecular}, using a local
-GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
+spacer DNA region (ITS) alignment that was used to construct the tree from \cite{gottlieb2005molecular},
+using a local GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
 analyses taking an additional 13hrs.
 %% <!-- MTH: you probably need some details about the hardware, given the fact that you are discussing running times -->
-The updated Gottlieb2005 tree (Figure 2) displays
-all 41 distinct
+The Gottlieb2005 tree \cite{gottlieb2005molecular} updated using Physcraper (Figure 2;
+Physcraper updated tree from now on) displays all 41 distinct
 taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional \textit{Ilex} taxa. The best RaxML tree is 99\% resolved, with 25\% of nodes
-with bootstrap support $<$ 0.1 and 48\% nodes with bootstrap support $>$ 0.75.
+84 additional \textit{Ilex} taxa. The best ML tree from the RAxML analysis is 99\%
+resolved, with 25\% of nodes with bootstrap support $<$ 0.1 and 48\% nodes with
+bootstrap support $>$ 0.75.
 A large portion of internal branches are negligibly small, with 30 branches
@@
 (Figure 2).
-The Yao2020 tree \cite{yao2020phylogeny} also contains all 41 distinct taxa
+As comparison with the Physraper updated tree, the Yao2020 tree \cite{yao2020phylogeny}
+also contains all 41 distinct taxa
 sampled in the Gottlieb2005 tree \cite{gottlieb2005molecular},
 while contributing phylogenetic data to 134 additional \textit{Ilex} taxa. From these,
-67 taxa are also in the Physcraper updated Gottlieb2005 tree \cite{gottlieb2005molecular}.
+67 taxa are also in the Physcraper updated tree.
 While the Yao2020 tree \cite{yao2020phylogeny} was also constructed using
@@
 scrape GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
-incorporate these 18 additional \textit{Ilex} taxa.
+incorporate these 18 additional \textit{Ilex} taxa in the updated tree (Figure 2).
 

### 2021-03-30 sources/bmc_template_submission/bmc_article.tex
pylint implementation
pylint scores

@@
 \section*{Implementation}
+
+Physcraper is implemented with Python and can also be run on the command line.
+It currently consists of 12 modules.
+%% , classes, methods and functions.%%
+For testing and improving the quality of the Python code syntax in the Phsycraper code,
+we used the Pylint \cite{pylintWeb, pylintManual} software with a configuration file.
+We improved the code syntax on modules with low Pylint scores and fixed errors by
+following Pylint's recommendations.
+In some cases, we overruled Pylint's recommendations by using its check-disabling
+system. As of now, all modules have a Pylint score of 10/10.
 

### 2021-03-30 sources/bmc_template_submission/bmc_article.tex
reviewer 1

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-Updating existing phylogenies with publicly available molecular sequence data that
-has never been
-incorporated into any phylogenetic estimate provides the opportunity to study
+Updating existing phylogenies with publicly available molecular sequence data provides the opportunity to study
 the evolutionary history of many taxa in a reproducible and continuous manner.
@@
 Updated and original tree are compared with Robinson-Foulds weighted and unweighted
-metrics estimated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
+metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
 comparison between the synthetic OpenTree and original and updated tree individually,
@@
 (Figure 2).
-As comparison with the Physraper updated tree, the Yao2020 tree \cite{yao2020phylogeny}
+As comparison with the Physcraper updated tree, the Yao2020 tree \cite{yao2020phylogeny}
 also contains all 41 distinct taxa
@@
 incorporate these 18 additional \textit{Ilex} taxa in the updated tree (Figure 2).
+
+%% Maybe show conflict results? and then discuss in the discussion.
 
@@
 they contain a 100 bp long gap of unidentified nucleotides (Ns) in GenBank that
-was manually curated and trimmed in the original alignment.
+was manually curated and removed from in the original alignment.
+
+These sequences do appear in the Physcraper results in the list of matches from GenBank which did not fit the sequence length cutoffs set in the configuration. 
+This ``seqlen\_mismatch.txt'' file includes the accession number, taxon, and sequence length of all sequences culled based on sequence length.
+
+We encourage users to look at this file, and potentially modify their sequence length cutoff parameters based on the excluded sequences.
+The default relative sequence length thresholds are arbitrary, but we hope that by making the process of locating homologous sequences online reproducible, and tracking what cutoffs are used, we make it easier for researchers to delve into the effect that different choices have on their inferences.
+This is in contrast to ‘manual’ searches for taxa, where similarly arbitrary cutoffs are applied, but are difficult to trace.
+As many studies have shown \cite{huang_what_2009} the effect of missing data can be enigmatic, and interact with the true phylogenetic relationships for the data set at hand. 
+There is not currently strong support in the literature for any cutoffs, and rather than prescribe specific approaches, we encourage users to explore the effects of different choices on inference.
+
+Organellar genome sequences, such as chloroplasts and mitochondria will also generally be excluded from automatic addition based on length cutoffs.
+Multiple sequence alignment of loci of drastically different lengths is unfeasible, and we have found in testing often returns incorrect results, splitting shorter sequences with many long gaps to align with exact matches across the entire longer locus.
+While it would be possible to directly extract the blast match from genomes, this match would exclude potentially homologous flanking regions which are not matched by blasts local search algorithm, but may be important for phylogenetic inference.
+Instead we list the accession numbers for these matches in the seqlen\_mismatch file, for users to assess appropriate homologous regions to their alignment of interest.
+
+Single gene phylogenies with very high numbers of taxa may lack sufficient signal for accurate resolution \cite{morel_phylogenetic_2020}.
+The Physcraper workflow avoids some of these challenges, by focussing on ingroup taxa of an exisiting phylogeny, using markers that have been assessed and proven apporpraite for that phylogenetic scope in past publictions.
+In addition, we thin alignments by removing identical sequences, and by setting a maximum number of sequences per taxon.
+Nonetheless, it is encumbent on users to assess their final inference with respect to statistical support and biological plausibility.
+
+Most of the default thresholds parameter values and software choices in the Physcraper pipeline are arbitrary. 
+We hope that by making the process of locating homologous sequences online reproducible, 
+and tracking what cutoffs are used, we make it easier for researchers to delve into the effect that different choices have on their inferences.
+In addition, by providing the output files at each step of the analysis, 
+it is straightforward to assess how changing parameter and software choices do or do not drive differences in phylogenetic inference.
+By gathering the sequences, and making the unaligned files easily available to users, researchers can compare if applying any alternate alignment tool of their choice affects inferences. 
+Once sequences are aligned, they can apply and compare inferences from any phylogenetics software.
 
@@
 that together can be used to reconstruct species trees taking into account coalescent
-processes with ASTRAL \cite{mirarab2014astral},
-BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
+processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}).
 Rigorous analyses of multiple loci allows for more complex evolutionary models
 than analyses of large genomic data sets, and can provide better evolutionary estimates.
+For example \cite{zhu_complexity_2021} show that when applying coalescent models, there is more information in two genes of 300 bp each than in 600 independent sites 
+We have included a script (``multi_locus.py'') and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
 
@@
 geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
+These taxonomic links, and comparisons to existing published phylogenies in the OpenTree data store can also help flag paralogous sequences.
+Accidentally including paralogs as homologs is a risk in phylogenetic analyses, and can be more prevalent in automated analyses than in manually curated analyses.
+We provide users with several tools to try to assess homology of their aligned sequences. 
+The estimated gene tree itself is an evolutionarily explicit way to visualize gene evolution, which in concert with taxonomic labelling can reveal paralogy.
+Our conflict analysis tool informs the users of whether their tree contains major conflicts with established taxonomy. 
+This conflict tool also returns information on if these taxonomic conflicts existed in the original input tree. 
+These conflicts may be a sign that taxonomy needs to be updated, or may be a sign that non-homologous sequences have been included in the analysis. 
+These taxonomic and phylogenetic conflicts flag regions for the tree for the researcher to more closely examine and assess homology.
+
+%% Biological plausibility of example??
 

### 2021-03-31 sources/bmc_template_submission/bmc_article.tex
some edits

@@
 These sequences do appear in the Physcraper results in the list of matches from GenBank which did not fit the sequence length cutoffs set in the configuration.
-This ``seqlen\_mismatch.txt'' file includes the accession number, taxon, and sequence length of all sequences culled based on sequence length.
+This "seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence length of all sequences filtered based on sequence length.
 
@@
 Organellar genome sequences, such as chloroplasts and mitochondria will also generally be excluded from automatic addition based on length cutoffs.
-Multiple sequence alignment of loci of drastically different lengths is unfeasible, and we have found in testing often returns incorrect results, splitting shorter sequences with many long gaps to align with exact matches across the entire longer locus.
+Multiple sequence alignment of loci of drastically different lengths is unfeasible, and we have found in testing that it often returns incorrect results, splitting shorter sequences with many long gaps to align with exact matches across the entire longer locus.
 While it would be possible to directly extract the blast match from genomes, this match would exclude potentially homologous flanking regions which are not matched by blasts local search algorithm, but may be important for phylogenetic inference.
-Instead we list the accession numbers for these matches in the seqlen\_mismatch file, for users to assess appropriate homologous regions to their alignment of interest.
+Instead we list the accession numbers for these matches in the "seqlen\_mismatch.txt" file, for users to assess appropriate homologous regions for their alignment of interest.
 
 Single gene phylogenies with very high numbers of taxa may lack sufficient signal for accurate resolution \cite{morel_phylogenetic_2020}.
-The Physcraper workflow avoids some of these challenges, by focussing on ingroup taxa of an exisiting phylogeny, using markers that have been assessed and proven apporpraite for that phylogenetic scope in past publictions.
-In addition, we thin alignments by removing identical sequences, and by setting a maximum number of sequences per taxon.
-Nonetheless, it is encumbent on users to assess their final inference with respect to statistical support and biological plausibility.
+The Physcraper workflow avoids some of these challenges, by focussing on ingroup taxa of an existing phylogeny, using markers that have been assessed and proven apporpraite for that phylogenetic scope in past publictions.
+In addition, Physcraper thins alignments by removing identical sequences, and by setting a maximum number of sequences per taxon.
+Nonetheless, it is incumbent on users to assess their final inference with respect to statistical support and biological plausibility.
 
@@
 By gathering the sequences, and making the unaligned files easily available to users, researchers can compare if applying any alternate alignment tool of their choice affects inferences.
-Once sequences are aligned, they can apply and compare inferences from any phylogenetics software.
+Once sequences are aligned, they can apply and compare inferences from any Phylogenetic software.
 
@@
 For example \cite{zhu_complexity_2021} show that when applying coalescent models, there is more information in two genes of 300 bp each than in 600 independent sites
-We have included a script (``multi_locus.py'') and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
+We have included a script ("multi_locus.py") and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
 
@@
 geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
-These taxonomic links, and comparisons to existing published phylogenies in the OpenTree data store can also help flag paralogous sequences.
+Taxonomic links, and comparisons to existing published phylogenies in the OpenTree data store can also help flag paralogous sequences.
 Accidentally including paralogs as homologs is a risk in phylogenetic analyses, and can be more prevalent in automated analyses than in manually curated analyses.
@@
 Initial ideas about the data are constantly changed by results from new analyses.
-Physcraper provides a framework for reproducible phylogenetics that has the
+Physcraper provides a framework for reproducible Phylogenetics that has the
 potential to consistently provide context for these ideas, highlighting the

### 2021-03-31 sources/bmc_template_submission/bmc_article.tex
intro increased taxon sampling

@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-Updating existing phylogenies with publicly available molecular sequence data provides the opportunity to study
+Updating existing phylogenies with publicly available molecular sequence data provides
+the opportunity to study
 the evolutionary history of many taxa in a reproducible and continuous manner.
 Here, we introduce Physcraper, a tool that establishes a data interoperability
-framework for biological databases to automate data connections across databases,
-with the main goal of building upon published alignments to extend existing
-phylogenetic inferences with more data and taxa.
+framework for biological databases to automate data connections across databases.
+Physcraper's main goal is to build upon published alignments to extend existing
+phylogenetic inferences with more data and taxa, which have been proved to improve
+phylogenetic reconstruction \cite{hillis1996inferring, natsidis2019phylogenomics},
+time of divergence estimates \cite{schulte2013undersampling, soares2015influence} and
+biogeographic analyses \cite{kayaalp2017back},
+as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
 Physcraper updates a starting tree and single locus alignments
@@
 This decreases error (requiring less manual downstream processing) and
-eases comparison with previous phylogenetic
-knowledge.
+eases comparison with previous phylogenetic knowledge.
 

### 2021-04-01 sources/bmc_template_submission/bmc_article.tex
escape underscore

@@
 than analyses of large genomic data sets, and can provide better evolutionary estimates.
+
 For example \cite{zhu_complexity_2021} show that when applying coalescent models, there is more information in two genes of 300 bp each than in 600 independent sites 
-We have included a script ("multi_locus.py") and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
+We have included a script (``multi\_locus.py'') and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
 

### 2021-04-02 sources/bmc_template_submission/bmc_article.tex
some edits

@@
 Physcraper's main goal is to build upon published alignments to extend existing
-phylogenetic inferences with more data and taxa, which have been proved to improve
+phylogenetic inferences with more data and taxa, which improve
 phylogenetic reconstruction \cite{hillis1996inferring, natsidis2019phylogenomics},
-time of divergence estimates \cite{schulte2013undersampling, soares2015influence} and
+time of divergence estimates \cite{schulte2013undersampling, soares2015influence},
 biogeographic analyses \cite{kayaalp2017back},
 as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
-Physcraper updates a starting tree and single locus alignments
-with public DNA data, and links the tips in these trees to a unified, interoperable
+Physcraper updates a starting single locus alignment and corresponding tree
+with public DNA data, and links the tips in the updated trees to a unified, interoperable
 taxonomic resource \cite{rees2017automated}.
@@
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-However, even thirteen years later, phylogenetic estimates for many of these taxa are not available \cite{mctavish_how_2017}.
-OpenTree's comprehensive tree of life comprises 2.3 million tips, of which around
-90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed
+However, even thirteen years later, phylogenetic estimates for most of these taxa
+are still not available \cite{mctavish_how_2017}.
+The Open Tree of Life project assembles a comprehensive synthetic tree of life comprising
+2.3 million tips, of which around only
+90,000 are supported by phylogenetic data - the remaining 1.4 million taxa are placed
 in the tree based on taxonomy.
@@
 around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
+can be leveraged as seeds to mine molecular databases, and as "jump-start" alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 following Pylint's recommendations.
-In some cases, we overruled Pylint's recommendations by using its check-disabling
-system. As of now, all modules have a Pylint score of 10/10.
+Based on Physcraper's software design choices, some of Pylint's recommendations were overruled
+by using its check-disabling system, and are explained in the code.
+As of now, all modules have a Pylint score of 10/10.
 

### 2021-04-06 sources/bmc_template_submission/bmc_article.tex
answering soem more questions

@@
 %% Biological plausibility of example??
+The updated holly tree is based on a single marker, so we expected for it to be
+not as well resolved as trees using multiple markers. Nevertheless, all samples
+corresponding to the ingroup cluster together in a monophyletic group. We also
+note that most samples belonging to the same species also cluster together forming
+monophyletic groups. Notable exceptions are Ilex theeizans, which also appears as
+non-monophyletic in the original Gottlieb et al. 2005 tree. A visual comparison
+of the Yao et al. 2020 tree and the original Gottlieb et al. 2005 tree suggests
+that the relationships within the genus Ilex are still being determined, and that
+increased taxon sampling is helping with it.
 

### 2021-04-06 2.4-discussion-ideas.Rmd
tidying up

@@
+There is ample evidence that increased taxon sampling benefits phylogenetic inference.
+It breaks long branches (Hillis 1996), increases phylogenetic accuracy (Zwickl and Hillis, 2002), resolves phylogenetic conflict between gene trees (Hedtke et al.
+2006), and resolves contentious phylogenetic relationships (Townsend & Lopez-Giraldez 2010; Natsidis et al.
+2019), reveal hidden orthologs and paralogs (Martin-Duran et al.
+2017).
+It has been the focus to give insight into phylogenetic relationships of many different taxa.
+Applied phylogenetic analyses have largely focused on improving taxon sampling to gain better insight in a group’s evolutionary history:
+in fish (Salmonidae, Crête et al. 2012; flounders, Vinnikov et al. 2018),
+birds (Wang et al. 2017)
+Gastropods (Oskars et al. 2015)
+Insects (Hemiptera, Song et al. 2019; Diptera, Shin et al. 2018; Lepidopetra, Wilson 2011; Orthoptera, Vandergast et al. 2017),
+amphibians (Peloso et al. 2016)
+Mammals (Cetartiodactyla, Agnarsson & May-Collado 2008)
+Squamate reptiles (Ketchum & Benson 2010)
+turtles (Vitek et al. 2018)
+flowering plants (Cornales, Xiang et al. 2011; Caryophyllales, Crawley & Hilu 2012; Campanula, Mansion et al. 2012, Rosids, Sun et al. 2020)
+unicellular organisms (Ciliates, Tang & Zhao 2016; Xu et al. 2021; chlamydomonas, Nakada et al. 2019)
+Spiders (Kallal et al. 2020)
+Sponges (Alvizu et al. 2018)
+fungi (Prsanna et al. 2020)
+biogeography (bees, Kayaalp et al. 2017; gallus, )
+Increased taxon sampling also improves time of divergence estimates (Schulte 2013; Soares & Schrago 2015)
+
 <!-- databases designed to preserve and democratize access to biological data constitute an amazing resource for evolutionary estimation -->

### 2021-04-06 sources/bmc_template_submission/bmc_article.tex
tidying up

@@
 the genus \textit{Ilex}.
+In order to showcase Physcraper's performance, we updated
+the tree with the smallest number of tips (the Gottlieb2005 tree) with the expectation
+to use the tree with the largest number of tips (the Yao2020 tree) as a standard
+for comparison and verification of Physcraper's results.
 
@@
 that together can be used to reconstruct species trees taking into account coalescent
-processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}).
+processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}.
 Rigorous analyses of multiple loci allows for more complex evolutionary models

### 2021-04-06 sources/bmc_template_submission/bmc_article.tex
fixing italics

@@
 %% Biological plausibility of example??
-The updated holly tree is based on a single marker, so we expected for it to be
-not as well resolved as trees using multiple markers. Nevertheless, all samples
-corresponding to the ingroup cluster together in a monophyletic group. We also
-note that most samples belonging to the same species also cluster together forming
-monophyletic groups. Notable exceptions are Ilex theeizans, which also appears as
-non-monophyletic in the original Gottlieb et al. 2005 tree. A visual comparison
+The updated holly tree is based on a single marker, so we expect for it to be
+not as well resolved as trees from studies using multiple markers. Although not perfect,
+we think the tree seems biologically reasonable in different ways. All samples
+corresponding to the ingroup cluster together forming a monophyletic group. We also
+note that samples belonging to the same \textit{Ilex} species also cluster together forming
+monophyletic groups. A notable exception is \textit{Ilex theeizans}, which appears as
+non-monophyletic in the updated tree and the original Gottlieb et al. 2005 tree.
+%This suggests that the samples might represent two ifferent species, but further analyses should be conducted
+A visual comparison
 of the Yao et al. 2020 tree and the original Gottlieb et al. 2005 tree suggests
-that the relationships within the genus Ilex are still being determined, and that
+that the relationships within the genus \textit{Ilex} are still being determined, and that
 increased taxon sampling is helping with it.

### 2021-04-06 sources/bmc_template_submission/bmc_article.tex
fixing uppercases

@@
 By gathering the sequences, and making the unaligned files easily available to users, researchers can compare if applying any alternate alignment tool of their choice affects inferences.
-Once sequences are aligned, they can apply and compare inferences from any Phylogenetic software.
+Once sequences are aligned, they can apply and compare inferences from any phylogenetic software.
 
@@
 Initial ideas about the data are constantly changed by results from new analyses.
-Physcraper provides a framework for reproducible Phylogenetics that has the
+Physcraper provides a framework for reproducible phylogenetics that has the
 potential to consistently provide context for these ideas, highlighting the

### 2021-04-19 sources/bmc_template_submission/bmc_article.tex
last changes before sending for co-author approval

@@
   corref={aff1},                       % id of corresponding address, if any
-% noteref={n1},                        % id's of article notes, if any
   email={sanchez.reyes.luna@gmail.com}   % email address
@@
 All authors contributed to the manuscript and gave final approval for publication.
-\ldots
 
@@
 and Engineering Building 1, UCM, joint lab paper discussion group for valuable
-comments on this manuscript.
+comments on this manuscript. We thank the valuable comments of David Posada, Rutger Vos,
+and an anonymous reviewer that greatly improved earlier versions of this manuscript.
 
@@
   \caption{A) Phylogeny updated with Physcraper from original
-  \cite{gottlieb2005molecular} tree in B.
+  \cite{gottlieb2005molecular} tree.
   Tips in original alignment and new tips added with Physcraper are depicted in
-  black and red, respectively. Physcraper obtained sequences from the GenBank
+  black and golden, respectively. Physcraper obtained sequences from the GenBank
   database via local BLAST of all sequences in the original alignment that generated
@@
   filtering", aligned them to original alignment using MUSCLE and performed a
-  phylogenetic reconstruction using RAxML with 100 bootstraps. B-D conflict analyses
-  performed with OpenTree tools.}
+  phylogenetic reconstruction using RAxML with 100 bootstraps.}
 \end{figure}

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
leaving just one corresponding author
pylint's design choices

@@
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-  corref={aff1},                       % id of corresponding address, if any
+  % corref={aff1},                       % id of corresponding address, if any
   email={sanchez.reyes.luna@gmail.com}   % email address
@@
 %% , classes, methods and functions.%%
-For testing and improving the quality of the Python code syntax in the Phsycraper code,
-we used the Pylint \cite{pylintWeb, pylintManual} software with a configuration file.
-We improved the code syntax on modules with low Pylint scores and fixed errors by
+For testing and improving Python code’s syntax quality of the Phsycraper code, we
+used the Pylint software following instructions from their website \citep{pylintWeb}
+ and manual \citep{pylintManual}, with a "".pylintrc" configuration file.
+
+We improved code's syntax on modules with low Pylint scores and fixed code errors by
 following Pylint's recommendations.
@@
 by using its check-disabling system, and are explained in the code.
-As of now, all modules have a Pylint score of 10/10.
+As of now, all Physcraper modules have a Pylint score of 10/10.
+
+% Pylint score for each module and locally disabled Pylint reccomendations
+% physcraper/aligntreetax.py		10/10	pylint disabling “too many branches” and “too many locals”
+% physcraper/configobj.py 		10/10  pylint global disabling attributes defined outside init
+% physcraper/helpers.py		10/10
+% physcraper/ids.py			10/10 	pylint disabled unused-variables that are used later by other functions
+% physcraper/ncbi_data_parser.py	10/10
+% physcraper/opentree_helpers.py 10/10		we pylint-disabled “too many local variables”, “too many branches” and “too many statements”
+% physcraper/scrape.py		10/10	commented out a bare except and pylint-disabled ”globals”, “too many public methods”, “too many local variables”, “too many lines in module”, “too many branches”, “too many statements” and “no self use”
+% physcraper/treetaxon.py		10/10
+% bin/find_trees.py			10/10
+% bin/physcraper.py			10/10
+% bin/tree_comparison.py		10/10
+% bin/multi_loci.py			10/10 		Global disabled “redefined-outer-name”, “protected-access”, “too-many-branches”, “too-many-statements”, “too-many-locals”
 

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
fixing cites

@@
 For testing and improving Python code’s syntax quality of the Phsycraper code, we
-used the Pylint software following instructions from their website \citep{pylintWeb}
- and manual \citep{pylintManual}, with a "".pylintrc" configuration file.
+used the Pylint software following instructions from their website \cite{pylintWeb}
+ and manual \cite{pylintManual}, with a ".pylintrc" configuration file.
 

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
fixing quotations

@@
 around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be leveraged as seeds to mine molecular databases, and as "jump-start" alignments
+can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 %% , classes, methods and functions.%%
-For testing and improving Python code’s syntax quality of the Phsycraper code, we
+For testing and improving Physcraper's Python code syntax quality, we
 used the Pylint software following instructions from their website \cite{pylintWeb}
- and manual \cite{pylintManual}, with a ".pylintrc" configuration file.
+ and manual \cite{pylintManual}, with a ``.pylintrc" configuration file.
 
-We improved code's syntax on modules with low Pylint scores and fixed code errors by
+We improved code syntax of Physcraper's modules with low Pylint scores, and fixed code errors by
 following Pylint's recommendations.
 Based on Physcraper's software design choices, some of Pylint's recommendations were overruled
-by using its check-disabling system, and are explained in the code.
+by using its check-disabling system, and are explained along the code.
 As of now, all Physcraper modules have a Pylint score of 10/10.
@@
 These sequences do appear in the Physcraper results in the list of matches from GenBank which did not fit the sequence length cutoffs set in the configuration.
-This "seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence length of all sequences filtered based on sequence length.
+This ``seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence length of all sequences filtered based on sequence length.
 
@@
 The default relative sequence length thresholds are arbitrary, but we hope that by making the process of locating homologous sequences online reproducible, and tracking what cutoffs are used, we make it easier for researchers to delve into the effect that different choices have on their inferences.
-This is in contrast to ‘manual’ searches for taxa, where similarly arbitrary cutoffs are applied, but are difficult to trace.
+This is in contrast to ``manual" searches for taxa, where similarly arbitrary cutoffs are applied, but are difficult to trace.
 As many studies have shown \cite{huang_what_2009} the effect of missing data can be enigmatic, and interact with the true phylogenetic relationships for the data set at hand.
@@
 While it would be possible to directly extract the blast match from genomes, this match would exclude potentially homologous flanking regions which are not matched by blasts local search algorithm, but may be important for phylogenetic inference.
-Instead we list the accession numbers for these matches in the "seqlen\_mismatch.txt" file, for users to assess appropriate homologous regions for their alignment of interest.
+Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt" file, for users to assess appropriate homologous regions for their alignment of interest.
 

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
page 1 MK edits

@@
 \kwd{DNA alignment}
+\kwd{Open Tree of Life}
+\kwd{otol}
 \end{keyword}
@@
 framework for biological databases to automate data connections across databases.
-Physcraper's main goal is to build upon published alignments to extend existing
-phylogenetic inferences with more data and taxa, which improve
+Physcraper's main goal is to build upon published alignments and published phylogenies
+to extend existing phylogenetic inferences with more data and taxa, which improve
 phylogenetic reconstruction \cite{hillis1996inferring, natsidis2019phylogenomics},

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
page 2 MK edits

@@
 2.3 million tips, of which around only
-90,000 are supported by phylogenetic data - the remaining 1.4 million taxa are placed
+90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem) by curators - the remaining 1.4 million taxa are placed
 in the tree based on taxonomy.
@@
 (TNRS) tool \cite{TNRStool}. Users can upload their
-own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
-Phylesystem \cite{phylesystemGithub, mctavish2015phylesystem} that also have
+own tree, or choose from among the 2, 950 curated trees stored in OpenTree's
+Phylesystem database \cite{phylesystemGithub, mctavish2015phylesystem} that also have
 alignments available on TreeBASE \cite{piel2009treebase}.

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
MK edits page 3 and 4

@@
 
-Physcraper is implemented with Python and can also be run on the command line.
+Physcraper is implemented with Python and can be run on a Python interactive session,
+as a Python script or using the command line interface we developped for it.
 It currently consists of 12 modules.
@@
 For testing and improving Physcraper's Python code syntax quality, we
-used the Pylint software following instructions from their website \cite{pylintWeb}
+used the Pylint software following instructions from its website \cite{pylintWeb}
  and manual \cite{pylintManual}, with a ``.pylintrc" configuration file.
@@
 and max length threshold, defined as the proportion of the average length without
-gaps of all sequences in input alignment (default values of 80% and 120%,
+gaps of all sequences in input alignment (default values of 80\% and 120\%,
 respectively); 4) or if they are either identical to or shorter than an existing
@@
 A user is interested in phylogenetic relationships within the genus \textit{Ilex}. Commonly
-known as "hollies", the genus encompasses between 400-700 living species, and is
+known as ``hollies", the genus encompasses between 400-700 living species, and is
 the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
@@
 
-An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
+An online literature review in June 2020 (Google scholar search for ``ilex phylogeny")
 reveals that there are several published phylogenies showing relationships within

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
MK edits verification test

@@
 
-We pruned 9 out of the 41 original tips of the Gottlieb et al. tree, corresponding
-to a 20\% trim, excluding the outgroups. We then performed a Physcraper run to test
-if we would recover the dropped tips.
+To test the accuracy of Physcraper we designed a verification test.
+We pruned 9 out of the 41 original tips of the Gottlieb2005 tree \cite{gottlieb2005molecular},
+corresponding to a 20\% trim, excluding the outgroups. We then performed a Physcraper
+run to test if we would recover the dropped tips.
 We successfully recovered 6 out of 9 pruned tips in the updated tree.
-Closer examination of results revealed that the 3 missing tips were correctly retrieved
-with BLAST along with the other 6 tips that were included in the updated tree. Examining
-the ITS sequences belonging to the tips missed by the test, by following the GenBank
-accession numbers reported in the original publication, showed that these sequences
-contain a 100 bp long gap of unidentified nucleotides (Ns). These Ns were absent
-in the 3 sequences published in the original alignment.
+Closer examination of results revealed that sequences for the 3 missing tips were correctly retrieved
+with BLAST along with sequences from the other 6 tips recovered in the updated tree.
+By following the GenBank accession numbers reported in the original publication
+belonging to the ITS sequences of the missing tips, we observed that these three sequences
+contain a 100 bp long gap of unidentified nucleotides (Ns) that is completely absent
+from any of the sequences in the original alignment. This caused these three GenBank sequences
+in particular to exceed Physcraper's default sequence length cutoff of 120\%, being thus
+filtered and excluded from the alignment step onwards.
+These sequences do appear in the Physcraper results in the list of matches from
+GenBank which did not fit the sequence length cutoffs set in the configuration file.
+This ``seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence
+length of all sequences filtered based on sequence length.
 
@@
 of ignoring the information they contain and building homology statements fully \textit{de novo}.
-The three sequences that were in the original tree but were not included in
-the updated tree by Physcraper in the verification test were excluded because
-they contain a 100 bp long gap of unidentified nucleotides (Ns) in GenBank that
-was manually curated and removed from in the original alignment.
-
-These sequences do appear in the Physcraper results in the list of matches from GenBank which did not fit the sequence length cutoffs set in the configuration.
-This ``seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence length of all sequences filtered based on sequence length.
 

### 2021-04-20 sources/bmc_template_submission/bmc_article.tex
MK edits page 6

@@
 Physcraper is a tool that builds upon previous knowledge stored in published
-alignments, with the main goal to extend our knowledge of phylogenetic relationships
+alignments and phylogenies, taking advantage of OpenTree's services to facilitate comparison
+of phylogenies, with the main goal of extending our knowledge of phylogenetic relationships
 across the tree of life.
@@
 
-We encourage users to look at this file, and potentially modify their sequence length cutoff parameters based on the excluded sequences.
-The default relative sequence length thresholds are arbitrary, but we hope that by making the process of locating homologous sequences online reproducible, and tracking what cutoffs are used, we make it easier for researchers to delve into the effect that different choices have on their inferences.
-This is in contrast to ``manual" searches for taxa, where similarly arbitrary cutoffs are applied, but are difficult to trace.
-As many studies have shown \cite{huang_what_2009} the effect of missing data can be enigmatic, and interact with the true phylogenetic relationships for the data set at hand.
-There is not currently strong support in the literature for any cutoffs, and rather than prescribe specific approaches, we encourage users to explore the effects of different choices on inference.
+We encourage users to look at the output files containing information about the
+filtered sequences, and potentially modify configuration parameters such as the
+sequence length cutoff parameter, based on the filtered sequences.
+Default filtering parameters are arbitrary, but we hope that by making the process
+of locating homologous sequences online reproducible, and tracking what filters
+are used, we make it easier for researchers to delve into the effect that different
+choices have on their inferences.
+This is in contrast to ``manual" searches for taxa, where similarly arbitrary filters
+are applied, but are difficult to trace.
+As many studies have shown \cite{huang_what_2009} the effect of missing data can
+be enigmatic, and interact with the true phylogenetic relationships for the data
+set at hand.
+There is not currently strong support in the literature for any particular cutoffs
+value, and rather than prescribe specific approaches, we encourage users to explore
+the effects of different choices on their phylogenetic inferences.
+In addition, by providing the output files at each step of the analysis,
+it is straightforward to assess how changing parameter and software choices do or
+do not drive differences in phylogenetic inference.
+By gathering the sequences, and making the unaligned files easily available to users,
+researchers can compare if applying any alternate alignment tool of their choice
+affects inferences.
+Once sequences are aligned, they can apply and compare inferences from any phylogenetic
+software.
 
-Organellar genome sequences, such as chloroplasts and mitochondria will also generally be excluded from automatic addition based on length cutoffs.
-Multiple sequence alignment of loci of drastically different lengths is unfeasible, and we have found in testing that it often returns incorrect results, splitting shorter sequences with many long gaps to align with exact matches across the entire longer locus.
-While it would be possible to directly extract the blast match from genomes, this match would exclude potentially homologous flanking regions which are not matched by blasts local search algorithm, but may be important for phylogenetic inference.
-Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt" file, for users to assess appropriate homologous regions for their alignment of interest.
+Organellar genome sequences, such as chloroplasts and mitochondria will also generally
+be excluded from automatic addition based on length cutoffs.
+Multiple sequence alignment of loci of drastically different lengths is unfeasible,
+and we have found in testing that it often returns incorrect results, splitting
+shorter sequences with many long gaps to align with exact matches across the entire
+longer locus.
+While it would be possible to directly extract the blast match from genomes, this
+match would exclude potentially homologous flanking regions which are not matched
+by blasts local search algorithm, but may be important for phylogenetic inference.
+Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt"
+file, for users to assess appropriate homologous regions for their alignment of interest.
 
-Single gene phylogenies with very high numbers of taxa may lack sufficient signal for accurate resolution \cite{morel_phylogenetic_2020}.
-The Physcraper workflow avoids some of these challenges, by focussing on ingroup taxa of an existing phylogeny, using markers that have been assessed and proven apporpraite for that phylogenetic scope in past publictions.
-In addition, Physcraper thins alignments by removing identical sequences, and by setting a maximum number of sequences per taxon.
-Nonetheless, it is incumbent on users to assess their final inference with respect to statistical support and biological plausibility.
-
-Most of the default thresholds parameter values and software choices in the Physcraper pipeline are arbitrary.
-We hope that by making the process of locating homologous sequences online reproducible,
-and tracking what cutoffs are used, we make it easier for researchers to delve into the effect that different choices have on their inferences.
-In addition, by providing the output files at each step of the analysis,
-it is straightforward to assess how changing parameter and software choices do or do not drive differences in phylogenetic inference.
-By gathering the sequences, and making the unaligned files easily available to users, researchers can compare if applying any alternate alignment tool of their choice affects inferences.
-Once sequences are aligned, they can apply and compare inferences from any phylogenetic software.
+Single gene phylogenies with very high numbers of taxa may lack sufficient signal
+for accurate resolution \cite{morel_phylogenetic_2020}.
+The Physcraper workflow avoids some of these challenges, by focussing on ingroup
+taxa of an existing phylogeny, using markers that have been assessed and proven
+apporpraite for that phylogenetic scope in past publictions.
+In addition, Physcraper thins alignments by removing identical sequences, and by
+setting a maximum number of sequences per taxon.
+Nonetheless, it is incumbent on users to assess their final inference with respect
+to statistical support and biological plausibility.
 
@@
 
-For example \cite{zhu_complexity_2021} show that when applying coalescent models, there is more information in two genes of 300 bp each than in 600 independent sites
-We have included a script (``multi\_locus.py'') and examples of how to automatically merge the outputs of Physcraper runs at different loci into input files for ASTRAL, SVD Quartets, or as concatenated alignments for concatenated analyses.
+For example \cite{zhu_complexity_2021} show that when applying coalescent models,
+there is more information in two genes of 300 bp each than in 600 independent sites
+We have included a script (``multi\_locus.py'') and examples of how to automatically
+merge the outputs of Physcraper runs at different loci into input files for ASTRAL,
+SVD Quartets, or as concatenated alignments for concatenated analyses.
 
@@
 geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
-Taxonomic links, and comparisons to existing published phylogenies in the OpenTree data store can also help flag paralogous sequences.
-Accidentally including paralogs as homologs is a risk in phylogenetic analyses, and can be more prevalent in automated analyses than in manually curated analyses.
+Taxonomic links, and comparisons to existing published phylogenies in the OpenTree
+data store can also help flag paralogous sequences.
+Accidentally including paralogs as homologs is a risk in phylogenetic analyses,
+and can be more prevalent in automated analyses than in manually curated analyses.
 We provide users with several tools to try to assess homology of their aligned sequences.
-The estimated gene tree itself is an evolutionarily explicit way to visualize gene evolution, which in concert with taxonomic labelling can reveal paralogy.
-Our conflict analysis tool informs the users of whether their tree contains major conflicts with established taxonomy.
-This conflict tool also returns information on if these taxonomic conflicts existed in the original input tree.
-These conflicts may be a sign that taxonomy needs to be updated, or may be a sign that non-homologous sequences have been included in the analysis.
-These taxonomic and phylogenetic conflicts flag regions for the tree for the researcher to more closely examine and assess homology.
+The estimated gene tree itself is an evolutionarily explicit way to visualize gene
+evolution, which in concert with taxonomic labelling can reveal paralogy.
+Our conflict analysis tool informs the users of whether their tree contains major
+conflicts with established taxonomy.
+This conflict tool also returns information on if these taxonomic conflicts existed
+in the original input tree.
+These conflicts may be a sign that taxonomy needs to be updated, or may be a sign
+that non-homologous sequences have been included in the analysis.
+These taxonomic and phylogenetic conflicts flag regions for the tree for the researcher
+to more closely examine and assess homology.
 
@@
 non-monophyletic in the updated tree and the original Gottlieb et al. 2005 tree.
-%This suggests that the samples might represent two ifferent species, but further analyses should be conducted
-A visual comparison
-of the Yao et al. 2020 tree and the original Gottlieb et al. 2005 tree suggests
+%This suggests that the samples might represent two ifferent species, but further
+analyses should be conducted
+A visual comparison of the Yao et al. 2020 tree and the original Gottlieb et al. 2005 tree suggests
 that the relationships within the genus \textit{Ilex} are still being determined, and that

### 2021-04-22 sources/bmc_template_submission/bmc_article.tex
MK edits page 7

@@
 
-Physcraper improves on previous work in automating phylogenetic reconstruction by leveraging
-the power of existing homology hypotheses that taxon specialists have assessed
+Physcraper improves on previous work to automate phylogenetic reconstruction by leveraging
+the power of existing homology hypotheses that phylogeneticists and taxon specialists have assessed
 and deemed appropriate for a specific phylogenetic scope.
@@
 around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
+can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 Physcraper is implemented with Python and can be run on a Python interactive session,
-as a Python script or using the command line interface we developped for it.
-It currently consists of 12 modules.
+as a Python script or using the command line interface developed for it.
+It currently consists of 13 modules.
 %% , classes, methods and functions.%%
@@
 set at hand.
-There is not currently strong support in the literature for any particular cutoffs
+There is not currently strong support in the literature for any particular cutoff
 value, and rather than prescribe specific approaches, we encourage users to explore
@@
 Organellar genome sequences, such as chloroplasts and mitochondria will also generally
-be excluded from automatic addition based on length cutoffs.
+be excluded from automatic addition based on default Physcraper length cutoffs.
 Multiple sequence alignment of loci of drastically different lengths is unfeasible,
@@
 longer locus.
-While it would be possible to directly extract the blast match from genomes, this
-match would exclude potentially homologous flanking regions which are not matched
-by blasts local search algorithm, but may be important for phylogenetic inference.
+While it would be possible to directly extract the BLAST match from genomes, this
+ would exclude potentially homologous flanking regions which are not matched
+by BLAST's local search algorithm, but that may be important for phylogenetic inference.
 Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt"
-file, for users to assess appropriate homologous regions for their alignment of interest.
-
-Single gene phylogenies with very high numbers of taxa may lack sufficient signal
-for accurate resolution \cite{morel_phylogenetic_2020}.
-The Physcraper workflow avoids some of these challenges, by focussing on ingroup
-taxa of an existing phylogeny, using markers that have been assessed and proven
-apporpraite for that phylogenetic scope in past publictions.
-In addition, Physcraper thins alignments by removing identical sequences, and by
-setting a maximum number of sequences per taxon.
-Nonetheless, it is incumbent on users to assess their final inference with respect
-to statistical support and biological plausibility.
+file, for users to assess and incorporate appropriate homologous regions to their
+alignment of interest.
 
@@
 Physcraper generates gene trees, which individually do not capture the full complexity
-of species' evolutionary history \cite{song2012resolving}.
-However, Physcraper facilitates
+of species' evolutionary history \cite{song2012resolving}. In addition, single gene
+phylogenies with very high numbers of taxa may lack sufficient signal
+for accurate phylogenetic resolution \cite{morel_phylogenetic_2020}.
+The Physcraper workflow avoids this challenge by focusing on ingroup
+taxa of an existing phylogeny, using markers that have been assessed and proven
+appropriate for that phylogenetic scope in past publications.
+Also, Physcraper thins alignments by removing sequences identical to original
+and newly added sequences, and by setting a maximum number of sequences per taxon.
+Nonetheless, it is incumbent on users to assess their final inference with respect
+to statistical support and biological plausibility.
+
+Rigorous analyses of multiple loci allow for more complex evolutionary models
+than analyses of large genomic data sets, and can provide better evolutionary estimates.
+For example \cite{zhu_complexity_2021} show that when applying coalescent models,
+there is more information in two genes of 300 bp each than in 600 independent sites.
+Physcraper is desighed to facilitate
 gathering alignments and gene trees for multiple loci from a group of interest,
@@
 processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}.
-Rigorous analyses of multiple loci allows for more complex evolutionary models
-than analyses of large genomic data sets, and can provide better evolutionary estimates.
-
-For example \cite{zhu_complexity_2021} show that when applying coalescent models,
-there is more information in two genes of 300 bp each than in 600 independent sites
 We have included a script (``multi\_locus.py'') and examples of how to automatically
-merge the outputs of Physcraper runs at different loci into input files for ASTRAL,
-SVD Quartets, or as concatenated alignments for concatenated analyses.
+merge the outputs of Physcraper runs at different loci into input files for these
+two software, or as concatenated alignments for supermatrix analyses.
 
@@
 conflicts with established taxonomy.
-This conflict tool also returns information on if these taxonomic conflicts existed
+This tool also returns information on taxonomic conflicts that exist
 in the original input tree.
-These conflicts may be a sign that taxonomy needs to be updated, or may be a sign
+Detected conflicts may be a sign that taxonomy needs to be updated, or may be a sign
 that non-homologous sequences have been included in the analysis.
-These taxonomic and phylogenetic conflicts flag regions for the tree for the researcher
+These taxonomic and phylogenetic conflicts flag regions of the tree for the researcher
 to more closely examine and assess homology.

### 2021-04-22 sources/bmc_template_submission/bmc_article.tex
MK edits page 8 and 9

@@
   %\postcode{}
-  \city{Prague },
+  \city{Prague},
   \cny{Czech Republic}
@@
 
+%% Biological plausibility of example??
+Our example application of Physcraper to update a holly phylogeny is based on a
+single marker, so we expect for it to be not as well resolved as trees from studies
+using multiple markers. Although not perfect,
+we think the updated phylogeny seems biologically reasonable in different ways. All samples
+corresponding to the ingroup cluster together forming a monophyletic group, and
+samples belonging to the same \textit{Ilex} species also form
+monophyletic groups. A notable exception is \textit{Ilex theeizans}, which appears as
+non-monophyletic in the Physcraper updated phylogeny as well as in the original Gottlieb2005 tree.
+%This suggests that the samples might represent two different species, but further
+analyses should be conducted
+A visual comparison of the Yao2020 tree and the original Gottlieb2005 tree suggests
+that the relationships within the genus \textit{Ilex} are still being actively
+determined, and that increased taxon sampling might be key to resolve them.
+
 Physcraper has the added advantage of facilitating the linkage of taxonomic information about tips
@@
 
-%% Biological plausibility of example??
-The updated holly tree is based on a single marker, so we expect for it to be
-not as well resolved as trees from studies using multiple markers. Although not perfect,
-we think the tree seems biologically reasonable in different ways. All samples
-corresponding to the ingroup cluster together forming a monophyletic group. We also
-note that samples belonging to the same \textit{Ilex} species also cluster together forming
-monophyletic groups. A notable exception is \textit{Ilex theeizans}, which appears as
-non-monophyletic in the updated tree and the original Gottlieb et al. 2005 tree.
-%This suggests that the samples might represent two ifferent species, but further
-analyses should be conducted
-A visual comparison of the Yao et al. 2020 tree and the original Gottlieb et al. 2005 tree suggests
-that the relationships within the genus \textit{Ilex} are still being determined, and that
-increased taxon sampling is helping with it.
-
 The Physcraper workflow can be used to rapidly (in a matter of hours) create
@@
 potential to consistently provide context for these ideas, highlighting the
-importance of data sharing and open science in the field, biology and science.
+importance of data sharing and open science in phylogenetics, biology and science.
 
@@
 
-\textbf{Project name:} Physcraper
+\par{\textbf{Project name:} Physcraper}
 
@@
 
-\textbf{OpenTree:} The Open Tree of Life project
+\par{\textbf{OpenTree:} The Open Tree of Life project}
 
@@
 
-The datasets generated and analysed during the current study are available in
+Code and datasets generated and analysed during the current study are available in
 the repositories "physcraper" containing the source code,
@@
 and Engineering Building 1, UCM, joint lab paper discussion group for valuable
-comments on this manuscript. We thank the valuable comments of David Posada, Rutger Vos,
-and an anonymous reviewer that greatly improved earlier versions of this manuscript.
+comments on this manuscript. We also thank the valuable comments of David Posada, Rutger Vos,
+and an anonymous reviewer that greatly improved an earlier version of this manuscript.
 

### 2021-04-22 sources/bmc_template_submission/bmc_article.tex
MK several edits to references

@@
 There are almost 8,200 publicly available, peer-reviewed curated alignments, covering
-around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
-can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
+around 100,000 distinct taxa in the TreeBASE database
+\cite{piel2009treebase, vos2012nexml, piel2018treebasedmp, treebase_website, supertreebase}, 
+which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 to alignments
-from TreeBASE \cite{sanderson1994treebase}, and data from GenBank \cite{benson2000genbank}.
+from TreeBASE, and data from GenBank \cite{benson2000genbank}.
 Its utility and functionalities are presented with a case-study analysis of a
@@
 Phylesystem database \cite{phylesystemGithub, mctavish2015phylesystem} that also have
-alignments available on TreeBASE \cite{piel2009treebase}.
+alignments available on TreeBASE.
 

### 2021-04-23 sources/bmc_template_submission/bmc_article.tex
more edits to references

@@
 BLAST is performed using the `blastn` algorithm \cite{camacho2009blast}
-implemented in BioPython's \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
+implemented in BioPython 1.71 \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
 modified to accept an alternative BLAST address.

### 2021-04-28 sources/bmc_template_submission/bmc_article.tex
final edits to figure caption

@@
 
-Rigorous analyses of multiple loci allow for more complex evolutionary models
+In the era of phylogenomics, rigorous analyses of multiple loci still allow for
+more complex evolutionary models
 than analyses of large genomic data sets, and can provide better evolutionary estimates.
@@
 evolution, which in concert with taxonomic labelling can reveal paralogy.
-Our conflict analysis tool informs the users of whether their tree contains major
-conflicts with established taxonomy.
-This tool also returns information on taxonomic conflicts that exist
+OpenTree's conflict analysis tool informs the users of whether their tree contains major
+conflicts with established taxonomy and any phylogenetic context they wish to compare to.
+This tool also returns information on taxonomic and phylogenetic conflicts that exist
 in the original input tree.
@@
 \begin{figure}[h!]
-  \caption{A) Phylogeny updated with Physcraper from original
-  \cite{gottlieb2005molecular} tree.
+  \caption{A) Phylogeny updated with Physcraper using a starting phylogeny and alignment from
+  \cite{gottlieb2005molecular} (Gottlieb2005 data in text).
   Tips in original alignment and new tips added with Physcraper are depicted in
-  black and golden, respectively. Physcraper obtained sequences from the GenBank
-  database via local BLAST of all sequences in the original alignment that generated
-  tree in B), filtered them following criteria from section "DNA sequence search and
-  filtering", aligned them to original alignment using MUSCLE and performed a
-  phylogenetic reconstruction using RAxML with 100 bootstraps.}
+  black and red, respectively. Physcraper obtained sequences from the GenBank
+  database via local BLAST of all sequences in the Gottlieb2005 internal transcribed
+  spacer DNA region (ITS) alignment,
+  filtered them following criteria described in section "DNA sequence search and
+  filtering", aligned them to the original alignment using MUSCLE, and performed a
+  phylogenetic reconstruction using RAxML, with 100 bootstraps. B) Results of conflict
+  analysis performed using OpenTree's conflict tool \cite{redelings2017supertree}.
+  The Physcraper updated Gottlieb2005 phylogeny in (A) was compared to an \textit{Ilex}
+  OpenTree synthetic subtree v. 12.3 \cite{ott727571} constructed using taxonomy of the genus
+  as backbone and resolving branches based on phylogenetic data from the original
+  Gottlieb2005 phylogeny.
 \end{figure}

### 2021-04-28 sources/bmc_template_submission/bmc_article_submission_2021-01-08.tex
final edits to figure caption

--- 2021-04-22 sources/bmc_template_submission/bmc_article.tex
@@
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-  % corref={aff1},                       % id of corresponding address, if any
+  corref={aff1},                       % id of corresponding address, if any
+% noteref={n1},                        % id's of article notes, if any
   email={sanchez.reyes.luna@gmail.com}   % email address
@@
   %\postcode{}
-  \city{Prague},
+  \city{Prague },
   \cny{Czech Republic}
@@
 \kwd{DNA alignment}
-\kwd{Open Tree of Life}
-\kwd{otol}
 \end{keyword}
@@
 context for our biological observations \cite{dobzhansky1973nothing}.
-Updating existing phylogenies with publicly available molecular sequence data provides
-the opportunity to study
+Updating existing phylogenies with publicly available molecular sequence data that
+has never been
+incorporated into any phylogenetic estimate provides the opportunity to study
 the evolutionary history of many taxa in a reproducible and continuous manner.
 Here, we introduce Physcraper, a tool that establishes a data interoperability
-framework for biological databases to automate data connections across databases.
-Physcraper's main goal is to build upon published alignments and published phylogenies
-to extend existing phylogenetic inferences with more data and taxa, which improve
-phylogenetic reconstruction \cite{hillis1996inferring, natsidis2019phylogenomics},
-time of divergence estimates \cite{schulte2013undersampling, soares2015influence},
-biogeographic analyses \cite{kayaalp2017back},
-as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
-Physcraper updates a starting single locus alignment and corresponding tree
-with public DNA data, and links the tips in the updated trees to a unified, interoperable
+framework for biological databases to automate data connections across databases,
+with the main goal of building on published alignments and extending existing
+phylogenetic inferences with more data and taxa.
+Physcraper updates a starting tree and single locus alignments
+with public DNA data, and links the tips in these trees to a unified, interoperable
 taxonomic resource \cite{rees2017automated}.
@@
 distinct taxa in the NCBI taxonomy at the time \cite{sanderson2008phylota}.
-However, even thirteen years later, phylogenetic estimates for most of these taxa
-are still not available \cite{mctavish_how_2017}.
-The Open Tree of Life project assembles a comprehensive synthetic tree of life comprising
-2.3 million tips, of which around only
-90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem) by curators - the remaining 1.4 million taxa are placed
+However, even thirteen years later, phylogenetic estimates for many of these taxa are not available \cite{mctavish_how_2017}.
+OpenTree's comprehensive tree of life comprises 2.3 million tips, of which around
+90,000 are supported by phylogenies - the remaining 1.4 million taxa are placed
 in the tree based on taxonomy.
@@
 
-Physcraper improves on previous work to automate phylogenetic reconstruction by leveraging
-the power of existing homology hypotheses that phylogeneticists and taxon specialists have assessed
+Physcraper improves on previous work in automating phylogenetic reconstruction by leveraging
+the power of existing homology hypotheses that taxon specialists have assessed
 and deemed appropriate for a specific phylogenetic scope.
 There are almost 8,200 publicly available, peer-reviewed curated alignments, covering
-around 100,000 distinct taxa in the TreeBASE database
-\cite{piel2009treebase, vos2012nexml, piel2018treebasedmp, treebase_website, supertreebase}, 
-which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
+around 100,000 distinct taxa in the TreeBASE database \cite{piel2009treebase}, which
+can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
@@
 to alignments
-from TreeBASE, and data from GenBank \cite{benson2000genbank}.
+from TreeBASE \cite{sanderson1994treebase}, and data from GenBank \cite{benson2000genbank}.
 Its utility and functionalities are presented with a case-study analysis of a
@@
 \section*{Implementation}
-
-Physcraper is implemented with Python and can be run on a Python interactive session,
-as a Python script or using the command line interface developed for it.
-It currently consists of 13 modules.
-%% , classes, methods and functions.%%
-For testing and improving Physcraper's Python code syntax quality, we
-used the Pylint software following instructions from its website \cite{pylintWeb}
- and manual \cite{pylintManual}, with a ``.pylintrc" configuration file.
-
-We improved code syntax of Physcraper's modules with low Pylint scores, and fixed code errors by
-following Pylint's recommendations.
-Based on Physcraper's software design choices, some of Pylint's recommendations were overruled
-by using its check-disabling system, and are explained along the code.
-As of now, all Physcraper modules have a Pylint score of 10/10.
-
-% Pylint score for each module and locally disabled Pylint reccomendations
-% physcraper/aligntreetax.py		10/10	pylint disabling “too many branches” and “too many locals”
-% physcraper/configobj.py 		10/10  pylint global disabling attributes defined outside init
-% physcraper/helpers.py		10/10
-% physcraper/ids.py			10/10 	pylint disabled unused-variables that are used later by other functions
-% physcraper/ncbi_data_parser.py	10/10
-% physcraper/opentree_helpers.py 10/10		we pylint-disabled “too many local variables”, “too many branches” and “too many statements”
-% physcraper/scrape.py		10/10	commented out a bare except and pylint-disabled ”globals”, “too many public methods”, “too many local variables”, “too many lines in module”, “too many branches”, “too many statements” and “no self use”
-% physcraper/treetaxon.py		10/10
-% bin/find_trees.py			10/10
-% bin/physcraper.py			10/10
-% bin/tree_comparison.py		10/10
-% bin/multi_loci.py			10/10 		Global disabled “redefined-outer-name”, “protected-access”, “too-many-branches”, “too-many-statements”, “too-many-locals”
 
@@
 (TNRS) tool \cite{TNRStool}. Users can upload their
-own tree, or choose from among the 2, 950 curated trees stored in OpenTree's
-Phylesystem database \cite{phylesystemGithub, mctavish2015phylesystem} that also have
-alignments available on TreeBASE.
+own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
+Phylesystem \cite{phylesystemGithub, mctavish2015phylesystem} that also have
+alignments available on TreeBASE \cite{piel2009treebase}.
 
@@
 and max length threshold, defined as the proportion of the average length without
-gaps of all sequences in input alignment (default values of 80\% and 120\%,
+gaps of all sequences in input alignment (default values of 80% and 120%,
 respectively); 4) or if they are either identical to or shorter than an existing
@@
 Updated and original tree are compared with Robinson-Foulds weighted and unweighted
-metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
+metrics estimated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
 comparison between the synthetic OpenTree and original and updated tree individually,
@@
 A user is interested in phylogenetic relationships within the genus \textit{Ilex}. Commonly
-known as ``hollies", the genus encompasses between 400-700 living species, and is
+known as "hollies", the genus encompasses between 400-700 living species, and is
 the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
@@
 
-An online literature review in June 2020 (Google scholar search for ``ilex phylogeny")
+An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
 reveals that there are several published phylogenies showing relationships within
@@
 \cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
-(study 1091 \cite{treebase1091}). This tree (Gottlieb2005 tree from now on) samples 41 species, is available from
+(study 1091 \cite{treebase1091}). The tree sampling 41 species is also available from
 OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
 and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
-The most recent \textit{Ilex} tree \cite{yao2020phylogeny} (Yao2020 tree from now on) is available in
+The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
 OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
 and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the Yao2020 tree \cite{yao2020phylogeny} is the best sampled phylogeny available for
-the genus \textit{Ilex}.
-In order to showcase Physcraper's performance, we updated
-the tree with the smallest number of tips (the Gottlieb2005 tree) with the expectation
-to use the tree with the largest number of tips (the Yao2020 tree) as a standard
-for comparison and verification of Physcraper's results.
+With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
+the hollies.
 
 We ran Physcraper on a laptop Linux computer to update an internal transcribed
-spacer DNA region (ITS) alignment that was used to construct the tree from \cite{gottlieb2005molecular},
-using a local GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
+spacer DNA region (ITS) alignment from \cite{gottlieb2005molecular}, using a local
+GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
 analyses taking an additional 13hrs.
 %% <!-- MTH: you probably need some details about the hardware, given the fact that you are discussing running times -->
-The Gottlieb2005 tree \cite{gottlieb2005molecular} updated using Physcraper (Figure 2;
-Physcraper updated tree from now on) displays all 41 distinct
+The updated \cite{gottlieb2005molecular} tree (Figure 2) displays
+all 41 distinct
 taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional \textit{Ilex} taxa. The best ML tree from the RAxML analysis is 99\%
-resolved, with 25\% of nodes with bootstrap support $<$ 0.1 and 48\% nodes with
-bootstrap support $>$ 0.75.
+84 additional \textit{Ilex} taxa. The best RaxML tree is 99\% resolved, with 25\% of nodes
+with bootstrap support $<$ 0.1 and 48\% nodes with bootstrap support $>$ 0.75.
 A large portion of internal branches are negligibly small, with 30 branches
@@
 (Figure 2).
-As comparison with the Physcraper updated tree, the Yao2020 tree \cite{yao2020phylogeny}
-also contains all 41 distinct taxa
-sampled in the Gottlieb2005 tree \cite{gottlieb2005molecular},
-while contributing phylogenetic data to 134 additional \textit{Ilex} taxa. From these,
-67 taxa are also in the Physcraper updated tree.
-While the Yao2020 tree \cite{yao2020phylogeny} was also constructed using
-ITS as a marker, their GenBank data is not released yet. Hence, Physcraper was unable
-to incorporate 68 taxa that are only on the Yao2020 tree because the DNA data is unavailable.
-We also note that Physcraper incorporates 18 \textit{Ilex} taxa that are not in the Yao2020
-tree \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex} species
-and visual inspection of the DNA sequences suggests they are correctly assigned as \textit{Ilex}.
-The ITS alignment that underlies the Yao2020 tree was constructed without any tool to
-scrape GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
-incorporate these 18 additional \textit{Ilex} taxa in the updated tree (Figure 2).
-
-%% Maybe show conflict results? and then discuss in the discussion.
-
-\subsection*{Verification test}
-
-To test the accuracy of Physcraper we designed a verification test.
-We pruned 9 out of the 41 original tips of the Gottlieb2005 tree \cite{gottlieb2005molecular},
-corresponding to a 20\% trim, excluding the outgroups. We then performed a Physcraper
-run to test if we would recover the dropped tips.
-We successfully recovered 6 out of 9 pruned tips in the updated tree.
-Closer examination of results revealed that sequences for the 3 missing tips were correctly retrieved
-with BLAST along with sequences from the other 6 tips recovered in the updated tree.
-By following the GenBank accession numbers reported in the original publication
-belonging to the ITS sequences of the missing tips, we observed that these three sequences
-contain a 100 bp long gap of unidentified nucleotides (Ns) that is completely absent
-from any of the sequences in the original alignment. This caused these three GenBank sequences
-in particular to exceed Physcraper's default sequence length cutoff of 120\%, being thus
-filtered and excluded from the alignment step onwards.
-These sequences do appear in the Physcraper results in the list of matches from
-GenBank which did not fit the sequence length cutoffs set in the configuration file.
-This ``seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence
-length of all sequences filtered based on sequence length.
+For comparison, \cite{yao2020phylogeny} also contains all 41 distinct taxa from
+the original
+\cite{gottlieb2005molecular} study,
+and contributes phylogenetic data to 134 additional \textit{Ilex} taxa, from which
+67 are also in updated \cite{gottlieb2005molecular}. While \cite{yao2020phylogeny} also used
+ITS as a marker, their GenBank data is not released yet, so Physcraper was unable
+to incorporate 68 additional taxa into the analysis. However, Physcraper was able
+to incorporate 18 taxa that were not in \cite{yao2020phylogeny}.
 
@@
 existent evolutionary knowledge contribute to the acceleration of scientific discovery.
-
-Physcraper is a tool that builds upon previous knowledge stored in published
-alignments and phylogenies, taking advantage of OpenTree's services to facilitate comparison
-of phylogenies, with the main goal of extending our knowledge of phylogenetic relationships
-across the tree of life.
 
@@
 This decreases error (requiring less manual downstream processing) and
-eases comparison with previous phylogenetic knowledge.
-
-Results from the verification test highlight the importance of incorporating existing expertly
-curated homology statements to automatically update phylogenetic relationships, instead
-of ignoring the information they contain and building homology statements fully \textit{de novo}.
-
-We encourage users to look at the output files containing information about the
-filtered sequences, and potentially modify configuration parameters such as the
-sequence length cutoff parameter, based on the filtered sequences.
-Default filtering parameters are arbitrary, but we hope that by making the process
-of locating homologous sequences online reproducible, and tracking what filters
-are used, we make it easier for researchers to delve into the effect that different
-choices have on their inferences.
-This is in contrast to ``manual" searches for taxa, where similarly arbitrary filters
-are applied, but are difficult to trace.
-As many studies have shown \cite{huang_what_2009} the effect of missing data can
-be enigmatic, and interact with the true phylogenetic relationships for the data
-set at hand.
-There is not currently strong support in the literature for any particular cutoff
-value, and rather than prescribe specific approaches, we encourage users to explore
-the effects of different choices on their phylogenetic inferences.
-In addition, by providing the output files at each step of the analysis,
-it is straightforward to assess how changing parameter and software choices do or
-do not drive differences in phylogenetic inference.
-By gathering the sequences, and making the unaligned files easily available to users,
-researchers can compare if applying any alternate alignment tool of their choice
-affects inferences.
-Once sequences are aligned, they can apply and compare inferences from any phylogenetic
-software.
-
-Organellar genome sequences, such as chloroplasts and mitochondria will also generally
-be excluded from automatic addition based on default Physcraper length cutoffs.
-Multiple sequence alignment of loci of drastically different lengths is unfeasible,
-and we have found in testing that it often returns incorrect results, splitting
-shorter sequences with many long gaps to align with exact matches across the entire
-longer locus.
-While it would be possible to directly extract the BLAST match from genomes, this
- would exclude potentially homologous flanking regions which are not matched
-by BLAST's local search algorithm, but that may be important for phylogenetic inference.
-Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt"
-file, for users to assess and incorporate appropriate homologous regions to their
-alignment of interest.
-
+eases comparison with previous phylogenetic
+knowledge.
 Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
@@
 Physcraper generates gene trees, which individually do not capture the full complexity
-of species' evolutionary history \cite{song2012resolving}. In addition, single gene
-phylogenies with very high numbers of taxa may lack sufficient signal
-for accurate phylogenetic resolution \cite{morel_phylogenetic_2020}.
-The Physcraper workflow avoids this challenge by focusing on ingroup
-taxa of an existing phylogeny, using markers that have been assessed and proven
-appropriate for that phylogenetic scope in past publications.
-Also, Physcraper thins alignments by removing sequences identical to original
-and newly added sequences, and by setting a maximum number of sequences per taxon.
-Nonetheless, it is incumbent on users to assess their final inference with respect
-to statistical support and biological plausibility.
-
-Rigorous analyses of multiple loci allow for more complex evolutionary models
-than analyses of large genomic data sets, and can provide better evolutionary estimates.
-For example \cite{zhu_complexity_2021} show that when applying coalescent models,
-there is more information in two genes of 300 bp each than in 600 independent sites.
-Physcraper is desighed to facilitate
+of species' evolutionary history \cite{song2012resolving}.
+However, Physcraper facilitates
 gathering alignments and gene trees for multiple loci from a group of interest,
 that together can be used to reconstruct species trees taking into account coalescent
-processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}.
-We have included a script (``multi\_locus.py'') and examples of how to automatically
-merge the outputs of Physcraper runs at different loci into input files for these
-two software, or as concatenated alignments for supermatrix analyses.
-
-%% Biological plausibility of example??
-Our example application of Physcraper to update a holly phylogeny is based on a
-single marker, so we expect for it to be not as well resolved as trees from studies
-using multiple markers. Although not perfect,
-we think the updated phylogeny seems biologically reasonable in different ways. All samples
-corresponding to the ingroup cluster together forming a monophyletic group, and
-samples belonging to the same \textit{Ilex} species also form
-monophyletic groups. A notable exception is \textit{Ilex theeizans}, which appears as
-non-monophyletic in the Physcraper updated phylogeny as well as in the original Gottlieb2005 tree.
-%This suggests that the samples might represent two different species, but further
-analyses should be conducted
-A visual comparison of the Yao2020 tree and the original Gottlieb2005 tree suggests
-that the relationships within the genus \textit{Ilex} are still being actively
-determined, and that increased taxon sampling might be key to resolve them.
+processes with ASTRAL \cite{mirarab2014astral},
+BEAST2 \cite{bouckaert2019beast}, or SVD Quartets \cite{chifman2014quartet}).
+Rigorous analyses of multiple loci allows for more complex evolutionary models
+than analyses of large genomic data sets, and can provide better evolutionary estimates.
 
@@
 geographical locations for taxa from the GBIF \cite{gbif_secretariat_gbif_2019}.
-Taxonomic links, and comparisons to existing published phylogenies in the OpenTree
-data store can also help flag paralogous sequences.
-Accidentally including paralogs as homologs is a risk in phylogenetic analyses,
-and can be more prevalent in automated analyses than in manually curated analyses.
-We provide users with several tools to try to assess homology of their aligned sequences.
-The estimated gene tree itself is an evolutionarily explicit way to visualize gene
-evolution, which in concert with taxonomic labelling can reveal paralogy.
-Our conflict analysis tool informs the users of whether their tree contains major
-conflicts with established taxonomy.
-This tool also returns information on taxonomic conflicts that exist
-in the original input tree.
-Detected conflicts may be a sign that taxonomy needs to be updated, or may be a sign
-that non-homologous sequences have been included in the analysis.
-These taxonomic and phylogenetic conflicts flag regions of the tree for the researcher
-to more closely examine and assess homology.
 
@@
 potential to consistently provide context for these ideas, highlighting the
-importance of data sharing and open science in phylogenetics, biology and science.
+importance of data sharing and open science in the field, biology and science.
 
@@
 
-Code and datasets generated and analysed during the current study are available in
+The datasets generated and analysed during the current study are available in
 the repositories "physcraper" containing the source code,
@@
 All authors contributed to the manuscript and gave final approval for publication.
+\ldots
 
@@
 and Engineering Building 1, UCM, joint lab paper discussion group for valuable
-comments on this manuscript. We also thank the valuable comments of David Posada, Rutger Vos,
-and an anonymous reviewer that greatly improved an earlier version of this manuscript.
+comments on this manuscript.
 
@@
   \caption{A) Phylogeny updated with Physcraper from original
-  \cite{gottlieb2005molecular} tree.
+  \cite{gottlieb2005molecular} tree in B.
   Tips in original alignment and new tips added with Physcraper are depicted in
-  black and golden, respectively. Physcraper obtained sequences from the GenBank
+  black and red, respectively. Physcraper obtained sequences from the GenBank
   database via local BLAST of all sequences in the original alignment that generated
@@
   filtering", aligned them to original alignment using MUSCLE and performed a
-  phylogenetic reconstruction using RAxML with 100 bootstraps.}
+  phylogenetic reconstruction using RAxML with 100 bootstraps. B-D conflict analyses
+  performed with OpenTree tools.}
 \end{figure}

### 2021-04-29 sources/bmc_template_submission/bmc_article_submission_2021-01-08.tex
fixing quotes and some citations

@@
 
-Assembling a DNA alignment from such a massive database as GenBank can be done "by hand",
+Assembling a DNA alignment from such a massive database as GenBank can be done ``by hand",
 but that is a time consuming approach which is not highly reproducible.
@@
 to alignments
-from TreeBASE \cite{sanderson1994treebase}, and data from GenBank \cite{benson2000genbank}.
+from TreeBASE, and data from GenBank.
 Its utility and functionalities are presented with a case-study analysis of a
@@
 sequence search on a remote or local GenBank database. It is constrained to a
-"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
+ ``search taxon", a taxonomic group in the NCBI taxonomy that is automatically
 identified using the OpenTree's taxonomic Most Recent Common Ancestor (MRCA) API
@@
 
-BLAST is performed using the `blastn` algorithm \cite{camacho2009blast}
+BLAST is performed using the blastn algorithm \cite{camacho2009blast}
 implemented in BioPython's \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
@@
 
-\par{\textbf{Project name:} Physcraper}
+\textbf{Project name:} Physcraper
 
@@
 
-\par{\textbf{OpenTree:} The Open Tree of Life project}
+\textbf{OpenTree:} The Open Tree of Life project
 
@@
 The datasets generated and analysed during the current study are available in
-the repositories "physcraper" containing the source code,
+the repositories ``physcraper" containing the source code,
 \href{https://github.com/McTavishLab/physcraper}{https://github.com/McTavishLab/physcraper};
-"physcraperex" containing the examples,
+ ``physcraperex" containing the examples,
 \href{https://github.com/McTavishLab/physcraperex}{https://github.com/McTavishLab/physcraperex};
-and, "physcraper\_ms" containing this reproducible manuscript,
+and, ``physcraper\_ms" containing this reproducible manuscript,
 \href{https://github.com/McTavishLab/physcraper\_ms}{https://github.com/McTavishLab/physcraper\_ms}.
@@
 \section*{Funding}
-This research was supported by the grant "Sustaining the Open Tree of Life", NSF ABI
+This research was supported by the grant ``Sustaining the Open Tree of Life", NSF ABI
 No. 1759838, and ABI No. 1759846.
@@
 All authors contributed to the manuscript and gave final approval for publication.
-\ldots
 
 \section*{Acknowledgements}
-We thank the members of the OpenTree development team and the "short bar" Science
+We thank the members of the OpenTree development team and the ``short bar" Science
 and Engineering Building 1, UCM, joint lab paper discussion group for valuable
@@
   database via local BLAST of all sequences in the original alignment that generated
-  tree in B), filtered them following criteria from section "DNA sequence search and
+  tree in B), filtered them following criteria from section ``DNA sequence search and
   filtering", aligned them to original alignment using MUSCLE and performed a

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
fixing a curly brace that was causing trouble
various final edits

@@
 Updating existing phylogenies with publicly available molecular sequence data provides
-the opportunity to study
+the opportunity to simultaneously study
 the evolutionary history of many taxa in a reproducible and continuous manner.
 Here, we introduce Physcraper, a tool that establishes a data interoperability
-framework for biological databases to automate data connections across databases.
+framework for biological databases to automate data connections across biological databases.
 Physcraper's main goal is to build upon published alignments and published phylogenies
@@
 Physcraper updates a starting single locus alignment and corresponding tree
-with public DNA data, and links the tips in the updated trees to a unified, interoperable
+with public DNA data from GenBank \cite{benson2000genbank}, and links the tips in the updated trees to a unified, interoperable
 taxonomic resource \cite{rees2017automated}.
@@
 geographical location, fossil ranges, and genetic and phenotypic information increasingly
-available in public databases constitute an amazing
-resource for biological discovery \cite{baxevanis2015importance}.
+available in public biological databases constitute an amazing
+resource for scientific discovery \cite{baxevanis2015importance}.
 One of the main challenges for automatic
-integration of biological data across databases are varying taxonomic idiosyncrasies.
+integration of data across biological databases are varying taxonomic idiosyncrasies.
 To address this challenge, the Open Tree of Life project (OpenTree)
-created a unified taxonomy for name standardization, by integrating taxonomic
+created a unified taxonomy for automatic taxonomic name standardization, by integrating taxonomic
 data from several databases \cite{rees2017automated}, including the USA National
 Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
-and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019} among others.
-By using the existing OpenTree taxonomy programmatic tools to map tip names,
-Physcraper has a framework for connecting updated phylogenies with data from any
-biological database.
+and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
+All of OpenTree tools and methods are carried out via OpenTree's Application Programming Interfaces (APIs),
+that are open as services for use by the general public \cite{opentreeAPIs}.
+Physcraper leverages on existing OpenTree's taxonomic APIs to
+standardize taxon names in any phylogeny, providing
+a framework for automatic connection of updated phylogenies with data from any biological database.
 
@@
 are still not available \cite{mctavish_how_2017}.
-The Open Tree of Life project assembles a comprehensive synthetic tree of life comprising
+Using its unified taxonomy, OpenTree assembles a comprehensive synthetic tree of life comprising
 2.3 million tips, of which around only
-90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem) by curators - the remaining 1.4 million taxa are placed
+90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem
+\cite{mctavish2015phylesystem}) by curators - the remaining 1.4 million taxa are placed
 in the tree based on taxonomy.
@@
 
-Assembling a DNA alignment from such a massive database as GenBank can be done "by hand",
-but that is a time consuming approach which is not highly reproducible.
-A variety of computational pipelines that mine DNA databases fast, efficiently, and reproducibly
-have been developed and used to infer phylogenetic relationships in a variety of organisms
+Assembling a DNA alignment from a massive database such as GenBank can be done ``by hand",
+but that is a time-consuming approach which is largely not reproducible.
+Various computational pipelines that mine DNA databases fast, efficiently, and reproducibly
+have been developed and used to infer phylogenetic relationships of many organisms
 (e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper, pearse2013phylogenerator}).
 While genomics has, and will continue to, revolutionize phylogenetic inference,
-the diversity of alternative genomic sequencing approaches implemented produce
-largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
+the diversity of alternative genomic sequencing approaches that are implemented produce
+widely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
 creating challenges for phylogenetic reconstruction.
@@
 
-Physcraper improves on previous work to automate phylogenetic reconstruction by leveraging
-the power of existing homology hypotheses that phylogeneticists and taxon specialists have assessed
+Physcraper improves on previous work that automates phylogenetic reconstruction, by leveraging
+on the knowledge contained in existing homology hypotheses that phylogeneticists and taxon specialists have assessed
 and deemed appropriate for a specific phylogenetic scope.
@@
 around 100,000 distinct taxa in the TreeBASE database
-\cite{piel2009treebase, vos2012nexml, piel2018treebasedmp, treebase_website, supertreebase},
+\cite{piel2009treebase, vos2012nexml, piel2018treebasedmp},
 which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
 for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
-update and compare existing phylogenetic knowledge.
+update and compare new phylogenetic hypotheses to existing knowledge.
 
-Physcraper is implemented as a Python pipeline that uses OpenTree's programmatic access
-protocols (API's) to automatically
-link any phylogeny mapped to OpenTree's standardized taxonomy \cite{mctavish2015phylesystem},
-to alignments
-from TreeBASE, and data from GenBank \cite{benson2000genbank}.
-Its utility and functionalities are presented with a case-study analysis of a
+Physcraper is implemented as a Python pipeline that uses OpenTree's APIs to automatically
+link any phylogeny mapped to OpenTree's standardized taxonomy,
+to alignments from TreeBASE, and data from GenBank.
+Physcraper's usage and functionalities are presented with a case-study analysis of a
 group of flowering plants, the hollies.
@@
 Physcraper is implemented with Python and can be run on a Python interactive session,
-as a Python script or using the command line interface developed for it.
+as a Python script, or using the command line interface we developed for it.
 It currently consists of 13 modules.
@@
 
-% Pylint score for each module and locally disabled Pylint reccomendations
+% Pylint score for each module and locally disabled Pylint reccommendations
 % physcraper/aligntreetax.py		10/10	pylint disabling “too many branches” and “too many locals”
@@
 
-Taxon names in the input tree must be standardized to OpenTree taxonomy
-\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
+Taxon names in the input tree must be standardized to the OpenTree taxonomy
+\cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
 (TNRS) tool \cite{TNRStool}. Users can upload their
-own tree, or choose from among the 2, 950 curated trees stored in OpenTree's
-Phylesystem database \cite{phylesystemGithub, mctavish2015phylesystem} that also have
-alignments available on TreeBASE.
+own phylogeny, or choose from among the 2, 950 curated phylogenies stored in OpenTree's
+Phylesystem database \cite{phylesystemGithub} that also have
+alignments available on the TreeBASE database \cite{treebase_website, supertreebase}.
 
 The input alignment is a single locus DNA dataset that was used in part or in
-whole to generate the input tree. Physcraper retrieves TreeBASE alignments
-automatically. Alternatively, users must provide the path to a local copy of the
-alignment.
-Only taxa that are both in the sequence alignment and in the tree are considered
+whole to generate the input phylogeny. Physcraper retrieves TreeBASE alignments
+automatically. Alternatively, users can provide the path to a local copy of the
+alignment of their choosing.
+Only taxa that are both in the sequence alignment and in the phylogeny are considered
 further for analysis; at least one taxon and its corresponding sequence are required.
@@
 sequence search on a remote or local GenBank database. It is constrained to a
-"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
-identified using the OpenTree's taxonomic Most Recent Common Ancestor (MRCA) API
-\cite{mrcaAPI, rees2017automated}, as the MRCA of all ingroup taxa that is
-also a named clade in the NCBI taxonomy (Figure 1).
+ ``search taxon", which corresponds to the Most Recent Common Ancestor (MRCA) of
+all ingroup taxa that is also a named clade in the NCBI taxonomy (Figure 1). The
+search taxon is identified using OpenTree's taxonomic API \cite{mrcaAPI}.
 
-BLAST is performed using the `blastn` algorithm \cite{camacho2009blast}
+BLAST is performed using the blastn algorithm \cite{camacho2009blast}
 implemented in BioPython 1.71 \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
@@
 New sequences are excluded for analysis if they 1) are not in the search taxon;
-2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
-and max length threshold, defined as the proportion of the average length without
-gaps of all sequences in input alignment (default values of 80\% and 120\%,
-respectively); 4) or if they are either identical to or shorter than an existing
-sequence in the input alignment and they represent the same taxon in OpenTree or
+2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a minimum
+and maximum sequence length threshold, defined as a proportion of the average sequence length without
+gaps of all sequences in the input alignment (default values of 80\% and 120\%,
+respectively); 4) or, if they are either identical to or shorter than an existing
+sequence in the input alignment and they represent the same taxon in OpenTree's or
 NCBI taxonomy.
@@
 A user is interested in phylogenetic relationships within the genus \textit{Ilex}. Commonly
-known as ``hollies", the genus encompasses between 400-700 living species, and is
-the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering
-plants.
+known as ``hollies", the genus encompasses between 400 \cite{gottlieb2005molecular}
+and 500 accepted living species \cite{plantlist}, and is
+the only extant taxon within the family Aquifoliaceae, order Aquifoliales of flowering
+plants \cite{plantlist}.
 
@@
 the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
-selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
-\cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
-(study 1091 \cite{treebase1091}). This tree (Gottlieb2005 tree from now on) samples 41 species, is available from
+selbach2009new}, but only two have data publicly available, the ``Gottlieb2005" study
+\cite{gottlieb2005molecular} and the ``Yao2020" study \cite{yao2020phylogeny}.
+Gottlieb2005 made their original published phylogeny and alignment openly available in TreeBASE
+(study 1091 \cite{treebase1091}). The Gottlieb2005 phylogeny samples 41 species, is available from
 OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
 and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
-The most recent \textit{Ilex} tree \cite{yao2020phylogeny} (Yao2020 tree from now on) is available in
-OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
+The Yao2020 \textit{Ilex} phylogeny is the most recent one for the genus \cite{yao2020phylogeny},
+and it is only available in OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
 and in the DRYAD repository \cite{yao2020dryad}.
-With 175 tips, the Yao2020 tree \cite{yao2020phylogeny} is the best sampled phylogeny available for
+With 175 tips, the Yao2020 phylogeny \cite{yao2020phylogeny} is the best sampled phylogeny available for
 the genus \textit{Ilex}.
-In order to showcase Physcraper's performance, we updated
-the tree with the smallest number of tips (the Gottlieb2005 tree) with the expectation
-to use the tree with the largest number of tips (the Yao2020 tree) as a standard
-for comparison and verification of Physcraper's results.
+In order to showcase Physcraper's performance, we used the Gottlieb2005 phylogeny
+and alignment to update relationships in the genus \textot{Ilex},
+with the expectation to use the Yao2020 phylogeny, the best sampled and most recent
+one, as a ``gold" standard to compare and verify Physcraper's results.
 
@@
 \textbf{GBIF:} Global Biodiversity Information Facility
+
+\textbf{API:} Application Programming Interface
 
@@
 Code and datasets generated and analysed during the current study are available in
-the repositories "physcraper" containing the source code,
+the repositories ``physcraper" containing the source code,
 \href{https://github.com/McTavishLab/physcraper}{https://github.com/McTavishLab/physcraper};
-"physcraperex" containing the examples,
+ ``physcraperex" containing the examples,
 \href{https://github.com/McTavishLab/physcraperex}{https://github.com/McTavishLab/physcraperex};
-and, "physcraper\_ms" containing this reproducible manuscript,
+and, ``physcraper\_ms" containing this reproducible manuscript,
 \href{https://github.com/McTavishLab/physcraper\_ms}{https://github.com/McTavishLab/physcraper\_ms}.
@@
 \section*{Acknowledgements}
-We thank the members of the OpenTree development team and the "short bar" Science
+We thank the members of the OpenTree development team and the ``short bar" Science
 and Engineering Building 1, UCM, joint lab paper discussion group for valuable
@@
 \begin{figure}[h!]
-  \caption{A) Phylogeny updated with Physcraper using a starting phylogeny and alignment from
+  \caption{A) Phylogeny updated with Physcraper using a starting phylogeny and an alignment from
   \cite{gottlieb2005molecular} (Gottlieb2005 data in text).
@@
   black and red, respectively. Physcraper obtained sequences from the GenBank
-  database via local BLAST of all sequences in the Gottlieb2005 internal transcribed
-  spacer DNA region (ITS) alignment,
-  filtered them following criteria described in section "DNA sequence search and
+  database via local BLAST of all sequences from the Gottlieb2005 internal transcribed
+  spacer DNA region (ITS) original alignment,
+  filtered them following criteria described in section ``DNA sequence search and
   filtering", aligned them to the original alignment using MUSCLE, and performed a
@@
   Gottlieb2005 phylogeny.
+}
 \end{figure}

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
various final edits

@@
 and 500 accepted living species \cite{plantlist}, and is
-the only extant taxon within the family Aquifoliaceae, order Aquifoliales of flowering
-plants \cite{plantlist}.
+the only extant taxon within the family Aquifoliaceae, in the order Aquifoliales of flowering
+plants \cite{chase2016update}.
 
@@
 the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
-selbach2009new}, but only two have data publicly available, the ``Gottlieb2005" study
+selbach2009new}, but only two have data openly available, the ``Gottlieb2005" study
 \cite{gottlieb2005molecular} and the ``Yao2020" study \cite{yao2020phylogeny}.
@@
 In order to showcase Physcraper's performance, we used the Gottlieb2005 phylogeny
-and alignment to update relationships in the genus \textot{Ilex},
+and alignment to update relationships in the genus \textit{Ilex},
 with the expectation to use the Yao2020 phylogeny, the best sampled and most recent
@@
 more complex evolutionary models
-than analyses of large genomic data sets, and can provide better evolutionary estimates.
+than analyses of large genomic data sets, and in many cases can provide better evolutionary estimates.
 For example \cite{zhu_complexity_2021} show that when applying coalescent models,
 there is more information in two genes of 300 bp each than in 600 independent sites.
-Physcraper is desighed to facilitate
+Physcraper is designed to facilitate
 gathering alignments and gene trees for multiple loci from a group of interest,
@@
 processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}.
-We have included a script (``multi\_locus.py'') and examples of how to automatically
-merge the outputs of Physcraper runs at different loci into input files for these
-two software, or as concatenated alignments for supermatrix analyses.
+Physcraper's ``multi\_locus.py'' module allows to automatically
+merge the outputs of Physcraper runs from different loci into input files for the
+two software mentioned above, or as concatenated alignments for supermatrix analyses.
 
 %% Biological plausibility of example??
-Our example application of Physcraper to update a holly phylogeny is based on a
-single marker, so we expect for it to be not as well resolved as trees from studies
-using multiple markers. Although not perfect,
-we think the updated phylogeny seems biologically reasonable in different ways. All samples
-corresponding to the ingroup cluster together forming a monophyletic group, and
+Our example application of Physcraper to update a phylogeny of the genus \textit{Ilex} is based on a
+single marker, so we expect for it to be not as well resolved as phylogenies resulting from analyses
+that used multiple markers. Although not perfect,
+we think the Physcraper updated \textit{Ilex} phylogeny seems biologically reasonable
+in different ways. All samples
+corresponding to the ingroup cluster together forming a monophyletic group (Figure 2A), and
 samples belonging to the same \textit{Ilex} species also form
-monophyletic groups. A notable exception is \textit{Ilex theeizans}, which appears as
-non-monophyletic in the Physcraper updated phylogeny as well as in the original Gottlieb2005 tree.
+monophyletic groups (Figure 2B). A notable exception is samples of the species \textit{Ilex theeizans}, which appear as
+non-monophyletic in the updated phylogeny as well as in the original Gottlieb2005 phylogeny.
 %This suggests that the samples might represent two different species, but further
 analyses should be conducted
-A visual comparison of the Yao2020 tree and the original Gottlieb2005 tree suggests
+A visual comparison of the Yao2020 phylogeny and the original Gottlieb2005 phylogeny suggests
 that the relationships within the genus \textit{Ilex} are still being actively
@@
 for the advantage of all areas of biology as well as the advancement of scientific
-policy and applications.
+policy, applications and education.
 Initial ideas about the data are constantly changed by results from new analyses.
@@
 potential to consistently provide context for these ideas, highlighting the
-importance of data sharing and open science in phylogenetics, biology and science.
+importance of data sharing and open science for phylogenetics, biology and science.
 

### 2021-04-29 sources/bmc_template_submission/bmc_article_submission_2021-01-08.tex
various final edits

@@
 the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
-selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
+selbach2009new}, but only two have data openly available \cite{gottlieb2005molecular, yao2020phylogeny}.
 \cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
homogenizing 'phylogeny' and 'tree' use

@@
 \kwd{gene tree}
+\kwd{gene phylogeny}
+\kwd{multilocus}
 \kwd{interoperability}
@@
 as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
-Physcraper updates a starting single locus alignment and corresponding tree
+Physcraper updates a starting single locus alignment and corresponding phylogeny
 with public DNA data from GenBank \cite{benson2000genbank}, and links the tips in the updated trees to a unified, interoperable
@@
 The general Physcraper framework (Figure 1) consists of 4 steps: 1) identifying and
-processing a tree and its underlying alignment; 2) performing a BLAST search of
+processing a phylogeny and its underlying alignment; 2) performing a BLAST search of
 DNA sequences from original alignment on GenBank, and filtering of new sequences;
 3) profile-aligning new sequences to original alignment; 4) performing a phylogenetic
-analysis and comparing the updated tree to existing phylogenies.
+analysis and comparing the updated results to existing phylogenies.
 
-\subsection*{The inputs: a tree and an alignment}
+\subsection*{The inputs: a phylogeny and an alignment}
 
-Taxon names in the input tree must be standardized to the OpenTree taxonomy
+Taxon names in the input phylogeny must be standardized to the OpenTree taxonomy
 \cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
@@
 
-\subsection*{Tree reconstruction and comparison}
+\subsection*{Phylogenetic reconstruction and comparison}
 
 RAxML \cite{stamatakis2014raxml} is implemented to reconstruct a Maximum Likelihood
-(ML) gene tree for each input alignment with default settings (GTRCAT model and
-100 bootstrap replicates with default algorithm), using input tree as starting
+(ML) gene phylogeny for each input alignment with default settings (GTRCAT model and
+100 bootstrap replicates with default algorithm), using the input phylogeny as starting
 tree for ML searches.
@@
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original tree are compared with Robinson-Foulds weighted and unweighted
+Updated and original phylpogeny are compared with Robinson-Foulds weighted and unweighted
 metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
-comparison between the synthetic OpenTree and original and updated tree individually,
+comparison between the synthetic OpenTree and the original and updated phylogenies individually,
 using OpenTree's conflict API \cite{redelings2017supertree}.
@@
 We ran Physcraper on a laptop Linux computer to update an internal transcribed
-spacer DNA region (ITS) alignment that was used to construct the tree from \cite{gottlieb2005molecular},
+spacer DNA region (ITS) alignment that was used to construct the phylogeny from \cite{gottlieb2005molecular},
 using a local GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
@@
 %% <!-- MTH: you probably need some details about the hardware, given the fact that you are discussing running times -->
-The Gottlieb2005 tree \cite{gottlieb2005molecular} updated using Physcraper (Figure 2;
-Physcraper updated tree from now on) displays all 41 distinct
+The Gottlieb2005 phylogeny \cite{gottlieb2005molecular} updated using Physcraper (Figure 2;
+Physcraper updated phylogeny from now on) displays all 41 distinct
 taxa from the original study plus 231 new tips, contributing phylogenetic data to
-84 additional \textit{Ilex} taxa. The best ML tree from the RAxML analysis is 99\%
+84 additional \textit{Ilex} taxa. The best ML phylogeny from the RAxML analysis is 99\%
 resolved, with 25\% of nodes with bootstrap support $<$ 0.1 and 48\% nodes with
@@
 (Figure 2).
-As comparison with the Physcraper updated tree, the Yao2020 tree \cite{yao2020phylogeny}
+As comparison with the Physcraper updated phylogeny, the Yao2020 phylogeny \cite{yao2020phylogeny}
 also contains all 41 distinct taxa
-sampled in the Gottlieb2005 tree \cite{gottlieb2005molecular},
+sampled in the Gottlieb2005 phylogeny \cite{gottlieb2005molecular},
 while contributing phylogenetic data to 134 additional \textit{Ilex} taxa. From these,
-67 taxa are also in the Physcraper updated tree.
-While the Yao2020 tree \cite{yao2020phylogeny} was also constructed using
+67 taxa are also in the Physcraper updated phylogeny.
+While the Yao2020 phylogeny \cite{yao2020phylogeny} was also constructed using
 ITS as a marker, their GenBank data is not released yet. Hence, Physcraper was unable
-to incorporate 68 taxa that are only on the Yao2020 tree because the DNA data is unavailable.
+to incorporate 68 taxa that are only on the Yao2020 phylogeny because the DNA data is unavailable.
 We also note that Physcraper incorporates 18 \textit{Ilex} taxa that are not in the Yao2020
-tree \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex} species
+phylogeny \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex} species
 and visual inspection of the DNA sequences suggests they are correctly assigned as \textit{Ilex}.
-The ITS alignment that underlies the Yao2020 tree was constructed without any tool to
+The ITS alignment that underlies the Yao2020 phylogeny was constructed without any tool to
 scrape GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
-incorporate these 18 additional \textit{Ilex} taxa in the updated tree (Figure 2).
+incorporate these 18 additional \textit{Ilex} taxa in the Physcraper updated phylogeny (Figure 2).
 
@@
 To test the accuracy of Physcraper we designed a verification test.
-We pruned 9 out of the 41 original tips of the Gottlieb2005 tree \cite{gottlieb2005molecular},
+We pruned 9 out of the 41 original tips of the Gottlieb2005 phylogeny \cite{gottlieb2005molecular},
 corresponding to a 20\% trim, excluding the outgroups. We then performed a Physcraper
 run to test if we would recover the dropped tips.
-We successfully recovered 6 out of 9 pruned tips in the updated tree.
+We successfully recovered 6 out of 9 pruned tips in the Physcraper updated phylogeny.
 Closer examination of results revealed that sequences for the 3 missing tips were correctly retrieved
-with BLAST along with sequences from the other 6 tips recovered in the updated tree.
+with BLAST along with sequences from the other 6 tips recovered in the updated phylogeny.
 By following the GenBank accession numbers reported in the original publication
@@
 Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
-which add new taxa without modifying the input tree, Physcraper estimates all
+which add new taxa without modifying the input phylogeny, Physcraper estimates all
 the relationships anew in the context of the new data.
@@
 We provide users with several tools to try to assess homology of their aligned sequences.
-The estimated gene tree itself is an evolutionarily explicit way to visualize gene
+The estimated gene phylogeny itself is an evolutionarily explicit way to visualize gene
 evolution, which in concert with taxonomic labelling can reveal paralogy.
-OpenTree's conflict analysis tool informs the users of whether their tree contains major
+OpenTree's conflict analysis tool informs the users of whether their phylogeny contains major
 conflicts with established taxonomy and any phylogenetic context they wish to compare to.
 This tool also returns information on taxonomic and phylogenetic conflicts that exist
-in the original input tree.
+in the original input phylogeny.
 Detected conflicts may be a sign that taxonomy needs to be updated, or may be a sign
 that non-homologous sequences have been included in the analysis.
-These taxonomic and phylogenetic conflicts flag regions of the tree for the researcher
+These taxonomic and phylogenetic conflicts flag regions of the phylogeny for the researcher
 to more closely examine and assess homology.

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
verification test missing taxa genbank accession numbers

@@
 
-To test the accuracy of Physcraper we designed a verification test.
-We pruned 9 out of the 41 original tips of the Gottlieb2005 phylogeny \cite{gottlieb2005molecular},
+To test the accuracy of Physcraper we performed a verification test in which
+we pruned 9 out of the 41 tips in the original Gottlieb2005 phylogeny \cite{gottlieb2005molecular},
 corresponding to a 20\% trim, excluding the outgroups. We then performed a Physcraper
-run to test if we would recover the dropped tips.
-We successfully recovered 6 out of 9 pruned tips in the Physcraper updated phylogeny.
+run to test if we would recover the pruned tips. Physcraper automatically dropped
+from the alignment the taxa pruned from the phylogeny to perform the BLAST searches.
+The Physcraper updated phylogeny successfully recovered only 6 out of 9 pruned tips.
 Closer examination of results revealed that sequences for the 3 missing tips were correctly retrieved
-with BLAST along with sequences from the other 6 tips recovered in the updated phylogeny.
-By following the GenBank accession numbers reported in the original publication
-belonging to the ITS sequences of the missing tips, we observed that these three sequences
+with BLAST along with the 6 sequences belonging to the remaining pruned tips.
+We then followed the GenBank accession numbers reported in \cite{gottlieb2005molecular}
+belonging to the sequences of the 3 missing tips, \textit{Ilex warburgii} (accession
+number reported: U92600/U92601; updated: AH007153.2), \textit{Ilex dimorphophylla}
+(accession number reported: U92592/U92593; updated: AH007149.2),
+and \textit{Ilex percoriacea} (accession number: AH007156.2). We note that these three sequences
 contain a 100 bp long gap of unidentified nucleotides (Ns) that is completely absent
-from any of the sequences in the original alignment. This caused these three GenBank sequences
-in particular to exceed Physcraper's default sequence length cutoff of 120\%, being thus
-filtered and excluded from the alignment step onwards.
-These sequences do appear in the Physcraper results in the list of matches from
-GenBank which did not fit the sequence length cutoffs set in the configuration file.
-This ``seqlen\_mismatch.txt" file includes the accession number, taxon, and sequence
-length of all sequences filtered based on sequence length.
+from the original alignment. This caused them to exceed Physcraper's default sequence
+length cutoff of 120\%, being thus filtered and excluded from further analyses.
+The missing ITS sequences do appear in Physcraper's output file ``seqlen\_mismatch.txt",
+that includes the accession number, taxon name, and sequence length of all BLAST matches
+that were filtered based on sequence length cutoffs set in the configuration file.
 

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
some more edits

@@
 known as ``hollies", the genus encompasses between 400 \cite{gottlieb2005molecular}
-and 500 accepted living species \cite{plantlist}, and is
+and 500 recognized living species \cite{plantlist}, and is
 the only extant taxon within the family Aquifoliaceae, in the order Aquifoliales of flowering
@@
 An online literature review in June 2020 (Google scholar search for ``ilex phylogeny")
-reveals that there are several published phylogenies showing relationships within
+revealed that there are several published phylogenies showing relationships within
 the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
@@
 the genus \textit{Ilex}.
-In order to showcase Physcraper's performance, we used the Gottlieb2005 phylogeny
-and alignment to update relationships in the genus \textit{Ilex},
-with the expectation to use the Yao2020 phylogeny, the best sampled and most recent
-one, as a ``gold" standard to compare and verify Physcraper's results.
+In order to showcase Physcraper's performance, we chose the Gottlieb2005 phylogeny
+and alignment as Physcraper inputs to update relationships in the genus \textit{Ilex}.
+Currently being the best sampled and most recent phylogenetic hypothesis for the genus,
+we used the Yao2020 phylogeny as a golden standard to compare our results and verify
+Physcraper's performance.
 
@@
 corresponding to a 20\% trim, excluding the outgroups. We then performed a Physcraper
-run to test if we would recover the pruned tips. Physcraper automatically dropped
-from the alignment the taxa pruned from the phylogeny to perform the BLAST searches.
+run to test if we would recover the pruned tips. To perform the BLAST searches,
+Physcraper automatically drops sequences from the alignment belonging to taxa
+absent from the input phylogeny.
 The Physcraper updated phylogeny successfully recovered only 6 out of 9 pruned tips.

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
cluster instead of laptop

@@
 An online literature review in June 2020 (Google scholar search for ``ilex phylogeny")
-revealed that there are several published phylogenies showing relationships within
-the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
-selbach2009new}, but only two have data openly available, the ``Gottlieb2005" study
-\cite{gottlieb2005molecular} and the ``Yao2020" study \cite{yao2020phylogeny}.
-Gottlieb2005 made their original published phylogeny and alignment openly available in TreeBASE
-(study 1091 \cite{treebase1091}). The Gottlieb2005 phylogeny samples 41 species, is available from
+revealed that there are several published studies addressing phylogenetic relationships within
+the hollies \cite{cuenoud2000molecular, gottlieb2005molecular, manen2010history,
+setoguchi2000intersectional, selbach2009new, yao2020phylogeny}, but only the ``Gottlieb2005"
+study \cite{gottlieb2005molecular} and the ``Yao2020" study \cite{yao2020phylogeny}
+have data openly available.
+The Gottlieb2005 phylogeny and alignment are available in TreeBASE study 1091 \cite{treebase1091}.
+The Gottlieb2005 phylogeny samples 41 species, is available in
 OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
@@
 In order to showcase Physcraper's performance, we chose the Gottlieb2005 phylogeny
-and alignment as Physcraper inputs to update relationships in the genus \textit{Ilex}.
-Currently being the best sampled and most recent phylogenetic hypothesis for the genus,
-we used the Yao2020 phylogeny as a golden standard to compare our results and verify
-Physcraper's performance.
+and a corresponding single locus alignment of the internal transcribed
+spacer DNA region (ITS) as Physcraper inputs, to update relationships in the genus \textit{Ilex}.
+Currently being the best sampled and most recent phylogenetic hypothesis for \textit{Ilex},
+we used the Yao2020 phylogeny as the ideal standard to compare results from this
+Physcraper case study.
 
-We ran Physcraper on a laptop Linux computer to update an internal transcribed
-spacer DNA region (ITS) alignment that was used to construct the phylogeny from \cite{gottlieb2005molecular},
-using a local GenBank database.  BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
+We ran Physcraper on a computer node of the Multi-Environment Research Computer for
+Exploration and Discovery (MERCED) Linux cluster from the University of California,
+Merced, using one core at 2301 MHz RAM. BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap
 analyses taking an additional 13hrs.
@@
 resolved, with 25\% of nodes with bootstrap support $<$ 0.1 and 48\% nodes with
-bootstrap support $>$ 0.75.
+bootstrap support $>$ 0.75 (figure 2A).
 A large portion of internal branches are negligibly small, with 30 branches
@@
 We also note that Physcraper incorporates 18 \textit{Ilex} taxa that are not in the Yao2020
-phylogeny \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex} species
-and visual inspection of the DNA sequences suggests they are correctly assigned as \textit{Ilex}.
+phylogeny \cite{yao2020phylogeny}. These taxa appear nested among other \textit{Ilex}
+species (Figure 2B) and visual inspection of the DNA sequences suggests they are
+correctly assigned as \textit{Ilex}.
 The ITS alignment that underlies the Yao2020 phylogeny was constructed without any tool to
-scrape GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
-incorporate these 18 additional \textit{Ilex} taxa in the Physcraper updated phylogeny (Figure 2).
+mine GenBank \cite{yao2020phylogeny}, which could explain why Physcraper was able to
+incorporate these 18 additional \textit{Ilex} taxa in the Physcraper updated phylogeny.
 
-%% Maybe show conflict results? and then discuss in the discussion.
+%% Maybe show conflict results? and then add to the discussion. YES
 
@@
 
-To test the accuracy of Physcraper we performed a verification test in which
+To test the performance of Physcraper we performed a verification test in which
 we pruned 9 out of the 41 tips in the original Gottlieb2005 phylogeny \cite{gottlieb2005molecular},

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
cluster instead of laptop

@@
 We ran Physcraper on a computer node of the Multi-Environment Research Computer for
-Exploration and Discovery (MERCED) Linux cluster from the University of California,
+Exploration and Discovery (MERCED) cluster from the University of California,
 Merced, using one core at 2301 MHz RAM. BLAST and RAxML analyses ran for 19hrs 45min, with bootstrap

### 2021-04-29 sources/bmc_template_submission/bmc_article.tex
tiny edits

--- 2021-04-29 sources/bmc_template_submission/bmc_article.tex
@@
 and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
-All of OpenTree tools and methods are carried out via OpenTree's Application Programming Interfaces (APIs),
-that are open as services for use by the general public \cite{opentreeAPIs}.
+All of OpenTree tools and methods are implemented via OpenTree's Application Programming Interfaces (APIs),
+as open programmatic access services for use by the general public \cite{opentreeAPIs}.
 Physcraper leverages on existing OpenTree's taxonomic APIs to
@@
 %% Biological plausibility of example??
-Our example application of Physcraper to update a phylogeny of the genus \textit{Ilex} is based on a
+Our case study application of Physcraper to update a phylogeny of the genus \textit{Ilex} is based on a
 single marker, so we expect for it to be not as well resolved as phylogenies resulting from analyses
@@
 non-monophyletic in the updated phylogeny as well as in the original Gottlieb2005 phylogeny.
-%This suggests that the samples might represent two different species, but further
-analyses should be conducted
+% This suggests that the samples might represent two different species, but further
+% analyses should be conducted.
 A visual comparison of the Yao2020 phylogeny and the original Gottlieb2005 phylogeny suggests

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last MK edits

@@
 framework for biological databases to automate data connections across biological databases.
-Physcraper's main goal is to build upon published alignments and published phylogenies
-to extend existing phylogenetic inferences with more data and taxa, which improve
-phylogenetic reconstruction \cite{hillis1996inferring, natsidis2019phylogenomics},
+Physcraper's main goal is to build upon published alignments and phylogenies
+to extend existing phylogenetic inferences with more data and taxa, which are known to improve
+phylogenetic reconstructions \cite{hillis1996inferring, natsidis2019phylogenomics},
 time of divergence estimates \cite{schulte2013undersampling, soares2015influence},
@@
 Physcraper updates a starting single locus alignment and corresponding phylogeny
-with public DNA data from GenBank \cite{benson2000genbank}, and links the tips in the updated trees to a unified, interoperable
+with public DNA data from GenBank \cite{benson2000genbank}, and links tip taxon labels in the updated trees to a unified, interoperable
 taxonomic resource \cite{rees2017automated}.
@@
  and manual \cite{pylintManual}, with a ``.pylintrc" configuration file.
-
-We improved code syntax of Physcraper's modules with low Pylint scores, and fixed code errors by
-following Pylint's recommendations.
-Based on Physcraper's software design choices, some of Pylint's recommendations were overruled
-by using its check-disabling system, and are explained along the code.
+% Based on Physcraper's software design choices, some of Pylint's recommendations were overruled
+% by using its check-disabling system, and are explained along the code.
 As of now, all Physcraper modules have a Pylint score of 10/10.
@@
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original phylpogeny are compared with Robinson-Foulds weighted and unweighted
+Updated and original phylogeny are compared with Robinson-Foulds weighted and unweighted
 metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last MK edits

@@
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original phylogeny are compared with Robinson-Foulds weighted and unweighted
+Updated and original phylogeies are compared with Robinson-Foulds weighted and unweighted
 metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
@@
 resolved, with 25\% of nodes with bootstrap support $<$ 0.1 and 48\% nodes with
-bootstrap support $>$ 0.75 (figure 2A).
+bootstrap support $>$ 0.75 (Figure 2A).
 A large portion of internal branches are negligibly small, with 30 branches

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last edits

@@
 framework for biological databases to automate data connections across biological databases.
-Physcraper's main goal is to build upon published alignments and phylogenies
-to extend existing phylogenetic inferences with more data and taxa, which are known to improve
+Physcraper's main goal is to use its interoperability framework to build upon published alignments and phylogenies
+to extend existing phylogenetic inferences with more data and taxa, which improve our
 phylogenetic reconstructions \cite{hillis1996inferring, natsidis2019phylogenomics},
@@
 biogeographic analyses \cite{kayaalp2017back},
-as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
+and help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
 Physcraper updates a starting single locus alignment and corresponding phylogeny
 with public DNA data from GenBank \cite{benson2000genbank}, and links tip taxon labels in the updated trees to a unified, interoperable
-taxonomic resource \cite{rees2017automated}.
+taxonomic resource \cite{rees2017automated}, streamlining connections across biological databases.
 
@@
 All of OpenTree tools and methods are implemented via OpenTree's Application Programming Interfaces (APIs),
-as open programmatic access services for use by the general public \cite{opentreeAPIs}.
+as open access services for use by the general public \cite{opentreeAPIs}.
 Physcraper leverages on existing OpenTree's taxonomic APIs to
@@
 2.3 million tips, of which around only
-90,000 are supported by phylogenetic data uploaded to OpenTrees' database (the Phylesystem
-\cite{mctavish2015phylesystem}) by curators - the remaining 1.4 million taxa are placed
-in the tree based on taxonomy.
+90,000 are supported by publicly available expert phylogenetic data, uploaded to OpenTrees' database (the Phylesystem
+\cite{mctavish2015phylesystem}) by human curators - the remaining 1.4 million taxa are placed
+in the synthetic tree based on taxonomy.
 There is a considerable amount of phylogenetically
@@
 Assembling a DNA alignment from a massive database such as GenBank can be done ``by hand",
-but that is a time-consuming approach which is largely not reproducible.
+but that is a time-consuming approach which is largely non reproducible.
 Various computational pipelines that mine DNA databases fast, efficiently, and reproducibly
@@
 the diversity of alternative genomic sequencing approaches that are implemented produce
-widely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
-creating challenges for phylogenetic reconstruction.
+widely non-overlapping homology hypotheses across taxa,
+creating challenges for phylogenetic reconstruction \cite{jones2016targeted}.
 Phylogenomics addresses this problem by focusing on targeted capture of informative
 regions \cite{andermann2020guide}.
-However, fine-grained curated markers and alignments can significantly improve phylogenetic
+However, carefully curated markers and alignments can significantly improve phylogenetic
 reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
@@
 which can be used as seeds to mine molecular databases, and as ``jump-start'' alignments
-for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
+for phylogenetic reconstruction \cite{morrison2006multiple} to continually enrich,
 update and compare new phylogenetic hypotheses to existing knowledge.

### 2021-05-01 sources/bmc_template_submission/bmc_article_submission_2021-01-08.tex
last edits

@@
 Assembling a DNA alignment from such a massive database as GenBank can be done ``by hand",
-but that is a time consuming approach which is not highly reproducible.
+but that is a time-consuming approach which is not highly reproducible.
 A variety of computational pipelines that mine DNA databases fast, efficiently, and reproducibly
@@
 regions \cite{andermann2020guide}.
-However, fine-grained curated markers and alignments can significantly improve phylogenetic
+However, carefully curated markers and alignments can significantly improve phylogenetic
 reconstructions, even in phylogenomic analyses \cite{fragoso2017pilot}.
@@
 can be leveraged as seeds to mine molecular databases, and as ``jump-start'' alignments
-for phylogenetic reconstructions \cite{morrison2006multiple} to continually enrich,
+for phylogenetic reconstruction \cite{morrison2006multiple} to continually enrich,
 update and compare existing phylogenetic knowledge.

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last edits

@@
 time of divergence estimates \cite{schulte2013undersampling, soares2015influence},
-biogeographic analyses \cite{kayaalp2017back},
+and biogeographic analyses \cite{kayaalp2017back},
 and help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
@@
 with public DNA data from GenBank \cite{benson2000genbank}, and links tip taxon labels in the updated trees to a unified, interoperable
-taxonomic resource \cite{rees2017automated}, streamlining connections across biological databases.
+taxonomic resource \cite{rees2017automated}, streamlining connections between phylogenies and biological data.
+Physcraper links tip taxon labels in the updated trees to a unified, interoperable
+taxonomic resource \cite{rees2017automated}, eefectively streamlining connections between phylogenies and biological data.
 
@@
 All of OpenTree tools and methods are implemented via OpenTree's Application Programming Interfaces (APIs),
-as open access services for use by the general public \cite{opentreeAPIs}.
+as open access services for the general public \cite{opentreeAPIs}.
 Physcraper leverages on existing OpenTree's taxonomic APIs to
@@
 are still not available \cite{mctavish_how_2017}.
-Using its unified taxonomy, OpenTree assembles a comprehensive synthetic tree of life comprising
-2.3 million tips, of which around only
+OpenTree assembles a comprehensive synthetic tree of life comprising
+2.3 million tips, of which around
 90,000 are supported by publicly available expert phylogenetic data, uploaded to OpenTrees' database (the Phylesystem
-\cite{mctavish2015phylesystem}) by human curators - the remaining 1.4 million taxa are placed
-in the synthetic tree based on taxonomy.
+\cite{mctavish2015phylesystem}) by volunteer curators - the remaining 1.4 million taxa are placed
+in the synthetic tree based on OpenTree's unified taxonomy.
 There is a considerable amount of phylogenetically
@@
 the tree of life, but this data either has not been analysed or the analyses
-have not been made publicly available \cite{mctavish_how_2017}.
+have not been made publicly available and accesible \cite{mctavish_how_2017}.
 

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last edits

@@
 created a unified taxonomy for automatic taxonomic name standardization, by integrating taxonomic
-data from several databases \cite{rees2017automated}, including the USA National
+data from several resources \cite{rees2017automated}, including the USA National
 Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
 and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
-All of OpenTree tools and methods are implemented via OpenTree's Application Programming Interfaces (APIs),
-as open access services for the general public \cite{opentreeAPIs}.
-Physcraper leverages on existing OpenTree's taxonomic APIs to
+OpenTree's unfified taxonomy, along with tools and methods to manage it, are available
+as Application Programming Interfaces (APIs),
+which are implemented as open access services for the general public \cite{opentreeAPIs}.
+Physcraper leverages on existing OpenTree's unified taxonomy APIs to automatically
 standardize taxon names in any phylogeny, providing
-a framework for automatic connection of updated phylogenies with data from any biological database.
+a framework for connection of updated phylogenies with data from any biological database.
 
@@
 (e.g., \cite{smith2009mega, antonelli2017toward, izquierdo2014pumper, pearse2013phylogenerator}).
-While genomics has, and will continue to, revolutionize phylogenetic inference,
+While genomics has, and will continue to revolutionize phylogenetic inference,
 the diversity of alternative genomic sequencing approaches that are implemented produce
@@
 for phylogenetic reconstruction \cite{morrison2006multiple} to continually enrich,
-update and compare new phylogenetic hypotheses to existing knowledge.
+update and compare phylogenetic hypotheses to existing evolutionary knowledge.
 
 Physcraper is implemented as a Python pipeline that uses OpenTree's APIs to automatically
-link any phylogeny mapped to OpenTree's standardized taxonomy,
+link any phylogeny mapped to OpenTree's unified taxonomy,
 to alignments from TreeBASE, and data from GenBank.
@@
 
-Taxon names in the input phylogeny must be standardized to the OpenTree taxonomy
+Taxon names in the input phylogeny must be standardized to OpenTree's unfified taxonomy
 \cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
@@
 The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
-sequence search on a remote or local GenBank database. It is constrained to a
+sequence search either on a remote or a local GenBank database. It is constrained to a
  ``search taxon", which corresponds to the Most Recent Common Ancestor (MRCA) of
 all ingroup taxa that is also a named clade in the NCBI taxonomy (Figure 1). The
-search taxon is identified using OpenTree's taxonomic API \cite{mrcaAPI}.
+search taxon is identified using OpenTree's unified taxonomic API \cite{mrcaAPI}.
 
@@
 modified to accept an alternative BLAST address.
-Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
-New sequences are excluded for analysis if they 1) are not in the search taxon;
+Each sequence in the alignment is BLASTed once against the GenBank database.
+Matching sequences are filtered and excluded from the analysis if they 1) are not in the search taxon;
 2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a minimum
@@
 respectively); 4) or, if they are either identical to or shorter than an existing
-sequence in the input alignment and they represent the same taxon in OpenTree's or
-NCBI taxonomy.
-An arbitrary maximum number of randomly chosen sequences per taxon are allowed
+sequence in the input alignment and they represent the same taxon in NCBI's or
+OpenTree's unififed taxonomy.
+By default, an arbitrary maximum number of 5 sequences per taxon are chosen at random
+from the set of matching sequences that passed the filtering step.
 (default to 5).
@@
   \begin{figure}[h!]
-  \caption{The Physcraper framework consists of four general steps. The software is fully
+  \caption{The Physcraper framework consists of four general steps. The star represents
+  the ``search taxon", the Most Recent Common Ancestor (MRCA) of all taxa in the
+  ingroup that is also a named clade in the NCBI taxonomy. The software is fully
   described on its documentation website at \href{https://physcraper.readthedocs.io}{https://physcraper.readthedocs.io},

### 2021-05-01 sources/bmc_template_submission/bmc_article_submission_2021-01-08.tex
last edits

@@
 created a unified taxonomy for name standardization, by integrating taxonomic
-data from several databases \cite{rees2017automated}, including the USA National
+data from several resources \cite{rees2017automated}, including the USA National
 Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
@@
 the diversity of alternative genomic sequencing approaches implemented produce
-largely non-overlapping homology hypotheses across taxa \cite{jones2016targeted},
-creating challenges for phylogenetic reconstruction.
+largely non-overlapping homology hypotheses across taxa,
+creating challenges for phylogenetic reconstruction \cite{jones2016targeted}.
 Phylogenomics addresses this problem by focusing on targeted capture of informative

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
last edits to captions

@@
 framework for biological databases to automate data connections across biological databases.
-Physcraper's main goal is to use its interoperability framework to build upon published alignments and phylogenies
-to extend existing phylogenetic inferences with more data and taxa, which improve our
+Physcraper's main goal is to use its interoperability framework to build upon published
+alignments and phylogenies
+to extend existing phylogenetic inferences with sequence data available in public
+DNA databases for more taxa. Increased taxonomic sampling improves
 phylogenetic reconstructions \cite{hillis1996inferring, natsidis2019phylogenomics},
@@
 and help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
-Physcraper updates a starting single locus alignment and corresponding phylogeny
-with public DNA data from GenBank \cite{benson2000genbank}, and links tip taxon labels in the updated trees to a unified, interoperable
-taxonomic resource \cite{rees2017automated}, streamlining connections between phylogenies and biological data.
+% Physcraper updates the starting single locus alignment and corresponding phylogeny
+% with public DNA data from GenBank \cite{benson2000genbank}, and links tip taxon
+% labels in the updated trees to a unified, interoperable
+% taxonomic resource \cite{rees2017automated}, streamlining connections between phylogenies and biological data.
 Physcraper links tip taxon labels in the updated trees to a unified, interoperable
-taxonomic resource \cite{rees2017automated}, eefectively streamlining connections between phylogenies and biological data.
+taxonomic resource \cite{rees2017automated}, effectively streamlining connections between phylogenies and biological data.
 
@@
 from the set of matching sequences that passed the filtering step.
-(default to 5).
 
@@
 to statistical support and biological plausibility.
+% Add citation of good things about thining alignments
 
@@
   as backbone and resolving branches based on phylogenetic data from the original
-  Gottlieb2005 phylogeny.
+  Gottlieb2005 phylogeny. Green branches are resolved by the updated phylogeny,
+  blue branches are in agreement between the updated phylogeny and the syntehtic subtree, and
+  orange branches are in conflict between the two.
 }

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
fixing 'unified' typos

@@
 and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
-OpenTree's unfified taxonomy, along with tools and methods to manage it, are available
+OpenTree's unified taxonomy, along with tools and methods to manage it, are available
 as Application Programming Interfaces (APIs),
@@
 
-Taxon names in the input phylogeny must be standardized to OpenTree's unfified taxonomy
+Taxon names in the input phylogeny must be standardized to OpenTree's unified taxonomy
 \cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
@@
 sequence in the input alignment and they represent the same taxon in NCBI's or
-OpenTree's unififed taxonomy.
+OpenTree's unified taxonomy.
 By default, an arbitrary maximum number of 5 sequences per taxon are chosen at random

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
tiny edits

@@
 the evolutionary history of many taxa in a reproducible and continuous manner.
-Here, we introduce Physcraper, a tool that establishes a data interoperability
-framework for biological databases to automate data connections across biological databases.
-Physcraper's main goal is to use its interoperability framework to build upon published
-alignments and phylogenies
-to extend existing phylogenetic inferences with sequence data available in public
-DNA databases for more taxa. Increased taxonomic sampling improves
+Increased taxonomic sampling improves
 phylogenetic reconstructions \cite{hillis1996inferring, natsidis2019phylogenomics},
@@
 and help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
+Here, we introduce Physcraper, a Python package that provides a data interoperability
+framework to automate data connections across biological databases,
+with the main goal to build upon published alignments and phylogenies
+to extend existing phylogenetic inferences with sequence data available for more taxa
+in public DNA databases.
 % Physcraper updates the starting single locus alignment and corresponding phylogeny
@@
 Physcraper links tip taxon labels in the updated trees to a unified, interoperable
-taxonomic resource \cite{rees2017automated}, effectively streamlining connections between phylogenies and biological data.
+taxonomic resource \cite{rees2017automated}, effectively streamlining connections
+between phylogenies and any type of biological data.
 
-Data such as
-geographical location, fossil ranges, and genetic and phenotypic information increasingly
+Information such as
+geographical location, fossil ranges, and genetic and phenotypic data increasingly
 available in public biological databases constitute an amazing

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
more tiny edits

@@
 the evolutionary history of many taxa in a reproducible and continuous manner.
-Increased taxonomic sampling improves
+Increased taxonomic sampling is known to improve
 phylogenetic reconstructions \cite{hillis1996inferring, natsidis2019phylogenomics},
@@
 and biogeographic analyses \cite{kayaalp2017back},
-and help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
+as well as help in resolving phylogenetic conflict \cite{hedtke2006resolution, townsend2010optimal, natsidis2019phylogenomics}.
 Here, we introduce Physcraper, a Python package that provides a data interoperability
 framework to automate data connections across biological databases,
-with the main goal to build upon published alignments and phylogenies
+with the main goal of building upon published alignments and phylogenies
 to extend existing phylogenetic inferences with sequence data available for more taxa

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
more tiny edits

@@
 % taxonomic resource \cite{rees2017automated}, streamlining connections between phylogenies and biological data.
-Physcraper links tip taxon labels in the updated trees to a unified, interoperable
+Physcraper links tip taxon labels in the updated trees to a unified
 taxonomic resource \cite{rees2017automated}, effectively streamlining connections
@@
 Physcraper leverages on existing OpenTree's unified taxonomy APIs to automatically
-standardize taxon names in any phylogeny, providing
-a framework for connection of updated phylogenies with data from any biological database.
+standardize taxon names in any phylogeny, a key step to streamline the connection
+of updated phylogenies with data from different and independent biological database.
 

### 2021-05-24 sources/bmc_template_submission/bmc_article.tex
second review

@@
 the tree of life, but this data either has not been analysed or the analyses
-have not been made publicly available and accesible \cite{mctavish_how_2017}.
+have not been made publicly available and accessible \cite{mctavish_how_2017}.
 
@@
 Physcraper's main result is an updated phylogenetic hypothesis for the search taxon.
-Updated and original phylogeies are compared with Robinson-Foulds weighted and unweighted
+Updated and original phylogenies are compared with Robinson-Foulds weighted and unweighted
 metrics calculated with Dendropy \cite{sukumaran2010dendropy}, and with a node by node
@@
 
-We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics,
-facilitating phylogenetic knowledge for ecologists and other non-specialists, effectively
-democratizing phylogenetic studies.
+We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics.
+It contributes to the democratization of the field by facilitating access to phylogenetic knowledge for ecologists and other non-specialists.
 
@@
 As many studies have shown \cite{huang_what_2009} the effect of missing data can
-be enigmatic, and interact with the true phylogenetic relationships for the data
-set at hand.
+be enigmatic, and interact with the true phylogenetic relationships for the dataset at hand.
 There is not currently strong support in the literature for any particular cutoff
@@
 do not drive differences in phylogenetic inference.
-By gathering the sequences, and making the unaligned files easily available to users,
-researchers can compare if applying any alternate alignment tool of their choice
-affects inferences.
-Once sequences are aligned, they can apply and compare inferences from any phylogenetic
-software.
+By gathering the DNA sequences, and making the unaligned files easily accessible and reusable,
+Physcraper facilitates the exploration of alternative aligning tools to determine their effect on phylogenetic inferences.
+Once sequences are aligned, users can apply practically any phylogenetic
+software and compare results using Physcraper tools.
 
@@
  would exclude potentially homologous flanking regions which are not matched
-by BLAST's local search algorithm, but that may be important for phylogenetic inference.
+using BLAST's local search algorithm, but that may be important for phylogenetic inference.
 Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt"
@@
 PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no
-longer under active development, is challenging to install and run, and has resulted
-in very few phylogenetic analyses since its publication.
+longer under active development, is challenging to install and run, and has produced
+ very few published phylogenetic analyses since its publication.
 
@@
 to statistical support and biological plausibility.
-% Add citation of good things about thining alignments
+% Add citation of good things about thinning alignments?
 
@@
 in different ways. All samples
-corresponding to the ingroup cluster together forming a monophyletic group (Figure 2A), and
+corresponding to the ingroup are clustered together, forming a monophyletic group (Figure 2A), and
 samples belonging to the same \textit{Ilex} species also form
-monophyletic groups (Figure 2B). A notable exception is samples of the species \textit{Ilex theeizans}, which appear as
+monophyletic groups (Figure 2B). A notable exception are samples of the species \textit{Ilex theeizans}, which appear as
 non-monophyletic in the updated phylogeny as well as in the original Gottlieb2005 phylogeny.
@@
 data store can also help flag paralogous sequences.
-Accidentally including paralogs as homologs is a risk in phylogenetic analyses,
-and can be more prevalent in automated analyses than in manually curated analyses.
-We provide users with several tools to try to assess homology of their aligned sequences.
+Accidentally including paralogs as homologs is a known risk of assembling a dataset for phlyogenetic analysis,
+and can be more prevalent in automatically assembled dtasets than in manually curated ones.
+We provide users with several tools to assess homology of aligned sequences.
 The estimated gene phylogeny itself is an evolutionarily explicit way to visualize gene
 evolution, which in concert with taxonomic labelling can reveal paralogy.
-OpenTree's conflict analysis tool informs the users of whether their phylogeny contains major
+OpenTree's conflict analysis tool implemented in Physcraper informs the users of whether their phylogeny contains major
 conflicts with established taxonomy and any phylogenetic context they wish to compare to.
@@
 policy, applications and education.
-Initial ideas about the data are constantly changed by results from new analyses.
+Initial understanding about scientific data is constantly challenged and reframed by new data and analyses.
 Physcraper provides a framework for reproducible phylogenetics that has the
-potential to consistently provide context for these ideas, highlighting the
-importance of data sharing and open science for phylogenetics, biology and science.
+potential to consistently contextualize new knowledge in the light of previous knowledge,
+showcasing the utility and importance of good data sharing practices and open science for phylogenetics, biology and science.
 
@@
 
-Code and datasets generated and analysed during the current study are available in
-the repositories ``physcraper" containing the source code,
+Code and datasets developed and analysed for this study are available at
+the following GitHub repositories: ``physcraper" -- contains the source code,
 \href{https://github.com/McTavishLab/physcraper}{https://github.com/McTavishLab/physcraper};
- ``physcraperex" containing the examples,
+ ``physcraperex" -- contains the examples,
 \href{https://github.com/McTavishLab/physcraperex}{https://github.com/McTavishLab/physcraperex};
-and, ``physcraper\_ms" containing this reproducible manuscript,
+and, ``physcraper\_ms" -- contains this reproducible manuscript,
 \href{https://github.com/McTavishLab/physcraper\_ms}{https://github.com/McTavishLab/physcraper\_ms}.
@@
 Merced (UCM), supported by the NSF Grant No. ACI-1429783.
+The funding body did not play any roles in the design of the study and collection,
+analysis, and interpretation of data nor in writing the manuscript.
 
@@
 LLSR wrote manuscript, alignment code, documentation, performed analyses and developed examples; MK wrote code for ncbidataparser module, filtering of sequences per OTU and using offline blast searches, wrote documentation and tests; EJM conceived study, wrote most of the code, documentation and tests.
-All authors contributed to the manuscript and gave final approval for publication.
+All authors contributed to the manuscript, and have read and approved the final manuscript for publication.
 
 \section*{Acknowledgements}
-We thank the members of the OpenTree development team and the ``short bar" Science
-and Engineering Building 1, UCM, joint lab paper discussion group for valuable
+We thank the members of the current OpenTree development team (MH, KC, MTH, and JA),
+and the ``short bar" Science
+and Engineering Building 1, UCM, Bloise-McTavish joint lab paper discussion group for valuable
 comments on this manuscript. We also thank the valuable comments of David Posada, Rutger Vos,
@@
   Tips in original alignment and new tips added with Physcraper are depicted in
-  black and red, respectively. Physcraper obtained sequences from the GenBank
-  database via local BLAST of all sequences from the Gottlieb2005 internal transcribed
-  spacer DNA region (ITS) original alignment,
-  filtered them following criteria described in section ``DNA sequence search and
-  filtering", aligned them to the original alignment using MUSCLE, and performed a
+  black and red, respectively. First, Physcraper obtained sequences from the GenBank
+  database via local BLAST of all sequences in the Gottlieb2005 original alignment of the internal transcribed
+  spacer DNA region (ITS). Then, it
+  filtered the obtained sequences following criteria described in section ``DNA sequence search and
+  filtering", and aligned them to the original Gottlieb2005 alignment using MUSCLE. Finally, Physcraper performed a
   phylogenetic reconstruction using RAxML, with 100 bootstraps. B) Results of conflict
@@
   Gottlieb2005 phylogeny. Green branches are resolved by the updated phylogeny,
-  blue branches are in agreement between the updated phylogeny and the syntehtic subtree, and
+  blue branches are in agreement between the updated phylogeny and the synthetic subtree, and
   orange branches are in conflict between the two.

### 2021-05-28 sources/bmc_template_submission/bmc_article.tex
final edits :tada:

@@
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-  % corref={aff1},                       % id of corresponding address, if any
+  corref={aff1},                       % id of corresponding address, if any
   email={sanchez.reyes.luna@gmail.com}   % email address
@@
   addressref={aff1},
-  corref={aff1},
+  % corref={aff1},
   email={ejmctavish@ucmerced.edu}
@@
 Accidentally including paralogs as homologs is a known risk of assembling a dataset for phlyogenetic analysis,
-and can be more prevalent in automatically assembled dtasets than in manually curated ones.
+and can be more prevalent in automatically assembled datasets than in manually curated ones.
 We provide users with several tools to assess homology of aligned sequences.
@@
 The Physcraper workflow can be used to rapidly (in a matter of hours) create
-phylogenies which can address challenges overarching both fields of ecology and evolution, such as
+phylogenies to address challenges overarching both fields of ecology and evolution, such as
 phylogenetically placing newly discovered species \cite{webb2010biodiversity},
@@
 policy, applications and education.
-Initial understanding about scientific data is constantly challenged and reframed by new data and analyses.
+Scientific understanding is constantly challenged and reframed by new data and analyses.
 Physcraper provides a framework for reproducible phylogenetics that has the
-potential to consistently contextualize new knowledge in the light of previous knowledge,
+potential to consistently contextualize new knowledge in the light of previous understanding,
 showcasing the utility and importance of good data sharing practices and open science for phylogenetics, biology and science.

### 2021-05-28 sources/bmc_template_submission/bmc_article.tex
some moretiny edits

@@
 Databases preserving and democratizing access to biological data
-have become essential resources for science.
-New molecular data keep accumulating and tools facilitating its integration into
+have become essential resources for science. New molecular data keep accumulating and tools facilitating its integration into
 existent evolutionary knowledge contribute to the acceleration of scientific discovery.
-
 Physcraper is a tool that builds upon previous knowledge stored in published
@@
 of phylogenies, with the main goal of extending our knowledge of phylogenetic relationships
-across the tree of life.
-
-We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics.
-It contributes to the democratization of the field by facilitating access to phylogenetic knowledge for ecologists and other non-specialists.
+across the tree of life. We believe this is a key step to successfully establish an open, reproducible workflow for phylogenetics.
+As such, it facilitates access to phylogenetic knowledge for non-specialists in diverse research areas (e.g., ecology, medicine),
+effectively contributing to the democratization of phylogenetic studies.
 
@@
 make evolutionary sense of the vast amount of
-public molecular data available.
-
-Several analysis tools create full phylogenies \textit{de novo} by mining of molecular
+public molecular data available. Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
+which add new taxa without modifying the input phylogeny, Physcraper estimates all
+the relationships anew in the context of the new data.
+The tool PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no
+longer under active development, is challenging to install and run, and has produced
+few published phylogenetic analyses since its publication.
+Several other existing tools create full phylogenies \textit{de novo} by mining of molecular
 databases \cite{antonelli2017toward, sanderson2008phylota, smith2019pyphlawd,
-bennett2018phylotar, pearse2013phylogenerator}.
-In particular, Phylota \cite{sanderson2008phylota},
+bennett2018phylotar, pearse2013phylogenerator}. In particular, Phylota \cite{sanderson2008phylota},
 and PHLAWD \cite{smith2009mega}, have been cited and used abundantly.
-
-Physcraper builds on this automated database mining concept by incorporating prior
+Physcraper adds to this automated database mining concept by incorporating prior
 phylogenetic work and existing taxonomic domain knowledge on appropriate markers
-and alignment construction.
+and alignment construction, to update existing phylogenetic knowledge.
 This decreases error (requiring less manual downstream processing) and
-eases comparison with previous phylogenetic knowledge.
-
+eases comparison with previous phylogenetic hypotheses.
 Results from the verification test highlight the importance of incorporating existing expertly
@@
 filtered sequences, and potentially modify configuration parameters such as the
-sequence length cutoff parameter, based on the filtered sequences.
-Default filtering parameters are arbitrary, but we hope that by making the process
+sequence length cutoff parameter, based on said information.
+Physcraper's default filtering parameters are arbitrary, but we hope that by making the process
 of locating homologous sequences online reproducible, and tracking what filters
 are used, we make it easier for researchers to delve into the effect that different
-choices have on their inferences.
+filtering choices have on their inferences.
 This is in contrast to ``manual" searches for taxa, where similarly arbitrary filters
 are applied, but are difficult to trace.
-As many studies have shown \cite{huang_what_2009} the effect of missing data can
+It has been shown \cite{huang_what_2009} that the effect of missing data in alignments can
 be enigmatic, and interact with the true phylogenetic relationships for the dataset at hand.
-There is not currently strong support in the literature for any particular cutoff
+There is not currently strong support in the literature for any particular filtering cutoff
 value, and rather than prescribe specific approaches, we encourage users to explore
-the effects of different choices on their phylogenetic inferences.
+the effects of different parameter values on resulting phylogenetic hypotheses.
 In addition, by providing the output files at each step of the analysis,
-it is straightforward to assess how changing parameter and software choices do or
-do not drive differences in phylogenetic inference.
+it is straightforward to assess how changing filtering cutoffs and software choices might drive differences in phylogenetic inference.
 By gathering the DNA sequences, and making the unaligned files easily accessible and reusable,
@@
 Once sequences are aligned, users can apply practically any phylogenetic
-software and compare results using Physcraper tools.
+software and compare results using Physcraper tools in a reproducible framework.
 
@@
  would exclude potentially homologous flanking regions which are not matched
-using BLAST's local search algorithm, but that may be important for phylogenetic inference.
+by BLAST's local search algorithm, but that may be important for phylogenetic inference.
 Instead we list the accession numbers for these matches in the ``seqlen\_mismatch.txt"
@@
 alignment of interest.
-
-Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
-which add new taxa without modifying the input phylogeny, Physcraper estimates all
-the relationships anew in the context of the new data.
-PUMPER \cite{izquierdo2014pumper} shares these conceptual strengths, but is no
-longer under active development, is challenging to install and run, and has produced
- very few published phylogenetic analyses since its publication.
 
@@
 The Physcraper workflow can be used to rapidly (in a matter of hours) create
-phylogenies to address challenges overarching both fields of ecology and evolution, such as
+phylogenies which can address challenges overarching both fields of ecology and evolution, such as
 phylogenetically placing newly discovered species \cite{webb2010biodiversity},

### 2021-05-28 sources/bmc_template_submission/bmc_article.tex
some moretiny edits

@@
 \section*{Acknowledgements}
-We thank the members of the current OpenTree development team (MH, KC, MTH, and JA),
+We thank the members of the current OpenTree development team (MH, KC, BR, and JA),
 and the ``short bar" Science
-and Engineering Building 1, UCM, Bloise-McTavish joint lab paper discussion group for valuable
+and Engineering Building 1, UCM, Blois-McTavish joint lab paper discussion group for valuable
 comments on this manuscript. We also thank the valuable comments of David Posada, Rutger Vos,

### 2021-05-28 sources/bmc_template_submission/bmc_article.tex
some moretiny edits

@@
 potential to consistently contextualize new knowledge in the light of previous understanding,
-showcasing the utility and importance of good data sharing practices and open science for phylogenetics, biology and science.
+showcasing the utility and importance of good data sharing practices and open research for phylogenetics, biology and science.
 

### 2021-05-29 sources/bmc_template_submission/bmc_article.tex
last tiny edits

--- 2021-05-28 sources/bmc_template_submission/bmc_article.tex
@@
 make evolutionary sense of the vast amount of
-public molecular data available. Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
+public genetic data available. Unlike phylogenetic placement approaches \cite{berger_performance_2011, matsen_pplacer_2010},
 which add new taxa without modifying the input phylogeny, Physcraper estimates all
@@
 
-We encourage users to look at the output files containing information about the
-filtered sequences, and potentially modify configuration parameters such as the
-sequence length cutoff parameter, based on said information.
+We encourage users to look at Physcraper's output files containing information about the
+filtered sequences, and use this results to potentially modify configuration parameters such as the
+sequence length cutoff parameter in subsequent Physcraper runs.
 Physcraper's default filtering parameters are arbitrary, but we hope that by making the process
@@
 filtering choices have on their inferences.
-This is in contrast to ``manual" searches for taxa, where similarly arbitrary filters
+This is in contrast to ``manual" searches for DNA sequences in molecular databases, where similarly arbitrary filters
 are applied, but are difficult to trace.
@@
 By gathering the DNA sequences, and making the unaligned files easily accessible and reusable,
-Physcraper facilitates the exploration of alternative aligning tools to determine their effect on phylogenetic inferences.
+Physcraper facilitates the exploration of alternative aligning tools.
 Once sequences are aligned, users can apply practically any phylogenetic
@@
 potential to consistently contextualize new knowledge in the light of previous understanding,
-showcasing the utility and importance of good data sharing practices and open science for phylogenetics, biology and science.
+showcasing the utility and importance of good data sharing practices and open science for the advancement of phylogenetics, biology and research.
 
